### 2014-02-05 report/paper.tex
eerste onderzoeksopzet

\documentclass{sig-alternate-br}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{xcolor,colortbl}
\usepackage[official]{eurosym}
\usepackage{hyperref}
% \url{} als link + fixed-width font
\usepackage{url}

% interpunctie, citations
\usepackage[english]{babel}
\usepackage{csquotes}% Recommended
\usepackage{natbib}

% tabel tool
\usepackage{tabu}

\usepackage{graphicx}
\usepackage{subfig}

\begin{document}
\CopyrightYear{2013}

\title{Methods for Large Scale Learning to Rank}

\numberofauthors{1}
\author{
\alignauthor Niek Tax\\
       \affaddr{University of Twente}\\
       \affaddr{P.O. Box 217, 7500AE Enschede}\\
       \affaddr{The Netherlands}\\
}

\maketitle
\begin{abstract}
Here comes the abstract
\end{abstract}

\keywords{Learning to Rank (LTR), Machine Learned Ranking (MLR), MapReduce, Parallel Computing}

\section{Introduction}
Several Learning to Rank (LTR) benchmark datasets have been proposed in recent years \cite{Chapelle2011a,Qin2010,Alcantara2010} (Yandex Internet Mathematics competition). Chapelle et al \cite{Chapelle2011b} observed that the top competitors' results in LTR competitions are very close to each other, which raises the presumption that high accuracy LTR is a 'solved problem'.\\

Chappelle et al \cite{Chapelle2011b} identified efficiency and scalability of LTR algorithms to be one of the most important directions of future research in the field of LTR. The rationale behind this observation is that IR datasets are becoming bigger and bigger and that algorithm efficiency and scalability has not played a role in any LTR benchmark so far.
\section{Problem Statement \& Research Questions}
Several attempts on parallel or distributed machine learning \cite{Chu2007,Chang2007} has been made. The work on parallel or distributed LTR is still very limited compared to the amount of work that has been done on parallel or distributed regular machine learning. The field of efficient LTR has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012} since Chapelle at all \cite{Chapelle2011b} stated its growing importance back in 2011, only few of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient LTR through the use of parallel programming paradigms. Lin \cite{Lin2013} observed that algorithms that are of iterative nature, like most LTR algorithms, are not amenable to the MapReduce framework. Lin \cite{Lin2013} argued that iterative algorithms can often be replaced with non-iterative alternatives or can still be optimized in such a way that its performance in a MapReduce setting is good enough.\\

\textbf{Research Question:} How can we efficiently run LTR algorithms in a distributed setting using MapReduce? (scope can possibly be extended to other parallel processing paradigms like MPI)

\section{Methodology}
\begin{enumerate}
\item Identify well performing LTR algorithms on benchmark datasets (using benchmark overview papers and other literature)
\item Search for literature on parallelisation of algorithms
\item Experiment with optimizing LTR algorithms with MapReduce (maybe include MPI to research scope) implementations of algorithms (compared to literature performance when existing, or compared to non-optimized performance otherwise).
\end{enumerate}
\section{Results \& Discussion}
\section{Conclusions}
\section{Future Work}

\section{Acknowledgement}

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.

\bibliographystyle{abbrv}
\bibliography{paper}

\balancecolumns

\onecolumn

\end{document}

### 2014-02-11 classicthesis/Chapters/Introduction.tex
introduction & problem statement in classicthesis

%********************************************************************
% Appendix
%*******************************************************
% If problems with the headers: get headings in appendix etc. right
%\markboth{\spacedlowsmallcaps{Appendix}}{\spacedlowsmallcaps{Appendix}}
\chapter{Motivation and Problem Statement}
Several Learning to Rank (LTR) benchmark datasets have been proposed in recent years \cite{Chapelle2011a,Qin2010,Alcantara2010} (Yandex Internet Mathematics competition). Chapelle et al \cite{Chapelle2011b} observed that the top competitors' results in LTR competitions are very close to each other, which raises the presumption that high accuracy LTR is a 'solved problem'.\\

Chappelle et al \cite{Chapelle2011b} identified efficiency and scalability of LTR algorithms to be one of the most important directions of future research in the field of LTR. The rationale behind this observation is that IR datasets are becoming bigger and bigger and that algorithm efficiency and scalability has not played a role in any LTR benchmark so far.\\

Several attempts on parallel or distributed machine learning \cite{Chu2007,Chang2007} has been made. The work on parallel or distributed LTR is still very limited compared to the amount of work that has been done on parallel or distributed regular machine learning. The field of efficient LTR has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012} since Chapelle at all \cite{Chapelle2011b} stated its growing importance back in 2011, only few of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient LTR through the use of parallel programming paradigms. Lin \cite{Lin2013} observed that algorithms that are of iterative nature, like most LTR algorithms, are not amenable to the MapReduce framework. Lin \cite{Lin2013} argued that iterative algorithms can often be replaced with non-iterative alternatives or can still be optimized in such a way that its performance in a MapReduce setting is good enough.\\

\chapter{Research Goals}
The objective of this thesis is the exploration of execution time speed-up of Learning-to-Rank algorithms through parallelization using Hadoop. 
This work focuses on those Learning-to-Rank algorithms that have shown leading performance on relevant benchmark datasets.
The research questions raised and answered in this work include:
\begin{itemize}
\item What is the speed-up of existing Learning-to-Rank algorithms when implemented and executed on Hadoop?
\item Can we adjust those Learning-to-Rank algorithms such that the parallel execution speed-up increases without decreasing accuracy?
\end{itemize}

\chapter{Approach}
Equidem detraxit cu nam, vix eu delenit periculis. Eos ut vero
constituto, no vidit propriae complectitur sea. Diceret nonummy in
has, no qui eligendi recteque consetetur. Mel eu dictas suscipiantur,
et sed placerat oporteat. At ipsum electram mei, ad aeque atomorum
mea.

\chapter{Thesis Overview}

\begin{table}
    \myfloatalign
  \begin{tabularx}{\textwidth}{Xll} \toprule
    \tableheadline{labitur bonorum pri no} & \tableheadline{que vista}
    & \tableheadline{human} \\ \midrule
    fastidii ea ius & germano &  demonstratea \\
    suscipit instructior & titulo & personas \\
    %postulant quo & westeuropee & sanctificatec \\
    \midrule
    quaestio philosophia & facto & demonstrated \\
    %autem vulputate ex & parola & romanic \\
    %usu mucius iisque & studio & sanctificatef \\
    \bottomrule
  \end{tabularx}
  \caption[Autem usu id]{Autem usu id.}
  \label{tab:moreexample}
\end{table}

Ei solet nemore consectetuer nam. Ad eam porro impetus, te choro omnes
evertitur mel. Molestie conclusionemque vel at, no qui omittam
expetenda efficiendi. Eu quo nobis offendit, verterem scriptorem ne
vix.

\begin{lstlisting}[float,caption=A floating example]
for i:=maxint to 0 do
begin
{ do nothing }
end;
\end{lstlisting}

### 2014-02-11 classicthesis/Chapters/Introduction.tex
thesis outline

@@
 
-\begin{table}
-    \myfloatalign
-  \begin{tabularx}{\textwidth}{Xll} \toprule
-    \tableheadline{labitur bonorum pri no} & \tableheadline{que vista}
-    & \tableheadline{human} \\ \midrule
-    fastidii ea ius & germano &  demonstratea \\
-    suscipit instructior & titulo & personas \\
-    %postulant quo & westeuropee & sanctificatec \\
-    \midrule
-    quaestio philosophia & facto & demonstrated \\
-    %autem vulputate ex & parola & romanic \\
-    %usu mucius iisque & studio & sanctificatef \\
-    \bottomrule
-  \end{tabularx}
-  \caption[Autem usu id]{Autem usu id.}
-  \label{tab:moreexample}
-\end{table}
-
-Ei solet nemore consectetuer nam. Ad eam porro impetus, te choro omnes
-evertitur mel. Molestie conclusionemque vel at, no qui omittam
-expetenda efficiendi. Eu quo nobis offendit, verterem scriptorem ne
-vix.
-
-\begin{lstlisting}[float,caption=A floating example]
-for i:=maxint to 0 do
-begin
-{ do nothing }
-end;
-\end{lstlisting}
+\begin{description}
+\item[Part II ]{introduces the reader to the basic principles and recent work in the fields of Learning-to-Rank}
+\item[Part III ]{describes a short overview work in the field of parallel machine learning and Learning-to-Rank.}
+\item[Part IV ]{describes implementation details of the Learning-to-Rank algorithms in the Hadoop framework.}
+\item[Part V ]{presents and discusses speed-up results for the implemented Learning-to-Rank methods.}
+\item[Part VI ]{summarizes the results. The limitations of our research as well as future research directions in the field are mentioned here.}
+\end{description}

### 2014-02-11 classicthesis/Chapters/Background.tex
witpaginas weg en structuur background

%************************************************
\chapter{Introduction to Learning-to-Rank}
\chapter{Benchmark Datasets}
\chapter{Pointwise Approach}
\chapter{Pairwise Approach}
\chapter{Listwise Approach}
\chapter{Summary}
%************************************************
Ei choro aeterno antiopam mea, labitur bonorum pri no. His no decore
nemore graecis. In eos meis nominavi, liber soluta vim cu. Sea commune
suavitate interpretaris eu, vix eu libris efficiantur.

\section{Some Formulas}
Due to the statistical nature of ionisation energy loss, large
fluctuations can occur in the amount of energy deposited by a particle
traversing an absorber element\footnote{Examples taken from Walter
Schmidt's great gallery: \\
\url{http://home.vrweb.de/~was/mathfonts.html}}.  Continuous processes
such as multiple
scattering and energy loss play a relevant role in the longitudinal
and lateral development of electromagnetic and hadronic
showers, and in the case of sampling calorimeters the
measured resolution can be significantly affected by such fluctuations
in their active layers.  The description of ionisation fluctuations is
characterised by the significance parameter $\kappa$, which is
proportional to the ratio of mean energy loss to the maximum allowed
energy transfer in a single collision with an atomic electron:
\graffito{You might get unexpected results using math in chapter or
section heads. Consider the \texttt{pdfspacing} option.}
\begin{equation}
\kappa =\frac{\xi}{E_{\textrm{max}}} %\mathbb{ZNR}
\end{equation}
$E_{\textrm{max}}$ is the maximum transferable energy in a single
collision with an atomic electron.
\[
E_{\textrm{max}} =\frac{2 m_{\textrm{e}} \beta^2\gamma^2 }{1 +
2\gamma m_{\textrm{e}}/m_{\textrm{x}} + \left ( m_{\textrm{e}}
/m_{\textrm{x}}\right)^2}\ ,
\]
where $\gamma = E/m_{\textrm{x}}$, $E$ is energy and
$m_{\textrm{x}}$ the mass of the incident particle,
$\beta^2 = 1 - 1/\gamma^2$ and $m_{\textrm{e}}$ is the electron mass.
$\xi$ comes from the Rutherford scattering cross section
and is defined as:
\begin{eqnarray*} \xi  = \frac{2\pi z^2 e^4 N_{\textrm{Av}} Z \rho
\delta x}{m_{\textrm{e}} \beta^2 c^2 A} =  153.4 \frac{z^2}{\beta^2}
\frac{Z}{A}
  \rho \delta x \quad\textrm{keV},
\end{eqnarray*}
where

\begin{tabular}{ll}
$z$          & charge of the incident particle \\
$N_{\textrm{Av}}$     & Avogadro's number \\
$Z$          & atomic number of the material \\
$A$          & atomic weight of the material \\
$\rho$       & density \\
$ \delta x$  & thickness of the material \\
\end{tabular}

$\kappa$ measures the contribution of the collisions with energy
transfer close to $E_{\textrm{max}}$.  For a given absorber, $\kappa$
tends
towards large values if $\delta x$ is large and/or if $\beta$ is
small.  Likewise, $\kappa$ tends towards zero if $\delta x $ is small
and/or if $\beta$ approaches $1$.

The value of $\kappa$ distinguishes two regimes which occur in the
description of ionisation fluctuations:

\begin{enumerate}
\item A large number of collisions involving the loss of all or most
  of the incident particle energy during the traversal of an absorber.

  As the total energy transfer is composed of a multitude of small
  energy losses, we can apply the central limit theorem and describe
  the fluctuations by a Gaussian distribution.  This case is
  applicable to non-relativistic particles and is described by the
  inequality $\kappa > 10 $ (\ie, when the mean energy loss in the
  absorber is greater than the maximum energy transfer in a single
  collision).

\item Particles traversing thin counters and incident electrons under
  any conditions.

  The relevant inequalities and distributions are $ 0.01 < \kappa < 10
  $,
  Vavilov distribution, and $\kappa < 0.01 $, Landau distribution.
\end{enumerate}

\section{Various Mathematical Examples}
If $n > 2$, the identity
\[
  t[u_1,\dots,u_n] = t\bigl[t[u_1,\dots,u_{n_1}], t[u_2,\dots,u_n]
  \bigr]
\]
defines $t[u_1,\dots,u_n]$ recursively, and it can be shown that the
alternative definition
\[
  t[u_1,\dots,u_n] = t\bigl[t[u_1,u_2],\dots,t[u_{n-1},u_n]\bigr]
\]
gives the same result.

%*****************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************

### 2014-02-11 classicthesis/Chapters/Background.tex
alle tekst die niet van mij was is nu echt weg :)

@@
 \chapter{Summary}
-%************************************************
-Ei choro aeterno antiopam mea, labitur bonorum pri no. His no decore
-nemore graecis. In eos meis nominavi, liber soluta vim cu. Sea commune
-suavitate interpretaris eu, vix eu libris efficiantur.
-
-\section{Some Formulas}
-Due to the statistical nature of ionisation energy loss, large
-fluctuations can occur in the amount of energy deposited by a particle
-traversing an absorber element\footnote{Examples taken from Walter
-Schmidt's great gallery: \\
-\url{http://home.vrweb.de/~was/mathfonts.html}}.  Continuous processes
-such as multiple
-scattering and energy loss play a relevant role in the longitudinal
-and lateral development of electromagnetic and hadronic
-showers, and in the case of sampling calorimeters the
-measured resolution can be significantly affected by such fluctuations
-in their active layers.  The description of ionisation fluctuations is
-characterised by the significance parameter $\kappa$, which is
-proportional to the ratio of mean energy loss to the maximum allowed
-energy transfer in a single collision with an atomic electron:
-\graffito{You might get unexpected results using math in chapter or
-section heads. Consider the \texttt{pdfspacing} option.}
-\begin{equation}
-\kappa =\frac{\xi}{E_{\textrm{max}}} %\mathbb{ZNR}
-\end{equation}
-$E_{\textrm{max}}$ is the maximum transferable energy in a single
-collision with an atomic electron.
-\[
-E_{\textrm{max}} =\frac{2 m_{\textrm{e}} \beta^2\gamma^2 }{1 +
-2\gamma m_{\textrm{e}}/m_{\textrm{x}} + \left ( m_{\textrm{e}}
-/m_{\textrm{x}}\right)^2}\ ,
-\]
-where $\gamma = E/m_{\textrm{x}}$, $E$ is energy and
-$m_{\textrm{x}}$ the mass of the incident particle,
-$\beta^2 = 1 - 1/\gamma^2$ and $m_{\textrm{e}}$ is the electron mass.
-$\xi$ comes from the Rutherford scattering cross section
-and is defined as:
-\begin{eqnarray*} \xi  = \frac{2\pi z^2 e^4 N_{\textrm{Av}} Z \rho
-\delta x}{m_{\textrm{e}} \beta^2 c^2 A} =  153.4 \frac{z^2}{\beta^2}
-\frac{Z}{A}
-  \rho \delta x \quad\textrm{keV},
-\end{eqnarray*}
-where
-
-\begin{tabular}{ll}
-$z$          & charge of the incident particle \\
-$N_{\textrm{Av}}$     & Avogadro's number \\
-$Z$          & atomic number of the material \\
-$A$          & atomic weight of the material \\
-$\rho$       & density \\
-$ \delta x$  & thickness of the material \\
-\end{tabular}
-
-$\kappa$ measures the contribution of the collisions with energy
-transfer close to $E_{\textrm{max}}$.  For a given absorber, $\kappa$
-tends
-towards large values if $\delta x$ is large and/or if $\beta$ is
-small.  Likewise, $\kappa$ tends towards zero if $\delta x $ is small
-and/or if $\beta$ approaches $1$.
-
-The value of $\kappa$ distinguishes two regimes which occur in the
-description of ionisation fluctuations:
-
-\begin{enumerate}
-\item A large number of collisions involving the loss of all or most
-  of the incident particle energy during the traversal of an absorber.
-
-  As the total energy transfer is composed of a multitude of small
-  energy losses, we can apply the central limit theorem and describe
-  the fluctuations by a Gaussian distribution.  This case is
-  applicable to non-relativistic particles and is described by the
-  inequality $\kappa > 10 $ (\ie, when the mean energy loss in the
-  absorber is greater than the maximum energy transfer in a single
-  collision).
-
-\item Particles traversing thin counters and incident electrons under
-  any conditions.
-
-  The relevant inequalities and distributions are $ 0.01 < \kappa < 10
-  $,
-  Vavilov distribution, and $\kappa < 0.01 $, Landau distribution.
-\end{enumerate}
-
-\section{Various Mathematical Examples}
-If $n > 2$, the identity
-\[
-  t[u_1,\dots,u_n] = t\bigl[t[u_1,\dots,u_{n_1}], t[u_2,\dots,u_n]
-  \bigr]
-\]
-defines $t[u_1,\dots,u_n]$ recursively, and it can be shown that the
-alternative definition
-\[
-  t[u_1,\dots,u_n] = t\bigl[t[u_1,u_2],\dots,t[u_{n-1},u_n]\bigr]
-\]
-gives the same result.
-
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************

### 2014-02-11 classicthesis/FrontBackmatter/Acknowledgments.tex
references fixed :)

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape    
    We have seen that computer programming is an art, \\ 
    because it applies accumulated knowledge to the world, \\ 
    because it requires skill and ingenuity, and especially \\
    because it produces objects of beauty.} \\ \medskip
\end{flushright}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
Put your acknowledgments here.

Many thanks to everybody who already sent me a postcard!

Regarding the typography and other help, many thanks go to Marco 
Kuhlmann, Philipp Lehman, Lothar Schlesier, Jim Young, Lorenzo 
Pantieri and Enrico Gregorio\footnote{Members of GuIT (Gruppo 
Italiano Utilizzatori di \TeX\ e \LaTeX )}, J\"org Sommer, 
Joachim K\"ostler, Daniel Gottschlag, Denis Aydin, Paride 
Legovini, Steffen Prochnow, Nicolas Repp, Hinrich Harms, 
 Roland Winkler, J\"org Weber, 
 and the whole \LaTeX-community for support, ideas and 
 some great software.

\bigskip

\noindent\emph{Regarding \mLyX}: The \mLyX\ port was intially done by 
\emph{Nicholas Mariette} in March 2009 and continued by 
\emph{Ivo Pletikosi\'c} in 2011. Thank you very much for your 
work and the contributions to the original style.

\endgroup

### 2014-02-11 classicthesis/Chapters/Background.tex
definition LTR

@@
 \chapter{Introduction to Learning-to-Rank}
+Different definitions of Learning-to-Rank exist. In general, all methods that use machine learning technologies to solve the problem of ranking are called Learning-to-Rank methods.\\
+Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning-to-Rank method when it meets the following two properties:
+\begin{description}
+\item[Feature Based]{\emph{Feature based} means that all documents under investigation are represented by feature vectors that reflect the relevance of the documents to the query.}
+\item[Discriminative Training]{\emph{Discriminative training} means that the learning process can be well described by the four components of discriminative learning, that is, a Learning-to-Rank method has its own input space, output space, hypothesis space, and loss function.
+Input space, output space, hypothesis space, and loss function are hereby defined as follows:
+	\begin{description}
+	\item[Input Space]{contains the objects under investigation. Usually objects are represented by feature vectors, extracted according to different applications.}
+	\item[Output Space]{contains the learning target with respect to the input objects.}
+	\item[Hypothesis Space]{defines the class of functions mapping the input space to the output space.
+The functions operate on the feature vectors of the input object, and make predictions according to the format of the output space.}
+	\item[Loss Function]{In order to learn the optimal hypothesis, a training set is usually used, which contains a number of objects and their ground truth labels, sampled from the product of the input and output spaces.}
+	\end{description}
+Figure \ref{fig:discriminative_training} shows the relationships between input space, output space, hypothesis space and loss function.
+}
+
+\begin{figure}
+\includegraphics[scale=0.5]{gfx/descriminative_training}
+\caption{Machine learning framework for Learning-to-Rank, from Lui\cite{Liu2007}}
+\label{fig:discriminative_training}
+\end{figure}
+
+\end{description}
 \chapter{Benchmark Datasets}

### 2014-02-11 classicthesis/Chapters/Background.tex
uitbreiding background

@@
 	\end{description}
+\begin{figure}
+\includegraphics[scale=0.4]{gfx/descriminative_training}
+\caption{Machine learning framework for Learning-to-Rank, from Liu\cite{Liu2007}}
+\label{fig:discriminative_training}
+\end{figure}
 Figure \ref{fig:discriminative_training} shows the relationships between input space, output space, hypothesis space and loss function.
@@
 
+Figure \ref{ref:ltr_framework} shows how the learning process in Learning-to-Rank typically happens.
+A set of queries $q_i$ with $n > i > 1$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgments of those documents $y_i$ are inputted to a learning algorithm that learns a ranking model such that the output of the ranking model can predict relevant judgment labels in this training set as accurately as possible.
+Later on we can test the accuracy of this ranking model by predicting the relevant judgments new documents given a new query.
 \begin{figure}
-\includegraphics[scale=0.5]{gfx/descriminative_training}
-\caption{Machine learning framework for Learning-to-Rank, from Lui\cite{Liu2007}}
-\label{fig:discriminative_training}
+\includegraphics[scale=0.3]{gfx/ltr_framework}
+\caption{The Learning-to-Rank framework, from Liu\cite{Liu2007}}
+\label{ref:ltr_framework}
 \end{figure}

### 2014-02-12 classicthesis/FrontBackmatter/Acknowledgments.tex
acknowledgement geschreven en dedication aangepast

@@
 \chapter*{Acknowledgments}
-Put your acknowledgments here.
+Many thanks go to Dr. ir. Djoerd Hiemstra of the University of Twente and to Sander Bockting Msc. of Avanade Netherlands BV for their valuable insights and great supervision.
 
-Many thanks to everybody who already sent me a postcard!
-
-Regarding the typography and other help, many thanks go to Marco 
-Kuhlmann, Philipp Lehman, Lothar Schlesier, Jim Young, Lorenzo 
-Pantieri and Enrico Gregorio\footnote{Members of GuIT (Gruppo 
-Italiano Utilizzatori di \TeX\ e \LaTeX )}, J\"org Sommer, 
-Joachim K\"ostler, Daniel Gottschlag, Denis Aydin, Paride 
-Legovini, Steffen Prochnow, Nicolas Repp, Hinrich Harms, 
- Roland Winkler, J\"org Weber, 
- and the whole \LaTeX-community for support, ideas and 
- some great software.
-
-\bigskip
-
-\noindent\emph{Regarding \mLyX}: The \mLyX\ port was intially done by 
-\emph{Nicholas Mariette} in March 2009 and continued by 
-\emph{Ivo Pletikosi\'c} in 2011. Thank you very much for your 
-work and the contributions to the original style.
+In addition I would like to thank all other graduate interns at Avanade as well as all the Avanade employees who were so kind to provide me with valuable insights.
 

### 2014-02-12 classicthesis/Chapters/Background.tex
acknowledgement geschreven en dedication aangepast
introductie begint beter te lopen

@@
 	\end{description}
-\begin{figure}
-\includegraphics[scale=0.4]{gfx/descriminative_training}
+	}
+\end{description}
+\begin{figure}[!h]
+\includegraphics[scale=0.26]{gfx/descriminative_training}
 \caption{Machine learning framework for Learning-to-Rank, from Liu\cite{Liu2007}}
@@
 Figure \ref{fig:discriminative_training} shows the relationships between input space, output space, hypothesis space and loss function.
-}
 
@@
 Later on we can test the accuracy of this ranking model by predicting the relevant judgments new documents given a new query.
-\begin{figure}
-\includegraphics[scale=0.3]{gfx/ltr_framework}
-\caption{The Learning-to-Rank framework, from Liu\cite{Liu2007}}
+\begin{figure}[!h]
+\includegraphics[scale=0.25]{gfx/ltr_framework}
+\caption{A typical Learning-to-Rank setting, from Liu\cite{Liu2007}}
 \label{ref:ltr_framework}
-\end{figure}
+\end{figure}\\
 
-\end{description}
+Learning-to-Rank algorithms can be classified into three approaches: the pointwise approach, the pairwise approach and the listwise approach, on which we will elaborate in the subsequent chapters.
+
 \chapter{Benchmark Datasets}

### 2014-02-12 classicthesis/Chapters/Introduction.tex
introductie begint beter te lopen

@@
 \chapter{Motivation and Problem Statement}
-Several Learning to Rank (LTR) benchmark datasets have been proposed in recent years \cite{Chapelle2011a,Qin2010,Alcantara2010} (Yandex Internet Mathematics competition).
-Chapelle et al \cite{Chapelle2011b} observed that the top competitors' results in LTR competitions are very close to each other, which raises the presumption that high accuracy LTR is a 'solved problem'.\\
+Ranking is a core problem in the field of information retrieval.
+The ranking task in information retrieval entails the ranking of candidate documents according to their relevance to a query.
+Heuristic ranking models have long been around in the information retrieval field.
+The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models.
+Learning-to-Rank is the relatively new research area covering the use of machine learning models for the ranking task.\\
 
-Chappelle et al \cite{Chapelle2011b} identified efficiency and scalability of LTR algorithms to be one of the most important directions of future research in the field of LTR.
-The rationale behind this observation is that IR datasets are becoming bigger and bigger and that algorithm efficiency and scalability has not played a role in any LTR benchmark so far.\\
+In recent years several Learning-to-Rank benchmark datasets have been proposed that allow us to compare the performance of different Learning-to-Rank methods.
+Well-known benchmark datasets include the \emph{Yahoo!
+Learning to Rank Challenge} dataset\cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR dataset\cite{Qin2010} that was build by Microsoft Research.
+One of the concluding observations of the \emph{Yahoo!
+Learning to Rank Challenge} was that almost all work in the Learning-to-Rank field focuses on ranking accuracy and that efficiency and scalability of Learning-to-Rank algorithms is still an underexposed research area of which its importance is likely to increase in the near future as training sets are becoming larger and larger\cite{Chapelle2011b}.
+Liu\cite{Liu2007} confirms the observation that efficiency and scalability of Learning-to-Rank methods has so far been an overlooked research area in his influential book on Learning-to-Rank.\\
 
-Several attempts on parallel or distributed machine learning \cite{Chu2007,Chang2007} has been made.
-The work on parallel or distributed LTR is still very limited compared to the amount of work that has been done on parallel or distributed regular machine learning.
+Some research has been done in the area of on parallel or distributed machine learning \cite{Chu2007,Chang2007}.
+However, almost none of these studies include the Learning-to-Rank subfield of machine learning.
 The field of efficient LTR has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012} since Chapelle at all \cite{Chapelle2011b} stated its growing importance back in 2011, only few of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient LTR through the use of parallel programming paradigms.

### 2014-02-12 classicthesis/FrontBackmatter/Acknowledgments.tex
mooie acknowledgement quote :)

@@
 \begin{flushright}{\slshape    
-    We have seen that computer programming is an art, \\ 
-    because it applies accumulated knowledge to the world, \\ 
-    because it requires skill and ingenuity, and especially \\
-    because it produces objects of beauty.} \\ \medskip
+   God gave you a gift of 86 400 seconds today. \\
+   Have you used one to say thank you?.} \\ \medskip
+   - William Arthur Ward
 \end{flushright}
@@
 \chapter*{Acknowledgments}
-Many thanks go to Dr. ir. Djoerd Hiemstra of the University of Twente and to Sander Bockting Msc. of Avanade Netherlands BV for their valuable insights and great supervision.
+Many thanks go to Dr. ir. Djoerd Hiemstra of the University of Twente and to Sander Bockting Msc. of Avanade Netherlands BV for their great supervision throughout the project.
 
-In addition I would like to thank all other graduate interns at Avanade as well as all the Avanade employees who were so kind to provide me with valuable insights.
+In addition I would like to thank all other graduate interns at Avanade as well as all the Avanade employees for the great talks at the coffee machine or elsewhere.
+In particular I owe many thanks to Michel Barneveld for giving me insight in potential applications of Learning-to-Rank within Avanade.
 

### 2014-02-12 classicthesis/Chapters/Introduction.tex
meer werk aan introductie

@@
 However, almost none of these studies include the Learning-to-Rank subfield of machine learning.
-The field of efficient LTR has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012} since Chapelle at all \cite{Chapelle2011b} stated its growing importance back in 2011, only few of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient LTR through the use of parallel programming paradigms.
-Lin \cite{Lin2013} observed that algorithms that are of iterative nature, like most LTR algorithms, are not amenable to the MapReduce framework.
-Lin \cite{Lin2013} argued that iterative algorithms can often be replaced with non-iterative alternatives or can still be optimized in such a way that its performance in a MapReduce setting is good enough.\\
+The field of efficient Learning-to-Rank has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012} since Chapelle at all \cite{Chapelle2011b} stated its growing importance back in 2011, but only few of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient Learning-to-Rank through the use of parallel programming paradigms.
+Lin \cite{Lin2013} observed that algorithms that are of iterative nature, which most Learning-to-Rank algorithms are, are not amenable to the MapReduce framework.
+Lin argued that as a solution to the non-amenability of iterative algorithms to the MapReduce framework, iterative algorithms can often be replaced with non-iterative alternatives or can still be optimized in such a way that its performance in a MapReduce setting is good enough.\\
+
+The appearance of benchmark datasets gave insight in the performance of different Learning-to-Rank approaches and resulted in increasing popularity of those methods that showed to perform well on one or more benchmark datasets.
+Up to now it remains unknown whether popular existing Learning-to-Rank methods scale well when they are used in a parallel manner using the MapReduce paradigm.\\
 

### 2014-02-12 classicthesis/Chapters/Introduction.tex
approach getypt

@@
 \chapter{Research Goals}
-The objective of this thesis is the exploration of execution time speed-up of Learning-to-Rank algorithms through parallelization using Hadoop. 
+The objective of this thesis is the exploration of execution time speed-up of Learning-to-Rank algorithms through parallelization using the MapReduce paradigm. 
 This work focuses on those Learning-to-Rank algorithms that have shown leading performance on relevant benchmark datasets.
@@
 \begin{itemize}
-\item What is the speed-up of existing Learning-to-Rank algorithms when implemented and executed on Hadoop?
+\item What is the speed-up of existing Learning-to-Rank algorithms when executed using the MapReduce paradigm?
 \item Can we adjust those Learning-to-Rank algorithms such that the parallel execution speed-up increases without decreasing accuracy?
@@
 \chapter{Approach}
-Equidem detraxit cu nam, vix eu delenit periculis. Eos ut vero
-constituto, no vidit propriae complectitur sea. Diceret nonummy in
-has, no qui eligendi recteque consetetur. Mel eu dictas suscipiantur,
-et sed placerat oporteat. At ipsum electram mei, ad aeque atomorum
-mea.
+The first research question will be answered by implementing Learning-to-Rank methods in the MapReduce model and measuring the runtime as a factor of the number of cluster nodes used to complete the computation.
+To implement the Learning-to-Rank algorithms I will use cloud based MapReduce implementation from Microsoft was used that is called HDInsight, which is based on the popular MapReduce open source implementation Hadoop\footnote{http://hadoop.apache.org/}.
+The algorithms that we include in the measurements will be determined based on experimental results on the \emph{Yahoo!
+Learning to Rank Challenge}\cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, the LETOR\cite{Qin2010} dataset and the LETOR successor MSLR-WEB30k.
 

### 2014-02-12 classicthesis/BenchmarkResults.tex
structuurswijziging + parttext

\chapter{Yahoo! Learning to Rank Challenge}
\section{Benchmark characteristics}
\section{Results}
\chapter{Yandex Internet Mathematics competition}
\section{Benchmark characteristics}
\section{Results}
\chapter{LETOR}
\section{Benchmark characteristics}
\section{Results}
\chapter{MSLR-WEB30k}
\section{Benchmark characteristics}
\section{Results}

### 2014-02-12 classicthesis/Chapters/Introduction.tex
structuurswijziging + parttext
update thesis outline

@@
 \chapter{Approach}
-The first research question will be answered by implementing Learning-to-Rank methods in the MapReduce model and measuring the runtime as a factor of the number of cluster nodes used to complete the computation.
+The first research question will be answered by implementing Learning-to-Rank methods in the MapReduce model and measuring the runtime as a factor of the number of cluster nodes used to complete the computation.\\
+
 To implement the Learning-to-Rank algorithms I will use cloud based MapReduce implementation from Microsoft was used that is called HDInsight, which is based on the popular MapReduce open source implementation Hadoop\footnote{http://hadoop.apache.org/}.
@@
 \begin{description}
-\item[Part II ]{introduces the reader to the basic principles and recent work in the fields of Learning-to-Rank}
-\item[Part III ]{describes a short overview work in the field of parallel machine learning and Learning-to-Rank.}
-\item[Part IV ]{describes implementation details of the Learning-to-Rank algorithms in the Hadoop framework.}
-\item[Part V ]{presents and discusses speed-up results for the implemented Learning-to-Rank methods.}
-\item[Part VI ]{summarizes the results. The limitations of our research as well as future research directions in the field are mentioned here.}
+\item[Part II ]{introduces the reader to the basic principles and recent work in the fields of Learning-to-Rank.}
+\item[Part III ]{sketches the performance of existing Learning-to-Rank methods on several benchmark datasets.}
+\item[Part IV ]{concisely describes existing work in the field of parallel machine learning and parallel Learning-to-Rank.}
+\item[Part V ]{describes implementation details of the Learning-to-Rank algorithms in the Hadoop framework.}
+\item[Part VI ]{presents and discusses speed-up results for the implemented Learning-to-Rank methods.}
+\item[Part VII ]{summarizes the results. The limitations of our research as well as future research directions in the field are mentioned here.}
 \end{description}

### 2014-02-12 classicthesis/Chapters/Background.tex
structuurswijziging + parttext
logischere opzet background

@@
 
-Learning-to-Rank algorithms can be classified into three approaches: the pointwise approach, the pairwise approach and the listwise approach, on which we will elaborate in the subsequent chapters.
+Learning-to-Rank algorithms can be classified into three approaches: the pointwise approach, the pairwise approach and the listwise approach.
+In the following chapters we will summarize the different approaches in Learning-to-Rank.
 
-\chapter{Benchmark Datasets}
-\chapter{Pointwise Approach}
-\chapter{Pairwise Approach}
-\chapter{Listwise Approach}
-\chapter{Summary}
+\section{Pointwise Approach}
+
+\section{Pairwise Approach}
+\section{Listwise Approach}

### 2014-02-12 classicthesis/Chapters/Introduction.tex
feedback Jurjen tm chapter 5 verwerkt

@@
 Ranking is a core problem in the field of information retrieval.
-The ranking task in information retrieval entails the ranking of candidate documents according to their relevance to a query.
+The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a query.
 Heuristic ranking models have long been around in the information retrieval field.
@@
 
-In recent years several Learning-to-Rank benchmark datasets have been proposed that allow us to compare the performance of different Learning-to-Rank methods.
+In recent years several Learning-to-Rank benchmark datasets have been proposed that enable comparison of the performance of different Learning-to-Rank methods.
 Well-known benchmark datasets include the \emph{Yahoo!
@@
 One of the concluding observations of the \emph{Yahoo!
-Learning to Rank Challenge} was that almost all work in the Learning-to-Rank field focuses on ranking accuracy and that efficiency and scalability of Learning-to-Rank algorithms is still an underexposed research area of which its importance is likely to increase in the near future as training sets are becoming larger and larger\cite{Chapelle2011b}.
+Learning to Rank Challenge} was that almost all work in the Learning-to-Rank field focuses on ranking accuracy, while efficiency and scalability of Learning-to-Rank algorithms is still an underexposed research area that is likely to become more important in the near future as training sets are becoming larger and larger\cite{Chapelle2011b}.
 Liu\cite{Liu2007} confirms the observation that efficiency and scalability of Learning-to-Rank methods has so far been an overlooked research area in his influential book on Learning-to-Rank.\\
@@
 However, almost none of these studies include the Learning-to-Rank subfield of machine learning.
-The field of efficient Learning-to-Rank has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012} since Chapelle at all \cite{Chapelle2011b} stated its growing importance back in 2011, but only few of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient Learning-to-Rank through the use of parallel programming paradigms.
+The field of efficient Learning-to-Rank has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012}, since Liu \cite{Liu2007} first stated its growing importance back in 2007.
+Only several of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient Learning-to-Rank through the use of parallel programming paradigms.\\
+
+MapReduce\cite{Dean2004} is a parallel programming framework that is inspired by the \emph{Map} and \emph{Reduce} functions commonly used in functional programming.
+Since Google developed the MapReduce parallel programming framework back in 2004 it has since grown to be the industry standard model for parallel programming.
 Lin \cite{Lin2013} observed that algorithms that are of iterative nature, which most Learning-to-Rank algorithms are, are not amenable to the MapReduce framework.
@@
 
-The appearance of benchmark datasets gave insight in the performance of different Learning-to-Rank approaches and resulted in increasing popularity of those methods that showed to perform well on one or more benchmark datasets.
-Up to now it remains unknown whether popular existing Learning-to-Rank methods scale well when they are used in a parallel manner using the MapReduce paradigm.\\
+The appearance of benchmark datasets gave insight in the performance of different Learning-to-Rank approaches, which resulted in increasing popularity of those methods that showed to perform well on one or more benchmark datasets.
+Up to now it remains unknown whether popular existing Learning-to-Rank methods scale well when they are used in a parallel manner using the MapReduce framework.
+This thesis aims to be an explorative start in this little researched area of parallel Learning-to-Rank.
+A more extensive overview of my research goals and questions are described in chapter \ref{chap:goals}.\\
 
 \chapter{Research Goals}
-The objective of this thesis is the exploration of execution time speed-up of Learning-to-Rank algorithms through parallelization using the MapReduce paradigm. 
+\label{chap:goals}
+The objective of this thesis is to explore the speed-up in execution time of Learning-to-Rank algorithms through parallelization using the MapReduce framework. 
 This work focuses on those Learning-to-Rank algorithms that have shown leading performance on relevant benchmark datasets.
@@
 \begin{itemize}
-\item What is the speed-up of existing Learning-to-Rank algorithms when executed using the MapReduce paradigm?
+\item What is the speed-up of existing Learning-to-Rank algorithms when executed using the MapReduce framework?
 \item Can we adjust those Learning-to-Rank algorithms such that the parallel execution speed-up increases without decreasing accuracy?
@@
 \chapter{Approach}
-The first research question will be answered by implementing Learning-to-Rank methods in the MapReduce model and measuring the runtime as a factor of the number of cluster nodes used to complete the computation.\\
+To answer the first research question I will implement Learning-to-Rank methods in the MapReduce framework and measuring the runtime as a factor of the number of cluster nodes used to complete the computation.\\
 
-To implement the Learning-to-Rank algorithms I will use cloud based MapReduce implementation from Microsoft was used that is called HDInsight, which is based on the popular MapReduce open source implementation Hadoop\footnote{http://hadoop.apache.org/}.
+To implement the Learning-to-Rank algorithms I will use cloud based MapReduce implementation from Microsoft was used that is called HDInsight.
+It which is based on the popular MapReduce open source implementation Hadoop\footnote{http://hadoop.apache.org/}.
 The algorithms that we include in the measurements will be determined based on experimental results on the \emph{Yahoo!
@@
 \item[Part VI ]{presents and discusses speed-up results for the implemented Learning-to-Rank methods.}
-\item[Part VII ]{summarizes the results. The limitations of our research as well as future research directions in the field are mentioned here.}
+\item[Part VII ]{summarizes the results and answers our research questions based on the results. The limitations of our research as well as future research directions in the field are mentioned here.}
 \end{description}

### 2014-02-12 classicthesis/Chapters/Background.tex
referentiefix & kleine fixes

@@
 \chapter{Introduction to Learning-to-Rank}
-Different definitions of Learning-to-Rank exist. In general, all methods that use machine learning technologies to solve the problem of ranking are called Learning-to-Rank methods.\\
+Different definitions of Learning-to-Rank exist. In general, all ranking methods that use machine learning technologies to solve the problem of ranking are called Learning-to-Rank methods.\\
 Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning-to-Rank method when it meets the following two properties:

### 2014-02-12 classicthesis/Chapters/Background.tex
feedback background sectie verwerkt

--- 2014-02-12 classicthesis/Chapters/Background.tex
@@
 \chapter{Introduction to Learning-to-Rank}
-Different definitions of Learning-to-Rank exist. In general, all methods that use machine learning technologies to solve the problem of ranking are called Learning-to-Rank methods.\\
-Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning-to-Rank method when it meets the following two properties:
+Different definitions of Learning-to-Rank exist.
+In general, all ranking methods that use machine learning technologies to solve the problem of ranking are called Learning-to-Rank methods.
+Figure \ref{fig:discriminative_training} describes the general process of machine learning.
+It shows training elements from an input space that are mapped to an output space using a model such that the difference between the actual labels of the training elements and the labels predicted with with the model are minimal in terms of a loss function.
+\begin{figure}[!h]
+\includegraphics[scale=0.26]{gfx/descriminative_training}
+\caption{Machine learning framework for Learning-to-Rank, from Liu\cite{Liu2007}}
+\label{fig:discriminative_training}
+\end{figure}\\
+Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning-to-Rank method when it is \emph{feature based} and uses \emph{discriminative training}, which are itself defined as follows:
 \begin{description}
 \item[Feature Based]{\emph{Feature based} means that all documents under investigation are represented by feature vectors that reflect the relevance of the documents to the query.}
-\item[Discriminative Training]{\emph{Discriminative training} means that the learning process can be well described by the four components of discriminative learning, that is, a Learning-to-Rank method has its own input space, output space, hypothesis space, and loss function.
-Input space, output space, hypothesis space, and loss function are hereby defined as follows:
+\item[Discriminative Training]{\emph{Discriminative training} means that the learning process can be well described by the four components of discriminative learning.
+That is, a Learning-to-Rank method has its own \emph{input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function}, like the machine learning process described by Figure \ref{fig:discriminative_training}.
+\emph{Input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function} are hereby defined as follows:
 	\begin{description}
@@
 The functions operate on the feature vectors of the input object, and make predictions according to the format of the output space.}
-	\item[Loss Function]{In order to learn the optimal hypothesis, a training set is usually used, which contains a number of objects and their ground truth labels, sampled from the product of the input and output spaces.}
+	\item[Loss Function]{in order to learn the optimal hypothesis, a training set is usually used, which contains a number of objects and their ground truth labels, sampled from the product of the input and output spaces.}
 	\end{description}
@@
 \end{description}
-\begin{figure}[!h]
-\includegraphics[scale=0.26]{gfx/descriminative_training}
-\caption{Machine learning framework for Learning-to-Rank, from Liu\cite{Liu2007}}
-\label{fig:discriminative_training}
-\end{figure}
-Figure \ref{fig:discriminative_training} shows the relationships between input space, output space, hypothesis space and loss function.
 
-Figure \ref{ref:ltr_framework} shows how the learning process in Learning-to-Rank typically happens.
-A set of queries $q_i$ with $n > i > 1$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgments of those documents $y_i$ are inputted to a learning algorithm that learns a ranking model such that the output of the ranking model can predict relevant judgment labels in this training set as accurately as possible.
-Later on we can test the accuracy of this ranking model by predicting the relevant judgments new documents given a new query.
+Figure \ref{ref:ltr_framework} shows how the machine learning process as described in Figure \ref{fig:discriminative_training} typically takes place in a ranking scenario.
+A set of queries $q_i$ with $n > i > 1$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgments of those documents $y_i$ are used together to train a model $h$, that can predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is are minimal in terms of a loss function.
 \begin{figure}[!h]
@@
 
-Learning-to-Rank algorithms can be classified into three approaches: the pointwise approach, the pairwise approach and the listwise approach.
-In the following chapters we will summarize the different approaches in Learning-to-Rank.
+The predictions and the loss function might either be defined for:
+\begin{enumerate}
+\item the relevance of a single document
+\item the classification of the most relevant document out of a document-pair
+\item the ranking of documents directly
+\end{enumerate}
+These three approaches are in literature respectively called the pointwise approach, the pairwise approach and the listwise approach.
+In the following sections we will further describe the three different approaches in Learning-to-Rank.
 

### 2014-02-12 classicthesis/Chapters/Introduction.tex
laatste puntjes netjes gemaakt voordat Sander de draft kreeg

@@
 \begin{description}
-\item[Part II ]{introduces the reader to the basic principles and recent work in the fields of Learning-to-Rank.}
-\item[Part III ]{sketches the performance of existing Learning-to-Rank methods on several benchmark datasets.}
-\item[Part IV ]{concisely describes existing work in the field of parallel machine learning and parallel Learning-to-Rank.}
-\item[Part V ]{describes implementation details of the Learning-to-Rank algorithms in the Hadoop framework.}
-\item[Part VI ]{presents and discusses speed-up results for the implemented Learning-to-Rank methods.}
-\item[Part VII ]{summarizes the results and answers our research questions based on the results. The limitations of our research as well as future research directions in the field are mentioned here.}
+\item[Part II: Background ]{introduces the reader to the basic principles and recent work in the fields of Learning-to-Rank.}
+\item[Part III: Related Work]{concisely describes existing work in the field of parallel machine learning and parallel Learning-to-Rank.}
+\item[Part IV: Benchmark Results]{sketches the performance of existing Learning-to-Rank methods on several benchmark datasets and describes the selection of Learning-to-Rank methods for the parallelization experiments.}
+\item[Part V: Selected Learning-to-Rank Methods]{describes the algorithms and details of the selected Learning-to-Rank methods.}
+\item[Part VI: Implementation]{describes implementation details of the Learning-to-Rank algorithms in the Hadoop framework.}
+\item[Part VII: Results \& Discussion]{presents and discusses speed-up results for the implemented Learning-to-Rank methods.}
+\item[Part VIII: Conclusion]{summarizes the results and answers our research questions based on the results.
+The limitations of our research as well as future research directions in the field are mentioned here.}
 \end{description}

### 2014-02-13 classicthesis/BenchmarkResults.tex
begin Yahoo benchmark + stukje IR geschiedenisles

@@
 \chapter{Yahoo! Learning to Rank Challenge}
+Yahoo's observation that all existing benchmark datasets were too small to draw reliable conclusions, especially in comparison with datasets used in commercial search engines, prompted Yahoo to release two internal datasets from Yahoo!
+search.
+The Yahoo!
+Learning to Rank Challenge\cite{Chapelle2011a} is a public Learning-to-Rank competition which took place from March to May 2010, with the goal to promote the datasets and encourage the research community to develop new Learning-to-Rank algorithms.\\
+
+The Yahoo!
+Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning-to-Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country.
+For this experiment I will only look at the standard Learning-to-Rank dataset because transfer learning is a separate research area that is not included in this thesis.
 \section{Benchmark characteristics}
+\begin{description}
+\item 36k queries
+\item 883k documents
+\item 700 features
+\end{description}
+
 \section{Results}

### 2014-02-13 classicthesis/Chapters/Introduction.tex
begin Yahoo benchmark + stukje IR geschiedenisles

@@
 The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a query.
-Heuristic ranking models have long been around in the information retrieval field.
+Research in the field of ranking models has for a long time been based on manually designed ranking functions and has been an active area of research since Luhn\cite{Luhn1957} was the first to propose model that assigned relevance scores to documents given a query back in 1957.
 The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models.

### 2014-02-13 classicthesis/Chapters/Introduction.tex
typo gefixt

--- 2014-02-12 classicthesis/Chapters/Introduction.tex
@@
 The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a query.
-Heuristic ranking models have long been around in the information retrieval field.
+Research in the field of ranking models has for a long time been based on manually designed ranking functions and has been an active area of research since Luhn\cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
 The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models.

### 2014-02-13 classicthesis/BenchmarkResults.tex
meer over benchmarks

@@
 Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning-to-Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country.
-For this experiment I will only look at the standard Learning-to-Rank dataset because transfer learning is a separate research area that is not included in this thesis.
-\section{Benchmark characteristics}
-\begin{description}
-\item 36k queries
-\item 883k documents
-\item 700 features
-\end{description}
+For this experiment I will only look at the standard Learning-to-Rank dataset because transfer learning is a separate research area that is not included in this thesis.\\
 
+Model validation on the Learning-to-Rank methods perticipanting in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}.
+The large number of documents, queries and features compared to other benchmark datasets makes the Yahoo!
+Learning to Rank Challenge dataset interesting.
+Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo.
+Instead high level descriptions of feature categories are provided.\\
+
+The performance on the test-set is measured in terms of nDCG and ERR.
+
+\begin{table}
+\begin{tabular}{llll}
+ & Train & Validation & Test \\ 
+Queries & 19,994 & 2,994 & 6,983 \\ 
+Documents & 473,134 & 71,083 & 165,660 \\ 
+Features & 519 & - & - \\ 
+\end{tabular}
+\caption{Yahoo! Learning to Rank Challende dataset characteristics, as described in the challenge overview paper\cite{Chapelle2011a}}
+\label{tab:yahoo_characteristics}
+\end{table}
 \section{Results}
+\begin{table}
+\begin{tabular}{lll}
+ & Authors & ERR \\ 
+1 & Burges et al (Microsoft Research) & 0.46861 \\ 
+2 & Gottschalk (Activision Blizzard) \& Vogel (Data Mining Solutions) & 0.46786 \\ 
+3 & Parakhin (Microsoft) & 0.46695 \\ 
+4 & Pavlov \& Brunk (Yandex Labs) & 0.46678 \\ 
+5 & Sorokina (Yandex Labs) & 0.46616 \\ 
+\end{tabular}
+\caption{Final standings of the Yahoo! Learning to Rank Challenge, as presented in the challenge overview paper\cite{Chapelle2011a}}
+\end{table}
 \chapter{Yandex Internet Mathematics competition}
@@
 \chapter{MSLR-WEB30k}
+Contrary to the Yahoo!
+Learning to Rank Challenge dataset, the MSLR-WEB30k provides detailed feature descriptions.
+The MSLR-WEB30k dataset however contains no proprietary features but only features that are commonly used in the research community.
 \section{Benchmark characteristics}

### 2014-02-13 classicthesis/Chapters/Background.tex
meer over benchmarks

@@
 
-Figure \ref{ref:ltr_framework} shows how the machine learning process as described in Figure \ref{fig:discriminative_training} typically takes place in a ranking scenario.
+Figure \ref{fig:ltr_framework} shows how the machine learning process as described in Figure \ref{fig:discriminative_training} typically takes place in a ranking scenario.
 A set of queries $q_i$ with $n > i > 1$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgments of those documents $y_i$ are used together to train a model $h$, that can predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is are minimal in terms of a loss function.
@@
 \caption{A typical Learning-to-Rank setting, from Liu\cite{Liu2007}}
-\label{ref:ltr_framework}
+\label{fig:ltr_framework}
 \end{figure}\\
@@
 \section{Listwise Approach}
+\section{Evaluation Methods}
+\subsection{Normalized Discounted Cumulative Gain}
+\subsubsection{Cumulative Gain}
+\subsubsection{Discounted Cumulative Gain}
+\subsubsection{Normalized Discounted Cumulative Gain}

### 2014-02-13 classicthesis/BenchmarkResults.tex
begin feature omschrijving

@@
 Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo.
-Instead high level descriptions of feature categories are provided.\\
+Instead high level descriptions of feature categories are provided.
+The following categories of features are described in the challenge overview paper\cite{Chapelle2011a}:\\
+\begin{description}
+\item[Web graph]{Quality and popularity metrics of web documents, e.g. PageRank\cite{Page1999}}.
+\item[Document statistics]{Basic document statistics such as the number of words and url characteristics.}
+\item[Document classifier]{Results of various classifiers on the documents. These classifiers amongst others include: spam, adult, language, main topic, and quality classifiers.}
+\item[Query]{Basic query statistics, such as the number of terms, query frequency, and click-through rate.}
+\item[Text match]{Textual similarity metrics between query and document. Includes TF-IDF, BM25\cite{Robertson2009} and other metrics for different sections of the document.}
+\item[Topical matching]{}
+\item[Click]{}
+\item[External references]{}
+\item[Time]{}
+\end{description}
 

### 2014-02-13 classicthesis/BenchmarkResults.tex
begin acronyms

@@
 For this experiment I will only look at the standard Learning-to-Rank dataset because transfer learning is a separate research area that is not included in this thesis.\\
-
+\begin{table}[!h]
+\begin{tabular}{l|lll}
+ & Train & Validation & Test \\ 
+ \hline
+Queries & 19,994 & 2,994 & 6,983 \\ 
+Documents & 473,134 & 71,083 & 165,660 \\ 
+Features & 519 & - & - \\ 
+\end{tabular}
+\caption{Yahoo! Learning to Rank Challende dataset characteristics, as described in the challenge overview paper\cite{Chapelle2011a}}
+\label{tab:yahoo_characteristics}
+\end{table}\\
 Model validation on the Learning-to-Rank methods perticipanting in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}.
@@
 \item[Query]{Basic query statistics, such as the number of terms, query frequency, and click-through rate.}
-\item[Text match]{Textual similarity metrics between query and document. Includes TF-IDF, BM25\cite{Robertson2009} and other metrics for different sections of the document.}
-\item[Topical matching]{}
-\item[Click]{}
-\item[External references]{}
-\item[Time]{}
+\item[Text match]{Textual similarity metrics between query and document. Includes \ac{TF-IDF}, BM25\cite{Robertson2009} and other metrics for different sections of the document.}
+\item[Topical matching]{These features go beyond similarity at word level and compute similarity on topic level.
+For example by classifying both the document and the query in a large topical taxonomy.}
+\item[Click]{Click-based user feedback.}
+\item[External references]{Document meta-information such as Delicious\footnote{https://delicious.com/} tags}
+\item[Time]{Document age and historical in- and outlink data that might help for time sensitive queries.}
 \end{description}
@@
 
+\section{Results}
+Figure \ref{fig:yahoo_results} shows the top five participants in the Yahoo! Learning to Rank Challenge in terms of ERR score.
 \begin{table}
-\begin{tabular}{llll}
- & Train & Validation & Test \\ 
-Queries & 19,994 & 2,994 & 6,983 \\ 
-Documents & 473,134 & 71,083 & 165,660 \\ 
-Features & 519 & - & - \\ 
-\end{tabular}
-\caption{Yahoo! Learning to Rank Challende dataset characteristics, as described in the challenge overview paper\cite{Chapelle2011a}}
-\label{tab:yahoo_characteristics}
-\end{table}
-\section{Results}
-\begin{table}
-\begin{tabular}{lll}
+\begin{tabular}{l|p{6.3cm}|l}
  & Authors & ERR \\
+ \hline 
 1 & Burges et al (Microsoft Research) & 0.46861 \\ 
@@
 \caption{Final standings of the Yahoo! Learning to Rank Challenge, as presented in the challenge overview paper\cite{Chapelle2011a}}
+\label{fig:yahoo_results}
 \end{table}

### 2014-02-13 classicthesis/Chapters/Background.tex
begin acronyms

@@
 \includegraphics[scale=0.26]{gfx/descriminative_training}
-\caption{Machine learning framework for Learning-to-Rank, from Liu\cite{Liu2007}}
+\caption{Machine learning framework for Learning-to-Rank, obtained from Liu\cite{Liu2007}}
 \label{fig:discriminative_training}
@@
 \includegraphics[scale=0.25]{gfx/ltr_framework}
-\caption{A typical Learning-to-Rank setting, from Liu\cite{Liu2007}}
+\caption{A typical Learning-to-Rank setting, obtained from Liu\cite{Liu2007}}
 \label{fig:ltr_framework}
@@
 \subsubsection{Discounted Cumulative Gain}
+\ac{DCG}
 \subsubsection{Normalized Discounted Cumulative Gain}
+\ac{nDCG}
+\subsection{Expected Reciprocal Rank}
+\ac{ERR}

### 2014-02-13 classicthesis/BenchmarkResults.tex
nu loopt het een beetje

@@
 \end{table}\\
+Both \ac{nDCG} and \ac{ERR} are measured as performance metrics, but the final standings of the challenge were based on the \ac{ERR} values.
 Model validation on the Learning-to-Rank methods perticipanting in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}.
+Competitors could train on the training set and get immediate feedback on their performance on the validation set.
+The test set performance is used to create the final standings and is only measured after the competition has ended to avoid overfitting on the test set.
 The large number of documents, queries and features compared to other benchmark datasets makes the Yahoo!
@@
 \end{description}
-
-The performance on the test-set is measured in terms of nDCG and ERR.
 

### 2014-02-13 classicthesis/BenchmarkResults.tex
yahoo results + bibliography

@@
 \section{Results}
-Figure \ref{fig:yahoo_results} shows the top five participants in the Yahoo! Learning to Rank Challenge in terms of ERR score.
+Figure \ref{fig:yahoo_results} shows the top five participants in the Yahoo!
+Learning to Rank Challenge in terms of ERR score.
+The top five participants all used decision trees and ensemble methods.
+Burges et al\cite{Burges2011} created a linear combination ensemble of eight LambdaMART\cite{Burges2010}, two LambdaRank and two Logistic Regression models.
+Gottschalk and Vogel used a combination of RandomForest models and \ac{GBDT} models.
+Pavlov and Brunk used a regression based model using the BagBoo\cite{Pavlov2010} ensemble technique, which combines bagging and boosting.
+Sorokina used a similar combination of bagging and boosting that is called Additive Groves\cite{Sorokina2007}.\\
+
 \begin{table}

### 2014-02-14 classicthesis/BenchmarkResults.tex
yahoo klaar
klein begin LETOR + approach literature study

@@
 
+The challenge overview paper states as one of the lessons of the challenge that the simple baseline \ac{GBDT} model performed very well with a small performance gap to the complex ensemble submissions at the top of the table.
+
 \begin{table}
@@
 \end{table}
+
 \chapter{Yandex Internet Mathematics competition}
@@
 \section{Results}
+
 \chapter{LETOR}
+The LETOR dataset was first released by Microsoft Research Asia in April 2007 to solve the absence of a experimental platform for Learning-to-Rank at that time.
+LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0
 \section{Benchmark characteristics}
 \section{Results}
+
 \chapter{MSLR-WEB30k}

### 2014-02-14 classicthesis/Chapters/Introduction.tex
klein begin LETOR + approach literature study

@@
 \chapter{Approach}
+A literature study will be performed to get insight in relevant existing techniques for large scale Learning-to-Rank. The literature study will be performed by using the following query:
+\begin{itemize}
+\item ("learning to rank" OR "learning-to-rank" OR "machine learned ranking") AND ("large scale" OR "parallel" OR "distributed")
+\end{itemize}
+and the following bibliographic databases:
+\begin{itemize}
+\item Scopus
+\item Web of Science
+\end{itemize}
+
 To answer the first research question I will implement Learning-to-Rank methods in the MapReduce framework and measuring the runtime as a factor of the number of cluster nodes used to complete the computation.\\

### 2014-02-17 classicthesis/Chapters/Background.tex
US english -> GB english

@@
 Figure \ref{fig:ltr_framework} shows how the machine learning process as described in Figure \ref{fig:discriminative_training} typically takes place in a ranking scenario.
-A set of queries $q_i$ with $n > i > 1$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgments of those documents $y_i$ are used together to train a model $h$, that can predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is are minimal in terms of a loss function.
+A set of queries $q_i$ with $n > i > 1$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgements of those documents $y_i$ are used together to train a model $h$, that can predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is are minimal in terms of a loss function.
 \begin{figure}[!h]

### 2014-02-17 classicthesis/Chapters/Introduction.tex
US english -> GB english

@@
 Some research has been done in the area of on parallel or distributed machine learning \cite{Chu2007,Chang2007}.
-However, almost none of these studies include the Learning-to-Rank subfield of machine learning.
+However, almost none of these studies include the Learning-to-Rank sub-field of machine learning.
 The field of efficient Learning-to-Rank has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012}, since Liu \cite{Liu2007} first stated its growing importance back in 2007.
@@
 Up to now it remains unknown whether popular existing Learning-to-Rank methods scale well when they are used in a parallel manner using the MapReduce framework.
-This thesis aims to be an explorative start in this little researched area of parallel Learning-to-Rank.
+This thesis aims to be an exploratory start in this little researched area of parallel Learning-to-Rank.
 A more extensive overview of my research goals and questions are described in chapter \ref{chap:goals}.\\
@@
 \label{chap:goals}
-The objective of this thesis is to explore the speed-up in execution time of Learning-to-Rank algorithms through parallelization using the MapReduce framework. 
+The objective of this thesis is to explore the speed-up in execution time of Learning-to-Rank algorithms through parallelisation using the MapReduce framework. 
 This work focuses on those Learning-to-Rank algorithms that have shown leading performance on relevant benchmark datasets.
@@
 \begin{description}
-\item[Part II: Background ]{introduces the reader to the basic principles and recent work in the fields of Learning-to-Rank.}
+\item[Part II: Background]{introduces the reader to the basic principles and recent work in the fields of Learning-to-Rank.}
 \item[Part III: Related Work]{concisely describes existing work in the field of parallel machine learning and parallel Learning-to-Rank.}
-\item[Part IV: Benchmark Results]{sketches the performance of existing Learning-to-Rank methods on several benchmark datasets and describes the selection of Learning-to-Rank methods for the parallelization experiments.}
+\item[Part IV: Benchmark Results]{sketches the performance of existing Learning-to-Rank methods on several benchmark datasets and describes the selection of Learning-to-Rank methods for the parallelisation experiments.}
 \item[Part V: Selected Learning-to-Rank Methods]{describes the algorithms and details of the selected Learning-to-Rank methods.}

### 2014-02-17 classicthesis/RelatedWork.tex
US english -> GB english

\chapter{Related Work}
\section{Search characteristics}
The literature research is performed by using the bibliographic databases Scopus and Web of Science with the following search query: \emph{("learning to rank" OR "learning-to-rank" OR "machine learned ranking") AND ("large scale" OR "parallel" OR "distributed")}. An abstract based manual filtering step is applied where I filter those results that actually use the \emph{large scale}, \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}. Studies focusing on efficient query evaluation instead of efficient model training are likely to meet all criteria listed. As a last step I will filter out studies focusing on efficient query evaluation.
\subsection{Scopus}
Defined search query resulted in 65 documents.
\subsection{Web of Science}
\section{CCRank}
Wang et al\cite{Wang2011} propose a parallel evolutionary algorithm based on \ac{CC}, a type of evolutionary algorithm. The \ac{CC} algorithm is capable of directly optimizing non-differentiable functions, as \ac{nDCG}, in contrary to many optimization algorithms.  In \ac{CC} algorithms to evolution process can be naturally parallelised

### 2014-02-17 classicthesis/RelatedWork.tex
spark & cooperative coevolution

@@
 The \ac{CC} algorithm is capable of directly optimizing non-differentiable functions, as \ac{nDCG}, in contrary to many optimization algorithms.
-In \ac{CC} algorithms to evolution process can be naturally parallelised
+the divide-and-conquer nature of the \ac{CC} algorithm enables parallelisation.
+CCRank showed an increase in both accuracy and efficiency on the LETOR 4.0 benchmark dataset compared to the baselines.
+It must be stated however that the increased efficiency was achieved through speed-up and not scale-up.
+Two reasons have been identified for not achieving linear scale-up with CCRank: 1) parallel execution is suspended after each generation to perform combination in order to produce the candidate solution, 2) Combination has to wait until all parallel tasks have finished, which may spend different running time.
+\section{Parallel ListNet using Spark}
+Shukla et al\cite{Shukla2012} explored the parallelisation of the well-known ListNet Learning-to-Rank method using Spark.
+Spark is a parallel computing model that is designed for cyclic data flows which makes it more suitable for iterative algorithms.
+Spark is incorporated into Hadoop since Hadoop 2.0.
+The Spark implementation of ListNet showed near a linear training time reduction.

### 2014-02-17 classicthesis/RelatedWork.tex
GBDT & nDCG-annealing

@@
 \section{CCRank}
-Wang et al\cite{Wang2011} propose a parallel evolutionary algorithm based on \ac{CC}, a type of evolutionary algorithm.
+Wang et al\cite{Wang2011a,Wang2011b} propose a parallel evolutionary algorithm based on \ac{CC}, a type of evolutionary algorithm.
 The \ac{CC} algorithm is capable of directly optimizing non-differentiable functions, as \ac{nDCG}, in contrary to many optimization algorithms.
@@
 The Spark implementation of ListNet showed near a linear training time reduction.
+\section{Gradient Boosted Distributed Decision Trees}
+Ye et al\cite{Ye2009} described how to implement the \ac{GBDT} process in a parallel manner using both MPI and Hadoop.
+\ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise\cite{Zheng2007} or listwise\cite{Wang2011} setting.
+Experiments showed the Hadoop implementation to result into too expensive communication cost to be useful.
+Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
+The MPI approach proved to be successful and obtained near linear speed-ups.
+\section{nDCG-Annealing}
+Karimzadeghan el al\cite{Karimzadehgan2011} proposed a method using Simulated Annealing along with the Simplex method for its parameter search.
+This method directly optimises the often non-differentiable Learning-to-Rank evaluation metrics like \ac{nDCG} and \ac{MAP}.
+The authors successfully parallelised their method in the MapReduce paradigm using Hadoop.
+The approach showed to be effective on both the LETOR 3.0 dataset and their own dataset with contextual advertising data.
+Unfortunately their work does not directly report on the speed-up obtained by parallelising  with Hadoop, but it is mentioned that further work needs to be done to effectively leverage parallel execution.

### 2014-02-17 classicthesis/RelatedWork.tex
low complexity learning to rank

@@
 Unfortunately their work does not directly report on the speed-up obtained by parallelising  with Hadoop, but it is mentioned that further work needs to be done to effectively leverage parallel execution.
+\section{Low complexity learning}
+Designing Learning-to-Rank algorithms with low time complexity for training is a different another approach towards large scale Learning-to-Rank.
+Pahikkala et al\cite{Pahikkala2009} describe a pairwise \ac{RLS} type of ranking function, RankRLS, with time complexity $\mathcal{O}(n^3+n^2m)$, with $n$ the number of features and $m$ the number of training documents.
+The RankRLS ranking function showed ranking performance similar to RankSVM on the BioInfer corpus\cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.

### 2014-02-17 classicthesis/RelatedWork.tex
distributed stochastic gradient descent

@@
 Unfortunately their work does not directly report on the speed-up obtained by parallelising  with Hadoop, but it is mentioned that further work needs to be done to effectively leverage parallel execution.
-\section{Low complexity learning}
+\section{Low complexity Learning-to-Rank}
 Designing Learning-to-Rank algorithms with low time complexity for training is a different another approach towards large scale Learning-to-Rank.
@@
 The RankRLS ranking function showed ranking performance similar to RankSVM on the BioInfer corpus\cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.
+\section{Distributed Stochastic Gradient Descent}
+Long et al\cite{Long2012} describe special case of their pairwise cross-domain factor Learning-to-Rank method using distributed optimization of \ac{SGD} based on Hadoop MapReduce.
+Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce based method described by  Zinkevich et al\cite{Zinkevich2010} has been used.
+Real world data from Yahoo!
+has been used to show that the model is effective.
+Unfortunately the speed-up obtained by training their model in parallel is not reported.

### 2014-02-18 classicthesis/RelatedWork.tex
late night work, WoS
FPGA apprach

@@
 \subsection{Scopus}
-Defined search query resulted in 65 documents.
+The defined search query resulted in 65 documents.
 \subsection{Web of Science}
+The defined search query resulted in 16 documents. Three of those documents were also present in the set of 65 documents found using Scopus.
 \section{CCRank}
@@
 Unfortunately the speed-up obtained by training their model in parallel is not reported.
+\section{FPGA-based LambdaRank}
+Yan et al\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run the Neural Network based LambdaRank Learning-to-Rank algorithm.
+This architecture achieved a 29.3X speed-up compared to the software implementation when evaluated on data from a commercial search engine.
+The exploration of \ac{FPGA} for Learning-to-Rank showed other advantages of the \ac{FPGA} approach next to faster model training.
+In their latest publication\cite{Yan2012} the FPGA based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.

### 2014-02-18 classicthesis/RelatedWork.tex
related work 'klaar'

@@
 Studies focusing on efficient query evaluation instead of efficient model training are likely to meet all criteria listed.
-As a last step I will filter out studies focusing on efficient query evaluation.
+As a last step I will filter out studies based on the whole document that only focus on efficient query evaluation and not on parallel or distributed learning of ranking functions.
 \subsection{Scopus}
 The defined search query resulted in 65 documents.
+Only 14 of those documents used \emph{large scale}, \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}.
+10 out of those 14 documents focussed on parallel or distributed learning of ranking functions.
 \subsection{Web of Science}
-The defined search query resulted in 16 documents. Three of those documents were also present in the set of 65 documents found using Scopus.
+The defined search query resulted in 16 documents.
+Four of those documents were also present in the set of 65 documents found using Scopus, leaving 61 unique documents.
+Only four of those 61 documents used \emph{large scale}, \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}, none of them focussed on parallel or distributed learning of ranking functions.
 \section{CCRank}
@@
 In their latest publication\cite{Yan2012} the FPGA based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.
+\section{GPGPU for Learning-to-Rank}
+De Sousa et al\cite{DeSousa2012} proposed a \ac{GPGPU} approach using the \ac{GPU} both learning the ranking function and for query processing, thereby improving both training time and query time.
+An association rule based Learning-to-Rank approach proposed by \cite{Veloso2008} has been implemented using the GPU in such a way that the set of rules van be computed in parallel, in different threads, for each document.
+A speed-up of 127X in query processing time is reported based on evaluation on the LETOR dataset.
+The speed-up achieved at learning the ranking function was unfortunately not stated.

### 2014-02-20 report/BenchmarkResults.tex
hernoeming mappen

--- 2014-02-14 classicthesis/BenchmarkResults.tex
@@
 \end{tabular}
-\caption{Yahoo! Learning to Rank Challende dataset characteristics, as described in the challenge overview paper\cite{Chapelle2011a}}
+\caption{Yahoo! Learning to Rank Challenge dataset characteristics, as described in the challenge overview paper\cite{Chapelle2011a}}
 \label{tab:yahoo_characteristics}
@@
 Both \ac{nDCG} and \ac{ERR} are measured as performance metrics, but the final standings of the challenge were based on the \ac{ERR} values.
-Model validation on the Learning-to-Rank methods perticipanting in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}.
+Model validation on the Learning-to-Rank methods participating in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}.
 Competitors could train on the training set and get immediate feedback on their performance on the validation set.

### 2014-02-20 report/Chapters/Background.tex
hernoeming mappen

--- 2014-02-17 classicthesis/Chapters/Background.tex
@@
 \section{Listwise Approach}
-\section{Evaluation Methods}
+\chapter{Evaluation Methods}
+
+\section{Normalized Discounted Cumulative Gain}
+Cumulative gain, or its predecessor discounted cumulative gain and normalized discounted cumulative gain, is one of the most widely used measures for effectiveness of ranking methods.
+\subsection{Discounted Cumulative Gain}
+The \ac{DCG}\cite{Burges2005} at a position $p$ is defined as\\
+$DCG_p = \sum\nolimits_{i=1}^p \frac{2^{rel_i-1}}{log_2(i+1)}$\\
+with $rel_i$ the graded relevance of the result at position $i$.
+The idea is that highly relevant documents that appear lower in a search result should be penalized (discounted).
+This discounting is done by reducing the graded relevance  logarithmically proportional to the position of the result.
 \subsection{Normalized Discounted Cumulative Gain}
-\subsubsection{Cumulative Gain}
-\subsubsection{Discounted Cumulative Gain}
-\ac{DCG}
-\subsubsection{Normalized Discounted Cumulative Gain}
-\ac{nDCG}
-\subsection{Expected Reciprocal Rank}
-\ac{ERR}
+\ac{nDCG} normalizes the \ac{DCG} metric to a value in the [0,1] interval by dividing by the \ac{DCG} value of the optimal rank.
+This optimal rank is obtained by sorting documents on relevance for a given query.
+We can write the definition of \ac{nDCG} down mathematically as\\
+$nDCG_p = \frac{DCG_p}{IDCG_p}$
+\section{Expected Reciprocal Rank}
+\ac{ERR}\cite{Chapelle2009} was designed based on the observation that \ac{nDCG} is based on the false assumption that the usefulness of a document at rank $i$ is independent of the usefulness of the documents at rank less than $i$.
+\ac{ERR} is based on the reasoning that users are likely to stop exploring the result list once they have found a document that satisfied their information need.
+The \ac{ERR} metric is defined as the expected reciprocal length of time that the user will take to find a relevant document.
+\ac{ERR} is formally defined as\\
+$ERR = \sum\nolimits_{r=1}^n \frac{1}{r} \prod\nolimits_{i=1}^{r-1}(1-R_i)R_i$\\
+where the product sequence part of the formula represents the chance that the user will stop at position $r$.\\
+An algorithm to compute \ac{ERR} is shown in Listing \ref{lst:err}.
+The algorithm requires relevance grades $g_i$, $1 \le i \le n$ and mapping function $R$ that maps relevance grades to probability of relevance.
+\begin{lstlisting}[caption={Algorithm to compute the ERR metric, obtained from \cite{Chapelle2009}}, label={lst:err},language=Ada]
+p <- 1, ERR <- 0
+for r=1 to n do
+	R <- R(g[r])
+	ERR <- ERR + p * R/r
+	p <- p * (1-R)
+end for
+return ERR
+\end{lstlisting}
+\section{Mean Average Precision}
+\ac{AP}\cite{Zhu2004} is an often used binary relevance judgement based metric that can be seen as a trade-off between precision and recall that is defined as\\
+$AP = \frac{\sum\nolimits_{k=1}^{n}Precision(k)*relevance(k)}{\text{number of relevant docs}}$\\
+With $k$ being the positions in the result set between 1 and $n$.
+\ac{MAP} is the average \ac{AP} for a set of queries.\\
+$MAP = \frac{\sum\nolimits_{q=1}^{Q}AP(q)}{Q}$

### 2014-02-20 report/RelatedWork.tex
hernoeming mappen

--- 2014-02-18 classicthesis/RelatedWork.tex
@@
 The exploration of \ac{FPGA} for Learning-to-Rank showed other advantages of the \ac{FPGA} approach next to faster model training.
-In their latest publication\cite{Yan2012} the FPGA based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.
+In their latest publication\cite{Yan2012} the \ac{FPGA} based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.
 \section{GPGPU for Learning-to-Rank}
 De Sousa et al\cite{DeSousa2012} proposed a \ac{GPGPU} approach using the \ac{GPU} both learning the ranking function and for query processing, thereby improving both training time and query time.
-An association rule based Learning-to-Rank approach proposed by \cite{Veloso2008} has been implemented using the GPU in such a way that the set of rules van be computed in parallel, in different threads, for each document.
+An association rule based Learning-to-Rank approach proposed by \cite{Veloso2008} has been implemented using the \ac{GPU} in such a way that the set of rules van be computed in parallel, in different threads, for each document.
 A speed-up of 127X in query processing time is reported based on evaluation on the LETOR dataset.

### 2014-02-21 report/Chapters/Introduction.tex
referenties opgeschoond + definitie speed-up

--- 2014-02-17 classicthesis/Chapters/Introduction.tex
@@
 
-Some research has been done in the area of on parallel or distributed machine learning \cite{Chu2007,Chang2007}.
+Some research has been done in the area of parallel or distributed machine learning \cite{Chu2007,Chang2007}.
 However, almost none of these studies include the Learning-to-Rank sub-field of machine learning.
@@
 \begin{itemize}
-\item What is the speed-up of existing Learning-to-Rank algorithms when executed using the MapReduce framework?
+\item What is the speed-up of existing Learning-to-Rank algorithms when executed using the MapReduce framework?\\
+In which Nussbaum and Agarwal's definition of asymptotic speed-up will be used for speed-up $S(s)$\cite{Nussbaum1991}:\\
+
+\begin{math}
+s \equiv \text{problem size}\\
+T_{seq}(s) \equiv \Theta(\text{Serial Running Time})\\
+T_{par}(s) \equiv \Theta(\text{Minimum Parallel Running Time})\\
+S(s) \equiv \frac{T_{seq}(s)}{T_{par}(s)}\\
+\end{math}
+
 \item Can we adjust those Learning-to-Rank algorithms such that the parallel execution speed-up increases without decreasing accuracy?

### 2014-02-21 report/Chapters/Introduction.tex
nu wel een goede speedup definitie :)

--- 2014-02-17 classicthesis/Chapters/Introduction.tex
@@
 \begin{itemize}
-\item What is the speed-up of existing Learning-to-Rank algorithms when executed using the MapReduce framework?
+\item What is the speed-up of existing Learning-to-Rank algorithms when executed using the MapReduce framework?\\
+Where the definition of \emph{relative speed-up} is used for speed-up\cite{Sun1991}:\\
+
+$S_N = \frac{\text{execution time using one core}}{\text{execution time using \emph{N} cores}}$
+
 \item Can we adjust those Learning-to-Rank algorithms such that the parallel execution speed-up increases without decreasing accuracy?

### 2014-02-21 report/Chapters/Background.tex
referenties opgeschoond + definitie speed-up
begin uiteenzetting approaches L2R :)

@@
 %************************************************
-\chapter{Introduction to Learning-to-Rank}
+\chapter{A basic introduction to Learning-to-Rank}
 Different definitions of Learning-to-Rank exist.
@@
 These three approaches are in literature respectively called the pointwise approach, the pairwise approach and the listwise approach.
-In the following sections we will further describe the three different approaches in Learning-to-Rank.
+These three approaches to Learning-to-Rank will be described in more detail further on in this part.
 
-\section{Pointwise Approach}
-
-\section{Pairwise Approach}
-\section{Listwise Approach}
-\chapter{Evaluation Methods}
-
+\chapter{Evaluating a ranking}
 \section{Normalized Discounted Cumulative Gain}
@@
 $MAP = \frac{\sum\nolimits_{q=1}^{Q}AP(q)}{Q}$
+
+\chapter{Approaches to Learning-to-Rank}
+%TODO: Pointwise, pairwise & listwise beter uitwerken
+\section{Pointwise Approach}
+Learn from each document in isolation.
+Through:
+	\begin{enumerate}
+	\item regression, when relevance is real-valued
+	\item classification, when relevance is categorical
+	\item ordinal regression
+	\end{enumerate}
+\section{Pairwise Approach}
+Pointwise approach Learning-to-Rank methods optimise real valued expected relevance, while evaluation metrics like \ac{nDCG} and \ac{ERR} are only impacted by a change in expected relevance when that change impacts a pairwise preference.
+The pairwise approach solves this drawback of the pointwise approach by regarding ranking as pairwise classification.
+Aggregating a set of predicted pairwise preferences into the corresponding optimal rank is shown to be a NP-Hard problem.
+\section{Listwise Approach}
+Listwise ranking optimises the actual evaluation metric.
+The learner learns to predict an actual ranking itself without using an intermediate step like in pointwise or pairwise Learning-to-Rank.
+The main challenge in this approach is that most evaluation metrics are not differentiable.
+\ac{MAP}, \ac{ERR} and \ac{nDCG} are all discontinuous and non-convex functions, what makes them very hard to optimize.

### 2014-02-21 report/Chapters/Introduction.tex
begin uiteenzetting approaches L2R :)

@@
 The algorithms that we include in the measurements will be determined based on experimental results on the \emph{Yahoo!
-Learning to Rank Challenge}\cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, the LETOR\cite{Qin2010} dataset and the LETOR successor MSLR-WEB30k.
+Learning to Rank Challenge}\cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, the LETOR\cite{Qin2010} dataset and the LETOR successors MSLR-WEB10k and MSLR-WEB30k.
 

### 2014-02-21 report/Chapters/Background.tex
example computations evaluation metrics

@@
 \chapter{Evaluating a ranking}
+Since the 
 \section{Normalized Discounted Cumulative Gain}
@@
 We can write the definition of \ac{nDCG} down mathematically as\\
-$nDCG_p = \frac{DCG_p}{IDCG_p}$
+$nDCG_p = \frac{DCG_p}{IDCG_p}$\\
+Table \ref{tab:example_calculation_nDCG} shows an example calculation for \ac{nDCG}.
+
+\begin{table}
+\begin{tabular}{llllllllllll}
+ & \multicolumn{10}{c}{Rank} &  \\ 
+\cline{2-11}
+ & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & Sum \\ 
+\hline
+rel(i) & 10 & 7 & 6 & 8 & 9 & 5 & 1 & 3 & 2 & 4 &  \\ 
+$\frac{2^{rel_i-1}}{log_2(i+1)}$ & 512 & 40.4 & 16 & 55.1 & 99.0 & 5.7 & 0.3 & 1.3 & 0.6 & 2.3 & 732.7 \\ 
+\hline
+optimal rank & 10 & 9 & 8 & 7 & 6 & 5 & 4 & 3 & 2 & 1 &  \\ 
+$\frac{2^{rel_i-1}}{log_2(i+1)}$ & 512 & 161.5 & 64 & 27.6 & 12.4 & 5.7 & 2.7 & 1.3 & 0.6 & 0.2 & 788.0 \\ 
+\hline
+ &  &  &  &  &  &  &  &  &  &  &  \\ 
+ & \multicolumn{10}{c}{nDCG = $\frac{732.7}{788.0} = 0.93$} &  \\ 
+\end{tabular}
+\caption{Example calculation for \ac{nDCG}}
+\label{tab:example_calculation_nDCG}
+\end{table}
 \section{Expected Reciprocal Rank}
@@
 With $k$ being the positions in the result set between 1 and $n$.
+Table \ref{tab:example_calculation_AP} provides an example calculation of average precision where de documents at positions 1, 5, 6 and 8 in the ranking are relevant.
 \ac{MAP} is the average \ac{AP} for a set of queries.\\
 $MAP = \frac{\sum\nolimits_{q=1}^{Q}AP(q)}{Q}$
+
+\begin{table}
+\begin{tabular}{lllllllllllll}
+ & \multicolumn{10}{c}{Rank} &  & Sum \\ 
+\cline{2-11}
+ & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 &  &  \\ 
+\hline
+$r_i$ & 1 & 0 & 0 & 0 & 1 & 1 & 0 & 1 & 0 & 0 &  &  \\ 
+P@$i$ & 1 &  &  &  & 0.4 & 0.5 &  & 0.5 &  &  &  & 2.4 \\ 
+\hline
+ &  &  &  &  &  &  &  &  &  & /\#Relevant docs & = & 7 \\ 
+ &  &  &  &  &  &  &  &  &  & AP@10 & = & 0.34 \\ 
+\end{tabular}
+\caption{Average Precision example calculation. The number of relevant documents \emph{R} is assumed to be seven.}
+\label{tab:example_calculation_AP}
+\end{table}
 

### 2014-02-21 report/BenchmarkResults.tex
begin uiteenzetting approaches L2R :)
beginnetje method selection + referentie-fix

@@
 
-\chapter{MSLR-WEB30k}
+\chapter{MSLR-WEB10k/MSLR-WEB30k}
 Contrary to the Yahoo!
@@
 \section{Results}
+
+\chapter{Selecting Learning-to-Rank methods}
+The Accuracies of the Learning-to-Rank methods described in the preceding chapters must only be compared within the benchmark and not between benchmarks for the following reasons:
+\begin{enumerate}
+\item Differences in feature sets between data sets detract from fair comparison
+\item Although the \ac{nDCG} definition is unambiguous, Busa-Fekete et al\cite{Busa-Fekete2012} found that \ac{nDCG} evaluation tools of benchmark data sets produced different scores
+\end{enumerate}

### 2014-02-21 report/RelatedWork.tex
beginnetje method selection + referentie-fix

@@
 Ye et al\cite{Ye2009} described how to implement the \ac{GBDT} process in a parallel manner using both MPI and Hadoop.
-\ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise\cite{Zheng2007} or listwise\cite{Wang2011} setting.
+\ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise\cite{Zheng2007} or listwise\cite{Chen2008} setting.
 Experiments showed the Hadoop implementation to result into too expensive communication cost to be useful.

### 2014-02-21 report/Chapters/Background.tex
metricsectie ziet er beter uit :)

@@
 \chapter{Evaluating a ranking}
-Since the 
+Evaluation metrics have long been studied in the field of information retrieval.
+First in the form of evaluation of unranked retrieval sets and later, when the information retrieval field started focussing more on ranked retrieval, in the form of ranked retrieval evaluation.
+In the succeeding section several frequently used evaluation metrics for ranked results will be described.
 \section{Normalized Discounted Cumulative Gain}
@@
 The \ac{DCG}\cite{Burges2005} at a position $p$ is defined as\\
+
 $DCG_p = \sum\nolimits_{i=1}^p \frac{2^{rel_i-1}}{log_2(i+1)}$\\
+
 with $rel_i$ the graded relevance of the result at position $i$.
@@
 We can write the definition of \ac{nDCG} down mathematically as\\
+
 $nDCG_p = \frac{DCG_p}{IDCG_p}$\\
+
 Table \ref{tab:example_calculation_nDCG} shows an example calculation for \ac{nDCG}.
@@
 \ac{ERR} is formally defined as\\
+
 $ERR = \sum\nolimits_{r=1}^n \frac{1}{r} \prod\nolimits_{i=1}^{r-1}(1-R_i)R_i$\\
+
 where the product sequence part of the formula represents the chance that the user will stop at position $r$.\\
@@
 \ac{AP}\cite{Zhu2004} is an often used binary relevance judgement based metric that can be seen as a trade-off between precision and recall that is defined as\\
+
 $AP = \frac{\sum\nolimits_{k=1}^{n}Precision(k)*relevance(k)}{\text{number of relevant docs}}$\\
+
 With $k$ being the positions in the result set between 1 and $n$.
@@
 \ac{MAP} is the average \ac{AP} for a set of queries.\\
-$MAP = \frac{\sum\nolimits_{q=1}^{Q}AP(q)}{Q}$
+
+$MAP = \frac{\sum\nolimits_{q=1}^{Q}AP(q)}{Q}$\\
 

### 2014-02-21 report/BenchmarkResults.tex
gov feature table

@@
 \chapter{LETOR}
-The LETOR dataset was first released by Microsoft Research Asia in April 2007 to solve the absence of a experimental platform for Learning-to-Rank at that time.
-LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0
+The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 to solve the absence of a experimental platform for Learning-to-Rank at that time.
+LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0.
+The original benchmark\cite{Liu2007b} as released in 2007 contained
 \section{Benchmark characteristics}
+\begin{table}[!h]
+\scalebox{0.87}{
+\begin{tabular}{p{0.29cm}|p{7.26cm}||p{0.29cm}|p{9.55cm}}
+ \textbf{ID} & \textbf{Feature Description} & \textbf{ID} & \textbf{Feature Description}\\ 
+ \hline
+1 & \ac{TF} of body					& 36& LMIR.JM of body\\
+2 &	\ac{TF} of anchor				& 37& LMIR.JM of anchor\\
+3 & \ac{TF} of title				& 38& LMIR.JM of title\\
+4 & \ac{TF} of \ac{URL}				& 39& LMIR.JM of \ac{URL}\\
+5 & \ac{TF} of whole document		& 40& LMIR.JM of whole document\\
+
+6 & \ac{IDF} of body 				& 41& Sitemap based term propagation\\
+7 & \ac{IDF} of anchor 				& 42& Sitemap based score propagation\\
+8 & \ac{IDF} of title 				& 43& Hyperlink based score propagation: weighted in-link\\
+9 & \ac{IDF} of \ac{URL} 			& 44& Hyperlink based score propagation: weighted out-link\\
+10& \ac{IDF} of whole document 		& 45& Hyperlink based score propagation: uniform out-link\\
+
+11& \ac{TF-IDF} of body 			& 46& Hyperlink based feature propagation: weighted in-link\\
+12& \ac{TF-IDF} of anchor 			& 47& Hyperlink based feature propagation: weighted out-link\\
+13& \ac{TF-IDF} of title 			& 48& Hyperlink based feature propagation: uniform out-link\\
+14& \ac{TF-IDF} of \ac{URL}			& 49& HITS authority\\
+15& \ac{TF-IDF} of whole document	& 50& HITS hub\\
+
+16& Document length of body			& 51& PageRank\\
+17& Document length of anchor		& 52& HostRank\\
+18& Document length of title		& 53& Topical PageRank\\
+19& Document length of \ac{URL}		& 54& Topical HITS authority\\
+20& Document length of whole document	& 55& Topical HITS hub\\
+
+21& BM25 of body					& 56& In-link number\\
+22& BM25 of anchor					& 57& Out-link number\\
+23& BM25 of title					& 58& Number of slashes in \ac{URL}\\
+24& BM25 of \ac{URL}				& 59& Length of \ac{URL}\\
+25& BM25 of whole document			& 60& Number of child page\\
+
+26& LMIR.ABS\cite{Zhai2001} of body	& 61& BM25 of extracted title\\
+27& LMIR.ABS of anchor				& 62& LMIR.ABS of extracted title\\
+28& LMIR.ABS of title				& 63& LMIR.DIR of extracted title\\
+29& LMIR.ABS of \ac{URL}			& 64& LMIR.JM of extracted title\\
+30& LMIR.ABS of whole document\\
+
+31& LMIR.DIR of body\\
+32& LMIR.DIR of anchor\\
+33& LMIR.DIR of title\\
+34& LMIR.DIR of \ac{URL}\\
+35& LMIR.DIR of whole document\\
+\end{tabular}
+}
+\caption{Features of the LETOR .GOV dataset, obtained from Qin et al\cite{Qin2010}}
+\label{tbl:features_gov}
+\end{table}
 \section{Results}

### 2014-02-24 report/BenchmarkResults.tex
feature tabel met formules af :)

@@
 The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 to solve the absence of a experimental platform for Learning-to-Rank at that time.
-LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0.
-The original benchmark\cite{Liu2007b} as released in 2007 contained
+LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009.
+The original LETOR benchmark collection\cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.
+The OHSUMED collection is a subset of the medical publication database MEDLINE and contains medical publications from 270 journals that were published between 1987 and 1991.
+The .gov collection is a web crawl obtained in January 2002, which was used for the TREC web track in 2003 and 2004.\\
 \section{Benchmark characteristics}
@@
 \end{table}
+
+\begin{table}[!h]
+\scalebox{0.87}{
+\begin{tabular}{p{0.29cm}|p{6.88cm}||p{0.29cm}|p{6.90cm}}
+ \textbf{ID} & \textbf{Feature Description} & \textbf{ID} & \textbf{Feature Description}\\ 
+ \hline
+1 & $\sum\nolimits_{q_i \in q \cap d}c(q_i, d)$ of title					& 24&  $\sum\nolimits_{q_i \in q \cap d}c(q_i,d)\cdot log(\frac{|C|}{df(q_i)})$ of abstract\\
+2 &	$\sum\nolimits_{q_i \in q \cap d}log(c(q_i, d)+1)$ of title				& 25& $\sum\nolimits_{q_i \in q \cap d} log(\frac{c(q_i,d)}{|d|}) \cdot \frac{|C|}{c(q_i,C)}+1$ of abstract\\
+3 & $\sum\nolimits_{q_i \in q \cap d}\frac{c(q_i,d)}{|d|}$ of title				& 26& BM25 of abstract\\
+4 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|}+1)$ of title				& 27& $log($BM25$)$ of abstract\\
+5 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{df(q_i)})$ of title		& 28& LMIR.DIR of abstract\\
+
+6 & $\sum\nolimits_{q_i \in q \cap d}log(log(\frac{|C|}{df(q_i)}))$ of title 				& 29& LMIR.JM of abstract\\
+7 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{c(q_i,C)}+1)$ of title 				& 30& LMIR.ABS of abstract\\
+8 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|} \cdot \frac{|C|}{df(q_i)}+1)$ of title 				& 31& $\sum\nolimits_{q_i \in q \cap d}c(q_i, d)$ of title + abstract\\
+9 & $\sum\nolimits_{q_i \in q \cap d}c(q_i,d)\cdot log(\frac{|C|}{df(q_i)})$ of title 			& 32&	$\sum\nolimits_{q_i \in q \cap d}log(c(q_i, d)+1)$ of title + abstract\\
+10& $\sum\nolimits_{q_i \in q \cap d} log(\frac{c(q_i,d)}{|d|}) \cdot \frac{|C|}{c(q_i,C)}+1$ of title		& 33& $\sum\nolimits_{q_i \in q \cap d}\frac{c(q_i,d)}{|d|}$ of title + abstract\\
+
+11& BM25 of title 			& 34& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|}+1)$ of title + abstract\\
+12& $log($BM25$)$ of title 			& 35& $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{df(q_i)})$ of title + abstract\\
+13& LMIR.DIR of title 			& 36& $\sum\nolimits_{q_i \in q \cap d}log(log(\frac{|C|}{df(q_i)}))$ of title + abstract\\
+14& LMIR.JM of title			& 37& $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{c(q_i,C)}+1)$ of title + abstract\\
+15& LMIR.ABS of title	& 38& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|} \cdot \frac{|C|}{df(q_i)}+1)$ of title + abstract\\
+
+16& $\sum\nolimits_{q_i \in q \cap d}c(q_i, d)$ of abstract		& 39& $\sum\nolimits_{q_i \in q \cap d}c(q_i,d)\cdot log(\frac{|C|}{df(q_i)})$ of title + abstract\\
+17& $\sum\nolimits_{q_i \in q \cap d}log(c(q_i, d)+1)$ of abstract		& 40& $\sum\nolimits_{q_i \in q \cap d} log(\frac{c(q_i,d)}{|d|}) \cdot \frac{|C|}{c(q_i,C)}+1$ of title + abstract\\
+18& $\sum\nolimits_{q_i \in q \cap d}\frac{c(q_i,d)}{|d|}$ of abstract		& 41& BM25 of title + abstract\\
+19& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|}+1)$ of abstract		& 42& $log($BM25$)$ of title + abstract\\
+20& $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{df(q_i)})$ of abstract	& 43& LMIR.DIR of title + abstract\\
+
+21& $\sum\nolimits_{q_i \in q \cap d}log(log(\frac{|C|}{df(q_i)}))$ of abstract					& 44& LMIR.JM of title + abstract\\
+22&  $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{c(q_i,C)}+1)$ of abstract					& 45& LMIR.ABS of title + abstract\\
+23& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|} \cdot \frac{|C|}{df(q_i)}+1)$ of abstract					& & \\
+\end{tabular}
+}
+\caption{Features of the LETOR OHSUMED dataset, obtained from Qin et al\cite{Qin2010}}
+\label{tbl:features_gov}
+\end{table}
 \section{Results}

### 2014-02-24 report/BenchmarkResults.tex
feedback jurrien verwerkt

@@
 LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009.
-The original LETOR benchmark collection\cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.
+The original LETOR benchmark collection\cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.\\
+
 The OHSUMED collection is a subset of the medical publication database MEDLINE and contains medical publications from 270 journals that were published between 1987 and 1991.
-The .gov collection is a web crawl obtained in January 2002, which was used for the TREC web track in 2003 and 2004.\\
+The .gov collection is a web crawl obtained in January 2002, which was used for the TREC web track in 2003 and 2004.
+Tables \ref{tbl:features_gov} and \ref{tbl:features_ohsumed} provide the descriptions of the features of the .gov and the OHSUMED collections respectively.\\
 \section{Benchmark characteristics}
@@
 \caption{Features of the LETOR OHSUMED dataset, obtained from Qin et al\cite{Qin2010}}
-\label{tbl:features_gov}
+\label{tbl:features_ohsumed}
 \end{table}

### 2014-02-24 report/Chapters/Introduction.tex
feedback jurrien verwerkt

@@
 This work focuses on those Learning-to-Rank algorithms that have shown leading performance on relevant benchmark datasets.
-The research questions raised and answered in this work include:
-\begin{itemize}
-\item What is the speed-up of existing Learning-to-Rank algorithms when executed using the MapReduce framework?\\
+This thesis addresses the following research questions:
+\begin{description}
+\item[RQ1] What is the speed-up of existing Learning-to-Rank algorithms when executed using the MapReduce framework?\\
 Where the definition of \emph{relative speed-up} is used for speed-up\cite{Sun1991}:\\
@@
 
-\item Can we adjust those Learning-to-Rank algorithms such that the parallel execution speed-up increases without decreasing accuracy?
-\end{itemize}
+\item[RQ2] Can we adjust those Learning-to-Rank algorithms such that the parallel execution speed-up increases without decreasing accuracy?
+\end{description}
 
@@
 \end{itemize}
+
+The query incorporates different ways of writing of Learning-to-Rank, with and without hyphens, and the synonymous term \emph{machine learned ranking} to increase search recall, i.e.
+to make sure that no relevant studies are missed.
+For the same reason the terms \emph{parallel} and \emph{distributed} are included in the search query.
+Even though \emph{parallel} and \emph{distributed} are not always synonymous is all definitions, we are interested in both approaches in non-sequential data processing.
+\emph{large scale} is a term in the search query that is likely to result in a decrease of precision, because this term is also often used in highly efficient sequential approaches.
+However, the \emph{large scale} term is still included in the query to prevent missing relevant results.\\
 

### 2014-02-25 report/BenchmarkResults.tex
werk aan LETOR

@@
 \chapter{Yandex Internet Mathematics competition}
-\section{Benchmark characteristics}
+
 \section{Results}
@@
 Tables \ref{tbl:features_gov} and \ref{tbl:features_ohsumed} provide the descriptions of the features of the .gov and the OHSUMED collections respectively.\\
-\section{Benchmark characteristics}
+
+Different query sets exists for the .gov corpus.
+Those query sets can be categorized in \emph{topic distillation} (TD), \emph{named page finding} (NP) and \emph{home page finding} (HP).
+With query sets available from both the TREC 2003 and 2004 conferences there are six query sets in total: TD2003, TD2004, NP2003, NP2004, HP2003 and HP2004.\\
+
+The evaluation metrics used are \ac{nDCG} and \ac{MAP}.
+The \emph{winning number} metric is defined as the number of other algorithms that it can beat over all of the seven datasets (six .gov sets + OHSUMED).
+The LETOR organisation implemented and evaluated several well-known Learning-to-Rank algorithms themselves and in addition gathers and publications and results of new algorithms evaluated on the LETOR benchmark.
+The baseline Learning-to-Rank algorithms evaluated by the organisation are:
+\begin{description}
+\item[pointwise]{\leavevmode
+	\begin{itemize}
+	\item Linear regression
+	\end{itemize}}
+\item[pairwise]{\leavevmode
+	\begin{itemize}
+	\item RankingSVM\cite{Herbrich1999,Joachims2002}
+	\item RankBoost\cite{Freund2003}
+	\item FRank\cite{Tsai2007}
+	\end{itemize}}
+\item[listwise]{\leavevmode
+	\begin{itemize}
+	\item ListNet\cite{Cao2007}
+	\item AdaRank-MAP\cite{Xu2007}
+	\item AdaRank-nDCG\cite{Xu2007}
+	\item SVMMAP\cite{Yue2007} 
+	\end{itemize}}
+\end{description} 
 \begin{table}[!h]
@@
 \section{Results}
+\begin{figure}[!h]
+\includegraphics[scale=0.30]{gfx/ndcg_winning_number}
+\caption{Comparison across the seven datasets in LETOR by \ac{nDCG}, obtained from Qin et al\cite{Qin2010}}
+\label{fig:ndcg_winning_number}
+\end{figure}
+
+\begin{figure}[!h]
+\includegraphics[scale=0.30]{gfx/map_winning_number}
+\caption{Comparison across the seven datasets in LETOR by \ac{MAP}, obtained from Qin et al\cite{Qin2010}}
+\label{fig:map_winning_number}
+\end{figure}
 
@@
 The MSLR-WEB30k dataset however contains no proprietary features but only features that are commonly used in the research community.
-\section{Benchmark characteristics}
 \section{Results}

### 2014-02-25 report/Chapters/Introduction.tex
versteviging introduciton

@@
 The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a query.
+Ranking is useful in many other application domains as well, amongst others in drug discovery, to determine the order of maintenance operations\cite{Rudin2009}.
+Learning-to-Rank has also been argued to be a more fitting approach to recommender systems than rating prediction\cite{Adomavicius2005,McNee2006}.\\
+
 Research in the field of ranking models has for a long time been based on manually designed ranking functions and has been an active area of research since Luhn\cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.

### 2014-02-25 report/Chapters/Introduction.tex
misleidende komma vervangen

--- 2014-02-24 report/Chapters/Introduction.tex
@@
 The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a query.
+Ranking is useful in many other application domains as well, amongst others in drug discovery and to determine the order of maintenance operations\cite{Rudin2009}.
+Learning-to-Rank has also been argued to be a more fitting approach to recommender systems than rating prediction\cite{Adomavicius2005,McNee2006}.\\
+
 Research in the field of ranking models has for a long time been based on manually designed ranking functions and has been an active area of research since Luhn\cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.

### 2014-02-25 report/BenchmarkResults.tex
selectiecriterium 99 papers beschreven

@@
 \section{Results}
+The LETOR paper\cite{Qin2010} describes the performance of the LETOR baseline methods.
+Figures \ref{fig:ndcg_winning_number} and \ref{fig:map_winning_number} show ListNet to be the best performing baseline in terms of winning number on both \ac{nDCG} and \ac{MAP}.
+The LETOR website lists 99\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/paper.aspx} papers that use the LETOR benchmark for evaluation of their Learning-to-Rank method.\\
+
 \begin{figure}[!h]
@@
 \end{figure}
+
+In this section I will list those Learning-to-Rank methods that beat the ListNet baseline in at least four of the seven datasets in \ac{nDCG} value.
+Note that this does not necessarily imply that these methods would have scored a higher \ac{nDCG} winning number than ListNet.
+Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR datasets in terms of \ac{nDCG}@10 and \ac{MAP}.
+\begin{table}
+\begin{tabular}{l|ll}
+ & \ac{nDCG}@10 & \ac{MAP} \\ 
+ \hline
+TD2003 & 0.348 & 0.2753 \\ 
+TD2004 & 0.317 & 0.2231 \\ 
+NP2003 & 0.801 & 0.6895 \\ 
+NP2004 & 0.812 & 0.6720 \\ 
+HP2003 & 0.837 & 0.7659 \\ 
+HP2004 & 0.784 & 0.6899 \\ 
+OHSUMED & 0.441 & 0.4457 \\ 
+\end{tabular}
+\caption{Performance of ListNet on LETOR}
+\label{tbl:LETOR_ListNet}
+\end{table}
 

### 2014-02-25 report/BenchmarkResults.tex
structuur LETOR aangepast

@@
 LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009.
-The original LETOR benchmark collection\cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.\\
+The LETOR3.0 benchmark collection\cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.\\
 
+\section{LETOR3.0}
 The OHSUMED collection is a subset of the medical publication database MEDLINE and contains medical publications from 270 journals that were published between 1987 and 1991.
@@
 \end{table}
-\section{Results}
+\subsection{Results}
 The LETOR paper\cite{Qin2010} describes the performance of the LETOR baseline methods.
@@
 
+\section{LETOR4.0}
+\subsection{Results}
+
 \chapter{MSLR-WEB10k/MSLR-WEB30k}

### 2014-02-25 report/BenchmarkResults.tex
tabel tweak

@@
 Figures \ref{fig:ndcg_winning_number} and \ref{fig:map_winning_number} show ListNet to be the best performing baseline in terms of winning number on both \ac{nDCG} and \ac{MAP}.
-The LETOR website lists 99\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/paper.aspx} papers that use the LETOR benchmark for evaluation of their Learning-to-Rank method.\\
+The LETOR website lists\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} a few algorithms that have since been evaluated on the LETOR benchmark collection.\\
 
@@
 
-In this section I will list those Learning-to-Rank methods that beat the ListNet baseline in at least four of the seven datasets in \ac{nDCG} value.
+I will describe the performance of the algorithms that were evaluated on LETOR3.0 after publication of the LETOR paper by Qin et al\cite{Qin2010} as listed on the LETOR website\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} by comparing their performance with the ListNet baseline.
+We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven datasets in \ac{nDCG} value.
 Note that this does not necessarily imply that these methods would have scored a higher \ac{nDCG} winning number than ListNet.
@@
 
+\begin{table}
+\begin{tabular}{l|p{1.2cm}p{1.2cm}p{1.2cm}p{1.4cm}||l}
+ & \rotatebox{55}{Regression + L2 regulatisation} & \rotatebox{55}{RankSVM-Primal} &\rotatebox{55}{RankSVM-Struct} & \rotatebox{55}{SmoothRank} & \rotatebox{55}{ListNet} \\
+ \hline
+TD2003 & 0.3297 & 0.3571 & 0.3467 & 0.3367 & 0.348 \\ 
+TD2004 & 0.2832 & 0.2913 & 0.3090 & 0.3343 & 0.317 \\ 
+NP2003 & 0.8025 & 0.7894 & 0.7955 & 0.7986 & 0.801 \\ 
+NP2004 & 0.8040 & 0.7950 & 0.7977 & 0.8075 & 0.812 \\ 
+HP2003 & 0.8216 & 0.8180 & 0.8162 & 0.8325 & 0.837 \\ 
+HP2004 & 0.7188 & 0.7720 & 0.7666 & 0.8221 & 0.784 \\ 
+OHSUMED & 0.4436 & 0.4504 & 0.4523 & 0.4568 & 0.441 \\ 
+\# winning datasets & 2 & 2 & 1 & 3 & - \\ 
+\end{tabular}
+\caption{Comparison of algorithms recently evaluated on LETOR3.0 with the ListNet baselines}
+\label{tbl:LETOR_recently_added}
+\end{table}
 \section{LETOR4.0}

### 2014-02-26 report/BenchmarkResults.tex
nare realisatie mbt google scholar + inleiding pointwise + kritische noot yahoo winners

@@
 
-The challenge overview paper states as one of the lessons of the challenge that the simple baseline \ac{GBDT} model performed very well with a small performance gap to the complex ensemble submissions at the top of the table.
+The challenge overview paper states as one of the lessons of the challenge that the simple baseline \ac{GBDT} model performed very well with a small performance gap to the complex ensemble submissions at the top of the table.\\
 
@@
 \end{table}
+
+Although the winning \ac{GBDT} models performs very well in terms of \ac{nDCG} their high complexity makes them unsuitable to use in production.
+The winning models take weeks to train and are very slow during query evaluation.
 

### 2014-02-26 report/Chapters/Background.tex
nare realisatie mbt google scholar + inleiding pointwise + kritische noot yahoo winners

@@
 \chapter{Approaches to Learning-to-Rank}
-%TODO: Pointwise, pairwise & listwise beter uitwerken
 \section{Pointwise Approach}
-Learn from each document in isolation.
-Through:
+The pointwise approach can be seen as the most straightforward way of using machine learning for ranking.
+Pointwise Learning-to-Rank methods directly apply machine learning methods to the ranking problem by observing each document in isolation.
+They can be subdivided in the following approaches:
 	\begin{enumerate}
-	\item regression, when relevance is real-valued
-	\item classification, when relevance is categorical
-	\item ordinal regression
+	\item regression-based, which estimate the relevance of a considered document using a regression model.
+	\item classification-based, which classify the relevance category of the document using a classification model.
+	\item ordinal regression-based, which classify the relevance category of the document using a classification model in such a way that the order of relevance categories is taken into account. 
 	\end{enumerate}

### 2014-02-26 report/Chapters/Introduction.tex
nare realisatie mbt google scholar + inleiding pointwise + kritische noot yahoo winners

@@
 \begin{itemize}
-\item ("learning to rank" OR "learning-to-rank" OR "machine learned ranking") AND ("large scale" OR "parallel" OR "distributed")
+\item ("learning to rank" OR "learning-to-rank" OR "machine learned ranking") AND ("parallel" OR "distributed")
 \end{itemize}
@@
 For the same reason the terms \emph{parallel} and \emph{distributed} are included in the search query.
-Even though \emph{parallel} and \emph{distributed} are not always synonymous is all definitions, we are interested in both approaches in non-sequential data processing.
-\emph{large scale} is a term in the search query that is likely to result in a decrease of precision, because this term is also often used in highly efficient sequential approaches.
-However, the \emph{large scale} term is still included in the query to prevent missing relevant results.\\
+Even though \emph{parallel} and \emph{distributed} are not always synonymous is all definitions, we are interested in both approaches in non-sequential data processing.\\
 

### 2014-02-26 report/RelatedWork.tex
nare realisatie mbt google scholar + inleiding pointwise + kritische noot yahoo winners

@@
 \section{Search characteristics}
-The literature research is performed by using the bibliographic databases Scopus and Web of Science with the following search query: \emph{("learning to rank" OR "learning-to-rank" OR "machine learned ranking") AND ("large scale" OR "parallel" OR "distributed")}.
-An abstract based manual filtering step is applied where I filter those results that actually use the \emph{large scale}, \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}.
+The literature research is performed by using the bibliographic databases Scopus and Web of Science with the following search query: \emph{("learning to rank" OR "learning-to-rank" OR "machine learned ranking") AND ("parallel" OR "distributed")}.
+An abstract based manual filtering step is applied where I filter those results that actually use the \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}.
 Studies focusing on efficient query evaluation instead of efficient model training are likely to meet all criteria listed.
@@
 Only four of those 61 documents used \emph{large scale}, \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}, none of them focussed on parallel or distributed learning of ranking functions.
+\subsection{Google Scholar}
+The defined search query resulted in 3300 documents.
 \section{CCRank}

### 2014-02-27 report/BenchmarkResults.tex
bagboo

@@
 The winning models take weeks to train and are very slow during query evaluation.
-
+An exception in training time is the BagBoo\cite{Pavlov2010} method used by Pavlov \& Brunk.
+The bagging component of the BagBoo method enables it to achieve high scalability through parallelism.
+Pavlov et al \cite{Pavlov2010} managed to train half a million trees on 200 nodes in 2 hours.
 \chapter{Yandex Internet Mathematics competition}
@@
 \section{Results}
+The Yandex Internet Mathematics competition was won by
 

### 2014-02-27 report/RelatedWork.tex
bagboo

@@
 \subsection{Google Scholar}
-The defined search query resulted in 3300 documents.
+The defined search query resulted in 3300 documents. We evaluate the first 300 search results.
 \section{CCRank}
@@
 The speed-up achieved at learning the ranking function was unfortunately not stated.
+\section{BagBoo: A hybrid Bagging-the-Boosting Model}
+Pavlov et al\cite{Pavlov2010} at Yandex developed a hybrid model that combines two methods from ensemble learning: bagging and boosting.
+Bagging is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement.
+The bagging method is parallelisable by training each $D_i$ with $i \in \{1..m\}$ on a different node.
+Boosting is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.
+Boosting is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.

### 2014-02-27 report/RelatedWork.tex
BoostingTree

@@
 Boosting is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.
+\section{BoostingTree}
+Kocsis et al\cite{Kocsis2013} propose way to train multiple weak learners in parallel and extend those models that are likely to yield a good model when combined through boosting.
+Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular boosting.
+Using this parallel BoostingTree algorithm \ac{GBDT} models could be trained in parallel.

### 2014-02-27 report/BenchmarkResults.tex
BoostingTree
referenties geuniformiseerd

@@
 The Yahoo!
-Learning to Rank Challenge\cite{Chapelle2011a} is a public Learning-to-Rank competition which took place from March to May 2010, with the goal to promote the datasets and encourage the research community to develop new Learning-to-Rank algorithms.\\
+Learning to Rank Challenge \cite{Chapelle2011a} is a public Learning-to-Rank competition which took place from March to May 2010, with the goal to promote the datasets and encourage the research community to develop new Learning-to-Rank algorithms.\\
 
@@
 \end{tabular}
-\caption{Yahoo! Learning to Rank Challenge dataset characteristics, as described in the challenge overview paper\cite{Chapelle2011a}}
+\caption{Yahoo! Learning to Rank Challenge dataset characteristics, as described in the challenge overview paper \cite{Chapelle2011a}}
 \label{tab:yahoo_characteristics}
@@
 Instead high level descriptions of feature categories are provided.
-The following categories of features are described in the challenge overview paper\cite{Chapelle2011a}:\\
+The following categories of features are described in the challenge overview paper \cite{Chapelle2011a}:\\
 \begin{description}
-\item[Web graph]{Quality and popularity metrics of web documents, e.g. PageRank\cite{Page1999}}.
+\item[Web graph]{Quality and popularity metrics of web documents, e.g. PageRank \cite{Page1999}}.
 \item[Document statistics]{Basic document statistics such as the number of words and url characteristics.}
@@
 \item[Query]{Basic query statistics, such as the number of terms, query frequency, and click-through rate.}
-\item[Text match]{Textual similarity metrics between query and document. Includes \ac{TF-IDF}, BM25\cite{Robertson2009} and other metrics for different sections of the document.}
+\item[Text match]{Textual similarity metrics between query and document. Includes \ac{TF-IDF}, BM25 \cite{Robertson2009} and other metrics for different sections of the document.}
 \item[Topical matching]{These features go beyond similarity at word level and compute similarity on topic level.
@@
 The top five participants all used decision trees and ensemble methods.
-Burges et al\cite{Burges2011} created a linear combination ensemble of eight LambdaMART\cite{Burges2010}, two LambdaRank and two Logistic Regression models.
+Burges et al \cite{Burges2011} created a linear combination ensemble of eight LambdaMART \cite{Burges2010}, two LambdaRank and two Logistic Regression models.
 Gottschalk and Vogel used a combination of RandomForest models and \ac{GBDT} models.
-Pavlov and Brunk used a regression based model using the BagBoo\cite{Pavlov2010} ensemble technique, which combines bagging and boosting.
-Sorokina used a similar combination of bagging and boosting that is called Additive Groves\cite{Sorokina2007}.\\
+Pavlov and Brunk used a regression based model using the BagBoo \cite{Pavlov2010} ensemble technique, which combines bagging and boosting.
+Sorokina used a similar combination of bagging and boosting that is called Additive Groves \cite{Sorokina2007}.\\
 
@@
 \end{tabular}
-\caption{Final standings of the Yahoo! Learning to Rank Challenge, as presented in the challenge overview paper\cite{Chapelle2011a}}
+\caption{Final standings of the Yahoo! Learning to Rank Challenge, as presented in the challenge overview paper \cite{Chapelle2011a}}
 \label{fig:yahoo_results}
@@
 The winning models take weeks to train and are very slow during query evaluation.
-An exception in training time is the BagBoo\cite{Pavlov2010} method used by Pavlov \& Brunk.
+An exception in training time is the BagBoo \cite{Pavlov2010} method used by Pavlov \& Brunk.
 The bagging component of the BagBoo method enables it to achieve high scalability through parallelism.
@@
 \section{Results}
-The Yandex Internet Mathematics competition was won by
+The Yandex Internet Mathematics competition was won by Pavlov et al \cite{Pavlov2010} with the BagBoo method.
 
@@
 LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009.
-The LETOR3.0 benchmark collection\cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.\\
+The LETOR3.0 benchmark collection \cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.\\
 
@@
 	\begin{itemize}
-	\item RankingSVM\cite{Herbrich1999,Joachims2002}
-	\item RankBoost\cite{Freund2003}
-	\item FRank\cite{Tsai2007}
+	\item RankingSVM \cite{Herbrich1999,Joachims2002}
+	\item RankBoost \cite{Freund2003}
+	\item FRank \cite{Tsai2007}
 	\end{itemize}}
@@
 	\begin{itemize}
-	\item ListNet\cite{Cao2007}
-	\item AdaRank-MAP\cite{Xu2007}
-	\item AdaRank-nDCG\cite{Xu2007}
-	\item SVMMAP\cite{Yue2007} 
+	\item ListNet \cite{Cao2007}
+	\item AdaRank-MAP \cite{Xu2007}
+	\item AdaRank-nDCG \cite{Xu2007}
+	\item SVMMAP \cite{Yue2007} 
 	\end{itemize}}
@@
 
-26& LMIR.ABS\cite{Zhai2001} of body	& 61& BM25 of extracted title\\
+26& LMIR.ABS \cite{Zhai2001} of body	& 61& BM25 of extracted title\\
 27& LMIR.ABS of anchor				& 62& LMIR.ABS of extracted title\\
@@
 }
-\caption{Features of the LETOR .GOV dataset, obtained from Qin et al\cite{Qin2010}}
+\caption{Features of the LETOR .GOV dataset, obtained from Qin et al \cite{Qin2010}}
 \label{tbl:features_gov}
@@
 }
-\caption{Features of the LETOR OHSUMED dataset, obtained from Qin et al\cite{Qin2010}}
+\caption{Features of the LETOR OHSUMED dataset, obtained from Qin et al \cite{Qin2010}}
 \label{tbl:features_ohsumed}
@@
 \subsection{Results}
-The LETOR paper\cite{Qin2010} describes the performance of the LETOR baseline methods.
+The LETOR paper \cite{Qin2010} describes the performance of the LETOR baseline methods.
 Figures \ref{fig:ndcg_winning_number} and \ref{fig:map_winning_number} show ListNet to be the best performing baseline in terms of winning number on both \ac{nDCG} and \ac{MAP}.
@@
 \includegraphics[scale=0.30]{gfx/ndcg_winning_number}
-\caption{Comparison across the seven datasets in LETOR by \ac{nDCG}, obtained from Qin et al\cite{Qin2010}}
+\caption{Comparison across the seven datasets in LETOR by \ac{nDCG}, obtained from Qin et al \cite{Qin2010}}
 \label{fig:ndcg_winning_number}
@@
 \includegraphics[scale=0.30]{gfx/map_winning_number}
-\caption{Comparison across the seven datasets in LETOR by \ac{MAP}, obtained from Qin et al\cite{Qin2010}}
+\caption{Comparison across the seven datasets in LETOR by \ac{MAP}, obtained from Qin et al \cite{Qin2010}}
 \label{fig:map_winning_number}
@@
 
-I will describe the performance of the algorithms that were evaluated on LETOR3.0 after publication of the LETOR paper by Qin et al\cite{Qin2010} as listed on the LETOR website\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} by comparing their performance with the ListNet baseline.
+I will describe the performance of the algorithms that were evaluated on LETOR3.0 after publication of the LETOR paper by Qin et al \cite{Qin2010} as listed on the LETOR website\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} by comparing their performance with the ListNet baseline.
 We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven datasets in \ac{nDCG} value.
@@
 \subsection{Results}
+BoostedTree model \cite{Kocsis2013} showed an \ac{nDCG} of 0.5071 on the MQ2007 dataset and thereby beat all baseline models.
 
@@
 \item Differences in feature sets between data sets detract from fair comparison
-\item Although the \ac{nDCG} definition is unambiguous, Busa-Fekete et al\cite{Busa-Fekete2012} found that \ac{nDCG} evaluation tools of benchmark data sets produced different scores
+\item Although the \ac{nDCG} definition is unambiguous, Busa-Fekete et al \cite{Busa-Fekete2012} found that \ac{nDCG} evaluation tools of benchmark data sets produced different scores
 \end{enumerate}

### 2014-02-27 report/Chapters/Background.tex
referenties geuniformiseerd

@@
 \includegraphics[scale=0.26]{gfx/descriminative_training}
-\caption{Machine learning framework for Learning-to-Rank, obtained from Liu\cite{Liu2007}}
+\caption{Machine learning framework for Learning-to-Rank, obtained from Liu \cite{Liu2007}}
 \label{fig:discriminative_training}
@@
 \includegraphics[scale=0.25]{gfx/ltr_framework}
-\caption{A typical Learning-to-Rank setting, obtained from Liu\cite{Liu2007}}
+\caption{A typical Learning-to-Rank setting, obtained from Liu \cite{Liu2007}}
 \label{fig:ltr_framework}
@@
 \subsection{Discounted Cumulative Gain}
-The \ac{DCG}\cite{Burges2005} at a position $p$ is defined as\\
+The \ac{DCG} \cite{Burges2005} at a position $p$ is defined as\\
 
@@
 \section{Expected Reciprocal Rank}
-\ac{ERR}\cite{Chapelle2009} was designed based on the observation that \ac{nDCG} is based on the false assumption that the usefulness of a document at rank $i$ is independent of the usefulness of the documents at rank less than $i$.
+\ac{ERR} \cite{Chapelle2009} was designed based on the observation that \ac{nDCG} is based on the false assumption that the usefulness of a document at rank $i$ is independent of the usefulness of the documents at rank less than $i$.
 \ac{ERR} is based on the reasoning that users are likely to stop exploring the result list once they have found a document that satisfied their information need.
@@
 \section{Mean Average Precision}
-\ac{AP}\cite{Zhu2004} is an often used binary relevance judgement based metric that can be seen as a trade-off between precision and recall that is defined as\\
+\ac{AP} \cite{Zhu2004} is an often used binary relevance judgement based metric that can be seen as a trade-off between precision and recall that is defined as\\
 

### 2014-02-27 report/Chapters/Introduction.tex
referenties geuniformiseerd

@@
 The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a query.
-Ranking is useful in many other application domains as well, amongst others in drug discovery and to determine the order of maintenance operations\cite{Rudin2009}.
-Learning-to-Rank has also been argued to be a more fitting approach to recommender systems than rating prediction\cite{Adomavicius2005,McNee2006}.\\
+Ranking is useful in many other application domains as well, amongst others in drug discovery and to determine the order of maintenance operations \cite{Rudin2009}.
+Learning-to-Rank has also been argued to be a more fitting approach to recommender systems than rating prediction \cite{Adomavicius2005,McNee2006}.\\
 
-Research in the field of ranking models has for a long time been based on manually designed ranking functions and has been an active area of research since Luhn\cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
+Research in the field of ranking models has for a long time been based on manually designed ranking functions and has been an active area of research since Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
 The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models.
@@
 Well-known benchmark datasets include the \emph{Yahoo!
-Learning to Rank Challenge} dataset\cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR dataset\cite{Qin2010} that was build by Microsoft Research.
+Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR dataset \cite{Qin2010} that was build by Microsoft Research.
 One of the concluding observations of the \emph{Yahoo!
-Learning to Rank Challenge} was that almost all work in the Learning-to-Rank field focuses on ranking accuracy, while efficiency and scalability of Learning-to-Rank algorithms is still an underexposed research area that is likely to become more important in the near future as training sets are becoming larger and larger\cite{Chapelle2011b}.
-Liu\cite{Liu2007} confirms the observation that efficiency and scalability of Learning-to-Rank methods has so far been an overlooked research area in his influential book on Learning-to-Rank.\\
+Learning to Rank Challenge} was that almost all work in the Learning-to-Rank field focuses on ranking accuracy, while efficiency and scalability of Learning-to-Rank algorithms is still an underexposed research area that is likely to become more important in the near future as training sets are becoming larger and larger \cite{Chapelle2011b}.
+Liu \cite{Liu2007} confirms the observation that efficiency and scalability of Learning-to-Rank methods has so far been an overlooked research area in his influential book on Learning-to-Rank.\\
 
@@
 
-MapReduce\cite{Dean2004} is a parallel programming framework that is inspired by the \emph{Map} and \emph{Reduce} functions commonly used in functional programming.
+MapReduce \cite{Dean2004} is a parallel programming framework that is inspired by the \emph{Map} and \emph{Reduce} functions commonly used in functional programming.
 Since Google developed the MapReduce parallel programming framework back in 2004 it has since grown to be the industry standard model for parallel programming.
@@
 \item[RQ1] What is the speed-up of existing Learning-to-Rank algorithms when executed using the MapReduce framework?\\
-Where the definition of \emph{relative speed-up} is used for speed-up\cite{Sun1991}:\\
+Where the definition of \emph{relative speed-up} is used for speed-up \cite{Sun1991}:\\
 
@@
 The algorithms that we include in the measurements will be determined based on experimental results on the \emph{Yahoo!
-Learning to Rank Challenge}\cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, the LETOR\cite{Qin2010} dataset and the LETOR successors MSLR-WEB10k and MSLR-WEB30k.
+Learning to Rank Challenge} \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, the LETOR \cite{Qin2010} dataset and the LETOR successors MSLR-WEB10k and MSLR-WEB30k.
 

### 2014-02-27 report/RelatedWork.tex
referenties geuniformiseerd

@@
 \section{CCRank}
-Wang et al\cite{Wang2011a,Wang2011b} propose a parallel evolutionary algorithm based on \ac{CC}, a type of evolutionary algorithm.
+Wang et al \cite{Wang2011a,Wang2011b} propose a parallel evolutionary algorithm based on \ac{CC}, a type of evolutionary algorithm.
 The \ac{CC} algorithm is capable of directly optimizing non-differentiable functions, as \ac{nDCG}, in contrary to many optimization algorithms.
@@
 \section{Parallel ListNet using Spark}
-Shukla et al\cite{Shukla2012} explored the parallelisation of the well-known ListNet Learning-to-Rank method using Spark.
+Shukla et al \cite{Shukla2012} explored the parallelisation of the well-known ListNet Learning-to-Rank method using Spark.
 Spark is a parallel computing model that is designed for cyclic data flows which makes it more suitable for iterative algorithms.
@@
 \section{Gradient Boosted Distributed Decision Trees}
-Ye et al\cite{Ye2009} described how to implement the \ac{GBDT} process in a parallel manner using both MPI and Hadoop.
-\ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise\cite{Zheng2007} or listwise\cite{Chen2008} setting.
+Ye et al \cite{Ye2009} described how to implement the \ac{GBDT} process in a parallel manner using both MPI and Hadoop.
+\ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
 Experiments showed the Hadoop implementation to result into too expensive communication cost to be useful.
@@
 \section{nDCG-Annealing}
-Karimzadeghan el al\cite{Karimzadehgan2011} proposed a method using Simulated Annealing along with the Simplex method for its parameter search.
+Karimzadeghan el al \cite{Karimzadehgan2011} proposed a method using Simulated Annealing along with the Simplex method for its parameter search.
 This method directly optimises the often non-differentiable Learning-to-Rank evaluation metrics like \ac{nDCG} and \ac{MAP}.
@@
 Designing Learning-to-Rank algorithms with low time complexity for training is a different another approach towards large scale Learning-to-Rank.
-Pahikkala et al\cite{Pahikkala2009} describe a pairwise \ac{RLS} type of ranking function, RankRLS, with time complexity $\mathcal{O}(n^3+n^2m)$, with $n$ the number of features and $m$ the number of training documents.
-The RankRLS ranking function showed ranking performance similar to RankSVM on the BioInfer corpus\cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.
+Pahikkala et al \cite{Pahikkala2009} describe a pairwise \ac{RLS} type of ranking function, RankRLS, with time complexity $\mathcal{O}(n^3+n^2m)$, with $n$ the number of features and $m$ the number of training documents.
+The RankRLS ranking function showed ranking performance similar to RankSVM on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.
 \section{Distributed Stochastic Gradient Descent}
-Long et al\cite{Long2012} describe special case of their pairwise cross-domain factor Learning-to-Rank method using distributed optimization of \ac{SGD} based on Hadoop MapReduce.
-Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce based method described by  Zinkevich et al\cite{Zinkevich2010} has been used.
+Long et al \cite{Long2012} describe special case of their pairwise cross-domain factor Learning-to-Rank method using distributed optimization of \ac{SGD} based on Hadoop MapReduce.
+Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce based method described by  Zinkevich et al \cite{Zinkevich2010} has been used.
 Real world data from Yahoo!
@@
 \section{FPGA-based LambdaRank}
-Yan et al\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run the Neural Network based LambdaRank Learning-to-Rank algorithm.
+Yan et al \cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run the Neural Network based LambdaRank Learning-to-Rank algorithm.
 This architecture achieved a 29.3X speed-up compared to the software implementation when evaluated on data from a commercial search engine.
 The exploration of \ac{FPGA} for Learning-to-Rank showed other advantages of the \ac{FPGA} approach next to faster model training.
-In their latest publication\cite{Yan2012} the \ac{FPGA} based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.
+In their latest publication \cite{Yan2012} the \ac{FPGA} based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.
 \section{GPGPU for Learning-to-Rank}
-De Sousa et al\cite{DeSousa2012} proposed a \ac{GPGPU} approach using the \ac{GPU} both learning the ranking function and for query processing, thereby improving both training time and query time.
+De Sousa et al \cite{DeSousa2012} proposed a \ac{GPGPU} approach using the \ac{GPU} both learning the ranking function and for query processing, thereby improving both training time and query time.
 An association rule based Learning-to-Rank approach proposed by \cite{Veloso2008} has been implemented using the \ac{GPU} in such a way that the set of rules van be computed in parallel, in different threads, for each document.
@@
 \section{BagBoo: A hybrid Bagging-the-Boosting Model}
-Pavlov et al\cite{Pavlov2010} at Yandex developed a hybrid model that combines two methods from ensemble learning: bagging and boosting.
+Pavlov et al \cite{Pavlov2010} at Yandex developed a hybrid model that combines two methods from ensemble learning: bagging and boosting.
 Bagging is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement.
@@
 \section{BoostingTree}
-Kocsis et al\cite{Kocsis2013} propose way to train multiple weak learners in parallel and extend those models that are likely to yield a good model when combined through boosting.
+Kocsis et al \cite{Kocsis2013} propose way to train multiple weak learners in parallel and extend those models that are likely to yield a good model when combined through boosting.
 Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular boosting.

### 2014-02-27 report/RelatedWork.tex
wat uitleg + referenties

@@
 Pavlov et al \cite{Pavlov2010} at Yandex developed a hybrid model that combines two methods from ensemble learning: bagging and boosting.
-Bagging is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement.
+Bagging \cite{Breiman1996}, also called bootstrap aggregating, is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement.
 The bagging method is parallelisable by training each $D_i$ with $i \in \{1..m\}$ on a different node.
-Boosting is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.
+Boosting \cite{Friedman2002} is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.
 Boosting is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.
@@
 Using this parallel BoostingTree algorithm \ac{GBDT} models could be trained in parallel.
+\section{Deep Stacking Networks}
+A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.\ac{DSN} is based on the stacked generalization ensemble learning method \cite{Wolpert1992},

### 2014-02-28 report/RelatedWork.tex
ADMM omschreven

@@
 \section{Deep Stacking Networks}
-A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.\ac{DSN} is based on the stacked generalization ensemble learning method \cite{Wolpert1992},
+A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.\ac{DSN} is based on the stacked generalization ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack.
+Each layer in the stack learns has the task of learning the weights of the inputs of that layer.
+The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.
+\section{Alternating Direction Method of Multipliers}
+Duh et al \cite{Duh2011} propose the use of \ac{ADMM} for the Learning-to-Rank task. \ac{ADMM} is a general optimization method that solves problems of the form
+\begin{alignat*}{2}
+\text{minimize }   &  f(x) + g(x) \\
+\text{subject to } &  Ax + Bz = c
+\end{alignat*}
+by updating $x$ and $z$ in an alternating fashion.
+It holds the nice properties that it can be executed in parallel and that it allows for incremental model updates without full retraining.
+Duh et al \cite{Duh2011} show how to use \ac{ADMM} to train a RankSVM model in parallel.
+Experiments showed the \ac{ADMM} implementation to achieve a 13.1x training time speed-up on 72 workers, compared to training on a single worker.

### 2014-02-28 report/RelatedWork.tex
meer related work

@@
 Duh et al \cite{Duh2011} show how to use \ac{ADMM} to train a RankSVM model in parallel.
-Experiments showed the \ac{ADMM} implementation to achieve a 13.1x training time speed-up on 72 workers, compared to training on a single worker.
+Experiments showed the \ac{ADMM} implementation to achieve a 13.1x training time speed-up on 72 workers, compared to training on a single worker.\\
+
+Another \ac{ADMM} Learning-to-Rank based approach was proposed by Boyd et al \cite{Boyd2012}.
+Boyd et al \cite{Boyd2012} implemented an \ac{ADMM} based Learning-to-Rank method in Pregel \cite{Malewicz2010}, a parallel computation framework for graph computations.
+No experimental results on parallelisation speed-up were reported on this Pregel based approach.
+\section{Distributed Hyper-parameter Tuning using MapReduce}
+Ganjisaffar et al \cite{Ganjisaffar2011} observed that long training times are often a result of training many models for hyperparameter optimisation.
+Grid search is the \emph{de facto} standard of hyperparameter optimisation and is simply an exhaustive search through a manually specified subset of hyperparameter combinations.
+The authors show how to perform parallel grid search on MapReduce clusters, which is easy because grid search is an embarrassingly parallel method as hyperparameter combinations are mutually independent.
+They apply their grid search on MapReduce approach in a Learning-to-Rank setting to train a LambdaMART \cite{Wu2008} ranking model, a model that uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
+Experiments showed that the solution scales linearly in number of hyperparameter combinations.
+However, the risk of overfitting grows overwhelmingly as the number of hyperparameter combinations grow, even when validation sets grow large.

### 2014-03-04 report/BenchmarkResults.tex
feedback jurjen

@@
 \chapter{Yahoo! Learning to Rank Challenge}
-Yahoo's observation that all existing benchmark datasets were too small to draw reliable conclusions, especially in comparison with datasets used in commercial search engines, prompted Yahoo to release two internal datasets from Yahoo!
+Yahoo's observation that all existing benchmark datasets were too small to draw reliable conclusions prompted Yahoo to release two internal datasets from Yahoo!
 search.
+Datasets used at commercial search engines are many times larger than available benchmark datasets.
+Yahoo!
+published a subset of their own commercial training data and launched a Learning-to-Rank competition based on this data.
 The Yahoo!
@@
 Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning-to-Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country.
-For this experiment I will only look at the standard Learning-to-Rank dataset because transfer learning is a separate research area that is not included in this thesis.\\
+For this experiment, I will only look at the standard Learning-to-Rank dataset, because transfer learning is a separate research area that is not included in this thesis.\\
 \begin{table}[!h]
@@
 
-Different query sets exists for the .gov corpus.
-Those query sets can be categorized in \emph{topic distillation} (TD), \emph{named page finding} (NP) and \emph{home page finding} (HP).
+Different query sets exists for the .gov corpus. Those query sets are categorized in the following categories \cite{}:
+\begin{description}
+\item[topic distillation (TD)]these queries involve finding relevant pages given a broad query.
+E.g., the query 'cotton industry' which is entered with the aim to find pages that provide information about the cotton industry.
+\item[named page finding (NP)]in these queries the user asks for one specific page by name. E.g., the user queries for 'ADA Enforcement' to find the page $http://www.usdoj.gov/crt/ada/enforce.htm$.
+\item[home page finding (HP)]these queries are similar to NP queries in that the user is looking for one specific page, but now this specific page is a homepage.
+E.g., the user queries for 'Tennessee Valley Authority' to find the homepage $http://www.tva.gov$.
+\end{description}
 With query sets available from both the TREC 2003 and 2004 conferences there are six query sets in total: TD2003, TD2004, NP2003, NP2004, HP2003 and HP2004.\\

### 2014-03-04 report/Chapters/Background.tex
feedback jurjen

@@
 \begin{description}
-\item[Feature Based]{\emph{Feature based} means that all documents under investigation are represented by feature vectors that reflect the relevance of the documents to the query.}
-\item[Discriminative Training]{\emph{Discriminative training} means that the learning process can be well described by the four components of discriminative learning.
+\item[Feature Based]{means that all documents under investigation are represented by feature vectors that reflect the relevance of the documents to the query.}
+\item[Discriminative Training]{means that the learning process can be well described by the four components of discriminative learning.
 That is, a Learning-to-Rank method has its own \emph{input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function}, like the machine learning process described by Figure \ref{fig:discriminative_training}.
@@
 Figure \ref{fig:ltr_framework} shows how the machine learning process as described in Figure \ref{fig:discriminative_training} typically takes place in a ranking scenario.
-A set of queries $q_i$ with $n > i > 1$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgements of those documents $y_i$ are used together to train a model $h$, that can predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is are minimal in terms of a loss function.
+A set of queries $q_i$ with $n > i > 1$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgements of those documents $y_i$ are used together to train a model $h$.
+Model $h$ can after training be used to predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is are minimal in terms of a loss function.
 \begin{figure}[!h]
@@
 First in the form of evaluation of unranked retrieval sets and later, when the information retrieval field started focussing more on ranked retrieval, in the form of ranked retrieval evaluation.
-In the succeeding section several frequently used evaluation metrics for ranked results will be described.
+In the succeeding section several frequently used evaluation metrics for ranked results will be described.\\
+
+No single evaluation metric that we are going to describe is indisputably better or worse than any of the other metrics.
+Different benchmarking settings have used different evaluation metrics.
+Metrics introduced in this section will be used to express the evaluation results in the Benchmark Results part of this thesis.
 \section{Normalized Discounted Cumulative Gain}
@@
 This optimal rank is obtained by sorting documents on relevance for a given query.
-We can write the definition of \ac{nDCG} down mathematically as\\
+The definition of \ac{nDCG} can be written down mathematically as\\
 
@@
 \section{Pairwise Approach}
-Pointwise approach Learning-to-Rank methods optimise real valued expected relevance, while evaluation metrics like \ac{nDCG} and \ac{ERR} are only impacted by a change in expected relevance when that change impacts a pairwise preference.
+Pointwise Learning-to-Rank methods have the drawback that they optimise real valued expected relevance, while evaluation metrics like \ac{nDCG} and \ac{ERR} are only impacted by a change in expected relevance when that change impacts a pairwise preference.
 The pairwise approach solves this drawback of the pointwise approach by regarding ranking as pairwise classification.

### 2014-03-04 report/Chapters/Introduction.tex
feedback jurjen

@@
 Ranking is a core problem in the field of information retrieval.
-The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a query.
-Ranking is useful in many other application domains as well, amongst others in drug discovery and to determine the order of maintenance operations \cite{Rudin2009}.
+The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a given query.
+Ranking is useful in many other application domains outside information retrieval as well, amongst others in drug discovery and to determine the order of maintenance operations \cite{Rudin2009}.
 Learning-to-Rank has also been argued to be a more fitting approach to recommender systems than rating prediction \cite{Adomavicius2005,McNee2006}.\\
 
-Research in the field of ranking models has for a long time been based on manually designed ranking functions and has been an active area of research since Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
+Research in the field of ranking models has for a long time been based on manually designed ranking functions and has been an active area of research.
+Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
 The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models.
-Learning-to-Rank is the relatively new research area covering the use of machine learning models for the ranking task.\\
+Learning-to-Rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\
 
@@
 MapReduce \cite{Dean2004} is a parallel programming framework that is inspired by the \emph{Map} and \emph{Reduce} functions commonly used in functional programming.
-Since Google developed the MapReduce parallel programming framework back in 2004 it has since grown to be the industry standard model for parallel programming.
+Since Google developed the MapReduce parallel programming framework back in 2004 it has grown to be the industry standard model for parallel programming.
 Lin \cite{Lin2013} observed that algorithms that are of iterative nature, which most Learning-to-Rank algorithms are, are not amenable to the MapReduce framework.
@@
 \begin{description}
-\item[RQ1] What is the speed-up of existing Learning-to-Rank algorithms when executed using the MapReduce framework?\\
+\item[RQ1] What are the best performing Learning-to-Rank algorithms on relevant benchmark datasets?\\
+
+\item[RQ2] What is the speed-up of those Learning-to-Rank algorithms when executed using the MapReduce framework?\\
 Where the definition of \emph{relative speed-up} is used for speed-up \cite{Sun1991}:\\
@@
 
-\item[RQ2] Can we adjust those Learning-to-Rank algorithms such that the parallel execution speed-up increases without decreasing accuracy?
+\item[RQ3] Can those Learning-to-Rank algorithms be adjusted in such a way that the parallel execution speed-up increases without decreasing accuracy?
 \end{description}
@@
 
-To answer the first research question I will implement Learning-to-Rank methods in the MapReduce framework and measuring the runtime as a factor of the number of cluster nodes used to complete the computation.\\
+To answer the first research question, I will implement Learning-to-Rank methods in the MapReduce framework and measure runtime as a factor of the number of cluster nodes used to complete the computation.\\
 
-To implement the Learning-to-Rank algorithms I will use cloud based MapReduce implementation from Microsoft was used that is called HDInsight.
+To implement the Learning-to-Rank algorithms, I will use cloud based MapReduce implementation from Microsoft was used that is called HDInsight.
 It which is based on the popular MapReduce open source implementation Hadoop\footnote{http://hadoop.apache.org/}.

### 2014-03-04 report/RelatedWork.tex
feedback jurjen

@@
 \chapter{Related Work}
-\section{Search characteristics}
 The literature research is performed by using the bibliographic databases Scopus and Web of Science with the following search query: \emph{("learning to rank" OR "learning-to-rank" OR "machine learned ranking") AND ("parallel" OR "distributed")}.
@@
 Studies focusing on efficient query evaluation instead of efficient model training are likely to meet all criteria listed.
-As a last step I will filter out studies based on the whole document that only focus on efficient query evaluation and not on parallel or distributed learning of ranking functions.
-\subsection{Scopus}
-The defined search query resulted in 65 documents.
+As a last step I will filter out studies based on the whole document that only focus on efficient query evaluation and not on parallel or distributed learning of ranking functions.\\
+
+on Scopus the defined search query resulted in 65 documents.
 Only 14 of those documents used \emph{large scale}, \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}.
-10 out of those 14 documents focussed on parallel or distributed learning of ranking functions.
-\subsection{Web of Science}
-The defined search query resulted in 16 documents.
+10 out of those 14 documents focussed on parallel or distributed learning of ranking functions.\\
+
+The defined search query resulted in 16 documents on Web of Science.
 Four of those documents were also present in the set of 65 documents found using Scopus, leaving 61 unique documents.
-Only four of those 61 documents used \emph{large scale}, \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}, none of them focussed on parallel or distributed learning of ranking functions.
-\subsection{Google Scholar}
-The defined search query resulted in 3300 documents. We evaluate the first 300 search results.
+Only four of those 61 documents used \emph{large scale}, \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}, none of them focussed on parallel or distributed learning of ranking functions.\\
+
+On Google Scholar the defined search query resulted in 3300 documents. We evaluate the first 300 search results.\\
+
+The following sections describe the literature study results found.
+
 \section{CCRank}
-Wang et al \cite{Wang2011a,Wang2011b} propose a parallel evolutionary algorithm based on \ac{CC}, a type of evolutionary algorithm.
+Wang et al.
+\cite{Wang2011a,Wang2011b} propose a parallel evolutionary algorithm based on \ac{CC}, a type of evolutionary algorithm.
 The \ac{CC} algorithm is capable of directly optimizing non-differentiable functions, as \ac{nDCG}, in contrary to many optimization algorithms.
@@
 \section{Parallel ListNet using Spark}
-Shukla et al \cite{Shukla2012} explored the parallelisation of the well-known ListNet Learning-to-Rank method using Spark.
+Shukla et al.
+\cite{Shukla2012} explored the parallelisation of the well-known ListNet Learning-to-Rank method using Spark.
 Spark is a parallel computing model that is designed for cyclic data flows which makes it more suitable for iterative algorithms.
@@
 \section{Gradient Boosted Distributed Decision Trees}
-Ye et al \cite{Ye2009} described how to implement the \ac{GBDT} process in a parallel manner using both MPI and Hadoop.
+Ye et al.
+\cite{Ye2009} described how to implement the \ac{GBDT} process in a parallel manner using both MPI and Hadoop.
 \ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
@@
 \section{nDCG-Annealing}
-Karimzadeghan el al \cite{Karimzadehgan2011} proposed a method using Simulated Annealing along with the Simplex method for its parameter search.
+Karimzadeghan et al.
+\cite{Karimzadehgan2011} proposed a method using Simulated Annealing along with the Simplex method for its parameter search.
 This method directly optimises the often non-differentiable Learning-to-Rank evaluation metrics like \ac{nDCG} and \ac{MAP}.
@@
 \section{Low complexity Learning-to-Rank}
-Designing Learning-to-Rank algorithms with low time complexity for training is a different another approach towards large scale Learning-to-Rank.
-Pahikkala et al \cite{Pahikkala2009} describe a pairwise \ac{RLS} type of ranking function, RankRLS, with time complexity $\mathcal{O}(n^3+n^2m)$, with $n$ the number of features and $m$ the number of training documents.
+Designing Learning-to-Rank algorithms with low time complexity for training is another approach towards large scale Learning-to-Rank.
+Pahikkala et al.
+\cite{Pahikkala2009} describe a pairwise \ac{RLS} type of ranking function, RankRLS, with time complexity $\mathcal{O}(n^3+n^2m)$, with $n$ the number of features and $m$ the number of training documents.
 The RankRLS ranking function showed ranking performance similar to RankSVM on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.
 \section{Distributed Stochastic Gradient Descent}
-Long et al \cite{Long2012} describe special case of their pairwise cross-domain factor Learning-to-Rank method using distributed optimization of \ac{SGD} based on Hadoop MapReduce.
-Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce based method described by  Zinkevich et al \cite{Zinkevich2010} has been used.
+Long et al.
+\cite{Long2012} describe special case of their pairwise cross-domain factor Learning-to-Rank method using distributed optimization of \ac{SGD} based on Hadoop MapReduce.
+Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce based method described by  Zinkevich et al.
+\cite{Zinkevich2010} has been used.
 Real world data from Yahoo!
@@
 \section{FPGA-based LambdaRank}
-Yan et al \cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run the Neural Network based LambdaRank Learning-to-Rank algorithm.
+Yan et al.
+\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run the Neural Network based LambdaRank Learning-to-Rank algorithm.
 This architecture achieved a 29.3X speed-up compared to the software implementation when evaluated on data from a commercial search engine.
@@
 \section{GPGPU for Learning-to-Rank}
-De Sousa et al \cite{DeSousa2012} proposed a \ac{GPGPU} approach using the \ac{GPU} both learning the ranking function and for query processing, thereby improving both training time and query time.
+De Sousa et al.
+\cite{DeSousa2012} proposed a \ac{GPGPU} approach using the \ac{GPU} both learning the ranking function and for query processing, thereby improving both training time and query time.
 An association rule based Learning-to-Rank approach proposed by \cite{Veloso2008} has been implemented using the \ac{GPU} in such a way that the set of rules van be computed in parallel, in different threads, for each document.
@@
 \section{BagBoo: A hybrid Bagging-the-Boosting Model}
-Pavlov et al \cite{Pavlov2010} at Yandex developed a hybrid model that combines two methods from ensemble learning: bagging and boosting.
+Pavlov et al.
+\cite{Pavlov2010} at Yandex developed a hybrid model that combines two methods from ensemble learning: bagging and boosting.
 Bagging \cite{Breiman1996}, also called bootstrap aggregating, is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement.
@@
 \section{BoostingTree}
-Kocsis et al \cite{Kocsis2013} propose way to train multiple weak learners in parallel and extend those models that are likely to yield a good model when combined through boosting.
+Kocsis et al.
+\cite{Kocsis2013} propose way to train multiple weak learners in parallel and extend those models that are likely to yield a good model when combined through boosting.
 Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular boosting.
@@
 \section{Deep Stacking Networks}
-A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.\ac{DSN} is based on the stacked generalization ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack.
+A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.
+\ac{DSN} is based on the stacked generalization ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack.
 Each layer in the stack learns has the task of learning the weights of the inputs of that layer.
@@
 \section{Alternating Direction Method of Multipliers}
-Duh et al \cite{Duh2011} propose the use of \ac{ADMM} for the Learning-to-Rank task. \ac{ADMM} is a general optimization method that solves problems of the form
+Duh et al. \cite{Duh2011} propose the use of \ac{ADMM} for the Learning-to-Rank task. \ac{ADMM} is a general optimization method that solves problems of the form
 \begin{alignat*}{2}
@@
 It holds the nice properties that it can be executed in parallel and that it allows for incremental model updates without full retraining.
-Duh et al \cite{Duh2011} show how to use \ac{ADMM} to train a RankSVM model in parallel.
+Duh et al.
+\cite{Duh2011} show how to use \ac{ADMM} to train a RankSVM model in parallel.
 Experiments showed the \ac{ADMM} implementation to achieve a 13.1x training time speed-up on 72 workers, compared to training on a single worker.\\
 
-Another \ac{ADMM} Learning-to-Rank based approach was proposed by Boyd et al \cite{Boyd2012}.
-Boyd et al \cite{Boyd2012} implemented an \ac{ADMM} based Learning-to-Rank method in Pregel \cite{Malewicz2010}, a parallel computation framework for graph computations.
+Another \ac{ADMM} Learning-to-Rank based approach was proposed by Boyd et al.
+\cite{Boyd2012}.
+Boyd et al.
+\cite{Boyd2012} implemented an \ac{ADMM} based Learning-to-Rank method in Pregel \cite{Malewicz2010}, a parallel computation framework for graph computations.
 No experimental results on parallelisation speed-up were reported on this Pregel based approach.
 \section{Distributed Hyper-parameter Tuning using MapReduce}
-Ganjisaffar et al \cite{Ganjisaffar2011} observed that long training times are often a result of training many models for hyperparameter optimisation.
+Ganjisaffar et al.
+\cite{Ganjisaffar2011} observed that long training times are often a result of training many models for hyperparameter optimisation.
 Grid search is the \emph{de facto} standard of hyperparameter optimisation and is simply an exhaustive search through a manually specified subset of hyperparameter combinations.
@@
 However, the risk of overfitting grows overwhelmingly as the number of hyperparameter combinations grow, even when validation sets grow large.
+\section{Distributed LambdaRank}
+Svore and Burges \cite{Svore2010,Svore2012} designed two approaches to train the LambdaMART \cite{Wu2008} model in a distributed manner.
+Their first approach is applicable in case the whole training set fits in main memory on every node, in that case the tree split computations of the regressions trees in the LambdaMART tree ensemble are split and the data is not.
+The second approach distributes the training data and corresponding training computations and is therefore scalable to high amounts of training data.
+Both approaches achieves a six times speed-up over LambdaMART on 32 nodes compared to a single node.
+\section{Learning-to-Rank with Bregman Divergences and Monotone Retargeting}
+Acharyya et al.
+\cite{Acharyya} propose a Learning-to-Rank method searches for an order preserving transformation of the target scores that is easier for a regressor to fit.
+This approach is based on the observation that it is not necessary to fit scores exactly, since the evaluation is dependent on the order and not on the pointwise predictions themselves.

### 2014-03-04 report/RelatedWork.tex
versie bespreking djoerd, sander, niek

@@
 Acharyya et al.
-\cite{Acharyya} propose a Learning-to-Rank method searches for an order preserving transformation of the target scores that is easier for a regressor to fit.
-This approach is based on the observation that it is not necessary to fit scores exactly, since the evaluation is dependent on the order and not on the pointwise predictions themselves.
+\cite{Acharyya2012} propose a Learning-to-Rank method searches for an order preserving transformation (\emph{monotone retargeting}) of the target scores that is easier for a regressor to fit.
+This approach is based on the observation that it is not necessary to fit scores exactly, since the evaluation is dependent on the order and not on the pointwise predictions themselves.\\
+
+Bregman divergences are a family of distance-like functions that do not satisfy symmetry nor the triangle inequality.
+A well-known member of the class of Bregman divergences is Kullback-Leibler divergence, also known as \emph{information gain}.\\
+
+Acharyya et al.
+\cite{Acharyya2012} defined a parallel algorithm that optimises a Bregman divergence function as surrogate of \ac{nDCG} that is claimed to be well suited for implementation of a \ac{GPGPU}.
+No experiments on speed-up have been performed.

### 2014-03-05 report/BenchmarkResults.tex
TREC2003 reference
parallel decision trees categorie in related work

@@
 \section{Results}
+1055 teams send in at least one submission to the Yahoo!
+Learning to Rank challenge.
+The top eight participants of the Yahoo!
+Learning to Rank challenge all used decision trees combined with ensemble methods.
+The mainly used ensemble method within these top performers is boosting.
+The combination of boosting with decision tree learners is often called \acf{GBDT}.
 Figure \ref{fig:yahoo_results} shows the top five participants in the Yahoo!
 Learning to Rank Challenge in terms of ERR score.
-The top five participants all used decision trees and ensemble methods.
 Burges et al \cite{Burges2011} created a linear combination ensemble of eight LambdaMART \cite{Burges2010}, two LambdaRank and two Logistic Regression models.
@@
 
-Different query sets exists for the .gov corpus. Those query sets are categorized in the following categories \cite{}:
+Different query sets exists for the .gov corpus. Those query sets are categorized in the following categories \cite{Craswell2003}:
 \begin{description}

### 2014-03-05 report/RelatedWork.tex
parallel decision trees categorie in related work

@@
 The Spark implementation of ListNet showed near a linear training time reduction.
-\section{Gradient Boosted Distributed Decision Trees}
-Ye et al.
-\cite{Ye2009} described how to implement the \ac{GBDT} process in a parallel manner using both MPI and Hadoop.
-\ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
-Experiments showed the Hadoop implementation to result into too expensive communication cost to be useful.
-Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
-The MPI approach proved to be successful and obtained near linear speed-ups.
 \section{nDCG-Annealing}
@@
 Boosting is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.
-\section{BoostingTree}
+\section{Parallel Decision Trees}
+Ye et al.
+\cite{Ye2009} described how to implement the \ac{GBDT} process in a parallel manner using both MPI and Hadoop.
+\ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
+Experiments showed the Hadoop implementation to result into too expensive communication cost to be useful.
+Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
+The MPI approach proved to be successful and obtained near linear speed-ups.\\
+
 Kocsis et al.
@@
 Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular boosting.
-Using this parallel BoostingTree algorithm \ac{GBDT} models could be trained in parallel.
+Using this parallel BoostingTree algorithm \ac{GBDT} models could be trained in parallel.\\
+
+Tyree et al.
+\cite{Tyree2011} describe a way of parallelising \ac{GBDT} models for Learning-to-Rank where the construction of the regression trees themselves are parallelised, based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}.
+Decision trees are build layer-by-layer where the calculations needed for building each new layer in the tree are divided over the nodes using a master-worker paradigm.
 \section{Deep Stacking Networks}

### 2014-03-05 report/Chapters/Background.tex
uitbreiding listwise ltr uitleg

@@
 The main challenge in this approach is that most evaluation metrics are not differentiable.
-\ac{MAP}, \ac{ERR} and \ac{nDCG} are all discontinuous and non-convex functions, what makes them very hard to optimize.
+\ac{MAP}, \ac{ERR} and \ac{nDCG} are non-differentiable, non-convex and discontinuous functions, what makes them very hard to optimize.\\
+
+Although the properties of \ac{MAP}, \ac{ERR} and \ac{nDCG} are not ideal for direct optimisation, some listwise approaches do focus on direct metric optimisation \cite{Yue2007, Taylor2008, Chapelle2010}.
+Most listwise approaches work around optimisation of the non-differentiable, non-convex and discontinuous metrics by optimising surrogate cost functions that mimic the behaviour of \ac{MAP}, \ac{ERR} or \ac{nDCG}, but have nicer properties for optimisation.

### 2014-03-05 report/RelatedWork.tex
uitbreiding listwise ltr uitleg

@@
 Decision trees are build layer-by-layer where the calculations needed for building each new layer in the tree are divided over the nodes using a master-worker paradigm.
+The data is partitioned of the workers, who compress their share into histograms and send these to the master.
+The master uses those histograms to approximate the split and build the next layer.
+The master than communicates this new layer to the workers who can use this new layer to compute new histograms.
+this process is repeated until the tree depth limit is reached.
+The tree construction algorithm parallelised with described master-worker method is \ac{CART} \cite{Breiman1984}.
+Speed-up experiments on the LETOR and the Yahoo!
+Learning to Rank challenge data sets were performed.
+This parallel \ac{CART}-tree building approach showed speed-up of up to 42x on shared memory machines and up to 25x on distributed memory machines.
 \section{Deep Stacking Networks}

### 2014-03-05 report/Chapters/Background.tex
uitleg NP-Hard probleem pairwise ranking

@@
 Pointwise Learning-to-Rank methods have the drawback that they optimise real valued expected relevance, while evaluation metrics like \ac{nDCG} and \ac{ERR} are only impacted by a change in expected relevance when that change impacts a pairwise preference.
-The pairwise approach solves this drawback of the pointwise approach by regarding ranking as pairwise classification.
-Aggregating a set of predicted pairwise preferences into the corresponding optimal rank is shown to be a NP-Hard problem.
+The pairwise approach solves this drawback of the pointwise approach by regarding ranking as pairwise classification.\\
+
+Aggregating a set of predicted pairwise preferences into the corresponding optimal rank is shown to be a NP-Hard problem \cite{Feldman2012}.
+An often used solution to this problem is to upper bound the number of classification mistakes by an easy to optimise function \cite{Bartlett2006}.
 \section{Listwise Approach}

### 2014-03-05 report/RelatedWork.tex
uitleg NP-Hard probleem pairwise ranking

@@
 Tyree et al.
-\cite{Tyree2011} describe a way of parallelising \ac{GBDT} models for Learning-to-Rank where the construction of the regression trees themselves are parallelised, based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}.
+\cite{Tyree2011} describe a way of parallelising \ac{GBDT} models for Learning-to-Rank where the boosting step is still executed sequentially, instead the construction of the regression trees themselves are parallelised.
+The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}.
 Decision trees are build layer-by-layer where the calculations needed for building each new layer in the tree are divided over the nodes using a master-worker paradigm.

### 2014-03-05 report/RelatedWork.tex
robust statistics referentie
RoBiRank

@@
 No experiments on speed-up have been performed.
+\section{Parallel robust Learning-to-Rank}
+Robust learning \cite{Huber1981} is defined as the task to lean a model in the presence of outliers.
+Yun et al.
+describe a \cite{Yun2014} robust Learning-to-Rank model called RoBiRank that has the additional advantage that it is executable in parallel.
+Their RoBiRank model uses parameter expansion to linearise a surrogate loss function, after which the elements of the linear combination can be divided over the available nodes.
+The Parameter expansion trick was proposed by Gopal and Yang \cite{Gopal2013}, who originally proposed it for multinomial logistic regression.
+Unfortunately, no speed-up experiments were mentioned for the RoBiRank method, since Yun et al.
+focussed their research more on robust ranking than on parallel ranking.
+It is stated however that the training time of RoBiRank on a computing cluster is comparable to the parallel implementation by Lee and Lin \cite{Lee2014} of a linear kernel RankSVM \cite{Herbrich1999, Joachims2002} that is also described in this related work section.

### 2014-03-05 report/RelatedWork.tex
lee and lin efficient rankSVM aangepast

@@
 focussed their research more on robust ranking than on parallel ranking.
-It is stated however that the training time of RoBiRank on a computing cluster is comparable to the parallel implementation by Lee and Lin \cite{Lee2014} of a linear kernel RankSVM \cite{Herbrich1999, Joachims2002} that is also described in this related work section.
+The only reference to performance of RoBiRank in terms of speed is the statement that its training time on a computing cluster is comparable to the more efficient implementation by Lee and Lin \cite{Lee2014} of a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.

### 2014-03-05 report/Chapters/Introduction.tex
introduction sterker gemaakt

@@
 The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a given query.
-Ranking is useful in many other application domains outside information retrieval as well, amongst others in drug discovery and to determine the order of maintenance operations \cite{Rudin2009}.
-Learning-to-Rank has also been argued to be a more fitting approach to recommender systems than rating prediction \cite{Adomavicius2005,McNee2006}.\\
+Ranking has become a vital part of web search, where commercial search engines help users find their need in the extremely large collection of the World Wide Web.
+One can find useful applications of ranking in many application domains outside web search as well.
+For example, it plays a vital role in amongst others: document summarisation, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}.
+In addition, the ranking task has been argued to be more fitting to recommender systems than rating prediction on a continuous scale \cite{Adomavicius2005,McNee2006}.\\
 

### 2014-03-05 report/Chapters/Introduction.tex
verduidelijking

@@
 One can find useful applications of ranking in many application domains outside web search as well.
-For example, it plays a vital role in amongst others: document summarisation, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}.
+For example, it plays a vital role in amongst others: automatic document summarisation, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}.
 In addition, the ranking task has been argued to be more fitting to recommender systems than rating prediction on a continuous scale \cite{Adomavicius2005,McNee2006}.\\

### 2014-03-07 report/Chapters/Background.tex
twee versies ndcg

@@
 \begin{description}
-\item[Feature Based]{means that all documents under investigation are represented by feature vectors that reflect the relevance of the documents to the query.}
+\item[Feature Based]{means that all documents under investigation are represented by feature vectors.
+Those feature vectors reflect the relevance of the documents to the query, or the importance of the document in itself.}
 \item[Discriminative Training]{means that the learning process can be well described by the four components of discriminative learning.
@@
 \subsection{Discounted Cumulative Gain}
-The \ac{DCG} \cite{Burges2005} at a position $p$ is defined as\\
+There are two definitions of \ac{DCG} used in practice. \ac{DCG} at position $p$ was originally defined by J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} as\\
+
+$DCG_p = \sum\nolimits_{i=1}^p \frac{rel_i-1}{log_2(i+1)}$\\
+
+with $rel_i$ the graded relevance of the result at position $i$.
+The idea is that highly relevant documents that appear lower in a search result should be penalized (discounted).
+This discounting is done by reducing the graded relevance  logarithmically proportional to the position of the result.\\
+
+Burges et al. \cite{Burges2005} proposed an alternative definition of \ac{DCG} that puts stronger emphasis on document relevance\\
 
@@
 
-with $rel_i$ the graded relevance of the result at position $i$.
-The idea is that highly relevant documents that appear lower in a search result should be penalized (discounted).
-This discounting is done by reducing the graded relevance  logarithmically proportional to the position of the result.
 \subsection{Normalized Discounted Cumulative Gain}
@@
 
-Table \ref{tab:example_calculation_nDCG} shows an example calculation for \ac{nDCG}.
+Table \ref{tab:example_calculation_nDCG} shows an example calculation for \ac{nDCG} for both the J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} and Burges et al.
+\cite{Burges2005} version of \ac{DCG}.
 
-\begin{table}
+\begin{table}[!h]
 \begin{tabular}{llllllllllll}
@@
 \hline
+\hline
 rel(i) & 10 & 7 & 6 & 8 & 9 & 5 & 1 & 3 & 2 & 4 &  \\
+\hline
 $\frac{2^{rel_i-1}}{log_2(i+1)}$ & 512 & 40.4 & 16 & 55.1 & 99.0 & 5.7 & 0.3 & 1.3 & 0.6 & 2.3 & 732.7 \\
 \hline
+$\frac{rel_i}{log_2(i+1)}$ & 10 & 4.42 & 3 & 3.45 & 3.48 & 1.78 & 0.33 & 0.95 & 0.6 & 1.16 & 29.17 \\  
+\hline
+\hline
 optimal rank & 10 & 9 & 8 & 7 & 6 & 5 & 4 & 3 & 2 & 1 &  \\
+\hline 
 $\frac{2^{rel_i-1}}{log_2(i+1)}$ & 512 & 161.5 & 64 & 27.6 & 12.4 & 5.7 & 2.7 & 1.3 & 0.6 & 0.2 & 788.0 \\
 \hline
+$\frac{rel_i}{log_2(i+1)}$ & 10 & 5.68 & 4 & 3.01 & 2.32 & 1.78 & 1.33 & 0.95 & 0.6 & 0.29 & 29.96 \\   
+\hline
  &  &  &  &  &  &  &  &  &  &  &  \\
- & \multicolumn{10}{c}{nDCG = $\frac{732.7}{788.0} = 0.93$} &  \\ 
+ & \multicolumn{10}{c}{J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} \ac{nDCG} = $\frac{29.17}{29.96} = 0.97$} &  \\  
+ & \multicolumn{10}{c}{Burges \cite{Burges2005} et al. \ac{nDCG} = $\frac{732.7}{788.0} = 0.93$} &  \\ 
 \end{tabular}

### 2014-03-07 report/Chapters/Introduction.tex
twee versies ndcg

@@
 \begin{description}
-\item[RQ1] What are the best performing Learning-to-Rank algorithms on relevant benchmark datasets?\\
+\item[RQ1] What are the best performing Learning-to-Rank algorithms in terms of accuracy on relevant benchmark datasets?\\
 
@@
 
-To implement the Learning-to-Rank algorithms, I will use cloud based MapReduce implementation from Microsoft was used that is called HDInsight.
-It which is based on the popular MapReduce open source implementation Hadoop\footnote{http://hadoop.apache.org/}.
+I will use the HDInsight cloud-based MapReduce implementation from Microsoft to implement the Learning-to-Rank algorithms.
+HDInsight is based on the popular open source MapReduce implementation Hadoop\footnote{http://hadoop.apache.org/}.
 The algorithms that we include in the measurements will be determined based on experimental results on the \emph{Yahoo!
@@
 \item[Part III: Related Work]{concisely describes existing work in the field of parallel machine learning and parallel Learning-to-Rank.}
-\item[Part IV: Benchmark Results]{sketches the performance of existing Learning-to-Rank methods on several benchmark datasets and describes the selection of Learning-to-Rank methods for the parallelisation experiments.}
+\item[Part IV: Benchmark Results]{sketches the accuracy of existing Learning-to-Rank methods on several benchmark datasets and describes the selection of Learning-to-Rank methods for the parallelisation experiments.}
 \item[Part V: Selected Learning-to-Rank Methods]{describes the algorithms and details of the selected Learning-to-Rank methods.}

### 2014-03-07 report/Chapters/Background.tex
laatste puntjes feedback djoerd

@@
 A set of queries $q_i$ with $n > i > 1$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgements of those documents $y_i$ are used together to train a model $h$.
-Model $h$ can after training be used to predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is are minimal in terms of a loss function.
+Model $h$ can after training be used to predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is minimal in terms of a loss function.
 \begin{figure}[!h]

### 2014-03-07 report/BenchmarkResults.tex
laatste puntjes feedback djoerd
structuur zonder parts

@@
-\chapter{Yahoo! Learning to Rank Challenge}
+\label{chap:benchmark_results}
+\section{Yahoo! Learning to Rank Challenge}
 Yahoo's observation that all existing benchmark datasets were too small to draw reliable conclusions prompted Yahoo to release two internal datasets from Yahoo!
@@
 Documents & 473,134 & 71,083 & 165,660 \\ 
-Features & 519 & - & - \\ 
+Features & 519 & 519 & 519 \\ 
 \end{tabular}
@@
 
-\section{Results}
+\subsection{Results}
 1055 teams send in at least one submission to the Yahoo!
@@
 Pavlov et al \cite{Pavlov2010} managed to train half a million trees on 200 nodes in 2 hours.
-\chapter{Yandex Internet Mathematics competition}
+\section{Yandex Internet Mathematics competition}
 
-\section{Results}
+\subsection{Results}
 The Yandex Internet Mathematics competition was won by Pavlov et al \cite{Pavlov2010} with the BagBoo method.
 
-\chapter{LETOR}
+\section{LETOR}
 The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 to solve the absence of a experimental platform for Learning-to-Rank at that time.
@@
 
-\section{LETOR3.0}
+\subsection{LETOR3.0}
 The OHSUMED collection is a subset of the medical publication database MEDLINE and contains medical publications from 270 journals that were published between 1987 and 1991.
@@
 \end{table}
-\section{LETOR4.0}
-\subsection{Results}
+\subsection{LETOR4.0}
+\subsubsection{Results}
 BoostedTree model \cite{Kocsis2013} showed an \ac{nDCG} of 0.5071 on the MQ2007 dataset and thereby beat all baseline models.
 
-\chapter{MSLR-WEB10k/MSLR-WEB30k}
+\section{MSLR-WEB10k/MSLR-WEB30k}
 Contrary to the Yahoo!
@@
 The MSLR-WEB30k dataset however contains no proprietary features but only features that are commonly used in the research community.
-\section{Results}
+\subsection{Results}
 
-\chapter{Selecting Learning-to-Rank methods}
-The Accuracies of the Learning-to-Rank methods described in the preceding chapters must only be compared within the benchmark and not between benchmarks for the following reasons:
+\section{Selecting Learning-to-Rank methods}
+The accuracies of the Learning-to-Rank methods described in the preceding sections must only be compared within the benchmark and not between benchmarks for the following reasons:
 \begin{enumerate}

### 2014-03-07 report/Chapters/Background.tex
structuur zonder parts

@@
 %************************************************
-\chapter{A basic introduction to Learning-to-Rank}
+\section{A basic introduction to Learning-to-Rank}
 Different definitions of Learning-to-Rank exist.
@@
 These three approaches are in literature respectively called the pointwise approach, the pairwise approach and the listwise approach.
-These three approaches to Learning-to-Rank will be described in more detail further on in this part.
+These three approaches to Learning-to-Rank will be described in more detail in section \ref{sec:ltr_approaches}.
 
-\chapter{Evaluating a ranking}
+\section{Evaluating a ranking}
 Evaluation metrics have long been studied in the field of information retrieval.
 First in the form of evaluation of unranked retrieval sets and later, when the information retrieval field started focussing more on ranked retrieval, in the form of ranked retrieval evaluation.
-In the succeeding section several frequently used evaluation metrics for ranked results will be described.\\
+In this section several frequently used evaluation metrics for ranked results will be described.\\
 
@@
 Different benchmarking settings have used different evaluation metrics.
-Metrics introduced in this section will be used to express the evaluation results in the Benchmark Results part of this thesis.
-\section{Normalized Discounted Cumulative Gain}
+Metrics introduced in this section will be used to express the evaluation results in chapter \ref{chap:benchmark_results} of this thesis.
+\subsection{Normalized Discounted Cumulative Gain}
 Cumulative gain, or its predecessor discounted cumulative gain and normalized discounted cumulative gain, is one of the most widely used measures for effectiveness of ranking methods.
-\subsection{Discounted Cumulative Gain}
+\subsubsection{Discounted Cumulative Gain}
 There are two definitions of \ac{DCG} used in practice. \ac{DCG} at position $p$ was originally defined by J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} as\\
@@
 
-\subsection{Normalized Discounted Cumulative Gain}
+\subsubsection{Normalized Discounted Cumulative Gain}
 \ac{nDCG} normalizes the \ac{DCG} metric to a value in the [0,1] interval by dividing by the \ac{DCG} value of the optimal rank.
@@
 \end{table}
-\section{Expected Reciprocal Rank}
+\subsection{Expected Reciprocal Rank}
 \ac{ERR} \cite{Chapelle2009} was designed based on the observation that \ac{nDCG} is based on the false assumption that the usefulness of a document at rank $i$ is independent of the usefulness of the documents at rank less than $i$.
@@
 \end{lstlisting}
-\section{Mean Average Precision}
+\subsection{Mean Average Precision}
 \ac{AP} \cite{Zhu2004} is an often used binary relevance judgement based metric that can be seen as a trade-off between precision and recall that is defined as\\
@@
 
-\chapter{Approaches to Learning-to-Rank}
-\section{Pointwise Approach}
+\section{Approaches to Learning-to-Rank}
+\label{sec:ltr_approaches}
+\subsection{Pointwise Approach}
 The pointwise approach can be seen as the most straightforward way of using machine learning for ranking.
@@
 	\end{enumerate}
-\section{Pairwise Approach}
+\subsection{Pairwise Approach}
 Pointwise Learning-to-Rank methods have the drawback that they optimise real valued expected relevance, while evaluation metrics like \ac{nDCG} and \ac{ERR} are only impacted by a change in expected relevance when that change impacts a pairwise preference.
@@
 An often used solution to this problem is to upper bound the number of classification mistakes by an easy to optimise function \cite{Bartlett2006}.
-\section{Listwise Approach}
+\subsection{Listwise Approach}
 Listwise ranking optimises the actual evaluation metric.

### 2014-03-07 report/Chapters/Introduction.tex
verduidelijking Thesis Overview
structuur zonder parts

@@
 %\markboth{\spacedlowsmallcaps{Appendix}}{\spacedlowsmallcaps{Appendix}}
-\chapter{Motivation and Problem Statement}
+\section{Motivation and Problem Statement}
 Ranking is a core problem in the field of information retrieval.
@@
 This thesis aims to be an exploratory start in this little researched area of parallel Learning-to-Rank.
-A more extensive overview of my research goals and questions are described in chapter \ref{chap:goals}.\\
+A more extensive overview of my research goals and questions are described in section \ref{sec:goals}.\\
 
-\chapter{Research Goals}
-\label{chap:goals}
+\section{Research Goals}
+\label{sec:goals}
 The objective of this thesis is to explore the speed-up in execution time of Learning-to-Rank algorithms through parallelisation using the MapReduce framework.
@@
 
-\chapter{Approach}
+\section{Approach}
 A literature study will be performed to get insight in relevant existing techniques for large scale Learning-to-Rank. The literature study will be performed by using the following query:
@@
 
-\chapter{Thesis Overview}
+\section{Thesis Overview}
 
 \begin{description}
-\item[Part II: Background]{introduces the reader to the basic principles and recent work in the fields of Learning-to-Rank.}
-\item[Part III: Related Work]{concisely describes existing work in the field of parallel machine learning and parallel Learning-to-Rank.}
-\item[Part IV: Benchmark Results]{sketches the accuracy of existing Learning-to-Rank methods on several benchmark datasets and describes the selection of Learning-to-Rank methods for the parallelisation experiments.}
-\item[Part V: Selected Learning-to-Rank Methods]{describes the algorithms and details of the selected Learning-to-Rank methods.}
-\item[Part VI: Implementation]{describes implementation details of the Learning-to-Rank algorithms in the Hadoop framework.}
-\item[Part VII: Results \& Discussion]{presents and discusses speed-up results for the implemented Learning-to-Rank methods.}
-\item[Part VIII: Conclusion]{summarizes the results and answers our research questions based on the results.
+\item[Chapter II: Background]{introduces the reader to the basic principles and recent work in the fields of Learning-to-Rank.}
+\item[Chapter III: Related Work]{concisely describes existing work in the field of parallel machine learning and parallel Learning-to-Rank.}
+\item[Chapter IV: Benchmark Results]{sketches the accuracy of existing Learning-to-Rank methods on several benchmark datasets and describes the selection of Learning-to-Rank methods for the parallelisation experiments.}
+\item[Chapter V: Selected Learning-to-Rank Methods]{describes the algorithms and details of the Learning-to-Rank methods selected in Chapter IV.}
+\item[Chapter VI: Implementation]{describes implementation details of the Learning-to-Rank algorithms in the Hadoop framework.}
+\item[Chapter VII: Results \& Discussion]{presents and discusses speed-up results for the implemented Learning-to-Rank methods.}
+\item[Chapter VIII: Conclusion]{summarizes the results and answers our research questions based on the results.
 The limitations of our research as well as future research directions in the field are mentioned here.}

### 2014-03-07 report/RelatedWork.tex
structuur zonder parts

@@
-\chapter{Related Work}
+\section{Literature study characteristics}
 The literature research is performed by using the bibliographic databases Scopus and Web of Science with the following search query: \emph{("learning to rank" OR "learning-to-rank" OR "machine learned ranking") AND ("parallel" OR "distributed")}.
@@
 
-The following sections describe the literature study results found.
-
-\section{CCRank}
+Section \ref{sec:lit_descr} describe the approaches on parallel or distributed Learning-to-Rank found in literature.
+\section{Literature description}
+\label{sec:lit_descr}
+\subsection{CCRank}
 Wang et al.
@@
 Two reasons have been identified for not achieving linear scale-up with CCRank: 1) parallel execution is suspended after each generation to perform combination in order to produce the candidate solution, 2) Combination has to wait until all parallel tasks have finished, which may spend different running time.
-\section{Parallel ListNet using Spark}
+\subsection{Parallel ListNet using Spark}
 Shukla et al.
@@
 The Spark implementation of ListNet showed near a linear training time reduction.
-\section{nDCG-Annealing}
+\subsection{nDCG-Annealing}
 Karimzadeghan et al.
@@
 Unfortunately their work does not directly report on the speed-up obtained by parallelising  with Hadoop, but it is mentioned that further work needs to be done to effectively leverage parallel execution.
-\section{Low complexity Learning-to-Rank}
+\subsection{Low complexity Learning-to-Rank}
 Designing Learning-to-Rank algorithms with low time complexity for training is another approach towards large scale Learning-to-Rank.
@@
 The RankRLS ranking function showed ranking performance similar to RankSVM on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.
-\section{Distributed Stochastic Gradient Descent}
+\subsection{Distributed Stochastic Gradient Descent}
 Long et al.
@@
 Unfortunately the speed-up obtained by training their model in parallel is not reported.
-\section{FPGA-based LambdaRank}
+\subsection{FPGA-based LambdaRank}
 Yan et al.
@@
 In their latest publication \cite{Yan2012} the \ac{FPGA} based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.
-\section{GPGPU for Learning-to-Rank}
+\subsection{GPGPU for Learning-to-Rank}
 De Sousa et al.
@@
 The speed-up achieved at learning the ranking function was unfortunately not stated.
-\section{BagBoo: A hybrid Bagging-the-Boosting Model}
+\subsection{BagBoo: A hybrid Bagging-the-Boosting Model}
 Pavlov et al.
@@
 Boosting is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.
-\section{Parallel Decision Trees}
+\subsection{Parallel Decision Trees}
 Ye et al.
@@
 This parallel \ac{CART}-tree building approach showed speed-up of up to 42x on shared memory machines and up to 25x on distributed memory machines.
-\section{Deep Stacking Networks}
+\subsection{Deep Stacking Networks}
 A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.
@@
 The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.
-\section{Alternating Direction Method of Multipliers}
+\subsection{Alternating Direction Method of Multipliers}
 Duh et al. \cite{Duh2011} propose the use of \ac{ADMM} for the Learning-to-Rank task. \ac{ADMM} is a general optimization method that solves problems of the form
@@
 No experimental results on parallelisation speed-up were reported on this Pregel based approach.
-\section{Distributed Hyper-parameter Tuning using MapReduce}
+\subsection{Distributed Hyper-parameter Tuning using MapReduce}
 Ganjisaffar et al.
@@
 However, the risk of overfitting grows overwhelmingly as the number of hyperparameter combinations grow, even when validation sets grow large.
-\section{Distributed LambdaRank}
+\subsection{Distributed LambdaRank}
 Svore and Burges \cite{Svore2010,Svore2012} designed two approaches to train the LambdaMART \cite{Wu2008} model in a distributed manner.
@@
 Both approaches achieves a six times speed-up over LambdaMART on 32 nodes compared to a single node.
-\section{Learning-to-Rank with Bregman Divergences and Monotone Retargeting}
+\subsection{Learning-to-Rank with Bregman Divergences and Monotone Retargeting}
 Acharyya et al.
@@
 No experiments on speed-up have been performed.
-\section{Parallel robust Learning-to-Rank}
+\subsection{Parallel robust Learning-to-Rank}
 Robust learning \cite{Huber1981} is defined as the task to lean a model in the presence of outliers.

### 2014-03-07 report/Chapters/AppendixA.tex
cranfield referentie + grote tabellen naar appendix

--- 2014-03-04 report/BenchmarkResults.tex
@@
-\chapter{Yahoo! Learning to Rank Challenge}
-Yahoo's observation that all existing benchmark datasets were too small to draw reliable conclusions prompted Yahoo to release two internal datasets from Yahoo!
-search.
-Datasets used at commercial search engines are many times larger than available benchmark datasets.
-Yahoo!
-published a subset of their own commercial training data and launched a Learning-to-Rank competition based on this data.
-The Yahoo!
-Learning to Rank Challenge \cite{Chapelle2011a} is a public Learning-to-Rank competition which took place from March to May 2010, with the goal to promote the datasets and encourage the research community to develop new Learning-to-Rank algorithms.\\
+\chapter{Appendix A: LETOR Feature set}
+\label{chap:appendix_a}
+\begin{table}[!htp]
+\scalebox{0.87}{
+\begin{tabular}{p{0.29cm}|p{6.88cm}||p{0.29cm}|p{6cm}}
+ \textbf{ID} & \textbf{Feature Description} & \textbf{ID} & \textbf{Feature Description}\\ 
+ \hline
+1 & $\sum\nolimits_{q_i \in q \cap d}c(q_i, d)$ of title					& 24&  $\sum\nolimits_{q_i \in q \cap d}c(q_i,d)\cdot log(\frac{|C|}{df(q_i)})$ of abstract\\
+2 &	$\sum\nolimits_{q_i \in q \cap d}log(c(q_i, d)+1)$ of title				& 25& $\sum\nolimits_{q_i \in q \cap d} log(\frac{c(q_i,d)}{|d|}) \cdot \frac{|C|}{c(q_i,C)}+1$ of abstract\\
+3 & $\sum\nolimits_{q_i \in q \cap d}\frac{c(q_i,d)}{|d|}$ of title				& 26& BM25 of abstract\\
+4 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|}+1)$ of title				& 27& $log($BM25$)$ of abstract\\
+5 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{df(q_i)})$ of title		& 28& LMIR.DIR of abstract\\
 
-The Yahoo!
-Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning-to-Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country.
-For this experiment, I will only look at the standard Learning-to-Rank dataset, because transfer learning is a separate research area that is not included in this thesis.\\
-\begin{table}[!h]
-\begin{tabular}{l|lll}
- & Train & Validation & Test \\ 
- \hline
-Queries & 19,994 & 2,994 & 6,983 \\ 
-Documents & 473,134 & 71,083 & 165,660 \\ 
-Features & 519 & - & - \\ 
+6 & $\sum\nolimits_{q_i \in q \cap d}log(log(\frac{|C|}{df(q_i)}))$ of title 				& 29& LMIR.JM of abstract\\
+7 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{c(q_i,C)}+1)$ of title 				& 30& LMIR.ABS of abstract\\
+8 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|} \cdot \frac{|C|}{df(q_i)}+1)$ of title 				& 31& $\sum\nolimits_{q_i \in q \cap d}c(q_i, d)$ of title + abstract\\
+9 & $\sum\nolimits_{q_i \in q \cap d}c(q_i,d)\cdot log(\frac{|C|}{df(q_i)})$ of title 			& 32&	$\sum\nolimits_{q_i \in q \cap d}log(c(q_i, d)+1)$ of title + abstract\\
+10& $\sum\nolimits_{q_i \in q \cap d} log(\frac{c(q_i,d)}{|d|}) \cdot \frac{|C|}{c(q_i,C)}+1$ of title		& 33& $\sum\nolimits_{q_i \in q \cap d}\frac{c(q_i,d)}{|d|}$ of title + abstract\\
+
+11& BM25 of title 			& 34& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|}+1)$ of title + abstract\\
+12& $log($BM25$)$ of title 			& 35& $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{df(q_i)})$ of title + abstract\\
+13& LMIR.DIR of title 			& 36& $\sum\nolimits_{q_i \in q \cap d}log(log(\frac{|C|}{df(q_i)}))$ of title + abstract\\
+14& LMIR.JM of title			& 37& $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{c(q_i,C)}+1)$ of title + abstract\\
+15& LMIR.ABS of title	& 38& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|} \cdot \frac{|C|}{df(q_i)}+1)$ of title + abstract\\
+
+16& $\sum\nolimits_{q_i \in q \cap d}c(q_i, d)$ of abstract		& 39& $\sum\nolimits_{q_i \in q \cap d}c(q_i,d)\cdot log(\frac{|C|}{df(q_i)})$ of title + abstract\\
+17& $\sum\nolimits_{q_i \in q \cap d}log(c(q_i, d)+1)$ of abstract		& 40& $\sum\nolimits_{q_i \in q \cap d} log(\frac{c(q_i,d)}{|d|}) \cdot \frac{|C|}{c(q_i,C)}+1$ of title + abstract\\
+18& $\sum\nolimits_{q_i \in q \cap d}\frac{c(q_i,d)}{|d|}$ of abstract		& 41& BM25 of title + abstract\\
+19& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|}+1)$ of abstract		& 42& $log($BM25$)$ of title + abstract\\
+20& $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{df(q_i)})$ of abstract	& 43& LMIR.DIR of title + abstract\\
+
+21& $\sum\nolimits_{q_i \in q \cap d}log(log(\frac{|C|}{df(q_i)}))$ of abstract					& 44& LMIR.JM of title + abstract\\
+22&  $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{c(q_i,C)}+1)$ of abstract					& 45& LMIR.ABS of title + abstract\\
+23& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|} \cdot \frac{|C|}{df(q_i)}+1)$ of abstract					& & \\
 \end{tabular}
-\caption{Yahoo! Learning to Rank Challenge dataset characteristics, as described in the challenge overview paper \cite{Chapelle2011a}}
-\label{tab:yahoo_characteristics}
-\end{table}\\
-Both \ac{nDCG} and \ac{ERR} are measured as performance metrics, but the final standings of the challenge were based on the \ac{ERR} values.
-Model validation on the Learning-to-Rank methods participating in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}.
-Competitors could train on the training set and get immediate feedback on their performance on the validation set.
-The test set performance is used to create the final standings and is only measured after the competition has ended to avoid overfitting on the test set.
-The large number of documents, queries and features compared to other benchmark datasets makes the Yahoo!
-Learning to Rank Challenge dataset interesting.
-Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo.
-Instead high level descriptions of feature categories are provided.
-The following categories of features are described in the challenge overview paper \cite{Chapelle2011a}:\\
-\begin{description}
-\item[Web graph]{Quality and popularity metrics of web documents, e.g. PageRank \cite{Page1999}}.
-\item[Document statistics]{Basic document statistics such as the number of words and url characteristics.}
-\item[Document classifier]{Results of various classifiers on the documents. These classifiers amongst others include: spam, adult, language, main topic, and quality classifiers.}
-\item[Query]{Basic query statistics, such as the number of terms, query frequency, and click-through rate.}
-\item[Text match]{Textual similarity metrics between query and document. Includes \ac{TF-IDF}, BM25 \cite{Robertson2009} and other metrics for different sections of the document.}
-\item[Topical matching]{These features go beyond similarity at word level and compute similarity on topic level.
-For example by classifying both the document and the query in a large topical taxonomy.}
-\item[Click]{Click-based user feedback.}
-\item[External references]{Document meta-information such as Delicious\footnote{https://delicious.com/} tags}
-\item[Time]{Document age and historical in- and outlink data that might help for time sensitive queries.}
-\end{description}
-
-\section{Results}
-Figure \ref{fig:yahoo_results} shows the top five participants in the Yahoo!
-Learning to Rank Challenge in terms of ERR score.
-The top five participants all used decision trees and ensemble methods.
-Burges et al \cite{Burges2011} created a linear combination ensemble of eight LambdaMART \cite{Burges2010}, two LambdaRank and two Logistic Regression models.
-Gottschalk and Vogel used a combination of RandomForest models and \ac{GBDT} models.
-Pavlov and Brunk used a regression based model using the BagBoo \cite{Pavlov2010} ensemble technique, which combines bagging and boosting.
-Sorokina used a similar combination of bagging and boosting that is called Additive Groves \cite{Sorokina2007}.\\
-
-The challenge overview paper states as one of the lessons of the challenge that the simple baseline \ac{GBDT} model performed very well with a small performance gap to the complex ensemble submissions at the top of the table.\\
+}
+\caption{Features of the LETOR OHSUMED dataset, obtained from Qin et al \cite{Qin2010}}
+\label{tbl:features_ohsumed}
+\end{table}
 
 \begin{table}
-\begin{tabular}{l|p{6.3cm}|l}
- & Authors & ERR \\
- \hline 
-1 & Burges et al (Microsoft Research) & 0.46861 \\ 
-2 & Gottschalk (Activision Blizzard) \& Vogel (Data Mining Solutions) & 0.46786 \\ 
-3 & Parakhin (Microsoft) & 0.46695 \\ 
-4 & Pavlov \& Brunk (Yandex Labs) & 0.46678 \\ 
-5 & Sorokina (Yandex Labs) & 0.46616 \\ 
-\end{tabular}
-\caption{Final standings of the Yahoo! Learning to Rank Challenge, as presented in the challenge overview paper \cite{Chapelle2011a}}
-\label{fig:yahoo_results}
-\end{table}
-
-Although the winning \ac{GBDT} models performs very well in terms of \ac{nDCG} their high complexity makes them unsuitable to use in production.
-The winning models take weeks to train and are very slow during query evaluation.
-An exception in training time is the BagBoo \cite{Pavlov2010} method used by Pavlov \& Brunk.
-The bagging component of the BagBoo method enables it to achieve high scalability through parallelism.
-Pavlov et al \cite{Pavlov2010} managed to train half a million trees on 200 nodes in 2 hours.
-\chapter{Yandex Internet Mathematics competition}
-
-\section{Results}
-The Yandex Internet Mathematics competition was won by Pavlov et al \cite{Pavlov2010} with the BagBoo method.
-
-\chapter{LETOR}
-The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 to solve the absence of a experimental platform for Learning-to-Rank at that time.
-LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009.
-The LETOR3.0 benchmark collection \cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.\\
-
-\section{LETOR3.0}
-The OHSUMED collection is a subset of the medical publication database MEDLINE and contains medical publications from 270 journals that were published between 1987 and 1991.
-The .gov collection is a web crawl obtained in January 2002, which was used for the TREC web track in 2003 and 2004.
-Tables \ref{tbl:features_gov} and \ref{tbl:features_ohsumed} provide the descriptions of the features of the .gov and the OHSUMED collections respectively.\\
-
-Different query sets exists for the .gov corpus. Those query sets are categorized in the following categories \cite{}:
-\begin{description}
-\item[topic distillation (TD)]these queries involve finding relevant pages given a broad query.
-E.g., the query 'cotton industry' which is entered with the aim to find pages that provide information about the cotton industry.
-\item[named page finding (NP)]in these queries the user asks for one specific page by name. E.g., the user queries for 'ADA Enforcement' to find the page $http://www.usdoj.gov/crt/ada/enforce.htm$.
-\item[home page finding (HP)]these queries are similar to NP queries in that the user is looking for one specific page, but now this specific page is a homepage.
-E.g., the user queries for 'Tennessee Valley Authority' to find the homepage $http://www.tva.gov$.
-\end{description}
-With query sets available from both the TREC 2003 and 2004 conferences there are six query sets in total: TD2003, TD2004, NP2003, NP2004, HP2003 and HP2004.\\
-
-The evaluation metrics used are \ac{nDCG} and \ac{MAP}.
-The \emph{winning number} metric is defined as the number of other algorithms that it can beat over all of the seven datasets (six .gov sets + OHSUMED).
-The LETOR organisation implemented and evaluated several well-known Learning-to-Rank algorithms themselves and in addition gathers and publications and results of new algorithms evaluated on the LETOR benchmark.
-The baseline Learning-to-Rank algorithms evaluated by the organisation are:
-\begin{description}
-\item[pointwise]{\leavevmode
-	\begin{itemize}
-	\item Linear regression
-	\end{itemize}}
-\item[pairwise]{\leavevmode
-	\begin{itemize}
-	\item RankingSVM \cite{Herbrich1999,Joachims2002}
-	\item RankBoost \cite{Freund2003}
-	\item FRank \cite{Tsai2007}
-	\end{itemize}}
-\item[listwise]{\leavevmode
-	\begin{itemize}
-	\item ListNet \cite{Cao2007}
-	\item AdaRank-MAP \cite{Xu2007}
-	\item AdaRank-nDCG \cite{Xu2007}
-	\item SVMMAP \cite{Yue2007} 
-	\end{itemize}}
-\end{description} 
-\begin{table}[!h]
 \scalebox{0.87}{
@@
 \end{table}
-
-\begin{table}[!h]
-\scalebox{0.87}{
-\begin{tabular}{p{0.29cm}|p{6.88cm}||p{0.29cm}|p{6.90cm}}
- \textbf{ID} & \textbf{Feature Description} & \textbf{ID} & \textbf{Feature Description}\\ 
- \hline
-1 & $\sum\nolimits_{q_i \in q \cap d}c(q_i, d)$ of title					& 24&  $\sum\nolimits_{q_i \in q \cap d}c(q_i,d)\cdot log(\frac{|C|}{df(q_i)})$ of abstract\\
-2 &	$\sum\nolimits_{q_i \in q \cap d}log(c(q_i, d)+1)$ of title				& 25& $\sum\nolimits_{q_i \in q \cap d} log(\frac{c(q_i,d)}{|d|}) \cdot \frac{|C|}{c(q_i,C)}+1$ of abstract\\
-3 & $\sum\nolimits_{q_i \in q \cap d}\frac{c(q_i,d)}{|d|}$ of title				& 26& BM25 of abstract\\
-4 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|}+1)$ of title				& 27& $log($BM25$)$ of abstract\\
-5 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{df(q_i)})$ of title		& 28& LMIR.DIR of abstract\\
-
-6 & $\sum\nolimits_{q_i \in q \cap d}log(log(\frac{|C|}{df(q_i)}))$ of title 				& 29& LMIR.JM of abstract\\
-7 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{c(q_i,C)}+1)$ of title 				& 30& LMIR.ABS of abstract\\
-8 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|} \cdot \frac{|C|}{df(q_i)}+1)$ of title 				& 31& $\sum\nolimits_{q_i \in q \cap d}c(q_i, d)$ of title + abstract\\
-9 & $\sum\nolimits_{q_i \in q \cap d}c(q_i,d)\cdot log(\frac{|C|}{df(q_i)})$ of title 			& 32&	$\sum\nolimits_{q_i \in q \cap d}log(c(q_i, d)+1)$ of title + abstract\\
-10& $\sum\nolimits_{q_i \in q \cap d} log(\frac{c(q_i,d)}{|d|}) \cdot \frac{|C|}{c(q_i,C)}+1$ of title		& 33& $\sum\nolimits_{q_i \in q \cap d}\frac{c(q_i,d)}{|d|}$ of title + abstract\\
-
-11& BM25 of title 			& 34& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|}+1)$ of title + abstract\\
-12& $log($BM25$)$ of title 			& 35& $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{df(q_i)})$ of title + abstract\\
-13& LMIR.DIR of title 			& 36& $\sum\nolimits_{q_i \in q \cap d}log(log(\frac{|C|}{df(q_i)}))$ of title + abstract\\
-14& LMIR.JM of title			& 37& $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{c(q_i,C)}+1)$ of title + abstract\\
-15& LMIR.ABS of title	& 38& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|} \cdot \frac{|C|}{df(q_i)}+1)$ of title + abstract\\
-
-16& $\sum\nolimits_{q_i \in q \cap d}c(q_i, d)$ of abstract		& 39& $\sum\nolimits_{q_i \in q \cap d}c(q_i,d)\cdot log(\frac{|C|}{df(q_i)})$ of title + abstract\\
-17& $\sum\nolimits_{q_i \in q \cap d}log(c(q_i, d)+1)$ of abstract		& 40& $\sum\nolimits_{q_i \in q \cap d} log(\frac{c(q_i,d)}{|d|}) \cdot \frac{|C|}{c(q_i,C)}+1$ of title + abstract\\
-18& $\sum\nolimits_{q_i \in q \cap d}\frac{c(q_i,d)}{|d|}$ of abstract		& 41& BM25 of title + abstract\\
-19& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|}+1)$ of abstract		& 42& $log($BM25$)$ of title + abstract\\
-20& $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{df(q_i)})$ of abstract	& 43& LMIR.DIR of title + abstract\\
-
-21& $\sum\nolimits_{q_i \in q \cap d}log(log(\frac{|C|}{df(q_i)}))$ of abstract					& 44& LMIR.JM of title + abstract\\
-22&  $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{c(q_i,C)}+1)$ of abstract					& 45& LMIR.ABS of title + abstract\\
-23& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|} \cdot \frac{|C|}{df(q_i)}+1)$ of abstract					& & \\
-\end{tabular}
-}
-\caption{Features of the LETOR OHSUMED dataset, obtained from Qin et al \cite{Qin2010}}
-\label{tbl:features_ohsumed}
-\end{table}
-\subsection{Results}
-The LETOR paper \cite{Qin2010} describes the performance of the LETOR baseline methods.
-Figures \ref{fig:ndcg_winning_number} and \ref{fig:map_winning_number} show ListNet to be the best performing baseline in terms of winning number on both \ac{nDCG} and \ac{MAP}.
-The LETOR website lists\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} a few algorithms that have since been evaluated on the LETOR benchmark collection.\\
-
-\begin{figure}[!h]
-\includegraphics[scale=0.30]{gfx/ndcg_winning_number}
-\caption{Comparison across the seven datasets in LETOR by \ac{nDCG}, obtained from Qin et al \cite{Qin2010}}
-\label{fig:ndcg_winning_number}
-\end{figure}
-
-\begin{figure}[!h]
-\includegraphics[scale=0.30]{gfx/map_winning_number}
-\caption{Comparison across the seven datasets in LETOR by \ac{MAP}, obtained from Qin et al \cite{Qin2010}}
-\label{fig:map_winning_number}
-\end{figure}
-
-I will describe the performance of the algorithms that were evaluated on LETOR3.0 after publication of the LETOR paper by Qin et al \cite{Qin2010} as listed on the LETOR website\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} by comparing their performance with the ListNet baseline.
-We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven datasets in \ac{nDCG} value.
-Note that this does not necessarily imply that these methods would have scored a higher \ac{nDCG} winning number than ListNet.
-Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR datasets in terms of \ac{nDCG}@10 and \ac{MAP}.
-\begin{table}
-\begin{tabular}{l|ll}
- & \ac{nDCG}@10 & \ac{MAP} \\ 
- \hline
-TD2003 & 0.348 & 0.2753 \\ 
-TD2004 & 0.317 & 0.2231 \\ 
-NP2003 & 0.801 & 0.6895 \\ 
-NP2004 & 0.812 & 0.6720 \\ 
-HP2003 & 0.837 & 0.7659 \\ 
-HP2004 & 0.784 & 0.6899 \\ 
-OHSUMED & 0.441 & 0.4457 \\ 
-\end{tabular}
-\caption{Performance of ListNet on LETOR}
-\label{tbl:LETOR_ListNet}
-\end{table}
-
-\begin{table}
-\begin{tabular}{l|p{1.2cm}p{1.2cm}p{1.2cm}p{1.4cm}||l}
- & \rotatebox{55}{Regression + L2 regulatisation} & \rotatebox{55}{RankSVM-Primal} &\rotatebox{55}{RankSVM-Struct} & \rotatebox{55}{SmoothRank} & \rotatebox{55}{ListNet} \\
- \hline
-TD2003 & 0.3297 & 0.3571 & 0.3467 & 0.3367 & 0.348 \\ 
-TD2004 & 0.2832 & 0.2913 & 0.3090 & 0.3343 & 0.317 \\ 
-NP2003 & 0.8025 & 0.7894 & 0.7955 & 0.7986 & 0.801 \\ 
-NP2004 & 0.8040 & 0.7950 & 0.7977 & 0.8075 & 0.812 \\ 
-HP2003 & 0.8216 & 0.8180 & 0.8162 & 0.8325 & 0.837 \\ 
-HP2004 & 0.7188 & 0.7720 & 0.7666 & 0.8221 & 0.784 \\ 
-OHSUMED & 0.4436 & 0.4504 & 0.4523 & 0.4568 & 0.441 \\ 
-\# winning datasets & 2 & 2 & 1 & 3 & - \\ 
-\end{tabular}
-\caption{Comparison of algorithms recently evaluated on LETOR3.0 with the ListNet baselines}
-\label{tbl:LETOR_recently_added}
-\end{table}
-\section{LETOR4.0}
-\subsection{Results}
-BoostedTree model \cite{Kocsis2013} showed an \ac{nDCG} of 0.5071 on the MQ2007 dataset and thereby beat all baseline models.
-
-\chapter{MSLR-WEB10k/MSLR-WEB30k}
-Contrary to the Yahoo!
-Learning to Rank Challenge dataset, the MSLR-WEB30k provides detailed feature descriptions.
-The MSLR-WEB30k dataset however contains no proprietary features but only features that are commonly used in the research community.
-\section{Results}
-
-\chapter{Selecting Learning-to-Rank methods}
-The Accuracies of the Learning-to-Rank methods described in the preceding chapters must only be compared within the benchmark and not between benchmarks for the following reasons:
-\begin{enumerate}
-\item Differences in feature sets between data sets detract from fair comparison
-\item Although the \ac{nDCG} definition is unambiguous, Busa-Fekete et al \cite{Busa-Fekete2012} found that \ac{nDCG} evaluation tools of benchmark data sets produced different scores
-\end{enumerate}

### 2014-03-07 report/Chapters/BenchmarkResults.tex
cranfield referentie + grote tabellen naar appendix

--- 2014-03-07 report/BenchmarkResults.tex
@@
 \label{chap:benchmark_results}
+This chapter describes benchmark characteristics like collection size and features and gives an overview of the performance of different Learning-to-Rank methods.
+Performance differences for a given Learning-to-Rank method are explained in terms of differences in benchmark characteristics.
+This chapter is concluded with a selection of Learning-to-Rank methods to include in the experiments.\\
+
+Benchmarking datasets enable fair comparisons of ranking models over a fixed set of documents and queries.
+The approach of using fixed benchmarking datasets to compare the performance of Information Retrieval systems under equal circumstances has been the standard evaluation methodology in the Information Retrieval field since the release of the Cranfield collection \cite{Cleverdon1966}.
 \section{Yahoo! Learning to Rank Challenge}
@@
 The .gov collection is a web crawl obtained in January 2002, which was used for the TREC web track in 2003 and 2004.
-Tables \ref{tbl:features_gov} and \ref{tbl:features_ohsumed} provide the descriptions of the features of the .gov and the OHSUMED collections respectively.\\
+Tables \ref{tbl:features_gov} and \ref{tbl:features_ohsumed} in Appendix \ref{chap:appendix_a} provide the descriptions of the features of the .gov and the OHSUMED collections respectively.\\
 
@@
 \end{description}
-\begin{table}[!h]
-\scalebox{0.87}{
-\begin{tabular}{p{0.29cm}|p{7.26cm}||p{0.29cm}|p{9.55cm}}
- \textbf{ID} & \textbf{Feature Description} & \textbf{ID} & \textbf{Feature Description}\\ 
- \hline
-1 & \ac{TF} of body					& 36& LMIR.JM of body\\
-2 &	\ac{TF} of anchor				& 37& LMIR.JM of anchor\\
-3 & \ac{TF} of title				& 38& LMIR.JM of title\\
-4 & \ac{TF} of \ac{URL}				& 39& LMIR.JM of \ac{URL}\\
-5 & \ac{TF} of whole document		& 40& LMIR.JM of whole document\\
 
-6 & \ac{IDF} of body 				& 41& Sitemap based term propagation\\
-7 & \ac{IDF} of anchor 				& 42& Sitemap based score propagation\\
-8 & \ac{IDF} of title 				& 43& Hyperlink based score propagation: weighted in-link\\
-9 & \ac{IDF} of \ac{URL} 			& 44& Hyperlink based score propagation: weighted out-link\\
-10& \ac{IDF} of whole document 		& 45& Hyperlink based score propagation: uniform out-link\\
-
-11& \ac{TF-IDF} of body 			& 46& Hyperlink based feature propagation: weighted in-link\\
-12& \ac{TF-IDF} of anchor 			& 47& Hyperlink based feature propagation: weighted out-link\\
-13& \ac{TF-IDF} of title 			& 48& Hyperlink based feature propagation: uniform out-link\\
-14& \ac{TF-IDF} of \ac{URL}			& 49& HITS authority\\
-15& \ac{TF-IDF} of whole document	& 50& HITS hub\\
-
-16& Document length of body			& 51& PageRank\\
-17& Document length of anchor		& 52& HostRank\\
-18& Document length of title		& 53& Topical PageRank\\
-19& Document length of \ac{URL}		& 54& Topical HITS authority\\
-20& Document length of whole document	& 55& Topical HITS hub\\
-
-21& BM25 of body					& 56& In-link number\\
-22& BM25 of anchor					& 57& Out-link number\\
-23& BM25 of title					& 58& Number of slashes in \ac{URL}\\
-24& BM25 of \ac{URL}				& 59& Length of \ac{URL}\\
-25& BM25 of whole document			& 60& Number of child page\\
-
-26& LMIR.ABS \cite{Zhai2001} of body	& 61& BM25 of extracted title\\
-27& LMIR.ABS of anchor				& 62& LMIR.ABS of extracted title\\
-28& LMIR.ABS of title				& 63& LMIR.DIR of extracted title\\
-29& LMIR.ABS of \ac{URL}			& 64& LMIR.JM of extracted title\\
-30& LMIR.ABS of whole document\\
-
-31& LMIR.DIR of body\\
-32& LMIR.DIR of anchor\\
-33& LMIR.DIR of title\\
-34& LMIR.DIR of \ac{URL}\\
-35& LMIR.DIR of whole document\\
-\end{tabular}
-}
-\caption{Features of the LETOR .GOV dataset, obtained from Qin et al \cite{Qin2010}}
-\label{tbl:features_gov}
-\end{table}
-
-\begin{table}[!h]
-\scalebox{0.87}{
-\begin{tabular}{p{0.29cm}|p{6.88cm}||p{0.29cm}|p{6.90cm}}
- \textbf{ID} & \textbf{Feature Description} & \textbf{ID} & \textbf{Feature Description}\\ 
- \hline
-1 & $\sum\nolimits_{q_i \in q \cap d}c(q_i, d)$ of title					& 24&  $\sum\nolimits_{q_i \in q \cap d}c(q_i,d)\cdot log(\frac{|C|}{df(q_i)})$ of abstract\\
-2 &	$\sum\nolimits_{q_i \in q \cap d}log(c(q_i, d)+1)$ of title				& 25& $\sum\nolimits_{q_i \in q \cap d} log(\frac{c(q_i,d)}{|d|}) \cdot \frac{|C|}{c(q_i,C)}+1$ of abstract\\
-3 & $\sum\nolimits_{q_i \in q \cap d}\frac{c(q_i,d)}{|d|}$ of title				& 26& BM25 of abstract\\
-4 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|}+1)$ of title				& 27& $log($BM25$)$ of abstract\\
-5 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{df(q_i)})$ of title		& 28& LMIR.DIR of abstract\\
-
-6 & $\sum\nolimits_{q_i \in q \cap d}log(log(\frac{|C|}{df(q_i)}))$ of title 				& 29& LMIR.JM of abstract\\
-7 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{c(q_i,C)}+1)$ of title 				& 30& LMIR.ABS of abstract\\
-8 & $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|} \cdot \frac{|C|}{df(q_i)}+1)$ of title 				& 31& $\sum\nolimits_{q_i \in q \cap d}c(q_i, d)$ of title + abstract\\
-9 & $\sum\nolimits_{q_i \in q \cap d}c(q_i,d)\cdot log(\frac{|C|}{df(q_i)})$ of title 			& 32&	$\sum\nolimits_{q_i \in q \cap d}log(c(q_i, d)+1)$ of title + abstract\\
-10& $\sum\nolimits_{q_i \in q \cap d} log(\frac{c(q_i,d)}{|d|}) \cdot \frac{|C|}{c(q_i,C)}+1$ of title		& 33& $\sum\nolimits_{q_i \in q \cap d}\frac{c(q_i,d)}{|d|}$ of title + abstract\\
-
-11& BM25 of title 			& 34& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|}+1)$ of title + abstract\\
-12& $log($BM25$)$ of title 			& 35& $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{df(q_i)})$ of title + abstract\\
-13& LMIR.DIR of title 			& 36& $\sum\nolimits_{q_i \in q \cap d}log(log(\frac{|C|}{df(q_i)}))$ of title + abstract\\
-14& LMIR.JM of title			& 37& $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{c(q_i,C)}+1)$ of title + abstract\\
-15& LMIR.ABS of title	& 38& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|} \cdot \frac{|C|}{df(q_i)}+1)$ of title + abstract\\
-
-16& $\sum\nolimits_{q_i \in q \cap d}c(q_i, d)$ of abstract		& 39& $\sum\nolimits_{q_i \in q \cap d}c(q_i,d)\cdot log(\frac{|C|}{df(q_i)})$ of title + abstract\\
-17& $\sum\nolimits_{q_i \in q \cap d}log(c(q_i, d)+1)$ of abstract		& 40& $\sum\nolimits_{q_i \in q \cap d} log(\frac{c(q_i,d)}{|d|}) \cdot \frac{|C|}{c(q_i,C)}+1$ of title + abstract\\
-18& $\sum\nolimits_{q_i \in q \cap d}\frac{c(q_i,d)}{|d|}$ of abstract		& 41& BM25 of title + abstract\\
-19& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|}+1)$ of abstract		& 42& $log($BM25$)$ of title + abstract\\
-20& $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{df(q_i)})$ of abstract	& 43& LMIR.DIR of title + abstract\\
-
-21& $\sum\nolimits_{q_i \in q \cap d}log(log(\frac{|C|}{df(q_i)}))$ of abstract					& 44& LMIR.JM of title + abstract\\
-22&  $\sum\nolimits_{q_i \in q \cap d}log(\frac{|C|}{c(q_i,C)}+1)$ of abstract					& 45& LMIR.ABS of title + abstract\\
-23& $\sum\nolimits_{q_i \in q \cap d}log(\frac{c(q_i,d)}{|d|} \cdot \frac{|C|}{df(q_i)}+1)$ of abstract					& & \\
-\end{tabular}
-}
-\caption{Features of the LETOR OHSUMED dataset, obtained from Qin et al \cite{Qin2010}}
-\label{tbl:features_ohsumed}
-\end{table}
 \subsection{Results}

### 2014-03-07 report/Chapters/Implementation.tex
cranfield referentie + grote tabellen naar appendix

Describes implementation details of the Learning-to-Rank algorithm parallel implementations in HDInsight that are either Hadoop, Microsoft Azure, or HDInsight and are not part of the algorithm specification itself.

### 2014-03-07 report/Chapters/LTRMethods.tex
cranfield referentie + grote tabellen naar appendix

Algorithms and details of the well-performing Learning-to-Rank methods as selected in the afore-going part are presented and explained in this part.

### 2014-03-10 report/Chapters/Background.tex
cranfield referentie + grote tabellen naar appendix
voorbeelden met referenties bij ltr approaches

@@
 %************************************************
+This chapter provides a background in the Learning-to-Rank field with the goal of making the subsequential chapters of this thesis understandable for non-experts in the field.
 \section{A basic introduction to Learning-to-Rank}
@@
 	\end{enumerate}
+Well-known algorithms that belong to the pointwise approach include McRank \cite{Li2007} and PRank \cite{Crammer2001}.
 \subsection{Pairwise Approach}
@@
 Aggregating a set of predicted pairwise preferences into the corresponding optimal rank is shown to be a NP-Hard problem \cite{Feldman2012}.
-An often used solution to this problem is to upper bound the number of classification mistakes by an easy to optimise function \cite{Bartlett2006}.
+An often used solution to this problem is to upper bound the number of classification mistakes by an easy to optimise function \cite{Bartlett2006}.\\
+
+Well-known pairwise Learning-to-Rank algorithms include FRank \cite{Tsai2007}, GBRank \cite{Zheng2007}, LambdaRank \cite{Burges2006}, RankBoost \cite{Freund2003}, RankNet \cite{Burges2005}, Ranking SVM \cite{Herbrich1999b,Joachims2002}, and SortNet \cite{Rigutini2008}.
 \subsection{Listwise Approach}
@@
 Although the properties of \ac{MAP}, \ac{ERR} and \ac{nDCG} are not ideal for direct optimisation, some listwise approaches do focus on direct metric optimisation \cite{Yue2007, Taylor2008, Chapelle2010}.
-Most listwise approaches work around optimisation of the non-differentiable, non-convex and discontinuous metrics by optimising surrogate cost functions that mimic the behaviour of \ac{MAP}, \ac{ERR} or \ac{nDCG}, but have nicer properties for optimisation.
+Most listwise approaches work around optimisation of the non-differentiable, non-convex and discontinuous metrics by optimising surrogate cost functions that mimic the behaviour of \ac{MAP}, \ac{ERR} or \ac{nDCG}, but have nicer properties for optimisation.\\
+
+Well-known algorithms that belong to the listwise approach include AdaRank \cite{Xu2007}, BoltzRank \cite{Volkovs2009}, ListMLE \cite{Xia2008}, ListNet \cite{Cao2007}, RankCosine \cite{Qin2008}, SmoothRank \cite{Chapelle2010}, SoftRank \cite{Taylor2008}, SVM$^map$ \cite{Yue2007}.

### 2014-03-10 report/Chapters/RelatedWork.tex
voorbeelden met referenties bij ltr approaches
begin categorisatie related work + inleiding ensemble methods

--- 2014-03-07 report/RelatedWork.tex
@@
 \label{sec:lit_descr}
+
+\subsection{Parallelising Learning-to-Rank through ensemble learning}
+Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure that he called boosting \cite{Schapire1990}.
+The invention of the boosting method resulted in an increasing focus in machine learning research on methods to combine multiple weak models, which are called ensemble learning methods.
+Well-known ensemble methods include boosting \cite{Schapire1990}, bagging \cite{Breiman1996}, AdaBoost \cite{Freund1997} and stacked generalisation (stacking) \cite{Wolpert1992}.\\
+
+Ensemble methods can be used to parallelise learning methods by training the weak models in the ensemble on different nodes.
+In the Learning-to-Rank field parallelisation of model training through ensemble learning is frequently achieved using decision tree learners and often using the boosting ensemble method, jointly called \acf{GBDT}.
+\ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
+\subsubsection{Stochastic Boosting}
+Ye et al.
+\cite{Ye2009} described how to implement the \ac{SGBDT} process in a parallel manner using both MPI and Hadoop.
+Experiments showed the Hadoop implementation to result into too expensive communication cost to be useful.
+Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
+The MPI approach proved to be successful and obtained near linear speed-ups.\\
+
+Kocsis et al.
+\cite{Kocsis2013} propose way to train multiple weak learners in parallel and extend those models that are likely to yield a good model when combined through boosting.
+Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular boosting.
+Using this parallel BoostingTree algorithm \ac{GBDT} models could be trained in parallel.\\
+
+Tyree et al.
+\cite{Tyree2011} describe a way of parallelising \ac{GBDT} models for Learning-to-Rank where the boosting step is still executed sequentially, instead the construction of the regression trees themselves are parallelised.
+The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}.
+Decision trees are build layer-by-layer where the calculations needed for building each new layer in the tree are divided over the nodes using a master-worker paradigm.
+The data is partitioned of the workers, who compress their share into histograms and send these to the master.
+The master uses those histograms to approximate the split and build the next layer.
+The master than communicates this new layer to the workers who can use this new layer to compute new histograms.
+this process is repeated until the tree depth limit is reached.
+The tree construction algorithm parallelised with described master-worker method is \ac{CART} \cite{Breiman1984}.
+Speed-up experiments on the LETOR and the Yahoo!
+Learning to Rank challenge data sets were performed.
+This parallel \ac{CART}-tree building approach showed speed-up of up to 42x on shared memory machines and up to 25x on distributed memory machines.
+
+\subsubsection{Stacked generalisation}
+A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.
+\ac{DSN} is based on the stacked generalization ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack.
+Each layer in the stack learns has the task of learning the weights of the inputs of that layer.
+The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.
+
+\subsubsection{Boosting wrapped in Bagging}
+Pavlov et al.
+\cite{Pavlov2010} at Yandex developed a hybrid model that combines two methods from ensemble learning: bagging and boosting.
+Bagging \cite{Breiman1996}, also called bootstrap aggregating, is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement.
+The bagging method is parallelisable by training each $D_i$ with $i \in \{1..m\}$ on a different node.
+Boosting \cite{Friedman2002} is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.
+Boosting is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.
+
 \subsection{CCRank}
@@
 The speed-up achieved at learning the ranking function was unfortunately not stated.
-\subsection{BagBoo: A hybrid Bagging-the-Boosting Model}
-Pavlov et al.
-\cite{Pavlov2010} at Yandex developed a hybrid model that combines two methods from ensemble learning: bagging and boosting.
-Bagging \cite{Breiman1996}, also called bootstrap aggregating, is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement.
-The bagging method is parallelisable by training each $D_i$ with $i \in \{1..m\}$ on a different node.
-Boosting \cite{Friedman2002} is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.
-Boosting is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.
-\subsection{Parallel Decision Trees}
-Ye et al.
-\cite{Ye2009} described how to implement the \ac{GBDT} process in a parallel manner using both MPI and Hadoop.
-\ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
-Experiments showed the Hadoop implementation to result into too expensive communication cost to be useful.
-Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
-The MPI approach proved to be successful and obtained near linear speed-ups.\\
 
-Kocsis et al.
-\cite{Kocsis2013} propose way to train multiple weak learners in parallel and extend those models that are likely to yield a good model when combined through boosting.
-Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular boosting.
-Using this parallel BoostingTree algorithm \ac{GBDT} models could be trained in parallel.\\
-
-Tyree et al.
-\cite{Tyree2011} describe a way of parallelising \ac{GBDT} models for Learning-to-Rank where the boosting step is still executed sequentially, instead the construction of the regression trees themselves are parallelised.
-The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}.
-Decision trees are build layer-by-layer where the calculations needed for building each new layer in the tree are divided over the nodes using a master-worker paradigm.
-The data is partitioned of the workers, who compress their share into histograms and send these to the master.
-The master uses those histograms to approximate the split and build the next layer.
-The master than communicates this new layer to the workers who can use this new layer to compute new histograms.
-this process is repeated until the tree depth limit is reached.
-The tree construction algorithm parallelised with described master-worker method is \ac{CART} \cite{Breiman1984}.
-Speed-up experiments on the LETOR and the Yahoo!
-Learning to Rank challenge data sets were performed.
-This parallel \ac{CART}-tree building approach showed speed-up of up to 42x on shared memory machines and up to 25x on distributed memory machines.
-\subsection{Deep Stacking Networks}
-A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.
-\ac{DSN} is based on the stacked generalization ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack.
-Each layer in the stack learns has the task of learning the weights of the inputs of that layer.
-The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.
 \subsection{Alternating Direction Method of Multipliers}
@@
 Ganjisaffar et al.
-\cite{Ganjisaffar2011} observed that long training times are often a result of training many models for hyperparameter optimisation.
+\cite{Ganjisaffar2011, Ganjisaffar2011b} observed that long training times are often a result of training many models for hyperparameter optimisation.
 Grid search is the \emph{de facto} standard of hyperparameter optimisation and is simply an exhaustive search through a manually specified subset of hyperparameter combinations.

### 2014-03-10 report/Chapters/RelatedWork.tex
reference fixes + herschuiving related work

@@
 Boosting \cite{Friedman2002} is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.
-Boosting is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.
+Boosting is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.\\
+
+\subsection{Low complexity Learning-to-Rank}
+Designing Learning-to-Rank algorithms with low time complexity for training is another approach towards large scale Learning-to-Rank.
+Pahikkala et al.
+\cite{Pahikkala2009} describe a pairwise \ac{RLS} type of ranking function, RankRLS, with time complexity $\mathcal{O}(n^3+n^2m)$, with $n$ the number of features and $m$ the number of training documents.
+The RankRLS ranking function showed ranking performance similar to RankSVM on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.
 
@@
 Unfortunately their work does not directly report on the speed-up obtained by parallelising  with Hadoop, but it is mentioned that further work needs to be done to effectively leverage parallel execution.
-\subsection{Low complexity Learning-to-Rank}
-Designing Learning-to-Rank algorithms with low time complexity for training is another approach towards large scale Learning-to-Rank.
-Pahikkala et al.
-\cite{Pahikkala2009} describe a pairwise \ac{RLS} type of ranking function, RankRLS, with time complexity $\mathcal{O}(n^3+n^2m)$, with $n$ the number of features and $m$ the number of training documents.
-The RankRLS ranking function showed ranking performance similar to RankSVM on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.
+
 \subsection{Distributed Stochastic Gradient Descent}

### 2014-03-10 report/Chapters/RelatedWork.tex
algemene inleiding related work

@@
 \label{sec:lit_descr}
+Different approaches to parallelising Learning-to-Rank can be identified in literature. The following subsections describe the related work in the field categorised by approach.
 
@@
 Boosting \cite{Friedman2002} is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.
-Boosting is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.\\
+The boosting step itself is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.\\
+
+Ganjisaffar \cite{Ganjisaffar2011b} proposed a pairwise boosting-wrapped-in-bagging model called BL-MART, in contrast to BagBoo which is a pointwise model.
+BL-MART adds a bagging step to the LambdaMART \cite{Wu2008} ranking model, a model that uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
+In contract to BagBoo, BL-MART is limited in the number of trees.
+An experiment on the TD2004 dataset resulted in 4500 trees using BL-MART while 1.1 million trees were created with the BagBoo model.
 

### 2014-03-10 report/Chapters/RelatedWork.tex
indeling ensemble learning related work klaar

@@
 The invention of the boosting method resulted in an increasing focus in machine learning research on methods to combine multiple weak models, which are called ensemble learning methods.
-Well-known ensemble methods include boosting \cite{Schapire1990}, bagging \cite{Breiman1996}, AdaBoost \cite{Freund1997} and stacked generalisation (stacking) \cite{Wolpert1992}.\\
+Well-known ensemble methods include gradient boosting \cite{Friedman2001}, bagging \cite{Breiman1996}, AdaBoost \cite{Freund1997} and stacked generalisation (stacking) \cite{Wolpert1992}.\\
 
@@
 \ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
-\subsubsection{Stochastic Boosting}
-Ye et al.
-\cite{Ye2009} described how to implement the \ac{SGBDT} process in a parallel manner using both MPI and Hadoop.
-Experiments showed the Hadoop implementation to result into too expensive communication cost to be useful.
-Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
-The MPI approach proved to be successful and obtained near linear speed-ups.\\
+
+\subsubsection{Gradient Boosting}
+Gradient Boosting \cite{Friedman2001} is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.\\
 
 Kocsis et al.
-\cite{Kocsis2013} propose way to train multiple weak learners in parallel and extend those models that are likely to yield a good model when combined through boosting.
-Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular boosting.
-Using this parallel BoostingTree algorithm \ac{GBDT} models could be trained in parallel.\\
+\cite{Kocsis2013} propose way to train multiple weak learners in parallel by extending those models that are likely to yield a good model when combined through boosting.
+Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular gradient boosting.
+\ac{GBDT} models could be trained in parallel using their BoostingTree algorithm.\\
 
@@
 Learning to Rank challenge data sets were performed.
-This parallel \ac{CART}-tree building approach showed speed-up of up to 42x on shared memory machines and up to 25x on distributed memory machines.
+This parallel \ac{CART}-tree building approach showed speed-up of up to 42x on shared memory machines and up to 25x on distributed memory machines.\\
+
+Ye et al.
+\cite{Ye2009} described how to implement the stochastic \ac{GBDT} model in a parallel manner using both MPI and Hadoop.
+Stochastic \ac{GBDT} differs from regular \ac{GBDT} models by using stochastic gradient boosting instead of regular gradient boosting.
+Stochastic gradient boosting is a minor alteration to regular gradient boosting that, at each iteration of the algorithm, trains a base learner on a sub-sample of the training set that is drawn at random without replacement \cite{Friedman2002}.
+Experiments showed the Hadoop implementation to result into too expensive communication cost to be useful.
+Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
+The MPI approach proved to be successful and obtained near linear speed-ups.
+
+\subsubsection{Boosting wrapped in Bagging}
+Some approaches combine both boosting and bagging.
+Bagging \cite{Breiman1996}, also called bootstrap aggregating, is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement.
+The bagging method is parallelisable by training each $D_i$ with $i \in \{1..m\}$ on a different node.\\
+
+Pavlov et al.
+\cite{Pavlov2010} from Yandex were the first to propose a boosting-wrapped-in-bagging approach, which they called BagBoo.
+The boosting step itself is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.
+
+Ganjisaffar \cite{Ganjisaffar2011b} proposed a pairwise boosting-wrapped-in-bagging model called BL-MART, in contrast to BagBoo which is a pointwise model.
+BL-MART adds a bagging step to the LambdaMART \cite{Wu2008} ranking model, a model that uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
+In contract to BagBoo, BL-MART is limited in the number of trees.
+An experiment on the TD2004 dataset resulted in 4500 trees using BL-MART while 1.1 million trees were created with the BagBoo model.
 
@@
 Each layer in the stack learns has the task of learning the weights of the inputs of that layer.
-The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.
-
-\subsubsection{Boosting wrapped in Bagging}
-Pavlov et al.
-\cite{Pavlov2010} at Yandex developed a hybrid model that combines two methods from ensemble learning: bagging and boosting.
-Bagging \cite{Breiman1996}, also called bootstrap aggregating, is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement.
-The bagging method is parallelisable by training each $D_i$ with $i \in \{1..m\}$ on a different node.
-Boosting \cite{Friedman2002} is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.
-The boosting step itself is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.\\
-
-Ganjisaffar \cite{Ganjisaffar2011b} proposed a pairwise boosting-wrapped-in-bagging model called BL-MART, in contrast to BagBoo which is a pointwise model.
-BL-MART adds a bagging step to the LambdaMART \cite{Wu2008} ranking model, a model that uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
-In contract to BagBoo, BL-MART is limited in the number of trees.
-An experiment on the TD2004 dataset resulted in 4500 trees using BL-MART while 1.1 million trees were created with the BagBoo model.
+The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.\\
 

### 2014-03-10 report/Chapters/RelatedWork.tex
hardware accelerator categorie related work

@@
 
+\subsection{Learning-to-Rank parallelisation by hardware acceleration}
+Hardware accelerators are special purpose processors designed to speed up compute-intensive tasks.
+An \ac{FPGA} and a \ac{GPU} are two different types of hardware that can achieve better performance on some tasks though parallel computing.
+In general, \ac{FPGA}s provide better performance while \ac{GPU}s tend to be easier to program \cite{Che2008}.
+Some research has been done in parallelising Learning-to-Rank methods using hardware accelerators.
+\subsubsection{FPGA-based parallel Learning-to-Rank}
+Yan et al.
+\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run the Neural Network based LambdaRank Learning-to-Rank algorithm.
+This architecture achieved a 29.3X speed-up compared to the software implementation when evaluated on data from a commercial search engine.
+The exploration of \ac{FPGA} for Learning-to-Rank showed other advantages of the \ac{FPGA} approach next to faster model training.
+In their latest publication \cite{Yan2012} the \ac{FPGA} based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.
+
+\subsubsection{GPGPU for Learning-to-Rank}
+De Sousa et al.
+\cite{DeSousa2012} proposed a \ac{GPGPU} approach using the \ac{GPU} both learning the ranking function and for query processing, thereby improving both training time and query time.
+An association rule based Learning-to-Rank approach proposed by \cite{Veloso2008} has been implemented using the \ac{GPU} in such a way that the set of rules van be computed in parallel, in different threads, for each document.
+A speed-up of 127X in query processing time is reported based on evaluation on the LETOR dataset.
+The speed-up achieved at learning the ranking function was unfortunately not stated.\\
+
 \subsection{Low complexity Learning-to-Rank}
@@
 Unfortunately the speed-up obtained by training their model in parallel is not reported.
-\subsection{FPGA-based LambdaRank}
-Yan et al.
-\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run the Neural Network based LambdaRank Learning-to-Rank algorithm.
-This architecture achieved a 29.3X speed-up compared to the software implementation when evaluated on data from a commercial search engine.
-The exploration of \ac{FPGA} for Learning-to-Rank showed other advantages of the \ac{FPGA} approach next to faster model training.
-In their latest publication \cite{Yan2012} the \ac{FPGA} based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.
-\subsection{GPGPU for Learning-to-Rank}
-De Sousa et al.
-\cite{DeSousa2012} proposed a \ac{GPGPU} approach using the \ac{GPU} both learning the ranking function and for query processing, thereby improving both training time and query time.
-An association rule based Learning-to-Rank approach proposed by \cite{Veloso2008} has been implemented using the \ac{GPU} in such a way that the set of rules van be computed in parallel, in different threads, for each document.
-A speed-up of 127X in query processing time is reported based on evaluation on the LETOR dataset.
-The speed-up achieved at learning the ranking function was unfortunately not stated.
 

### 2014-03-10 report/Chapters/RelatedWork.tex
duidelijkere structuur related work

@@
 
-Section \ref{sec:lit_descr} describe the approaches on parallel or distributed Learning-to-Rank found in literature.
-\section{Literature description}
-\label{sec:lit_descr}
-Different approaches to parallelising Learning-to-Rank can be identified in literature. The following subsections describe the related work in the field categorised by approach.
+Different approaches to parallelising Learning-to-Rank can be identified in related literature. The following subsections describe the related work in the field categorised by approach.
 
-\subsection{Parallelising Learning-to-Rank through ensemble learning}
+\section{Parallelising Learning-to-Rank through ensemble learning}
 Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure that he called boosting \cite{Schapire1990}.
@@
 
-\subsubsection{Gradient Boosting}
+\subsection{Gradient Boosting}
 Gradient Boosting \cite{Friedman2001} is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.\\
@@
 
-\subsubsection{Boosting wrapped in Bagging}
+\subsection{Boosting wrapped in Bagging}
 Some approaches combine both boosting and bagging.
@@
 
-\subsubsection{Stacked generalisation}
+\subsection{Stacked generalisation}
 A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.
@@
 
-\subsection{Learning-to-Rank parallelisation by hardware acceleration}
+\section{Learning-to-Rank parallelisation by hardware acceleration}
 Hardware accelerators are special purpose processors designed to speed up compute-intensive tasks.
@@
 Some research has been done in parallelising Learning-to-Rank methods using hardware accelerators.
-\subsubsection{FPGA-based parallel Learning-to-Rank}
+\subsection{FPGA-based parallel Learning-to-Rank}
 Yan et al.
@@
 
-\subsubsection{GPGPU for Learning-to-Rank}
+\subsection{GPGPU for Learning-to-Rank}
 De Sousa et al.
@@
 
-\subsection{Low complexity Learning-to-Rank}
+\section{Low complexity Learning-to-Rank}
 Designing Learning-to-Rank algorithms with low time complexity for training is another approach towards large scale Learning-to-Rank.
@@
 
-\subsection{CCRank}
+\section{CCRank}
 Wang et al.
@@
 Two reasons have been identified for not achieving linear scale-up with CCRank: 1) parallel execution is suspended after each generation to perform combination in order to produce the candidate solution, 2) Combination has to wait until all parallel tasks have finished, which may spend different running time.
-\subsection{Parallel ListNet using Spark}
+\section{Parallel ListNet using Spark}
 Shukla et al.
@@
 The Spark implementation of ListNet showed near a linear training time reduction.
-\subsection{nDCG-Annealing}
+\section{nDCG-Annealing}
 Karimzadeghan et al.
@@
 
-\subsection{Distributed Stochastic Gradient Descent}
+\section{Distributed Stochastic Gradient Descent}
 Long et al.
@@
 No experimental results on parallelisation speed-up were reported on this Pregel based approach.
-\subsection{Distributed Hyper-parameter Tuning using MapReduce}
+\section{Distributed Hyper-parameter Tuning using MapReduce}
 Ganjisaffar et al.
@@
 However, the risk of overfitting grows overwhelmingly as the number of hyperparameter combinations grow, even when validation sets grow large.
-\subsection{Distributed LambdaRank}
+\section{Distributed LambdaRank}
 Svore and Burges \cite{Svore2010,Svore2012} designed two approaches to train the LambdaMART \cite{Wu2008} model in a distributed manner.
@@
 Both approaches achieves a six times speed-up over LambdaMART on 32 nodes compared to a single node.
-\subsection{Learning-to-Rank with Bregman Divergences and Monotone Retargeting}
+\section{Learning-to-Rank with Bregman Divergences and Monotone Retargeting}
 Acharyya et al.
@@
 No experiments on speed-up have been performed.
-\subsection{Parallel robust Learning-to-Rank}
+\section{Parallel robust Learning-to-Rank}
 Robust learning \cite{Huber1981} is defined as the task to lean a model in the presence of outliers.

### 2014-03-10 report/Chapters/RelatedWork.tex
eerste opzet categorisering related work

@@
 Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
-The MPI approach proved to be successful and obtained near linear speed-ups.
+The MPI approach proved to be successful and obtained near linear speed-ups.\\
+
+Svore and Burges \cite{Svore2010,Svore2012} designed two approaches to train the LambdaMART \cite{Wu2008} model in a distributed manner.
+Their first approach is applicable in case the whole training set fits in main memory on every node, in that case the tree split computations of the regressions trees in the LambdaMART tree ensemble are split and the data is not.
+The second approach distributes the training data and corresponding training computations and is therefore scalable to high amounts of training data.
+Both approaches achieves a six times speed-up over LambdaMART on 32 nodes compared to a single node.
 
@@
 Some research has been done in parallelising Learning-to-Rank methods using hardware accelerators.
+
 \subsection{FPGA-based parallel Learning-to-Rank}
@@
 
+\section{Parallelised strong learners in Learning-to-Rank}
+Some research focused on parallelising the steps Learning-to-Rank algorithms that can be characterised as strong learners.
+\subsection{Parallel ListNet using Spark}
+Shukla et al.
+\cite{Shukla2012} explored the parallelisation of the well-known ListNet Learning-to-Rank method using Spark.
+Spark is a parallel computing model that is designed for cyclic data flows which makes it more suitable for iterative algorithms.
+Spark is incorporated into Hadoop since Hadoop 2.0.
+The Spark implementation of ListNet showed near a linear training time reduction.\\
+
 \section{Low complexity Learning-to-Rank}
@@
 \cite{Pahikkala2009} describe a pairwise \ac{RLS} type of ranking function, RankRLS, with time complexity $\mathcal{O}(n^3+n^2m)$, with $n$ the number of features and $m$ the number of training documents.
-The RankRLS ranking function showed ranking performance similar to RankSVM on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.
+The RankRLS ranking function showed ranking performance similar to RankSVM on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.\\
 
-\section{CCRank}
+\section{Parallelisable Search Heuristics for Listwise ranking}
+%TODO:
+Here comes a short intro to genetic/evolutionary algorithms, stochastic gradient descent and simulated annealing.
+\subsection{CCRank}
 Wang et al.
@@
 Two reasons have been identified for not achieving linear scale-up with CCRank: 1) parallel execution is suspended after each generation to perform combination in order to produce the candidate solution, 2) Combination has to wait until all parallel tasks have finished, which may spend different running time.
-\section{Parallel ListNet using Spark}
-Shukla et al.
-\cite{Shukla2012} explored the parallelisation of the well-known ListNet Learning-to-Rank method using Spark.
-Spark is a parallel computing model that is designed for cyclic data flows which makes it more suitable for iterative algorithms.
-Spark is incorporated into Hadoop since Hadoop 2.0.
-The Spark implementation of ListNet showed near a linear training time reduction.
-\section{nDCG-Annealing}
+\subsection{nDCG-Annealing}
 Karimzadeghan et al.
@@
 Unfortunately their work does not directly report on the speed-up obtained by parallelising  with Hadoop, but it is mentioned that further work needs to be done to effectively leverage parallel execution.
-
-\section{Distributed Stochastic Gradient Descent}
+\subsection{Distributed Stochastic Gradient Descent}
 Long et al.
@@
 has been used to show that the model is effective.
-Unfortunately the speed-up obtained by training their model in parallel is not reported.
+Unfortunately the speed-up obtained by training their model in parallel is not reported.\\
 
+\section{Paralelly optimisable surrogate loss functions}
+%TODO:
 \subsection{Alternating Direction Method of Multipliers}
@@
 No experimental results on parallelisation speed-up were reported on this Pregel based approach.
-\section{Distributed Hyper-parameter Tuning using MapReduce}
-Ganjisaffar et al.
-\cite{Ganjisaffar2011, Ganjisaffar2011b} observed that long training times are often a result of training many models for hyperparameter optimisation.
-Grid search is the \emph{de facto} standard of hyperparameter optimisation and is simply an exhaustive search through a manually specified subset of hyperparameter combinations.
-The authors show how to perform parallel grid search on MapReduce clusters, which is easy because grid search is an embarrassingly parallel method as hyperparameter combinations are mutually independent.
-They apply their grid search on MapReduce approach in a Learning-to-Rank setting to train a LambdaMART \cite{Wu2008} ranking model, a model that uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
-Experiments showed that the solution scales linearly in number of hyperparameter combinations.
-However, the risk of overfitting grows overwhelmingly as the number of hyperparameter combinations grow, even when validation sets grow large.
-\section{Distributed LambdaRank}
-Svore and Burges \cite{Svore2010,Svore2012} designed two approaches to train the LambdaMART \cite{Wu2008} model in a distributed manner.
-Their first approach is applicable in case the whole training set fits in main memory on every node, in that case the tree split computations of the regressions trees in the LambdaMART tree ensemble are split and the data is not.
-The second approach distributes the training data and corresponding training computations and is therefore scalable to high amounts of training data.
-Both approaches achieves a six times speed-up over LambdaMART on 32 nodes compared to a single node.
-\section{Learning-to-Rank with Bregman Divergences and Monotone Retargeting}
+\subsection{Learning-to-Rank with Bregman Divergences and Monotone Retargeting}
 Acharyya et al.
@@
 focussed their research more on robust ranking than on parallel ranking.
-The only reference to performance of RoBiRank in terms of speed is the statement that its training time on a computing cluster is comparable to the more efficient implementation by Lee and Lin \cite{Lee2014} of a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.
+The only reference to performance of RoBiRank in terms of speed is the statement that its training time on a computing cluster is comparable to the more efficient implementation by Lee and Lin \cite{Lee2014} of a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.\\
+
+\section{Distributed Hyper-parameter Tuning using MapReduce}
+Ganjisaffar et al.
+\cite{Ganjisaffar2011, Ganjisaffar2011b} observed that long training times are often a result of training many models for hyperparameter optimisation.
+Grid search is the \emph{de facto} standard of hyperparameter optimisation and is simply an exhaustive search through a manually specified subset of hyperparameter combinations.
+The authors show how to perform parallel grid search on MapReduce clusters, which is easy because grid search is an embarrassingly parallel method as hyperparameter combinations are mutually independent.
+They apply their grid search on MapReduce approach in a Learning-to-Rank setting to train a LambdaMART \cite{Wu2008} ranking model, a model that uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
+Experiments showed that the solution scales linearly in number of hyperparameter combinations.
+However, the risk of overfitting grows overwhelmingly as the number of hyperparameter combinations grow, even when validation sets grow large.

### 2014-03-10 report/Chapters/RelatedWork.tex
visuele categorisatie large scale L2R training

@@
 
-Different approaches to parallelising Learning-to-Rank can be identified in related literature. The following subsections describe the related work in the field categorised by approach.
+Research in large scale Learning-to-Rank model training can be categorised according to the approach that they take in scaling up the training step of Learning-to-Rank models.
+Figure \ref{fig:related_work_categories} illustrates the categories of scalable training approaches in Learning-to-Rank.
+The following sections describe the related work in the field following the categorisation depicted in Figure \ref{fig:related_work_categories}.
+
+\begin{figure}
+\includegraphics{gfx/related_work_categories}
+\caption{Categories in research on large scale training of  Learning-to-Rank models}
+\label{fig:related_work_categories}
+\end{figure}
 
@@
 \section{Parallelisable Search Heuristics for Listwise ranking}
+The hard part in 
 %TODO:

### 2014-03-10 report/Chapters/RelatedWork.tex
verbetering categorisatie related work

@@
 
-\section{Parallelised strong learners in Learning-to-Rank}
+\section{Parallelised execution of algorithm steps in Learning-to-Rank methods}
 Some research focused on parallelising the steps Learning-to-Rank algorithms that can be characterised as strong learners.
@@
 \section{Parallelisable Search Heuristics for Listwise ranking}
-The hard part in 
-%TODO:
-Here comes a short intro to genetic/evolutionary algorithms, stochastic gradient descent and simulated annealing.
+Direct minimisation of ranking metrics is a hard problem due to the non-continuous, non-differentiable and non-convex nature of \ac{nDCG}, \ac{ERR} and \ac{MAP}.
+This optimisation problem is generally addressed either by replacing the ranking metric with a convex surrogate, or by heuristic optimisation methods.
+One heuristic optimisation method that is successfully used in direct optimisation of ranking evaluation functions is the \ac{GA} \cite{Yeh2007}.
+\ac{GA}s are search heuristic functions that mimic the process of natural selection, consisting of mutation and cross-over steps \cite{Holland1995}.
+The following subsection describe related work that uses search heuristics for parallel/distributed training.
 \subsection{CCRank}
@@
 The approach showed to be effective on both the LETOR 3.0 dataset and their own dataset with contextual advertising data.
-Unfortunately their work does not directly report on the speed-up obtained by parallelising  with Hadoop, but it is mentioned that further work needs to be done to effectively leverage parallel execution.
-\subsection{Distributed Stochastic Gradient Descent}
-Long et al.
-\cite{Long2012} describe special case of their pairwise cross-domain factor Learning-to-Rank method using distributed optimization of \ac{SGD} based on Hadoop MapReduce.
-Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce based method described by  Zinkevich et al.
-\cite{Zinkevich2010} has been used.
-Real world data from Yahoo!
-has been used to show that the model is effective.
-Unfortunately the speed-up obtained by training their model in parallel is not reported.\\
+Unfortunately their work does not directly report on the speed-up obtained by parallelising  with Hadoop, but it is mentioned that further work needs to be done to effectively leverage parallel execution.\\
 
@@
 focussed their research more on robust ranking than on parallel ranking.
-The only reference to performance of RoBiRank in terms of speed is the statement that its training time on a computing cluster is comparable to the more efficient implementation by Lee and Lin \cite{Lee2014} of a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.\\
+The only reference to performance of RoBiRank in terms of speed is the statement that its training time on a computing cluster is comparable to the more efficient implementation by Lee and Lin \cite{Lee2014} of a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.
+\subsection{Distributed Stochastic Gradient Descent}
+Long et al.
+\cite{Long2012} describe special case of their pairwise cross-domain factor Learning-to-Rank method using distributed optimization of \ac{SGD} based on Hadoop MapReduce.
+Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce based method described by  Zinkevich et al.
+\cite{Zinkevich2010} has been used.
+Real world data from Yahoo!
+has been used to show that the model is effective.
+Unfortunately the speed-up obtained by training their model in parallel is not reported.\\
 

### 2014-03-11 report/Chapters/RelatedWork.tex
deng et al

@@
 On Google Scholar the defined search query resulted in 3300 documents. We evaluate the first 300 search results.\\
+
+A one-level forward and backward reference search is used to find relevant papers missed so far.
+To handle the large volume of studies involved in the backward and forward reference search, relevance of the studies will be evaluated solely on the title of the study.\\
 
@@
 \subsection{Stacked generalisation}
+% TODO: Kijken naar deze subsection, is erg zwak %
 A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.
@@
 Each layer in the stack learns has the task of learning the weights of the inputs of that layer.
-The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.\\
+The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.
+Deng et al.
+\cite{Deng2013} % TODO: beschrijven Deng et al.%\\
 
@@
 \cite{DeSousa2012} proposed a \ac{GPGPU} approach using the \ac{GPU} both learning the ranking function and for query processing, thereby improving both training time and query time.
-An association rule based Learning-to-Rank approach proposed by \cite{Veloso2008} has been implemented using the \ac{GPU} in such a way that the set of rules van be computed in parallel, in different threads, for each document.
+An association rule based Learning-to-Rank approach proposed by Veloso et al.
+\cite{Veloso2008} has been implemented using the \ac{GPU} in such a way that the set of rules van be computed in parallel, in different threads, for each document.
 A speed-up of 127X in query processing time is reported based on evaluation on the LETOR dataset.

### 2014-03-11 report/Chapters/RelatedWork.tex
volgende stap backward search

@@
 A one-level forward and backward reference search is used to find relevant papers missed so far.
-To handle the large volume of studies involved in the backward and forward reference search, relevance of the studies will be evaluated solely on the title of the study.\\
+To handle the large volume of studies involved in the backward and forward reference search, relevance of the studies will be evaluated solely on the title of the study.
+Backward reference search resulted in 10 studies regarded as potentially relevant based on the title of which x where actually relevant.\\
 

### 2014-03-11 report/Chapters/RelatedWork.tex
eerste nuttige backward search resultaat verwerkt

@@
 
-\section{Distributed Hyper-parameter Tuning using MapReduce}
+\section{Distributed Hyperparameter Tuning using MapReduce}
 Ganjisaffar et al.
@@
 Experiments showed that the solution scales linearly in number of hyperparameter combinations.
-However, the risk of overfitting grows overwhelmingly as the number of hyperparameter combinations grow, even when validation sets grow large.
+However, the risk of overfitting grows overwhelmingly as the number of hyperparameter combinations grow, even when validation sets grow large.\\
+
+Burges et al.
+\cite{Burges2011} describe their Yahoo!
+Learning to Rank Challenge submission that was build by performing an extensive hyperparameter search on a 122-node MPI cluster, running Microsoft HPC Server 2008.
+The hyperparameter optimisation was performed on a linear combination ensemble of eight LambdaMART models, two LambdaRank models and two MART models using a logistic regression cost.
+This submission achieved the highest \ac{ERR} score of all Yahoo!
+Learning to Rank Challenge submissions.

### 2014-03-11 report/Chapters/RelatedWork.tex
backward search referentie twee

@@
 The exploration of \ac{FPGA} for Learning-to-Rank showed other advantages of the \ac{FPGA} approach next to faster model training.
-In their latest publication \cite{Yan2012} the \ac{FPGA} based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.
+In their latest publication \cite{Yan2012} the \ac{FPGA} based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.\\
+
+Xu et al.
+\cite{Xu2007b,Xu2009} designed an \ac{FPGA}-based accelerator to reduce the training time of the RankBoost algorithm \cite{Freund2003}, a pairwise ranking function based on Freund and Schapire's AdaBoost ensemble learning method \cite{Freund1997}.
+Xu et al.
+\cite{Xu2009} state that RankBoost is a Learning-to-Rank method that is not widely used in practice because of its long training time.
+Experiments on MSN search engine data showed the implementation on a \ac{FPGA} with \ac{SIMD} architecture to be 170.6x faster than the original software implementation \cite{Xu2007b}.
+In a second experiment in which a much more powerful \ac{FPGA} accelerator board was used, the speed-up even increased to 1800x compared to the original software implementation \cite{Xu2009}.\\
 

### 2014-03-11 report/Chapters/RelatedWork.tex
derde relevante backward search resultaat verwerkt

@@
 \subsection{GPGPU for Learning-to-Rank}
+Wang et al.
+\cite{Wang2009} experimented with a \ac{GPGPU} approach for parallelising RankBoost.
+Nvida \ac{CUDA} and ATI Stream are the two main \ac{GPGPU} computing platform and are released by the two main \ac{GPU} vendors Nvidia and AMD.
+Experiments show a 22.9x speed-up on Nvida \ac{CUDA} and a 9.2x speed-up on ATI Stream.\\
+
 De Sousa et al.
-\cite{DeSousa2012} proposed a \ac{GPGPU} approach using the \ac{GPU} both learning the ranking function and for query processing, thereby improving both training time and query time.
+\cite{DeSousa2012} proposed a \ac{GPGPU} approach to improve both training time and query evaluation through \ac{GPU} use.
 An association rule based Learning-to-Rank approach proposed by Veloso et al.

### 2014-03-11 report/Chapters/RelatedWork.tex
vierde relevante backward search resultaat verwerkt

@@
 Direct minimisation of ranking metrics is a hard problem due to the non-continuous, non-differentiable and non-convex nature of \ac{nDCG}, \ac{ERR} and \ac{MAP}.
-This optimisation problem is generally addressed either by replacing the ranking metric with a convex surrogate, or by heuristic optimisation methods.
-One heuristic optimisation method that is successfully used in direct optimisation of ranking evaluation functions is the \ac{GA} \cite{Yeh2007}.
+This optimisation problem is generally addressed either by replacing the ranking metric with a convex surrogate, or by heuristic optimisation methods such as Simulated Annealing or a \ac{EA}.
+One \ac{EA} heuristic optimisation method that is successfully used in direct rank evaluation functions optimisation is the \ac{GA} \cite{Yeh2007}.
 \ac{GA}s are search heuristic functions that mimic the process of natural selection, consisting of mutation and cross-over steps \cite{Holland1995}.
 The following subsection describe related work that uses search heuristics for parallel/distributed training.
+\subsection{Immune Programming}
+Wang et al.
+\cite{Wang2009b} proposed a \ac{IP} solution to direct ranking metric optimisation and observed that all \ac{EA}s are generally easy to implement such that it can be carried out on different computers distributively.
+However, no statements on the possible speed-up of a distributed implementation of the \ac{IP} solution has been made and no speed-up experiments have been conducted.
 \subsection{CCRank}

### 2014-03-11 report/Chapters/RelatedWork.tex
backward search voltooid

@@
 To handle the large volume of studies involved in the backward and forward reference search, relevance of the studies will be evaluated solely on the title of the study.
-Backward reference search resulted in 10 studies regarded as potentially relevant based on the title of which x where actually relevant.\\
+Backward reference search resulted in 10 studies regarded as potentially relevant based on the title of which four were actually relevant and included in the literature description.\\
 

### 2014-03-12 report/Chapters/RelatedWork.tex
eerste relevante paper forward search verwerkt

--- 2014-03-11 report/Chapters/RelatedWork.tex
@@
 To handle the large volume of studies involved in the backward and forward reference search, relevance of the studies will be evaluated solely on the title of the study.
-Backward reference search resulted in 10 studies regarded as potentially relevant based on the title of which x where actually relevant.\\
+Backward reference search resulted in 10 studies regarded as potentially relevant based on the title of which four were actually relevant and included in the literature description.
+Forward reference search resulted in 10 potentially relevant titles, of which x studies turned out to be relevant.\\
 
@@
 \section{Low complexity Learning-to-Rank}
+\label{sec:related_work_low_complexity}
 Designing Learning-to-Rank algorithms with low time complexity for training is another approach towards large scale Learning-to-Rank.
@@
 \cite{Pahikkala2009} describe a pairwise \ac{RLS} type of ranking function, RankRLS, with time complexity $\mathcal{O}(n^3+n^2m)$, with $n$ the number of features and $m$ the number of training documents.
-The RankRLS ranking function showed ranking performance similar to RankSVM on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.\\
+The RankRLS ranking function showed ranking performance similar to RankSVM \cite{Herbrich1999, Joachims2002} on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.\\
+
+Lee and Lin \cite{Lee2014} describe a lower time complexity method to train a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.
+The authors observed that linear kernel RankSVMs are inferior in accuracy compared to nonlinear kernel RankSVMs and \ac{GBDT}s and are mainly useful to quickly produce a baseline model.
+Details of the lower time complexity version of the linear kernel RankSVM will not be discussed as it is shown to be an inferior Learning-to-Rank method in terms of accuracy.
 
@@
 Duh et al.
-\cite{Duh2011} show how to use \ac{ADMM} to train a RankSVM model in parallel.
+\cite{Duh2011} show how to use \ac{ADMM} to train a RankSVM \cite{Herbrich1999, Joachims2002} model in parallel.
 Experiments showed the \ac{ADMM} implementation to achieve a 13.1x training time speed-up on 72 workers, compared to training on a single worker.\\
@@
 focussed their research more on robust ranking than on parallel ranking.
-The only reference to performance of RoBiRank in terms of speed is the statement that its training time on a computing cluster is comparable to the more efficient implementation by Lee and Lin \cite{Lee2014} of a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.
+The only reference to performance of RoBiRank in terms of speed is the statement that its training time on a computing cluster is comparable to the more efficient implementation by Lee and Lin \cite{Lee2014} of a linear kernel RankSVM \cite{Herbrich1999, Joachims2002} described in section \ref{sec:related_work_low_complexity}.
 \subsection{Distributed Stochastic Gradient Descent}

### 2014-03-12 report/Chapters/RelatedWork.tex
tweede relevante forward search resultaat beschreven

@@
 A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.
-\ac{DSN} is based on the stacked generalization ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack.
+\ac{DSN} is based on the stacked generalization (stacking) ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack.
 Each layer in the stack learns has the task of learning the weights of the inputs of that layer.
@@
 \cite{Deng2013} % TODO: beschrijven Deng et al.%\\
+
+\subsection{Randomisation}
+Most ensemble learning based Learning-to-Rank methods are based on boosting.
+Although boosting have shown to be an effective method in Learning-to-Rank, it has the drawback that the models in the ensemble are not completely mutually independent, which makes the process harder to parallelise.
+Geurts and Louppe \cite{Geurts2011} propose a tree based ensemble method in which the trees are build using random subsets of the available features.
+The authors called this method \ac{ET} and originally proposed this method for classification and regression settings instead of ranking settings \cite{Geurts2006}.
+The \ac{ET} algorithm is similar to the well-known Random Forest algorithm, with two differences: 1) in \ac{ET} each tree is build on the complete dataset instead of random subsets of the data and 2) \ac{ET} sets the discretisation thresholds that define the splits in the tree randomly, instead of based on the sample.
+The randomisations in the process make the models in the ensemble mutually independent, which makes it trivial to parallelise by training each tree on a different node.
+In contrary to what one might think, \ac{ET}s show very good performance on benchmark data with a 10th place in the Yahoo!
+Learning to Rank Challenge.\\
 

### 2014-03-12 report/Chapters/RelatedWork.tex
naar aanleiding van forward search, een nieuwe referentie voor BL-MART

@@
 
-Ganjisaffar \cite{Ganjisaffar2011b} proposed a pairwise boosting-wrapped-in-bagging model called BL-MART, in contrast to BagBoo which is a pointwise model.
+Ganjisaffar et al.
+\cite{Ganjisaffar2011c, Ganjisaffar2011b} proposed a pairwise boosting-wrapped-in-bagging model called BL-MART, in contrast to BagBoo which is a pointwise model.
 BL-MART adds a bagging step to the LambdaMART \cite{Wu2008} ranking model, a model that uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
@@
 The authors observed that linear kernel RankSVMs are inferior in accuracy compared to nonlinear kernel RankSVMs and \ac{GBDT}s and are mainly useful to quickly produce a baseline model.
-Details of the lower time complexity version of the linear kernel RankSVM will not be discussed as it is shown to be an inferior Learning-to-Rank method in terms of accuracy.
+Details of the lower time complexity version of the linear kernel RankSVM will not be discussed as it is shown to be an inferior Learning-to-Rank method in terms of accuracy.\\
 

### 2014-03-12 report/Chapters/RelatedWork.tex
5e forward search resultaat verwerkt

@@
 
-Lee and Lin \cite{Lee2014} describe a lower time complexity method to train a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.
-The authors observed that linear kernel RankSVMs are inferior in accuracy compared to nonlinear kernel RankSVMs and \ac{GBDT}s and are mainly useful to quickly produce a baseline model.
+Airola et al.
+\cite{Airola2011} and Lee and Lin \cite{Lee2014} have both described lower time complexity methods to train a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.
+Lee and Lin \cite{Lee2014} observed that linear kernel RankSVMs are inferior in accuracy compared to nonlinear kernel RankSVMs and \ac{GBDT}s and are mainly useful to quickly produce a baseline model.
 Details of the lower time complexity version of the linear kernel RankSVM will not be discussed as it is shown to be an inferior Learning-to-Rank method in terms of accuracy.\\

### 2014-03-12 report/Chapters/RelatedWork.tex
zesde forward search resultaat verwerkt

@@
 Pahikkala et al.
-\cite{Pahikkala2009} describe a pairwise \ac{RLS} type of ranking function, RankRLS, with time complexity $\mathcal{O}(n^3+n^2m)$, with $n$ the number of features and $m$ the number of training documents.
+\cite{Pahikkala2009} describe a pairwise \ac{RLS} type of ranking function, RankRLS, with low time complexity.
+Airola et al.
+\cite{Airola2010} further improved the training time complexity of RankRLS to $\mathcal{O}(tms)$, where $t$ is the number of needed iterations, $m$ the number of training documents and $s$ the number of features.
 The RankRLS ranking function showed ranking performance similar to RankSVM \cite{Herbrich1999, Joachims2002} on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.\\

### 2014-03-12 report/Chapters/RelatedWork.tex
zevende relevante forward search paper verwerkt

@@
 Acharyya et al.
-\cite{Acharyya2012} propose a Learning-to-Rank method searches for an order preserving transformation (\emph{monotone retargeting}) of the target scores that is easier for a regressor to fit.
+\cite{Acharyya2012, Acharyya2013} propose a Learning-to-Rank method searches for an order preserving transformation (\emph{monotone retargeting}) of the target scores that is easier for a regressor to fit.
 This approach is based on the observation that it is not necessary to fit scores exactly, since the evaluation is dependent on the order and not on the pointwise predictions themselves.\\
@@
 Acharyya et al.
-\cite{Acharyya2012} defined a parallel algorithm that optimises a Bregman divergence function as surrogate of \ac{nDCG} that is claimed to be well suited for implementation of a \ac{GPGPU}.
+\cite{Acharyya2012, Acharyya2013} defined a parallel algorithm that optimises a Bregman divergence function as surrogate of \ac{nDCG} that is claimed to be well suited for implementation of a \ac{GPGPU}.
 No experiments on speed-up have been performed.

### 2014-03-13 report/Chapters/Background.tex
verwerking self-check

@@
 %************************************************
-This chapter provides a background in the Learning-to-Rank field with the goal of making the subsequential chapters of this thesis understandable for non-experts in the field.
+This chapter is written with the goal to provide the reader with a subtle introduction in the Learning-to-Rank field.
+Models and theories explained in this chapter can be regarded as prior knowledge needed to understand the subsequent chapters of this thesis.
 \section{A basic introduction to Learning-to-Rank}
@@
 
-\section{Evaluating a ranking}
+\section{How to evaluate a ranking}
 Evaluation metrics have long been studied in the field of information retrieval.
@@
 
-Well-known algorithms that belong to the listwise approach include AdaRank \cite{Xu2007}, BoltzRank \cite{Volkovs2009}, ListMLE \cite{Xia2008}, ListNet \cite{Cao2007}, RankCosine \cite{Qin2008}, SmoothRank \cite{Chapelle2010}, SoftRank \cite{Taylor2008}, SVM$^map$ \cite{Yue2007}.
+Well-known algorithms that belong to the listwise approach include AdaRank \cite{Xu2007}, BoltzRank \cite{Volkovs2009}, ListMLE \cite{Xia2008}, ListNet \cite{Cao2007}, RankCosine \cite{Qin2008}, SmoothRank \cite{Chapelle2010}, SoftRank \cite{Taylor2008}, SVM$^{map}$ \cite{Yue2007}.

### 2014-03-13 report/Chapters/Introduction.tex
verwerking self-check

@@
 \section{Approach}
-A literature study will be performed to get insight in relevant existing techniques for large scale Learning-to-Rank. The literature study will be performed by using the following query:
-\begin{itemize}
-\item ("learning to rank" OR "learning-to-rank" OR "machine learned ranking") AND ("parallel" OR "distributed")
-\end{itemize}
+A literature study will be performed to get insight in relevant existing techniques for large scale Learning-to-Rank. The literature study will be performed by using the following query:\\
+
+\emph{("learning to rank" \emph{OR} "learning-to-rank" \emph{OR} "machine learned ranking") \emph{AND} ("parallel" \emph{OR} "distributed")}\\
+
 and the following bibliographic databases:
@@
 \item Web of Science
+\item Google Scholar
 \end{itemize}
@@
 Even though \emph{parallel} and \emph{distributed} are not always synonymous is all definitions, we are interested in both approaches in non-sequential data processing.\\
+
+A one-level forward and backward reference search is used to find relevant papers missed so far.
+To handle the large volume of studies involved in the backward and forward reference search, relevance of the studies will be evaluated solely on the title of the study.\\
 
@@
 \item[Chapter II: Background]{introduces the reader to the basic principles and recent work in the fields of Learning-to-Rank.}
-\item[Chapter III: Related Work]{concisely describes existing work in the field of parallel machine learning and parallel Learning-to-Rank.}
+\item[Chapter III: Related Work]{concisely describes existing work in the field of parallel Learning-to-Rank.}
 \item[Chapter IV: Benchmark Results]{sketches the accuracy of existing Learning-to-Rank methods on several benchmark datasets and describes the selection of Learning-to-Rank methods for the parallelisation experiments.}

### 2014-03-13 report/Chapters/RelatedWork.tex
forward reference search verwerkt
verwerking self-check

@@
 \section{Literature study characteristics}
-The literature research is performed by using the bibliographic databases Scopus and Web of Science with the following search query: \emph{("learning to rank" OR "learning-to-rank" OR "machine learned ranking") AND ("parallel" OR "distributed")}.
-An abstract based manual filtering step is applied where I filter those results that actually use the \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}.
-Studies focusing on efficient query evaluation instead of efficient model training are likely to meet all criteria listed.
-As a last step I will filter out studies based on the whole document that only focus on efficient query evaluation and not on parallel or distributed learning of ranking functions.\\
+The literature research is performed by using the bibliographic databases Scopus and Web of Science with the following search query:\\
 
-on Scopus the defined search query resulted in 65 documents.
+\emph{("learning to rank" \emph{OR} "learning-to-rank" \emph{OR} "machine learned ranking") \emph{AND} ("parallel" \emph{OR} "distributed")}\\
+
+An abstract-based manual filtering step is applied where those results are filtered that use the terms \emph{parallel} or \emph{distributed} in context to \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}.
+As a last step I will filter out studies based on the whole document that only focus on efficient query evaluation and not on parallel or distributed learning of ranking functions, as those studies are likely to meet listed search terms.\\
+
+On Scopus, the defined search query resulted in 65 documents.
 Only 14 of those documents used \emph{large scale}, \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}.
@@
 The defined search query resulted in 16 documents on Web of Science.
-Four of those documents were also present in the set of 65 documents found using Scopus, leaving 61 unique documents.
-Only four of those 61 documents used \emph{large scale}, \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}, none of them focussed on parallel or distributed learning of ranking functions.\\
+Four of those documents were part of the 10 relevant documents found using Scopus, leaving 12 new potentially relevant documents to consider.
+Four of those 12 documents used \emph{large scale}, \emph{parallel} or \emph{distributed} terms in context to the \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}, none of them focused on parallel or distributed learning of ranking functions.\\
 
-On Google Scholar the defined search query resulted in 3300 documents. We evaluate the first 300 search results.\\
+On Google Scholar, the defined search query resulted in 3300 documents. Because it infeasible to evaluate all 3300 studies we focus on the first 300 search results.\\
 
-A one-level forward and backward reference search is used to find relevant papers missed so far.
-To handle the large volume of studies involved in the backward and forward reference search, relevance of the studies will be evaluated solely on the title of the study.
 Backward reference search resulted in 10 studies regarded as potentially relevant based on the title of which four were actually relevant and included in the literature description.
-Forward reference search resulted in 10 potentially relevant titles, of which x studies turned out to be relevant.\\
+Forward reference search resulted in 10 potentially relevant titles, of which seven studies turned out to be relevant.\\
 
-Research in large scale Learning-to-Rank model training can be categorised according to the approach that they take in scaling up the training step of Learning-to-Rank models.
+Research in scaling up the training phase of Learning-to-Rank models can be categorised according to the approach in scaling up.
 Figure \ref{fig:related_work_categories} illustrates the categories of scalable training approaches in Learning-to-Rank.
-The following sections describe the related work in the field following the categorisation depicted in Figure \ref{fig:related_work_categories}.
+The following sections describe the related work in the field following this categorisation.
 
@@
 
-\section{Parallelising Learning-to-Rank through ensemble learning}
+\section{Ensemble learning for parallel Learning-to-Rank}
 Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure that he called boosting \cite{Schapire1990}.
 The invention of the boosting method resulted in an increasing focus in machine learning research on methods to combine multiple weak models, which are called ensemble learning methods.
-Well-known ensemble methods include gradient boosting \cite{Friedman2001}, bagging \cite{Breiman1996}, AdaBoost \cite{Freund1997} and stacked generalisation (stacking) \cite{Wolpert1992}.\\
+Well-known ensemble methods include gradient boosting \cite{Friedman2001}, bagging \cite{Breiman1996}, AdaBoost \cite{Freund1997} and stacked generalisation (often called stacking) \cite{Wolpert1992}.\\
 
@@
 In the Learning-to-Rank field parallelisation of model training through ensemble learning is frequently achieved using decision tree learners and often using the boosting ensemble method, jointly called \acf{GBDT}.
-\ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
+\ac{GBDT}s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
 
@@
 Kocsis et al.
-\cite{Kocsis2013} propose way to train multiple weak learners in parallel by extending those models that are likely to yield a good model when combined through boosting.
+\cite{Kocsis2013} propose a way to train multiple weak learners in parallel by extending those models that are likely to yield a good model when combined through boosting.
 Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular gradient boosting.
 \ac{GBDT} models could be trained in parallel using their BoostingTree algorithm.\\
-
-Tyree et al.
-\cite{Tyree2011} describe a way of parallelising \ac{GBDT} models for Learning-to-Rank where the boosting step is still executed sequentially, instead the construction of the regression trees themselves are parallelised.
-The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}.
-Decision trees are build layer-by-layer where the calculations needed for building each new layer in the tree are divided over the nodes using a master-worker paradigm.
-The data is partitioned of the workers, who compress their share into histograms and send these to the master.
-The master uses those histograms to approximate the split and build the next layer.
-The master than communicates this new layer to the workers who can use this new layer to compute new histograms.
-this process is repeated until the tree depth limit is reached.
-The tree construction algorithm parallelised with described master-worker method is \ac{CART} \cite{Breiman1984}.
-Speed-up experiments on the LETOR and the Yahoo!
-Learning to Rank challenge data sets were performed.
-This parallel \ac{CART}-tree building approach showed speed-up of up to 42x on shared memory machines and up to 25x on distributed memory machines.\\
 
@@
 
-\section{Learning-to-Rank parallelisation by hardware acceleration}
+\section{Hardware accelerated Learning-to-Rank}
 Hardware accelerators are special purpose processors designed to speed up compute-intensive tasks.
@@
 
-\section{Parallelised execution of algorithm steps in Learning-to-Rank methods}
+\section{Parallelisation through execution of algorithm steps in Learning-to-Rank methods}
 Some research focused on parallelising the steps Learning-to-Rank algorithms that can be characterised as strong learners.
+
+Tyree et al.
+\cite{Tyree2011} describe a way of parallelising \ac{GBDT} models for Learning-to-Rank where the boosting step is still executed sequentially and instead the construction of the regression trees themselves is parallelised.
+The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}.
+Decision trees are build layer-by-layer where the calculations needed for building each new layer in the tree are divided over the nodes using a master-worker paradigm.
+The data is partitioned of the workers, who compress their share into histograms and send these to the master.
+The master uses those histograms to approximate the split and build the next layer.
+The master than communicates this new layer to the workers who can use this new layer to compute new histograms.
+this process is repeated until the tree depth limit is reached.
+The tree construction algorithm parallelised with described master-worker method is \ac{CART} \cite{Breiman1984}.
+Speed-up experiments on the LETOR and the Yahoo!
+Learning to Rank challenge data sets were performed.
+This parallel \ac{CART}-tree building approach showed speed-up of up to 42x on shared memory machines and up to 25x on distributed memory machines.\\
+
 \subsection{Parallel ListNet using Spark}
@@
 
-\section{Parallelisable Search Heuristics for Listwise ranking}
+\section{Parallelisable search heuristics for Listwise ranking}
 Direct minimisation of ranking metrics is a hard problem due to the non-continuous, non-differentiable and non-convex nature of \ac{nDCG}, \ac{ERR} and \ac{MAP}.
@@
 No experimental results on parallelisation speed-up were reported on this Pregel based approach.
-\subsection{Learning-to-Rank with Bregman Divergences and Monotone Retargeting}
+\subsection{Bregman Divergences and Monotone Retargeting}
 Acharyya et al.
@@
 
-\section{Distributed Hyperparameter Tuning using MapReduce}
+\section{Distributed hyperparameter tuning using MapReduce}
 Ganjisaffar et al.

### 2014-03-13 report/Chapters/RelatedWork.tex
related work feedback jurjen

@@
 
-Backward reference search resulted in 10 studies regarded as potentially relevant based on the title of which four were actually relevant and included in the literature description.
+Backward reference search resulted in 10 studies regarded as potentially relevant based on the title, of which four were actually relevant and included in the literature description.
 Forward reference search resulted in 10 potentially relevant titles, of which seven studies turned out to be relevant.\\
@@
 Figure \ref{fig:related_work_categories} illustrates the categories of scalable training approaches in Learning-to-Rank.
-The following sections describe the related work in the field following this categorisation.
+The following sections describe the related work in the field based on this categorisation.
 
@@
 \section{Ensemble learning for parallel Learning-to-Rank}
-Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure that he called boosting \cite{Schapire1990}.
+Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure called boosting \cite{Schapire1990}.
 The invention of the boosting method resulted in an increasing focus in machine learning research on methods to combine multiple weak models, which are called ensemble learning methods.
@@
 Ensemble methods can be used to parallelise learning methods by training the weak models in the ensemble on different nodes.
-In the Learning-to-Rank field parallelisation of model training through ensemble learning is frequently achieved using decision tree learners and often using the boosting ensemble method, jointly called \acf{GBDT}.
+Parallelisation of the training phase of Learning-to-Rank models is often achieved through ensemble learning, mainly boosting, and combined with decision trees, jointly called \acf{GBDT}.
 \ac{GBDT}s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
@@
 Kocsis et al.
-\cite{Kocsis2013} propose a way to train multiple weak learners in parallel by extending those models that are likely to yield a good model when combined through boosting.
+\cite{Kocsis2013} proposed a way to train multiple weak learners in parallel by extending those models that are likely to yield a good model when combined through boosting.
 Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular gradient boosting.
@@
 Stochastic gradient boosting is a minor alteration to regular gradient boosting that, at each iteration of the algorithm, trains a base learner on a sub-sample of the training set that is drawn at random without replacement \cite{Friedman2002}.
-Experiments showed the Hadoop implementation to result into too expensive communication cost to be useful.
+Experiments showed the Hadoop implementation to resulted in too expensive communication cost to be useful.
 Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
@@
 Svore and Burges \cite{Svore2010,Svore2012} designed two approaches to train the LambdaMART \cite{Wu2008} model in a distributed manner.
-Their first approach is applicable in case the whole training set fits in main memory on every node, in that case the tree split computations of the regressions trees in the LambdaMART tree ensemble are split and the data is not.
+Their first approach is applicable in case the whole training set fits in main memory on every node, in that case the tree split computations of the trees in the LambdaMART tree-ensemble are split instead of the data.
 The second approach distributes the training data and corresponding training computations and is therefore scalable to high amounts of training data.
@@
 
-Pavlov et al.
-\cite{Pavlov2010} from Yandex were the first to propose a boosting-wrapped-in-bagging approach, which they called BagBoo.
-The boosting step itself is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.
+Yandex researchers Pavlov et al.
+\cite{Pavlov2010} were the first to proposed a boosting-wrapped-in-bagging approach, which they called BagBoo.
+The boosting step itself is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.\\
 
 Ganjisaffar et al.
-\cite{Ganjisaffar2011c, Ganjisaffar2011b} proposed a pairwise boosting-wrapped-in-bagging model called BL-MART, in contrast to BagBoo which is a pointwise model.
-BL-MART adds a bagging step to the LambdaMART \cite{Wu2008} ranking model, a model that uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
+\cite{Ganjisaffar2011c, Ganjisaffar2011b} proposed a pairwise boosting-wrapped-in-bagging model called BL-MART, contrasting the pointwise BagBoo model.
+BL-MART adds a bagging step to the LambdaMART \cite{Wu2008} ranking model, that uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
 In contract to BagBoo, BL-MART is limited in the number of trees.
@@
 Although boosting have shown to be an effective method in Learning-to-Rank, it has the drawback that the models in the ensemble are not completely mutually independent, which makes the process harder to parallelise.
-Geurts and Louppe \cite{Geurts2011} propose a tree based ensemble method in which the trees are build using random subsets of the available features.
+Geurts and Louppe \cite{Geurts2011} proposed a tree based ensemble method in which the trees are built using random subsets of the available features.
 The authors called this method \ac{ET} and originally proposed this method for classification and regression settings instead of ranking settings \cite{Geurts2006}.
-The \ac{ET} algorithm is similar to the well-known Random Forest algorithm, with two differences: 1) in \ac{ET} each tree is build on the complete dataset instead of random subsets of the data and 2) \ac{ET} sets the discretisation thresholds that define the splits in the tree randomly, instead of based on the sample.
+The \ac{ET} algorithm is similar to the well-known Random Forest algorithm, but with two differences: 1) in \ac{ET} each tree is build on the complete dataset instead of random subsets of the data and 2) \ac{ET} sets the discretisation thresholds that define the splits in the tree randomly, instead of based on the sample.
 The randomisations in the process make the models in the ensemble mutually independent, which makes it trivial to parallelise by training each tree on a different node.
@@
 Yan et al.
-\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run the Neural Network based LambdaRank Learning-to-Rank algorithm.
-This architecture achieved a 29.3X speed-up compared to the software implementation when evaluated on data from a commercial search engine.
-The exploration of \ac{FPGA} for Learning-to-Rank showed other advantages of the \ac{FPGA} approach next to faster model training.
+\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run, the Neural Network based LambdaRank Learning-to-Rank algorithm.
+This architecture achieved a 29.3X speed-up compared to the software implementation, when evaluated on data from a commercial search engine.
+The exploration of \ac{FPGA} for Learning-to-Rank showed additional benefits other than the speed-up originally aimed for.
 In their latest publication \cite{Yan2012} the \ac{FPGA} based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.\\
@@
 \cite{Wang2009} experimented with a \ac{GPGPU} approach for parallelising RankBoost.
-Nvida \ac{CUDA} and ATI Stream are the two main \ac{GPGPU} computing platform and are released by the two main \ac{GPU} vendors Nvidia and AMD.
-Experiments show a 22.9x speed-up on Nvida \ac{CUDA} and a 9.2x speed-up on ATI Stream.\\
+Nvidia \ac{CUDA} and ATI Stream are the two main \ac{GPGPU} computing platform and are released by the two main \ac{GPU} vendors Nvidia and AMD.
+Experiments show a 22.9x speed-up on Nvidia \ac{CUDA} and a 9.2x speed-up on ATI Stream.\\
 
@@
 \cite{DeSousa2012} proposed a \ac{GPGPU} approach to improve both training time and query evaluation through \ac{GPU} use.
-An association rule based Learning-to-Rank approach proposed by Veloso et al.
-\cite{Veloso2008} has been implemented using the \ac{GPU} in such a way that the set of rules van be computed in parallel, in different threads, for each document.
+An association rule based Learning-to-Rank approach, proposed by Veloso et al.
+\cite{Veloso2008}, has been implemented using the \ac{GPU} in such a way that the set of rules van be computed simultaneously for each document.
 A speed-up of 127X in query processing time is reported based on evaluation on the LETOR dataset.
@@
 
-\section{Parallelisation through execution of algorithm steps in Learning-to-Rank methods}
+\section{Parallel execution of Learning-to-Rank algorithm steps}
 Some research focused on parallelising the steps Learning-to-Rank algorithms that can be characterised as strong learners.
@@
 Tyree et al.
-\cite{Tyree2011} describe a way of parallelising \ac{GBDT} models for Learning-to-Rank where the boosting step is still executed sequentially and instead the construction of the regression trees themselves is parallelised.
-The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}.
-Decision trees are build layer-by-layer where the calculations needed for building each new layer in the tree are divided over the nodes using a master-worker paradigm.
+\cite{Tyree2011} described a way of parallelising \ac{GBDT} models for Learning-to-Rank where the boosting step is still executed sequentially, but instead the construction of the regression trees themselves is parallelised.
+The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}, which are build layer-by-layer.
+The calculations needed for building each new layer in the tree are divided among the nodes, using a master-worker paradigm.
 The data is partitioned of the workers, who compress their share into histograms and send these to the master.
 The master uses those histograms to approximate the split and build the next layer.
-The master than communicates this new layer to the workers who can use this new layer to compute new histograms.
-this process is repeated until the tree depth limit is reached.
-The tree construction algorithm parallelised with described master-worker method is \ac{CART} \cite{Breiman1984}.
+The master then communicates this new layer to the workers who can use this new layer to compute new histograms.
+This process is repeated until the tree depth limit is reached.
+The tree construction algorithm parallelised with this master-worker approach is the well-known \ac{CART} \cite{Breiman1984} algorithm.
 Speed-up experiments on the LETOR and the Yahoo!
@@
 Spark is incorporated into Hadoop since Hadoop 2.0.
-The Spark implementation of ListNet showed near a linear training time reduction.\\
+The Spark implementation of ListNet showed near linear training time reduction.\\
 
@@
 Pahikkala et al.
-\cite{Pahikkala2009} describe a pairwise \ac{RLS} type of ranking function, RankRLS, with low time complexity.
+\cite{Pahikkala2009} described a pairwise \ac{RLS} type of ranking function, RankRLS, with low time complexity.
 Airola et al.
@@
 Airola et al.
-\cite{Airola2011} and Lee and Lin \cite{Lee2014} have both described lower time complexity methods to train a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.
+\cite{Airola2011} and Lee and Lin \cite{Lee2014} both described lower time complexity methods to train a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.
 Lee and Lin \cite{Lee2014} observed that linear kernel RankSVMs are inferior in accuracy compared to nonlinear kernel RankSVMs and \ac{GBDT}s and are mainly useful to quickly produce a baseline model.
@@
 \section{Parallelisable search heuristics for Listwise ranking}
-Direct minimisation of ranking metrics is a hard problem due to the non-continuous, non-differentiable and non-convex nature of \ac{nDCG}, \ac{ERR} and \ac{MAP}.
+Direct minimisation of ranking metrics is a hard problem due to the non-continuous, non-differentiable and non-convex nature of the \ac{nDCG}, \ac{ERR} and \ac{MAP} evaluation metrics.
 This optimisation problem is generally addressed either by replacing the ranking metric with a convex surrogate, or by heuristic optimisation methods such as Simulated Annealing or a \ac{EA}.
@@
 Wang et al.
-\cite{Wang2011a,Wang2011b} propose a parallel evolutionary algorithm based on \ac{CC}, a type of evolutionary algorithm.
+\cite{Wang2011a,Wang2011b} proposed a parallel evolutionary algorithm based on \ac{CC}, a type of evolutionary algorithm.
 The \ac{CC} algorithm is capable of directly optimizing non-differentiable functions, as \ac{nDCG}, in contrary to many optimization algorithms.
 the divide-and-conquer nature of the \ac{CC} algorithm enables parallelisation.
-CCRank showed an increase in both accuracy and efficiency on the LETOR 4.0 benchmark dataset compared to the baselines.
-It must be stated however that the increased efficiency was achieved through speed-up and not scale-up.
+CCRank showed an increase in both accuracy and efficiency on the LETOR 4.0 benchmark dataset compared to its baselines.
+However, the increased efficiency was achieved through speed-up and not scale-up.
 Two reasons have been identified for not achieving linear scale-up with CCRank: 1) parallel execution is suspended after each generation to perform combination in order to produce the candidate solution, 2) Combination has to wait until all parallel tasks have finished, which may spend different running time.
@@
 \subsection{Alternating Direction Method of Multipliers}
-Duh et al. \cite{Duh2011} propose the use of \ac{ADMM} for the Learning-to-Rank task. \ac{ADMM} is a general optimization method that solves problems of the form
+Duh et al. \cite{Duh2011} proposed the use of \ac{ADMM} for the Learning-to-Rank task. \ac{ADMM} is a general optimization method that solves problems of the form
 \begin{alignat*}{2}
@@
 Duh et al.
-\cite{Duh2011} show how to use \ac{ADMM} to train a RankSVM \cite{Herbrich1999, Joachims2002} model in parallel.
+\cite{Duh2011} showed how to use \ac{ADMM} to train a RankSVM \cite{Herbrich1999, Joachims2002} model in parallel.
 Experiments showed the \ac{ADMM} implementation to achieve a 13.1x training time speed-up on 72 workers, compared to training on a single worker.\\
@@
 \cite{Boyd2012}.
-Boyd et al.
-\cite{Boyd2012} implemented an \ac{ADMM} based Learning-to-Rank method in Pregel \cite{Malewicz2010}, a parallel computation framework for graph computations.
+They implemented an \ac{ADMM} based Learning-to-Rank method in Pregel \cite{Malewicz2010}, a parallel computation framework for graph computations.
 No experimental results on parallelisation speed-up were reported on this Pregel based approach.
@@
 Acharyya et al.
-\cite{Acharyya2012, Acharyya2013} propose a Learning-to-Rank method searches for an order preserving transformation (\emph{monotone retargeting}) of the target scores that is easier for a regressor to fit.
+\cite{Acharyya2012, Acharyya2013} proposed a Learning-to-Rank method searches for an order preserving transformation (\emph{monotone retargeting}) of the target scores that is easier for a regressor to fit.
 This approach is based on the observation that it is not necessary to fit scores exactly, since the evaluation is dependent on the order and not on the pointwise predictions themselves.\\
@@
 Yun et al.
-describe a \cite{Yun2014} robust Learning-to-Rank model called RoBiRank that has the additional advantage that it is executable in parallel.
-Their RoBiRank model uses parameter expansion to linearise a surrogate loss function, after which the elements of the linear combination can be divided over the available nodes.
+described a \cite{Yun2014} robust Learning-to-Rank model called RoBiRank that has the additional advantage that it is executable in parallel.
+RoBiRank uses parameter expansion to linearise a surrogate loss function, after which the elements of the linear combination can be divided over the available nodes.
 The Parameter expansion trick was proposed by Gopal and Yang \cite{Gopal2013}, who originally proposed it for multinomial logistic regression.
@@
 Long et al.
-\cite{Long2012} describe special case of their pairwise cross-domain factor Learning-to-Rank method using distributed optimization of \ac{SGD} based on Hadoop MapReduce.
+\cite{Long2012} described special case of their pairwise cross-domain factor Learning-to-Rank method using distributed optimization of \ac{SGD} based on Hadoop MapReduce.
 Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce based method described by  Zinkevich et al.
@@
 The authors show how to perform parallel grid search on MapReduce clusters, which is easy because grid search is an embarrassingly parallel method as hyperparameter combinations are mutually independent.
-They apply their grid search on MapReduce approach in a Learning-to-Rank setting to train a LambdaMART \cite{Wu2008} ranking model, a model that uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
+They apply their grid search on MapReduce approach in a Learning-to-Rank setting to train a LambdaMART \cite{Wu2008} ranking model, which uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
 Experiments showed that the solution scales linearly in number of hyperparameter combinations.
@@
 Burges et al.
-\cite{Burges2011} describe their Yahoo!
-Learning to Rank Challenge submission that was build by performing an extensive hyperparameter search on a 122-node MPI cluster, running Microsoft HPC Server 2008.
+\cite{Burges2011} described their Yahoo!
+Learning to Rank Challenge submission that was built by performing an extensive hyperparameter search on a 122-node MPI cluster, running Microsoft HPC Server 2008.
 The hyperparameter optimisation was performed on a linear combination ensemble of eight LambdaMART models, two LambdaRank models and two MART models using a logistic regression cost.

### 2014-03-17 report/Chapters/RelatedWork.tex
betere opbouw related work

@@
 
-\section{Ensemble learning for parallel Learning-to-Rank}
-Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure called boosting \cite{Schapire1990}.
-The invention of the boosting method resulted in an increasing focus in machine learning research on methods to combine multiple weak models, which are called ensemble learning methods.
-Well-known ensemble methods include gradient boosting \cite{Friedman2001}, bagging \cite{Breiman1996}, AdaBoost \cite{Freund1997} and stacked generalisation (often called stacking) \cite{Wolpert1992}.\\
+\section{Low computational complexity Learning-to-Rank}
+\label{sec:related_work_low_complexity}
+One approach for handling large volumes of training data for Learning-to-Rank is through design of low time complexity Learning-to-Rank methods.
+Pahikkala et al.
+\cite{Pahikkala2009} described a pairwise \ac{RLS} type of ranking function, RankRLS, with low time complexity.
+Airola et al.
+\cite{Airola2010} further improved the training time complexity of RankRLS to $\mathcal{O}(tms)$, where $t$ is the number of needed iterations, $m$ the number of training documents and $s$ the number of features.
+The RankRLS ranking function showed ranking performance similar to RankSVM \cite{Herbrich1999, Joachims2002} on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.\\
 
-Ensemble methods can be used to parallelise learning methods by training the weak models in the ensemble on different nodes.
-Parallelisation of the training phase of Learning-to-Rank models is often achieved through ensemble learning, mainly boosting, and combined with decision trees, jointly called \acf{GBDT}.
-\ac{GBDT}s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
+Airola et al.
+\cite{Airola2011} and Lee and Lin \cite{Lee2014} both described lower time complexity methods to train a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.
+Lee and Lin \cite{Lee2014} observed that linear kernel RankSVMs are inferior in accuracy compared to nonlinear kernel RankSVMs and \ac{GBDT}s and are mainly useful to quickly produce a baseline model.
+Details of the lower time complexity version of the linear kernel RankSVM will not be discussed as it is shown to be an inferior Learning-to-Rank method in terms of accuracy.\\
 
-\subsection{Gradient Boosting}
-Gradient Boosting \cite{Friedman2001} is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.\\
+Learning-to-Rank methods that are specifically designed for their low computational complexity, like RankRLS and the linear kernel RankSVM methods described in this section, are generally not among the top achieving models in terms of accuracy.
+From results on benchmarks and competitions it can be observed that the best generalisation accuracy are often more complex ones.
+This makes low time complexity models as a solution for large scale Learning-to-Rank less applicable and increases the relevance of the search for efficient training of more complex Learning-to-Rank models.\\
 
-Kocsis et al.
-\cite{Kocsis2013} proposed a way to train multiple weak learners in parallel by extending those models that are likely to yield a good model when combined through boosting.
-Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular gradient boosting.
-\ac{GBDT} models could be trained in parallel using their BoostingTree algorithm.\\
+\section{Distributed hyperparameter tuning of Learning-to-Rank models}
+Hyperparameter optimisation is the task of selecting the combination of hyperparameters such that the Learning-to-Rank model shows optimal generalisation accuracy.
+Ganjisaffar et al.
+\cite{Ganjisaffar2011, Ganjisaffar2011b} observed that long training times are often a result of hyperparameter optimisation, because it results in training multiple Learning-to-Rank models.
+Grid search is the \emph{de facto} standard of hyperparameter optimisation and is simply an exhaustive search through a manually specified subset of hyperparameter combinations.
+The authors show how to perform parallel grid search on MapReduce clusters, which is easy because grid search is an embarrassingly parallel method as hyperparameter combinations are mutually independent.
+They apply their grid search on MapReduce approach in a Learning-to-Rank setting to train a LambdaMART \cite{Wu2008} ranking model, which uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
+Experiments showed that the solution scales linearly in number of hyperparameter combinations.
+However, the risk of overfitting grows overwhelmingly as the number of hyperparameter combinations grow, even when validation sets grow large.\\
 
-Ye et al.
-\cite{Ye2009} described how to implement the stochastic \ac{GBDT} model in a parallel manner using both MPI and Hadoop.
-Stochastic \ac{GBDT} differs from regular \ac{GBDT} models by using stochastic gradient boosting instead of regular gradient boosting.
-Stochastic gradient boosting is a minor alteration to regular gradient boosting that, at each iteration of the algorithm, trains a base learner on a sub-sample of the training set that is drawn at random without replacement \cite{Friedman2002}.
-Experiments showed the Hadoop implementation to resulted in too expensive communication cost to be useful.
-Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
-The MPI approach proved to be successful and obtained near linear speed-ups.\\
+Burges et al.
+\cite{Burges2011} described their Yahoo!
+Learning to Rank Challenge submission that was built by performing an extensive hyperparameter search on a 122-node MPI cluster, running Microsoft HPC Server 2008.
+The hyperparameter optimisation was performed on a linear combination ensemble of eight LambdaMART models, two LambdaRank models and two MART models using a logistic regression cost.
+This submission achieved the highest \ac{ERR} score of all Yahoo!
+Learning to Rank Challenge submissions.\\
 
-Svore and Burges \cite{Svore2010,Svore2012} designed two approaches to train the LambdaMART \cite{Wu2008} model in a distributed manner.
-Their first approach is applicable in case the whole training set fits in main memory on every node, in that case the tree split computations of the trees in the LambdaMART tree-ensemble are split instead of the data.
-The second approach distributes the training data and corresponding training computations and is therefore scalable to high amounts of training data.
-Both approaches achieves a six times speed-up over LambdaMART on 32 nodes compared to a single node.
-
-\subsection{Boosting wrapped in Bagging}
-Some approaches combine both boosting and bagging.
-Bagging \cite{Breiman1996}, also called bootstrap aggregating, is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement.
-The bagging method is parallelisable by training each $D_i$ with $i \in \{1..m\}$ on a different node.\\
-
-Yandex researchers Pavlov et al.
-\cite{Pavlov2010} were the first to proposed a boosting-wrapped-in-bagging approach, which they called BagBoo.
-The boosting step itself is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.\\
-
-Ganjisaffar et al.
-\cite{Ganjisaffar2011c, Ganjisaffar2011b} proposed a pairwise boosting-wrapped-in-bagging model called BL-MART, contrasting the pointwise BagBoo model.
-BL-MART adds a bagging step to the LambdaMART \cite{Wu2008} ranking model, that uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
-In contract to BagBoo, BL-MART is limited in the number of trees.
-An experiment on the TD2004 dataset resulted in 4500 trees using BL-MART while 1.1 million trees were created with the BagBoo model.
-
-\subsection{Stacked generalisation}
-% TODO: Kijken naar deze subsection, is erg zwak %
-A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.
-\ac{DSN} is based on the stacked generalization (stacking) ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack.
-Each layer in the stack learns has the task of learning the weights of the inputs of that layer.
-The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.
-Deng et al.
-\cite{Deng2013} % TODO: beschrijven Deng et al.%\\
-
-\subsection{Randomisation}
-Most ensemble learning based Learning-to-Rank methods are based on boosting.
-Although boosting have shown to be an effective method in Learning-to-Rank, it has the drawback that the models in the ensemble are not completely mutually independent, which makes the process harder to parallelise.
-Geurts and Louppe \cite{Geurts2011} proposed a tree based ensemble method in which the trees are built using random subsets of the available features.
-The authors called this method \ac{ET} and originally proposed this method for classification and regression settings instead of ranking settings \cite{Geurts2006}.
-The \ac{ET} algorithm is similar to the well-known Random Forest algorithm, but with two differences: 1) in \ac{ET} each tree is build on the complete dataset instead of random subsets of the data and 2) \ac{ET} sets the discretisation thresholds that define the splits in the tree randomly, instead of based on the sample.
-The randomisations in the process make the models in the ensemble mutually independent, which makes it trivial to parallelise by training each tree on a different node.
-In contrary to what one might think, \ac{ET}s show very good performance on benchmark data with a 10th place in the Yahoo!
-Learning to Rank Challenge.\\
+Notice that methods described in this section train multiple Learning-to-Rank models at the same time to find the optimal set of parameters for a model, but that the Learning-to-Rank models itself are still trained sequentially.
+In the next sections we will present literature focusing on training Learning-to-Rank models in such a way that steps in this training process can be executed simultaneously.\\
 
@@
 
-\subsection{GPGPU for Learning-to-Rank}
+\subsection{GPGPU for parallel Learning-to-Rank}
 Wang et al.
@@
 The calculations needed for building each new layer in the tree are divided among the nodes, using a master-worker paradigm.
-The data is partitioned of the workers, who compress their share into histograms and send these to the master.
+The data is partitioned and the data parts are divided between the workers, who compress their share into histograms and send these to the master.
 The master uses those histograms to approximate the split and build the next layer.
@@
 The Spark implementation of ListNet showed near linear training time reduction.\\
-
-\section{Low complexity Learning-to-Rank}
-\label{sec:related_work_low_complexity}
-Designing Learning-to-Rank algorithms with low time complexity for training is another approach towards large scale Learning-to-Rank.
-Pahikkala et al.
-\cite{Pahikkala2009} described a pairwise \ac{RLS} type of ranking function, RankRLS, with low time complexity.
-Airola et al.
-\cite{Airola2010} further improved the training time complexity of RankRLS to $\mathcal{O}(tms)$, where $t$ is the number of needed iterations, $m$ the number of training documents and $s$ the number of features.
-The RankRLS ranking function showed ranking performance similar to RankSVM \cite{Herbrich1999, Joachims2002} on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.\\
-
-Airola et al.
-\cite{Airola2011} and Lee and Lin \cite{Lee2014} both described lower time complexity methods to train a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.
-Lee and Lin \cite{Lee2014} observed that linear kernel RankSVMs are inferior in accuracy compared to nonlinear kernel RankSVMs and \ac{GBDT}s and are mainly useful to quickly produce a baseline model.
-Details of the lower time complexity version of the linear kernel RankSVM will not be discussed as it is shown to be an inferior Learning-to-Rank method in terms of accuracy.\\
 
@@
 
-\section{Distributed hyperparameter tuning using MapReduce}
+\section{Ensemble learning for parallel Learning-to-Rank}
+Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure called boosting \cite{Schapire1990}.
+The invention of the boosting method resulted in an increasing focus in machine learning research on methods to combine multiple weak models, which are called ensemble learning methods.
+Well-known ensemble methods include gradient boosting \cite{Friedman2001}, bagging \cite{Breiman1996}, AdaBoost \cite{Freund1997} and stacked generalisation (often called stacking) \cite{Wolpert1992}.\\
+
+Ensemble methods can be used to parallelise learning methods by training the weak models in the ensemble on different nodes.
+Parallelisation of the training phase of Learning-to-Rank models is often achieved through ensemble learning, mainly boosting, and combined with decision trees, jointly called \acf{GBDT}.
+\ac{GBDT}s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
+
+\subsection{Gradient Boosting}
+Gradient Boosting \cite{Friedman2001} is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.\\
+
+Kocsis et al.
+\cite{Kocsis2013} proposed a way to train multiple weak learners in parallel by extending those models that are likely to yield a good model when combined through boosting.
+Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular gradient boosting.
+\ac{GBDT} models could be trained in parallel using their BoostingTree algorithm.\\
+
+Ye et al.
+\cite{Ye2009} described how to implement the stochastic \ac{GBDT} model in a parallel manner using both MPI and Hadoop.
+Stochastic \ac{GBDT} differs from regular \ac{GBDT} models by using stochastic gradient boosting instead of regular gradient boosting.
+Stochastic gradient boosting is a minor alteration to regular gradient boosting that, at each iteration of the algorithm, trains a base learner on a sub-sample of the training set that is drawn at random without replacement \cite{Friedman2002}.
+Experiments showed the Hadoop implementation to resulted in too expensive communication cost to be useful.
+Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
+The MPI approach proved to be successful and obtained near linear speed-ups.\\
+
+Svore and Burges \cite{Svore2010,Svore2012} designed two approaches to train the LambdaMART \cite{Wu2008} model in a distributed manner.
+Their first approach is applicable in case the whole training set fits in main memory on every node, in that case the tree split computations of the trees in the LambdaMART tree-ensemble are split instead of the data.
+The second approach distributes the training data and corresponding training computations and is therefore scalable to high amounts of training data.
+Both approaches achieves a six times speed-up over LambdaMART on 32 nodes compared to a single node.
+
+\subsection{Boosting wrapped in Bagging}
+Some approaches combine both boosting and bagging.
+Bagging \cite{Breiman1996}, also called bootstrap aggregating, is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement.
+The bagging method is parallelisable by training each $D_i$ with $i \in \{1..m\}$ on a different node.\\
+
+Yandex researchers Pavlov et al.
+\cite{Pavlov2010} were the first to proposed a boosting-wrapped-in-bagging approach, which they called BagBoo.
+The boosting step itself is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.\\
+
 Ganjisaffar et al.
-\cite{Ganjisaffar2011, Ganjisaffar2011b} observed that long training times are often a result of training many models for hyperparameter optimisation.
-Grid search is the \emph{de facto} standard of hyperparameter optimisation and is simply an exhaustive search through a manually specified subset of hyperparameter combinations.
-The authors show how to perform parallel grid search on MapReduce clusters, which is easy because grid search is an embarrassingly parallel method as hyperparameter combinations are mutually independent.
-They apply their grid search on MapReduce approach in a Learning-to-Rank setting to train a LambdaMART \cite{Wu2008} ranking model, which uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
-Experiments showed that the solution scales linearly in number of hyperparameter combinations.
-However, the risk of overfitting grows overwhelmingly as the number of hyperparameter combinations grow, even when validation sets grow large.\\
+\cite{Ganjisaffar2011c, Ganjisaffar2011b} proposed a pairwise boosting-wrapped-in-bagging model called BL-MART, contrasting the pointwise BagBoo model.
+BL-MART adds a bagging step to the LambdaMART \cite{Wu2008} ranking model, that uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
+In contract to BagBoo, BL-MART is limited in the number of trees.
+An experiment on the TD2004 dataset resulted in 4500 trees using BL-MART while 1.1 million trees were created with the BagBoo model.
 
-Burges et al.
-\cite{Burges2011} described their Yahoo!
-Learning to Rank Challenge submission that was built by performing an extensive hyperparameter search on a 122-node MPI cluster, running Microsoft HPC Server 2008.
-The hyperparameter optimisation was performed on a linear combination ensemble of eight LambdaMART models, two LambdaRank models and two MART models using a logistic regression cost.
-This submission achieved the highest \ac{ERR} score of all Yahoo!
-Learning to Rank Challenge submissions.
+\subsection{Stacked generalisation}
+% TODO: Kijken naar deze subsection, is erg zwak %
+A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.
+\ac{DSN} is based on the stacked generalization (stacking) ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack.
+Each layer in the stack learns has the task of learning the weights of the inputs of that layer.
+The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.
+Deng et al.
+\cite{Deng2013} % TODO: beschrijven Deng et al.%\\
+
+\subsection{Randomisation}
+Most ensemble learning based Learning-to-Rank methods are based on boosting.
+Although boosting have shown to be an effective method in Learning-to-Rank, it has the drawback that the models in the ensemble are not completely mutually independent, which makes the process harder to parallelise.
+Geurts and Louppe \cite{Geurts2011} proposed a tree based ensemble method in which the trees are built using random subsets of the available features.
+The authors called this method \ac{ET} and originally proposed this method for classification and regression settings instead of ranking settings \cite{Geurts2006}.
+The \ac{ET} algorithm is similar to the well-known Random Forest algorithm, but with two differences: 1) in \ac{ET} each tree is build on the complete dataset instead of random subsets of the data and 2) \ac{ET} sets the discretisation thresholds that define the splits in the tree randomly, instead of based on the sample.
+The randomisations in the process make the models in the ensemble mutually independent, which makes it trivial to parallelise by training each tree on a different node.
+In contrary to what one might think, \ac{ET}s show very good performance on benchmark data with a 10th place in the Yahoo!
+Learning to Rank Challenge.\\

### 2014-03-17 report/Chapters/BenchmarkResults.tex
begin LETOR 4.0 results

@@
  \hline
-Queries & 19,994 & 2,994 & 6,983 \\ 
-Documents & 473,134 & 71,083 & 165,660 \\ 
-Features & 519 & 519 & 519 \\ 
+\# of queries & 19,994 & 2,994 & 6,983 \\ 
+\# of documents & 473,134 & 71,083 & 165,660 \\ 
+\# of features & 519 & 519 & 519 \\ 
 \end{tabular}
@@
 The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 to solve the absence of a experimental platform for Learning-to-Rank at that time.
-LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009.
-The LETOR3.0 benchmark collection \cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.\\
+LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009.\\
 
-\subsection{LETOR3.0}
+\subsection{LETOR 3.0}
+The LETOR 3.0 benchmark collection \cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.
 The OHSUMED collection is a subset of the medical publication database MEDLINE and contains medical publications from 270 journals that were published between 1987 and 1991.
-The .gov collection is a web crawl obtained in January 2002, which was used for the TREC web track in 2003 and 2004.
+The .gov collection is a web crawl obtained in January 2002, which was used for the \ac{TREC} web track in 2003 and 2004.
 Tables \ref{tbl:features_gov} and \ref{tbl:features_ohsumed} in Appendix \ref{chap:appendix_a} provide the descriptions of the features of the .gov and the OHSUMED collections respectively.\\
@@
 \end{description}
-With query sets available from both the TREC 2003 and 2004 conferences there are six query sets in total: TD2003, TD2004, NP2003, NP2004, HP2003 and HP2004.\\
+With query sets available from both the \ac{TREC} 2003 and 2004 conferences there are six query sets in total: TD2003, TD2004, NP2003, NP2004, HP2003 and HP2004.\\
 
@@
 
-I will describe the performance of the algorithms that were evaluated on LETOR3.0 after publication of the LETOR paper by Qin et al \cite{Qin2010} as listed on the LETOR website\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} by comparing their performance with the ListNet baseline.
+I will describe the performance of the algorithms that were evaluated on LETOR 3.0 after publication of the LETOR paper by Qin et al \cite{Qin2010} as listed on the LETOR website\footnotemark[2] by comparing their performance with the ListNet baseline.
 We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven datasets in \ac{nDCG} value.
@@
 \end{tabular}
-\caption{Performance of ListNet on LETOR}
+\caption{Performance of ListNet on LETOR 3.0}
 \label{tbl:LETOR_ListNet}
@@
 \end{tabular}
-\caption{Comparison of algorithms recently evaluated on LETOR3.0 with the ListNet baselines}
+\caption{\ac{nDCG}@10 comparison of algorithms recently evaluated on LETOR 3.0 with the ListNet baselines}
 \label{tbl:LETOR_recently_added}
 \end{table}
-\subsection{LETOR4.0}
+\subsection{LETOR 4.0}
+The LETOR 4.0 benchmark collection\footnote{http://http://research.microsoft.com/en-us/um/beijing/projects/letor/} consists of the Gov-2 document collection and two query datasets from the Million Query Track at \ac{TREC} 2007 (MQ2007) and \ac{TREC} 2008 (MQ2008).
+LETOR 4.0 consists of a semi-supervised ranking task, a rank aggregation task and a listwise ranking task next to the supervised ranking task.
+Table \ref{tab:letor4_characteristics} shows the collection characteristics in number of queries, documents and features for LETOR 4.0.
+Evaluation on this collection is performed using a five-fold cross-validation, where partitioning of the data into folds was performed beforehand by the creators of the MQ2007 and MQ2008 datasets.
+Documents in the dataset were
+
+\begin{table}
+\begin{tabular}{l|ll}
+ Dataset & MQ2007 & MQ2008 \\ 
+ \hline
+ Queries & 1692 & 784 \\ 
+ Documents & 69,622 & 15,211 \\
+ Features & 46 & 46 \\
+\end{tabular}
+\label{tab:letor4_characteristics}
+\caption{Characteristics of the LETOR 4.0 collection}
+\end{table}
 \subsubsection{Results}
+RankSVM-Struct, ListNet, AdaRank-\ac{nDCG}, AdaRank-\ac{MAP} and RankBoost were used as baseline models on the LETOR 4.0 dataset and were implemented and evaluated by the publishers of LETOR 4.0.
+Table \ref{tab:letor4_baseline_results} shows the performance of those baseline models on the LETOR 4.0 benchmark collection.\\
+
+\begin{table}
+\begin{tabular}{l|lll}
+Model & Dataset & Mean \ac{nDCG} & \ac{nDCG}@10 \\ 
+\hline
+RankSVM-Struct & MQ2007 & 0.4966 & 0.4439 \\ 
+ & MQ2008 & 0.4832 & 0.2279 \\ 
+\hline
+ListNet & MQ2007 & 0.4988 & 0.4440 \\ 
+ & MQ2008 & 0.4914 & 0.2303 \\ 
+\hline
+AdaRank-\ac{nDCG} & MQ2007 & 0.4914 & 0.4369 \\ 
+ & MQ2008 & 0.4950 & 0.2307 \\ 
+\hline
+AdaRank-\ac{MAP} & MQ2007 & 0.4891 & 0.4335 \\ 
+ & MQ2008 & 0.4915 & 0.2288 \\ 
+\hline
+RankBoost & MQ2007 & 0.5003 & 0.4464 \\ 
+ & MQ2008 & 0.4850 & 0.2255 \\ 
+\end{tabular}
+\caption{Comparison of LETOR 4.0 baseline models}
+\label{tab:letor4_baseline_results}
+\end{table}
+
 BoostedTree model \cite{Kocsis2013} showed an \ac{nDCG} of 0.5071 on the MQ2007 dataset and thereby beat all baseline models.

### 2014-03-17 report/Chapters/Background.tex
SVM opgenomen in acronyms

@@
 
-Well-known pairwise Learning-to-Rank algorithms include FRank \cite{Tsai2007}, GBRank \cite{Zheng2007}, LambdaRank \cite{Burges2006}, RankBoost \cite{Freund2003}, RankNet \cite{Burges2005}, Ranking SVM \cite{Herbrich1999b,Joachims2002}, and SortNet \cite{Rigutini2008}.
+Well-known pairwise Learning-to-Rank algorithms include FRank \cite{Tsai2007}, GBRank \cite{Zheng2007}, LambdaRank \cite{Burges2006}, RankBoost \cite{Freund2003}, RankNet \cite{Burges2005}, Ranking \acs{SVM} \cite{Herbrich1999b,Joachims2002}, and SortNet \cite{Rigutini2008}.
 \subsection{Listwise Approach}
@@
 
-Well-known algorithms that belong to the listwise approach include AdaRank \cite{Xu2007}, BoltzRank \cite{Volkovs2009}, ListMLE \cite{Xia2008}, ListNet \cite{Cao2007}, RankCosine \cite{Qin2008}, SmoothRank \cite{Chapelle2010}, SoftRank \cite{Taylor2008}, SVM$^{map}$ \cite{Yue2007}.
+Well-known algorithms that belong to the listwise approach include AdaRank \cite{Xu2007}, BoltzRank \cite{Volkovs2009}, ListMLE \cite{Xia2008}, ListNet \cite{Cao2007}, RankCosine \cite{Qin2008}, SmoothRank \cite{Chapelle2010}, SoftRank \cite{Taylor2008}, \acs{SVM}$^{map}$ \cite{Yue2007}.

### 2014-03-17 report/Chapters/BenchmarkResults.tex
SVM opgenomen in acronyms

@@
 	\begin{itemize}
-	\item RankingSVM \cite{Herbrich1999,Joachims2002}
+	\item Rank\ac{SVM} \cite{Herbrich1999,Joachims2002}
 	\item RankBoost \cite{Freund2003}
@@
 	\item AdaRank-nDCG \cite{Xu2007}
-	\item SVMMAP \cite{Yue2007} 
+	\item \ac{SVM}$^{MAP}$ \cite{Yue2007} 
 	\end{itemize}}
@@
 \begin{tabular}{l|p{1.2cm}p{1.2cm}p{1.2cm}p{1.4cm}||l}
- & \rotatebox{55}{Regression + L2 regulatisation} & \rotatebox{55}{RankSVM-Primal} &\rotatebox{55}{RankSVM-Struct} & \rotatebox{55}{SmoothRank} & \rotatebox{55}{ListNet} \\
+ & \rotatebox{55}{Regression + L2 regulatisation} & \rotatebox{55}{Rank\ac{SVM}-Primal} &\rotatebox{55}{Rank\ac{SVM}-Struct} & \rotatebox{55}{SmoothRank} & \rotatebox{55}{ListNet} \\
  \hline
@@
 \subsubsection{Results}
-RankSVM-Struct, ListNet, AdaRank-\ac{nDCG}, AdaRank-\ac{MAP} and RankBoost were used as baseline models on the LETOR 4.0 dataset and were implemented and evaluated by the publishers of LETOR 4.0.
+Rank\ac{SVM}-Struct, ListNet, AdaRank-\ac{nDCG}, AdaRank-\ac{MAP} and RankBoost were used as baseline models on the LETOR 4.0 dataset and were implemented and evaluated by the publishers of LETOR 4.0.
 Table \ref{tab:letor4_baseline_results} shows the performance of those baseline models on the LETOR 4.0 benchmark collection.\\
@@
 \hline
-RankSVM-Struct & MQ2007 & 0.4966 & 0.4439 \\ 
+Rank\ac{SVM}-Struct & MQ2007 & 0.4966 & 0.4439 \\ 
  & MQ2008 & 0.4832 & 0.2279 \\ 

### 2014-03-17 report/Chapters/RelatedWork.tex
related work zin aangepast
SVM opgenomen in acronyms

@@
 Figure \ref{fig:related_work_categories} illustrates the categories of scalable training approaches in Learning-to-Rank.
-The following sections describe the related work in the field based on this categorisation.
+The numbers in Figure \ref{fig:related_work_categories} correspond to the sections that describe the related work belonging to these categories.
 
@@
 \cite{Airola2010} further improved the training time complexity of RankRLS to $\mathcal{O}(tms)$, where $t$ is the number of needed iterations, $m$ the number of training documents and $s$ the number of features.
-The RankRLS ranking function showed ranking performance similar to RankSVM \cite{Herbrich1999, Joachims2002} on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.\\
+The RankRLS ranking function showed ranking performance similar to Rank\acs{SVM} \cite{Herbrich1999, Joachims2002} on the BioInfer corpus \cite{Pyysalo2007}, a corpus for information extraction in the biomedical domain.\\
 
 Airola et al.
-\cite{Airola2011} and Lee and Lin \cite{Lee2014} both described lower time complexity methods to train a linear kernel RankSVM \cite{Herbrich1999, Joachims2002}.
-Lee and Lin \cite{Lee2014} observed that linear kernel RankSVMs are inferior in accuracy compared to nonlinear kernel RankSVMs and \ac{GBDT}s and are mainly useful to quickly produce a baseline model.
-Details of the lower time complexity version of the linear kernel RankSVM will not be discussed as it is shown to be an inferior Learning-to-Rank method in terms of accuracy.\\
+\cite{Airola2011} and Lee and Lin \cite{Lee2014} both described lower time complexity methods to train a linear kernel ranking \ac{SVM} \cite{Herbrich1999, Joachims2002}.
+Lee and Lin \cite{Lee2014} observed that linear kernel Rank\ac{SVM}s are inferior in accuracy compared to nonlinear kernel Rank\ac{SVM}s and \ac{GBDT}s and are mainly useful to quickly produce a baseline model.
+Details of the lower time complexity version of the linear kernel Rank\ac{SVM} will not be discussed as it is shown to be an inferior Learning-to-Rank method in terms of accuracy.\\
 
-Learning-to-Rank methods that are specifically designed for their low computational complexity, like RankRLS and the linear kernel RankSVM methods described in this section, are generally not among the top achieving models in terms of accuracy.
+Learning-to-Rank methods that are specifically designed for their low computational complexity, like RankRLS and the linear kernel Rank\ac{SVM} methods described in this section, are generally not among the top achieving models in terms of accuracy.
 From results on benchmarks and competitions it can be observed that the best generalisation accuracy are often more complex ones.
@@
 Duh et al.
-\cite{Duh2011} showed how to use \ac{ADMM} to train a RankSVM \cite{Herbrich1999, Joachims2002} model in parallel.
+\cite{Duh2011} showed how to use \ac{ADMM} to train a Rank\ac{SVM} \cite{Herbrich1999, Joachims2002} model in parallel.
 Experiments showed the \ac{ADMM} implementation to achieve a 13.1x training time speed-up on 72 workers, compared to training on a single worker.\\
@@
 focussed their research more on robust ranking than on parallel ranking.
-The only reference to performance of RoBiRank in terms of speed is the statement that its training time on a computing cluster is comparable to the more efficient implementation by Lee and Lin \cite{Lee2014} of a linear kernel RankSVM \cite{Herbrich1999, Joachims2002} described in section \ref{sec:related_work_low_complexity}.
+The only reference to performance of RoBiRank in terms of speed is the statement that its training time on a computing cluster is comparable to the more efficient implementation by Lee and Lin \cite{Lee2014} of a linear kernel Rank\ac{SVM} \cite{Herbrich1999, Joachims2002} described in section \ref{sec:related_work_low_complexity}.
 \subsection{Distributed Stochastic Gradient Descent}

### 2014-03-18 report/Chapters/BenchmarkResults.tex
kleine afwerkingsdingetjes gefixt in de trein

@@
 Learning to Rank Challenge in terms of ERR score.
-Burges et al \cite{Burges2011} created a linear combination ensemble of eight LambdaMART \cite{Burges2010}, two LambdaRank and two Logistic Regression models.
+Burges  \cite{Burges2011} created a linear combination ensemble of eight LambdaMART \cite{Burges2010}, two LambdaRank and two Logistic Regression models.
 Gottschalk and Vogel used a combination of RandomForest models and \ac{GBDT} models.
@@
 
-The challenge overview paper states as one of the lessons of the challenge that the simple baseline \ac{GBDT} model performed very well with a small performance gap to the complex ensemble submissions at the top of the table.\\
+The challenge overview paper \cite{Chapelle2011a} states as one of the lessons of the challenge that the simple baseline \ac{GBDT} model performed very well with a small performance gap to the complex ensemble submissions at the top of the table.\\
 
@@
  \hline 
-1 & Burges et al (Microsoft Research) & 0.46861 \\ 
+1 & Burges et al. (Microsoft Research) & 0.46861 \\ 
 2 & Gottschalk (Activision Blizzard) \& Vogel (Data Mining Solutions) & 0.46786 \\ 
@@
 The bagging component of the BagBoo method enables it to achieve high scalability through parallelism.
-Pavlov et al \cite{Pavlov2010} managed to train half a million trees on 200 nodes in 2 hours.
+Pavlov et al.
+\cite{Pavlov2010} managed to train half a million trees on 200 nodes in 2 hours.
 \section{Yandex Internet Mathematics competition}
@@
 \subsection{Results}
-The Yandex Internet Mathematics competition was won by Pavlov et al \cite{Pavlov2010} with the BagBoo method.
+The Yandex Internet Mathematics competition was won by Pavlov et al. \cite{Pavlov2010} with the BagBoo method.
 
 \section{LETOR}
-The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 to solve the absence of a experimental platform for Learning-to-Rank at that time.
+The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 \cite{Liu2007b} to solve the absence of a experimental platform for Learning-to-Rank at that time.
 LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009.\\
@@
 \subsection{LETOR 3.0}
-The LETOR 3.0 benchmark collection \cite{Liu2007b} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.
+The LETOR 3.0 benchmark collection \cite{Qin2010} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.
 The OHSUMED collection is a subset of the medical publication database MEDLINE and contains medical publications from 270 journals that were published between 1987 and 1991.
@@
 \includegraphics[scale=0.30]{gfx/ndcg_winning_number}
-\caption{Comparison across the seven datasets in LETOR by \ac{nDCG}, obtained from Qin et al \cite{Qin2010}}
+\caption{Comparison across the seven datasets in LETOR by \ac{nDCG}, obtained from Qin et al. \cite{Qin2010}}
 \label{fig:ndcg_winning_number}
@@
 \includegraphics[scale=0.30]{gfx/map_winning_number}
-\caption{Comparison across the seven datasets in LETOR by \ac{MAP}, obtained from Qin et al \cite{Qin2010}}
+\caption{Comparison across the seven datasets in LETOR by \ac{MAP}, obtained from Qin et al. \cite{Qin2010}}
 \label{fig:map_winning_number}
@@
 
-I will describe the performance of the algorithms that were evaluated on LETOR 3.0 after publication of the LETOR paper by Qin et al \cite{Qin2010} as listed on the LETOR website\footnotemark[2] by comparing their performance with the ListNet baseline.
+I will describe the performance of the algorithms that were evaluated by the LETOR team on LETOR 3.0 after publication of the LETOR paper by Qin et al.
+\cite{Qin2010}, as listed on the LETOR website\footnotemark[2] by comparing their performance with the ListNet baseline.
 We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven datasets in \ac{nDCG} value.
@@
 
+Since LETOR is arguably the most well-known benchmark collection in the field, it is conceivable that creators of new Learning-to-Rank methods evaluate their new method on the LETOR collection to show how well their new method works.
+To create an overview of the performance of newer Learning-to-Rank methods on LETOR 3.0 we perform a forward reference search on the LETOR 3.0 paper \cite{Qin2010}.
+It should be noted that results of evaluations individually performed by the creator of the evaluated Learning-to-Rank method can not be regarded as equally reliable as the evaluations performed by the LETOR team.
+TODO: 307 citations, x evaluated a new Learning-to-Rank method on LETOR 3.0, x were usable.
+The unusable studies did test a new Learning-to-Rank method on one or more datasets of the LETOR collection, but 1 ) did not use a ranking metric used in LETOR (\ac{nDCG} or \ac{MAP}) or 2) did not list evaluation results at all.\\
+
 \begin{table}
@@
 \end{table}
+
 \subsection{LETOR 4.0}
@@
 \item Differences in feature sets between data sets detract from fair comparison
-\item Although the \ac{nDCG} definition is unambiguous, Busa-Fekete et al \cite{Busa-Fekete2012} found that \ac{nDCG} evaluation tools of benchmark data sets produced different scores
+\item Although the \ac{nDCG} definition is unambiguous, Busa-Fekete et al. \cite{Busa-Fekete2012} found that \ac{nDCG} evaluation tools of benchmark data sets produced different scores
 \end{enumerate}

### 2014-03-18 report/Chapters/Introduction.tex
kleine afwerkingsdingetjes gefixt in de trein

@@
 For example, it plays a vital role in amongst others: automatic document summarisation, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}.
-In addition, the ranking task has been argued to be more fitting to recommender systems than rating prediction on a continuous scale \cite{Adomavicius2005,McNee2006}.\\
+In addition, the ranking task has been argued to be more fitting to recommender systems than regression-based rating prediction on a continuous scale \cite{Adomavicius2005,McNee2006}.\\
 
@@
 Well-known benchmark datasets include the \emph{Yahoo!
-Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR dataset \cite{Qin2010} that was build by Microsoft Research.
+Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR dataset \cite{Qin2010} that was published by Microsoft Research.
 One of the concluding observations of the \emph{Yahoo!

### 2014-03-20 report/Chapters/BenchmarkResults.tex
WCL2R dataset

@@
 \end{tabular}
+\caption{Characteristics of the LETOR 4.0 collection}
 \label{tab:letor4_characteristics}
-\caption{Characteristics of the LETOR 4.0 collection}
 \end{table}
@@
 \end{enumerate}
+
+\section{Other datasets}
+\subsection{WCL2R}
+The WCL2R collection, released by Alc{\^a}ntara et al \cite{Alcantara2010}, contains of two datasets from the Chilean search engine TodoCL\footnote{www.todocl.cl}.
+Both datasets contain approximately 3 million documents.
+WCL2R is the only benchmark collection known that provides click-through data on user level.
+The collection contains 277,095 queries and logs of in total 1,5 million clicks by 16,829 users.
+The WCL2R paper provides evaluation results for the well-known Rank\ac{SVM} \cite{Herbrich1999,Joachims2002} and RankBoost \cite{Freund2003} algorithms.
+In addition two ranking methods developed at the same university of the WCL2R paper were evaluated: LAC \cite{Veloso2008} and a ranking algorithm based on \ac{GP} \cite{DeAlmeida2007}.

### 2014-03-21 report/Chapters/Background.tex
graffito + begin benchmark_results.csv

@@
 \end{tabular}
-\caption{Example calculation for \ac{nDCG}}
+\caption{Example calculation of \acs{nDCG}}
 \label{tab:example_calculation_nDCG}
@@
 The algorithm requires relevance grades $g_i$, $1 \le i \le n$ and mapping function $R$ that maps relevance grades to probability of relevance.
-\begin{lstlisting}[caption={Algorithm to compute the ERR metric, obtained from \cite{Chapelle2009}}, label={lst:err},language=Ada]
+\begin{lstlisting}[caption={Algorithm to compute the \acs{ERR} metric, obtained from \cite{Chapelle2009}}, label={lst:err},language=Ada]
 p <- 1, ERR <- 0

### 2014-03-21 report/Chapters/BenchmarkResults.tex
WCL2R resultaten
graffito + begin benchmark_results.csv

@@
 \includegraphics[scale=0.30]{gfx/ndcg_winning_number}
-\caption{Comparison across the seven datasets in LETOR by \ac{nDCG}, obtained from Qin et al. \cite{Qin2010}}
+\caption{Comparison across the seven datasets in LETOR by \acs{nDCG}, obtained from Qin et al. \cite{Qin2010}}
 \label{fig:ndcg_winning_number}
@@
 \includegraphics[scale=0.30]{gfx/map_winning_number}
-\caption{Comparison across the seven datasets in LETOR by \ac{MAP}, obtained from Qin et al. \cite{Qin2010}}
+\caption{Comparison across the seven datasets in LETOR by \acs{MAP}, obtained from Qin et al. \cite{Qin2010}}
 \label{fig:map_winning_number}
@@
 \end{tabular}
-\caption{\ac{nDCG}@10 comparison of algorithms recently evaluated on LETOR 3.0 with the ListNet baselines}
+\caption{\acs{nDCG}@10 comparison of algorithms recently evaluated on LETOR 3.0 with the ListNet baselines}
 \label{tbl:LETOR_recently_added}
@@
 In addition two ranking methods developed at the same university of the WCL2R paper were evaluated: LAC \cite{Veloso2008} and a ranking algorithm based on \ac{GP} \cite{DeAlmeida2007}.
+Table \ref{tab:results_WCL2R} shows the performance of those algorithms on the WCL2R collection.
+No evaluations of other algorithms on the WCL2R collection are known.
+
+\begin{table}
+\begin{tabular}{l|lll|lll}
+ &  & FS &  &  & NC &  \\ 
+\hline
+Algorithm & @1 & @3 & @10 & @1 & @3 & @10 \\ 
+\hline
+RankSVM & 0.314 & 0.353 & 0.395 & 0.265 & 0.301 & 0.339 \\ 
+LAC & 0.296 & 0.360 & 0.403 & 0.244 & 0.266 & 0.315 \\ 
+GP & 0.288 & 0.344 & 0.396 & 0.221 & 0.262 & 0.318 \\ 
+RankBoost & 0.295 & 0.328 & 0.375 & 0.247 & 0.264 & 0.305 \\ 
+\end{tabular}
+\caption{\acs{nDCG} results of baseline methods on the WCL2R collection, obtained from \cite{Alcantara2010}}
+\label{tab:results_WCL2R}
+\end{table}

### 2014-03-24 report/Chapters/BenchmarkResults.tex
begin cross dataset comparison

@@
 \end{table}
+
+\subsection{AOL}
+TODO: AOL dataset beschrijven \cite{Pass2006}.
+
+\section{comparing Learning-to-Rank models across datasets}
+\begin{table}[!h]
+\begin{tabular}{l|l|l}
+Benchmark & Paper & \# of forward references \\
+\hline
+LETOR 1.0 \& 2.0 & Liu et al. \cite{Liu2007b} & 307\\
+LETOR 3.0 & Qin et al. \cite{Qin2010} & 105\\
+Yahoo! Learning to Rank Challenge & Chapelle et al. \cite{Chapelle2011a} & 102 forward references\\
+AOL dataset & Pass et al. \cite{Pass2006} & 339\\
+WCL2R & Alc{\^a}ntara et al. \cite{Alcantara2010} & 2\\
+\end{tabular}
+\caption{Forward references of Learning-to-Rank benchmark papers}
+\label{tbl:ltr_benchmark_forref}
+\end{table}
+Google scholar search results for benchmark data sets without a release/overview paper.
+\begin{table}
+\begin{tabular}{l|l}
+Benchmark & Google scholar search results \\
+\hline
+LETOR 4.0 & 75 results \\
+MSLR-web10k & 16 results \\
+MSLR-web30k & 15 results \\
+Yandex Internet Mathematics Competition & 1 result \\ 
+\end{tabular}
+\caption{Google scholar search results for Learning-to-Rank benchmarks}
+\label{tbl:ltr_benchmark_searchres}
+\end{table}

### 2014-03-25 method_selection/winning_number.R
begin 'winning number'-R script. data cleaning map_result.csv

@@ -1,0 +1,2 @@
+setwd("C:/Git-data/Afstuderen/method_selection")
+ltr_data = read.csv("map_result.csv")

### 2014-03-25 method_selection/winning_number.R
nog meer foute naamgeving

@@ -1,2 +1,2 @@
 setwd("C:/Git-data/Afstuderen/method_selection")
-ltr_data = read.csv("map_result.csv")
+ltr_data = read.csv("ndcg5_results.csv")

### 2014-03-25 method_selection/winning_number.R
winning numbers kunnen worden berekend :)

@@ -1,2 +1,4 @@
 setwd("C:/Git-data/Afstuderen/method_selection")
-ltr_data = read.csv("ndcg5_results.csv")
+ltr_data = read.csv("ndcg10_results.csv")
+temp <- ltr_data[,c(2:(length(ltr_data)-2))]
+winnum <- apply(temp, 2, function(x) sapply(x, function(y) if(is.na(y)){0}else{sum(x[!is.na(x)]<y)}))

### 2014-03-25 method_selection/winning_number.R
winning number en ideal winning number geimplementeerd

@@ -1,4 +1,6 @@
 setwd("C:/Git-data/Afstuderen/method_selection")
 ltr_data = read.csv("ndcg10_results.csv")
 temp <- ltr_data[,c(2:(length(ltr_data)-2))]
-winnum <- apply(temp, 2, function(x) sapply(x, function(y) if(is.na(y)){0}else{sum(x[!is.na(x)]<y)}))
+winnum <- apply(temp, 2, function(x) sapply(x, function(y) if(is.na(y)){NA}else{sum(x[!is.na(x)]<y)}))
+winnum <- data.frame(winnum)
+ideal_winnum <- apply(winnum, 2, function(x) sapply(x, function(y) if(is.na(y)){0}else{if(length(x[!is.na(x)]) == 0){0}else{max(x[!is.na(x)])}}))

### 2014-03-25 method_selection/winning_number.R
winning number visualisation

@@ -15,4 +15,15 @@ nr_measurements <- apply(temp,1,function(x) sum(!is.na(x)))
 
 output <- data.frame("Method" = ltr_data[,1], "Normalised_winning_number" = norm_winnum, "Nr_measurements"=nr_measurements)
 
-ggplot(output, aes(Nr_measurements,Normalised_winning_number)) + geom_point() + geom_text(aes(label=Method))
+# add ID's to output
+output$id <- seq_len(nrow(output))
+nc <- ncol(output)
+output <- output[, c(nc, 1:(nc-1))]
+
+# plot graph
+ggplot(output, aes(Nr_measurements,Normalised_winning_number)) + geom_point(size=2) + 
+  geom_text(aes(label=id), hjust=-0.5, vjust=-0.5) + 
+  theme_grey(base_size=12, base_family="") + 
+  scale_x_continuous("# of datasets evaluated on", expand = c(0,0), breaks=c(0,2,4,6,8,10)) +
+  scale_y_continuous("Normalised Winning Number", expand = c(0,0), breaks=c(0,0.2,0.4,0.6,0.8,1.0)) +
+  expand_limits(x = c(0,9.2), y = c(0,1.01))

### 2014-03-25 method_selection/winning_number.R
uitlijning nu altijd goed

@@ -1,7 +1,7 @@
 setwd("C:/Git-data/Afstuderen/method_selection")
 require(ggplot2)
 
-ltr_data = read.csv("map_results.csv")
+ltr_data = read.csv("ndcg5_results.csv")
 temp <- ltr_data[,c(2:(length(ltr_data)-2))]
 winnum <- apply(temp, 2, function(x) sapply(x, function(y) if(is.na(y)){NA}else{sum(x[!is.na(x)]<y)}))
 winnum <- data.frame(winnum)
@@ -26,4 +26,4 @@ ggplot(output, aes(Nr_measurements,Normalised_winning_number)) + geom_point(size
   theme_grey(base_size=12, base_family="") + 
   scale_x_continuous("# of datasets evaluated on", expand = c(0,0), breaks=c(0,2,4,6,8,10)) +
   scale_y_continuous("Normalised Winning Number", expand = c(0,0), breaks=c(0,0.2,0.4,0.6,0.8,1.0)) +
-  expand_limits(x = c(0,9.2), y = c(0,1.01))
+  expand_limits(x = c(0,9.35), y = c(0,1.03))

### 2014-03-25 method_selection/winning_number.R
uitlijning nu echt goed

@@ -1,7 +1,7 @@
 setwd("C:/Git-data/Afstuderen/method_selection")
 require(ggplot2)
 
-ltr_data = read.csv("ndcg5_results.csv")
+ltr_data = read.csv("ndcg10_results.csv")
 temp <- ltr_data[,c(2:(length(ltr_data)-2))]
 winnum <- apply(temp, 2, function(x) sapply(x, function(y) if(is.na(y)){NA}else{sum(x[!is.na(x)]<y)}))
 winnum <- data.frame(winnum)
@@ -26,4 +26,4 @@ ggplot(output, aes(Nr_measurements,Normalised_winning_number)) + geom_point(size
   theme_grey(base_size=12, base_family="") + 
   scale_x_continuous("# of datasets evaluated on", expand = c(0,0), breaks=c(0,2,4,6,8,10)) +
   scale_y_continuous("Normalised Winning Number", expand = c(0,0), breaks=c(0,0.2,0.4,0.6,0.8,1.0)) +
-  expand_limits(x = c(0,9.35), y = c(0,1.03))
+  expand_limits(x = c(0,10.35), y = c(0,1.03))

### 2014-03-26 report/Chapters/BenchmarkResults.tex
realisatie dat OHSUMED, TD2003 en TD2004 niet hetzelfde zijn voor LETOR 2.0 en LETOR 3.0. LETOR 2.0 baselinewaarden ingevoerd
Yahoo references tm Rank-PMBGP

@@
 \section{comparing Learning-to-Rank models across datasets}
+LETOR 2.0 baseline results from LETOR website \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor2.0/baseline.aspx}
 \begin{table}[!h]
@@
 LETOR 3.0 & Qin et al. \cite{Qin2010} & 105\\
-Yahoo! Learning to Rank Challenge & Chapelle et al. \cite{Chapelle2011a} & 102 forward references\\
+Yahoo! Learning to Rank Challenge & Chapelle et al. \cite{Chapelle2011a} & 102\\
 AOL dataset & Pass et al. \cite{Pass2006} & 339\\

### 2014-03-26 method_selection/winning_number.R

@@ -1,7 +1,7 @@
 setwd("C:/Git-data/Afstuderen/method_selection")
 require(ggplot2)
 
-ltr_data = read.csv("ndcg10_results.csv")
+ltr_data = read.csv("map_results.csv")
 temp <- ltr_data[,c(2:(length(ltr_data)-2))]
 winnum <- apply(temp, 2, function(x) sapply(x, function(y) if(is.na(y)){NA}else{sum(x[!is.na(x)]<y)}))
 winnum <- data.frame(winnum)
@@ -24,6 +24,6 @@ output <- output[, c(nc, 1:(nc-1))]
 ggplot(output, aes(Nr_measurements,Normalised_winning_number)) + geom_point(size=2) + 
   geom_text(aes(label=id), hjust=-0.5, vjust=-0.5) + 
   theme_grey(base_size=12, base_family="") + 
-  scale_x_continuous("# of datasets evaluated on", expand = c(0,0), breaks=c(0,2,4,6,8,10)) +
+  scale_x_continuous("# of datasets evaluated on", expand = c(0,0), breaks=c(0,2,4,6,8,10,12)) +
   scale_y_continuous("Normalised Winning Number", expand = c(0,0), breaks=c(0,0.2,0.4,0.6,0.8,1.0)) +
-  expand_limits(x = c(0,10.35), y = c(0,1.03))
+  expand_limits(x = c(0,13.35), y = c(0,1.03))

### 2014-03-27 method_selection/winning_number.R
LETOR 4.0 resultaten tm FSMRank

@@ -1,7 +1,7 @@
 setwd("C:/Git-data/Afstuderen/method_selection")
 require(ggplot2)
 
-ltr_data = read.csv("map_results.csv")
+ltr_data = read.csv("ndcg10_results.csv")
 temp <- ltr_data[,c(2:(length(ltr_data)-2))]
 winnum <- apply(temp, 2, function(x) sapply(x, function(y) if(is.na(y)){NA}else{sum(x[!is.na(x)]<y)}))
 winnum <- data.frame(winnum)
@@ -24,6 +24,6 @@ output <- output[, c(nc, 1:(nc-1))]
 ggplot(output, aes(Nr_measurements,Normalised_winning_number)) + geom_point(size=2) + 
   geom_text(aes(label=id), hjust=-0.5, vjust=-0.5) + 
   theme_grey(base_size=12, base_family="") + 
-  scale_x_continuous("# of datasets evaluated on", expand = c(0,0), breaks=c(0,2,4,6,8,10,12)) +
+  scale_x_continuous("# of datasets evaluated on", expand = c(0,0), breaks=c(0,2,4,6,8,10,12,14, 16)) +
   scale_y_continuous("Normalised Winning Number", expand = c(0,0), breaks=c(0,0.2,0.4,0.6,0.8,1.0)) +
-  expand_limits(x = c(0,13.35), y = c(0,1.03))
+  expand_limits(x = c(0,16.35), y = c(0,1.03))

### 2014-04-02 report/Chapters/BenchmarkResults.tex
eerste begin beschrijving comparative LTR study

--- 2014-03-24 report/Chapters/BenchmarkResults.tex
@@
 \section{Yandex Internet Mathematics competition}
-
+Yandex Internet Mathematics Competition
 \subsection{Results}
@@
 
-\section{comparing Learning-to-Rank models across datasets}
+\section{Comparing Learning-to-Rank models across datasets}
+As we have seen in the foregoing sections, evaluation of Learning-to-Rank methods is spread out over the several benchmark datasets.
+However, as the evaluated Learning-to-Rank methods differ from benchmark to benchmark, no single benchmark comparison can be regarded as a conclusive argument on which Learning-to-Rank method is most accurate.
+Several studies make a small start in considering Learning-to-Rank method performance over multiple benchmark datasets.
+Gomes et al.
+\cite{Gomes2013} analysed ranking accuracy of a set of models over both LETOR 3.0 and LETOR 4.0.
+Busa-Fekete et al.
+\cite{Busa-Fekete2013} compared the accuracy of a small set of models over the LETOR 4.0 datasets, both MSLR datasets, both Yahoo!
+Learning to Rank Challenge datasets and the OHSUMED dataset from LETOR 3.0.
+To my knowledge, no large-scale accuracy meta-analysis has been conducted where published evaluation results on several benchmark datasets are taken into account.\\
+
+With a meta-analysis I will compare the performance of Learning-to-Rank methods across the Learning-to-Rank benchmark datasets described in foregoing sections.
+With a literature review I will collect evaluation results on the datasets/collections.
+The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
+\begin{itemize}
+\item LETOR 2.0
+\item LETOR 3.0
+\item LETOR 4.0
+\item Yahoo! Learning to Rank Challenge
+\item Yandex Internet Mathematics Competition 2009
+\item MSLR-web10/30k
+\item WCL2R
+\item AOL
+\end{itemize}
+
+For the LETOR collections, the evaluation results of the baseline models will be used from LETOR 2.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor2.0/baseline.aspx}, LETOR 3.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and LETOR 4.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} as listed on the LETOR website.\\
+
+LETOR 1.0, LETOR 3.0, Yahoo!
+Learning to Rank Challenge, WCL2R and AOL have an accompanying paper that was published together with the benchmark collection.
+Users of those benchmark collections are encouraged to cite these papers.
+A forward literature search on these studies is performed to collect studies that present measurements of evaluations of Learning-to-Rank methods on these benchmark collections.
+Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search.
+
 \begin{table}[!h]
@@
 \end{table}
-Google scholar search results for benchmark data sets without a release/overview paper.
+
+The LETOR 4.0, MSLR-web10/30k and Yandex Internet Mathematics Competition 2009 benchmark collections were not accompanied with a describing study.
+To collect measurements of Learning-to-Rank methods evaluated on these benchmarks, a Google Scholar search is performed on the name of the benchmark.
+Table~\ref{chap:benchmark_results} shows the results of this literature search.
+
 \begin{table}

### 2014-04-02 report/Chapters/BenchmarkResults.tex
werk tm vergaderring

@@
 
+\subsection{Comparison methodology}
 With a meta-analysis I will compare the performance of Learning-to-Rank methods across the Learning-to-Rank benchmark datasets described in foregoing sections.
@@
 \end{table}
+
+The LETOR 3.0 paper \cite{Qin2010} states that it may differ between datasets what the most accurate ranking methods are.
+To evaluate the overall performance of Learning-to-Rank methods over the multiple datasets in the LETOR 3.0 collections, Qin et al.
+\cite{Qin2010} proposed a measure called \emph{winning number} as the number of other algorithms that an algorithm can beat over the set of datasets.
+Formally the winning number measure is defined as
+
+$S_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$
+
+where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{nDCG} or \ac{MAP}), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that
+
+$I_{\{M_i(j)>M_k(j)\}} = \begin{cases}
+1 \text{if} M_i(j) > M_k(k),
+0 \text{otherwise}0
+\end{cases}$

### 2014-04-03 report/Chapters/BenchmarkResults.tex
TODO's er in geknald

@@
 \cite{Qin2010} proposed a measure called \emph{winning number} as the number of other algorithms that an algorithm can beat over the set of datasets.
-Formally the winning number measure is defined as
+Formally the winning number measure is defined as\\
 
-$S_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$
+$S_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$\\
 
-where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{nDCG} or \ac{MAP}), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that
+where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{nDCG} or \ac{MAP}), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that\\
 
 $I_{\{M_i(j)>M_k(j)\}} = \begin{cases}
-1 \text{if} M_i(j) > M_k(k),
-0 \text{otherwise}0
-\end{cases}$
+1 & \text{if } M_i(j) > M_k(k), \\
+0 & \text{otherwise}
+\end{cases}$\\
+
+In contrast to the winning number comparison on LETOR 3.0, there will not be accuracy measurements for each algorithm on each dataset in our meta-analysis. To compare
+
+$TODO: wat is normalised winning number? waarom is dit nodig?$
+$TODO: waarom ndcg@3,5,10 en MAP?$
+$TODO: limitations van deze vergelijkingsmethodologie$

### 2014-04-03 report/Chapters/BenchmarkResults.tex
cross benchmark comparison naar nieuwe chapter

--- 2014-04-02 report/Chapters/BenchmarkResults.tex
@@
 TODO: AOL dataset beschrijven \cite{Pass2006}.
-
-\section{Comparing Learning-to-Rank models across datasets}
-As we have seen in the foregoing sections, evaluation of Learning-to-Rank methods is spread out over the several benchmark datasets.
-However, as the evaluated Learning-to-Rank methods differ from benchmark to benchmark, no single benchmark comparison can be regarded as a conclusive argument on which Learning-to-Rank method is most accurate.
-Several studies make a small start in considering Learning-to-Rank method performance over multiple benchmark datasets.
-Gomes et al.
-\cite{Gomes2013} analysed ranking accuracy of a set of models over both LETOR 3.0 and LETOR 4.0.
-Busa-Fekete et al.
-\cite{Busa-Fekete2013} compared the accuracy of a small set of models over the LETOR 4.0 datasets, both MSLR datasets, both Yahoo!
-Learning to Rank Challenge datasets and the OHSUMED dataset from LETOR 3.0.
-To my knowledge, no large-scale accuracy meta-analysis has been conducted where published evaluation results on several benchmark datasets are taken into account.\\
-
-With a meta-analysis I will compare the performance of Learning-to-Rank methods across the Learning-to-Rank benchmark datasets described in foregoing sections.
-With a literature review I will collect evaluation results on the datasets/collections.
-The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
-\begin{itemize}
-\item LETOR 2.0
-\item LETOR 3.0
-\item LETOR 4.0
-\item Yahoo! Learning to Rank Challenge
-\item Yandex Internet Mathematics Competition 2009
-\item MSLR-web10/30k
-\item WCL2R
-\item AOL
-\end{itemize}
-
-For the LETOR collections, the evaluation results of the baseline models will be used from LETOR 2.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor2.0/baseline.aspx}, LETOR 3.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and LETOR 4.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} as listed on the LETOR website.\\
-
-LETOR 1.0, LETOR 3.0, Yahoo!
-Learning to Rank Challenge, WCL2R and AOL have an accompanying paper that was published together with the benchmark collection.
-Users of those benchmark collections are encouraged to cite these papers.
-A forward literature search on these studies is performed to collect studies that present measurements of evaluations of Learning-to-Rank methods on these benchmark collections.
-Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search.
-
-\begin{table}[!h]
-\begin{tabular}{l|l|l}
-Benchmark & Paper & \# of forward references \\
-\hline
-LETOR 1.0 \& 2.0 & Liu et al. \cite{Liu2007b} & 307\\
-LETOR 3.0 & Qin et al. \cite{Qin2010} & 105\\
-Yahoo! Learning to Rank Challenge & Chapelle et al. \cite{Chapelle2011a} & 102\\
-AOL dataset & Pass et al. \cite{Pass2006} & 339\\
-WCL2R & Alc{\^a}ntara et al. \cite{Alcantara2010} & 2\\
-\end{tabular}
-\caption{Forward references of Learning-to-Rank benchmark papers}
-\label{tbl:ltr_benchmark_forref}
-\end{table}
-
-The LETOR 4.0, MSLR-web10/30k and Yandex Internet Mathematics Competition 2009 benchmark collections were not accompanied with a describing study.
-To collect measurements of Learning-to-Rank methods evaluated on these benchmarks, a Google Scholar search is performed on the name of the benchmark.
-Table~\ref{chap:benchmark_results} shows the results of this literature search.
-
-\begin{table}
-\begin{tabular}{l|l}
-Benchmark & Google scholar search results \\
-\hline
-LETOR 4.0 & 75 results \\
-MSLR-web10k & 16 results \\
-MSLR-web30k & 15 results \\
-Yandex Internet Mathematics Competition & 1 result \\ 
-\end{tabular}
-\caption{Google scholar search results for Learning-to-Rank benchmarks}
-\label{tbl:ltr_benchmark_searchres}
-\end{table}

### 2014-04-03 report/Chapters/CrossBenchmarkComparison.tex
cross benchmark comparison naar nieuwe chapter

As we have seen in chapter~\ref{chap:benchmark_results}, evaluation of Learning-to-Rank methods is spread out over the several benchmark datasets. However, as the evaluated Learning-to-Rank methods differ from benchmark to benchmark, no single benchmark comparison can be regarded as a conclusive argument on which Learning-to-Rank method is most accurate. Several studies make a small start in considering Learning-to-Rank method performance over multiple benchmark datasets. Gomes et al. \cite{Gomes2013} analysed ranking accuracy of a set of models over both LETOR 3.0 and LETOR 4.0. Busa-Fekete et al. \cite{Busa-Fekete2013} compared the accuracy of a small set of models over the LETOR 4.0 datasets, both MSLR datasets, both Yahoo! Learning to Rank Challenge datasets and the OHSUMED dataset from LETOR 3.0. To my knowledge, no large-scale accuracy meta-analysis has been conducted where published evaluation results on several benchmark datasets are taken into account.\\

\subsection{Comparison methodology}
With a meta-analysis I will compare the performance of Learning-to-Rank methods across the Learning-to-Rank benchmark datasets described in foregoing sections. With a literature review I will collect evaluation results on the datasets/collections. The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
\begin{itemize}
\item LETOR 2.0
\item LETOR 3.0
\item LETOR 4.0
\item Yahoo! Learning to Rank Challenge
\item Yandex Internet Mathematics Competition 2009
\item MSLR-web10/30k
\item WCL2R
\item AOL
\end{itemize}

For the LETOR collections, the evaluation results of the baseline models will be used from LETOR 2.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor2.0/baseline.aspx}, LETOR 3.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and LETOR 4.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} as listed on the LETOR website.\\

LETOR 1.0, LETOR 3.0, Yahoo! Learning to Rank Challenge, WCL2R and AOL have an accompanying paper that was published together with the benchmark collection. Users of those benchmark collections are encouraged to cite these papers. A forward literature search on these studies is performed to collect studies that present measurements of evaluations of Learning-to-Rank methods on these benchmark collections. Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search.

\begin{table}[!h]
\begin{tabular}{l|l|l}
Benchmark & Paper & \# of forward references \\
\hline
LETOR 1.0 \& 2.0 & Liu et al. \cite{Liu2007b} & 307\\
LETOR 3.0 & Qin et al. \cite{Qin2010} & 105\\
Yahoo! Learning to Rank Challenge & Chapelle et al. \cite{Chapelle2011a} & 102\\
AOL dataset & Pass et al. \cite{Pass2006} & 339\\
WCL2R & Alc{\^a}ntara et al. \cite{Alcantara2010} & 2\\
\end{tabular}
\caption{Forward references of Learning-to-Rank benchmark papers}
\label{tbl:ltr_benchmark_forref}
\end{table}

The LETOR 4.0, MSLR-web10/30k and Yandex Internet Mathematics Competition 2009 benchmark collections were not accompanied with a describing study. To collect measurements of Learning-to-Rank methods evaluated on these benchmarks, a Google Scholar search is performed on the name of the benchmark. Table~\ref{chap:benchmark_results} shows the results of this literature search.

\begin{table}
\begin{tabular}{l|l}
Benchmark & Google scholar search results \\
\hline
LETOR 4.0 & 75 results \\
MSLR-web10k & 16 results \\
MSLR-web30k & 15 results \\
Yandex Internet Mathematics Competition & 1 result \\ 
\end{tabular}
\caption{Google scholar search results for Learning-to-Rank benchmarks}
\label{tbl:ltr_benchmark_searchres}
\end{table}

The LETOR 3.0 paper \cite{Qin2010} states that it may differ between datasets what the most accurate ranking methods are. To evaluate the overall performance of Learning-to-Rank methods over the multiple datasets in the LETOR 3.0 collections, Qin et al. \cite{Qin2010} proposed a measure called \emph{winning number} as the number of other algorithms that an algorithm can beat over the set of datasets. Formally the winning number measure is defined as\\

$\text{Winning Number}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$\\

where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{nDCG} or \ac{MAP}), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that\\

$I_{\{M_i(j)>M_k(j)\}} = \begin{cases}
1 & \text{if } M_i(j) > M_k(j), \\
0 & \text{otherwise}
\end{cases}$\\

In contrast to the winning number comparison on LETOR 3.0, there will not be accuracy measurements for each algorithm on each dataset in our meta-analysis. To compare algorithms based on a sparse set of evaluation measurements, a normalised version of the winning number metric will be used which we calculate by only taking those datasets into account that an algorithm is evaluated on and then dividing this by the ideal winning number that an algorithm could have had if it would have been the best algorithm on all datasets on which it has been evaluated. I will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as \\

$I_{\{M_i(j)>M_k(j)\}} = \begin{cases}
1 & \text{if } M_i(j) \text{ and } M_k(j) \text{ are both defined and } M_i(j) > M_k(j), \\
0 & \text{otherwise}
\end{cases}$\\

From now on this adjusted version of winning number will be references to as \emph{normalised winning number}. The mathematical definition of normalised winning number is\\

$\text{Normalised Winning Number}_i(M) = \frac{\text{Winning Number}_i(M)}{\text{Ideal Winning Number}_i(M)}$\\

\noindent
where ideal winning number is defined as\\

$\text{Ideal Winning Number}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m D_{\{M_i(j),M_k(j)\}}$\\

where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{nDCG} or \ac{MAP}), and $D_{\{M_i(j),M_k(j)\}}$ is an evaluation definition function such that\\

$D_{\{M_i(j),M_k(j)\}} = \begin{cases}
1 & \text{if } M_i(j) \text{ and } M_k(j) \text{ are both defined}, \\
0 & \text{otherwise}
\end{cases}$\\

%TODO: waarom ndcg@3,5,10 en MAP?$
%TODO: limitations van deze vergelijkingsmethodologie$

### 2014-04-03 report/Chapters/CrossBenchmarkComparison.tex
Structuur Cross Benchmark comparison verbeterd

@@
 As we have seen in chapter~\ref{chap:benchmark_results}, evaluation of Learning-to-Rank methods is spread out over the several benchmark datasets.
-However, as the evaluated Learning-to-Rank methods differ from benchmark to benchmark, no single benchmark comparison can be regarded as a conclusive argument on which Learning-to-Rank method is most accurate.
+However, as the evaluated Learning-to-Rank methods differ from benchmark to benchmark, no single benchmark comparison can be regarded as a conclusive argument on which Learning-to-Rank method is most accurate.\\
+
 Several studies make a small start in considering Learning-to-Rank method performance over multiple benchmark datasets.
@@
 Learning to Rank Challenge datasets and the OHSUMED dataset from LETOR 3.0.
-To my knowledge, no large-scale accuracy meta-analysis has been conducted where published evaluation results on several benchmark datasets are taken into account.\\
+To my knowledge, no large-scale accuracy meta-analysis has been conducted where published evaluation results on several benchmark datasets are taken into account.
+With a meta-analysis I will compare the performance of Learning-to-Rank methods across the Learning-to-Rank benchmark datasets described in foregoing sections.
 
-\subsection{Comparison methodology}
-With a meta-analysis I will compare the performance of Learning-to-Rank methods across the Learning-to-Rank benchmark datasets described in foregoing sections.
-With a literature review I will collect evaluation results on the datasets/collections.
-The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
+\section{Collecting evaluation measurements}
+With a literature review I will collect evaluation results on the datasets/collections. The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
 \begin{itemize}
@@
 
-\begin{table}
+\begin{table}[!h]
 \begin{tabular}{l|l}
@@
 
+\section{Comparison methodology}
 The LETOR 3.0 paper \cite{Qin2010} states that it may differ between datasets what the most accurate ranking methods are.
@@
 
+\section{Results \& Discussion}
+\begin{figure}[!h]
+\includegraphics[scale=0.40]{gfx/combined_normalized_winnum}
+\caption{Cross-benchmark comparison of Learning-to-Rank methods}
+\label{fig:normalised_winning_number_all}
+\end{figure}
 %TODO: waarom ndcg@3,5,10 en MAP?$
 %TODO: limitations van deze vergelijkingsmethodologie$
+
+\section{Limitations}

### 2014-04-03 report/Chapters/CrossBenchmarkComparison.tex
LAC == LAC-MR-OR

@@
 
-\section{Collecting evaluation measurements}
+\section{Collecting Evaluation Results}
 With a literature review I will collect evaluation results on the datasets/collections. The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
@@
 
-\section{Comparison methodology}
+\section{Comparison Methodology}
 The LETOR 3.0 paper \cite{Qin2010} states that it may differ between datasets what the most accurate ranking methods are.
@@
 
+\section{Evaluation Results Found in Literature}
+
 \section{Results \& Discussion}
+\subsection{MAP}
+
+\subsection{NDCG@3}
+
+\subsection{NDCG@5}
+
+\subsection{NDCG@10}
+
+\subsection{All metrics}
 \begin{figure}[!h]

### 2014-04-03 report/Chapters/CrossBenchmarkComparison.tex
tabel

@@
 \section{Evaluation Results Found in Literature}
+\begin{tabular}{llll}
+AdaRank-\ac{MAP} & FenchelRank & LambdaRank & RankDE \\ 
+AdaRank-\ac{nDCG} & FocusedBoost & LARF & Ranking Committee Perceptron \\ 
+ADMM & FocusedNet & ListMLE & RankMGP \\ 
+Approx\ac{AP} & FocusedSVM & ListNet & RankNet \\ 
+Approx\ac{nDCG} & FPRank & ListReg & RankRLS \\ 
+BagBoo & FRank & LRUF & RankSVM \\ 
+Best Single Feature & FSMRank & MBP & RankSVM-Primal \\ 
+BL-MART & FSM$^{MAP}$ & MHR & RankSVM-Struct \\ 
+BoltzRank-Single & GAS-E & MultiStageBoost & RE-QR \\ 
+BoltzRank-Pair & GPRank & NewLoss & REG-SHF-SDCG \\ 
+BT & GRankRLS & OWPC & Regression \\ 
+C-CRF & GroupCE & RankELM (pointwise) & RSRank \\ 
+CA & GroupMLE & RankELM (pairwise) & SmoothGrad \\ 
+ &  & TM & VFLR \\ 
+CCRank & IntervalRank & PERF-MAP & SmoothRank \\ 
+CoList & KeepRank & PermuRank & SoftRank \\ 
+Consistent-RankCosine & Kernel-PCA & Q.D.KNN & SortNet \\ 
+CRR & KL-CRF & RandomForest & SparseRank \\ 
+DCMP & Regression with L2-regularisation & Rank-PMBGP & SVD-RankBoost \\ 
+DirectRank & LAC-MR-OR & RankAgg\ac{nDCG} & SVM$^{MAP}$ \\ 
+Energy\ac{nDCG} & LambdaMART & RankBoost & SwarmRank \\ 
+FBPCRank & LambdaNeuralRank & RankCSA & TGRank \\ 
+\end{tabular}
 
@@
 \end{figure}
-%TODO: waarom ndcg@3,5,10 en MAP?$
-%TODO: limitations van deze vergelijkingsmethodologie$
+%TODO: waarom ndcg@3,5,10 en MAP?
+%TODO: limitations van deze vergelijkingsmethodologie
 

### 2014-04-03 method_selection/winning_number.R

@@ -130,7 +130,9 @@ combined_output <- combined_output[, c(ncol(combined_output), 1:(ncol(combined_o
 #  expand_limits(x = c(0,16.35), y = c(0,1.03))
 
 ggplot(combined_aggregate, aes(Summed_ideal_winnum,Norm_winnum)) + geom_point(size=2) +  
-  geom_text(aes(label=Method)) +
-  scale_x_continuous("Summed Ideal Winning Number", expand = c(0,0)) + 
-  scale_y_continuous("Normalized Winnin Number", expand=c(0,0)) +
-  expand_limits(x = c(0,1000), y = c(0,1.01))
+#  geom_text(aes(label=Method)) +
+  scale_x_continuous("Ideal Winning Number", expand = c(0,0)) + 
+  scale_y_continuous("Normalized Winning Number", expand=c(0,0)) +
+  expand_limits(x = c(0,1000), y = c(0,1.01)) +
+  theme(axis.title.y = element_text(size = rel(1.8), angle = 90)) +
+  theme(axis.title.x = element_text(size = rel(1.8), angle = 00))

### 2014-04-04 report/Chapters/CrossBenchmarkComparison.tex
veel pijn en moeite in overzichtstabel. CCRank == CCR

--- 2014-04-03 report/Chapters/CrossBenchmarkComparison.tex
@@
 \section{Evaluation Results Found in Literature}
+\begin{table}[!h]
+\begin{tabular}{l|l|l}
+AdaRank-\acs{MAP} & \acs{GP}Rank & RankAgg\ac{nDCG} \\ 
+AdaRank-\acs{nDCG} & GRank\acs{RLS} & RankBoost \\ 
+ADMM & GroupCE & RankCSA \\ 
+Approx\acs{AP} & Group\acs{MLE} & RankDE \\ 
+Approx\acs{nDCG} & IntervalRank & RankELM (pairwise) \\ 
+BagBoo & \acs{IP}Rank & RankELM (pointwise) \\ 
+Best Single Feature & KeepRank & Ranking Committee Perceptron \\ 
+BL-MART & Kernel-\acs{PCA} & RankMGP \\ 
+BoltzRank-Single & KL-\acs{CRF} & RankNet \\ 
+BoltzRank-Pair & LAC-MR-OR & Rank\ac{RLS} \\ 
+BT & LambdaMART & Rank\ac{SVM} \\ 
+C-\acs{CRF} & LambdaNeuralRank & Rank\ac{SVM}-Primal \\ 
+\acs{CA} & LambdaRank & Rank\acs{SVM}-Struct \\ 
+\ac{CC}Rank & LARF & RE-QR \\ 
+CoList & Linear Regression & REG-SHF-SDCG \\ 
+Consistent-RankCosine & List\acs{MLE} & Regression \\ 
+DCMP & ListNet & Ridge Regression \\ 
+DirectRank & ListReg & RSRank \\ 
+Energy\ac{nDCG} & LRUF & SmoothGrad \\ 
+FBPCRank & MBP & SmoothRank \\ 
+FenchelRank & MHR & SoftRank \\ 
+FocusedBoost & MultiStageBoost & SortNet \\ 
+FocusedNet & NewLoss & SparseRank \\ 
+Focused\ac{SVM} & OWPC & \acs{SVD}-RankBoost \\ 
+FPRank & PERF-\ac{MAP} & \ac{SVM}$^{\ac{MAP}}$ \\ 
+FRank & PermuRank & SwarmRank \\ 
+FSMRank & Q.D.\acs{KNN} & TGRank \\ 
+FSM$^{\acs{MAP}}$ & RandomForest & TM \\ 
+GAS-E & Rank-PMBGP & VFLR \\ 
+\end{tabular}
+\caption{Learning-to-Rank algorithms with measurements on benchmark datasets}
+\label{ltr_methods_used}
+\end{table}
 

### 2014-04-04 report/Chapters/CrossBenchmarkComparison.tex
begin toevoegen references aan lijst methods

--- 2014-04-03 report/Chapters/CrossBenchmarkComparison.tex
@@
 \section{Evaluation Results Found in Literature}
+\begin{table}[!h]
+\begin{tabular}{l|l|l}
+AdaRank-\acs{MAP} & \acs{GP}Rank & RankAgg\ac{nDCG} \\ 
+AdaRank-\acs{nDCG} & GRank\acs{RLS} & RankBoost \\ 
+ADMM & Group\acs{CE} \cite{Lin2011} & RankCSA \cite{He2010} \\ 
+Approx\acs{AP} \cite{Qin2010b} & Group\acs{MLE} & RankDE \\ 
+Approx\acs{nDCG} \cite{Qin2010b} & IntervalRank \cite{Moon2010} & RankELM (pairwise) \\ 
+BagBoo & \acs{IP}Rank & RankELM (pointwise) \\ 
+Best Single Feature & KeepRank & Ranking Committee Perceptron \cite{Elsas2008} \\ 
+BL-MART \cite{Ganjisaffar2011c} & Kernel-\acs{PCA} & RankMGP \\ 
+BoltzRank-Single \cite{Volkovs2009} & KL-\acs{CRF} & RankNet \\ 
+BoltzRank-Pair \cite{Volkovs2009} & LAC-MR-OR \cite{Veloso2008} & Rank\ac{RLS} \\ 
+BT \cite{Zhou2008} & LambdaMART & Rank\ac{SVM} \\ 
+C-\acs{CRF} \cite{Qin2008b} & LambdaNeuralRank \cite{Papini2012} & Rank\ac{SVM}-Primal \\ 
+\acs{CA} & LambdaRank & Rank\acs{SVM}-Struct \\ 
+\ac{CC}Rank & LARF & RE-QR \\ 
+CoList & Linear Regression & REG-SHF-SDCG \\ 
+Consistent-RankCosine & List\acs{MLE} & Regression \\ 
+DCMP \cite{Renjifo2012} & ListNet & Ridge Regression \\ 
+DirectRank & ListReg & RSRank \\ 
+Energy\ac{nDCG} \cite{Freno2011} & LRUF & SmoothGrad \\ 
+FBPCRank \cite{Lai2011} & MBP & SmoothRank \\ 
+FenchelRank \cite{Lai2013} & MHR & SoftRank \\ 
+FocusedBoost \cite{Niu2012} & MultiStageBoost & SortNet \cite{Rigutini2008} \\ 
+FocusedNet \cite{Niu2012} & NewLoss & SparseRank \\ 
+Focused\ac{SVM} \cite{Niu2012} & OWPC & \acs{SVD}-RankBoost \cite{Lin2009} \\ 
+FPRank & PERF-\ac{MAP} & \ac{SVM}$^{\ac{MAP}}$ \\ 
+FRank & PermuRank \cite{Xu2008} & SwarmRank \\ 
+FSMRank & Q.D.\acs{KNN} & TGRank \\ 
+FSM$^{\acs{MAP}}$ & RandomForest & TM \cite{Zhou2008} \\ 
+GAS-E & Rank-PMBGP \cite{Sato2013} & VFLR \\ 
+\end{tabular}
+\caption{Learning-to-Rank algorithms with measurements on benchmark datasets}
+\label{ltr_methods_used}
+\end{table}
 

### 2014-04-04 report/Chapters/CrossBenchmarkComparison.tex
described en evaluated kolommen gesplitst

--- 2014-04-03 report/Chapters/CrossBenchmarkComparison.tex
@@
 \section{Evaluation Results Found in Literature}
-\begin{tabular}{llll}
-AdaRank-\ac{MAP} & FenchelRank & LambdaRank & RankDE \\ 
-AdaRank-\ac{nDCG} & FocusedBoost & LARF & Ranking Committee Perceptron \\ 
-ADMM & FocusedNet & ListMLE & RankMGP \\ 
-Approx\ac{AP} & FocusedSVM & ListNet & RankNet \\ 
-Approx\ac{nDCG} & FPRank & ListReg & RankRLS \\ 
-BagBoo & FRank & LRUF & RankSVM \\ 
-Best Single Feature & FSMRank & MBP & RankSVM-Primal \\ 
-BL-MART & FSM$^{MAP}$ & MHR & RankSVM-Struct \\ 
-BoltzRank-Single & GAS-E & MultiStageBoost & RE-QR \\ 
-BoltzRank-Pair & GPRank & NewLoss & REG-SHF-SDCG \\ 
-BT & GRankRLS & OWPC & Regression \\ 
-C-CRF & GroupCE & RankELM (pointwise) & RSRank \\ 
-CA & GroupMLE & RankELM (pairwise) & SmoothGrad \\ 
- &  & TM & VFLR \\ 
-CCRank & IntervalRank & PERF-MAP & SmoothRank \\ 
-CoList & KeepRank & PermuRank & SoftRank \\ 
-Consistent-RankCosine & Kernel-PCA & Q.D.KNN & SortNet \\ 
-CRR & KL-CRF & RandomForest & SparseRank \\ 
-DCMP & Regression with L2-regularisation & Rank-PMBGP & SVD-RankBoost \\ 
-DirectRank & LAC-MR-OR & RankAgg\ac{nDCG} & SVM$^{MAP}$ \\ 
-Energy\ac{nDCG} & LambdaMART & RankBoost & SwarmRank \\ 
-FBPCRank & LambdaNeuralRank & RankCSA & TGRank \\ 
-\end{tabular}
+\begin{table}[!h]
+\scalebox{0.85}{
+\begin{tabular}{|l|l|l||l|l|l|}
+Method & Described & Evaluated & Method & Described & Evaluated \\ 
+AdaRank-\acs{MAP} &  &  & Linear Regression &  &  \\ 
+AdaRank-\acs{nDCG} &  &  & List\acs{MLE} &  &  \\ 
+ADMM &  &  & ListNet &  &  \\ 
+Approx\acs{AP} & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
+Approx\acs{nDCG} & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
+BagBoo &  &  & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
+Best Single Feature &  &  & MHR &  &  \\ 
+BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
+BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009} & NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
+BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
+BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-\acs{MAP} &  &  \\ 
+C-\acs{CRF} & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
+CA &  &  & Q.D.\acs{KNN} &  &  \\ 
+CCRank &  &  & RandomForest & \cite{Gomes2013} & \cite{Gomes2013} \\ 
+CoList &  &  & Rank-PMBGP &  &  \\ 
+Consistent-RankCosine &  &  & RankAgg\acs{nDCG} &  &  \\ 
+DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost &  &  \\ 
+DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
+Energy\acs{nDCG} & \cite{Freno2011} & \cite{Freno2011} & RankDE &  &  \\ 
+FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FenchelRank & \cite{Lai2013} & \cite{Lai2013} & RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FocusedBoost & \cite{Niu2012} & \cite{Niu2012} & Ranking Committee Perceptron & \cite{Elsas2008} & \cite{Elsas2008} \\ 
+FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankMGP &  &  \\ 
+Focused\acs{SVM} & \cite{Niu2012} & \cite{Niu2012} & RankNet &  &  \\ 
+FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{RLS} &  &  \\ 
+FRank &  &  & Rank\acs{SVM} &  &  \\ 
+FSMRank & \cite{Lai2013c} & \cite{Lai2013c} & Rank\acs{SVM}-Primal &  &  \\ 
+FSM$^{\acs{MAP}}$ &  &  & Rank\acs{SVM}-Struct &  &  \\ 
+GAS-E &  &  & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\ 
+\acs{GP}Rank &  &  & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\ 
+GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & Ridge Regression &  &  \\ 
+Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & RSRank &  &  \\ 
+Group\acs{MLE} &  &  & SmoothGrad &  &  \\ 
+IntervalRank & \cite{Moon2010} & \cite{Moon2010} & SmoothRank & \cite{Chapelle2010} & \cite{Chapelle2010} \\ 
+\acs{IP}Rank &  &  & SoftRank &  &  \\ 
+KeepRank & \cite{Chen2009} & \cite{Chen2009} & SoftNet & \cite{Rigutini2008} & \cite{Rigutini2008} \\ 
+Kernel-\acs{PCA} &  &  & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
+KL-\acs{CRF} &  &  & \acs{SVD}-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\ 
+LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & \acs{SVM}$^{\acs{MAP}}$ &  &  \\ 
+LambdaMART &  &  & SwarmRank &  &  \\ 
+LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank &  &  \\ 
+LambdaRank &  &  & TM & \cite{Zhou2008} & \cite{Zhou2008} \\ 
+LARF & \cite{Torkestani2012} & \cite{Torkestani2012} & VFLR &  &  \\ 
+\end{tabular}}
+\caption{Learning-to-Rank algorithms with measurements on benchmark datasets}
+\label{ltr_methods_used}
+\end{table}
 

### 2014-04-04 report/Chapters/CrossBenchmarkComparison.tex
tabel references

--- 2014-04-04 report/Chapters/CrossBenchmarkComparison.tex
@@
 \begin{table}[!h]
-\begin{tabular}{l|l|l}
-AdaRank-\acs{MAP} & \acs{GP}Rank & RankAgg\ac{nDCG} \\ 
-AdaRank-\acs{nDCG} & GRank\acs{RLS} & RankBoost \\ 
-ADMM & GroupCE & RankCSA \\ 
-Approx\acs{AP} & Group\acs{MLE} & RankDE \\ 
-Approx\acs{nDCG} & IntervalRank & RankELM (pairwise) \\ 
-BagBoo & \acs{IP}Rank & RankELM (pointwise) \\ 
-Best Single Feature & KeepRank & Ranking Committee Perceptron \\ 
-BL-MART & Kernel-\acs{PCA} & RankMGP \\ 
-BoltzRank-Single & KL-\acs{CRF} & RankNet \\ 
-BoltzRank-Pair & LAC-MR-OR & Rank\ac{RLS} \\ 
-BT & LambdaMART & Rank\ac{SVM} \\ 
-C-\acs{CRF} & LambdaNeuralRank & Rank\ac{SVM}-Primal \\ 
-\acs{CA} & LambdaRank & Rank\acs{SVM}-Struct \\ 
-\ac{CC}Rank & LARF & RE-QR \\ 
-CoList & Linear Regression & REG-SHF-SDCG \\ 
-Consistent-RankCosine & List\acs{MLE} & Regression \\ 
-DCMP & ListNet & Ridge Regression \\ 
-DirectRank & ListReg & RSRank \\ 
-Energy\ac{nDCG} & LRUF & SmoothGrad \\ 
-FBPCRank & MBP & SmoothRank \\ 
-FenchelRank & MHR & SoftRank \\ 
-FocusedBoost & MultiStageBoost & SortNet \\ 
-FocusedNet & NewLoss & SparseRank \\ 
-Focused\ac{SVM} & OWPC & \acs{SVD}-RankBoost \\ 
-FPRank & PERF-\ac{MAP} & \ac{SVM}$^{\ac{MAP}}$ \\ 
-FRank & PermuRank & SwarmRank \\ 
-FSMRank & Q.D.\acs{KNN} & TGRank \\ 
-FSM$^{\acs{MAP}}$ & RandomForest & TM \\ 
-GAS-E & Rank-PMBGP & VFLR \\ 
-\end{tabular}
+\scalebox{0.85}{
+\begin{tabular}{|l|l|l||l|l|l|}
+Method & Described & Evaluated & Method & Described & Evaluated \\
+\hline 
+AdaRank-\acs{MAP} &  &  & Linear Regression &  & \cite{Wang2012} \\ 
+AdaRank-\acs{nDCG} &  & \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} &  &  \\ 
+ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet &  &  \\ 
+Approx\acs{AP} & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
+Approx\acs{nDCG} & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
+BagBoo &  &  & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
+Best Single Feature &  &  & MHR &  &  \\ 
+BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
+BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009} & NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
+BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
+BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-\acs{MAP} &  &  \\ 
+C-\acs{CRF} & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
+CA &  & \cite{Busa-Fekete2013} & Q.D.\acs{KNN} &  &  \\ 
+CCRank &  &  & RandomForest & \cite{Gomes2013} & \cite{Gomes2013} \\ 
+CoList & \cite{Gao2014} & \cite{Gao2014} & Rank-PMBGP &  &  \\ 
+Consistent-RankCosine &  &  & RankAgg\acs{nDCG} &  &  \\ 
+DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost &  & \cite{Busa-Fekete2013} \\ 
+DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
+Energy\acs{nDCG} & \cite{Freno2011} & \cite{Freno2011} & RankDE &  &  \\ 
+FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FenchelRank & \cite{Lai2013} & \cite{Lai2013} & RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FocusedBoost & \cite{Niu2012} & \cite{Niu2012} & RankMGP & \cite{Lin2012} & \cite{Lin2012} \\ 
+FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankNet &  & \cite{Busa-Fekete2013} \\ 
+Focused\acs{SVM} & \cite{Niu2012} & \cite{Niu2012} & Rank\acs{RLS} &  &  \\ 
+FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{SVM} &  & \cite{Busa-Fekete2013, Freno2011} \\ 
+FRank &  & \cite{Wang2012} & Rank\acs{SVM}-Primal &  &  \\ 
+FSMRank & \cite{Lai2013c} & \cite{Lai2013c} & Rank\acs{SVM}-Struct &  &  \\
+FSM$^{\acs{MAP}}$ & & & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
+GAS-E &  &  & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\ 
+\acs{GP}Rank &  &  & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\ 
+GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & Ridge Regression &  &  \\ 
+Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & RSRank &  &  \\ 
+Group\acs{MLE} &  &  & SmoothGrad &  &  \\ 
+IntervalRank & \cite{Moon2010} & \cite{Moon2010} & SmoothRank & \cite{Chapelle2010} & \cite{Chapelle2010} \\ 
+\acs{IP}Rank &  &  & SoftRank &  & \cite{Qin2010b} \\ 
+KeepRank & \cite{Chen2009} & \cite{Chen2009} & SoftNet & \cite{Rigutini2008} & \cite{Rigutini2008} \\ 
+Kernel-\acs{PCA} &  &  & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
+KL-\acs{CRF} & \cite{Volkovs2011} &  & \acs{SVD}-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\ 
+LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & \acs{SVM}$^{\acs{MAP}}$ &  & \cite{Wang2012} \\ 
+LambdaMART & \cite{Burges2010} & \cite{Asadi2013a} & SwarmRank &  &  \\ 
+LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank &  &  \\ 
+LambdaRank &  &  & TM & \cite{Zhou2008} & \cite{Zhou2008} \\ 
+LARF & \cite{Torkestani2012} & \cite{Torkestani2012} & VFLR & \cite{Cai2012} & \cite{Cai2012} \\ 
+\end{tabular}}
 \caption{Learning-to-Rank algorithms with measurements on benchmark datasets}

### 2014-04-04 report/Chapters/CrossBenchmarkComparison.tex
ndcg10 verwerkt in tabel

--- 2014-04-04 report/Chapters/CrossBenchmarkComparison.tex
@@
 \begin{table}[!h]
-\begin{tabular}{l|l|l}
-AdaRank-\acs{MAP} & \acs{GP}Rank & RankAgg\ac{nDCG} \\ 
-AdaRank-\acs{nDCG} & GRank\acs{RLS} & RankBoost \\ 
-ADMM & Group\acs{CE} \cite{Lin2011} & RankCSA \cite{He2010} \\ 
-Approx\acs{AP} \cite{Qin2010b} & Group\acs{MLE} & RankDE \\ 
-Approx\acs{nDCG} \cite{Qin2010b} & IntervalRank \cite{Moon2010} & RankELM (pairwise) \\ 
-BagBoo & \acs{IP}Rank & RankELM (pointwise) \\ 
-Best Single Feature & KeepRank & Ranking Committee Perceptron \cite{Elsas2008} \\ 
-BL-MART \cite{Ganjisaffar2011c} & Kernel-\acs{PCA} & RankMGP \\ 
-BoltzRank-Single \cite{Volkovs2009} & KL-\acs{CRF} & RankNet \\ 
-BoltzRank-Pair \cite{Volkovs2009} & LAC-MR-OR \cite{Veloso2008} & Rank\ac{RLS} \\ 
-BT \cite{Zhou2008} & LambdaMART & Rank\ac{SVM} \\ 
-C-\acs{CRF} \cite{Qin2008b} & LambdaNeuralRank \cite{Papini2012} & Rank\ac{SVM}-Primal \\ 
-\acs{CA} & LambdaRank & Rank\acs{SVM}-Struct \\ 
-\ac{CC}Rank & LARF & RE-QR \\ 
-CoList & Linear Regression & REG-SHF-SDCG \\ 
-Consistent-RankCosine & List\acs{MLE} & Regression \\ 
-DCMP \cite{Renjifo2012} & ListNet & Ridge Regression \\ 
-DirectRank & ListReg & RSRank \\ 
-Energy\ac{nDCG} \cite{Freno2011} & LRUF & SmoothGrad \\ 
-FBPCRank \cite{Lai2011} & MBP & SmoothRank \\ 
-FenchelRank \cite{Lai2013} & MHR & SoftRank \\ 
-FocusedBoost \cite{Niu2012} & MultiStageBoost & SortNet \cite{Rigutini2008} \\ 
-FocusedNet \cite{Niu2012} & NewLoss & SparseRank \\ 
-Focused\ac{SVM} \cite{Niu2012} & OWPC & \acs{SVD}-RankBoost \cite{Lin2009} \\ 
-FPRank & PERF-\ac{MAP} & \ac{SVM}$^{\ac{MAP}}$ \\ 
-FRank & PermuRank \cite{Xu2008} & SwarmRank \\ 
-FSMRank & Q.D.\acs{KNN} & TGRank \\ 
-FSM$^{\acs{MAP}}$ & RandomForest & TM \cite{Zhou2008} \\ 
-GAS-E & Rank-PMBGP \cite{Sato2013} & VFLR \\ 
-\end{tabular}
+\scalebox{0.85}{
+\begin{tabular}{|l|l|l||l|l|l|}
+Method & Described & Evaluated & Method & Described & Evaluated \\
+\hline 
+AdaRank-\acs{MAP} &  &  & Linear Regression &  & \cite{Wang2012} \\ 
+AdaRank-\acs{nDCG} &  & \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} &  & \cite{Lin2010, Gao2014} \\ 
+ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet &  &  \\ 
+Approx\acs{AP} & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
+Approx\acs{nDCG} & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
+BagBoo &  &  & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
+Best Single Feature &  & \cite{Gomes2013} & MHR &  &  \\ 
+BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
+BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009} & NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
+BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
+BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-\acs{MAP} &  & \cite{Torkestani2012b} \\ 
+C-\acs{CRF} & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
+CA &  & \cite{Busa-Fekete2013,Tan2013} & Q.D.\acs{KNN} &  &  \\ 
+CCRank &  &  & RandomForest &  & \cite{Gomes2013} \\ 
+CoList & \cite{Gao2014} & \cite{Gao2014} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
+Consistent-RankCosine &  & \cite{Tan2013} & RankAgg\acs{nDCG} &  &  \\ 
+DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost &  & \cite{Busa-Fekete2013} \\ 
+DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
+Energy\acs{nDCG} & \cite{Freno2011} & \cite{Freno2011} & RankDE &  & \cite{Sato2013} \\ 
+FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b,Laporte2013} & RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FocusedBoost & \cite{Niu2012} & \cite{Niu2012} & RankMGP & \cite{Lin2012} & \cite{Lin2012} \\ 
+FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankNet &  & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
+Focused\acs{SVM} & \cite{Niu2012} & \cite{Niu2012} & Rank\acs{RLS} &  &  \\ 
+FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{SVM} &  & \cite{Busa-Fekete2013, Freno2011} \\ 
+FRank &  & \cite{Wang2012} & Rank\acs{SVM}-Primal &  &  \\ 
+FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & Rank\acs{SVM}-Struct &  &  \\
+FSM$^{\acs{SVM}}$ & & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
+GAS-E &  & \cite{Lai2013c} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\ 
+\acs{GP}Rank &  & \cite{Torkestani2012} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\ 
+GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & Ridge Regression &  &  \\ 
+Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & RSRank &  &  \\ 
+Group\acs{MLE} &  &  & SmoothGrad &  & \cite{Tan2013} \\ 
+IntervalRank & \cite{Moon2010} & \cite{Moon2010} & SmoothRank & \cite{Chapelle2010} & \cite{Chapelle2010} \\ 
+\acs{IP}Rank &  & \cite{Torkestani2012} & SoftRank &  & \cite{Qin2010b} \\ 
+KeepRank & \cite{Chen2009} & \cite{Chen2009} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\ 
+Kernel-\acs{PCA} &  &  & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
+KL-\acs{CRF} & \cite{Volkovs2011} &  & \acs{SVD}-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\ 
+LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & \acs{SVM}$^{\acs{MAP}}$ &  & \cite{Wang2012} \\ 
+LambdaMART & \cite{Burges2010} & \cite{Asadi2013a} & SwarmRank &  & \cite{Sato2013} \\ 
+LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank &  &  \\ 
+LambdaRank &  &  & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012,Tan2013} \\ 
+LARF & \cite{Torkestani2012} & \cite{Torkestani2012} & VFLR & \cite{Cai2012} & \cite{Cai2012} \\ 
+\end{tabular}}
 \caption{Learning-to-Rank algorithms with measurements on benchmark datasets}

### 2014-04-06 report/Chapters/CrossBenchmarkComparison.tex
LETOR 2.0 metingen IPRank gevonden + begin overzicht unusable papers

@@
 AdaRank-\acs{MAP} &  &  & Linear Regression &  & \cite{Wang2012} \\ 
-AdaRank-\acs{nDCG} &  & \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} &  & \cite{Lin2010, Gao2014} \\ 
+AdaRank-\acs{nDCG} &  & \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} &  & \cite{Lin2010, Lin2011, Gao2014} \\ 
 ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet &  &  \\ 
@@
 GAS-E &  & \cite{Lai2013c} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\ 
-\acs{GP}Rank &  & \cite{Torkestani2012} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\ 
+\acs{GP}Rank & \cite{Silva2009} & \cite{Torkestani2012} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\ 
 GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & Ridge Regression &  &  \\ 
 Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & RSRank &  &  \\ 
-Group\acs{MLE} &  &  & SmoothGrad &  & \cite{Tan2013} \\ 
+Group\acs{MLE} & \cite{Lin2010} & \cite{Lin2011} & SmoothGrad &  & \cite{Tan2013} \\ 
 IntervalRank & \cite{Moon2010} & \cite{Moon2010} & SmoothRank & \cite{Chapelle2010} & \cite{Chapelle2010} \\ 
-\acs{IP}Rank &  & \cite{Torkestani2012} & SoftRank &  & \cite{Qin2010b} \\ 
+\acs{IP}Rank & \cite{Wang2009b} & \cite{Torkestani2012} & SoftRank &  & \cite{Qin2010b} \\ 
 KeepRank & \cite{Chen2009} & \cite{Chen2009} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\ 
-Kernel-\acs{PCA} &  &  & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
+Kernel-\acs{PCA} RankBoost & \cite{Duh2008} & \cite{Duh2008} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
 KL-\acs{CRF} & \cite{Volkovs2011} &  & \acs{SVD}-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\ 

### 2014-04-06 report/Chapters/CrossBenchmarkComparison.tex
referenties van alle result-files verwerkt

@@
 \hline 
-AdaRank-\acs{MAP} &  &  & Linear Regression &  & \cite{Wang2012} \\ 
+AdaRank-\acs{MAP} &  &  & Linear Regression &  & \cite{Wang2012, Volkovs2011} \\ 
 AdaRank-\acs{nDCG} &  & \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} &  & \cite{Lin2010, Lin2011, Gao2014} \\ 
@@
 Approx\acs{nDCG} & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
-BagBoo &  &  & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
+BagBoo &  & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
 Best Single Feature &  & \cite{Gomes2013} & MHR &  &  \\ 
@@
 BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009} & NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
-BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
+BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
 BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-\acs{MAP} &  & \cite{Torkestani2012b} \\ 
 C-\acs{CRF} & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
-CA &  & \cite{Busa-Fekete2013,Tan2013} & Q.D.\acs{KNN} &  &  \\ 
+CA &  & \cite{Busa-Fekete2013,Tan2013} & Q.D.\acs{KNN} &  & \cite{Wang2013} \\ 
 CCRank &  &  & RandomForest &  & \cite{Gomes2013} \\ 
 CoList & \cite{Gao2014} & \cite{Gao2014} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
-Consistent-RankCosine &  & \cite{Tan2013} & RankAgg\acs{nDCG} &  &  \\ 
+Consistent-RankCosine &  & \cite{Tan2013} & RankAgg\acs{nDCG} & \cite{Wang2013} & \cite{Wang2013} \\ 
 DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost &  & \cite{Busa-Fekete2013} \\ 
@@
 FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
-FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b,Laporte2013} & RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b, Laporte2013} & RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
 FocusedBoost & \cite{Niu2012} & \cite{Niu2012} & RankMGP & \cite{Lin2012} & \cite{Lin2012} \\ 
@@
 Focused\acs{SVM} & \cite{Niu2012} & \cite{Niu2012} & Rank\acs{RLS} &  &  \\ 
-FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{SVM} &  & \cite{Busa-Fekete2013, Freno2011} \\ 
-FRank &  & \cite{Wang2012} & Rank\acs{SVM}-Primal &  &  \\ 
+FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{SVM} &  & \cite{Busa-Fekete2013, Freno2011, He2010} \\ 
+FRank &  & \cite{Wang2012} & Rank\acs{SVM}-Primal &  & \cite{Lai2011} \\ 
 FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & Rank\acs{SVM}-Struct &  &  \\
@@
 GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & Ridge Regression &  &  \\ 
-Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & RSRank &  &  \\ 
+Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & RSRank &  & \cite{Lai2013} \\ 
 Group\acs{MLE} & \cite{Lin2010} & \cite{Lin2011} & SmoothGrad &  & \cite{Tan2013} \\ 
-IntervalRank & \cite{Moon2010} & \cite{Moon2010} & SmoothRank & \cite{Chapelle2010} & \cite{Chapelle2010} \\ 
-\acs{IP}Rank & \cite{Wang2009b} & \cite{Torkestani2012} & SoftRank &  & \cite{Qin2010b} \\ 
+IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SmoothRank & \cite{Chapelle2010} & \cite{Chapelle2010} \\ 
+\acs{IP}Rank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SoftRank &  & \cite{Qin2010b} \\ 
 KeepRank & \cite{Chen2009} & \cite{Chen2009} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\ 
-Kernel-\acs{PCA} RankBoost & \cite{Duh2008} & \cite{Duh2008} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
-KL-\acs{CRF} & \cite{Volkovs2011} &  & \acs{SVD}-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\ 
-LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & \acs{SVM}$^{\acs{MAP}}$ &  & \cite{Wang2012} \\ 
-LambdaMART & \cite{Burges2010} & \cite{Asadi2013a} & SwarmRank &  & \cite{Sato2013} \\ 
-LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank &  &  \\ 
-LambdaRank &  &  & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012,Tan2013} \\ 
+Kernel-\acs{PCA} RankBoost & \cite{Duh2008} & \cite{Duh2008, Sato2013} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
+KL-\acs{CRF} & \cite{Volkovs2011} & \cite{Volkovs2011} & \acs{SVD}-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\ 
+LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & \acs{SVM}$^{\acs{MAP}}$ &  & \cite{Wang2012, Xu2008, Niu2012} \\ 
+LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} & SwarmRank &  & \cite{Sato2013} \\ 
+LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank &  & \cite{Lai2013} \\ 
+LambdaRank &  &  & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
 LARF & \cite{Torkestani2012} & \cite{Torkestani2012} & VFLR & \cite{Cai2012} & \cite{Cai2012} \\ 

### 2014-04-06 report/Chapters/CrossBenchmarkComparison.tex
begin werk aan 'proposed by'-references

@@
 \hline 
-AdaRank-\acs{MAP} &  &  & Linear Regression &  & \cite{Wang2012, Volkovs2011} \\ 
-AdaRank-\acs{nDCG} &  & \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} &  & \cite{Lin2010, Lin2011, Gao2014} \\ 
+AdaRank-\acs{MAP} & \cite{Xu2007} &  & Linear Regression &  & \cite{Wang2012, Volkovs2011} \\ 
+AdaRank-\acs{nDCG} & \cite{Xu2007} & \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} &  & \cite{Lin2010, Lin2011, Gao2014} \\ 
 ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet &  &  \\ 
@@
 Approx\acs{nDCG} & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
-BagBoo &  & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
+BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
 Best Single Feature &  & \cite{Gomes2013} & MHR &  &  \\ 
@@
 C-\acs{CRF} & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
-CA &  & \cite{Busa-Fekete2013,Tan2013} & Q.D.\acs{KNN} &  & \cite{Wang2013} \\ 
-CCRank &  &  & RandomForest &  & \cite{Gomes2013} \\ 
+CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & Q.D.\acs{KNN} &  & \cite{Wang2013} \\ 
+CCRank & \cite{Wang2011c} & \cite{Wang2011c} & RandomForest &  & \cite{Gomes2013} \\ 
 CoList & \cite{Gao2014} & \cite{Gao2014} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 

### 2014-04-06 report/Chapters/CrossBenchmarkComparison.tex
verder werk aan 'proposed by'-referenties

@@
 AdaRank-\acs{MAP} & \cite{Xu2007} &  & Linear Regression &  & \cite{Wang2012, Volkovs2011} \\ 
-AdaRank-\acs{nDCG} & \cite{Xu2007} & \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} &  & \cite{Lin2010, Lin2011, Gao2014} \\ 
-ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet &  &  \\ 
+AdaRank-\acs{nDCG} & \cite{Xu2007} & \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
+ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet & \cite{Cao2007} &  \\ 
 Approx\acs{AP} & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
@@
 BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
-Best Single Feature &  & \cite{Gomes2013} & MHR &  &  \\ 
+Best Single Feature &  & \cite{Gomes2013} & MHR & \cite{Qin2007} &  \\ 
 BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
@@
 BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
-BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-\acs{MAP} &  & \cite{Torkestani2012b} \\ 
+BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-\acs{MAP} & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
 C-\acs{CRF} & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
-CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & Q.D.\acs{KNN} &  & \cite{Wang2013} \\ 
+CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & Q.D.\acs{KNN} & \cite{Geng2008} & \cite{Wang2013} \\ 
 CCRank & \cite{Wang2011c} & \cite{Wang2011c} & RandomForest &  & \cite{Gomes2013} \\ 
 CoList & \cite{Gao2014} & \cite{Gao2014} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
-Consistent-RankCosine &  & \cite{Tan2013} & RankAgg\acs{nDCG} & \cite{Wang2013} & \cite{Wang2013} \\ 
+Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAgg\acs{nDCG} & \cite{Wang2013} & \cite{Wang2013} \\ 
 DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost &  & \cite{Busa-Fekete2013} \\ 
@@
 FocusedBoost & \cite{Niu2012} & \cite{Niu2012} & RankMGP & \cite{Lin2012} & \cite{Lin2012} \\ 
-FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankNet &  & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
+FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
 Focused\acs{SVM} & \cite{Niu2012} & \cite{Niu2012} & Rank\acs{RLS} &  &  \\ 
-FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{SVM} &  & \cite{Busa-Fekete2013, Freno2011, He2010} \\ 
-FRank &  & \cite{Wang2012} & Rank\acs{SVM}-Primal &  & \cite{Lai2011} \\ 
+FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{SVM} & \cite{Herbrich1999, Joachims2002} & \cite{Busa-Fekete2013, Freno2011, He2010} \\ 
+FRank & \cite{Tsai2007} & \cite{Wang2012} & Rank\acs{SVM}-Primal &  & \cite{Lai2011} \\ 
 FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & Rank\acs{SVM}-Struct &  &  \\
-FSM$^{\acs{SVM}}$ & & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
-GAS-E &  & \cite{Lai2013c} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\ 
+FSM$^{\acs{SVM}}$ & \cite{Lai2013c} & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
+GAS-E & \cite{Geng2007} & \cite{Lai2013c} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\ 
 \acs{GP}Rank & \cite{Silva2009} & \cite{Torkestani2012} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\ 
 GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & Ridge Regression &  &  \\ 
-Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & RSRank &  & \cite{Lai2013} \\ 
-Group\acs{MLE} & \cite{Lin2010} & \cite{Lin2011} & SmoothGrad &  & \cite{Tan2013} \\ 
+Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
+Group\acs{MLE} & \cite{Lin2010} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
 IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SmoothRank & \cite{Chapelle2010} & \cite{Chapelle2010} \\ 
-\acs{IP}Rank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SoftRank &  & \cite{Qin2010b} \\ 
+\acs{IP}Rank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
 KeepRank & \cite{Chen2009} & \cite{Chen2009} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\ 
@@
 LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank &  & \cite{Lai2013} \\ 
-LambdaRank &  &  & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
+LambdaRank & \cite{Burges2006} &  & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
 LARF & \cite{Torkestani2012} & \cite{Torkestani2012} & VFLR & \cite{Cai2012} & \cite{Cai2012} \\ 

### 2014-04-07 report/Chapters/CrossBenchmarkComparison.tex
linear regression reference
method references geplaatst

@@
 \hline 
-AdaRank-\acs{MAP} & \cite{Xu2007} &  & Linear Regression &  & \cite{Wang2012, Volkovs2011} \\ 
+AdaRank-\acs{MAP} & \cite{Xu2007} &  & Linear Regression & \cite{Cossock2006} & \cite{Wang2012, Volkovs2011} \\ 
 AdaRank-\acs{nDCG} & \cite{Xu2007} & \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
@@
 Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAgg\acs{nDCG} & \cite{Wang2013} & \cite{Wang2013} \\ 
-DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost &  & \cite{Busa-Fekete2013} \\ 
+DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost & \cite{Freund2003} & \cite{Busa-Fekete2013} \\ 
 DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
-Energy\acs{nDCG} & \cite{Freno2011} & \cite{Freno2011} & RankDE &  & \cite{Sato2013} \\ 
+Energy\acs{nDCG} & \cite{Freno2011} & \cite{Freno2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\ 
 FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
@@
 FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
-Focused\acs{SVM} & \cite{Niu2012} & \cite{Niu2012} & Rank\acs{RLS} &  &  \\ 
+Focused\acs{SVM} & \cite{Niu2012} & \cite{Niu2012} & Rank\acs{RLS} & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
 FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{SVM} & \cite{Herbrich1999, Joachims2002} & \cite{Busa-Fekete2013, Freno2011, He2010} \\ 
@@
 \acs{GP}Rank & \cite{Silva2009} & \cite{Torkestani2012} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\ 
-GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & Ridge Regression &  &  \\ 
+GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & Ridge Regression & \cite{Cossock2006} &  \\ 
 Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
@@
 KL-\acs{CRF} & \cite{Volkovs2011} & \cite{Volkovs2011} & \acs{SVD}-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\ 
-LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & \acs{SVM}$^{\acs{MAP}}$ &  & \cite{Wang2012, Xu2008, Niu2012} \\ 
-LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} & SwarmRank &  & \cite{Sato2013} \\ 
-LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank &  & \cite{Lai2013} \\ 
+LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & \acs{SVM}$^{\acs{MAP}}$ & \cite{Yue2007} & \cite{Wang2012, Xu2008, Niu2012} \\ 
+LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} & SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
+LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank & \cite{Lai2013} & \cite{Lai2013} \\ 
 LambdaRank & \cite{Burges2006} &  & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 

### 2014-04-07 report/Chapters/CrossBenchmarkComparison.tex
twee smaken Genetic Programming voor ranking!

@@
 Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAgg\acs{nDCG} & \cite{Wang2013} & \cite{Wang2013} \\ 
-DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost & \cite{Freund2003} & \cite{Busa-Fekete2013} \\ 
+DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost & \cite{Freund2003} & \cite{Busa-Fekete2013, Alcantara2010} \\ 
 DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
@@
 Focused\acs{SVM} & \cite{Niu2012} & \cite{Niu2012} & Rank\acs{RLS} & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
-FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{SVM} & \cite{Herbrich1999, Joachims2002} & \cite{Busa-Fekete2013, Freno2011, He2010} \\ 
+FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{SVM} & \cite{Herbrich1999, Joachims2002} & \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010} \\ 
 FRank & \cite{Tsai2007} & \cite{Wang2012} & Rank\acs{SVM}-Primal &  & \cite{Lai2011} \\ 
@@
 GAS-E & \cite{Geng2007} & \cite{Lai2013c} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\ 
-\acs{GP}Rank & \cite{Silva2009} & \cite{Torkestani2012} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\ 
+\acs{GP}Rank & \cite{Silva2009} & \cite{Torkestani2012, Alcantara2010} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\ 
 GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & Ridge Regression & \cite{Cossock2006} &  \\ 

### 2014-04-07 report/Chapters/CrossBenchmarkComparison.tex
GP uitgesplitst in tabel
tabel weer helemaal mooi :)

@@
 GAS-E & \cite{Geng2007} & \cite{Lai2013c} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\
-\acs{GP}Rank & \cite{Silva2009} & \cite{Torkestani2012, Alcantara2010} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\ 
-GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & Ridge Regression & \cite{Cossock2006} &  \\ 
-Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
-Group\acs{MLE} & \cite{Lin2010} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
-IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SmoothRank & \cite{Chapelle2010} & \cite{Chapelle2010} \\ 
-\acs{IP}Rank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
-KeepRank & \cite{Chen2009} & \cite{Chen2009} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\ 
-Kernel-\acs{PCA} RankBoost & \cite{Duh2008} & \cite{Duh2008, Sato2013} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
-KL-\acs{CRF} & \cite{Volkovs2011} & \cite{Volkovs2011} & \acs{SVD}-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\ 
-LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & \acs{SVM}$^{\acs{MAP}}$ & \cite{Yue2007} & \cite{Wang2012, Xu2008, Niu2012} \\ 
-LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} & SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
-LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank & \cite{Lai2013} & \cite{Lai2013} \\ 
-LambdaRank & \cite{Burges2006} &  & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
-LARF & \cite{Torkestani2012} & \cite{Torkestani2012} & VFLR & \cite{Cai2012} & \cite{Cai2012} \\ 
+\acs{GP} & \cite{DeAlmeida2007} & \cite{Alcantara2010} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\  
+\acs{GP}Rank & \cite{Silva2009} & \cite{Torkestani2012} & Ridge Regression & \cite{Cossock2006} &  \\
+GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
+Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
+Group\acs{MLE} & \cite{Lin2010} & \cite{Lin2011} & SmoothRank & \cite{Chapelle2010} & \cite{Chapelle2010} \\
+IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
+\acs{IP}Rank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\
+KeepRank & \cite{Chen2009} & \cite{Chen2009} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
+Kernel-\acs{PCA} RankBoost & \cite{Duh2008} & \cite{Duh2008, Sato2013} & \acs{SVD}-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\
+KL-\acs{CRF} & \cite{Volkovs2011} & \cite{Volkovs2011} & \acs{SVM}$^{\acs{MAP}}$ & \cite{Yue2007} & \cite{Wang2012, Xu2008, Niu2012} \\ 
+LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
+LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} & TGRank & \cite{Lai2013} & \cite{Lai2013} \\ 
+LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
+LambdaRank & \cite{Burges2006} &  & VFLR & \cite{Cai2012} & \cite{Cai2012} \\ 
+LARF & \cite{Torkestani2012} & \cite{Torkestani2012} &  &  &  \\ 
 \end{tabular}}

### 2014-04-07 report/Chapters/CrossBenchmarkComparison.tex
veel repetitief werk gedaan :(

@@
 \section{Collecting Evaluation Results}
+\label{sec:collecting_evaluation_results}
 With a literature review I will collect evaluation results on the datasets/collections. The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
@@
 \section{Evaluation Results Found in Literature}
-\begin{table}[!h]
-\scalebox{0.85}{
+Table~\ref{tab:ltr_methods_used} gives an overview of the Learning-to-Rank methods for which evaluation results were found for one of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.\\
+
+Some studies with evaluation results found through in literature review were not usable for the meta-analysis because they 1) used a different evaluation methodology than other studies using the same benchmark \cite{Geng2011, Lin2012}, 2) provided results for different tasks (e.g.
+rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, Pan2011b, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}, 3) added additional features to a benchmark dataset \cite{Bidoki2009, Ding2010}, 4) did not provide exact data for evaluation results (often only in graph form) (), 5) reported evaluation results in a different metric than the metrics chosen for this meta-analysis (), 6) reported a higher performance on baseline methods than official runs (), or 7) did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
+\begin{table}[!h!p]
+\scalebox{0.80}{
 \begin{tabular}{|l|l|l||l|l|l|}
@@
 \caption{Learning-to-Rank algorithms with measurements on benchmark datasets}
-\label{ltr_methods_used}
+\label{tab:ltr_methods_used}
 \end{table}

### 2014-04-07 method_selection/winning_number.R

@@ -130,7 +130,7 @@ combined_output <- combined_output[, c(ncol(combined_output), 1:(ncol(combined_o
 #  expand_limits(x = c(0,16.35), y = c(0,1.03))
 
 ggplot(combined_aggregate, aes(Summed_ideal_winnum,Norm_winnum)) + geom_point(size=2) +  
-  geom_text(aes(label=Method)) +
+#  geom_text(aes(label=Method)) +
   scale_x_continuous("Ideal Winning Number", expand = c(0,0)) + 
   scale_y_continuous("Normalized Winning Number", expand=c(0,0)) +
   expand_limits(x = c(0,1000), y = c(0,1.01)) +

### 2014-04-07 report/Chapters/CrossBenchmarkComparison.tex
holy shit wat een feutenwerk!

@@
 Some studies with evaluation results found through in literature review were not usable for the meta-analysis because they 1) used a different evaluation methodology than other studies using the same benchmark \cite{Geng2011, Lin2012}, 2) provided results for different tasks (e.g.
-rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, Pan2011b, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}, 3) added additional features to a benchmark dataset \cite{Bidoki2009, Ding2010}, 4) did not provide exact data for evaluation results (often only in graph form) (), 5) reported evaluation results in a different metric than the metrics chosen for this meta-analysis (), 6) reported a higher performance on baseline methods than official runs (), or 7) did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
+rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, Pan2011b, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}, 3) added additional features to a benchmark dataset \cite{Bidoki2009, Ding2010}, 4) did not provide exact data for evaluation results (often only in graph form) \cite{Wang2008, Wang2010, Xu2010, Kuo2009, Li2008, Xia2008, Zhou2011, Wu2011, Zhu2009, Karimzadehgan2011, Swersky2012, Pan2011, Ni2008, Ciaramita2008, Stewart2012, Petterson2009, Agarwal2010, Chang2009, Qin2008c, Adams2011, Sculley2009, Huang2008, Alejo2010, Sun2011, He2010b, Benbouzid2012, Geng2012, Chen2012, Xu2012, Shivaswamy2011}, 5) reported evaluation results in a different metric than the metrics chosen for this meta-analysis (), 6) reported a higher performance on baseline methods than official runs (), or 7) did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
 \begin{table}[!h!p]

### 2014-04-07 report/Chapters/CrossBenchmarkComparison.tex
glsdefs in gitignore

--- 2014-04-07 report/Chapters/CrossBenchmarkComparison.tex
@@
 Some studies with evaluation results found through in literature review were not usable for the meta-analysis because they 1) used a different evaluation methodology than other studies using the same benchmark \cite{Geng2011, Lin2012}, 2) provided results for different tasks (e.g.
-rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, Pan2011b, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}, 3) added additional features to a benchmark dataset \cite{Bidoki2009, Ding2010}, 4) did not provide exact data for evaluation results (often only in graph form) (), 5) reported evaluation results in a different metric than the metrics chosen for this meta-analysis (), 6) reported a higher performance on baseline methods than official runs (), or 7) did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
+rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}, 3) added additional features to a benchmark dataset \cite{Bidoki2009, Ding2010}, 4) did not provide exact data for evaluation results (often only in graph form) \cite{Wang2008, Wang2010, Xu2010, Kuo2009, Li2008, Xia2008, Zhou2011, Wu2011, Zhu2009, Karimzadehgan2011, Swersky2012, Pan2011, Ni2008, Ciaramita2008, Stewart2012, Petterson2009, Agarwal2010, Chang2009, Qin2008c, Adams2011, Sculley2009, Huang2008, Alejo2010, Sun2011, He2010b, Benbouzid2012, Geng2012, Chen2012, Xu2012, Shivaswamy2011}, 5) reported evaluation results in a different metric than the metrics chosen for this meta-analysis \cite{Yu2009, Thuy2009, Pahikkala2009, Kersting2009, Mohan2011}, 6) reported a higher performance on baseline methods than official runs \cite{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}, or 7) did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
 \begin{table}[!h!p]

### 2014-04-07 report/Chapters/CrossBenchmarkComparison.tex
opmaak enumaration

--- 2014-04-07 report/Chapters/CrossBenchmarkComparison.tex
@@
 
-Some studies with evaluation results found through in literature review were not usable for the meta-analysis because they 1) used a different evaluation methodology than other studies using the same benchmark \cite{Geng2011, Lin2012}, 2) provided results for different tasks (e.g.
-rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, Pan2011b, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}, 3) added additional features to a benchmark dataset \cite{Bidoki2009, Ding2010}, 4) did not provide exact data for evaluation results (often only in graph form) (), 5) reported evaluation results in a different metric than the metrics chosen for this meta-analysis (), 6) reported a higher performance on baseline methods than official runs (), or 7) did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
+Some studies with evaluation results found through in literature review were not usable for the meta-analysis because one of the following properties made them unsuited for the meta-analysis:
+
+\begin{enumerate}
+\item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011, Lin2012}
+\item The study focussed on a different Learning-to-Rank task (e.g.
+rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}
+\item The study used an altered version of a benchmark that contained additional features \cite{Bidoki2009, Ding2010}
+\item The study provides no exact data of the evaluation results (e.g.
+results are only in graphical form) \cite{Wang2008, Wang2010, Xu2010, Kuo2009, Li2008, Xia2008, Zhou2011, Wu2011, Zhu2009, Karimzadehgan2011, Swersky2012, Pan2011, Ni2008, Ciaramita2008, Stewart2012, Petterson2009, Agarwal2010, Chang2009, Qin2008c, Adams2011, Sculley2009, Huang2008, Alejo2010, Sun2011, He2010b, Benbouzid2012, Geng2012, Chen2012, Xu2012, Shivaswamy2011}
+\item The study reported evaluation results in a different metric than the metrics chosen for this meta-analysis \cite{Yu2009, Thuy2009, Pahikkala2009, Kersting2009, Mohan2011}
+\item The study reported a higher performance on baseline methods than official benchmark runs \cite{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}
+\item The study did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
+\end{enumerate}
+
 \begin{table}[!h!p]

### 2014-04-07 method_selection/winning_number.R
nieuwe export graphic

@@ -130,7 +130,7 @@ combined_output <- combined_output[, c(ncol(combined_output), 1:(ncol(combined_o
 #  expand_limits(x = c(0,16.35), y = c(0,1.03))
 
 ggplot(combined_aggregate, aes(Summed_ideal_winnum,Norm_winnum)) + geom_point(size=2) +  
-#  geom_text(aes(label=Method)) +
+  geom_text(aes(label=Method)) +
   scale_x_continuous("Ideal Winning Number", expand = c(0,0)) + 
   scale_y_continuous("Normalized Winning Number", expand=c(0,0)) +
   expand_limits(x = c(0,1000), y = c(0,1.01)) +

### 2014-04-07 report/Chapters/CrossBenchmarkComparison.tex
nieuwe export graphic
raw data in appendix

@@
 \subsection{All metrics}
+Figure \ref{fig:normalised_winning_number_all} shows the normalised winning number as function of ideal winning number for the methods described in Table \ref{tab:ltr_methods_used}.
+Table \ref{tab:raw_data_norm_winnum} and Table \ref{tab:raw_data_norm_winnum_cntd} in Appendix \ref{app:norm_winnum_all} provide the raw data plotted in Figure \ref{fig:normalised_winning_number_all}.
 \begin{figure}[!h]
-\includegraphics[scale=0.40]{gfx/combined_normalized_winnum}
+\includegraphics[scale=0.35]{gfx/combined_normalized_winnum}
 \caption{Cross-benchmark comparison of Learning-to-Rank methods}

### 2014-04-07 method_selection/winning_number.R
ndcg5 en 10 winnum plots

@@ -122,17 +122,17 @@ combined_output <- combined_output[, c(ncol(combined_output), 1:(ncol(combined_o
 ### PLOT ###
 ### ADJUST HERE ###
 # plot graph
-#ggplot(ndcg5_output, aes(Nr_measurements,Normalised_winning_number)) + geom_point(size=2) + 
-#  geom_text(aes(label=id), hjust=-0.5, vjust=-0.5) + 
-#  theme_grey(base_size=12, base_family="") + 
-#  scale_x_continuous("# of datasets evaluated on", expand = c(0,0), breaks=c(0,2,4,6,8,10,12,14, 16)) +
-#  scale_y_continuous("Normalised Winning Number", expand = c(0,0), breaks=c(0,0.2,0.4,0.6,0.8,1.0)) +
-#  expand_limits(x = c(0,16.35), y = c(0,1.03))
+ggplot(ndcg10_output, aes(Nr_measurements,Normalised_winning_number)) + geom_point(size=2) + 
+  geom_text(aes(label=Method), hjust=-0.1, vjust=-0.1) + 
+  theme_grey(base_size=12, base_family="") + 
+  scale_x_continuous("# of datasets evaluated on", expand = c(0,0), breaks=c(0,2,4,6,8,10,12,14, 16)) +
+  scale_y_continuous("Normalised Winning Number", expand = c(0,0), breaks=c(0,0.2,0.4,0.6,0.8,1.0)) +
+  expand_limits(x = c(0,18.03), y = c(0,1.03))
 
-ggplot(combined_aggregate, aes(Summed_ideal_winnum,Norm_winnum)) + geom_point(size=2) +  
-  geom_text(aes(label=Method)) +
-  scale_x_continuous("Ideal Winning Number", expand = c(0,0)) + 
-  scale_y_continuous("Normalized Winning Number", expand=c(0,0)) +
-  expand_limits(x = c(0,1000), y = c(0,1.01)) +
-  theme(axis.title.y = element_text(size = rel(1.8), angle = 90)) +
-  theme(axis.title.x = element_text(size = rel(1.8), angle = 00))
+#ggplot(combined_aggregate, aes(Summed_ideal_winnum,Norm_winnum)) + geom_point(size=2) +  
+#  geom_text(aes(label=Method)) +
+#  scale_x_continuous("Ideal Winning Number", expand = c(0,0)) + 
+#  scale_y_continuous("Normalized Winning Number", expand=c(0,0)) +
[…]

### 2014-04-07 method_selection/winning_number.R
plots ge-exporteerd, ge-annoteerd en ge-integreerd

@@ -122,12 +122,12 @@ combined_output <- combined_output[, c(ncol(combined_output), 1:(ncol(combined_o
 ### PLOT ###
 ### ADJUST HERE ###
 # plot graph
-ggplot(ndcg10_output, aes(Nr_measurements,Normalised_winning_number)) + geom_point(size=2) + 
+ggplot(map_output, aes(Nr_measurements,Normalised_winning_number)) + geom_point(size=2) + 
   geom_text(aes(label=Method), hjust=-0.1, vjust=-0.1) + 
   theme_grey(base_size=12, base_family="") + 
   scale_x_continuous("# of datasets evaluated on", expand = c(0,0), breaks=c(0,2,4,6,8,10,12,14, 16)) +
   scale_y_continuous("Normalised Winning Number", expand = c(0,0), breaks=c(0,0.2,0.4,0.6,0.8,1.0)) +
-  expand_limits(x = c(0,18.03), y = c(0,1.03))
+  expand_limits(x = c(0,15.03), y = c(0,1.03))
 
 #ggplot(combined_aggregate, aes(Summed_ideal_winnum,Norm_winnum)) + geom_point(size=2) +  
 #  geom_text(aes(label=Method)) +

### 2014-04-08 report/Chapters/CrossBenchmarkComparison.tex
plots ge-exporteerd, ge-annoteerd en ge-integreerd
TODO gevinkt: uitleg gebruikte metrics

@@
 
+\acs{nDCG}@3, \acs{nDCG}@5, \acs{nDCG}@10 and \acs{MAP} are chosen as metrics on which the meta-analysis will be performed.
+These metrics seem to be the most frequently used evaluation metrics in most of the benchmark datasets used.
+An exception are the Yahoo!
+Learning to Rank Challenge datasets on which mainly \ac{ERR} is used as main evaluation metric.
+The lack of use of \ac{ERR} in other benchmarks makes it unsuitable for a cross-benchmark comparison.
+
 \section{Evaluation Results Found in Literature}
@@
 \subsection{MAP}
+\begin{figure}[!h]
+\includegraphics[scale=0.35]{gfx/map_winnum}
+\caption{\acs{MAP} comparison of Learning-to-Rank methods}
+\label{fig:normalised_winning_number_map}
+\end{figure}
 
 \subsection{NDCG@3}
+\begin{figure}[!h]
+\includegraphics[scale=0.35]{gfx/ndcg3_winnum}
+\caption{\acs{nDCG}@3 comparison of Learning-to-Rank methods}
+\label{fig:normalised_winning_number_ndcg3}
+\end{figure}
 
 \subsection{NDCG@5}
+\begin{figure}[!h]
+\includegraphics[scale=0.35]{gfx/ndcg5_winnum}
+\caption{\acs{nDCG}@5 comparison of Learning-to-Rank methods}
+\label{fig:normalised_winning_number_ndcg5}
+\end{figure}
 
 \subsection{NDCG@10}
+\begin{figure}[!h]
+\includegraphics[scale=0.35]{gfx/ndcg10_winnum}
+\caption{\acs{nDCG}@10 comparison of Learning-to-Rank methods}
+\label{fig:normalised_winning_number_ndcg10}
+\end{figure}
 
@@
 \end{figure}
-%TODO: waarom ndcg@3,5,10 en MAP?
-%TODO: limitations van deze vergelijkingsmethodologie
-
-\section{Limitations}

### 2014-04-08 report/Chapters/CrossBenchmarkComparison.tex
boltzrank-metingen gemist :(. opzetje raw data appendices

@@
 In contrast to the winning number comparison on LETOR 3.0, there will not be accuracy measurements for each algorithm on each dataset in our meta-analysis.
-To compare algorithms based on a sparse set of evaluation measurements, a normalised version of the winning number metric will be used which we calculate by only taking those datasets into account that an algorithm is evaluated on and then dividing this by the ideal winning number that an algorithm could have had if it would have been the best algorithm on all datasets on which it has been evaluated.
+To compare algorithms based on a sparse set of evaluation measurements, a normalised version of the Winning Number metric will be used which we calculate by only taking those datasets into account that an algorithm is evaluated on and then dividing this by the ideal Winning Number that an algorithm could have had if it would have been the best algorithm on all datasets on which it has been evaluated.
 I will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as \\
@@
 
-From now on this adjusted version of winning number will be references to as \emph{normalised winning number}. The mathematical definition of normalised winning number is\\
+From now on this adjusted version of Winning Number will be references to as \emph{Normalised Winning Number}. The mathematical definition of Normalised Winning Number is\\
 
@@
 \noindent
-where ideal winning number is defined as\\
+where Ideal Winning Number is defined as\\
 
@@
 Learning to Rank Challenge datasets on which mainly \ac{ERR} is used as main evaluation metric.
-The lack of use of \ac{ERR} in other benchmarks makes it unsuitable for a cross-benchmark comparison.
+The lack of use of the \ac{ERR}-metric in other benchmarks makes it unsuitable for a cross-benchmark comparison.
 
@@
 
-Some studies with evaluation results found through in literature review were not usable for the meta-analysis because one of the following properties made them unsuited for the meta-analysis:
+Some studies with evaluation results found through in literature review were not usable for the meta-analysis.
+The following enumeration lists the properties that made those studies unusable for the meta-analysis, with between brackets the studies that these properties apply to.
 
@@
 BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
-BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009} & NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
-BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
+BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009, Volkovs2013} & NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
+BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c, Volkovs2013} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
 BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-\acs{MAP} & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
@@
 \section{Results \& Discussion}
-\subsection{MAP}
+The following subsections provide the performance of Learning-to-Rank methods in terms of Normalised Winning Number for \ac{nDCG}@3, \ac{nDCG}@5, \ac{nDCG}@10 and \ac{MAP}.
+Performance of the Learning-to-Rank methods is plotted with Normalised Winning Number on the vertical axis and the number of datasets on which the method has been evaluated on the horizontal axis.
+The further to the right, the more certain we can be about the performance of the Learning-to-Rank method.
+The methods for which it holds that there is no other method that has 1) a higher Normalised Winning Number and 2) a higher number datasets evaluated, are identified as the best performing methods and are labelled with the name of the method.
+\subsection{NDCG@3}
+Figure \ref{fig:normalised_winning_number_ndcg3} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@3 metric.
 \begin{figure}[!h]
-\includegraphics[scale=0.35]{gfx/map_winnum}
-\caption{\acs{MAP} comparison of Learning-to-Rank methods}
-\label{fig:normalised_winning_number_map}
-\end{figure}
-
-\subsection{NDCG@3}
-\begin{figure}[!h]
-\includegraphics[scale=0.35]{gfx/ndcg3_winnum}
+\includegraphics[scale=0.40]{gfx/ndcg3_winnum}
 \caption{\acs{nDCG}@3 comparison of Learning-to-Rank methods}
@@
 \subsection{NDCG@5}
+Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@5 metric. Raw data for the comparison
 \begin{figure}[!h]
@@
 \subsection{NDCG@10}
+Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@10 metric.
 \begin{figure}[!h]
@@
 
-\subsection{All metrics}
-Figure \ref{fig:normalised_winning_number_all} shows the normalised winning number as function of ideal winning number for the methods described in Table \ref{tab:ltr_methods_used}.
+\subsection{MAP}
+Figure \ref{fig:normalised_winning_number_map} shows the performance of Learning-to-Rank methods for the \ac{MAP} metric.
+\begin{figure}[!h]
+\includegraphics[scale=0.35]{gfx/map_winnum}
+\caption{\acs{MAP} comparison of Learning-to-Rank methods}
+\label{fig:normalised_winning_number_map}
+\end{figure}
+
+\subsection{Cross-metric}
+Figure \ref{fig:normalised_winning_number_all} shows the Normalised Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}.
 Table \ref{tab:raw_data_norm_winnum} and Table \ref{tab:raw_data_norm_winnum_cntd} in Appendix \ref{app:norm_winnum_all} provide the raw data plotted in Figure \ref{fig:normalised_winning_number_all}.

### 2014-04-08 report/Chapters/CrossBenchmarkComparison.tex
tex lockfile in gitignore

@@
 \end{figure}
+
+\section{Limitations}
+In the Normalised Winning Number calculation the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
+By calculating it in this way, we implicitly make the assumption that the Learning-to-Rank methods are (approximately) distributed uniformly over the benchmarks, such that the average Learning-to-Rank method tested are approximately equally hard for each data set.
+It could be the case however that this assumption is false and that one some datasets significantly better Learning-to-Rank methods are evaluated than on some other datasets.

### 2014-04-08 report/Chapters/CrossBenchmarkComparison.tex
references naar appendices

@@
 Figure \ref{fig:normalised_winning_number_ndcg3} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@3 metric.
+Table \ref{tab:raw_data_norm_winnum_ndcg3} in Appendix \ref{app:norm_winnum_ndcg3} provides the raw Normalised Winning Number data for the Learning-to-Rank methods for which \ac{nDCG}@3 evaluation results were available.
 \begin{figure}[!h]
@@
 \end{figure}
+%TODO: Wat valt er te zeggen over de NDCG@3 resultaten? %
 
 \subsection{NDCG@5}
-Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@5 metric. Raw data for the comparison
+Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@5 metric.
+Table \ref{tab:raw_data_norm_winnum_ndcg5} in Appendix \ref{app:norm_winnum_ndcg5} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.
 \begin{figure}[!h]
@@
 \end{figure}
+%TODO: Wat valt er te zeggen over de NDCG@5 resultaten? %
 
@@
 Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@10 metric.
+Table \ref{tab:raw_data_norm_winnum_ndcg10} in Appendix \ref{app:norm_winnum_ndcg10} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.
 \begin{figure}[!h]
@@
 \end{figure}
+%TODO: Wat valt er te zeggen over de NDCG@10 resultaten? %
 
@@
 Figure \ref{fig:normalised_winning_number_map} shows the performance of Learning-to-Rank methods for the \ac{MAP} metric.
+Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.
 \begin{figure}[!h]
@@
 \end{figure}
+%TODO: Wat valt er te zeggen over de MAP resultaten? %
 
@@
 Figure \ref{fig:normalised_winning_number_all} shows the Normalised Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}.
-Table \ref{tab:raw_data_norm_winnum} and Table \ref{tab:raw_data_norm_winnum_cntd} in Appendix \ref{app:norm_winnum_all} provide the raw data plotted in Figure \ref{fig:normalised_winning_number_all}.
+Table \ref{tab:raw_data_norm_winnum_all} in Appendix \ref{app:norm_winnum_all} provide the raw data plotted in Figure \ref{fig:normalised_winning_number_all}.
 \begin{figure}[!h]
@@
 \end{figure}
+%TODO: Wat valt er te zeggen over de cross-metric resultaten? %
 

### 2014-04-08 method_selection/winning_number.R
ook tweede plaatsen gelabeld in cross-metric analyse

@@ -122,17 +122,17 @@ combined_output <- combined_output[, c(ncol(combined_output), 1:(ncol(combined_o
 ### PLOT ###
 ### ADJUST HERE ###
 # plot graph
-ggplot(map_output, aes(Nr_measurements,Normalised_winning_number)) + geom_point(size=2) + 
-  geom_text(aes(label=Method), hjust=-0.1, vjust=-0.1) + 
-  theme_grey(base_size=12, base_family="") + 
-  scale_x_continuous("# of datasets evaluated on", expand = c(0,0), breaks=c(0,2,4,6,8,10,12,14, 16)) +
-  scale_y_continuous("Normalised Winning Number", expand = c(0,0), breaks=c(0,0.2,0.4,0.6,0.8,1.0)) +
-  expand_limits(x = c(0,15.03), y = c(0,1.03))
+# ggplot(ndcg10_output, aes(Nr_measurements,Normalised_winning_number)) + geom_point(size=2) + 
+#  geom_text(aes(label=Method), hjust=-0.1, vjust=-0.1) + 
+#  theme_grey(base_size=12, base_family="") + 
+#  scale_x_continuous("# of datasets evaluated on", expand = c(0,0), breaks=c(0,2,4,6,8,10,12,14, 16)) +
+#  scale_y_continuous("Normalised Winning Number", expand = c(0,0), breaks=c(0,0.2,0.4,0.6,0.8,1.0)) +
+#  expand_limits(x = c(0,18.03), y = c(0,1.03))
 
-#ggplot(combined_aggregate, aes(Summed_ideal_winnum,Norm_winnum)) + geom_point(size=2) +  
-#  geom_text(aes(label=Method)) +
-#  scale_x_continuous("Ideal Winning Number", expand = c(0,0)) + 
-#  scale_y_continuous("Normalized Winning Number", expand=c(0,0)) +
-#  expand_limits(x = c(0,1000), y = c(0,1.01)) +
-#  theme(axis.title.y = element_text(size = rel(1.8), angle = 90)) +
-#  theme(axis.title.x = element_text(size = rel(1.8), angle = 00))
+ggplot(combined_aggregate, aes(Summed_ideal_winnum,Norm_winnum)) + geom_point(size=2) +  
+  geom_text(aes(label=Method)) +
+  scale_x_continuous("Ideal Winning Number", expand = c(0,0)) + 
+  scale_y_continuous("Normalized Winning Number", expand=c(0,0)) +
[…]

### 2014-04-08 report/Chapters/CrossBenchmarkComparison.tex
ook tweede plaatsen gelabeld in cross-metric analyse
tweede limitation :)

@@
 \begin{figure}[!h]
-\includegraphics[scale=0.40]{gfx/ndcg3_winnum}
+\includegraphics[scale=0.30]{gfx/ndcg3_winnum}
 \caption{\acs{nDCG}@3 comparison of Learning-to-Rank methods}
@@
 \begin{figure}[!h]
-\includegraphics[scale=0.35]{gfx/ndcg5_winnum}
+\includegraphics[scale=0.26]{gfx/ndcg5_winnum}
 \caption{\acs{nDCG}@5 comparison of Learning-to-Rank methods}
@@
 \begin{figure}[!h]
-\includegraphics[scale=0.35]{gfx/ndcg10_winnum}
+\includegraphics[scale=0.26]{gfx/ndcg10_winnum}
 \caption{\acs{nDCG}@10 comparison of Learning-to-Rank methods}
@@
 \begin{figure}[!h]
-\includegraphics[scale=0.35]{gfx/map_winnum}
+\includegraphics[scale=0.26]{gfx/map_winnum}
 \caption{\acs{MAP} comparison of Learning-to-Rank methods}
@@
 By calculating it in this way, we implicitly make the assumption that the Learning-to-Rank methods are (approximately) distributed uniformly over the benchmarks, such that the average Learning-to-Rank method tested are approximately equally hard for each data set.
-It could be the case however that this assumption is false and that one some datasets significantly better Learning-to-Rank methods are evaluated than on some other datasets.
+It could be the case however that this assumption is false and that one some datasets significantly better Learning-to-Rank methods are evaluated than on some other datasets.\\
+
+A second limitation is that the datasets on which Learning-to-Rank methods have been evaluated can not always be regarded a random choice.
+It might be the case that some researchers chose to publish results for exactly those benchmark datasets that showed the most positive results for their Learning-to-Rank method.\\

### 2014-04-09 report/Chapters/CrossBenchmarkComparison.tex
nieuwe limitation

@@
 It might be the case that some researchers chose to publish results for exactly those benchmark datasets that showed the most positive results for their Learning-to-Rank method.\\
+
+Another limitation lies in the meta-analysis methodology.
+Using evaluation results published by other researchers relies on the honesty of those researchers.
+Overly optimistic evaluation results published by other researchers do affect our Normalised Winning Number results.
+Only including studies in the meta-analysis that comparable results on one of the baseline methods of a benchmark reduces this limitation but does not solve it.
+By taking the Ideal Winning Number into account in Figure \ref{fig:normalised_winning_number_all} we further mitigate this limitation, as the Ideal Winning Number is loosely related with the number of studies that reported evaluation results for an algorithm.

### 2014-04-09 report/Chapters/CrossBenchmarkComparison.tex
baseline annotaties

@@
 \section{Evaluation Results Found in Literature}
-Table~\ref{tab:ltr_methods_used} gives an overview of the Learning-to-Rank methods for which evaluation results were found for one of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.\\
+Table \ref{tab:ltr_methods_used} gives an overview of the Learning-to-Rank methods for which evaluation results were found for one of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
+Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, LETOR 3.0 and LETOR 4.0 baselines respectively.\\
 
@@
 \begin{table}[!h!p]
-\scalebox{0.80}{
+\scalebox{0.75}{
 \begin{tabular}{|l|l|l||l|l|l|}
@@
 \hline 
-AdaRank-\acs{MAP} & \cite{Xu2007} &  & Linear Regression & \cite{Cossock2006} & \cite{Wang2012, Volkovs2011} \\ 
-AdaRank-\acs{nDCG} & \cite{Xu2007} & \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
-ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet & \cite{Cao2007} &  \\ 
+AdaRank-\acs{MAP} & \cite{Xu2007} & L2, L3, L4 & Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} \\ 
+AdaRank-\acs{nDCG} & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
+ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet & \cite{Cao2007} & L2, L3, L4 \\ 
 Approx\acs{AP} & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
@@
 BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
-Best Single Feature &  & \cite{Gomes2013} & MHR & \cite{Qin2007} &  \\ 
+Best Single Feature &  & \cite{Gomes2013} & MHR & L2, \cite{Qin2007} &  \\ 
 BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
@@
 Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAgg\acs{nDCG} & \cite{Wang2013} & \cite{Wang2013} \\ 
-DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost & \cite{Freund2003} & \cite{Busa-Fekete2013, Alcantara2010} \\ 
+DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010} \\ 
 DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
@@
 Focused\acs{SVM} & \cite{Niu2012} & \cite{Niu2012} & Rank\acs{RLS} & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
-FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{SVM} & \cite{Herbrich1999, Joachims2002} & \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010} \\ 
-FRank & \cite{Tsai2007} & \cite{Wang2012} & Rank\acs{SVM}-Primal &  & \cite{Lai2011} \\ 
-FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & Rank\acs{SVM}-Struct &  &  \\
+FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{SVM} & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010} \\ 
+FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & Rank\acs{SVM}-Primal &  & L3, \cite{Lai2011} \\ 
+FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & Rank\acs{SVM}-Struct &  & L3, L4 \\
 FSM$^{\acs{SVM}}$ & \cite{Lai2013c} & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
@@
 \acs{GP} & \cite{DeAlmeida2007} & \cite{Alcantara2010} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\  
-\acs{GP}Rank & \cite{Silva2009} & \cite{Torkestani2012} & Ridge Regression & \cite{Cossock2006} &  \\
+\acs{GP}Rank & \cite{Silva2009} & \cite{Torkestani2012} & Ridge Regression & \cite{Cossock2006} & L3 \\
 GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
 Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
-Group\acs{MLE} & \cite{Lin2010} & \cite{Lin2011} & SmoothRank & \cite{Chapelle2010} & \cite{Chapelle2010} \\
+Group\acs{MLE} & \cite{Lin2010} & \cite{Lin2011} & SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010} \\
 IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
@@
 Kernel-\acs{PCA} RankBoost & \cite{Duh2008} & \cite{Duh2008, Sato2013} & \acs{SVD}-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\
-KL-\acs{CRF} & \cite{Volkovs2011} & \cite{Volkovs2011} & \acs{SVM}$^{\acs{MAP}}$ & \cite{Yue2007} & \cite{Wang2012, Xu2008, Niu2012} \\ 
+KL-\acs{CRF} & \cite{Volkovs2011} & \cite{Volkovs2011} & \acs{SVM}$^{\acs{MAP}}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
 LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 

### 2014-04-09 report/Chapters/CrossBenchmarkComparison.tex
classicthesis marges aangepast

@@
 \end{figure}
-%TODO: Wat valt er te zeggen over de NDCG@3 resultaten? %
+LambdaNeuralRank and CoList both acquired a perfect Normalised Winning Number score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
+LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three datasets which in addition are more relevant datasets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 datasets.
+FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in Normalised Winning Number, but increasingly higher in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.
 
@@
 By taking the Ideal Winning Number into account in Figure \ref{fig:normalised_winning_number_all} we further mitigate this limitation, as the Ideal Winning Number is loosely related with the number of studies that reported evaluation results for an algorithm.
+
+\section{Conclusions}

### 2014-04-09 report/Chapters/CrossBenchmarkComparison.tex
ndcg5 results & discussion

--- 2014-04-09 report/Chapters/CrossBenchmarkComparison.tex
@@
 The methods for which it holds that there is no other method that has 1) a higher Normalised Winning Number and 2) a higher number datasets evaluated, are identified as the best performing methods and are labelled with the name of the method.
+
 \subsection{NDCG@3}
 Figure \ref{fig:normalised_winning_number_ndcg3} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@3 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg3} in Appendix \ref{app:norm_winnum_ndcg3} provides the raw Normalised Winning Number data for the Learning-to-Rank methods for which \ac{nDCG}@3 evaluation results were available.
+Table \ref{tab:raw_data_norm_winnum_ndcg3} in Appendix \ref{app:norm_winnum_ndcg3} provides the raw Normalised Winning Number data for the Learning-to-Rank methods for which \ac{nDCG}@3 evaluation results were available.\\
+
 \begin{figure}[!h]
@@
 \end{figure}
-%TODO: Wat valt er te zeggen over de NDCG@3 resultaten? %
+
+LambdaNeuralRank and CoList both acquired a perfect Normalised Winning Number score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
+LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three datasets which in addition are more relevant datasets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 datasets.
+FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in Normalised Winning Number, but increasingly higher in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
+
+LambdaNeuralRank, CoList, LARF, LRUF, OWPC and DCMP evaluation results are all based on one study, therefore are subjected to the risk of one overly optimistic study producing those results.
+FenchelRank evaluation result are based combined result from two studies, although those studies have overlap in authors.
+SmoothRank and ListNet have the most reliable evaluation result source, as they were official LETOR baseline runs.
 
@@
 Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@5 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg5} in Appendix \ref{app:norm_winnum_ndcg5} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.
+Table \ref{tab:raw_data_norm_winnum_ndcg5} in Appendix \ref{app:norm_winnum_ndcg5} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.\\
+
 \begin{figure}[!h]
@@
 \end{figure}
-%TODO: Wat valt er te zeggen over de NDCG@5 resultaten? %
+
+LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a Normalised Winning Number of 1.0.
+LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of datasets, but score decreasingly well in terms of Normalised Winning Number.
+These results are highly in agreement with the \ac{nDCG}@3 comparison, with the only modification being that OWPC showed to be a method with no better method performing better on both axes in the \ac{nDCG}@3 comparison, but not in the \ac{nDCG}@5 comparison.
+Also in this comparison SmoothRank and ListNet can be regarded as most reliable results, as the evaluation measurements for these methods are based on LETOR official baselines.
 
@@
 Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@10 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg10} in Appendix \ref{app:norm_winnum_ndcg10} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.
+Table \ref{tab:raw_data_norm_winnum_ndcg10} in Appendix \ref{app:norm_winnum_ndcg10} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.\\
+
 \begin{figure}[!h]
@@
 \end{figure}
+
 %TODO: Wat valt er te zeggen over de NDCG@10 resultaten? %
@@
 Figure \ref{fig:normalised_winning_number_map} shows the performance of Learning-to-Rank methods for the \ac{MAP} metric.
-Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.
+Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.\\
+
 \begin{figure}[!h]
@@
 \end{figure}
+
 %TODO: Wat valt er te zeggen over de MAP resultaten? %

### 2014-04-09 report/Chapters/CrossBenchmarkComparison.tex
ndcg@10 verhaal

@@
 \begin{figure}[!h]
-\includegraphics[scale=0.30]{gfx/ndcg3_winnum}
+\includegraphics[scale=0.285]{gfx/ndcg3_winnum}
 \caption{\acs{nDCG}@3 comparison of Learning-to-Rank methods}
@@
 \begin{figure}[!h]
-\includegraphics[scale=0.26]{gfx/ndcg5_winnum}
+\includegraphics[scale=0.285]{gfx/ndcg5_winnum}
 \caption{\acs{nDCG}@5 comparison of Learning-to-Rank methods}
@@
 LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of datasets, but score decreasingly well in terms of Normalised Winning Number.
-These results are highly in agreement with the \ac{nDCG}@3 comparison, with the only modification being that OWPC showed to be a method with no better method performing better on both axes in the \ac{nDCG}@3 comparison, but not in the \ac{nDCG}@5 comparison.
-Also in this comparison SmoothRank and ListNet can be regarded as most reliable results, as the evaluation measurements for these methods are based on LETOR official baselines.
+These results are highly in agreement with the \ac{nDCG}@3 comparison.
+The only modification compared to the \ac{nDCG}@3 comparison being that OWPC did show to be a method for which there were no methods performing better on both axes in the \ac{nDCG}@5 comparison, but not in the \ac{nDCG}@3 comparison.
+Like in the \ac{nDCG}@3 comparison, SmoothRank and ListNet can be regarded as most reliable results because the evaluation measurements for these methods are based on LETOR official baselines.
 
@@
 
+LambdaMART and LambdaNeuralRank score a Normalised Winning Number of 1.0 on the \ac{nDCG}@10 comparison.
+For LambdaNeuralRank these results are again based on AOL dataset measurements.
+LambdaMART showed the highest \ac{nDCG}@10 performance for the MSLR-WEB10k dataset.
+The set of algorithms for which there is no other algorithm with both a higher Normalised Winning Number and number of datasets evaluated on is partly in agreement with those for the \ac{nDCG}@3 and \ac{nDCG}@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
+SmoothRank and FSMRank were not present in this set for the \ac{nDCG}@3 and \ac{nDCG}@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_ndcg3} and \ref{tab:raw_data_norm_winnum_ndcg5} in the Appendices.
+DCMP is not in the set in contrast with the \ac{nDCG}@3 and \ac{nDCG}@5 comparison.
+
 \begin{figure}[!h]
-\includegraphics[scale=0.26]{gfx/ndcg10_winnum}
+\includegraphics[scale=0.285]{gfx/ndcg10_winnum}
 \caption{\acs{nDCG}@10 comparison of Learning-to-Rank methods}
@@
 \begin{figure}[!h]
-\includegraphics[scale=0.26]{gfx/map_winnum}
+\includegraphics[scale=0.285]{gfx/map_winnum}
 \caption{\acs{MAP} comparison of Learning-to-Rank methods}
@@
 \begin{figure}[!h]
-\includegraphics[scale=0.35]{gfx/combined_normalized_winnum}
+\includegraphics[scale=0.33]{gfx/combined_normalized_winnum}
 \caption{Cross-benchmark comparison of Learning-to-Rank methods}

### 2014-04-09 report/Chapters/CrossBenchmarkComparison.tex
verhaal MAP-results

@@
 
+\begin{figure}[!h]
+\includegraphics[scale=0.285]{gfx/ndcg10_winnum}
+\caption{\acs{nDCG}@10 comparison of Learning-to-Rank methods}
+\label{fig:normalised_winning_number_ndcg10}
+\end{figure}
+
 LambdaMART and LambdaNeuralRank score a Normalised Winning Number of 1.0 on the \ac{nDCG}@10 comparison.
@@
 DCMP is not in the set in contrast with the \ac{nDCG}@3 and \ac{nDCG}@5 comparison.
-
-\begin{figure}[!h]
-\includegraphics[scale=0.285]{gfx/ndcg10_winnum}
-\caption{\acs{nDCG}@10 comparison of Learning-to-Rank methods}
-\label{fig:normalised_winning_number_ndcg10}
-\end{figure}
-
-%TODO: Wat valt er te zeggen over de NDCG@10 resultaten? %
 
@@
 
-%TODO: Wat valt er te zeggen over de MAP resultaten? %
+Where comparisons on the \ac{nDCG}-metric at different cut-off points where highly in agreement in terms of the best performing algorithms, the comparison in terms of \ac{MAP} shows different results.
+RankDE scores a Normalised Winning Number of 1.0 on one dataset, like LambdaNeuralRank did on for all \ac{nDCG}-comparisons.
+In contrast to LambdaNeuralRank, RankDE achieved this highest score on the LETOR 2.0 TD2003, a dataset on which many methods are evaluated.\\
+
+LARF and LRUF score very high Normalised Winning Number scores, but based on only relatively few datasets, just as in the \ac{nDCG}-comparisons.
+Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all \ac{nDCG}-comparisons, are not within the best performing methods in the \ac{MAP}-comparison.
+A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_ndcg3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_ndcg3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more datasets for \ac{MAP} compared to \ac{nDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalised Winning Number.
+SmoothRank performed a Normalised Winning Number of around 0.53 over 7 datasets, which is still good in both certainty and accuracy, but not among the top methods.
+RE-QR is one of the best performers in the \ac{MAP}-comparison with a reasonable amount of benchmark evaluations.
+No reported \ac{nDCG} performance was found in the literature study for RE-QR.
+There is a lot of certainty on the accuracy of RankBoost and Rank\ac{SVM} as both models are evaluated on the majority of datasets included in the comparison for the \ac{MAP}-metric, but given their Normalised Winning Number it can said that both methods are not within the top performing Learning-to-Rank methods.
 

### 2014-04-09 report/Chapters/CrossBenchmarkComparison.tex
begin cross-metrric

@@
 Figure \ref{fig:normalised_winning_number_all} shows the Normalised Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}.
-Table \ref{tab:raw_data_norm_winnum_all} in Appendix \ref{app:norm_winnum_all} provide the raw data plotted in Figure \ref{fig:normalised_winning_number_all}.
+Table \ref{tab:raw_data_norm_winnum_all} in Appendix \ref{app:norm_winnum_all} provide the raw data plotted in Figure \ref{fig:normalised_winning_number_all}.\\
+
 \begin{figure}[!h]
@@
 \end{figure}
-%TODO: Wat valt er te zeggen over de cross-metric resultaten? %
+
+The cross-metric comparison is based on the \ac{nDCG}@3, \ac{nDCG}@5, \ac{nDCG}@10 and \ac{MAP} comparisons combined, which justifies analysing the comparison more thoroughly.
+Figure \ref{fig:normalised_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
 

### 2014-04-10 report/Chapters/CrossBenchmarkComparison.tex
cross-comparison
cross-metric comparison

@@
 Figure \ref{fig:normalised_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
+In addition, Linear Regression and the ranking method of simply sorting on the best single feature are labelled as baselines.\\
+
+LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both Ideal Winning Number and Normalised Winning Number.
+LRUF is the only methof that achieved this in all \ac{nDCG} comparisons, the \ac{MAP} comparison as well as the cross-metric comparison.
+With FenchelRank, FSMRank, SmoothRank and ListNet being among the top performing methods in all \ac{nDCG} comparisons as well as in the cross-metric comparison, it can be concluded that the cross-metric results are highly defined by the \ac{nDCG} performance as opposed to the \ac{MAP} performance.
+This was to be expected, because the cross-metric comparison input data of three \ac{nDCG} entries (@3, @5, and @10) enables it to have up to three times as many as many weight as the \ac{MAP} comparison.
 

### 2014-04-10 report/Chapters/CrossBenchmarkComparison.tex
hopefull methods

@@
 
-\acs{nDCG}@3, \acs{nDCG}@5, \acs{nDCG}@10 and \acs{MAP} are chosen as metrics on which the meta-analysis will be performed.
+\ac{nDCG}@3, \ac{nDCG}@5, \ac{nDCG}@10 and \ac{MAP} are chosen as metrics on which the meta-analysis will be performed.
 These metrics seem to be the most frequently used evaluation metrics in most of the benchmark datasets used.
@@
 The lack of use of the \ac{ERR}-metric in other benchmarks makes it unsuitable for a cross-benchmark comparison.
+By choosing including \ac{nDCG} at thee cut-off points and only a single \ac{MAP} entry we implicitly attain a higher weight for \ac{nDCG} compared to \ac{MAP} on an analysis that combines all measurements on the four metrics.
+This implicit weighting is a bit arbitrary and therefore undesirable, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.
 
@@
 With FenchelRank, FSMRank, SmoothRank and ListNet being among the top performing methods in all \ac{nDCG} comparisons as well as in the cross-metric comparison, it can be concluded that the cross-metric results are highly defined by the \ac{nDCG} performance as opposed to the \ac{MAP} performance.
-This was to be expected, because the cross-metric comparison input data of three \ac{nDCG} entries (@3, @5, and @10) enables it to have up to three times as many as many weight as the \ac{MAP} comparison.
+This was to be expected, because the cross-metric comparison input data of three \ac{nDCG} entries (@3, @5, and @10) enables it to have up to three times as many as many weight as the \ac{MAP} comparison.\\
+
+LARF, \ac{IP}Rank and DCMP and several variants of Rank\ac{SVM} performed very well on the cross-metric comparison, with all having only one method in its top right quadrant.
+LARF also performed among the top methods on the \ac{nDCG} and \ac{MAP} comparisons and DCMP was a top performer in a few of the \ac{nDCG} comparisons.\\
+
+C-CRF, DirectRank, FP-Rank, RankCSA, LambdaNeuralRank and VFLR all have near-perfect Normalised Winning Number measures, but have low Ideal Winning Number measures.
+Further evaluation runs of these methods on benchmark datasets that they are not yet evaluated on are desirable.
+The DirectRank paper \cite{Tan2013} shows that the method actually is evaluated datasets than is used in this paper.
+however measurements on some datasets were only available in graphical form and not in raw data, making these measurements unusable for this comparison.
 
@@
 \section{Conclusions}
+We will now look back on our first research question using the knowledge gained with the cross-benchmark comparison.
+\begin{description}
+\item[RQ1] What are the best performing Learning-to-Rank algorithms in terms of accuracy on relevant benchmark datasets?\\
+\end{description}
+Although no closing arguments can be formulated on which Learning-to-Rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.\\
+
+LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the Learning-to-Rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
+From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.
+For more definite conclusions on the relative performance of these five methods, more evaluation runs on are desirable for the methods on the left side on the list on benchmark datasets that these methods have not yet been evaluated on.\\

### 2014-04-10 report/Chapters/CrossBenchmarkComparison.tex
cross benchmark comparison

@@
 The DirectRank paper \cite{Tan2013} shows that the method actually is evaluated datasets than is used in this paper.
-however measurements on some datasets were only available in graphical form and not in raw data, making these measurements unusable for this comparison.
+however measurements on some datasets were only available in graphical form and not in raw data, making these measurements unusable for this comparison.\\
+
+LAC-MR-OR and RE-QR showed very good ranking accuracy in the \ac{MAP} comparison on multiple datasets.
+Both methods Because LAC-MR-OR is only evaluated on two datasets for \ac{nDCG}@10 and RE-QR is not evaluated for \ac{nDCG} at all
 
@@
 For more definite conclusions on the relative performance of these five methods, more evaluation runs on are desirable for the methods on the left side on the list on benchmark datasets that these methods have not yet been evaluated on.\\
+
+In the following chapters of this thesis, concerning parallel execution of the Learning-to-Rank training phase, the scope will be limited to the five methods that turned out to be superior Learning-to-Rank methods in terms of ranking accuracy and certainty about this ranking accuracy: LRUF, FSMRank, FenchelRank, SmoothRank and ListNet.
+Although it can not be concluded that these methods are inarguably the most accurate Learning-to-Rank methods, a strong presumption has been raised that these five Learning-to-Rank are accurate ranking methods.

### 2014-04-11 report/Chapters/LTRMethods.tex
cross benchmark comparison
begin uitleg ListNet

@@
 Algorithms and details of the well-performing Learning-to-Rank methods as selected in the afore-going part are presented and explained in this part.
+
+\section{ListNet}
+\begin{algorithm}[H]
+ \KwData{training data \{$(x^{(1)},y^{(1)}),(x^{(2)},y^{(2)}),...,(x^{(m)},y^{(m)})$\}}
+ Parameter: number of iterations $T$ and learning rate $\eta$\\
+ Initialize parameter $\omega$\\
+ \For{$t\leftarrow 1$ \KwTo $T$}{
+ 	\For{$i\leftarrow 1$ \KwTo $m$}{
+ 		Input $x^{(i)}$ of query $q^{(i)}$ to Neural Network and compute score list $z^{(i)}(f_\omega)$ with current $\omega$.\\
+ 		Compute gradient $\Delta\omega$ using Eq. (\ref{eq:gradient_descent}).\\
+ 		Update $\omega = \omega - \eta \times \Delta\omega$.
+ 	}
+ }
+ Output Neural Network model $\omega$.
+ \caption{Learning algorithm of ListNet, obtained from \cite{Cao2007}}
+ \label{alg:listnet}
+\end{algorithm}
+
+\begin{equation}
+\Delta\omega = \frac{\partial L(y^{(i)},z^{(i)}(f_\omega))}{\partial \omega} = - \sum_{\forall g \in \mathscr{G}_k}\limits\frac{\partial P_{z^{(i)}(f_\omega)}(g)}{\partial \omega}\frac{P_{y^{(i)}}(g)}{P_{z^{(i)}(f_\omega)}(g)}
+\label{eq:gradient_descent}
+\end{equation}
+\section{SmoothRank}
+
+\section{FenchelRank}
+
+\section{FSMRank}
+
+\section{LRUF}

### 2014-04-11 report/Chapters/Background.tex
begin cross dataset comparison
realisatie dat OHSUMED, TD2003 en TD2004 niet hetzelfde zijn voor LETOR 2.0 en LETOR 3.0. LETOR 2.0 baselinewaarden ingevoerd
list of algorithms

@@
 where the product sequence part of the formula represents the chance that the user will stop at position $r$.\\
-An algorithm to compute \ac{ERR} is shown in Listing \ref{lst:err}.
+An algorithm to compute \ac{ERR} is shown in Listing \ref{alg:err}.
 The algorithm requires relevance grades $g_i$, $1 \le i \le n$ and mapping function $R$ that maps relevance grades to probability of relevance.
-\begin{lstlisting}[caption={Algorithm to compute the \acs{ERR} metric, obtained from \cite{Chapelle2009}}, label={lst:err},language=Ada]
-p <- 1, ERR <- 0
-for r=1 to n do
-	R <- R(g[r])
-	ERR <- ERR + p * R/r
-	p <- p * (1-R)
-end for
-return ERR
-\end{lstlisting}
+
+\begin{algorithm}[H]
+ $p \leftarrow 1, ERR \leftarrow 0$\\
+ \For{$r\leftarrow 1$ \KwTo $n$}{
+	$R \leftarrow R(g[r])$\\
+	$ERR \leftarrow ERR + p * \frac{R}{r}$\\
+	$p \leftarrow p * (1-R)$
+ }
+ Output $ERR$
+ \caption{Algorithm to compute the \acs{ERR} metric, obtained from \cite{Chapelle2009}}
+ \label{alg:err}
+\end{algorithm}
+
 \subsection{Mean Average Precision}
@@
 \end{table}
+
+\subsection{Concluding notes on evaluation measures}
+TODO: Ingaan op verbanden/correlaties tussen metrics.
+TODO: Explain motivation for cutoff points in ranking evaluation measures.
 

### 2014-04-11 report/Chapters/CrossBenchmarkComparison.tex
feedback jurjen verwerkt

@@
-As we have seen in chapter~\ref{chap:benchmark_results}, evaluation of Learning-to-Rank methods is spread out over the several benchmark datasets.
-However, as the evaluated Learning-to-Rank methods differ from benchmark to benchmark, no single benchmark comparison can be regarded as a conclusive argument on which Learning-to-Rank method is most accurate.\\
+As we have seen in Chapter~\ref{chap:benchmark_results}, the evaluation of Learning-to-Rank methods is spread over the several benchmark datasets.
+However, as the Learning-to-Rank methods evaluated differs between benchmarks, no single benchmark comparison can be regarded as a conclusive argument on which Learning-to-Rank method is most accurate.\\
 
@@
 Gomes et al.
-\cite{Gomes2013} analysed ranking accuracy of a set of models over both LETOR 3.0 and LETOR 4.0.
+\cite{Gomes2013} analysed ranking accuracy of a set of models on both LETOR 3.0 and LETOR 4.0.
 Busa-Fekete et al.
@@
 Learning to Rank Challenge datasets and the OHSUMED dataset from LETOR 3.0.
-To my knowledge, no large-scale accuracy meta-analysis has been conducted where published evaluation results on several benchmark datasets are taken into account.
+To my knowledge, no structured meta-analysis on ranking accuracy has been conducted where evaluation results on several benchmark collections are taken into account.
 With a meta-analysis I will compare the performance of Learning-to-Rank methods across the Learning-to-Rank benchmark datasets described in foregoing sections.
@@
 LETOR 1.0, LETOR 3.0, Yahoo!
-Learning to Rank Challenge, WCL2R and AOL have an accompanying paper that was published together with the benchmark collection.
-Users of those benchmark collections are encouraged to cite these papers.
-A forward literature search on these studies is performed to collect studies that present measurements of evaluations of Learning-to-Rank methods on these benchmark collections.
+Learning to Rank Challenge, WCL2R and AOL have an accompanying paper which was published together with the benchmark collection.
+As users of those benchmark collections are encouraged to cite these papers, evaluated measurements for these benchmark collections will be gathered through a forward literature search.
 Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search.
+Google Scholar will be used for this forward reference search.
 
@@
 In contrast to the winning number comparison on LETOR 3.0, there will not be accuracy measurements for each algorithm on each dataset in our meta-analysis.
-To compare algorithms based on a sparse set of evaluation measurements, a normalised version of the Winning Number metric will be used which we calculate by only taking those datasets into account that an algorithm is evaluated on and then dividing this by the ideal Winning Number that an algorithm could have had if it would have been the best algorithm on all datasets on which it has been evaluated.
+To compare algorithms based on a sparse set of evaluation measurements, a normalised version of the Winning Number metric will be used.
+This Normalised Winning Number takes into account only those datasets that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all datasets on which it has been evaluated.
 I will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as \\
@@
 \ac{nDCG}@3, \ac{nDCG}@5, \ac{nDCG}@10 and \ac{MAP} are chosen as metrics on which the meta-analysis will be performed.
-These metrics seem to be the most frequently used evaluation metrics in most of the benchmark datasets used.
+These metrics seem to be the most frequently used evaluation metrics in most of the used benchmark datasets.
 An exception are the Yahoo!
@@
 The lack of use of the \ac{ERR}-metric in other benchmarks makes it unsuitable for a cross-benchmark comparison.
-By choosing including \ac{nDCG} at thee cut-off points and only a single \ac{MAP} entry we implicitly attain a higher weight for \ac{nDCG} compared to \ac{MAP} on an analysis that combines all measurements on the four metrics.
-This implicit weighting is a bit arbitrary and therefore undesirable, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.
+By making the decision to include \ac{nDCG} at thee cut-off points and only a single \ac{MAP} entry we implicitly attain a higher weight for \ac{nDCG} compared to \ac{MAP} on an analysis that combines all measurements on the four metrics.
+This implicit weighting is arbitrary and therefore undesirable, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.
 
 \section{Evaluation Results Found in Literature}
-Table \ref{tab:ltr_methods_used} gives an overview of the Learning-to-Rank methods for which evaluation results were found for one of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
+Table \ref{tab:ltr_methods_used} gives an overview of the Learning-to-Rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
 Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, LETOR 3.0 and LETOR 4.0 baselines respectively.\\
@@
 Further evaluation runs of these methods on benchmark datasets that they are not yet evaluated on are desirable.
-The DirectRank paper \cite{Tan2013} shows that the method actually is evaluated datasets than is used in this paper.
-however measurements on some datasets were only available in graphical form and not in raw data, making these measurements unusable for this comparison.\\
+The DirectRank paper \cite{Tan2013} shows that the method  is evaluated on more datasets than the number of datasets that we included evaluation results for in this meta-analysis.
+Some of the DirectRank measurements could not be used because measurements on some datasets were only available in graphical form and not in raw data.\\
 
@@
 Overly optimistic evaluation results published by other researchers do affect our Normalised Winning Number results.
-Only including studies in the meta-analysis that comparable results on one of the baseline methods of a benchmark reduces this limitation but does not solve it.
+Only including studies in the meta-analysis that comparable show results on one of the baseline methods of a benchmark reduces this limitation but does not solve it.
 By taking the Ideal Winning Number into account in Figure \ref{fig:normalised_winning_number_all} we further mitigate this limitation, as the Ideal Winning Number is loosely related with the number of studies that reported evaluation results for an algorithm.

### 2014-04-11 report/Chapters/AppendixA.tex
graffito + begin benchmark_results.csv
\ac{ -> \acs{ in tabel

@@
  \hline
-1 & \ac{TF} of body					& 36& LMIR.JM of body\\
-2 &	\ac{TF} of anchor				& 37& LMIR.JM of anchor\\
-3 & \ac{TF} of title				& 38& LMIR.JM of title\\
-4 & \ac{TF} of \ac{URL}				& 39& LMIR.JM of \ac{URL}\\
-5 & \ac{TF} of whole document		& 40& LMIR.JM of whole document\\
+1 & \acs{TF} of body					& 36& LMIR.JM of body\\
+2 &	\acs{TF} of anchor				& 37& LMIR.JM of anchor\\
+3 & \acs{TF} of title				& 38& LMIR.JM of title\\
+4 & \acs{TF} of \acs{URL}				& 39& LMIR.JM of \acs{URL}\\
+5 & \acs{TF} of whole document		& 40& LMIR.JM of whole document\\
 
-6 & \ac{IDF} of body 				& 41& Sitemap based term propagation\\
-7 & \ac{IDF} of anchor 				& 42& Sitemap based score propagation\\
-8 & \ac{IDF} of title 				& 43& Hyperlink based score propagation: weighted in-link\\
-9 & \ac{IDF} of \ac{URL} 			& 44& Hyperlink based score propagation: weighted out-link\\
-10& \ac{IDF} of whole document 		& 45& Hyperlink based score propagation: uniform out-link\\
+6 & \acs{IDF} of body 				& 41& Sitemap based term propagation\\
+7 & \acs{IDF} of anchor 				& 42& Sitemap based score propagation\\
+8 & \acs{IDF} of title 				& 43& Hyperlink based score propagation: weighted in-link\\
+9 & \acs{IDF} of \acs{URL} 			& 44& Hyperlink based score propagation: weighted out-link\\
+10& \acs{IDF} of whole document 		& 45& Hyperlink based score propagation: uniform out-link\\
 
-11& \ac{TF-IDF} of body 			& 46& Hyperlink based feature propagation: weighted in-link\\
-12& \ac{TF-IDF} of anchor 			& 47& Hyperlink based feature propagation: weighted out-link\\
-13& \ac{TF-IDF} of title 			& 48& Hyperlink based feature propagation: uniform out-link\\
-14& \ac{TF-IDF} of \ac{URL}			& 49& HITS authority\\
-15& \ac{TF-IDF} of whole document	& 50& HITS hub\\
+11& \acs{TF-IDF} of body 			& 46& Hyperlink based feature propagation: weighted in-link\\
+12& \acs{TF-IDF} of anchor 			& 47& Hyperlink based feature propagation: weighted out-link\\
+13& \acs{TF-IDF} of title 			& 48& Hyperlink based feature propagation: uniform out-link\\
+14& \acs{TF-IDF} of \acs{URL}			& 49& HITS authority\\
+15& \acs{TF-IDF} of whole document	& 50& HITS hub\\
 
@@
 18& Document length of title		& 53& Topical PageRank\\
-19& Document length of \ac{URL}		& 54& Topical HITS authority\\
+19& Document length of \acs{URL}		& 54& Topical HITS authority\\
 20& Document length of whole document	& 55& Topical HITS hub\\
@@
 22& BM25 of anchor					& 57& Out-link number\\
-23& BM25 of title					& 58& Number of slashes in \ac{URL}\\
-24& BM25 of \ac{URL}				& 59& Length of \ac{URL}\\
+23& BM25 of title					& 58& Number of slashes in \acs{URL}\\
+24& BM25 of \acs{URL}				& 59& Length of \acs{URL}\\
 25& BM25 of whole document			& 60& Number of child page\\
@@
 28& LMIR.ABS of title				& 63& LMIR.DIR of extracted title\\
-29& LMIR.ABS of \ac{URL}			& 64& LMIR.JM of extracted title\\
+29& LMIR.ABS of \acs{URL}			& 64& LMIR.JM of extracted title\\
 30& LMIR.ABS of whole document\\
@@
 33& LMIR.DIR of title\\
-34& LMIR.DIR of \ac{URL}\\
+34& LMIR.DIR of \acs{URL}\\
 35& LMIR.DIR of whole document\\

### 2014-04-14 report/Chapters/Background.tex
TODO gevinkt, tweede TODO ge-outcomment

@@
 Table \ref{tab:example_calculation_nDCG} shows an example calculation for \ac{nDCG} for both the J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} and Burges et al.
-\cite{Burges2005} version of \ac{DCG}.
+\cite{Burges2005} version of \ac{DCG}.\\
 
@@
 \end{table}
+
+One limitation of \ac{nDCG} that it does not penalise for missing documents in the result set.
+\acl{nDCG} is often used with a fixed set size for the result set to mitigate the missing document problem.
+\ac{nDCG} with a fixed set size is often called \ac{nDCG}@k, where $k$ represents the set size.
+
 \subsection{Expected Reciprocal Rank}
@@
 
-\subsection{Concluding notes on evaluation measures}
-TODO: Ingaan op verbanden/correlaties tussen metrics.
-TODO: Explain motivation for cutoff points in ranking evaluation measures.
+%\subsection{Concluding notes on evaluation measures}
+%TODO: Ingaan op verbanden/correlaties tussen metrics.
 

### 2014-04-14 report/Chapters/BenchmarkResults.tex
benchmarkresults inleiding

@@
 \label{chap:benchmark_results}
-This chapter describes benchmark characteristics like collection size and features and gives an overview of the performance of different Learning-to-Rank methods.
-Performance differences for a given Learning-to-Rank method are explained in terms of differences in benchmark characteristics.
-This chapter is concluded with a selection of Learning-to-Rank methods to include in the experiments.\\
+Benchmark datasets enable fair comparisons of ranking models over a fixed set of documents and queries.
+The approach of using fixed benchmark datasets to compare the performance of Information Retrieval systems under equal circumstances has been the standard evaluation methodology in the Information Retrieval field since the release of the Cranfield collection \cite{Cleverdon1966}.\\
 
-Benchmarking datasets enable fair comparisons of ranking models over a fixed set of documents and queries.
-The approach of using fixed benchmarking datasets to compare the performance of Information Retrieval systems under equal circumstances has been the standard evaluation methodology in the Information Retrieval field since the release of the Cranfield collection \cite{Cleverdon1966}.
+This chapter describes benchmark characteristics, e.g.
+collection size and features, of Learning-to-Rank benchmark collections and datasets and gives an overview of the performance of baseline algorithms on these collections and datasets.\\
+
 \section{Yahoo! Learning to Rank Challenge}

### 2014-04-14 report/Chapters/BenchmarkResults.tex
other datasets

@@
 \cite{Pavlov2010} managed to train half a million trees on 200 nodes in 2 hours.
-\section{Yandex Internet Mathematics competition}
-Yandex Internet Mathematics Competition
-\subsection{Results}
-The Yandex Internet Mathematics competition was won by Pavlov et al. \cite{Pavlov2010} with the BagBoo method.
 
@@
 \section{Other datasets}
+LETOR and the Yahoo! Learning to Rank Challenge datasets are the most used benchmark collections in the field. Several other benchmark collections for the Learning-to-Rank task have been proposed.
 \subsection{WCL2R}
@@
 \subsection{AOL}
-TODO: AOL dataset beschrijven \cite{Pass2006}.
+%TODO: AOL dataset beschrijven \cite{Pass2006}.
+
+\subsection{Yandex Internet Mathematics competition}
+%TODO: YIMC beschrijven
+The Yandex Internet Mathematics competition was won by Pavlov et al. \cite{Pavlov2010} with the BagBoo method.

### 2014-04-14 report/Chapters/BenchmarkResults.tex
OHSUMED LETOR 2.0 != LETOR 3.0, sampling

@@
 The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 \cite{Liu2007b} to solve the absence of a experimental platform for Learning-to-Rank at that time.
-LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009.\\
+LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009.
 
-\subsection{LETOR 3.0}
-The LETOR 3.0 benchmark collection \cite{Qin2010} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.
+\subsection{LETOR 2.0}
+LETOR 2.0 consists of three datasets: the OHSUMED dataset and two datasets of the on the .gov collection.
 The OHSUMED collection is a subset of the medical publication database MEDLINE and contains medical publications from 270 journals that were published between 1987 and 1991.
-The .gov collection is a web crawl obtained in January 2002, which was used for the \ac{TREC} web track in 2003 and 2004.
-Tables \ref{tbl:features_gov} and \ref{tbl:features_ohsumed} in Appendix \ref{chap:appendix_a} provide the descriptions of the features of the .gov and the OHSUMED collections respectively.\\
+The .gov collection is a web crawl obtained in January 2002, which was used for the \ac{TREC} web track in 2003 and 2004.\\
 
@@
 \end{description}
-With query sets available from both the \ac{TREC} 2003 and 2004 conferences there are six query sets in total: TD2003, TD2004, NP2003, NP2004, HP2003 and HP2004.\\
+
+LETOR 2.0 only uses the topic distillation queries. Baseline Learning-to-Rank algorithms evaluated on LETOR 2.0 by the organisation are:
+\begin{description}
+\item[pairwise]{\leavevmode
+	\begin{itemize}
+	\item Rank\ac{SVM} \cite{Herbrich1999,Joachims2002}
+	\item RankBoost \cite{Freund2003}
+	\item \ac{MHR} \cite{Qin2007}
+	\item FRank \cite{Tsai2007}
+	\end{itemize}}
+\item[listwise]{\leavevmode
+	\begin{itemize}
+	\item ListNet \cite{Cao2007}
+	\item AdaRank-MAP \cite{Xu2007}
+	\item AdaRank-nDCG \cite{Xu2007}
+	\end{itemize}}
+\end{description}
+
+\subsubsection{Results}
+
+\subsection{LETOR 3.0}
+The LETOR 3.0 benchmark collection \cite{Qin2010} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.
+Tables \ref{tbl:features_gov} and \ref{tbl:features_ohsumed} in Appendix \ref{chap:appendix_a} provide the descriptions of the features of the .gov and the OHSUMED collections for LETOR 3.0 respectively.
+Where LETOR 2.0 only used topic distillation (TD) queries on the .gov corpus, LETOR 3.0 also uses named page finding (NP) and home page finding (HP) queries.
+With query sets available from both the \ac{TREC} 2003 and 2004 conferences there are six query sets in total: TD2003, TD2004, NP2003, NP2004, HP2003 and HP2004.
+It is noteworthy that the OHSUMED, TD2003 and TD2004 datasets of LETOR 2.0 and LETOR 3.0 are not identical and therefore not comparable due to differences in sampling method.\\
 
@@
 The LETOR organisation implemented and evaluated several well-known Learning-to-Rank algorithms themselves and in addition gathers and publications and results of new algorithms evaluated on the LETOR benchmark.
-The baseline Learning-to-Rank algorithms evaluated by the organisation are:
+The baseline Learning-to-Rank algorithms evaluated by the LETOR team are:
 \begin{description}
@@
 
-\subsection{Results}
+\subsubsection{Results}
 The LETOR paper \cite{Qin2010} describes the performance of the LETOR baseline methods.

### 2014-04-14 report/Chapters/BenchmarkResults.tex
LETOR 2.0 results

@@
 \subsubsection{Results}
+Table \ref{tab:letor2.0_results} shows the performance of the baseline methods on the datasets in the LETOR 2.0 collection.
+\begin{table}
+\begin{tabular}{l|lll}
+ & OHSUMED & TD2003 & TD2004 \\
+ \hline 
+RankBoost & 0.4356 & 0.2851 & 0.4716 \\ 
+Rank\acs{SVM} & 0.4411 & 0.3410 & 0.4201 \\ 
+FRank & 0.4423 & 0.3357 & 0.4708 \\ 
+ListNet & 0.4489 & 0.3743 & 0.4579 \\ 
+AdaRank-\acs{MAP} & 0.4385 & 0.1940 & 0.4063 \\ 
+AdaRank-\acs{nDCG} & 0.4369 & 0.2702 & 0.3878 \\ 
+\acs{MHR} & 0.4423 &  &  \\ 
+\end{tabular}
+\caption{\acs{nDCG}@10 results of baseline methods on LETOR 2.0}
+\label{tab:letor2.0_results}
+\end{table}
 

### 2014-04-14 report/Chapters/BenchmarkResults.tex
LETOR 2.0 results

@@
 \subsubsection{Results}
-Table \ref{tab:letor2.0_results} shows the performance of the baseline methods on the datasets in the LETOR 2.0 collection.
-\begin{table}
+\begin{table}[!h]
 \begin{tabular}{l|lll}
@@
 \end{table}
+Table \ref{tab:letor2.0_results} shows the performance of the baseline methods on the datasets in the LETOR 2.0 collection.
+ListNet can be regarded as the winner of the comparison as it outperformed the other methods on two of the three datasets and ranked third on the third dataset.
+\acs{MHR} was only evaluated on the OHSUMED dataset, on which it performed second best, after ListNet.
 

### 2014-04-14 report/Chapters/BenchmarkResults.tex
stuk bullshit verwijderd

@@
 Since LETOR is arguably the most well-known benchmark collection in the field, it is conceivable that creators of new Learning-to-Rank methods evaluate their new method on the LETOR collection to show how well their new method works.
-To create an overview of the performance of newer Learning-to-Rank methods on LETOR 3.0 we perform a forward reference search on the LETOR 3.0 paper \cite{Qin2010}.
-It should be noted that results of evaluations individually performed by the creator of the evaluated Learning-to-Rank method can not be regarded as equally reliable as the evaluations performed by the LETOR team.
-TODO: 307 citations, x evaluated a new Learning-to-Rank method on LETOR 3.0, x were usable.
-The unusable studies did test a new Learning-to-Rank method on one or more datasets of the LETOR collection, but 1 ) did not use a ranking metric used in LETOR (\ac{nDCG} or \ac{MAP}) or 2) did not list evaluation results at all.\\
 

### 2014-04-14 report/Chapters/BenchmarkResults.tex
betere plaatsing tabellen LETOR 3.0 results
MSLR-WEB

@@
 collection size and features, of Learning-to-Rank benchmark collections and datasets and gives an overview of the performance of baseline algorithms on these collections and datasets.\\
+
+The accuracies of the Learning-to-Rank methods described in the following sections can only be compared within the benchmark and not between benchmarks for the following reasons:
+\begin{enumerate}
+\item Differences in feature sets between data sets detract from fair comparison
+\item Although the \ac{nDCG} definition is unambiguous, Busa-Fekete et al. \cite{Busa-Fekete2012} found that \ac{nDCG} evaluation tools of benchmark data sets produced different scores
+\end{enumerate}
 
@@
 Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR datasets in terms of \ac{nDCG}@10 and \ac{MAP}.
-\begin{table}
+\begin{table}[!h]
 \begin{tabular}{l|ll}
@@
 
-\begin{table}
+\begin{table}[!h]
 \begin{tabular}{l|p{1.2cm}p{1.2cm}p{1.2cm}p{1.4cm}||l}
@@
 
-\section{MSLR-WEB10k/MSLR-WEB30k}
-Contrary to the Yahoo!
-Learning to Rank Challenge dataset, the MSLR-WEB30k provides detailed feature descriptions.
-The MSLR-WEB30k dataset however contains no proprietary features but only features that are commonly used in the research community.
-\subsection{Results}
-
-\section{Selecting Learning-to-Rank methods}
-The accuracies of the Learning-to-Rank methods described in the preceding sections must only be compared within the benchmark and not between benchmarks for the following reasons:
-\begin{enumerate}
-\item Differences in feature sets between data sets detract from fair comparison
-\item Although the \ac{nDCG} definition is unambiguous, Busa-Fekete et al. \cite{Busa-Fekete2012} found that \ac{nDCG} evaluation tools of benchmark data sets produced different scores
-\end{enumerate}
-
 \section{Other datasets}
 LETOR and the Yahoo! Learning to Rank Challenge datasets are the most used benchmark collections in the field. Several other benchmark collections for the Learning-to-Rank task have been proposed.
+
+\subsection{MSLR-WEB10/30k}
+MSLR-WEB30k and the smaller subset MSLR WEB10k are two large datasets with 30,000 and 10,000 queries respectively.
+The data was obtained from the Microsoft Bing\footnote{http://www.bing.com/} commercial web search engine.
+A feature list and feature descriptions are available.
+The dataset only includes features that are commonly used in the research community, as proprietary features were excluded from the dataset.
+Microsoft published these datasets as unofficial successor to LETOR in June 2010, but no baseline evaluations were described by the MSLR-WEB10/30k team.
+The Microsoft Research website from which the dataset is obtainable\footnote{http://research.microsoft.com/en-us/projects/mslr/download.aspx} also offers an evaluation script for \ac{nDCG} and \ac{MAP}.
+The presence of an official evaluation script enables fair comparison of evaluation results from other researchers comparable.
+
 \subsection{WCL2R}

### 2014-04-14 report/Chapters/BenchmarkResults.tex
L2 regulatisation -> ridge

@@
 \begin{tabular}{l|p{1.2cm}p{1.2cm}p{1.2cm}p{1.4cm}||l}
- & \rotatebox{55}{Regression + L2 regulatisation} & \rotatebox{55}{Rank\ac{SVM}-Primal} &\rotatebox{55}{Rank\ac{SVM}-Struct} & \rotatebox{55}{SmoothRank} & \rotatebox{55}{ListNet} \\
+ & \rotatebox{55}{Ridge Regression} & \rotatebox{55}{Rank\ac{SVM}-Primal} &\rotatebox{55}{Rank\ac{SVM}-Struct} & \rotatebox{55}{SmoothRank} & \rotatebox{55}{ListNet} \\
  \hline

### 2014-04-14 report/Chapters/BenchmarkResults.tex
WCL2R results herplaatst
AOL fuckup

@@
 
-\begin{table}
+\begin{table}[!h]
 \begin{tabular}{l|lll|lll}
@@
 \subsection{AOL}
-%TODO: AOL dataset beschrijven \cite{Pass2006}.
+Pass et al.
+\cite{Pass2006} describe an America Online Learning-to-Rank dataset which they published in 2006.
+This dataset is unique in that it is the only large commercial web search dataset that contains user session and user behaviour information.
+The dataset contains 20 million search keywords for over 650,000 users over a three month time period.
+AOL later realised the publication of the dataset to be a mistake, as personally identifiable information turned out to be presented in some of the queries\footnote{http://select.nytimes.com/gst/abstract.html?res=F10612FC345B0C7A8CDDA10894DE404482}.
+AOL acknowledged the mistake of publishing the data and removed the data from its website.
+The removal of the data was too late, as the data was already downloadable from several mirror sites.
+This controversial AOL dataset is to date still used for Learning-to-Rank research.
+No official baseline evaluation is provided by the AOL team.
 

### 2014-04-14 report/Chapters/BenchmarkResults.tex
AOL mirror footnot
Yandex IMC

@@
 AOL acknowledged the mistake of publishing the data and removed the data from its website.
-The removal of the data was too late, as the data was already downloadable from several mirror sites.
+The removal of the data was too late, as the data was already downloadable from several mirror sites\footnote{http://gregsadetsky.com/aol-data/}.
 This controversial AOL dataset is to date still used for Learning-to-Rank research.
@@
 
-\subsection{Yandex Internet Mathematics competition}
-%TODO: YIMC beschrijven
-The Yandex Internet Mathematics competition was won by Pavlov et al. \cite{Pavlov2010} with the BagBoo method.
+\subsection{Yandex Internet Mathematics contest}
+Russian commercial web search engine Yandex dedicated their yearly Internet Mathematics contest to the task of Learning-to-Rank in the 2009 edition of the contest.
+Features (245 in total) are only numbered and their semantics are not revealed, equivalent to the Yahoo!
+Learning to Rank Challenge.
+The set is split into a 45\%-training, 10\%-validation, and 45\%-test data.
+The training set contains 9124 queries, with on average around 100 assessed documents per query.
+Yandex IMC 2009 has an online leaderboard\footnote{http://imat2009.yandex.ru/en/results} showing the best performing teams in the contest, but it is not traceable which methods each teams used.
+Pavlov et al.
+\cite{Pavlov2010} in their paper claim to have won the Yandex Internet Mathematics contest 2009 with their BagBoo method.

### 2014-04-15 report/Chapters/LTRMethods.tex
list of algorithms
ListNet uitgelegd :)

@@
 \section{ListNet}
+ListNet is a listwise ranking function whose loss function is not directly related to information retrieval evaluation metrics.
+ListNet's loss function is defined using the probability distribution on permutations.
+Probability distributions on permutations have been a research topic within the field of probability theory and has been extensively researched.
+One of the most well-known permutation probability distributions in the field, the Plackett-Luce model \cite{Plackett1975,Luce1959}, is used in ListNet.
+The Plackett-Luce model defines the probability of all possible permutations $\pi$, given all document ranking scores $S$, as shown in Equation \ref{eq:plackett_luce}.
+\begin{equation}
+P(\pi|S) = \prod\limits_{j=1}^{m}\frac{\phi(s_{\pi^{-1}(j)})}{\sum\nolimits_{u=1}^{m}\phi(s_{\pi^{-1}(u)})}
+\label{eq:plackett_luce}
+\end{equation}
+ListNet uses Gradient Descent to optimise neural network such that its Cross Entropy loss compared to the Plackett-Luce distribution over the ground truth is minimal.
+Note that some sources, including Liu \cite{Liu2007}, describe ListNet as using KL divergence as loss function.
+KL divergence and Cross-entropy are however identical up to an additive constant when comparing distribution $q$ against a fixed reference distribution $p$, since the cross-entropy definition can be formulated as in Equation \ref{eq:cross_entropy}.
+\begin{equation}
+H(p,q) = H(p) + D_{KL}(p||q)
+\label{eq:cross_entropy}
+\end{equation}
+where $H(p)$ is the entropy of $p$ and $D_{KL}(p||q)$ is the Kullback-Leibler divergence of $q$ from $p$.\\
+
+\noindent Equation \ref{eq:gradient_descent} describes the gradient descent step to minimise loss function $L(y^{(i)},z^{(i)}(f_\omega))$ with respect to parameter $\omega$. 
+\begin{equation}
+\Delta\omega = \frac{\partial L(y^{(i)},z^{(i)}(f_\omega))}{\partial \omega} = - \sum_{\forall g \in \mathscr{G}_k}\limits\frac{\partial P_{z^{(i)}(f_\omega)}(g)}{\partial \omega}\frac{P_{y^{(i)}}(g)}{P_{z^{(i)}(f_\omega)}(g)}
+\label{eq:gradient_descent}
+\end{equation}
+
+\noindent Algorithm \ref{alg:listnet} shows the pseudo-code of the ListNet training phase.\\
+
 \begin{algorithm}[H]
@@
 \end{algorithm}
-
-\begin{equation}
-\Delta\omega = \frac{\partial L(y^{(i)},z^{(i)}(f_\omega))}{\partial \omega} = - \sum_{\forall g \in \mathscr{G}_k}\limits\frac{\partial P_{z^{(i)}(f_\omega)}(g)}{\partial \omega}\frac{P_{y^{(i)}}(g)}{P_{z^{(i)}(f_\omega)}(g)}
-\label{eq:gradient_descent}
-\end{equation}
 \section{SmoothRank}

### 2014-04-15 report/Chapters/Introduction.tex
hopefull methods
BM25 in introduction verwerkt

@@
-%********************************************************************
-% Appendix
-%*******************************************************
-% If problems with the headers: get headings in appendix etc. right
-%\markboth{\spacedlowsmallcaps{Appendix}}{\spacedlowsmallcaps{Appendix}}
 \section{Motivation and Problem Statement}
@@
 
-Research in the field of ranking models has for a long time been based on manually designed ranking functions and has been an active area of research.
+Research in the field of ranking models has for a long time been based on manually designed ranking functions, such as the well-known BM25 model \cite{Robertson1994}.
 Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.

### 2014-04-15 report/Chapters/LTRMethods.tex
taalfout
BM25 in introduction verwerkt

@@
 \section{ListNet}
-ListNet is a listwise ranking function whose loss function is not directly related to information retrieval evaluation metrics.
-ListNet's loss function is defined using the probability distribution on permutations.
+ListNet \cite{Cao2007} is a listwise ranking function whose loss function is not directly related to information retrieval evaluation metrics.
+ListNet's loss function is defined using a probability distribution on permutations.
 Probability distributions on permutations have been a research topic within the field of probability theory and has been extensively researched.
@@
 \end{algorithm}
+
 \section{SmoothRank}
+SmoothRank \cite{Chapelle2010} is a listwise ranking method that directly optimises an information retrieval evaluation measure by smoothing, that is, approximating the rank position.
+In this section I will use \ac{nDCG} for illustration of SmoothRank, but the same procedure can also be applied to \ac{MAP} or any other information retrieval measure.
 

### 2014-04-15 report/Chapters/LTRMethods.tex
algorithm line numbers + SmoothRank algorithm

@@
 \noindent Algorithm \ref{alg:listnet} shows the pseudo-code of the ListNet training phase.\\
-
+\LinesNumbered
 \begin{algorithm}[H]
  \KwData{training data \{$(x^{(1)},y^{(1)}),(x^{(2)},y^{(2)}),...,(x^{(m)},y^{(m)})$\}}
- Parameter: number of iterations $T$ and learning rate $\eta$\\
+ \KwIn{number of iterations $T$ and learning rate $\eta$}
  Initialize parameter $\omega$\\
@@
 SmoothRank \cite{Chapelle2010} is a listwise ranking method that directly optimises an information retrieval evaluation measure by smoothing, that is, approximating the rank position.
-In this section I will use \ac{nDCG} for illustration of SmoothRank, but the same procedure can also be applied to \ac{MAP} or any other information retrieval measure.
+In this section I will use \ac{nDCG} for illustration of SmoothRank, but the same procedure can also be applied to \ac{MAP} or any other information retrieval measure.\\
+
+\LinesNumbered
+\begin{algorithm}[H]
+ Find an initial solution $w_0$ (by regression of Rank\acs{SVM}).\\
+ Set $w = w_0$ and $\sigma$ to a large value.\\
+ \While{Stopping condition not satisfied}{
+ 	 Starting from $w$, minimize by non-linear conjugate gradient descent:
+ 	 $\lambda||w-w_0||^2 - \sum\nolimits_{q}A_q(w,\sigma)$\\
+ 	 $\sigma = \sigma/2$
+ }
+ \caption{Learning algorithm of SmoothRank, obtained from \cite{Chapelle2010}}
+ \label{alg:smoothrank}
+\end{algorithm}
 

### 2014-04-22 report/Chapters/LTRMethods.tex
TODO list voor omschrijven SmoothRank
softmax activatie functie gefixt

@@
 SmoothRank \cite{Chapelle2010} is a listwise ranking method that directly optimises an information retrieval evaluation measure by smoothing, that is, approximating the rank position.
-In this section I will use \ac{nDCG} for illustration of SmoothRank, but the same procedure can also be applied to \ac{MAP} or any other information retrieval measure.\\
+In this section I will use \ac{nDCG} for illustration of SmoothRank, but the same procedure can also be applied to \ac{MAP} or any other information retrieval measure.
+The smoothing function used in SmoothRank based on the softmax activation function \cite{Bridle1990} that is often used in neural networks, which is formally defined as shown in \ref{eq:softmax_algorithm}.
+\begin{equation}
+p_i = \frac{e^{f_i/\sigma}}{\sum\nolimits_{j=1}^{m}e^{f_j/\sigma}}
+\label{eq:softmax_algorithm}
+\end{equation}
+
+% TODO: 1) Smoothing framework uitleggen en toepassen op nDCG 2) sigma trade-off behandelen 3)Nonlinear conjugate gradient descent uitleggen 4) startin point en regularization uitleggen 5) Annealing process uitleggen %
 

### 2014-04-22 report/Chapters/LTRMethods.tex
smoothrank uitgelegd :)

@@
 In this section I will use \ac{nDCG} for illustration of SmoothRank, but the same procedure can also be applied to \ac{MAP} or any other information retrieval measure.
-The smoothing function used in SmoothRank based on the softmax activation function \cite{Bridle1990} that is often used in neural networks, which is formally defined as shown in \ref{eq:softmax_algorithm}.
+The smoothing function used in SmoothRank based on the softmax activation function \cite{Bridle1990} that is often used in neural networks, which is formally defined as shown in Equation \ref{eq:softmax_algorithm}.
 \begin{equation}
@@
 
-% TODO: 1) Smoothing framework uitleggen en toepassen op nDCG 2) sigma trade-off behandelen 3)Nonlinear conjugate gradient descent uitleggen 4) startin point en regularization uitleggen 5) Annealing process uitleggen %
+where $\sigma$ is the smoothing parameter.
+Chapelle et al \cite{Chapelle2010} apply the softmax activation function to the \ac{nDCG} formula and thereby introduce a soft version of indicator variable $h_{i,j}$ as formulated in Equation \ref{eq:soft_ndcg}.
+
+\begin{equation}
+h_{i,j} = e^{-\frac{(f(x_i))-(f(x_{d(j)}))^2}{\sigma}}\Big/\sum\nolimits_{k=1}^{m}e^{-\frac{(f(x_k))-f(x_{d(j)})^2}{\sigma}}
+\label{eq:soft_ndcg}
+\end{equation}
+
+It can be shown that the derivative of the smoothed \ac{nDCG} version shown in Equation \ref{eq:soft_ndcg} and the smoothed versions of other \ac{IR} metrics can be calculated in $\mathcal{O}(m^2)$, which enable fast gradient descent optimisation.
+The optimisation step in SmoothRank uses the nonlinear conjugate gradient method  with Polak-Ribiere update \cite{Shewchuk1994}, which is a type of gradient descent method.
+The optimisation method is prone to local optima, which is alleviated by adding a good starting point and regulariser.\\
+
+The starting point in SmoothRank is set to either the solution of a simple linear regression, or alternatively to the solution of Rank\acs{SVM}.
+Since this starting point is expected to already be a good solution, a regulariser term is added to the SmoothRank objective function to prevent the solution from deviating to much from the starting point.
+The regularised smooth objective function is formulated as $\lambda||w-w_0||^2$ where $\lambda$ is a hyperparameter tuned on the validation set, and $w_0$ is the starting point solution.\\
+
+The choice of the smoothing parameter $\sigma$ in Equation \ref{eq:soft_ndcg} is important, because a too small value makes the function more non-smooth and therefore harder to optimise, while a too large value results in a optimisation function that substantially differs from the optimal rankings.
+SmoothRank uses an annealing method where the optimisation procedure starts with a large $\sigma$ and iteratively reduces it by dividing it by 2 at each step.
+Algorithm \ref{alg:smoothrank} shows the algorithm that summarises all steps of the SmoothRank method.
 

### 2014-04-23 report/Chapters/LTRMethods.tex
fenchelrank algorithm + nieuwe math functies

@@
 SmoothRank uses an annealing method where the optimisation procedure starts with a large $\sigma$ and iteratively reduces it by dividing it by 2 at each step.
-Algorithm \ref{alg:smoothrank} shows the algorithm that summarises all steps of the SmoothRank method.
-
+Algorithm \ref{alg:smoothrank} shows the algorithm that summarises all steps of the SmoothRank method.\\
 \LinesNumbered
@@
 \section{FenchelRank}
+FenchelRank \cite{Lai2013} learning method that addresses the sparse Learning-to-Rank problem, which is the problem of learning a ranking function with only a few non-zero coefficients with respect to the input features.
+FenchelRank is based on the theory of Fenchel Duality \cite{Rifkin2007} and uses the genetic algorithm framework proposed by Shalev-Shwartz and Singer \cite{Shalev-Shwartz2010}.
 
+FenchelRank optimises the objective function shown in Equation \ref{eq:pairwise_l1_loss} that is equivalent to the standard $\ell_1$-norm regularised pairwise loss function.
+
+\begin{equation}
+\min_w G(w) = \min_w I_{||w||_{1} \le 1}(w) + \frac{r^2}{p} \sum\limits_{i=1}^{p}\max(0,\frac{1}{r}-(Kw)_i)^2
+\label{eq:pairwise_l1_loss}
+\end{equation}
+
+The objective function in Equation \ref{eq:pairwise_l1_loss} is not differentiable everywhere because of its $\ell_1$-norm regularisation term.
+
+Let $D(w) = -G(w)$, then the objective function stated in \ref{eq:pairwise_l1_loss} can be rewritten as shown in Equation \ref{eq:fenchel_dual}.
+\begin{equation}
+\max_w D(w) = \max_w I_{||w||_{1} \le 1}(w) - \frac{r^2}{p} \sum\limits_{i=1}^{p}\max(0,\frac{1}{r}-(Kw)_i)^2
+\label{eq:fenchel_dual}
+\end{equation}
+
+\LinesNumbered
+\begin{algorithm}[H]
+ \KwData{pairwise training data matrix $K$}
+ \KwIn{desired accuracy $\epsilon$, maximum number of iterations $T$ and the radius $r$ of the $\ell1$ ball.}
+ Initialize: $w_1$ = $0_m$\\
+ \For{$t\leftarrow 1$ \KwTo $T$}{
+ 	//check if the early stopping criterion is satisfied\\
+ 	\If{$||g_{t}||_{\infty} + \langle d_{t}, -Kw_{t} \rangle \le \epsilon$}{
+ 	// Here $d_{t} = \nabla f^{*}(-Kw_{t}) = \frac{\partial f^{*}(-Kw)}{\partial(Kw)}|w=w_{t}$ and $g_{t} = d_{t}^{T}K$\\
+ 	return $w_{t}$ as ranking predictor $w$
+ 	}
+ //greedily choose a feature to update\\
+ Choose $j_{t} = \argmax_{j}|(g_t)_j)|$\\
+ //compute an appropriate step size\\
+ Let $\mu_t = \argmax_{0 \le \mu_{t} \le 1} D((1-\mu_{t})w_{t} + \mu_{t}\sign((g_{t})_{j_{t}})e^{j_{t}})$\\
+ //update the model with the chosen feature and step size\\
+ Update $w_{t+1}=(1-\mu_{t})w_{t} + \mu_{t}\sign((g_{t})j_{t})e^{j_{t}}$
+ }
+ return $w_{T}$ as ranking predictor for $w$
+ \caption{Learning algorithm of FenchelRank, obtained from \cite{Lai2013}}
+ \label{alg:fenchelrank}
+\end{algorithm}
 \section{FSMRank}
@@
 \section{LRUF}
+Learning automaton.

### 2014-04-23 report/Chapters/LTRMethods.tex
werk aan fenchelrank

@@
 FenchelRank \cite{Lai2013} learning method that addresses the sparse Learning-to-Rank problem, which is the problem of learning a ranking function with only a few non-zero coefficients with respect to the input features.
-FenchelRank is based on the theory of Fenchel Duality \cite{Rifkin2007} and uses the genetic algorithm framework proposed by Shalev-Shwartz and Singer \cite{Shalev-Shwartz2010}.
+FenchelRank is based on the theory of Fenchel Duality \cite{Rifkin2007} and uses the genetic algorithm framework proposed by Shalev-Shwartz and Singer \cite{Shalev-Shwartz2010}.\\
 
@@
 
-The objective function in Equation \ref{eq:pairwise_l1_loss} is not differentiable everywhere because of its $\ell_1$-norm regularisation term.
+\noindent where $I_{C}(w)$ is a function that is 0 if condition $C$ is satisfied, and $\infty$ otherwise.
+$m$ is the dimension of the data, $p$ is the number of pairs in set $P$, $K$ is a matrix in $\mathbb{R}^{p \times m}$ that contains pairwise information.
+The objective function in Equation \ref{eq:pairwise_l1_loss} is not differentiable everywhere because of its $\ell_1$-norm regularisation term.\\
 

### 2014-04-23 report/Chapters/LTRMethods.tex
fenchelrank wiskunde mooier

@@
 \noindent where $I_{C}(w)$ is a function that is 0 if condition $C$ is satisfied, and $\infty$ otherwise.
-$m$ is the dimension of the data, $p$ is the number of pairs in set $P$, $K$ is a matrix in $\mathbb{R}^{p \times m}$ that contains pairwise information.
+$m$ is the dimension of the data, $p$ is the number of comparable object pairs, $K$ is a matrix in $\mathbb{R}^{p \times m}$ that contains pairwise information.
 The objective function in Equation \ref{eq:pairwise_l1_loss} is not differentiable everywhere because of its $\ell_1$-norm regularisation term.\\
@@
  	\If{$||g_{t}||_{\infty} + \langle d_{t}, -Kw_{t} \rangle \le \epsilon$}{
- 	// Here $d_{t} = \nabla f^{*}(-Kw_{t}) = \frac{\partial f^{*}(-Kw)}{\partial(Kw)}|w=w_{t}$ and $g_{t} = d_{t}^{T}K$\\
+ 	// Here $d_{t} = \nabla f^{*}(-Kw_{t}) = \frac{\partial f^{*}(-Kw)}{\partial(Kw)}|w=w_{t}$, \\
+ 	// $\langle x,y \rangle$ represents the inner products of vectors $x$ and $y$, and\\
+ 	// $g_{t} = d_{t}^{T}K$\\
  	return $w_{t}$ as ranking predictor $w$

### 2014-04-24 report/Chapters/LTRMethods.tex
Fenchel Duality Theorem uitgelegd en foutjes uit hoofdstuk gehaald

@@
 \begin{algorithm}[H]
- Find an initial solution $w_0$ (by regression of Rank\acs{SVM}).\\
+ Find an initial solution $w_0$ (by regression or Rank\acs{SVM}).\\
  Set $w = w_0$ and $\sigma$ to a large value.\\
@@
 \section{FenchelRank}
-FenchelRank \cite{Lai2013} learning method that addresses the sparse Learning-to-Rank problem, which is the problem of learning a ranking function with only a few non-zero coefficients with respect to the input features.
-FenchelRank is based on the theory of Fenchel Duality \cite{Rifkin2007} and uses the genetic algorithm framework proposed by Shalev-Shwartz and Singer \cite{Shalev-Shwartz2010}.\\
-
-FenchelRank optimises the objective function shown in Equation \ref{eq:pairwise_l1_loss} that is equivalent to the standard $\ell_1$-norm regularised pairwise loss function.
-
+FenchelRank \cite{Lai2013} is a ranking method that addresses the sparse Learning-to-Rank problem, which is the problem of learning a ranking function with only a few non-zero coefficients with respect to the input features.
+FenchelRank is based on the theory of Fenchel Duality \cite{Rifkin2007} and uses the genetic algorithm framework proposed by Shalev-Shwartz and Singer \cite{Shalev-Shwartz2010}.
+FenchelRank optimises the objective function shown in Equation \ref{eq:pairwise_l1_loss} which is equivalent to the standard $\ell_1$-norm regularised pairwise loss function.
 \begin{equation}
@@
 The objective function in Equation \ref{eq:pairwise_l1_loss} is not differentiable everywhere because of its $\ell_1$-norm regularisation term.\\
+
+Fenchel's duality theorem \cite{Rifkin2007}, defined in Equation \ref{eq:fenchel_duality_theorem}, provides a way to approximate an optimisation problem by instead solving the Fenchel dual of the optimisation problem.
+\begin{equation}
+\min_x(f(x)-g(x)) = \max_p(f_*(p)-f^*(p))
+\label{eq:fenchel_duality_theorem}
+\end{equation}
+\noindent where $f^*$ is the convex conjugate of $f$ and $g_*$ is the concave conjugate of $g$.
 

### 2014-04-24 report/Chapters/LTRMethods.tex
fenchelrank geupdate

@@
 
-Let $D(w) = -G(w)$, then the objective function stated in \ref{eq:pairwise_l1_loss} can be rewritten as shown in Equation \ref{eq:fenchel_dual}.
+\noindent Let $D(w) = -G(w)$.
+This allows the FenchelRank objective function (Equation \ref{eq:pairwise_l1_loss}) to be rewritten using Fenchel's duality theorem (Equation \ref{eq:fenchel_duality_theorem}) to the form shown in Equation \ref{eq:fenchel_dual}.
 \begin{equation}

### 2014-04-24 report/Chapters/LTRMethods.tex
fenchelrank staat op papier :)

@@
 FenchelRank \cite{Lai2013} is a ranking method that addresses the sparse Learning-to-Rank problem, which is the problem of learning a ranking function with only a few non-zero coefficients with respect to the input features.
-FenchelRank is based on the theory of Fenchel Duality \cite{Rifkin2007} and uses the genetic algorithm framework proposed by Shalev-Shwartz and Singer \cite{Shalev-Shwartz2010}.
+FenchelRank is based on the theory of Fenchel Duality \cite{Rifkin2007} and uses a generic algorithm framework proposed by Shalev-Shwartz and Singer \cite{Shalev-Shwartz2010}.
 FenchelRank optimises the objective function shown in Equation \ref{eq:pairwise_l1_loss} which is equivalent to the standard $\ell_1$-norm regularised pairwise loss function.
@@
 $m$ is the dimension of the data, $p$ is the number of comparable object pairs, $K$ is a matrix in $\mathbb{R}^{p \times m}$ that contains pairwise information.
-The objective function in Equation \ref{eq:pairwise_l1_loss} is not differentiable everywhere because of its $\ell_1$-norm regularisation term.\\
-
-Fenchel's duality theorem \cite{Rifkin2007}, defined in Equation \ref{eq:fenchel_duality_theorem}, provides a way to approximate an optimisation problem by instead solving the Fenchel dual of the optimisation problem.
+The objective function in Equation \ref{eq:pairwise_l1_loss} is not differentiable everywhere because of its $\ell_1$-norm regularisation term.
+Fenchel's duality theorem \cite{Rifkin2007}, defined in Equation \ref{eq:fenchel_duality_theorem}, provides a way to approximate the optimal value of this non-differentiable optimisation problem by instead solving the Fenchel dual of the optimisation problem.
 \begin{equation}
@@
 \end{equation}
+
+Algorithm \ref{alg:fenchelrank} shows the FenchelRank training algorithm to optimise the Fenchel dual of the pairwise loss function.\\
 

### 2014-04-28 report/Chapters/LTRMethods.tex
begin fsmrank algorithm
fsmrank algoritme ingevoerd
fsmrank intro

@@
-Algorithms and details of the well-performing Learning-to-Rank methods as selected in the afore-going part are presented and explained in this part.
+Algorithms and details of the well-performing Learning-to-Rank methods as selected in the afore-going part are presented and explained in this chapter.
 
@@
  	\If{$||g_{t}||_{\infty} + \langle d_{t}, -Kw_{t} \rangle \le \epsilon$}{
- 	// Here $d_{t} = \nabla f^{*}(-Kw_{t}) = \frac{\partial f^{*}(-Kw)}{\partial(Kw)}|w=w_{t}$, \\
+ 	//here $d_{t} = \nabla f^{*}(-Kw_{t}) = \frac{\partial f^{*}(-Kw)}{\partial(Kw)}|w=w_{t}$, \\
  	// $\langle x,y \rangle$ represents the inner products of vectors $x$ and $y$, and\\
@@
 \section{FSMRank}
+Lai et al.
+\cite{Lai2013c} observed that existing feature selection methods in Learning-to-Rank all follow a two-stage paradigm: 1) select subset of features from the original features, and 2) learn a ranking model based on the selected features.
+It is stated as a limitation of this paradigm that the selected features in the first step are not necessarily the optimal features for the second stage where the ranking model is build.
+FSMRank \cite{Lai2013c} addresses this limitation by using a joint convex optimisation problem that minimises ranking errors while simultaneously selecting a subset of features.
+\LinesNumbered
+\begin{algorithm}[H]
+ \KwData{training set $S = {(q_k,\hat{X}^{(q_k)},Y^(q_k)}_{k=1}^{n}$}
+ \KwIn{$\lambda_1, \lambda_2, T \text{ and } L_0$}
+ Initialize: $w_0=z_1=0, \alpha_1=1, \gamma=2, L=L_0/\gamma^{10}$\\
+ \For{$t\leftarrow 1$ \KwTo $T$}{
+ 	Let $g = l^{'}(z_t)$\\
+ 	\While{true}{
+ 		//projection step\\
+ 		$w_t= \argmin_{w_t:w_t\in \mathbb{R}_{+}^{2d}} Q(w_t,z_t)$ s\\
+ 		\If{$l(w_t) \le l(z_t) + \langle l^{'}(z_t),w_t-z_t \rangle +\frac{L}{2}||w_t-z_t||^2$}{
+ 			break\\
+ 		}
+ 		$L = \lambda L$ 		
+ 	}
+ 	\If{$\frac{|F(w_t)-F(w_{t-1})|}{|F(w_{t-1})|} \le \epsilon_s$}{
+ 		//early stopping criterion\\
+ 		break
+ 	}
+ 	$\alpha_{t+1}=\frac{1+\sqrt{1+4\alpha_t^2}}{2}$\\
+ 	$z_{t+1} = w_t + \frac{\alpha_t-1}{\alpha_t+1}(w_t-w_{t-1})$ 
+ }
+
+ \caption{Learning algorithm of FSMRank, obtained from \cite{Lai2013c}}
+ \label{alg:fsmrank}
+\end{algorithm}
 

### 2014-04-28 report/Chapters/LTRMethods.tex
nesterov's accelerated gradient descent uitgelegd

@@
 It is stated as a limitation of this paradigm that the selected features in the first step are not necessarily the optimal features for the second stage where the ranking model is build.
-FSMRank \cite{Lai2013c} addresses this limitation by using a joint convex optimisation problem that minimises ranking errors while simultaneously selecting a subset of features.
+FSMRank \cite{Lai2013c} addresses this limitation by using a joint convex optimisation problem that minimises ranking errors while simultaneously selecting a subset of features.\\
+
+FSMRank uses an extended version of gradient descent optimisation, proposed by Yuri Nesterov, that enables faster convergence for convex problems \cite{Nesterov2004}.
+Nesterov's accelerated gradient descent can guarantee an $\epsilon$-accurate solution in at most $T$ iterations where $\epsilon=\mathcal{O}(1/T^2)$.
+
 \LinesNumbered

### 2014-04-28 report/Chapters/LTRMethods.tex
FSMRank optimisation function

@@
 FSMRank uses an extended version of gradient descent optimisation, proposed by Yuri Nesterov, that enables faster convergence for convex problems \cite{Nesterov2004}.
-Nesterov's accelerated gradient descent can guarantee an $\epsilon$-accurate solution in at most $T$ iterations where $\epsilon=\mathcal{O}(1/T^2)$.
+Nesterov's accelerated gradient descent can guarantee an $\epsilon$-accurate solution in at most $T$ iterations where $\epsilon=\mathcal{O}(1/T^2)$.\\
+
+Let $S = {(q_k, \hat{X}^{(q_k), Y^{(q_k)}})}_{k=1}^n$ be a training set with queries $q_k$, corresponding retrieval objects $\hat{X}^{(q_k)}$, and corresponding relevance labels $Y^{(q_k)}$.
+Let $||x||_1$ be the $l_1$-norm regularisation function that is defined as $||x||_1 = \sum\nolimits_i |x_i|$.
+Let $\oslash$ be the element-wise division operator.
+The convex joint optimisation function in FSMRank is defined as shown in Equation \ref{eq:fsmrank_optimisation_function}.
+
+\begin{equation}
+\min_{\hat{w}} \frac{\lambda_1}{2} \hat{w}^T \hat{A}\hat{w} + \lambda_2 ||\hat{w}\oslash\hat{s}||_1 + f(\hat{w}, (\hat{X}^{(q_k)}, Y^{(q_k)})_{k=1}^n)
+\label{eq:fsmrank_optimisation_function}
+\end{equation}
 

### 2014-04-28 report/Chapters/LTRMethods.tex
git status

@@
 FenchelRank optimises the objective function shown in Equation \ref{eq:pairwise_l1_loss} which is equivalent to the standard $\ell_1$-norm regularised pairwise loss function.
+The $\ell_1$-norm regularisation function in this equation is represented by $||x||_1$ and is defined as $||x||_1=\sum\nolimits_i|x_i|$.
 \begin{equation}
@@
 
-Algorithm \ref{alg:fenchelrank} shows the FenchelRank training algorithm to optimise the Fenchel dual of the pairwise loss function.\\
+Algorithm \ref{alg:fenchelrank} shows the FenchelRank training algorithm to optimise the Fenchel dual of the pairwise loss function.
+The $||x||_infty$ term in this algorithm represents an $\ell_\infty$-norm regularisation term and is defined as $||x||_\infty=\max_{i}|x_i|$.\\
 
@@
 The convex joint optimisation function in FSMRank is defined as shown in Equation \ref{eq:fsmrank_optimisation_function}.
-
 \begin{equation}
@@
 \end{equation}
+\noindent $\hat{A}$ is a $d \times d$ similarity matrix that contains similarity scores between features.
+The first term in Equation \ref{eq:fsmrank_optimisation_function} applies a penalty on redundancy of large weighted features by using the $\hat{A}$ matrix.
+The well-known Pearson correlation coefficient is used to calculate similarity matrix $\hat{A}$ based on the values for the features in the training data.
+$\lambda_1$ is a hyper-parameter of the model and can be used to set the weight of the similarity penalty term.\\
+
+The second term of Equation \ref{eq:fsmrank_optimisation_function} uses a $\ell_1$-norm regularisation term to select the effective features from the feature set.
+$\lambda_2$ is a hyper-parameter of the model and can be used to set the weight of the regularisation term.\\
+
+The last term represents the loss function in terms of ranking errors.
+The loss function $f$ can in theory be any convex loss function, but Lai et al.
+\cite{Lai2013c} used a squared hinge loss function for their evaluation measurements.
+
+Algorithm \ref{alg:fsmrank} shows the steps of the FSMRank algorithm.
+In this algorithm an extended version of Nesterov's accelerated gradient method is used that can handle optimisation problems in the form of $\min_w \l(w) + r(w)$ where $l(w)$ is a convex function with Lipschitz gradient and $r(w)$ is convex, but non-smooth.
+The optimization function of Equation \ref{eq:fsmrank_optimisation_function} is reformulated to match this form as presented in Equation \ref{eq:fsmrank_optimisation_reformulated}.
+\begin{equation}
+\argmin_{w_t:w_t\in \mathbb{R}} Q(w_t,z_t) = \lambda_2 \sum\nolimits_{i=1}^{2d}\frac{w_i}{s_i}+\langle l^{'}(z_t),w_t-z_t \rangle+\frac{L}{2}||w_t-z_t||^2
+\label{eq:fsmrank_optimisation_reformulated}
+\end{equation}
+\noindent where $Q(w_t,z_t)$ is a combination of the non-smooth part $r(w_t)$ and a quadratic approximation of the smooth part $l(w_t)$.
+$lambda_1$, $lambda_2$ and Lipschitz constant $L_0$ are input parameters of the algorithm and can be optimised through cross-validation or on a validation set.\\
 
@@
  		//projection step\\
- 		$w_t= \argmin_{w_t:w_t\in \mathbb{R}_{+}^{2d}} Q(w_t,z_t)$ s\\
+ 		$w_t= \argmin_{w_t:w_t\in \mathbb{R}_{+}^{2d}} Q(w_t,z_t)$\\
  		\If{$l(w_t) \le l(z_t) + \langle l^{'}(z_t),w_t-z_t \rangle +\frac{L}{2}||w_t-z_t||^2$}{

### 2014-04-28 report/Chapters/LTRMethods.tex
todo-lijst + formule LRUF
begin uitleg LRUF

@@
 Learning automaton.
+TODO:
+\begin{enumerate}
+\item Variable structure learning automata uitleggen
+\item $L_{R-p}, L_{R-\epsilon p} \text{en} L_{R-I}$ uitleggen
+\item $\epsilon$-optimality en absolutely expediency uitleggen en melden dat $L_{R-I}$ deze eigenschappen bezit
+\item LRUF-tuple uitleggen
+\end{enumerate}
+
+\begin{equation}
+p_j(k+1)=	\begin{cases}	p_j(k)+a[1-p_j(k)] 	&\mbox{if } j=i \\ 
+							(1-a)p_j(k) 		&\mbox{otherwise} 
+			\end{cases} 
+\label{eq:lruf_reward_update}
+\end{equation}
+
+\begin{equation}
+g_i=\frac{1}{N_i}\sum\nolimits_{d_{i}^{j}\in \underline{f}_{i}} a(r_{i}^{j})^{-1}
+\label{eq:lruf_average_relevance}
+\end{equation}
+
+\begin{algorithm}
+ \KwData{Query $q_i$, Number of results $n_i$}
+ Assume:
+ 	Let $A_i$ be the learning automaton corresponding to query $q_i$ with action-set $\alpha_i$\\
+ 	Action $\alpha_i^j \in \alpha_i$ is associated with document $d_i^j \in \underline{d_i}$\\
+ 	Let $k$ denote the stage number\\
+ 	Let $G$ be the total relevance score\\
+ Initialise: $k\leftarrow 1$, $T_i\leftarrow 0$\\
+ \While{$k \le n_i$}{
+ 	$A_i$ chooses one of its actions (e.g. $a_i^k$) at random\\
+ 	Document $d_i^j$ corresponding to selected action $\alpha_i^j$ is ranked at $K^{th}$ position of $\underline{R_i}$\\
+ 	Configuration of $A_i$ is updated by disabling action $\alpha_i^j$\\
+ 	$k\leftarrow k+1$
+ }
+ Ranking $\underline{R_i}$ is shown to the user\\
+ $N_i\leftarrow 0,\underline{f_i}\leftarrow \emptyset,G\leftarrow 0$\\
+ \Repeat{query session is expired}{
+ 	\For{every document $d_i^j$ visited by user}{
+ 		$\underline{f_i}\leftarrow \underline{f_i}+{d_i^j}$\\
+ 		$G\leftarrow G+a*(r_i^j)^{-1}$
+ 	}
+ 	$N_i\leftarrow N_i +1$
+ }
+ $g_i \leftarrow \frac{G}{N_i}$\\
+ Configuration of $A_i$ is updated by re-enabling all disabled actions\\
+ \If{$g_i\ge T_i$}{
+ 	Reward the actions corresponding to all visited documents by Equation \ref{eq:lruf_reward_update}\\
+ 	$T_i\leftarrow g_i$
+ }
+ \For{$\forall \alpha_i^j \in \underline{\alpha_i}$}{
+ 	\If{$p_i^j < T_{\epsilon}$}{
+ 		$d_i^j$ is replaced by another document of the searched results
+ 	}
+ }
+ Output $\underline{R_i}$
+ \caption{Learning algorithm of LRUF, obtained from \cite{Torkestani2012b}}
+ \label{alg:lruf}
+\end{algorithm}

### 2014-04-29 report/Chapters/LTRMethods.tex
LRUF methode verder uitgewerkt

@@
 \section{LRUF}
-Learning automaton.
-TODO:
-\begin{enumerate}
-\item Variable structure learning automata uitleggen
-\item $L_{R-p}, L_{R-\epsilon p} \text{en} L_{R-I}$ uitleggen
-\item $\epsilon$-optimality en absolutely expediency uitleggen en melden dat $L_{R-I}$ deze eigenschappen bezit
-\item LRUF-tuple uitleggen
-\end{enumerate}
+LRUF \cite{Torkestani2012b} is Learning-to-Rank method based on the theory of learning automata.
+Learning automata are adaptive decision-making units that learn the optimal set of actions through repeated interactions with its environment.
+Variable structure learning automata can be defined as a triple $<\beta,\alpha,L>$, with $\beta$ being the set of inputs, $\alpha$ the set of actions.
+$L$, the learning algorithm ,is a recurrence relation that modifies the action probability vector of the actions in $\alpha$.\\
 
+Three well-known learning algorithms $L$ used in variable structure learning automata are \emph{linear reward-penalty} ($L_{R-P}$), \emph{linear reward-$epsilon$-penalty} ($L_{R-\epsilon P}$) and \emph{linear reward-inaction} ($L_{R-I}$).
+LRUF uses the $L_{R-I}$ learning algorithm, which updates the action probability vector following Equation \ref{eq:lruf_reward_update} when the selected action $a_i(k)$ is rewarded by the environment.
 \begin{equation}
@@
 \end{equation}
+\noindent where $i$ and $j$ are indices of action in $\alpha$ and $p(k)$ is the probability vector over the action set at rank $k$. $\alpha$ $a$ is the learning rate of the model.\\
 
+A variable action-set learning automaton is an automaton in which the number of available actions change over time.
+It has been shown that a combination of a variable action-set learning automaton with the $L_{R-I}$ learning algorithm is absolutely expedient and $\epsilon$-optimal \cite{Thathachar1987}, which means that it is guaranteed to approximate the optimal solution to some value $\epsilon$ and each update step is guaranteed not to decrease performance of the model.
+A variable-action set learning automata has a finite set of $r$ actions $\alpha={\alpha_1,...,\alpha_r}$.
+$A$ is defined as the power set of $\alpha$, $A={A_1,..,A_m}$ with $m = 2^{r}-1$.
+$\psi(k)$ is a probability distribution over $A$ such that $\psi(k)=p(A(k)=A_i|A_i\in A, 1\le i \le 2^r-1)$.
+$\hat{p}_i(k)$ is the probability of choosing action $\alpha_i$ given that action subset $A(k)$ has already been selected and $\alpha_i \in A(k)$.\\
+
+LRUF uses an optimisation problem that can be illustrated as a quintuple $<q_i,A_i,\underline{d_i},\underline{R_i},\underline{f_i}>$, where $q_i$ is a submitted query, $A_i$ is a variable action-set learning automaton, $\underline{d_i}$ is a set of documents associated with $q_i$, and $\underline{R_i}={r_i^j|\forall d_i^j \in \underline{d_i}}$ is a ranking function that assigns rank $r_i^j$ to each document.
+$\underline{f_i}$ is a feedback set used for the update step described in Equation \ref{eq:lruf_reward_update}.
+For each rank $k$ the learning automaton $A_i$ chooses one of its actions following its probability vector, jointly forming $\underline{R_i}$.
+LRUF translates the feedback in $\underline{f_i}$ to an understandable value to use it to converge the action probability vector in optimal configuration.
+Therefore LRUF defines a $g_i: \underline{f_i} \rightarrow \mathbb{R}^{+}$ to be a mapping from the feedback set into a positive real number.
+The LRUF mapping function $g_i$ computes the average relevance score of ranking $R_i$ based on $f_i$ and is defined as shown in Equation \ref{eq:lruf_average_relevance}.
 \begin{equation}
-g_i=\frac{1}{N_i}\sum\nolimits_{d_{i}^{j}\in \underline{f}_{i}} a(r_{i}^{j})^{-1}
+g_i=\frac{1}{N_i}\sum\nolimits_{d_{i}^{j}\in \underline{f_i}} a(r_{i}^{j})^{-1}
 \label{eq:lruf_average_relevance}

### 2014-04-29 report/Chapters/LTRMethods.tex
LRUF uitgelegd :)

@@
 \end{equation}
+\noindent in which $N_i$ refers to the size of feedback set $\underline{f_i}$, $r_i^j$ is the rank of document $d_i^j$ in $R_i$ and, again, $a$ denotes the learning rate.\\
+
+Initially, before any feedback, the action probability factors are initialised with equal probability.
+
+In case the document set of the search engine changes, i.e.
+new documents are indexed or old documents are removed, LRUF has an Increase Action-Set Size (IAS) and a Reduce Action-Set Size (RAS) procedure defined to adapt to the new document set without needing a complete retraining of the model.
+Because this thesis focusses on model accuracy and scalability the IAS and RAS procedure of LRUF will not be explained in further detail.
+
+Torkestani \cite{Torkestani2012b} does not provide pseudo code specifically for the training phase of the LRUF algorithm.
+Instead an algorithm including both ranking and automaton update is presented, and included in Algorithm \ref{alg:lruf}.
+Initial training of the algorithm can be performed by using the training data relevance labels in the feedback set of the algorithm.
 

### 2014-04-29 report/Chapters/CrossBenchmarkComparison.tex
OHSUMED LETOR 2.0 != LETOR 3.0, sampling
self-review verwerkt tm fenchelrank

@@
+\label{chap:cross_comparison}
 As we have seen in Chapter~\ref{chap:benchmark_results}, the evaluation of Learning-to-Rank methods is spread over the several benchmark datasets.
@@
 BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
-Best Single Feature &  & \cite{Gomes2013} & MHR & L2, \cite{Qin2007} &  \\ 
+Best Single Feature &  & \cite{Gomes2013} & \acs{MHR} & L2, \cite{Qin2007} &  \\ 
 BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 

### 2014-04-29 report/Chapters/LTRMethods.tex
fix LRUF algo caption
self-review verwerkt tm fenchelrank

@@
-Algorithms and details of the well-performing Learning-to-Rank methods as selected in the afore-going part are presented and explained in this chapter.
+The learning algorithms of the well-performing Learning-to-Rank methods selected in Chapter \ref{chap:cross_comparison} are presented and explained in this Chapter.
+The Learning-to-Rank methods will be discussed in order of an increasing degree of certainty and decreasing ranking accuracy, as concluded in Chapter \ref{chap:cross_comparison}.
 
@@
 ListNet's loss function is defined using a probability distribution on permutations.
-Probability distributions on permutations have been a research topic within the field of probability theory and has been extensively researched.
-One of the most well-known permutation probability distributions in the field, the Plackett-Luce model \cite{Plackett1975,Luce1959}, is used in ListNet.
+Probability distributions on permutations have been a research topic within the field of probability theory which has been extensively researched.
+ListNet uses the Plackett-Luce model \cite{Plackett1975,Luce1959}, which is one of the most well-known permutation probability distributions.
 The Plackett-Luce model defines the probability of all possible permutations $\pi$, given all document ranking scores $S$, as shown in Equation \ref{eq:plackett_luce}.
@@
 \end{equation}
-ListNet uses Gradient Descent to optimise neural network such that its Cross Entropy loss compared to the Plackett-Luce distribution over the ground truth is minimal.
-Note that some sources, including Liu \cite{Liu2007}, describe ListNet as using KL divergence as loss function.
-KL divergence and Cross-entropy are however identical up to an additive constant when comparing distribution $q$ against a fixed reference distribution $p$, since the cross-entropy definition can be formulated as in Equation \ref{eq:cross_entropy}.
+ListNet uses Gradient Descent to optimise a neural network such that its Cross Entropy loss compared to the Plackett-Luce distribution over the training data relevance labels is minimal.
+Note that some sources, including Liu \cite{Liu2007}, describe ListNet as using \ac{KL divergence} as loss function.
+\ac{KL divergence} and Cross Entropy are however identical up to an additive constant when comparing distribution $q$ against a fixed reference distribution $p$.
+This follows from the definition of Cross Entropy Loss, which is shown in Equation \ref{eq:cross_entropy}.
 \begin{equation}
@@
 \end{equation}
-where $H(p)$ is the entropy of $p$ and $D_{KL}(p||q)$ is the Kullback-Leibler divergence of $q$ from $p$.\\
+where $H(p)$ is the entropy of $p$ and $D_{KL}(p||q)$ is the \ac{KL divergence} of $q$ from $p$.\\
 
@@
  	\For{$i\leftarrow 1$ \KwTo $m$}{
- 		Input $x^{(i)}$ of query $q^{(i)}$ to Neural Network and compute score list $z^{(i)}(f_\omega)$ with current $\omega$.\\
+ 		Input $x^{(i)}$ of query $q^{(i)}$ to Neural Network and, for the current value of $\omega$, compute score list $z^{(i)}(f_\omega)$.\\
  		Compute gradient $\Delta\omega$ using Eq. (\ref{eq:gradient_descent}).\\
@@
 \section{SmoothRank}
-SmoothRank \cite{Chapelle2010} is a listwise ranking method that directly optimises an information retrieval evaluation measure by smoothing, that is, approximating the rank position.
-In this section I will use \ac{nDCG} for illustration of SmoothRank, but the same procedure can also be applied to \ac{MAP} or any other information retrieval measure.
-The smoothing function used in SmoothRank based on the softmax activation function \cite{Bridle1990} that is often used in neural networks, which is formally defined as shown in Equation \ref{eq:softmax_algorithm}.
+SmoothRank \cite{Chapelle2010} is a listwise ranking method that, in contrast to ListNet, directly optimises an information retrieval evaluation measure.
+SmoothRank enables direct optimisation of the non-convex and discontinuous evaluation measure by smoothing, that is, approximating the rank position.
+In this section, I apply the SmoothRank approximation method to the \ac{nDCG} evaluation measure for illustration, but the same procedure can be applied to \ac{MAP} or any other information retrieval measure.
+The smoothing function used in SmoothRank is based on the softmax activation function \cite{Bridle1990}, which is often used in neural networks.
+The softmax function is formally defined as shown in Equation \ref{eq:softmax_algorithm}.
 \begin{equation}
@@
 where $\sigma$ is the smoothing parameter.
-Chapelle et al \cite{Chapelle2010} apply the softmax activation function to the \ac{nDCG} formula and thereby introduce a soft version of indicator variable $h_{i,j}$ as formulated in Equation \ref{eq:soft_ndcg}.
+Chapelle et al.
+\cite{Chapelle2010} apply this softmax function to the \ac{nDCG} formula and hereby introduce a soft version of indicator variable $h_{i,j}$ through the computation shown in Equation \ref{eq:soft_ndcg}.
 
@@
 The optimisation step in SmoothRank uses the nonlinear conjugate gradient method  with Polak-Ribiere update \cite{Shewchuk1994}, which is a type of gradient descent method.
-The optimisation method is prone to local optima, which is alleviated by adding a good starting point and regulariser.\\
+This optimisation method is prone to local optima, which is alleviated by adding a pre-calculated, better than naive, starting point and by adding a regularisation term.\\
 
@@
 Since this starting point is expected to already be a good solution, a regulariser term is added to the SmoothRank objective function to prevent the solution from deviating to much from the starting point.
-The regularised smooth objective function is formulated as $\lambda||w-w_0||^2$ where $\lambda$ is a hyperparameter tuned on the validation set, and $w_0$ is the starting point solution.\\
+The regularised smooth objective function is formulated as $\lambda||w-w_0||_2$ where $\lambda$ is a hyper-parameter tuned on the validation set, $w_0$ is the starting point solution, and $||x||_2$ is the $\ell2$-norm regularisation function defined as $||x||_2 = \sqrt{\sum\nolimits_ix_i^2}$.\\
 
-The choice of the smoothing parameter $\sigma$ in Equation \ref{eq:soft_ndcg} is important, because a too small value makes the function more non-smooth and therefore harder to optimise, while a too large value results in a optimisation function that substantially differs from the optimal rankings.
-SmoothRank uses an annealing method where the optimisation procedure starts with a large $\sigma$ and iteratively reduces it by dividing it by 2 at each step.
+The choice of the smoothing parameter $\sigma$ in Equation \ref{eq:soft_ndcg} is important, because a too small value makes the function more non-smooth and therefore harder to optimise, while a too large value results in a optimisation function with optima that substantially differ from the true optimal rankings.
+To deal with the problem of choosing the smoothing parameter, SmoothRank uses an annealing method where the optimisation procedure starts with a large $\sigma$ and iteratively reduces it by dividing it by two at each iteration.
 Algorithm \ref{alg:smoothrank} shows the algorithm that summarises all steps of the SmoothRank method.\\
@@
  	 Starting from $w$, minimize by non-linear conjugate gradient descent:
- 	 $\lambda||w-w_0||^2 - \sum\nolimits_{q}A_q(w,\sigma)$\\
+ 	 $\lambda||w-w_0||_2 - \sum\nolimits_{q}A_q(w,\sigma)$\\
  	 $\sigma = \sigma/2$
@@
  Output $\underline{R_i}$
- \caption{Learning algorithm of LRUF, obtained from \cite{Torkestani2012b}}
+ \caption{LRUF algorithm, obtained from \cite{Torkestani2012b}}
  \label{alg:lruf}

### 2014-04-30 report/Chapters/LTRMethods.tex
self-review

@@
 FenchelRank is based on the theory of Fenchel Duality \cite{Rifkin2007} and uses a generic algorithm framework proposed by Shalev-Shwartz and Singer \cite{Shalev-Shwartz2010}.
-FenchelRank optimises the objective function shown in Equation \ref{eq:pairwise_l1_loss} which is equivalent to the standard $\ell_1$-norm regularised pairwise loss function.
+FenchelRank optimises the objective function shown in Equation \ref{eq:pairwise_l1_loss} which is equivalent to the standard pairwise loss function with $\ell_1$-norm regularisation.
 The $\ell_1$-norm regularisation function in this equation is represented by $||x||_1$ and is defined as $||x||_1=\sum\nolimits_i|x_i|$.
@@
 \end{equation}
-\noindent where $f^*$ is the convex conjugate of $f$ and $g_*$ is the concave conjugate of $g$.
+\noindent where $f^*$ is the convex, and $f_*$ the concave conjugate of $f$.\\
 
-\noindent Let $D(w) = -G(w)$.
-This allows the FenchelRank objective function (Equation \ref{eq:pairwise_l1_loss}) to be rewritten using Fenchel's duality theorem (Equation \ref{eq:fenchel_duality_theorem}) to the form shown in Equation \ref{eq:fenchel_dual}.
+To ease applying Fenchel's duality theorem (Equation 7) to the $\ell_1$-regularised pairwise loss function (Equation \ref{eq:pairwise_l1_loss}), Lai et al.
+define $D(w)$ as $D(w) = -G(w)$.
+Equation \ref{eq:fenchel_dual} shown the resulting Fenchel dual of the $\ell_1$-regularised pairwise loss function, which is the loss function that is used in FenchelRank.
 \begin{equation}
@@
 Algorithm \ref{alg:fenchelrank} shows the FenchelRank training algorithm to optimise the Fenchel dual of the pairwise loss function.
-The $||x||_infty$ term in this algorithm represents an $\ell_\infty$-norm regularisation term and is defined as $||x||_\infty=\max_{i}|x_i|$.\\
+The $||x||_\infty$ term in this algorithm represents an $\ell_\infty$-norm regularisation term and is defined as $||x||_\infty=\max_{i}|x_i|$.\\
 
@@
 \end{algorithm}
+
 \section{FSMRank}
 Lai et al.
-\cite{Lai2013c} observed that existing feature selection methods in Learning-to-Rank all follow a two-stage paradigm: 1) select subset of features from the original features, and 2) learn a ranking model based on the selected features.
-It is stated as a limitation of this paradigm that the selected features in the first step are not necessarily the optimal features for the second stage where the ranking model is build.
-FSMRank \cite{Lai2013c} addresses this limitation by using a joint convex optimisation problem that minimises ranking errors while simultaneously selecting a subset of features.\\
+\cite{Lai2013c} observed that existing feature selection methods in Learning-to-Rank all follow a two-stage paradigm consisting of a first stage of selecting a subset of features from the original features, followed by a second stage where a ranking model learnt based on the selected features.
+Lai et al.
+\cite{Lai2013c} state it as a limitation of this paradigm that the selected features in the first step are not necessarily the optimal features for the second stage where the ranking model is build.
+FSMRank \cite{Lai2013c} addresses this limitation by formulating a joint convex optimisation function that minimises ranking errors while simultaneously selecting a subset of features.\\
 
@@
 Let $\oslash$ be the element-wise division operator.
+$\hat{A}$ is a $d \times d$ similarity matrix that contains similarity scores between features.
 The convex joint optimisation function in FSMRank is defined as shown in Equation \ref{eq:fsmrank_optimisation_function}.
@@
 \end{equation}
-\noindent $\hat{A}$ is a $d \times d$ similarity matrix that contains similarity scores between features.
-The first term in Equation \ref{eq:fsmrank_optimisation_function} applies a penalty on redundancy of large weighted features by using the $\hat{A}$ matrix.
-The well-known Pearson correlation coefficient is used to calculate similarity matrix $\hat{A}$ based on the values for the features in the training data.
+\noindent The first term in Equation \ref{eq:fsmrank_optimisation_function} applies a penalty on redundancy of large weighted features by using the $\hat{A}$ matrix.
+The well-known Pearson correlation coefficient is used to calculate similarity matrix $\hat{A}$, based on the values for the features in the training data.
 $\lambda_1$ is a hyper-parameter of the model and can be used to set the weight of the similarity penalty term.\\
 
-The second term of Equation \ref{eq:fsmrank_optimisation_function} uses a $\ell_1$-norm regularisation term to select the effective features from the feature set.
+The second term of Equation \ref{eq:fsmrank_optimisation_function} contains a $\ell_1$-norm regularisation term to select the effective features from the feature set.
 $\lambda_2$ is a hyper-parameter of the model and can be used to set the weight of the regularisation term.\\
 
-The last term represents the loss function in terms of ranking errors.
-The loss function $f$ can in theory be any convex loss function, but Lai et al.
+The third and last term of the equation represents the loss function in terms of ranking errors.
+Loss function $f$ can in theory be any convex loss function, but Lai et al.
 \cite{Lai2013c} used a squared hinge loss function for their evaluation measurements.
 
-Algorithm \ref{alg:fsmrank} shows the steps of the FSMRank algorithm.
-In this algorithm an extended version of Nesterov's accelerated gradient method is used that can handle optimisation problems in the form of $\min_w \l(w) + r(w)$ where $l(w)$ is a convex function with Lipschitz gradient and $r(w)$ is convex, but non-smooth.
+Algorithm \ref{alg:fsmrank} shows the steps of the FSMRank training algorithm.
+As stated, FSMRank uses an  extended version of Nesterov's accelerated gradient method.
+This optimisation method can handle optimisation problems in the form of $\min_w \l(w) + r(w)$ where $l(w)$ is a convex function with Lipschitz gradient and $r(w)$ is convex, but non-smooth.
 The optimization function of Equation \ref{eq:fsmrank_optimisation_function} is reformulated to match this form as presented in Equation \ref{eq:fsmrank_optimisation_reformulated}.
@@
 \noindent where $Q(w_t,z_t)$ is a combination of the non-smooth part $r(w_t)$ and a quadratic approximation of the smooth part $l(w_t)$.
-$lambda_1$, $lambda_2$ and Lipschitz constant $L_0$ are input parameters of the algorithm and can be optimised through cross-validation or on a validation set.\\
+$\lambda_1$, $\lambda_2$ and Lipschitz constant $L_0$ are input parameters of the algorithm and can be optimised through cross-validation or on a validation set.\\
 
@@
 
-Three well-known learning algorithms $L$ used in variable structure learning automata are \emph{linear reward-penalty} ($L_{R-P}$), \emph{linear reward-$epsilon$-penalty} ($L_{R-\epsilon P}$) and \emph{linear reward-inaction} ($L_{R-I}$).
+Three well-known learning algorithms $L$ used in variable structure learning automata are \emph{linear reward-penalty} ($L_{R-P}$), \emph{linear reward-$\epsilon$-penalty} ($L_{R-\epsilon P}$) and \emph{linear reward-inaction} ($L_{R-I}$).
 LRUF uses the $L_{R-I}$ learning algorithm, which updates the action probability vector following Equation \ref{eq:lruf_reward_update} when the selected action $a_i(k)$ is rewarded by the environment.
@@
 \end{equation}
-\noindent where $i$ and $j$ are indices of action in $\alpha$ and $p(k)$ is the probability vector over the action set at rank $k$. $\alpha$ $a$ is the learning rate of the model.\\
+\noindent where $i$ and $j$ are indices of action in $\alpha$ and $p(k)$ is the probability vector over the action set at rank $k$. $a$ is the learning rate of the model.\\
 

### 2014-05-01 report/Chapters/CrossBenchmarkComparison.tex
feedback Sander verwerkt

--- 2014-04-11 report/Chapters/CrossBenchmarkComparison.tex
@@
-As we have seen in Chapter~\ref{chap:benchmark_results}, the evaluation of Learning-to-Rank methods is spread over the several benchmark datasets.
+\label{chap:cross_comparison}
+As we have seen in Chapter~\ref{chap:benchmark_results}, the evaluation of Learning-to-Rank methods is spread over several benchmark datasets.
 However, as the Learning-to-Rank methods evaluated differs between benchmarks, no single benchmark comparison can be regarded as a conclusive argument on which Learning-to-Rank method is most accurate.\\
 
-Several studies make a small start in considering Learning-to-Rank method performance over multiple benchmark datasets.
+Several studies make a small start in considering Learning-to-Rank methods performance over multiple benchmark datasets.
 Gomes et al.
@@
 To my knowledge, no structured meta-analysis on ranking accuracy has been conducted where evaluation results on several benchmark collections are taken into account.
-With a meta-analysis I will compare the performance of Learning-to-Rank methods across the Learning-to-Rank benchmark datasets described in foregoing sections.
+With a meta-analysis we will compare the performance of Learning-to-Rank methods across the Learning-to-Rank benchmark datasets described in foregoing sections.
 
@@
 \label{sec:collecting_evaluation_results}
-With a literature review I will collect evaluation results on the datasets/collections. The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
+With a literature review we will collect evaluation results on the datasets/collections. The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
 \begin{itemize}
@@
 LETOR 1.0, LETOR 3.0, Yahoo!
-Learning to Rank Challenge, WCL2R and AOL have an accompanying paper which was published together with the benchmark collection.
-As users of those benchmark collections are encouraged to cite these papers, evaluated measurements for these benchmark collections will be gathered through a forward literature search.
+Learning to Rank Challenge, WCL2R and AOL have accompanying papers which were published together with these benchmark collections.
+Users of those benchmark collections are encouraged to cite these papers.
+Therefore, we collect evaluation measurements of Learning-to-Rank methods on these benchmark collections through forward literature search.
 Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search.
-Google Scholar will be used for this forward reference search.
+Google Scholar will be used to perform the forward reference search.
 
@@
 To compare algorithms based on a sparse set of evaluation measurements, a normalised version of the Winning Number metric will be used.
-This Normalised Winning Number takes into account only those datasets that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all datasets on which it has been evaluated.
-I will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as \\
+This Normalised Winning Number takes only those datasets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all datasets on which it has been evaluated.
+We will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as \\
 
@@
 These metrics seem to be the most frequently used evaluation metrics in most of the used benchmark datasets.
-An exception are the Yahoo!
+An exception is the Yahoo!
 Learning to Rank Challenge datasets on which mainly \ac{ERR} is used as main evaluation metric.
 The lack of use of the \ac{ERR}-metric in other benchmarks makes it unsuitable for a cross-benchmark comparison.
-By making the decision to include \ac{nDCG} at thee cut-off points and only a single \ac{MAP} entry we implicitly attain a higher weight for \ac{nDCG} compared to \ac{MAP} on an analysis that combines all measurements on the four metrics.
+By making the decision to include \ac{nDCG} at three cut-off points and only a single \ac{MAP} entry, we implicitly attain a higher weight for \ac{nDCG} compared to \ac{MAP} on an analysis that combines all measurements on the four metrics.
 This implicit weighting is arbitrary and therefore undesirable, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.
@@
 Some studies with evaluation results found through in literature review were not usable for the meta-analysis.
-The following enumeration lists the properties that made those studies unusable for the meta-analysis, with between brackets the studies that these properties apply to.
+The following enumeration enumerates those properties that made one or more studies unusable for the meta-analysis.
+Between brackets are the studies that these properties apply to.
 
@@
 BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
-Best Single Feature &  & \cite{Gomes2013} & MHR & L2, \cite{Qin2007} &  \\ 
+Best Single Feature &  & \cite{Gomes2013} & \acs{MHR} & \cite{Qin2007} & L2 \\ 
 BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
@@
 LambdaNeuralRank and CoList both acquired a perfect Normalised Winning Number score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
-LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three datasets which in addition are more relevant datasets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 datasets.
+LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three datasets which in addition are more relevant datasets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 datasets (see Table \ref{tab:ltr_methods_used}).
 FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in Normalised Winning Number, but increasingly higher in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
@@
 LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both Ideal Winning Number and Normalised Winning Number.
-LRUF is the only methof that achieved this in all \ac{nDCG} comparisons, the \ac{MAP} comparison as well as the cross-metric comparison.
+LRUF is the only method that achieved this in all \ac{nDCG} comparisons, the \ac{MAP} comparison as well as the cross-metric comparison.
 With FenchelRank, FSMRank, SmoothRank and ListNet being among the top performing methods in all \ac{nDCG} comparisons as well as in the cross-metric comparison, it can be concluded that the cross-metric results are highly defined by the \ac{nDCG} performance as opposed to the \ac{MAP} performance.
@@
 LAC-MR-OR and RE-QR showed very good ranking accuracy in the \ac{MAP} comparison on multiple datasets.
-Both methods Because LAC-MR-OR is only evaluated on two datasets for \ac{nDCG}@10 and RE-QR is not evaluated for \ac{nDCG} at all
+Because LAC-MR-OR is only evaluated on two datasets for \ac{nDCG}@10 and RE-QR is not evaluated for \ac{nDCG} at all, LAC-MR-OR and RE-QR are not within the top performing methods in the cross-metric comparison.
 
@@
 By calculating it in this way, we implicitly make the assumption that the Learning-to-Rank methods are (approximately) distributed uniformly over the benchmarks, such that the average Learning-to-Rank method tested are approximately equally hard for each data set.
-It could be the case however that this assumption is false and that one some datasets significantly better Learning-to-Rank methods are evaluated than on some other datasets.\\
+It could be the case however that this assumption is false and that significantly more accurate Learning-to-Rank methods are evaluated on some datasets than on other datasets.
+\\
 
@@
 Overly optimistic evaluation results published by other researchers do affect our Normalised Winning Number results.
-Only including studies in the meta-analysis that comparable show results on one of the baseline methods of a benchmark reduces this limitation but does not solve it.
+Limiting the meta-analysis to those studies that report comparable results on one of the baseline methods of a benchmark set reduces this limitation but does not solve it completely.
 By taking the Ideal Winning Number into account in Figure \ref{fig:normalised_winning_number_all} we further mitigate this limitation, as the Ideal Winning Number is loosely related with the number of studies that reported evaluation results for an algorithm.

### 2014-05-15 cikm_paper/paper.tex
eerste poging CIKM paper

--- 2014-05-01 report/Chapters/CrossBenchmarkComparison.tex
@@
+% This is "sig-alternate.tex" V2.0 May 2012
+% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
+%
+% This example file demonstrates the use of the 'sig-alternate.cls'
+% V2.5 LaTeX2e document class file. It is for those submitting
+% articles to ACM Conference Proceedings WHO DO NOT WISH TO
+% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
+% The 'sig-alternate.cls' file will produce a similar-looking,
+% albeit, 'tighter' paper resulting in, invariably, fewer pages.
+%
+% ----------------------------------------------------------------------------------------------------------------
+% This .tex file (and associated .cls V2.5) produces:
+%       1) The Permission Statement
+%       2) The Conference (location) Info information
+%       3) The Copyright Line with ACM data
+%       4) NO page numbers
+%
+% as against the acm_proc_article-sp.cls file which
+% DOES NOT produce 1) thru' 3) above.
+%
+% Using 'sig-alternate.cls' you have control, however, from within
+% the source .tex file, over both the CopyrightYear
+% (defaulted to 200X) and the ACM Copyright Data
+% (defaulted to X-XXXXX-XX-X/XX/XX).
+% e.g.
+% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
+% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
+%
+% ---------------------------------------------------------------------------------------------------------------
+% This .tex source is an example which *does* use
+% the .bib file (from which the .bbl file % is produced).
+% REMEMBER HOWEVER: After having produced the .bbl file,
+% and prior to final submission, you *NEED* to 'insert'
+% your .bbl file into your source .tex file so as to provide
+% ONE 'self-contained' source file.
+%
+% ================= IF YOU HAVE QUESTIONS =======================
+% Questions regarding the SIGS styles, SIGS policies and
+% procedures, Conferences etc. should be sent to
+% Adrienne Griscti (griscti@acm.org)
+%
+% Technical questions _only_ to
+% Gerald Murray (murray@hq.acm.org)
+% ===============================================================
+%
+% For tracking purposes - this is V2.0 - May 2012
+
+\documentclass{sig-alternate-2013}
+
+\begin{document}
+%
+% --- Author Metadata here ---
+\conferenceinfo{ACM International Conference on Information and Knowledge Management}{'14 Shanghai, China}
+%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
+%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
+% --- End of Author Metadata ---
+
+\title{A cross benchmark comparison of Learning to Rank methods}
+%
+% You need the command \numberofauthors to handle the 'placement
+% and alignment' of the authors beneath the title.
+%
+% For aesthetic reasons, we recommend 'three authors at a time'
+% i.e. three 'name/affiliation blocks' be placed beneath the title.
+%
+% NOTE: You are NOT restricted in how many 'rows' of
+% "name/affiliations" may appear. We just ask that you restrict
+% the number of 'columns' to three.
+%
+% Because of the available 'opening page real-estate'
+% we ask you to refrain from putting more than six authors
+% (two rows with three columns) beneath the article title.
+% More than six makes the first-page appear very cluttered indeed.
+%
+% Use the \alignauthor commands to handle the names
+% and affiliations for an 'aesthetic maximum' of six authors.
+% Add names, affiliations, addresses for
+% the seventh etc. author(s) as the argument for the
+% \additionalauthors command.
+% These 'additional authors' will be output/set for you
+% without further effort on your part as the last section in
+% the body of your article BEFORE References or any Appendices.
+
+\numberofauthors{3} %  in this sample file, there are a *total*
+% of EIGHT authors. SIX appear on the 'first-page' (for formatting
+% reasons) and the remaining two appear in the \additionalauthors section.
+%
+\author{
+% You can go ahead and credit any number of authors here,
+% e.g. one 'row of three' or two rows (consisting of one row of three
+% and a second row of one, two or three).
+%
+% The command \alignauthor (no curly braces needed) should
+% precede each author name, affiliation/snail-mail address and
+% e-mail address. Additionally, tag each line of
+% affiliation/address with \affaddr, and tag the
+% e-mail address with \email.
+%
+% 1st. author
+\alignauthor
+Niek Tax\titlenote{affiliated with University of Twente, but written during his time at Avanade Netherlands B.V.}\\
+       \affaddr{Avanade Netherlands B.V.}\\
+       \affaddr{Versterkerstraat 6}\\
+       \affaddr{1322 AP Almere}\\
+       \affaddr{The Netherlands}
+       \email{niek.tax@avanade.com}
+% 2nd. author
+\alignauthor
+Djoerd Hiemstra\\
+       \affaddr{University of Twente}\\
+       \affaddr{P.O. Box 217}\\
+       \affaddr{7500 AE Enschede}\\
+       \affaddr{The Netherlands}
+       \email{d.hiemstra@utwente.nl}
+% 3rd. author
+\alignauthor
+Sander Bockting\\
+       \affaddr{Avanade Netherlands B.V.}\\
+       \affaddr{Versterkerstraat 6}\\
+       \affaddr{1322 AP Almere}\\
+       \affaddr{The Netherlands}
+       \email{sander.bockting@avanade.com}
+}
+
+\maketitle
+\begin{abstract}
+
+\end{abstract}
+
+% A category with the (minimum) three required fields
+\category{H.4}{Information Systems Applications}{Miscellaneous}
+%A category including the fourth, optional field follows...
+\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+
+\terms{}
+
+\keywords{, }
+
+\section{Introduction}
+
+\section{The {\secit Body} of The Paper}
+
+\subsection{Type Changes and {\subsecit Special} Characters}
 \label{chap:cross_comparison}
@@
 With a literature review we will collect evaluation results on the datasets/collections. The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
-\begin{itemize}
+\begin{itemize}\itemsep0em
 \item LETOR 2.0
@@
 
-where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{nDCG} or \ac{MAP}), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that\\
+where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as NDCG or MAP), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that\\
 
@@
 
-where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{nDCG} or \ac{MAP}), and $D_{\{M_i(j),M_k(j)\}}$ is an evaluation definition function such that\\
+where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as NDCG or MAP), and $D_{\{M_i(j),M_k(j)\}}$ is an evaluation definition function such that\\
 
@@
 
-\ac{nDCG}@3, \ac{nDCG}@5, \ac{nDCG}@10 and \ac{MAP} are chosen as metrics on which the meta-analysis will be performed.
+NDCG@3, NDCG@5, NDCG@10 and MAP are chosen as metrics on which the meta-analysis will be performed.
 These metrics seem to be the most frequently used evaluation metrics in most of the used benchmark datasets.
 An exception is the Yahoo!
-Learning to Rank Challenge datasets on which mainly \ac{ERR} is used as main evaluation metric.
-The lack of use of the \ac{ERR}-metric in other benchmarks makes it unsuitable for a cross-benchmark comparison.
-By making the decision to include \ac{nDCG} at three cut-off points and only a single \ac{MAP} entry, we implicitly attain a higher weight for \ac{nDCG} compared to \ac{MAP} on an analysis that combines all measurements on the four metrics.
+Learning to Rank Challenge datasets on which mainly ERR is used as main evaluation metric.
+The lack of use of the ERR metric in other benchmarks makes it unsuitable for a cross-benchmark comparison.
+By making the decision to include NDCG at three cut-off points and only a single MAP entry, we implicitly attain a higher weight for NDCG compared to MAP on an analysis that combines all measurements on the four metrics.
 This implicit weighting is arbitrary and therefore undesirable, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.
@@
 \begin{table}[!h!p]
-\scalebox{0.75}{
 \begin{tabular}{|l|l|l||l|l|l|}
@@
 \hline 
-AdaRank-\acs{MAP} & \cite{Xu2007} & L2, L3, L4 & Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} \\ 
-AdaRank-\acs{nDCG} & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
+AdaRank-MAP & \cite{Xu2007} & L2, L3, L4 & Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} \\ 
+AdaRank-NDCG & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & ListMLE & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
 ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet & \cite{Cao2007} & L2, L3, L4 \\ 
-Approx\acs{AP} & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
-Approx\acs{nDCG} & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
+ApproxAP & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
+ApproxNDCG & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
 BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
-Best Single Feature &  & \cite{Gomes2013} & \acs{MHR} & \cite{Qin2007} & L2 \\ 
+Best Single Feature &  & \cite{Gomes2013} & MHR & \cite{Qin2007} & L2 \\ 
 BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
@@
 BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c, Volkovs2013} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
-BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-\acs{MAP} & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
-C-\acs{CRF} & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
-CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & Q.D.\acs{KNN} & \cite{Geng2008} & \cite{Wang2013} \\ 
+BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-MAP & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
+C-CRF & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
+CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & Q.D.KNN & \cite{Geng2008} & \cite{Wang2013} \\ 
 CCRank & \cite{Wang2011c} & \cite{Wang2011c} & RandomForest &  & \cite{Gomes2013} \\ 
 CoList & \cite{Gao2014} & \cite{Gao2014} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
-Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAgg\acs{nDCG} & \cite{Wang2013} & \cite{Wang2013} \\ 
+Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAggNDCG & \cite{Wang2013} & \cite{Wang2013} \\ 
 DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010} \\ 
 DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
-Energy\acs{nDCG} & \cite{Freno2011} & \cite{Freno2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\ 
+EnergyNDCG & \cite{Freno2011} & \cite{Freno2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\ 
 FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
@@
 FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
-Focused\acs{SVM} & \cite{Niu2012} & \cite{Niu2012} & Rank\acs{RLS} & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
-FP-Rank & \cite{Song2013} & \cite{Song2013} & Rank\acs{SVM} & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010} \\ 
-FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & Rank\acs{SVM}-Primal &  & L3, \cite{Lai2011} \\ 
-FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & Rank\acs{SVM}-Struct &  & L3, L4 \\
-FSM$^{\acs{SVM}}$ & \cite{Lai2013c} & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
+FocusedSVM & \cite{Niu2012} & \cite{Niu2012} & RankRLS & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
+FP-Rank & \cite{Song2013} & \cite{Song2013} & RankSVM & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010} \\ 
+FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & RankSVM-Primal &  & L3, \cite{Lai2011} \\ 
+FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & RankSVM-Struct &  & L3, L4 \\
+FSM$^{SVM}$ & \cite{Lai2013c} & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
 GAS-E & \cite{Geng2007} & \cite{Lai2013c} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\
-\acs{GP} & \cite{DeAlmeida2007} & \cite{Alcantara2010} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\  
-\acs{GP}Rank & \cite{Silva2009} & \cite{Torkestani2012} & Ridge Regression & \cite{Cossock2006} & L3 \\
-GRank\acs{RLS} & \cite{Pahikkala2010} & \cite{Pahikkala2010} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
-Group\acs{CE} & \cite{Lin2011} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
-Group\acs{MLE} & \cite{Lin2010} & \cite{Lin2011} & SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010} \\
+GP & \cite{DeAlmeida2007} & \cite{Alcantara2010} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\  
+GPRank & \cite{Silva2009} & \cite{Torkestani2012} & Ridge Regression & \cite{Cossock2006} & L3 \\
+GRankRLS & \cite{Pahikkala2010} & \cite{Pahikkala2010} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
+GroupCE & \cite{Lin2011} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
+GroupMLE & \cite{Lin2010} & \cite{Lin2011} & SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010} \\
 IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
-\acs{IP}Rank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\
+IPRank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\
 KeepRank & \cite{Chen2009} & \cite{Chen2009} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
-Kernel-\acs{PCA} RankBoost & \cite{Duh2008} & \cite{Duh2008, Sato2013} & \acs{SVD}-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\
-KL-\acs{CRF} & \cite{Volkovs2011} & \cite{Volkovs2011} & \acs{SVM}$^{\acs{MAP}}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
+Kernel-PCA RankBoost & \cite{Duh2008} & \cite{Duh2008, Sato2013} & SVD-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\
+KL-CRF & \cite{Volkovs2011} & \cite{Volkovs2011} & SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
 LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
@@
 LARF & \cite{Torkestani2012} & \cite{Torkestani2012} &  &  &  \\ 
-\end{tabular}}
+\end{tabular}
 \caption{Learning-to-Rank algorithms with measurements on benchmark datasets}
@@
 \section{Results \& Discussion}
-The following subsections provide the performance of Learning-to-Rank methods in terms of Normalised Winning Number for \ac{nDCG}@3, \ac{nDCG}@5, \ac{nDCG}@10 and \ac{MAP}.
+The following subsections provide the performance of Learning-to-Rank methods in terms of Normalised Winning Number for NDCG@3, NDCG@5, NDCG@10 and MAP.
 Performance of the Learning-to-Rank methods is plotted with Normalised Winning Number on the vertical axis and the number of datasets on which the method has been evaluated on the horizontal axis.
@@
 \subsection{NDCG@3}
-Figure \ref{fig:normalised_winning_number_ndcg3} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@3 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg3} in Appendix \ref{app:norm_winnum_ndcg3} provides the raw Normalised Winning Number data for the Learning-to-Rank methods for which \ac{nDCG}@3 evaluation results were available.\\
+Figure \ref{fig:normalised_winning_number_ndcg3} shows the performance of Learning-to-Rank methods for the NDCG@3 metric.
+Table \ref{tab:raw_data_norm_winnum_ndcg3} in Appendix \ref{app:norm_winnum_ndcg3} provides the raw Normalised Winning Number data for the Learning-to-Rank methods for which NDCG@3 evaluation results were available.\\
 
@@
 \includegraphics[scale=0.285]{gfx/ndcg3_winnum}
-\caption{\acs{nDCG}@3 comparison of Learning-to-Rank methods}
+\caption{NDCG@3 comparison of Learning-to-Rank methods}
 \label{fig:normalised_winning_number_ndcg3}
@@
 \subsection{NDCG@5}
-Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@5 metric.
+Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of Learning-to-Rank methods for the NDCG@5 metric.
 Table \ref{tab:raw_data_norm_winnum_ndcg5} in Appendix \ref{app:norm_winnum_ndcg5} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.\\
@@
 \includegraphics[scale=0.285]{gfx/ndcg5_winnum}
-\caption{\acs{nDCG}@5 comparison of Learning-to-Rank methods}
+\caption{NDCG@5 comparison of Learning-to-Rank methods}
 \label{fig:normalised_winning_number_ndcg5}
@@
 LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of datasets, but score decreasingly well in terms of Normalised Winning Number.
-These results are highly in agreement with the \ac{nDCG}@3 comparison.
-The only modification compared to the \ac{nDCG}@3 comparison being that OWPC did show to be a method for which there were no methods performing better on both axes in the \ac{nDCG}@5 comparison, but not in the \ac{nDCG}@3 comparison.
-Like in the \ac{nDCG}@3 comparison, SmoothRank and ListNet can be regarded as most reliable results because the evaluation measurements for these methods are based on LETOR official baselines.
+These results are highly in agreement with the NDCG@3 comparison.
+The only modification compared to the NDCG@3 comparison being that OWPC did show to be a method for which there were no methods performing better on both axes in the NDCG@5 comparison, but not in the NDCG@3 comparison.
+Like in the NDCG@3 comparison, SmoothRank and ListNet can be regarded as most reliable results because the evaluation measurements for these methods are based on LETOR official baselines.
 
 \subsection{NDCG@10}
-Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@10 metric.
+Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of Learning-to-Rank methods for the NDCG@10 metric.
 Table \ref{tab:raw_data_norm_winnum_ndcg10} in Appendix \ref{app:norm_winnum_ndcg10} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.\\
@@
 \includegraphics[scale=0.285]{gfx/ndcg10_winnum}
-\caption{\acs{nDCG}@10 comparison of Learning-to-Rank methods}
+\caption{NDCG@10 comparison of Learning-to-Rank methods}
 \label{fig:normalised_winning_number_ndcg10}
@@
 
-LambdaMART and LambdaNeuralRank score a Normalised Winning Number of 1.0 on the \ac{nDCG}@10 comparison.
+LambdaMART and LambdaNeuralRank score a Normalised Winning Number of 1.0 on the NDCG@10 comparison.
 For LambdaNeuralRank these results are again based on AOL dataset measurements.
-LambdaMART showed the highest \ac{nDCG}@10 performance for the MSLR-WEB10k dataset.
-The set of algorithms for which there is no other algorithm with both a higher Normalised Winning Number and number of datasets evaluated on is partly in agreement with those for the \ac{nDCG}@3 and \ac{nDCG}@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
-SmoothRank and FSMRank were not present in this set for the \ac{nDCG}@3 and \ac{nDCG}@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_ndcg3} and \ref{tab:raw_data_norm_winnum_ndcg5} in the Appendices.
-DCMP is not in the set in contrast with the \ac{nDCG}@3 and \ac{nDCG}@5 comparison.
+LambdaMART showed the highest NDCG@10 performance for the MSLR-WEB10k dataset.
+The set of algorithms for which there is no other algorithm with both a higher Normalised Winning Number and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and NDCG@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
+SmoothRank and FSMRank were not present in this set for the NDCG@3 and NDCG@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_ndcg3} and \ref{tab:raw_data_norm_winnum_ndcg5} in the Appendices.
+DCMP is not in the set in contrast with the NDCG@3 and NDCG@5 comparison.
 
 \subsection{MAP}
-Figure \ref{fig:normalised_winning_number_map} shows the performance of Learning-to-Rank methods for the \ac{MAP} metric.
+Figure \ref{fig:normalised_winning_number_map} shows the performance of Learning-to-Rank methods for the MAP metric.
 Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.\\
@@
 \includegraphics[scale=0.285]{gfx/map_winnum}
-\caption{\acs{MAP} comparison of Learning-to-Rank methods}
+\caption{MAP comparison of Learning-to-Rank methods}
 \label{fig:normalised_winning_number_map}
@@
 
-Where comparisons on the \ac{nDCG}-metric at different cut-off points where highly in agreement in terms of the best performing algorithms, the comparison in terms of \ac{MAP} shows different results.
-RankDE scores a Normalised Winning Number of 1.0 on one dataset, like LambdaNeuralRank did on for all \ac{nDCG}-comparisons.
+Where comparisons on the NDCG metric at different cut-off points where highly in agreement in terms of the best performing algorithms, the comparison in terms of MAP shows different results.
+RankDE scores a Normalised Winning Number of 1.0 on one dataset, like LambdaNeuralRank did on for all NDCG-comparisons.
 In contrast to LambdaNeuralRank, RankDE achieved this highest score on the LETOR 2.0 TD2003, a dataset on which many methods are evaluated.\\
 
-LARF and LRUF score very high Normalised Winning Number scores, but based on only relatively few datasets, just as in the \ac{nDCG}-comparisons.
-Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all \ac{nDCG}-comparisons, are not within the best performing methods in the \ac{MAP}-comparison.
-A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_ndcg3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_ndcg3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more datasets for \ac{MAP} compared to \ac{nDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalised Winning Number.
+LARF and LRUF score very high Normalised Winning Number scores, but based on only relatively few datasets, just as in the NDCG comparisons.
+Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all NDCG comparisons, are not within the best performing methods in the MAP comparison.
+A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_ndcg3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_ndcg3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more datasets for MAP compared to NDCG, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalised Winning Number.
 SmoothRank performed a Normalised Winning Number of around 0.53 over 7 datasets, which is still good in both certainty and accuracy, but not among the top methods.
-RE-QR is one of the best performers in the \ac{MAP}-comparison with a reasonable amount of benchmark evaluations.
-No reported \ac{nDCG} performance was found in the literature study for RE-QR.
-There is a lot of certainty on the accuracy of RankBoost and Rank\ac{SVM} as both models are evaluated on the majority of datasets included in the comparison for the \ac{MAP}-metric, but given their Normalised Winning Number it can said that both methods are not within the top performing Learning-to-Rank methods.
+RE-QR is one of the best performers in the MAP comparison with a reasonable amount of benchmark evaluations.
+No reported NDCG performance was found in the literature study for RE-QR.
+There is a lot of certainty on the accuracy of RankBoost and RankSVM as both models are evaluated on the majority of datasets included in the comparison for the MAP metric, but given their Normalised Winning Number it can said that both methods are not within the top performing Learning-to-Rank methods.
 
@@
 
-The cross-metric comparison is based on the \ac{nDCG}@3, \ac{nDCG}@5, \ac{nDCG}@10 and \ac{MAP} comparisons combined, which justifies analysing the comparison more thoroughly.
+The cross-metric comparison is based on the NDCG@3, NDCG@5, NDCG@10 and MAP comparisons combined, which justifies analysing the comparison more thoroughly.
 Figure \ref{fig:normalised_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
@@
 LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both Ideal Winning Number and Normalised Winning Number.
-LRUF is the only method that achieved this in all \ac{nDCG} comparisons, the \ac{MAP} comparison as well as the cross-metric comparison.
-With FenchelRank, FSMRank, SmoothRank and ListNet being among the top performing methods in all \ac{nDCG} comparisons as well as in the cross-metric comparison, it can be concluded that the cross-metric results are highly defined by the \ac{nDCG} performance as opposed to the \ac{MAP} performance.
-This was to be expected, because the cross-metric comparison input data of three \ac{nDCG} entries (@3, @5, and @10) enables it to have up to three times as many as many weight as the \ac{MAP} comparison.\\
+LRUF is the only method that achieved this in all NDCG comparisons, the MAP comparison as well as the cross-metric comparison.
+With FenchelRank, FSMRank, SmoothRank and ListNet being among the top performing methods in all NDCG comparisons as well as in the cross-metric comparison, it can be concluded that the cross-metric results are highly defined by the NDCG performance as opposed to the MAP performance.
+This was to be expected, because the cross-metric comparison input data of three NDCG entries (@3, @5, and @10) enables it to have up to three times as many as many weight as the MAP comparison.\\
 
-LARF, \ac{IP}Rank and DCMP and several variants of Rank\ac{SVM} performed very well on the cross-metric comparison, with all having only one method in its top right quadrant.
-LARF also performed among the top methods on the \ac{nDCG} and \ac{MAP} comparisons and DCMP was a top performer in a few of the \ac{nDCG} comparisons.\\
+LARF, IPRank and DCMP and several variants of RankSVM performed very well on the cross-metric comparison, with all having only one method in its top right quadrant.
+LARF also performed among the top methods on the NDCG and MAP comparisons and DCMP was a top performer in a few of the NDCG comparisons.\\
 
@@
 
-LAC-MR-OR and RE-QR showed very good ranking accuracy in the \ac{MAP} comparison on multiple datasets.
-Because LAC-MR-OR is only evaluated on two datasets for \ac{nDCG}@10 and RE-QR is not evaluated for \ac{nDCG} at all, LAC-MR-OR and RE-QR are not within the top performing methods in the cross-metric comparison.
+LAC-MR-OR and RE-QR showed very good ranking accuracy in the MAP comparison on multiple datasets.
+Because LAC-MR-OR is only evaluated on two datasets for NDCG@10 and RE-QR is not evaluated for NDCG at all, LAC-MR-OR and RE-QR are not within the top performing methods in the cross-metric comparison.
 
@@
 
-In the following chapters of this thesis, concerning parallel execution of the Learning-to-Rank training phase, the scope will be limited to the five methods that turned out to be superior Learning-to-Rank methods in terms of ranking accuracy and certainty about this ranking accuracy: LRUF, FSMRank, FenchelRank, SmoothRank and ListNet.
-Although it can not be concluded that these methods are inarguably the most accurate Learning-to-Rank methods, a strong presumption has been raised that these five Learning-to-Rank are accurate ranking methods.
+%
+% The following two commands are all you need in the
+% initial runs of your .tex file to
+% produce the bibliography for the citations in your paper.
+\bibliographystyle{abbrv}
+\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
+% You must have a proper ".bib" file
+%  and remember to run:
+% latex bibtex latex latex
+% to resolve all references
+%
+% ACM needs 'a single self-contained file'!
+%
+%APPENDICES are optional
+%\balancecolumns
+\appendix
+\section{Raw Data for NDCG@3 Comparison}
+\label{app:norm_winnum_NDCG3}
+
+\begin{table}
+\begin{tabular}{l|l|l}
+Method & Normalised Winning Number & \# of datasets \\
+\hline
+AdaRank-MAP & 0.34862385 & 12 \\ 
+AdaRank-NDCG & 0.30733945 & 12 \\ 
+ApproxNDCG & 0.79487179 & 1 \\ 
+BagBoo & 0.84782609 & 2 \\ 
+BL-MART & 0.95238095 & 2 \\ 
+BoltzRank-Pair & 0.83333333 & 4 \\ 
+BoltzRank-Single & 0.75490196 & 4 \\ 
+BT & 0.72727273 & 3 \\ 
+CoList & 1.00000000 & 1 \\ 
+DCMP & 0.54591837 & 9 \\ 
+EnergyNDCG & 0.36363636 & 2 \\ 
+FBPCRank & 0.41463415 & 3 \\ 
+FenchelRank & 0.77419355 & 5 \\ 
+FocusedBoost & 0.37500000 & 2 \\ 
+FocusedNet & 0.45833333 & 2 \\ 
+FocusedSVM & 0.25000000 & 2 \\ 
+FRank & 0.30845771 & 11 \\ 
+FSMRank & 0.83333333 & 4 \\ 
+FSMSVM & 0.22916667 & 4 \\ 
+GAS-E & 0.37500000 & 4 \\ 
+GPRank & 0.87096774 & 3 \\ 
+GroupCE & 0.73118280 & 3 \\ 
+GroupMLE & 0.52688172 & 3 \\ 
+IntervalRank & 0.58974359 & 1 \\ 
+IPRank & 0.93548387 & 3 \\ 
+KL-CRF & 0.59459459 & 2 \\ 
+LambdaMART & 0.57142857 & 2 \\ 
+LambdaNeuralRank & 1.00000000 & 1 \\ 
+LambdaRank & 0.20000000 & 1 \\ 
+LARF & 0.98924731 & 3 \\ 
+Linear Regression & 0.07142857 & 9 \\ 
+ListMLE & 0.00000000 & 1 \\ 
+ListNet & 0.44954128 & 12 \\ 
+ListReg & 0.73118280 & 3 \\ 
+LRUF & 0.98230088 & 4 \\ 
+MHR & 0.75000000 & 1 \\ 
+NewLoss & 0.51612903 & 3 \\ 
+OWPC & 0.65000000 & 6 \\ 
+PERF-MAP & 0.38938053 & 4 \\ 
+RankBoost & 0.32110092 & 12 \\ 
+RankELM (pairwise) & 0.61538462 & 1 \\ 
+RankELM (pointwise) & 0.69230769 & 1 \\ 
+RankNet & 0.18867925 & 1 \\ 
+RankSVM & 0.30097087 & 12 \\ 
+RankSVM-Primal & 0.39204545 & 8 \\ 
+RankSVM-Struct & 0.35204082 & 9 \\ 
+REG-SHF-SDCG & 0.38461538 & 1 \\ 
+Ridge Regression & 0.40880503 & 7 \\ 
+RSRank & 0.57291667 & 4 \\ 
+SmoothRank & 0.60377358 & 7 \\ 
+SoftRank & 0.23076923 & 1 \\ 
+SortNet & 0.25000000 & 2 \\ 
+SparseRank & 0.82242991 & 4 \\ 
+SVMMAP & 0.28930818 & 7 \\ 
+TGRank & 0.54166667 & 4 \\ 
+TM & 0.59090909 & 3 \\
+\end{tabular}
+\caption{Raw data of Normalised Winning Number comparison on NDCG@3}
+\label{tab:raw_data_norm_winnum_NDCG3}
+\end{table}
+
+\section{Raw Data for NDCG@5 Comparison}
+\label{app:norm_winnum_NDCG5}
+
+\begin{table}
+\begin{tabular}{l|l|l}
+Method & Normalised Winning Number & \# of datasets \\
+\hline
+AdaRank-MAP & 0.3883929 & 12 \\ 
+AdaRank-NDCG & 0.3258929 & 12 \\ 
+ApproxNDCG & 0.7500000 & 1 \\ 
+BagBoo & 0.8400000 & 1 \\ 
+BL-MART & 0.7200000 & 1 \\ 
+BoltzRank-Pair & 0.8349515 & 4 \\ 
+BoltzRank-Single & 0.7184466 & 4 \\ 
+BT & 0.7878788 & 3 \\ 
+C-CRF & 0.9500000 & 2 \\ 
+DCMP & 0.5078534 & 9 \\ 
+EnergyNDCG & 0.3777778 & 2 \\ 
+FBPCRank & 0.5529412 & 3 \\ 
+FenchelRank & 0.7500000 & 5 \\ 
+FocusedBoost & 0.4545455 & 2 \\ 
+FocusedNet & 0.6363636 & 2 \\ 
+FocusedSVM & 0.2727273 & 2 \\ 
+FPRank & 0.9000000 & 1 \\ 
+FRank & 0.2849462 & 10 \\ 
+FSMRank & 0.8775510 & 4 \\ 
+FSMSVM & 0.4081633 & 4 \\ 
+GAS-E & 0.4693878 & 4 \\ 
+GPRank & 0.7252747 & 3 \\ 
+IntervalRank & 0.3750000 & 1 \\ 
+IPRank & 0.8131868 & 3 \\ 
+Kernel-PCA & 0.2857143 & 3 \\ 
+KL-CRF & 0.5789474 & 2 \\ 
+LambdaNeuralRank & 1.0000000 & 1 \\ 
+LambdaRank & 0.2000000 & 1 \\ 
+LARF & 0.9890110 & 3 \\ 
+Linear Regression & 0.1099476 & 9 \\ 
+ListNet & 0.4910714 & 12 \\ 
+ListReg & 0.6923077 & 3 \\ 
+LRUF & 0.9816514 & 4 \\ 
+MHR & 0.6000000 & 1 \\ 
+NewLoss & 0.4285714 & 3 \\ 
+PERF-MAP & 0.2660550 & 4 \\ 
+Q.D.KNN & 0.3205128 & 3 \\ 
+RankAggNDCG & 0.5000000 & 3 \\ 
+RankBoost & 0.2794118 & 10 \\ 
+RankDE & 0.5384615 & 1 \\ 
+RankELM (pairwise) & 0.6500000 & 1 \\ 
+RankELM (pointwise) & 0.7000000 & 1 \\ 
+RankNet & 0.2857143 & 3 \\ 
+Rank-PMBGP & 0.7692308 & 1 \\ 
+RankSVM & 0.3612565 & 11 \\ 
+RankSVM-Primal & 0.4508671 & 8 \\ 
+RankSVM-Struct & 0.4136126 & 9 \\ 
+RCP & 0.5757576 & 3 \\ 
+REG-SHF-SDCG & 0.4500000 & 1 \\ 
+Ridge Regression & 0.3333333 & 7 \\ 
+RSRank & 0.5306122 & 4 \\ 
+SmoothRank & 0.6339869 & 7 \\ 
+SoftRank & 0.2750000 & 1 \\ 
+SortNet & 0.5147059 & 4 \\ 
+SparseRank & 0.8173077 & 4 \\ 
+SVD-RankBoost & 0.2727273 & 3 \\ 
+SVMMAP & 0.3801170 & 8 \\ 
+SwarmRank & 0.1538462 & 1 \\ 
+TGRank & 0.6122449 & 4 \\ 
+TM & 0.7575758 & 3 \\ 
+\end{tabular}
+\caption{Raw data of Normalised Winning Number comparison on NDCG@5}
+\label{tab:raw_data_norm_winnum_NDCG5}
+\end{table}
+
+\section{Raw Data for NDCG@10 Comparison}
+\label{app:norm_winnum_NDCG10}
+
+\begin{table}
+\begin{tabular}{l|l|l}
+Method & Normalised Winning Number & \# of datasets \\
+\hline
+AdaRank-MAP & 0.36480687 & 13 \\ 
+AdaRank-NDCG & 0.31578947 & 16 \\ 
+ADMM & 0.44444444 & 1 \\ 
+ApproxNDCG & 0.86111111 & 1 \\ 
+Best Single Feature & 0.16149068 & 8 \\ 
+CA & 0.65217391 & 4 \\ 
+CoList & 0.16666667 & 1 \\ 
+Consistent-RankCosine & 0.76923077 & 2 \\ 
+DCMP & 0.58883249 & 9 \\ 
+DirectRank & 0.92307692 & 2 \\ 
+EnergyNDCG & 0.41463415 & 2 \\ 
+FenchelRank & 0.76229508 & 5 \\ 
+FocusedBoost & 0.68627451 & 2 \\ 
+FocusedNet & 0.86274510 & 2 \\ 
+FocusedSVM & 0.60784314 & 2 \\ 
+FRank & 0.30288462 & 11 \\ 
+FSMRank & 0.86206897 & 5 \\ 
+FSMSVM & 0.54255319 & 4 \\ 
+GAS-E & 0.45744681 & 4 \\ 
+GP & 0.66666667 & 2 \\ 
+GPRank & 0.65909091 & 3 \\ 
+GRankRLS & 0.28947368 & 2 \\ 
+GroupCE & 0.72727273 & 3 \\ 
+GroupMLE & 0.62500000 & 3 \\ 
+IPRank & 0.79545455 & 3 \\ 
+LAC-MR-OR & 0.66666667 & 2 \\ 
+LambdaMART & 1.00000000 & 1 \\ 
+LambdaNeuralRank & 1.00000000 & 1 \\ 
+LambdaRank & 0.57142857 & 2 \\ 
+LARF & 0.98863636 & 3 \\ 
+Linear Regression & 0.08287293 & 8 \\ 
+ListMLE & 0.02127660 & 4 \\ 
+ListNet & 0.59821429 & 12 \\ 
+LRUF & 0.98181818 & 4 \\ 
+MHR & 0.62500000 & 1 \\ 
+NewLoss & 0.39772727 & 3 \\ 
+PERF-MAP & 0.20000000 & 4 \\ 
+Q.D.KNN & 0.50000000 & 3 \\ 
+RandomForest & 0.42236025 & 8 \\ 
+RankAggNDCG & 0.87837838 & 3 \\ 
+RankBoost & 0.39357430 & 17 \\ 
+RankDE & 0.18181818 & 1 \\ 
+RankELM (pairwise) & 0.69444444 & 1 \\ 
+RankELM (pointwise) & 0.80555556 & 1 \\ 
+RankNet & 0.59154930 & 5 \\ 
+Rank-PMBGP & 0.27272727 & 1 \\ 
+RankRLS & 0.36842105 & 2 \\ 
+RankSVM & 0.44957983 & 17 \\ 
+RankSVM-Primal & 0.45911950 & 7 \\ 
+RankSVM-Struct & 0.44670051 & 9 \\ 
+RCP & 0.74074074 & 3 \\ 
+Ridge Regression & 0.36477987 & 7 \\ 
+RSRank & 0.62765957 & 4 \\ 
+SmoothGrad & 0.38461538 & 2 \\ 
+SmoothRank & 0.64150943 & 7 \\ 
+SoftRank & 0.61111111 & 1 \\ 
+SortNet & 0.56666667 & 4 \\ 
+SparseRank & 0.79439252 & 4 \\ 
+SVD-RankBoost & 0.55555556 & 3 \\ 
+SVMMAP & 0.35911602 & 8 \\ 
+SwarmRank & 0.09090909 & 1 \\ 
+TGRank & 0.50000000 & 4 \\
+\end{tabular}
+\caption{Raw data of Normalised Winning Number comparison on NDCG@10}
+\label{tab:raw_data_norm_winnum_NDCG10}
+\end{table}
+
+\section{Raw Data for MAP Comparison}
+\label{app:norm_winnum_map}
+
+\begin{table}
+\begin{tabular}{l|l|l}
+Method & Normalised Winning Number & \# of datasets \\
+\hline
+Method & Normalised Winning Number & \# of datasets \\ 
+AdaRank-MAP & 0.320610687 & 12 \\ 
+AdaRank-NDCG & 0.286259542 & 12 \\ 
+ApproxAP & 0.500000000 & 2 \\ 
+BagBoo & 0.654545455 & 2 \\ 
+BL-MART & 0.803571429 & 3 \\ 
+BoltzRank-Pair & 0.580419580 & 5 \\ 
+BoltzRank-Single & 0.433566434 & 5 \\ 
+BT & 0.750000000 & 3 \\ 
+CCRank & 0.615384615 & 2 \\ 
+FenchelRank & 0.641791045 & 5 \\ 
+FRank & 0.262295082 & 11 \\ 
+FSMRank & 0.578947368 & 7 \\ 
+FSMSVM & 0.350000000 & 4 \\ 
+GAS-E & 0.410000000 & 4 \\ 
+GP & 0.500000000 & 2 \\ 
+GPRank & 0.817307692 & 3 \\ 
+GroupCE & 0.721153846 & 3 \\ 
+GroupMLE & 0.653846154 & 3 \\ 
+IntervalRank & 0.315789474 & 1 \\ 
+IPRank & 0.851351351 & 6 \\ 
+KeepRank & 0.538461538 & 3 \\ 
+LAC-MR-OR & 0.764192140 & 12 \\ 
+LambdaMART & 0.678571429 & 3 \\ 
+LARF & 0.980769231 & 3 \\ 
+Linear Regression & 0.065000000 & 8 \\ 
+ListMLE & 0.009615385 & 3 \\ 
+ListNet & 0.450381679 & 12 \\ 
+ListReg & 0.432692308 & 3 \\ 
+LRUF & 0.968000000 & 4 \\ 
+MCP & 0.571428571 & 2 \\ 
+MHR & 0.000000000 & 1 \\ 
+MultiStageBoost & 0.136363636 & 2 \\ 
+OWPC & 0.624113475 & 6 \\ 
+PERF-MAP & 0.768000000 & 4 \\ 
+PermuRank & 0.409090909 & 3 \\ 
+Q.D.KNN & 0.558441558 & 3 \\ 
+RandomForest & 0.438888889 & 8 \\ 
+RankAggNDCG & 0.792207792 & 3 \\ 
+RankBoost & 0.313432836 & 14 \\ 
+RankCSA & 0.916666667 & 2 \\ 
+RankDE & 1.000000000 & 1 \\ 
+RankELM (pairwise) & 0.514285714 & 2 \\ 
+RankELM (pointwise) & 0.542857143 & 2 \\ 
+RankMGP & 0.222222222 & 1 \\ 
+Rank-PMBGP & 0.875000000 & 1 \\ 
+RankSVM & 0.340000000 & 13 \\ 
+RankSVM-Primal & 0.351955307 & 7 \\ 
+RankSVM-Struct & 0.362385321 & 9 \\ 
+RCP & 0.363636364 & 3 \\ 
+REF-SHF-SDCG & 0.657894737 & 1 \\ 
+RE-QR & 0.865921788 & 7 \\ 
+Ridge Regression & 0.290502793 & 7 \\ 
+RSRank & 0.660000000 & 4 \\ 
+SmoothRank & 0.530726257 & 7 \\ 
+SortNet & 0.500000000 & 2 \\ 
+SVD-RankBoost & 0.568181818 & 3 \\ 
+SVMMAP & 0.349775785 & 10 \\ 
+SwarmRank & 0.125000000 & 1 \\ 
+TGRank & 0.460000000 & 4 \\ 
+TM & 0.613636364 & 3 \\ 
+VFLR & 0.974358974 & 2 \\
+\end{tabular}
+\caption{Raw data of Normalised Winning Number comparison on MAP}
+\label{tab:raw_data_norm_winnum_map}
+\end{table}
+
+\section{Raw Data on Normalised Winning Number for cross-comparison}
+\label{app:norm_winnum_all}
+
+\begin{table}
+\begin{tabular}{l|l|l|l}
+Method & Winning Number & Ideal Winning Number & Normalized Winning Number \\
+\hline
+AdaRank-MAP & 332 & 937 & 0.35432231 \\ 
+AdaRank-NDCG & 293 & 951 & 0.30809674 \\ 
+ADMM & 4 & 9 & 0.44444444 \\ 
+ApproxAP & 33 & 66 & 0.50000000 \\ 
+ApproxNDCG & 92 & 115 & 0.80000000 \\ 
+BagBoo & 96 & 126 & 0.76190476 \\ 
+Best Single Feature & 26 & 161 & 0.16149068 \\ 
+BL-MART & 83 & 102 & 0.81372549 \\ 
+BoltzRank-Pair & 254 & 348 & 0.72988506 \\ 
+BoltzRank-Single & 213 & 348 & 0.61206897 \\ 
+BT & 75 & 99 & 0.75757576 \\ 
+C-CRF & 19 & 20 & 0.95000000 \\ 
+CA & 15 & 23 & 0.65217391 \\ 
+CCRank & 24 & 39 & 0.61538462 \\ 
+CoList & 2 & 7 & 0.28571429 \\ 
+Consistent-RankCosine & 10 & 13 & 0.76923077 \\ 
+DCMP & 320 & 584 & 0.54794521 \\ 
+DirectRank & 12 & 13 & 0.92307692 \\ 
+EnergyNDCG & 50 & 130 & 0.38461538 \\ 
+FBPCRank & 81 & 167 & 0.48502994 \\ 
+FenchelRank & 368 & 504 & 0.73015873 \\ 
+FocusedBoost & 73 & 143 & 0.51048951 \\ 
+FocusedNet & 94 & 143 & 0.65734266 \\ 
+FocusedSVM & 55 & 143 & 0.38461538 \\ 
+FP-Rank & 18 & 20 & 0.90000000 \\ 
+FRank & 242 & 839 & 0.28843862 \\ 
+FSMRank & 365 & 481 & 0.75883576 \\ 
+FSM$^{SVM}$ & 148 & 388 & 0.38144330 \\ 
+GAS-E & 166 & 388 & 0.42783505 \\ 
+GP & 7 & 12 & 0.58333333 \\ 
+GPRank & 290 & 376 & 0.77127660 \\ 
+GRankRLS & 11 & 38 & 0.28947368 \\ 
+GroupCE & 207 & 285 & 0.72631579 \\ 
+GroupMLE & 172 & 285 & 0.60350877 \\ 
+IntervalRank & 50 & 117 & 0.42735043 \\ 
+IPRank & 357 & 420 & 0.85000000 \\ 
+KeepRank & 56 & 104 & 0.53846154 \\ 
+Kernel-PCA RankBoost & 26 &  91 & 0.28571429 \\ 
+KL-CRF &  44 & 75 & 0.58666667 \\ 
+LAC-MR-OR & 179 & 235  & 0.76170213 \\ 
+LambdaMART & 54 & 81 & 0.66666667 \\ 
+LambdaNeuralRank & 15 & 15 & 1.00000000 \\ 
+LambdaRank & 10 & 24 & 0.41666667 \\ 
+LARF & 371 & 376 & 0.98670213 \\ 
+Linear Regression & 63 & 761  & 0.08278581 \\ 
+ListMLE & 3 & 199 & 0.01507538 \\ 
+ListNet & 460 & 928 & 0.49568966 \\ 
+ListReg & 176 & 288 & 0.61111111 \\ 
+LRUF & 447 & 457 & 0.97811816 \\ 
+MCP & 40 & 70 & 0.57142857 \\ 
+MHR & 17 & 41 & 0.41463415 \\ 
+MultiStageBoost & 6  & 44 & 0.13636364 \\ 
+NewLoss & 122 & 272 & 0.44852941 \\ 
+OWPC & 166 & 261 & 0.63601533 \\ 
+PERF-MAP & 191 & 457 & 0.41794311 \\ 
+PermuRank & 18 & 44 & 0.40909091 \\ 
+Q.D.KNN & 105 & 229 & 0.45851528 \\ 
+RandomForest & 147 & 341 & 0.43108504 \\ 
+Rank-PMBGP & 27 & 40 & 0.67500000 \\ 
+RankAggNDCG & 165 & 229 & 0.72052402 \\ 
+RankBoost & 309 & 939 & 0.32907348 \\ 
+RankCSA & 33 & 36 & 0.91666667 \\ 
+RankDE & 25 & 40 & 0.62500000 \\ 
+RankELM (pairwise) & 111 & 185 & 0.60000000 \\ 
+RankELM (pointwise) & 122 & 185 & 0.65945946 \\ 
+RankMGP & 4 & 18 & 0.22222222 \\ 
+RankNet & 66 & 173 & 0.38150289 \\ 
+RankRLS & 14 & 38 & 0.36842105 \\ 
+RankSVM & 323 & 885 & 0.36497175 \\ 
+RankSVM-Primal & 283 & 687 & 0.41193595 \\ 
+RankSVM-Struct & 315 & 793 & 0.39276808 \\ 
+RCP & 55 & 104 & 0.52884615 \\ 
+RE-QR & 155 & 179 & 0.86592179 \\ 
+REG-SHF-SDCG & 58 & 117 & 0.49572650 \\ 
+Ridge Regression & 226 & 650 & 0.34769231 \\ 
+RSRank & 232 & 388 & 0.59793814 \\ 
+SmoothGrad & 5 & 13 & 0.38461538 \\ 
+SmoothRank & 390 & 650 & 0.60000000 \\ 
+SoftRank & 42 & 115 & 0.36521739 \\ 
+SortNet & 113 & 238 & 0.47478992 \\ 
+SparseRank & 258 & 318 & 0.81132075 \\ 
+SVD-RankBoost & 49 & 104 & 0.47115385 \\ 
+SVM$^{MAP}$ & 254 & 734 & 0.34604905 \\ 
+SwarmRank & 5 & 40 & 0.12500000 \\ 
+TGRank & 205 & 388 & 0.52835052 \\ 
+TM & 65 & 99 & 0.65656566 \\ 
+VFLR & 38 & 39 & 0.97435897 \\
+\end{tabular}
+\caption{Raw data of Normalised Winning Number comparison cross-metric}
+\label{tab:raw_data_norm_winnum_all}
+\end{table}
+
+\end{document}

### 2014-05-15 implementation/java/src/letor/ListNet.java

@@ -16,8 +16,8 @@ import java.util.List;
 
 public class ListNet {
     // Initialise hyper-parameters
-    private static final double   STEPSIZE   = 0.1;
-    private static final int      ITERATIONS = 10;
+    private static final double   STEPSIZE   = 0.0001;
+    private static final int      ITERATIONS = 100;
 
     public static void main(String[] args) throws Exception {
         // Connect to Pig
@@ -66,6 +66,7 @@ public class ListNet {
             for(int j=1; j<lossGradientTuple.size(); j++){
                 gradient[j-1] = Double.parseDouble(lossGradientTuple.get(i).toString());
             }
+            System.out.println("gradient: "+toParamString(gradient));
 
             // w -= gradient.value * stepSize
             for(int j=0; j<w.length; j++){

### 2014-05-15 cikm_paper/paper.tex
begin mooie tabel :)

@@
 \documentclass{sig-alternate-2013}
-
+\usepackage{longtable}
+\usepackage{booktabs}
 \begin{document}
@@
 \conferenceinfo{ACM International Conference on Information and Knowledge Management}{'14 Shanghai, China}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
+\CopyrightYear{2014} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
 %\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
@@
 
-\title{A cross benchmark comparison of Learning to Rank methods}
+\title{A cross-benchmark comparison of Learning to Rank methods}
 %
@@
 \alignauthor
-Niek Tax\titlenote{affiliated with University of Twente, but written during his time at Avanade Netherlands B.V.}\\
+Niek Tax\titlenote{author is affiliated with the University of Twente, but this paper written during his time at Avanade Netherlands B.V.}\\
        \affaddr{Avanade Netherlands B.V.}\\
@@
 
-\keywords{, }
+\keywords{}
 
@@
 
-\section{The {\secit Body} of The Paper}
-
-\subsection{Type Changes and {\subsecit Special} Characters}
+\section{Problem Statement \& Research Question}
 \label{chap:cross_comparison}
-As we have seen in Chapter~\ref{chap:benchmark_results}, the evaluation of Learning-to-Rank methods is spread over several benchmark datasets.
+Evaluation of Learning-to-Rank methods is spread over several benchmark datasets.
 However, as the Learning-to-Rank methods evaluated differs between benchmarks, no single benchmark comparison can be regarded as a conclusive argument on which Learning-to-Rank method is most accurate.\\
@@
 \begin{table}[!h]
-\begin{tabular}{l|l|l}
+\begin{tabular}{p{3.2cm}|l|p{1.8cm}}
 Benchmark & Paper & \# of forward references \\
@@
 \begin{table}[!h]
-\begin{tabular}{l|l}
+\begin{tabular}{p{4.2cm}|p{2.5cm}}
 Benchmark & Google scholar search results \\
@@
 
-\begin{table}[!h!p]
+\begin{table*}[!h!p]
 \begin{tabular}{|l|l|l||l|l|l|}
@@
 \label{tab:ltr_methods_used}
-\end{table}
+\end{table*}
 
@@
 
-\begin{figure}[!h]
+\begin{figure*}[!h]
 \includegraphics[scale=0.285]{gfx/ndcg3_winnum}
@@
 \label{fig:normalised_winning_number_ndcg3}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[!h]
+\begin{figure*}[!h]
 \includegraphics[scale=0.285]{gfx/ndcg5_winnum}
@@
 \label{fig:normalised_winning_number_ndcg5}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[!h]
+\begin{figure*}[!h]
 \includegraphics[scale=0.285]{gfx/ndcg10_winnum}
@@
 \label{fig:normalised_winning_number_ndcg10}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[!h]
+\begin{figure*}[!h]
 \includegraphics[scale=0.285]{gfx/map_winnum}
@@
 \label{fig:normalised_winning_number_map}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[!h]
+\begin{figure*}[!h]
 \includegraphics[scale=0.33]{gfx/combined_normalized_winnum}
@@
 \label{fig:normalised_winning_number_all}
-\end{figure}
+\end{figure*}
 
@@
 \bibliographystyle{abbrv}
-\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
+\bibliography{Bibliography}  % sigproc.bib is the name of the Bibliography in this case
 % You must have a proper ".bib" file
@@
 
-\begin{table}
+\begin{table*}
 \begin{tabular}{l|l|l}
@@
 \label{tab:raw_data_norm_winnum_NDCG3}
-\end{table}
+\end{table*}
 
@@
 \label{tab:raw_data_norm_winnum_NDCG5}
-\end{table}
+\end{table*}
 
@@
 \label{tab:raw_data_norm_winnum_NDCG10}
-\end{table}
+\end{table*}
 
@@
 \hline
-Method & Normalised Winning Number & \# of datasets \\ 
 AdaRank-MAP & 0.320610687 & 12 \\ 
@@
 \label{tab:raw_data_norm_winnum_map}
-\end{table}
+\end{table*}
 
@@
 \label{app:norm_winnum_all}
+\onecolumn
+\begin{table*}\centering
+\begin{longtable}{@{}rrrrrrrrrrrrrrrrr@{}}\toprule
+& \multicolumn{2}{p{0.5cm}}{NDCG@3} & \phantom{a} 
+& \multicolumn{2}{c}{NDCG@5} & \phantom{a} 
+& \multicolumn{2}{c}{NDG@10} & \phantom{a} 
+& \multicolumn{2}{c}{MAP}    & \phantom{a}
+& \multicolumn{3}{c}{CROSS}\\
+\cmidrule{2-3} \cmidrule{5-6} \cmidrule{8-9} \cmidrule{11-12} \cmidrule{14-16}
+Method & NWN & \#ds && NWN & \#ds && NWN & \#ds && NWN & \#ds && WM & IWN & NWN \\ \midrule
+AdaRank-MAP & 0.3486 & 12 && 0.3884 & 12 && 0.3648 & 13 && 0.3206 & 12 && 332 & 937 & 0.3543 \\
+\bottomrule
+\end{longtable}
+\caption{Caption}
+\end{table*}
 
-\begin{table}
-\begin{tabular}{l|l|l|l}
+\begin{longtable}{p{3.2cm}|p{1.3cm}|p{1.3cm}|p{1.4cm}}
 Method & Winning Number & Ideal Winning Number & Normalized Winning Number \\
@@
 VFLR & 38 & 39 & 0.97435897 \\
-\end{tabular}
+\bottomrule
 \caption{Raw data of Normalised Winning Number comparison cross-metric}
 \label{tab:raw_data_norm_winnum_all}
-\end{table}
+\end{longtable}
 

### 2014-05-16 cikm_paper/paper.tex
raamwerk meestertabel :)

@@
 \onecolumn
-\begin{table*}\centering
 \begin{longtable}{@{}rrrrrrrrrrrrrrrrr@{}}\toprule
@@
 AdaRank-MAP & 0.3486 & 12 && 0.3884 & 12 && 0.3648 & 13 && 0.3206 & 12 && 332 & 937 & 0.3543 \\
+AdaRank-NDCG & 0.3073 & 12 && 0.3259 & 12 && 0.3158 & 16 && 0.2863 & 12 && 293 & 951 & 0.3081 \\
+ADMM & & && & && & && & && & & \\
+ApproxAP & & && & && & && & && & & \\
+ApproxNDCG & & && & && & && & && & & \\
+BagBoo & & && & && & && & && & & \\
+Best Single Feature & & && & && & && & && & & \\
+BL-MART & & && & && & && & && & & \\
+BoltzRank-Pair & & && & && & && & && & & \\
+BoltzRank-Single & & && & && & && & && & & \\
+BT & & && & && & && & && & & \\
+C-CRF & & && & && & && & && & & \\
+CA & & && & && & && & && & & \\
+CCRank & & && & && & && & && & & \\
+CoList & & && & && & && & && & & \\
+Consistent-RankCosine & & && & && & && & && & & \\
+DCMP & & && & && & && & && & & \\
+DirectRank & & && & && & && & && & & \\
+EnergyNDCG & & && & && & && & && & & \\
+FBPCRank & & && & && & && & && & & \\
+FenchelRank & & && & && & && & && & & \\
+FocusedBoost & & && & && & && & && & & \\
+FocusedNet & & && & && & && & && & & \\
+FocusedSVM & & && & && & && & && & & \\
+FP-Rank & & && & && & && & && & & \\
+FRank & & && & && & && & && & & \\
+FSMRank & & && & && & && & && & & \\
+FSM$^{SVM}$ & & && & && & && & && & & \\
+GAS-E & & && & && & && & && & & \\
+GP & & && & && & && & && & & \\
+GPRank & & && & && & && & && & & \\
+GRankRLS & & && & && & && & && & & \\
+GroupCE & & && & && & && & && & & \\
+GroupMLE & & && & && & && & && & & \\
+IntervalRank & & && & && & && & && & & \\
+IPRank & & && & && & && & && & & \\
+KeepRank & & && & && & && & && & & \\
+Kernel-PCA RankBoost & & && & && & && & && & & \\
+KL-CRF & & && & && & && & && & & \\
+LAC-MR-OR & & && & && & && & && & & \\
+LambdaMART & & && & && & && & && & & \\
+LambdaNeuralRank & & && & && & && & && & & \\
+LambdaRank & & && & && & && & && & & \\
+LARF & & && & && & && & && & & \\
+Linear Regression & & && & && & && & && & & \\
+ListMLE & & && & && & && & && & & \\
+ListNet & & && & && & && & && & & \\
+ListReg & & && & && & && & && & & \\
+LRUF & & && & && & && & && & & \\
+MCP & & && & && & && & && & & \\
+MHR & & && & && & && & && & & \\
+MultiStageBoost & & && & && & && & && & & \\
+NewLoss & & && & && & && & && & & \\
+OWPC & & && & && & && & && & & \\
+PERF-MAP & & && & && & && & && & & \\
+PermuRank & & && & && & && & && & & \\
+Q.D.KNN & & && & && & && & && & & \\
+RandomForest & & && & && & && & && & & \\
+Rank-PMBGP & & && & && & && & && & & \\
+RankAggNDCG & & && & && & && & && & & \\
+RankBoost & & && & && & && & && & & \\
+RankCSA & & && & && & && & && & & \\
+RankDE & & && & && & && & && & & \\
+RankELM (pairwise) & & && & && & && & && & & \\
+RankELM (pointwise) & & && & && & && & && & & \\
+RankMGP & & && & && & && & && & & \\
+RankNet & & && & && & && & && & & \\
+RankRLS & & && & && & && & && & & \\
+RankSVM & & && & && & && & && & & \\
+RankSVM-Primal & & && & && & && & && & & \\
+RankSVM-Struct & & && & && & && & && & & \\
+RCP & & && & && & && & && & & \\
+RE-QR & & && & && & && & && & & \\
+REG-SHG-SDCG & & && & && & && & && & & \\
+Ridge Regression & & && & && & && & && & & \\
+RSRank & & && & && & && & && & & \\
+SmoothGrad & & && & && & && & && & & \\
+SmoothRank & & && & && & && & && & & \\
+SoftRank & & && & && & && & && & & \\
+SparseRank & & && & && & && & && & & \\
+SVD-RankBoost & & && & && & && & && & & \\
+SVM$^{MAP}$ & & && & && & && & && & & \\
+SwarmRank & & && & && & && & && & & \\
+TGRank & & && & && & && & && & & \\
+TM & & && & && & && & && & & \\
+VFLR & & && & && & && & && & & \\
 \bottomrule
+\caption{Raw data of cross-benchmark comparison}
+\label{tab:raw_data}
 \end{longtable}
-\caption{Caption}
-\end{table*}
 

### 2014-05-16 cikm_paper/paper.tex
ndcg3

--- 2014-05-15 cikm_paper/paper.tex
@@
 AdaRank-MAP & 0.3486 & 12 && 0.3884 & 12 && 0.3648 & 13 && 0.3206 & 12 && 332 & 937 & 0.3543 \\
+AdaRank-NDCG & 0.3073 & 12 && 0.3259 & 12 && 0.3158 & 16 && 0.2863 & 12 && 293 & 951 & 0.3081 \\
+ADMM & - & - && & && & && & && & & \\
+ApproxAP & - & - && & && & && & && & & \\
+ApproxNDCG & 0.7949 & 1 && & && & && & && & & \\
+BagBoo & 0.8478 & 2 && & && & && & && & & \\
+Best Single Feature & - & - && & && & && & && & & \\
+BL-MART & 0.9524 & 2 && & && & && & && & & \\
+BoltzRank-Pair & 0.8333 & 4 && & && & && & && & & \\
+BoltzRank-Single & 0.7549 & 4 && & && & && & && & & \\
+BT & 0.7273 & 3 && & && & && & && & & \\
+C-CRF & - & - && & && & && & && & & \\
+CA & - & - && & && & && & && & & \\
+CCRank & - & - && & && & && & && & & \\
+CoList & 1.0000 & 1 && & && & && & && & & \\
+Consistent-RankCosine & - & - && & && & && & && & & \\
+DCMP & 0.5459 & 9 && & && & && & && & & \\
+DirectRank & - & - && & && & && & && & & \\
+EnergyNDCG & 0.3636 & 2 && & && & && & && & & \\
+FBPCRank & 0.4146 & 3 && & && & && & && & & \\
+FenchelRank & 0.7742 & 5 && & && & && & && & & \\
+FocusedBoost & 0.3750 & 2 && & && & && & && & & \\
+FocusedNet & 0.4583 & 2 && & && & && & && & & \\
+FocusedSVM & 0.2500 & 2 && & && & && & && & & \\
+FP-Rank & - & - && & && & && & && & & \\
+FRank & 0.3085 & 11 && & && & && & && & & \\
+FSMRank & 0.8333 & 4 && & && & && & && & & \\
+FSM$^{SVM}$ & 0.2292 & 2 && & && & && & && & & \\
+GAS-E & 0.3750 & 4 && & && & && & && & & \\
+GP & - & - && & && & && & && & & \\
+GPRank & 0.8710 & 3 && & && & && & && & & \\
+GRankRLS & - & - && & && & && & && & & \\
+GroupCE & 0.7312 & 3 && & && & && & && & & \\
+GroupMLE & 0.5269 & 3 && & && & && & && & & \\
+IntervalRank & 0.5897 & 1 && & && & && & && & & \\
+IPRank & 0.9355 & 3 && & && & && & && & & \\
+KeepRank & - & - && & && & && & && & & \\
+Kernel-PCA RankBoost & & && & && & && & && & & \\
+KL-CRF & 0.5946 & 2 && & && & && & && & & \\
+LAC-MR-OR & - & - && & && & && & && & & \\
+LambdaMART & 0.5714 & 2 && & && & && & && & & \\
+LambdaNeuralRank & 1.0000 & 1 && & && & && & && & & \\
+LambdaRank & 0.2000 & 1 && & && & && & && & & \\
+LARF & 0.9892 & 3 && & && & && & && & & \\
+Linear Regression & 0.0714 & 9 && & && & && & && & & \\
+ListMLE & 0.0000 & 1 && & && & && & && & & \\
+ListNet & 0.4495 & 12 && & && & && & && & & \\
+ListReg & 0.7312 & 3 && & && & && & && & & \\
+LRUF & 0.9823 & 4 && & && & && & && & & \\
+MCP & - & - && & && & && & && & & \\
+MHR & 0.7500 & 1 && & && & && & && & & \\
+MultiStageBoost & - & - && & && & && & && & & \\
+NewLoss & 0.5161 & 3 && & && & && & && & & \\
+OWPC & 0.6500 & 6 && & && & && & && & & \\
+PERF-MAP & 0.3894 & 4 && & && & && & && & & \\
+PermuRank & - & - && & && & && & && & & \\
+Q.D.KNN & - & - && & && & && & && & & \\
+RandomForest & - & - && & && & && & && & & \\
+Rank-PMBGP & - & - && & && & && & && & & \\
+RankAggNDCG & - & - && & && & && & && & & \\
+RankBoost & 0.3211 & 12 && & && & && & && & & \\
+RankCSA & - & - && & && & && & && & & \\
+RankDE & - & - && & && & && & && & & \\
+RankELM (pairwise) & 0.6154 & 1 && & && & && & && & & \\
+RankELM (pointwise) & 0.6923 & 1 && & && & && & && & & \\
+RankMGP & - & - && & && & && & && & & \\
+RankNet & 0.1887 & 1 && & && & && & && & & \\
+RankRLS & - & - && & && & && & && & & \\
+RankSVM & 0.3010 & 12 && & && & && & && & & \\
+RankSVM-Primal & 0.3920 & 8 && & && & && & && & & \\
+RankSVM-Struct & 0.3520 & 9 && & && & && & && & & \\
+RCP & - & - && & && & && & && & & \\
+RE-QR & - & - && & && & && & && & & \\
+REG-SHG-SDCG & 0.3846 & 1 && & && & && & && & & \\
+Ridge Regression & 0.4088 & 7 && & && & && & && & & \\
+RSRank & 0.5729 & 4 && & && & && & && & & \\
+SmoothGrad & - & - && & && & && & && & & \\
+SmoothRank & 0.6038 & 7 && & && & && & && & & \\
+SoftRank & 0.2308 & 1 && & && & && & && & & \\
+sortNet & 0.2500 & 2 && & && & && & && & & \\
+SparseRank & 0.8224 & 4 && & && & && & && & & \\
+SVD-RankBoost & - & - && & && & && & && & & \\
+SVM$^{MAP}$ & 0.2893 & 7 && & && & && & && & & \\
+SwarmRank & - & - && & && & && & && & & \\
+TGRank & 0.5417 & 4 && & && & && & && & & \\
+TM & 0.5909 & 3 && & && & && & && & & \\
+VFLR & - & - && & && & && & && & & \\
 \bottomrule
+\caption{Raw data of cross-benchmark comparison}
+\label{tab:raw_data}
 \end{longtable}
-\caption{Caption}
-\end{table*}
 

### 2014-05-16 cikm_paper/paper.tex
ndcg5 in meestertabel

--- 2014-05-15 cikm_paper/paper.tex
@@
 AdaRank-MAP & 0.3486 & 12 && 0.3884 & 12 && 0.3648 & 13 && 0.3206 & 12 && 332 & 937 & 0.3543 \\
+AdaRank-NDCG & 0.3073 & 12 && 0.3259 & 12 && 0.3158 & 16 && 0.2863 & 12 && 293 & 951 & 0.3081 \\
+ADMM & - & - && - & - && & && & && & & \\
+ApproxAP & - & - && - & - && & && & && & & \\
+ApproxNDCG & 0.7949 & 1 && 0.7500 & 1 && & && & && & & \\
+BagBoo & 0.8478 & 2 && 0.8400 & 1 && & && & && & & \\
+Best Single Feature & - & - && - & - && & && & && & & \\
+BL-MART & 0.9524 & 2 && 0.7200 & 1 && & && & && & & \\
+BoltzRank-Pair & 0.8333 & 4 && 0.8350 & 4 && & && & && & & \\
+BoltzRank-Single & 0.7549 & 4 && 0.7184 & 4 && & && & && & & \\
+BT & 0.7273 & 3 && 0.7879 & 3 && & && & && & & \\
+C-CRF & - & - && 0.9500 & 2 && & && & && & & \\
+CA & - & - && - & - && & && & && & & \\
+CCRank & - & - && - & - && & && & && & & \\
+CoList & 1.0000 & 1 && - & - && & && & && & & \\
+Consistent-RankCosine & - & - && - & - && & && & && & & \\
+DCMP & 0.5459 & 9 && 0.5079 & 9 && & && & && & & \\
+DirectRank & - & - && - & - && & && & && & & \\
+EnergyNDCG & 0.3636 & 2 && 0.3778 & 2 && & && & && & & \\
+FBPCRank & 0.4146 & 3 && 0.5529 & 3 && & && & && & & \\
+FenchelRank & 0.7742 & 5 && 0.7500 & 5 && & && & && & & \\
+FocusedBoost & 0.3750 & 2 && 0.4545 & 2 && & && & && & & \\
+FocusedNet & 0.4583 & 2 && 0.6364 & 2 && & && & && & & \\
+FocusedSVM & 0.2500 & 2 && 0.2727 & 2 && & && & && & & \\
+FP-Rank & - & - && 0.9000 & 1 && & && & && & & \\
+FRank & 0.3085 & 11 && 0.2849 & 10 && & && & && & & \\
+FSMRank & 0.8333 & 4 && 0.8776 & 4 && & && & && & & \\
+FSM$^{SVM}$ & 0.2292 & 2 && 0.4082 & 4 && & && & && & & \\
+GAS-E & 0.3750 & 4 && 0.4694 & 4 && & && & && & & \\
+GP & - & - && - & - && & && & && & & \\
+GPRank & 0.8710 & 3 && 0.7253 & 3 && & && & && & & \\
+GRankRLS & - & - && - & - && & && & && & & \\
+GroupCE & 0.7312 & 3 && - & - && & && & && & & \\
+GroupMLE & 0.5269 & 3 && - & - && & && & && & & \\
+IntervalRank & 0.5897 & 1 && 0.3750 & 1 && & && & && & & \\
+IPRank & 0.9355 & 3 && 0.8132 & 3 && & && & && & & \\
+KeepRank & - & - && - & - && & && & && & & \\
+Kernel-PCA RankBoost & - & - && 0.2857 & 3 && & && & && & & \\
+KL-CRF & 0.5946 & 2 && 0.5789 & 2 && & && & && & & \\
+LAC-MR-OR & - & - && - & - && & && & && & & \\
+LambdaMART & 0.5714 & 2 && - & - && & && & && & & \\
+LambdaNeuralRank & 1.0000 & 1 && 1.0000 & 1 && & && & && & & \\
+LambdaRank & 0.2000 & 1 && 0.2000 & 1 && & && & && & & \\
+LARF & 0.9892 & 3 && 0.9890 & 3 && & && & && & & \\
+Linear Regression & 0.0714 & 9 && 0.1099 & 9 && & && & && & & \\
+ListMLE & 0.0000 & 1 && - & - && & && & && & & \\
+ListNet & 0.4495 & 12 && 0.4911 & 12 && & && & && & & \\
+ListReg & 0.7312 & 3 && 0.6923 & 3 && & && & && & & \\
+LRUF & 0.9823 & 4 && 0.9817 & 4 && & && & && & & \\
+MCP & - & - && - & - && & && & && & & \\
+MHR & 0.7500 & 1 && 0.6000 & 1 && & && & && & & \\
+MultiStageBoost & - & - && - & - && & && & && & & \\
+NewLoss & 0.5161 & 3 && 0.4286 & 3 && & && & && & & \\
+OWPC & 0.6500 & 6 && - & - && & && & && & & \\
+PERF-MAP & 0.3894 & 4 && 0.2661 & 4 && & && & && & & \\
+PermuRank & - & - && - & - && & && & && & & \\
+Q.D.KNN & - & - && 0.3205 & 3 && & && & && & & \\
+RandomForest & - & - && - & - && & && & && & & \\
+Rank-PMBGP & - & - && 0.7692 & 1 && & && & && & & \\
+RankAggNDCG & - & - && 0.5000 & 3 && & && & && & & \\
+RankBoost & 0.3211 & 12 && 0.2794 & 10 && & && & && & & \\
+RankCSA & - & - && - & - && & && & && & & \\
+RankDE & - & - && 0.5385 & 1 && & && & && & & \\
+RankELM (pairwise) & 0.6154 & 1 && 0.6500 & 1 && & && & && & & \\
+RankELM (pointwise) & 0.6923 & 1 && 0.7000 & 1 && & && & && & & \\
+RankMGP & - & - && - & - && & && & && & & \\
+RankNet & 0.1887 & 1 && 0.2857 & 3 && & && & && & & \\
+RankRLS & - & - && - & - && & && & && & & \\
+RankSVM & 0.3010 & 12 && 0.3613 & 11 && & && & && & & \\
+RankSVM-Primal & 0.3920 & 8 && 0.4509 & 8 && & && & && & & \\
+RankSVM-Struct & 0.3520 & 9 && 0.4136 & 9 && & && & && & & \\
+RCP & - & - && 0.5758 & 3 && & && & && & & \\
+RE-QR & - & - && - & - && & && & && & & \\
+REG-SHG-SDCG & 0.3846 & 1 && 0.4500 & 1 && & && & && & & \\
+Ridge Regression & 0.4088 & 7 && 0.3333 & 7 && & && & && & & \\
+RSRank & 0.5729 & 4 && 0.5306 & 4 && & && & && & & \\
+SmoothGrad & - & - && - & - && & && & && & & \\
+SmoothRank & 0.6038 & 7 && 0.6340 & 7 && & && & && & & \\
+SoftRank & 0.2308 & 1 && 0.2750 & 1 && & && & && & & \\
+sortNet & 0.2500 & 2 && 0.5147 & 4 && & && & && & & \\
+SparseRank & 0.8224 & 4 && 0.8173 & 4 && & && & && & & \\
+SVD-RankBoost & - & - && 0.2727 & 3 && & && & && & & \\
+SVM$^{MAP}$ & 0.2893 & 7 && 0.3801 & 8 && & && & && & & \\
+SwarmRank & - & - && 0.1538 & 1 && & && & && & & \\
+TGRank & 0.5417 & 4 && 0.6122 & 4 && & && & && & & \\
+TM & 0.5909 & 3 && 0.7576 & 3 && & && & && & & \\
+VFLR & - & - && - & - && & && & && & & \\
 \bottomrule
+\caption{Raw data of cross-benchmark comparison}
+\label{tab:raw_data}
 \end{longtable}
-\caption{Caption}
-\end{table*}
 

### 2014-05-16 cikm_paper/paper.tex
ndcg10 in meestertabel :)
map in meestertabel :)
meestertabel af :)
andere tabellen weg

--- 2014-05-16 cikm_paper/paper.tex
@@
 \appendix
-\section{Raw Data for NDCG@3 Comparison}
-\label{app:norm_winnum_NDCG3}
-
-\begin{table*}
-\begin{tabular}{l|l|l}
-Method & Normalised Winning Number & \# of datasets \\
-\hline
-AdaRank-MAP & 0.34862385 & 12 \\ 
-AdaRank-NDCG & 0.30733945 & 12 \\ 
-ApproxNDCG & 0.79487179 & 1 \\ 
-BagBoo & 0.84782609 & 2 \\ 
-BL-MART & 0.95238095 & 2 \\ 
-BoltzRank-Pair & 0.83333333 & 4 \\ 
-BoltzRank-Single & 0.75490196 & 4 \\ 
-BT & 0.72727273 & 3 \\ 
-CoList & 1.00000000 & 1 \\ 
-DCMP & 0.54591837 & 9 \\ 
-EnergyNDCG & 0.36363636 & 2 \\ 
-FBPCRank & 0.41463415 & 3 \\ 
-FenchelRank & 0.77419355 & 5 \\ 
-FocusedBoost & 0.37500000 & 2 \\ 
-FocusedNet & 0.45833333 & 2 \\ 
-FocusedSVM & 0.25000000 & 2 \\ 
-FRank & 0.30845771 & 11 \\ 
-FSMRank & 0.83333333 & 4 \\ 
-FSMSVM & 0.22916667 & 4 \\ 
-GAS-E & 0.37500000 & 4 \\ 
-GPRank & 0.87096774 & 3 \\ 
-GroupCE & 0.73118280 & 3 \\ 
-GroupMLE & 0.52688172 & 3 \\ 
-IntervalRank & 0.58974359 & 1 \\ 
-IPRank & 0.93548387 & 3 \\ 
-KL-CRF & 0.59459459 & 2 \\ 
-LambdaMART & 0.57142857 & 2 \\ 
-LambdaNeuralRank & 1.00000000 & 1 \\ 
-LambdaRank & 0.20000000 & 1 \\ 
-LARF & 0.98924731 & 3 \\ 
-Linear Regression & 0.07142857 & 9 \\ 
-ListMLE & 0.00000000 & 1 \\ 
-ListNet & 0.44954128 & 12 \\ 
-ListReg & 0.73118280 & 3 \\ 
-LRUF & 0.98230088 & 4 \\ 
-MHR & 0.75000000 & 1 \\ 
-NewLoss & 0.51612903 & 3 \\ 
-OWPC & 0.65000000 & 6 \\ 
-PERF-MAP & 0.38938053 & 4 \\ 
-RankBoost & 0.32110092 & 12 \\ 
-RankELM (pairwise) & 0.61538462 & 1 \\ 
-RankELM (pointwise) & 0.69230769 & 1 \\ 
-RankNet & 0.18867925 & 1 \\ 
-RankSVM & 0.30097087 & 12 \\ 
-RankSVM-Primal & 0.39204545 & 8 \\ 
-RankSVM-Struct & 0.35204082 & 9 \\ 
-REG-SHF-SDCG & 0.38461538 & 1 \\ 
-Ridge Regression & 0.40880503 & 7 \\ 
-RSRank & 0.57291667 & 4 \\ 
-SmoothRank & 0.60377358 & 7 \\ 
-SoftRank & 0.23076923 & 1 \\ 
-SortNet & 0.25000000 & 2 \\ 
-SparseRank & 0.82242991 & 4 \\ 
-SVMMAP & 0.28930818 & 7 \\ 
-TGRank & 0.54166667 & 4 \\ 
-TM & 0.59090909 & 3 \\
-\end{tabular}
-\caption{Raw data of Normalised Winning Number comparison on NDCG@3}
-\label{tab:raw_data_norm_winnum_NDCG3}
-\end{table*}
-
-\section{Raw Data for NDCG@5 Comparison}
-\label{app:norm_winnum_NDCG5}
-
-\begin{table*}
-\begin{tabular}{l|l|l}
-Method & Normalised Winning Number & \# of datasets \\
-\hline
-AdaRank-MAP & 0.3883929 & 12 \\ 
-AdaRank-NDCG & 0.3258929 & 12 \\ 
-ApproxNDCG & 0.7500000 & 1 \\ 
-BagBoo & 0.8400000 & 1 \\ 
-BL-MART & 0.7200000 & 1 \\ 
-BoltzRank-Pair & 0.8349515 & 4 \\ 
-BoltzRank-Single & 0.7184466 & 4 \\ 
-BT & 0.7878788 & 3 \\ 
-C-CRF & 0.9500000 & 2 \\ 
-DCMP & 0.5078534 & 9 \\ 
-EnergyNDCG & 0.3777778 & 2 \\ 
-FBPCRank & 0.5529412 & 3 \\ 
-FenchelRank & 0.7500000 & 5 \\ 
-FocusedBoost & 0.4545455 & 2 \\ 
-FocusedNet & 0.6363636 & 2 \\ 
-FocusedSVM & 0.2727273 & 2 \\ 
-FPRank & 0.9000000 & 1 \\ 
-FRank & 0.2849462 & 10 \\ 
-FSMRank & 0.8775510 & 4 \\ 
-FSMSVM & 0.4081633 & 4 \\ 
-GAS-E & 0.4693878 & 4 \\ 
-GPRank & 0.7252747 & 3 \\ 
-IntervalRank & 0.3750000 & 1 \\ 
-IPRank & 0.8131868 & 3 \\ 
-Kernel-PCA & 0.2857143 & 3 \\ 
-KL-CRF & 0.5789474 & 2 \\ 
-LambdaNeuralRank & 1.0000000 & 1 \\ 
-LambdaRank & 0.2000000 & 1 \\ 
-LARF & 0.9890110 & 3 \\ 
-Linear Regression & 0.1099476 & 9 \\ 
-ListNet & 0.4910714 & 12 \\ 
-ListReg & 0.6923077 & 3 \\ 
-LRUF & 0.9816514 & 4 \\ 
-MHR & 0.6000000 & 1 \\ 
-NewLoss & 0.4285714 & 3 \\ 
-PERF-MAP & 0.2660550 & 4 \\ 
-Q.D.KNN & 0.3205128 & 3 \\ 
-RankAggNDCG & 0.5000000 & 3 \\ 
-RankBoost & 0.2794118 & 10 \\ 
-RankDE & 0.5384615 & 1 \\ 
-RankELM (pairwise) & 0.6500000 & 1 \\ 
-RankELM (pointwise) & 0.7000000 & 1 \\ 
-RankNet & 0.2857143 & 3 \\ 
-Rank-PMBGP & 0.7692308 & 1 \\ 
-RankSVM & 0.3612565 & 11 \\ 
-RankSVM-Primal & 0.4508671 & 8 \\ 
-RankSVM-Struct & 0.4136126 & 9 \\ 
-RCP & 0.5757576 & 3 \\ 
-REG-SHF-SDCG & 0.4500000 & 1 \\ 
-Ridge Regression & 0.3333333 & 7 \\ 
-RSRank & 0.5306122 & 4 \\ 
-SmoothRank & 0.6339869 & 7 \\ 
-SoftRank & 0.2750000 & 1 \\ 
-SortNet & 0.5147059 & 4 \\ 
-SparseRank & 0.8173077 & 4 \\ 
-SVD-RankBoost & 0.2727273 & 3 \\ 
-SVMMAP & 0.3801170 & 8 \\ 
-SwarmRank & 0.1538462 & 1 \\ 
-TGRank & 0.6122449 & 4 \\ 
-TM & 0.7575758 & 3 \\ 
-\end{tabular}
-\caption{Raw data of Normalised Winning Number comparison on NDCG@5}
-\label{tab:raw_data_norm_winnum_NDCG5}
-\end{table*}
-
-\section{Raw Data for NDCG@10 Comparison}
-\label{app:norm_winnum_NDCG10}
-
-\begin{table*}
-\begin{tabular}{l|l|l}
-Method & Normalised Winning Number & \# of datasets \\
-\hline
-AdaRank-MAP & 0.36480687 & 13 \\ 
-AdaRank-NDCG & 0.31578947 & 16 \\ 
-ADMM & 0.44444444 & 1 \\ 
-ApproxNDCG & 0.86111111 & 1 \\ 
-Best Single Feature & 0.16149068 & 8 \\ 
-CA & 0.65217391 & 4 \\ 
-CoList & 0.16666667 & 1 \\ 
-Consistent-RankCosine & 0.76923077 & 2 \\ 
-DCMP & 0.58883249 & 9 \\ 
-DirectRank & 0.92307692 & 2 \\ 
-EnergyNDCG & 0.41463415 & 2 \\ 
-FenchelRank & 0.76229508 & 5 \\ 
-FocusedBoost & 0.68627451 & 2 \\ 
-FocusedNet & 0.86274510 & 2 \\ 
-FocusedSVM & 0.60784314 & 2 \\ 
-FRank & 0.30288462 & 11 \\ 
-FSMRank & 0.86206897 & 5 \\ 
-FSMSVM & 0.54255319 & 4 \\ 
-GAS-E & 0.45744681 & 4 \\ 
-GP & 0.66666667 & 2 \\ 
-GPRank & 0.65909091 & 3 \\ 
-GRankRLS & 0.28947368 & 2 \\ 
-GroupCE & 0.72727273 & 3 \\ 
-GroupMLE & 0.62500000 & 3 \\ 
-IPRank & 0.79545455 & 3 \\ 
-LAC-MR-OR & 0.66666667 & 2 \\ 
-LambdaMART & 1.00000000 & 1 \\ 
-LambdaNeuralRank & 1.00000000 & 1 \\ 
-LambdaRank & 0.57142857 & 2 \\ 
-LARF & 0.98863636 & 3 \\ 
-Linear Regression & 0.08287293 & 8 \\ 
-ListMLE & 0.02127660 & 4 \\ 
-ListNet & 0.59821429 & 12 \\ 
-LRUF & 0.98181818 & 4 \\ 
-MHR & 0.62500000 & 1 \\ 
-NewLoss & 0.39772727 & 3 \\ 
-PERF-MAP & 0.20000000 & 4 \\ 
-Q.D.KNN & 0.50000000 & 3 \\ 
-RandomForest & 0.42236025 & 8 \\ 
-RankAggNDCG & 0.87837838 & 3 \\ 
-RankBoost & 0.39357430 & 17 \\ 
-RankDE & 0.18181818 & 1 \\ 
-RankELM (pairwise) & 0.69444444 & 1 \\ 
-RankELM (pointwise) & 0.80555556 & 1 \\ 
-RankNet & 0.59154930 & 5 \\ 
-Rank-PMBGP & 0.27272727 & 1 \\ 
-RankRLS & 0.36842105 & 2 \\ 
-RankSVM & 0.44957983 & 17 \\ 
-RankSVM-Primal & 0.45911950 & 7 \\ 
-RankSVM-Struct & 0.44670051 & 9 \\ 
-RCP & 0.74074074 & 3 \\ 
-Ridge Regression & 0.36477987 & 7 \\ 
-RSRank & 0.62765957 & 4 \\ 
-SmoothGrad & 0.38461538 & 2 \\ 
-SmoothRank & 0.64150943 & 7 \\ 
-SoftRank & 0.61111111 & 1 \\ 
-SortNet & 0.56666667 & 4 \\ 
-SparseRank & 0.79439252 & 4 \\ 
-SVD-RankBoost & 0.55555556 & 3 \\ 
-SVMMAP & 0.35911602 & 8 \\ 
-SwarmRank & 0.09090909 & 1 \\ 
-TGRank & 0.50000000 & 4 \\
-\end{tabular}
-\caption{Raw data of Normalised Winning Number comparison on NDCG@10}
-\label{tab:raw_data_norm_winnum_NDCG10}
-\end{table*}
-
-\section{Raw Data for MAP Comparison}
-\label{app:norm_winnum_map}
-
-\begin{table*}
-\begin{tabular}{l|l|l}
-Method & Normalised Winning Number & \# of datasets \\
-\hline
-AdaRank-MAP & 0.320610687 & 12 \\ 
-AdaRank-NDCG & 0.286259542 & 12 \\ 
-ApproxAP & 0.500000000 & 2 \\ 
-BagBoo & 0.654545455 & 2 \\ 
-BL-MART & 0.803571429 & 3 \\ 
-BoltzRank-Pair & 0.580419580 & 5 \\ 
-BoltzRank-Single & 0.433566434 & 5 \\ 
-BT & 0.750000000 & 3 \\ 
-CCRank & 0.615384615 & 2 \\ 
-FenchelRank & 0.641791045 & 5 \\ 
-FRank & 0.262295082 & 11 \\ 
-FSMRank & 0.578947368 & 7 \\ 
-FSMSVM & 0.350000000 & 4 \\ 
-GAS-E & 0.410000000 & 4 \\ 
-GP & 0.500000000 & 2 \\ 
-GPRank & 0.817307692 & 3 \\ 
-GroupCE & 0.721153846 & 3 \\ 
-GroupMLE & 0.653846154 & 3 \\ 
-IntervalRank & 0.315789474 & 1 \\ 
-IPRank & 0.851351351 & 6 \\ 
-KeepRank & 0.538461538 & 3 \\ 
-LAC-MR-OR & 0.764192140 & 12 \\ 
-LambdaMART & 0.678571429 & 3 \\ 
-LARF & 0.980769231 & 3 \\ 
-Linear Regression & 0.065000000 & 8 \\ 
-ListMLE & 0.009615385 & 3 \\ 
-ListNet & 0.450381679 & 12 \\ 
-ListReg & 0.432692308 & 3 \\ 
-LRUF & 0.968000000 & 4 \\ 
-MCP & 0.571428571 & 2 \\ 
-MHR & 0.000000000 & 1 \\ 
-MultiStageBoost & 0.136363636 & 2 \\ 
-OWPC & 0.624113475 & 6 \\ 
-PERF-MAP & 0.768000000 & 4 \\ 
-PermuRank & 0.409090909 & 3 \\ 
-Q.D.KNN & 0.558441558 & 3 \\ 
-RandomForest & 0.438888889 & 8 \\ 
-RankAggNDCG & 0.792207792 & 3 \\ 
-RankBoost & 0.313432836 & 14 \\ 
-RankCSA & 0.916666667 & 2 \\ 
-RankDE & 1.000000000 & 1 \\ 
-RankELM (pairwise) & 0.514285714 & 2 \\ 
-RankELM (pointwise) & 0.542857143 & 2 \\ 
-RankMGP & 0.222222222 & 1 \\ 
-Rank-PMBGP & 0.875000000 & 1 \\ 
-RankSVM & 0.340000000 & 13 \\ 
-RankSVM-Primal & 0.351955307 & 7 \\ 
-RankSVM-Struct & 0.362385321 & 9 \\ 
-RCP & 0.363636364 & 3 \\ 
-REF-SHF-SDCG & 0.657894737 & 1 \\ 
-RE-QR & 0.865921788 & 7 \\ 
-Ridge Regression & 0.290502793 & 7 \\ 
-RSRank & 0.660000000 & 4 \\ 
-SmoothRank & 0.530726257 & 7 \\ 
-SortNet & 0.500000000 & 2 \\ 
-SVD-RankBoost & 0.568181818 & 3 \\ 
-SVMMAP & 0.349775785 & 10 \\ 
-SwarmRank & 0.125000000 & 1 \\ 
-TGRank & 0.460000000 & 4 \\ 
-TM & 0.613636364 & 3 \\ 
-VFLR & 0.974358974 & 2 \\
-\end{tabular}
-\caption{Raw data of Normalised Winning Number comparison on MAP}
-\label{tab:raw_data_norm_winnum_map}
-\end{table*}
-
-\section{Raw Data on Normalised Winning Number for cross-comparison}
-\label{app:norm_winnum_all}
+\section{Cross-Comparison Raw Data}
+\label{app:raw_data}
 \onecolumn
-\begin{longtable}{@{}rrrrrrrrrrrrrrrrr@{}}\toprule
+\begin{longtable}{@{}lrrrrrrrrrrrrrrrr@{}}\toprule
 & \multicolumn{2}{p{0.5cm}}{NDCG@3} & \phantom{a} 
@@
 AdaRank-NDCG & 0.3073 & 12 && 0.3259 & 12 && 0.3158 & 16 && 0.2863 & 12 && 293 & 951 & 0.3081 \\
-ADMM & & && & && & && & && & & \\
-ApproxAP & & && & && & && & && & & \\
-ApproxNDCG & & && & && & && & && & & \\
-BagBoo & & && & && & && & && & & \\
-Best Single Feature & & && & && & && & && & & \\
-BL-MART & & && & && & && & && & & \\
-BoltzRank-Pair & & && & && & && & && & & \\
-BoltzRank-Single & & && & && & && & && & & \\
-BT & & && & && & && & && & & \\
-C-CRF & & && & && & && & && & & \\
-CA & & && & && & && & && & & \\
-CCRank & & && & && & && & && & & \\
-CoList & & && & && & && & && & & \\
-Consistent-RankCosine & & && & && & && & && & & \\
-DCMP & & && & && & && & && & & \\
-DirectRank & & && & && & && & && & & \\
-EnergyNDCG & & && & && & && & && & & \\
-FBPCRank & & && & && & && & && & & \\
-FenchelRank & & && & && & && & && & & \\
-FocusedBoost & & && & && & && & && & & \\
-FocusedNet & & && & && & && & && & & \\
-FocusedSVM & & && & && & && & && & & \\
-FP-Rank & & && & && & && & && & & \\
-FRank & & && & && & && & && & & \\
-FSMRank & & && & && & && & && & & \\
-FSM$^{SVM}$ & & && & && & && & && & & \\
-GAS-E & & && & && & && & && & & \\
-GP & & && & && & && & && & & \\
-GPRank & & && & && & && & && & & \\
-GRankRLS & & && & && & && & && & & \\
-GroupCE & & && & && & && & && & & \\
-GroupMLE & & && & && & && & && & & \\
-IntervalRank & & && & && & && & && & & \\
-IPRank & & && & && & && & && & & \\
-KeepRank & & && & && & && & && & & \\
-Kernel-PCA RankBoost & & && & && & && & && & & \\
-KL-CRF & & && & && & && & && & & \\
-LAC-MR-OR & & && & && & && & && & & \\
-LambdaMART & & && & && & && & && & & \\
-LambdaNeuralRank & & && & && & && & && & & \\
-LambdaRank & & && & && & && & && & & \\
-LARF & & && & && & && & && & & \\
-Linear Regression & & && & && & && & && & & \\
-ListMLE & & && & && & && & && & & \\
-ListNet & & && & && & && & && & & \\
-ListReg & & && & && & && & && & & \\
-LRUF & & && & && & && & && & & \\
-MCP & & && & && & && & && & & \\
-MHR & & && & && & && & && & & \\
-MultiStageBoost & & && & && & && & && & & \\
-NewLoss & & && & && & && & && & & \\
-OWPC & & && & && & && & && & & \\
-PERF-MAP & & && & && & && & && & & \\
-PermuRank & & && & && & && & && & & \\
-Q.D.KNN & & && & && & && & && & & \\
-RandomForest & & && & && & && & && & & \\
-Rank-PMBGP & & && & && & && & && & & \\
-RankAggNDCG & & && & && & && & && & & \\
-RankBoost & & && & && & && & && & & \\
-RankCSA & & && & && & && & && & & \\
-RankDE & & && & && & && & && & & \\
-RankELM (pairwise) & & && & && & && & && & & \\
-RankELM (pointwise) & & && & && & && & && & & \\
-RankMGP & & && & && & && & && & & \\
-RankNet & & && & && & && & && & & \\
-RankRLS & & && & && & && & && & & \\
-RankSVM & & && & && & && & && & & \\
-RankSVM-Primal & & && & && & && & && & & \\
-RankSVM-Struct & & && & && & && & && & & \\
-RCP & & && & && & && & && & & \\
-RE-QR & & && & && & && & && & & \\
-REG-SHG-SDCG & & && & && & && & && & & \\
-Ridge Regression & & && & && & && & && & & \\
-RSRank & & && & && & && & && & & \\
-SmoothGrad & & && & && & && & && & & \\
-SmoothRank & & && & && & && & && & & \\
-SoftRank & & && & && & && & && & & \\
-SparseRank & & && & && & && & && & & \\
-SVD-RankBoost & & && & && & && & && & & \\
-SVM$^{MAP}$ & & && & && & && & && & & \\
-SwarmRank & & && & && & && & && & & \\
-TGRank & & && & && & && & && & & \\
-TM & & && & && & && & && & & \\
-VFLR & & && & && & && & && & & \\
+ADMM & - & - && - & - && 0.4444 & 1 && - & - && 4 & 9 & 0.4444 \\
+ApproxAP & - & - && - & - && - & - && 0.5000 & 2 && 33 & 66 & 0.5000 \\
+ApproxNDCG & 0.7949 & 1 && 0.7500 & 1 && 0.8611 & 1 && - & - && 92 & 115 & 0.8000 \\
+BagBoo & 0.8478 & 2 && 0.8400 & 1 && - & - && 0.6545 & 2 && 96 & 126 & 0.7619 \\
+Best Single Feature & - & - && - & - && 0.1615 & 8 && - & - && 26 & 161 & 0.1615 \\
+BL-MART & 0.9524 & 2 && 0.7200 & 1 && - & - && 0.8036 & 3 && 83 & 102 & 0.8137 \\
+BoltzRank-Pair & 0.8333 & 4 && 0.8350 & 4 && - & - && 0.5804 & 5 && 254 & 348 & 0.7299 \\
+BoltzRank-Single & 0.7549 & 4 && 0.7184 & 4 && - & - && 0.4336 & 5 && 213 & 348 & 0.6121 \\
+BT & 0.7273 & 3 && 0.7879 & 3 && - & - && 0.7500 & 3 && 75 & 99 & 0.7576 \\
+C-CRF & - & - && 0.9500 & 2 && - & - && - & - && 19 & 20 & 0.9500 \\
+CA & - & - && - & - && 0.6522 & 4 && - & - && 15 & 23 & 0.6522 \\
+CCRank & - & - && - & - && - & - && 0.6154 & 2 && 24 & 39 & 0.6154 \\
+CoList & 1.0000 & 1 && - & - && 0.1667 & 1 && - & - && 2 & 7 & 0.2857 \\
+Consistent-RankCosine & - & - && - & - && 0.7692 & 2 && - & - && 10 & 13 & 0.7692 \\
+DCMP & 0.5459 & 9 && 0.5079 & 9 && 0.5888 & 9 && - & - && 320 & 584 & 0.5479 \\
+DirectRank & - & - && - & - && 0.9231 & 2 && - & - && 12 & 13 & 0.9231 \\
+EnergyNDCG & 0.3636 & 2 && 0.3778 & 2 && 0.4146 & 2 && - & - && 50 & 130 & 0.3846 \\
+FBPCRank & 0.4146 & 3 && 0.5529 & 3 && - & - && - & - && 81 & 167 & 0.4850 \\
+FenchelRank & 0.7742 & 5 && 0.7500 & 5 && 0.7623 & 5 && 0.6418 & 5 && 368 & 504 & 0.7302 \\
+FocusedBoost & 0.3750 & 2 && 0.4545 & 2 && 0.6863 & 2 && - & - && 73 & 143 & 0.5105 \\
+FocusedNet & 0.4583 & 2 && 0.6364 & 2 && 0.8627 & 2 && - & - && 94 & 143 & 0.6573 \\
+FocusedSVM & 0.2500 & 2 && 0.2727 & 2 && 0.6078 & 2 && - & - && 55 & 143 & 0.3846 \\
+FP-Rank & - & - && 0.9000 & 1 && - & - && - & - && 18 & 20 & 0.9000 \\
+FRank & 0.3085 & 11 && 0.2849 & 10 && 0.3029 & 11 && 0.2623 & 11 && 242 & 839 & 0.2884 \\
+FSMRank & 0.8333 & 4 && 0.8776 & 4 && 0.8621 & 5 && 0.5789 & 7 && 365 & 481 & 0.7588 \\
+FSM$^{SVM}$ & 0.2292 & 2 && 0.4082 & 4 && 0.5426 & 4 && 0.3500 & 4 && 148 & 388 & 0.3814 \\
+GAS-E & 0.3750 & 4 && 0.4694 & 4 && 0.4574 & 4 && 0.4100 & 4 && 166 & 388 & 0.4278 \\
+GP & - & - && - & - && 0.6667 & 2 && 0.5000 & 2 && 7 & 12 & 0.5833 \\
+GPRank & 0.8710 & 3 && 0.7253 & 3 && 0.6591 & 3 && 0.8173 & 3 && 290 & 376 & 0.7713 \\
+GRankRLS & - & - && - & - && 0.2895 & 2 && - & - && 11 & 38 & 0.2895 \\
+GroupCE & 0.7312 & 3 && - & - && 0.7273 & 3 && 0.7212 & 3 && 207 & 285 & 0.7263 \\
+GroupMLE & 0.5269 & 3 && - & - && 0.6250 & 3 && 0.6538 & 3 && 172 & 285 & 0.6035 \\
+IntervalRank & 0.5897 & 1 && 0.3750 & 1 && - & - && 0.3158 & 1 && 50 & 117 & 0.4274 \\
+IPRank & 0.9355 & 3 && 0.8132 & 3 && 0.7955 & 3 && 0.8514 & 6 && 357 & 420 & 0.8500 \\
+KeepRank & - & - && - & - && - & - && 0.5385 & 3 && 56 & 104 & 0.5385 \\
+KL-CRF & 0.5946 & 2 && 0.5789 & 2 && - & - && - & - && 44 & 75 & 0.5867 \\
+LAC-MR-OR & - & - && - & - && 0.6667 & 2 && 0.7642 & 12 && 179 & 235 & 0.7617 \\
+LambdaMART & 0.5714 & 2 && - & - && 1.0000 & 1 && 0.6786 & 3 && 54 & 81 & 0.6667 \\
+LambdaNeuralRank & 1.0000 & 1 && 1.0000 & 1 && 1.0000 & 1 && - & - && 15 & 15 & 1.0000 \\
+LambdaRank & 0.2000 & 1 && 0.2000 & 1 && 0.5714 & 2 && - & - && 10 & 24 & 0.4167 \\
+LARF & 0.9892 & 3 && 0.9890 & 3 && 0.9886 & 3 && 0.9808 & 3 && 371 & 376 & 0.9867 \\
+Linear Regression & 0.0714 & 9 && 0.1099 & 9 && 0.0829 & 8 && 0.0650 & 8 && 63 & 761 & 0.0828 \\
+ListMLE & 0.0000 & 1 && - & - && 0.0213 & 4 && 0.00962 & 3 && 3 & 199 & 0.0151 \\
+ListNet & 0.4495 & 12 && 0.4911 & 12 && 0.5982 & 12 && 0.4504 & 12 && 460 & 928 & 0.496 \\
+ListReg & 0.7312 & 3 && 0.6923 & 3 && - & - && 0.4327 & 3 && 176 & 288 & 0.6111 \\
+LRUF & 0.9823 & 4 && 0.9817 & 4 && 0.9818 & 4 && 0.9680 & 4 && 447 & 457 & 0.9781 \\
+MCP & - & - && - & - && - & - && 0.5714 & 2 && 40 & 70 & 0.5714 \\
+MHR & 0.7500 & 1 && 0.6000 & 1 && 0.6250 & 1 && 0.0000 & 1 && 17 & 41 & 0.5714 \\
+MultiStageBoost & - & - && - & - && - & - && 0.1364 & 2 && 6 & 44 & 0.1364 \\
+NewLoss & 0.5161 & 3 && 0.4286 & 3 && 0.3977 & 3 && - & - && 122 & 272 & 0.4485 \\
+OWPC & 0.6500 & 6 && - & - && - & - && 0.6241 & 6 && 166 & 261 & 0.6360 \\
+PERF-MAP & 0.3894 & 4 && 0.2661 & 4 && 0.2000 & 4 && 0.7680 & 4 && 191 & 457 & 0.4179 \\
+PermuRank & - & - && - & - && - & - && 0.4091 & 3 && 18 & 44 & 0.4091 \\
+Q.D.KNN & - & - && 0.3205 & 3 && 0.5000 & 3 && 0.5584 & 3 && 105 & 229 & 0.4585 \\
+RandomForest & - & - && - & - && 0.4224 & 8 && 0.4389 & 8 && 147 & 341 & 0.4311 \\
+Rank-PMBGP & - & - && 0.7692 & 1 && 0.2727 & 1 && 0.8750 & 1 && 24 & 40 & 0.6750 \\
+RankAggNDCG & - & - && 0.5000 & 3 && 0.8784 & 3 && 0.7922 & 3 && 165 & 229 & 0.7205 \\
+RankBoost & 0.3211 & 12 && 0.2794 & 10 && 0.3936 & 17 && 0.3134 & 14 && 309 & 939 & 0.3291 \\
+RankBoost (Kernel-PCA) & - & - && 0.2857 & 3 && - & - && - & - && 26 & 91 & 0.2857 \\
+RankBoost (SVD) & - & - && 0.2727 & 3 && 0.5556 & 3 && 0.5682 & 3 && 49 & 104 & 0.4712 \\
+RankCSA & - & - && - & - && - & - && 0.9167 & 2 && 33 & 36 & 0.9167 \\
+RankDE & - & - && 0.5385 & 1 && 0.1818 & 1 && 1.0000 & 1 && 25 & 40 & 0.6250 \\
+RankELM (pairwise) & 0.6154 & 1 && 0.6500 & 1 && 0.6944 & 1 && 0.5143 & 2 && 111 & 185 & 0.6000 \\
+RankELM (pointwise) & 0.6923 & 1 && 0.7000 & 1 && 0.8056 & 1 && 0.5429 & 2 && 122 & 185 & 0.6595 \\
+RankMGP & - & - && - & - && - & - && 0.2222 & 1 && 4 & 18 & 0.2222 \\
+RankNet & 0.1887 & 1 && 0.2857 & 3 && 0.5915 & 5 && - & - && 66 & 173 & 0.3815 \\
+RankRLS & - & - && - & - && 0.3684 & 2 && - & - && 14 & 38 & 0.3684 \\
+RankSVM & 0.3010 & 12 && 0.3613 & 11 && 0.4496 & 17 && 0.3400 & 13 && 323 & 885 & 0.3650 \\
+RankSVM-Primal & 0.3920 & 8 && 0.4509 & 8 && 0.4591 & 7 && 0.3520 & 7 && 283 & 687 & 0.4119 \\
+RankSVM-Struct & 0.3520 & 9 && 0.4136 & 9 && 0.4467 & 9 && 0.3624 & 9 && 315 & 793 & 0.3928 \\
+RCP & - & - && 0.5758 & 3 && 0.7407 & 3 && 0.3636 & 3 && 55 & 104 & 0.5288 \\
+RE-QR & - & - && - & - && - & - && 0.8659 & 7 && 155 & 179 & 0.8659 \\
+REG-SHG-SDCG & 0.3846 & 1 && 0.4500 & 1 && - & - && 0.6579 & 1 && 58 & 117 & 0.4957 \\
+Ridge Regression & 0.4088 & 7 && 0.3333 & 7 && 0.3648 & 7 && 0.2905 & 7 && 226 & 650 & 0.3477 \\
+RSRank & 0.5729 & 4 && 0.5306 & 4 && 0.6277 & 4 && 0.6600 & 4 && 232 & 388 & 0.5979 \\
+SmoothGrad & - & - && - & - && 0.3846 & 2 && - & - && 5 & 13 & 0.3846 \\
+SmoothRank & 0.6038 & 7 && 0.6340 & 7 && 0.6415 & 7 && 0.5307 & 7 && 390 & 650 & 0.6000 \\
+SoftRank & 0.2308 & 1 && 0.2750 & 1 && 0.6111 & 1 && - & - && 42 & 115 & 0.3652 \\
+SortNet & 0.2500 & 2 && 0.5147 & 4 && 0.5667 & 4 && 0.5000 & 2 && 113 & 238 & 0.4748 \\
+SparseRank & 0.8224 & 4 && 0.8173 & 4 && 0.7944 & 4 && - & - && 258 & 318 & 0.8113 \\
+SVM$^{MAP}$ & 0.2893 & 7 && 0.3801 & 8 && 0.3591 & 8 && 0.3498 & 10 && 254 & 734 & 0.3460 \\
+SwarmRank & - & - && 0.1538 & 1 && 0.0909 & 1 && 0.1250 & 1 && 5 & 40 & 0.1250 \\
+TGRank & 0.5417 & 4 && 0.6122 & 4 && 0.5000 & 4 && 0.4600 & 4 && 205 & 388 & 0.5284 \\
+TM & 0.5909 & 3 && 0.7576 & 3 && - & - && 0.6136 & 3 && 65 & 99 & 0.6566 \\
+VFLR & - & - && - & - && - & - && 0.9744 & 2 && 38 & 39 & 0.9744 \\
 \bottomrule
@@
 \end{longtable}
-
-\begin{longtable}{p{3.2cm}|p{1.3cm}|p{1.3cm}|p{1.4cm}}
-Method & Winning Number & Ideal Winning Number & Normalized Winning Number \\
-\hline
-AdaRank-MAP & 332 & 937 & 0.35432231 \\ 
-AdaRank-NDCG & 293 & 951 & 0.30809674 \\ 
-ADMM & 4 & 9 & 0.44444444 \\ 
-ApproxAP & 33 & 66 & 0.50000000 \\ 
-ApproxNDCG & 92 & 115 & 0.80000000 \\ 
-BagBoo & 96 & 126 & 0.76190476 \\ 
-Best Single Feature & 26 & 161 & 0.16149068 \\ 
-BL-MART & 83 & 102 & 0.81372549 \\ 
-BoltzRank-Pair & 254 & 348 & 0.72988506 \\ 
-BoltzRank-Single & 213 & 348 & 0.61206897 \\ 
-BT & 75 & 99 & 0.75757576 \\ 
-C-CRF & 19 & 20 & 0.95000000 \\ 
-CA & 15 & 23 & 0.65217391 \\ 
-CCRank & 24 & 39 & 0.61538462 \\ 
-CoList & 2 & 7 & 0.28571429 \\ 
-Consistent-RankCosine & 10 & 13 & 0.76923077 \\ 
-DCMP & 320 & 584 & 0.54794521 \\ 
-DirectRank & 12 & 13 & 0.92307692 \\ 
-EnergyNDCG & 50 & 130 & 0.38461538 \\ 
-FBPCRank & 81 & 167 & 0.48502994 \\ 
-FenchelRank & 368 & 504 & 0.73015873 \\ 
-FocusedBoost & 73 & 143 & 0.51048951 \\ 
-FocusedNet & 94 & 143 & 0.65734266 \\ 
-FocusedSVM & 55 & 143 & 0.38461538 \\ 
-FP-Rank & 18 & 20 & 0.90000000 \\ 
-FRank & 242 & 839 & 0.28843862 \\ 
-FSMRank & 365 & 481 & 0.75883576 \\ 
-FSM$^{SVM}$ & 148 & 388 & 0.38144330 \\ 
-GAS-E & 166 & 388 & 0.42783505 \\ 
-GP & 7 & 12 & 0.58333333 \\ 
-GPRank & 290 & 376 & 0.77127660 \\ 
-GRankRLS & 11 & 38 & 0.28947368 \\ 
-GroupCE & 207 & 285 & 0.72631579 \\ 
-GroupMLE & 172 & 285 & 0.60350877 \\ 
-IntervalRank & 50 & 117 & 0.42735043 \\ 
-IPRank & 357 & 420 & 0.85000000 \\ 
-KeepRank & 56 & 104 & 0.53846154 \\ 
-Kernel-PCA RankBoost & 26 &  91 & 0.28571429 \\ 
-KL-CRF &  44 & 75 & 0.58666667 \\ 
-LAC-MR-OR & 179 & 235  & 0.76170213 \\ 
-LambdaMART & 54 & 81 & 0.66666667 \\ 
-LambdaNeuralRank & 15 & 15 & 1.00000000 \\ 
-LambdaRank & 10 & 24 & 0.41666667 \\ 
-LARF & 371 & 376 & 0.98670213 \\ 
-Linear Regression & 63 & 761  & 0.08278581 \\ 
-ListMLE & 3 & 199 & 0.01507538 \\ 
-ListNet & 460 & 928 & 0.49568966 \\ 
-ListReg & 176 & 288 & 0.61111111 \\ 
-LRUF & 447 & 457 & 0.97811816 \\ 
-MCP & 40 & 70 & 0.57142857 \\ 
-MHR & 17 & 41 & 0.41463415 \\ 
-MultiStageBoost & 6  & 44 & 0.13636364 \\ 
-NewLoss & 122 & 272 & 0.44852941 \\ 
-OWPC & 166 & 261 & 0.63601533 \\ 
-PERF-MAP & 191 & 457 & 0.41794311 \\ 
-PermuRank & 18 & 44 & 0.40909091 \\ 
-Q.D.KNN & 105 & 229 & 0.45851528 \\ 
-RandomForest & 147 & 341 & 0.43108504 \\ 
-Rank-PMBGP & 27 & 40 & 0.67500000 \\ 
-RankAggNDCG & 165 & 229 & 0.72052402 \\ 
-RankBoost & 309 & 939 & 0.32907348 \\ 
-RankCSA & 33 & 36 & 0.91666667 \\ 
-RankDE & 25 & 40 & 0.62500000 \\ 
-RankELM (pairwise) & 111 & 185 & 0.60000000 \\ 
-RankELM (pointwise) & 122 & 185 & 0.65945946 \\ 
-RankMGP & 4 & 18 & 0.22222222 \\ 
-RankNet & 66 & 173 & 0.38150289 \\ 
-RankRLS & 14 & 38 & 0.36842105 \\ 
-RankSVM & 323 & 885 & 0.36497175 \\ 
-RankSVM-Primal & 283 & 687 & 0.41193595 \\ 
-RankSVM-Struct & 315 & 793 & 0.39276808 \\ 
-RCP & 55 & 104 & 0.52884615 \\ 
-RE-QR & 155 & 179 & 0.86592179 \\ 
-REG-SHF-SDCG & 58 & 117 & 0.49572650 \\ 
-Ridge Regression & 226 & 650 & 0.34769231 \\ 
-RSRank & 232 & 388 & 0.59793814 \\ 
-SmoothGrad & 5 & 13 & 0.38461538 \\ 
-SmoothRank & 390 & 650 & 0.60000000 \\ 
-SoftRank & 42 & 115 & 0.36521739 \\ 
-SortNet & 113 & 238 & 0.47478992 \\ 
-SparseRank & 258 & 318 & 0.81132075 \\ 
-SVD-RankBoost & 49 & 104 & 0.47115385 \\ 
-SVM$^{MAP}$ & 254 & 734 & 0.34604905 \\ 
-SwarmRank & 5 & 40 & 0.12500000 \\ 
-TGRank & 205 & 388 & 0.52835052 \\ 
-TM & 65 & 99 & 0.65656566 \\ 
-VFLR & 38 & 39 & 0.97435897 \\
-\bottomrule
-\caption{Raw data of Normalised Winning Number comparison cross-metric}
-\label{tab:raw_data_norm_winnum_all}
-\end{longtable}
-
 \end{document}

### 2014-05-16 cikm_paper/paper.tex
Index term
formule past :)

@@
 % A category with the (minimum) three required fields
-\category{H.4}{Information Systems Applications}{Miscellaneous}
-%A category including the fourth, optional field follows...
-\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+\category{H.3.3}{Information Systems}{Information Storage and Retrieval}[Information Search and Retrieval]
 
-\terms{}
-
-\keywords{}
+\keywords{Learning to rank, Information retrieval, Benchmark datasets}
 
@@
 
-$I_{\{M_i(j)>M_k(j)\}} = \begin{cases}
-1 & \text{if } M_i(j) \text{ and } M_k(j) \text{ are both defined and } M_i(j) > M_k(j), \\
-0 & \text{otherwise}
-\end{cases}$\\
+\[
+I_{M_i(j)>M_k(j)} = \begin{cases}
+1 & \parbox[t]{5cm}{if $M_i(j)$ and $M_k(j$ are both defined and $M_i(j) > M_k(j)$,}\\
+0 & \parbox[t]{5cm}{otherwise}
+\end{cases}
+\]
 
@@
 
-\begin{figure*}[!h]
+\begin{figure*}[!htp]
 \includegraphics[scale=0.285]{gfx/ndcg3_winnum}
@@
 
-\begin{figure*}[!h]
+\begin{figure*}[!htp]
 \includegraphics[scale=0.285]{gfx/ndcg5_winnum}
@@
 
-\begin{figure*}[!h]
+\begin{figure*}[!htp]
 \includegraphics[scale=0.285]{gfx/ndcg10_winnum}
@@
 
-\begin{figure*}[!h]
+\begin{figure*}[!htp]
 \includegraphics[scale=0.285]{gfx/map_winnum}
@@
 
-\begin{figure*}[!h]
+\begin{figure*}[!htp]
 \includegraphics[scale=0.33]{gfx/combined_normalized_winnum}

### 2014-05-16 cikm_paper/paper.tex
Learning-to-Rank -> learning to rank

@@
 \section{Introduction}
+Ranking is a core problem in the field of information retrieval.
+The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a given query.
+Ranking has become a vital part of web search, where commercial search engines help users find their need in the extremely large collection of the World Wide Web.
+One can find useful applications of ranking in many application domains outside web search as well.
+For example, it plays a vital role in amongst others: automatic document summarisation, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}.
+In addition, the ranking task has been argued to be more fitting to recommender systems than regression-based rating prediction on a continuous scale \cite{Adomavicius2005,McNee2006}.\\
+
+Research in the field of ranking models has for a long time been based on manually designed ranking functions, such as the well-known BM25 model \cite{Robertson1994}.
+Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
+The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models.
+Learning to Rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\
+
+In recent years, several learning to rank benchmark datasets have been proposed that enable comparison of the performance of different learning to rank methods.
+Well-known benchmark datasets include the \emph{Yahoo!
+Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR dataset \cite{Qin2010} that was published by Microsoft Research.\\
+
+No agreement exists between authors in the learning to rank field on the benchmark sets used for model evaluation.
+Comparing ranking performance of learning to rank methods is largely hindered by the existence of this variety of evaluation datasets and the lack of a \emph{de facto} standard way of benchmarking new models.
 
@@
 \label{chap:cross_comparison}
-Evaluation of Learning-to-Rank methods is spread over several benchmark datasets.
-However, as the Learning-to-Rank methods evaluated differs between benchmarks, no single benchmark comparison can be regarded as a conclusive argument on which Learning-to-Rank method is most accurate.\\
+Because the learning to rank methods evaluated differs between benchmarks, no single benchmark comparison can be regarded as a conclusive argument on which learning to rank method is most accurate.\\
 
-Several studies make a small start in considering Learning-to-Rank methods performance over multiple benchmark datasets.
+Several studies make a small start in considering learning to rank methods performance over multiple benchmark datasets.
 Gomes et al.
@@
 To my knowledge, no structured meta-analysis on ranking accuracy has been conducted where evaluation results on several benchmark collections are taken into account.
-With a meta-analysis we will compare the performance of Learning-to-Rank methods across the Learning-to-Rank benchmark datasets described in foregoing sections.
+With a meta-analysis we will compare the performance of learning to rank methods across the learning to rank benchmark datasets described in foregoing sections.
 
@@
 Users of those benchmark collections are encouraged to cite these papers.
-Therefore, we collect evaluation measurements of Learning-to-Rank methods on these benchmark collections through forward literature search.
+Therefore, we collect evaluation measurements of learning to rank methods on these benchmark collections through forward literature search.
 Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search.
@@
 \end{tabular}
-\caption{Forward references of Learning-to-Rank benchmark papers}
+\caption{Forward references of learning to rank benchmark papers}
 \label{tbl:ltr_benchmark_forref}
@@
 The LETOR 4.0, MSLR-web10/30k and Yandex Internet Mathematics Competition 2009 benchmark collections were not accompanied with a describing study.
-To collect measurements of Learning-to-Rank methods evaluated on these benchmarks, a Google Scholar search is performed on the name of the benchmark.
+To collect measurements of learning to rank methods evaluated on these benchmarks, a Google Scholar search is performed on the name of the benchmark.
 Table~\ref{chap:benchmark_results} shows the results of this literature search.
@@
 \end{tabular}
-\caption{Google scholar search results for Learning-to-Rank benchmarks}
+\caption{Google scholar search results for learning to rank benchmarks}
 \label{tbl:ltr_benchmark_searchres}
@@
 The LETOR 3.0 paper \cite{Qin2010} states that it may differ between datasets what the most accurate ranking methods are.
-To evaluate the overall performance of Learning-to-Rank methods over the multiple datasets in the LETOR 3.0 collections, Qin et al.
+To evaluate the overall performance of learning to rank methods over the multiple datasets in the LETOR 3.0 collections, Qin et al.
 \cite{Qin2010} proposed a measure called \emph{winning number} as the number of other algorithms that an algorithm can beat over the set of datasets.
@@
 \section{Evaluation Results Found in Literature}
-Table \ref{tab:ltr_methods_used} gives an overview of the Learning-to-Rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
+Table \ref{tab:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
 Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, LETOR 3.0 and LETOR 4.0 baselines respectively.\\
@@
 \item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011, Lin2012}
-\item The study focussed on a different Learning-to-Rank task (e.g.
+\item The study focussed on a different learning to rank task (e.g.
 rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}
@@
 \end{tabular}
-\caption{Learning-to-Rank algorithms with measurements on benchmark datasets}
+\caption{Learning to rank algorithms with measurements on benchmark datasets}
 \label{tab:ltr_methods_used}
@@
 \section{Results \& Discussion}
-The following subsections provide the performance of Learning-to-Rank methods in terms of Normalised Winning Number for NDCG@3, NDCG@5, NDCG@10 and MAP.
-Performance of the Learning-to-Rank methods is plotted with Normalised Winning Number on the vertical axis and the number of datasets on which the method has been evaluated on the horizontal axis.
-The further to the right, the more certain we can be about the performance of the Learning-to-Rank method.
+The following subsections provide the performance of learning to rank methods in terms of Normalised Winning Number for NDCG@3, NDCG@5, NDCG@10 and MAP.
+Performance of the learning to rank methods is plotted with Normalised Winning Number on the vertical axis and the number of datasets on which the method has been evaluated on the horizontal axis.
+The further to the right, the more certain we can be about the performance of the learning to rank method.
 The methods for which it holds that there is no other method that has 1) a higher Normalised Winning Number and 2) a higher number datasets evaluated, are identified as the best performing methods and are labelled with the name of the method.
@@
 \subsection{NDCG@3}
-Figure \ref{fig:normalised_winning_number_ndcg3} shows the performance of Learning-to-Rank methods for the NDCG@3 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg3} in Appendix \ref{app:norm_winnum_ndcg3} provides the raw Normalised Winning Number data for the Learning-to-Rank methods for which NDCG@3 evaluation results were available.\\
+Figure \ref{fig:normalised_winning_number_ndcg3} shows the performance of learning to rank methods for the NDCG@3 metric.
+Table \ref{tab:raw_data_norm_winnum_ndcg3} in Appendix \ref{app:norm_winnum_ndcg3} provides the raw Normalised Winning Number data for the learning to rank methods for which NDCG@3 evaluation results were available.\\
 
@@
 \includegraphics[scale=0.285]{gfx/ndcg3_winnum}
-\caption{NDCG@3 comparison of Learning-to-Rank methods}
+\caption{NDCG@3 comparison of learning to rank methods}
 \label{fig:normalised_winning_number_ndcg3}
@@
 \subsection{NDCG@5}
-Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of Learning-to-Rank methods for the NDCG@5 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg5} in Appendix \ref{app:norm_winnum_ndcg5} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.\\
+Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of learning to rank methods for the NDCG@5 metric.
+Table \ref{tab:raw_data_norm_winnum_ndcg5} in Appendix \ref{app:norm_winnum_ndcg5} provides the raw Normalised Winning Number data for the learning to rank methods.\\
 
@@
 \includegraphics[scale=0.285]{gfx/ndcg5_winnum}
-\caption{NDCG@5 comparison of Learning-to-Rank methods}
+\caption{NDCG@5 comparison of learning to rank methods}
 \label{fig:normalised_winning_number_ndcg5}
@@
 \subsection{NDCG@10}
-Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of Learning-to-Rank methods for the NDCG@10 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg10} in Appendix \ref{app:norm_winnum_ndcg10} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.\\
+Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of learning to rank methods for the NDCG@10 metric.
+Table \ref{tab:raw_data_norm_winnum_ndcg10} in Appendix \ref{app:norm_winnum_ndcg10} provides the raw Normalised Winning Number data for the learning to rank methods.\\
 
@@
 \includegraphics[scale=0.285]{gfx/ndcg10_winnum}
-\caption{NDCG@10 comparison of Learning-to-Rank methods}
+\caption{NDCG@10 comparison of learning to rank methods}
 \label{fig:normalised_winning_number_ndcg10}
@@
 \subsection{MAP}
-Figure \ref{fig:normalised_winning_number_map} shows the performance of Learning-to-Rank methods for the MAP metric.
-Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.\\
+Figure \ref{fig:normalised_winning_number_map} shows the performance of learning to rank methods for the MAP metric.
+Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw Normalised Winning Number data for the learning to rank methods.\\
 
@@
 \includegraphics[scale=0.285]{gfx/map_winnum}
-\caption{MAP comparison of Learning-to-Rank methods}
+\caption{MAP comparison of learning to rank methods}
 \label{fig:normalised_winning_number_map}
@@
 No reported NDCG performance was found in the literature study for RE-QR.
-There is a lot of certainty on the accuracy of RankBoost and RankSVM as both models are evaluated on the majority of datasets included in the comparison for the MAP metric, but given their Normalised Winning Number it can said that both methods are not within the top performing Learning-to-Rank methods.
+There is a lot of certainty on the accuracy of RankBoost and RankSVM as both models are evaluated on the majority of datasets included in the comparison for the MAP metric, but given their Normalised Winning Number it can said that both methods are not within the top performing learning to rank methods.
 
@@
 \includegraphics[scale=0.33]{gfx/combined_normalized_winnum}
-\caption{Cross-benchmark comparison of Learning-to-Rank methods}
+\caption{Cross-benchmark comparison of learning to rank methods}
 \label{fig:normalised_winning_number_all}
@@
 In the Normalised Winning Number calculation the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
-By calculating it in this way, we implicitly make the assumption that the Learning-to-Rank methods are (approximately) distributed uniformly over the benchmarks, such that the average Learning-to-Rank method tested are approximately equally hard for each data set.
-It could be the case however that this assumption is false and that significantly more accurate Learning-to-Rank methods are evaluated on some datasets than on other datasets.
+By calculating it in this way, we implicitly make the assumption that the learning to rank methods are (approximately) distributed uniformly over the benchmarks, such that the average learning to rank method tested are approximately equally hard for each data set.
+It could be the case however that this assumption is false and that significantly more accurate learning to rank methods are evaluated on some datasets than on other datasets.
 \\
 
-A second limitation is that the datasets on which Learning-to-Rank methods have been evaluated can not always be regarded a random choice.
-It might be the case that some researchers chose to publish results for exactly those benchmark datasets that showed the most positive results for their Learning-to-Rank method.\\
+A second limitation is that the datasets on which learning to rank methods have been evaluated can not always be regarded a random choice.
+It might be the case that some researchers chose to publish results for exactly those benchmark datasets that showed the most positive results for their learning to rank method.\\
 
@@
 \begin{description}
-\item[RQ1] What are the best performing Learning-to-Rank algorithms in terms of accuracy on relevant benchmark datasets?\\
+\item[RQ1] What are the best performing learning to rank algorithms in terms of accuracy on relevant benchmark datasets?\\
 \end{description}
-Although no closing arguments can be formulated on which Learning-to-Rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.\\
+Although no closing arguments can be formulated on which learning to rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.\\
 
-LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the Learning-to-Rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
+LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the learning to rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
 From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.

### 2014-05-16 cikm_paper/paper.tex
tabel staat nu mooi na referenties

@@
 
-\title{A cross-benchmark comparison of Learning to Rank methods}
+\title{A Cross-Benchmark Comparison of Learning to Rank Methods}
 %
@@
 Ranking is a core problem in the field of information retrieval.
-The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a given query.
+The ranking task in information retrieval entails the ranking of candidate documents according to their relevance to a given query.
 Ranking has become a vital part of web search, where commercial search engines help users find their need in the extremely large collection of the World Wide Web.
 One can find useful applications of ranking in many application domains outside web search as well.
-For example, it plays a vital role in amongst others: automatic document summarisation, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}.
-In addition, the ranking task has been argued to be more fitting to recommender systems than regression-based rating prediction on a continuous scale \cite{Adomavicius2005,McNee2006}.\\
+For example, ranking plays a vital role in automatic text summarisation, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}.
+In addition, the ranking task has been argued to be a better fit for recommender systems compared to the regression task on a continuous scale \cite{Adomavicius2005,McNee2006}.\\
 
@@
 Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
-The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models.
-Learning to Rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\
+The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective ranking models.
+Learning to rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\
 
-In recent years, several learning to rank benchmark datasets have been proposed that enable comparison of the performance of different learning to rank methods.
-Well-known benchmark datasets include the \emph{Yahoo!
-Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR dataset \cite{Qin2010} that was published by Microsoft Research.\\
+In recent years, several learning to rank benchmark datasets have been proposed with the aim of enabling comparison of the performance of different learning to rank methods.
+Well-known benchmark datasets in the learning to rank field include the \emph{Yahoo!
+Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, the LETOR dataset \cite{Qin2010}, and LETOR's successor the MSLR (Microsoft Learning to Rank) datasets.\\
 
 No agreement exists between authors in the learning to rank field on the benchmark sets used for model evaluation.
-Comparing ranking performance of learning to rank methods is largely hindered by the existence of this variety of evaluation datasets and the lack of a \emph{de facto} standard way of benchmarking new models.
-
-\section{Problem Statement \& Research Question}
-\label{chap:cross_comparison}
-Because the learning to rank methods evaluated differs between benchmarks, no single benchmark comparison can be regarded as a conclusive argument on which learning to rank method is most accurate.\\
+Comparing ranking performance of learning to rank methods is largely hindered by this lack of a \emph{de facto} standard way of benchmarking new models.
 
@@
 \cite{Busa-Fekete2013} compared the accuracy of a small set of models over the LETOR 4.0 datasets, both MSLR datasets, both Yahoo!
-Learning to Rank Challenge datasets and the OHSUMED dataset from LETOR 3.0.
-To my knowledge, no structured meta-analysis on ranking accuracy has been conducted where evaluation results on several benchmark collections are taken into account.
-With a meta-analysis we will compare the performance of learning to rank methods across the learning to rank benchmark datasets described in foregoing sections.
+Learning to Rank Challenge datasets and one of the datasets from LETOR 3.0.\\
+
+To our knowledge, no structured meta-analysis on ranking accuracy has been conducted where evaluation results on several benchmark collections are taken into account.
+In this paper we will perform a meta-analysis with the aim of comparing the ranking accuracy of learning to rank methods.
 
 \section{Collecting Evaluation Results}
+
+\subsection{Evaluation Results Collection Methodology}
 \label{sec:collecting_evaluation_results}
@@
 
-\section{Comparison Methodology}
+\subsection{Comparison Methodology of Learning to Rank Methods}
 The LETOR 3.0 paper \cite{Qin2010} states that it may differ between datasets what the most accurate ranking methods are.
@@
 
-\section{Evaluation Results Found in Literature}
+\section{Results \& Discussion}
+\subsection{Collection of Evaluation Results}
 Table \ref{tab:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
@@
 
-\begin{table*}[!h!p]
+\begin{table*}
 \begin{tabular}{|l|l|l||l|l|l|}
@@
 
-\section{Results \& Discussion}
+\subsection{Comparison of Learning to Rank models}
 The following subsections provide the performance of learning to rank methods in terms of Normalised Winning Number for NDCG@3, NDCG@5, NDCG@10 and MAP.
@@
 
-\subsection{NDCG@3}
+\subsubsection{NDCG@3}
 Figure \ref{fig:normalised_winning_number_ndcg3} shows the performance of learning to rank methods for the NDCG@3 metric.
@@
 
-\subsection{NDCG@5}
+\subsubsection{NDCG@5}
 Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of learning to rank methods for the NDCG@5 metric.
@@
 
-\subsection{NDCG@10}
+\subsubsection{NDCG@10}
 Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of learning to rank methods for the NDCG@10 metric.
@@
 
-\begin{figure*}[!htp]
+\begin{figure*}
 \includegraphics[scale=0.285]{gfx/ndcg10_winnum}
@@
 
-\subsection{MAP}
+\subsubsection{MAP}
 Figure \ref{fig:normalised_winning_number_map} shows the performance of learning to rank methods for the MAP metric.
@@
 
-\begin{figure*}[!htp]
+\begin{figure*}
 \includegraphics[scale=0.285]{gfx/map_winnum}
@@
 
-\subsection{Cross-metric}
+\subsubsection{Cross-metric}
 Figure \ref{fig:normalised_winning_number_all} shows the Normalised Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}.
@@
 
-\begin{figure*}[!htp]
+\begin{figure*}
 \includegraphics[scale=0.33]{gfx/combined_normalized_winnum}
@@
 %\balancecolumns
+\onecolumn
+\clearpage
+\newpage
 \appendix
@@
 \label{app:raw_data}
-\onecolumn
-\begin{longtable}{@{}lrrrrrrrrrrrrrrrr@{}}\toprule
+
+\begin{longtable}[!h]{@{}lrrrrrrrrrrrrrrrr@{}}\toprule
 & \multicolumn{2}{p{0.5cm}}{NDCG@3} & \phantom{a} 

### 2014-05-19 cikm_paper/paper.tex
werk

@@
 Well-known benchmark datasets in the learning to rank field include the \emph{Yahoo!
-Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, the LETOR dataset \cite{Qin2010}, and LETOR's successor the MSLR (Microsoft Learning to Rank) datasets.\\
+Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the \emph{Yandex Internet Mathematics 2009} contest \footnote{http://imat2009.yandex.ru/en}, the LETOR dataset \cite{Qin2010}, and the MSLR (Microsoft Learning to Rank) datasets \footnote{http://research.microsoft.com/en-us/projects/mslr/}.\\
 
-No agreement exists between authors in the learning to rank field on the benchmark sets used for model evaluation.
-Comparing ranking performance of learning to rank methods is largely hindered by this lack of a \emph{de facto} standard way of benchmarking new models.
+No agreement exists between authors in the learning to rank field on which benchmark set(s) to use for model evaluation.
+Comparing ranking performance of learning to rank methods is largely hindered by this lack of a \emph{de facto} standard way of benchmarking new models.\\
 
-Several studies make a small start in considering learning to rank methods performance over multiple benchmark datasets.
 Gomes et al.
@@
 \cite{Busa-Fekete2013} compared the accuracy of a small set of models over the LETOR 4.0 datasets, both MSLR datasets, both Yahoo!
-Learning to Rank Challenge datasets and one of the datasets from LETOR 3.0.\\
-
+Learning to Rank Challenge datasets and one of the datasets from LETOR 3.0.
+Both studies did not aim to be complete in benchmark datasets and learning to rank methods included in their comparisons.
 To our knowledge, no structured meta-analysis on ranking accuracy has been conducted where evaluation results on several benchmark collections are taken into account.
 In this paper we will perform a meta-analysis with the aim of comparing the ranking accuracy of learning to rank methods.
+The paper will describe two stages in the meta-analysis process: 1) collection of evaluation results, and 2) comparison of learning to rank methods.
 
 \section{Collecting Evaluation Results}
-
-\subsection{Evaluation Results Collection Methodology}
 \label{sec:collecting_evaluation_results}
-With a literature review we will collect evaluation results on the datasets/collections. The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
+Through a literature review we will collect evaluation results on the datasets/collections.
+The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
 \begin{itemize}\itemsep0em
@@
 \item Yahoo! Learning to Rank Challenge
-\item Yandex Internet Mathematics Competition 2009
+\item Yandex Internet Mathematics 2009 contest
 \item MSLR-web10/30k
@@
 \begin{table}[!h]
-\begin{tabular}{p{3.2cm}|l|p{1.8cm}}
+\begin{tabular}{p{3.2cm}lp{1.8cm}}\toprule
 Benchmark & Paper & \# of forward references \\
-\hline
+\midrule
 LETOR 1.0 \& 2.0 & Liu et al. \cite{Liu2007b} & 307\\
@@
 WCL2R & Alc{\^a}ntara et al. \cite{Alcantara2010} & 2\\
+\bottomrule
 \end{tabular}
@@
 \begin{table}[!h]
-\begin{tabular}{p{4.2cm}|p{2.5cm}}
+\begin{tabular}{p{4.2cm}p{2.5cm}}\toprule
 Benchmark & Google scholar search results \\
-\hline
+\midrule
 LETOR 4.0 & 75 results \\
@@
 MSLR-web30k & 15 results \\
-Yandex Internet Mathematics Competition & 1 result \\ 
+Yandex Internet Mathematics & 3 results \\
+\bottomrule
 \end{tabular}
@@
 
-\subsection{Comparison Methodology of Learning to Rank Methods}
+\subsection{Literature selection}
+Table \ref{tab:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
+Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, LETOR 3.0 and LETOR 4.0 baselines respectively.\\
+
+Some studies with evaluation results found through in literature review were not usable for the meta-analysis.
+The following enumeration enumerates those properties that made one or more studies unusable for the meta-analysis.
+Between brackets are the studies that these properties apply to.
+
+\begin{enumerate}
+\item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011, Lin2012}
+\item The study focussed on a different learning to rank task (e.g.
+rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}
+\item The study used an altered version of a benchmark that contained additional features \cite{Bidoki2009, Ding2010}
+\item The study provides no exact data of the evaluation results (e.g.
+results are only in graphical form) \cite{Wang2008, Wang2010, Xu2010, Kuo2009, Li2008, Xia2008, Zhou2011, Wu2011, Zhu2009, Karimzadehgan2011, Swersky2012, Pan2011, Ni2008, Ciaramita2008, Stewart2012, Petterson2009, Agarwal2010, Chang2009, Qin2008c, Adams2011, Sculley2009, Huang2008, Alejo2010, Sun2011, He2010b, Benbouzid2012, Geng2012, Chen2012, Xu2012, Shivaswamy2011}
+\item The study reported evaluation results in a different metric than the metrics chosen for this meta-analysis \cite{Yu2009, Thuy2009, Pahikkala2009, Kersting2009, Mohan2011}
+\item The study reported a higher performance on baseline methods than official benchmark runs \cite{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}
+\item The study did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
+\end{enumerate}
+
+\begin{table*}
+\begin{tabular}{lll|lll}\toprule
+Method & Described in & Evaluated in & Method & Described in & Evaluated in\\
+\midrule
+AdaRank-MAP & \cite{Xu2007} & L2, L3, L4 & Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} \\ 
+AdaRank-NDCG & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & ListMLE & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
+ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet & \cite{Cao2007} & L2, L3, L4 \\ 
+ApproxAP & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
+ApproxNDCG & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
+BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
+Best Single Feature &  & \cite{Gomes2013} & MHR & \cite{Qin2007} & L2 \\ 
+BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
+BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009, Volkovs2013} & NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
+BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c, Volkovs2013} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
+BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-MAP & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
+C-CRF & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
+CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & Q.D.KNN & \cite{Geng2008} & \cite{Wang2013} \\ 
+CCRank & \cite{Wang2011c} & \cite{Wang2011c} & RandomForest &  & \cite{Gomes2013} \\ 
+CoList & \cite{Gao2014} & \cite{Gao2014} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
+Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAggNDCG & \cite{Wang2013} & \cite{Wang2013} \\ 
+DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010} \\ 
+DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
+EnergyNDCG & \cite{Freno2011} & \cite{Freno2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\ 
+FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b, Laporte2013} & RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FocusedBoost & \cite{Niu2012} & \cite{Niu2012} & RankMGP & \cite{Lin2012} & \cite{Lin2012} \\ 
+FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
+FocusedSVM & \cite{Niu2012} & \cite{Niu2012} & RankRLS & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
+FP-Rank & \cite{Song2013} & \cite{Song2013} & RankSVM & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010} \\ 
+FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & RankSVM-Primal &  & L3, \cite{Lai2011} \\ 
+FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & RankSVM-Struct &  & L3, L4 \\
+FSM$^{SVM}$ & \cite{Lai2013c} & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
+GAS-E & \cite{Geng2007} & \cite{Lai2013c} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\
+GP & \cite{DeAlmeida2007} & \cite{Alcantara2010} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\  
+GPRank & \cite{Silva2009} & \cite{Torkestani2012} & Ridge Regression & \cite{Cossock2006} & L3 \\
+GRankRLS & \cite{Pahikkala2010} & \cite{Pahikkala2010} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
+GroupCE & \cite{Lin2011} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
+GroupMLE & \cite{Lin2010} & \cite{Lin2011} & SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010} \\
+IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
+IPRank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\
+KeepRank & \cite{Chen2009} & \cite{Chen2009} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
+Kernel-PCA RankBoost & \cite{Duh2008} & \cite{Duh2008, Sato2013} & SVD-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\
+KL-CRF & \cite{Volkovs2011} & \cite{Volkovs2011} & SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
+LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
+LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} & TGRank & \cite{Lai2013} & \cite{Lai2013} \\ 
+LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
+LambdaRank & \cite{Burges2006} &  & VFLR & \cite{Cai2012} & \cite{Cai2012} \\ 
+LARF & \cite{Torkestani2012} & \cite{Torkestani2012} &  &  &  \\
+\bottomrule
+\end{tabular}
+\caption{Learning to rank algorithms with measurements on benchmark datasets}
+\label{tab:ltr_methods_used}
+\end{table*}
+
+\section{A Methodology for Comparing Learning to Rank Methods Cross-Benchmark}
 The LETOR 3.0 paper \cite{Qin2010} states that it may differ between datasets what the most accurate ranking methods are.
@@
 
-\section{Results \& Discussion}
-\subsection{Collection of Evaluation Results}
-Table \ref{tab:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
-Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, LETOR 3.0 and LETOR 4.0 baselines respectively.\\
-
-Some studies with evaluation results found through in literature review were not usable for the meta-analysis.
-The following enumeration enumerates those properties that made one or more studies unusable for the meta-analysis.
-Between brackets are the studies that these properties apply to.
-
-\begin{enumerate}
-\item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011, Lin2012}
-\item The study focussed on a different learning to rank task (e.g.
-rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}
-\item The study used an altered version of a benchmark that contained additional features \cite{Bidoki2009, Ding2010}
-\item The study provides no exact data of the evaluation results (e.g.
-results are only in graphical form) \cite{Wang2008, Wang2010, Xu2010, Kuo2009, Li2008, Xia2008, Zhou2011, Wu2011, Zhu2009, Karimzadehgan2011, Swersky2012, Pan2011, Ni2008, Ciaramita2008, Stewart2012, Petterson2009, Agarwal2010, Chang2009, Qin2008c, Adams2011, Sculley2009, Huang2008, Alejo2010, Sun2011, He2010b, Benbouzid2012, Geng2012, Chen2012, Xu2012, Shivaswamy2011}
-\item The study reported evaluation results in a different metric than the metrics chosen for this meta-analysis \cite{Yu2009, Thuy2009, Pahikkala2009, Kersting2009, Mohan2011}
-\item The study reported a higher performance on baseline methods than official benchmark runs \cite{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}
-\item The study did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
-\end{enumerate}
-
-\begin{table*}
-\begin{tabular}{|l|l|l||l|l|l|}
-Method & Described & Evaluated & Method & Described & Evaluated \\
-\hline 
-AdaRank-MAP & \cite{Xu2007} & L2, L3, L4 & Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} \\ 
-AdaRank-NDCG & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & ListMLE & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
-ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet & \cite{Cao2007} & L2, L3, L4 \\ 
-ApproxAP & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
-ApproxNDCG & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
-BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
-Best Single Feature &  & \cite{Gomes2013} & MHR & \cite{Qin2007} & L2 \\ 
-BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
-BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009, Volkovs2013} & NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
-BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c, Volkovs2013} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
-BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-MAP & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
-C-CRF & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
-CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & Q.D.KNN & \cite{Geng2008} & \cite{Wang2013} \\ 
-CCRank & \cite{Wang2011c} & \cite{Wang2011c} & RandomForest &  & \cite{Gomes2013} \\ 
-CoList & \cite{Gao2014} & \cite{Gao2014} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
-Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAggNDCG & \cite{Wang2013} & \cite{Wang2013} \\ 
-DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010} \\ 
-DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
-EnergyNDCG & \cite{Freno2011} & \cite{Freno2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\ 
-FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
-FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b, Laporte2013} & RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
-FocusedBoost & \cite{Niu2012} & \cite{Niu2012} & RankMGP & \cite{Lin2012} & \cite{Lin2012} \\ 
-FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
-FocusedSVM & \cite{Niu2012} & \cite{Niu2012} & RankRLS & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
-FP-Rank & \cite{Song2013} & \cite{Song2013} & RankSVM & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010} \\ 
-FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & RankSVM-Primal &  & L3, \cite{Lai2011} \\ 
-FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & RankSVM-Struct &  & L3, L4 \\
-FSM$^{SVM}$ & \cite{Lai2013c} & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
-GAS-E & \cite{Geng2007} & \cite{Lai2013c} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\
-GP & \cite{DeAlmeida2007} & \cite{Alcantara2010} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\  
-GPRank & \cite{Silva2009} & \cite{Torkestani2012} & Ridge Regression & \cite{Cossock2006} & L3 \\
-GRankRLS & \cite{Pahikkala2010} & \cite{Pahikkala2010} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
-GroupCE & \cite{Lin2011} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
-GroupMLE & \cite{Lin2010} & \cite{Lin2011} & SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010} \\
-IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
-IPRank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\
-KeepRank & \cite{Chen2009} & \cite{Chen2009} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
-Kernel-PCA RankBoost & \cite{Duh2008} & \cite{Duh2008, Sato2013} & SVD-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\
-KL-CRF & \cite{Volkovs2011} & \cite{Volkovs2011} & SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
-LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
-LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} & TGRank & \cite{Lai2013} & \cite{Lai2013} \\ 
-LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
-LambdaRank & \cite{Burges2006} &  & VFLR & \cite{Cai2012} & \cite{Cai2012} \\ 
-LARF & \cite{Torkestani2012} & \cite{Torkestani2012} &  &  &  \\ 
-\end{tabular}
-\caption{Learning to rank algorithms with measurements on benchmark datasets}
-\label{tab:ltr_methods_used}
-\end{table*}
-
-\subsection{Comparison of Learning to Rank models}
+\section{Results of Learning to Rank Comparison}
 The following subsections provide the performance of learning to rank methods in terms of Normalised Winning Number for NDCG@3, NDCG@5, NDCG@10 and MAP.
@@
 
-\subsubsection{NDCG@3}
+\subsection{NDCG@3}
 Figure \ref{fig:normalised_winning_number_ndcg3} shows the performance of learning to rank methods for the NDCG@3 metric.
@@
 
-\begin{figure*}[!htp]
+\begin{figure}[!h]
 \includegraphics[scale=0.285]{gfx/ndcg3_winnum}
@@
 \label{fig:normalised_winning_number_ndcg3}
-\end{figure*}
+\end{figure}
 
@@
 
-\subsubsection{NDCG@5}
+\subsection{NDCG@5}
 Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of learning to rank methods for the NDCG@5 metric.
@@
 
-\begin{figure*}[!htp]
+\begin{figure}[!h]
 \includegraphics[scale=0.285]{gfx/ndcg5_winnum}
@@
 \label{fig:normalised_winning_number_ndcg5}
-\end{figure*}
+\end{figure}
 
@@
 
-\subsubsection{NDCG@10}
+\subsection{NDCG@10}
 Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of learning to rank methods for the NDCG@10 metric.
@@
 
-\begin{figure*}
+\begin{figure}[!h]
 \includegraphics[scale=0.285]{gfx/ndcg10_winnum}
@@
 \label{fig:normalised_winning_number_ndcg10}
-\end{figure*}
+\end{figure}
 
@@
 
-\subsubsection{MAP}
+\subsection{MAP}
 Figure \ref{fig:normalised_winning_number_map} shows the performance of learning to rank methods for the MAP metric.
@@
 
-\begin{figure*}
+\begin{figure}[!h]
 \includegraphics[scale=0.285]{gfx/map_winnum}
@@
 \label{fig:normalised_winning_number_map}
-\end{figure*}
+\end{figure}
 
@@
 
-\subsubsection{Cross-metric}
+\subsection{Cross-Metric}
 Figure \ref{fig:normalised_winning_number_all} shows the Normalised Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}.
@@
 \begin{longtable}[!h]{@{}lrrrrrrrrrrrrrrrr@{}}\toprule
-& \multicolumn{2}{p{0.5cm}}{NDCG@3} & \phantom{a} 
+& \multicolumn{2}{c}{NDCG@3} & \phantom{a} 
 & \multicolumn{2}{c}{NDCG@5} & \phantom{a} 

### 2014-05-19 cikm_paper/paper.tex
eindelijk resultaat na geruzie met LaTeX

@@
 \usepackage{booktabs}
+\usepackage{dblfloatfix}
+\usepackage{fixltx2e}
+\renewcommand{\dblfloatpagefraction}{2.0}
+\makeatletter
+\makeatother
 \begin{document}
+
 %
@@
 
-\subsection{Literature selection}
+\subsection{Literature Selection}
 Table \ref{tab:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
@@
 
-\begin{table*}
+\begin{table*}[p]
 \begin{tabular}{lll|lll}\toprule
@@
 
+\begin{figure*}
+\includegraphics[scale=0.37]{gfx/ndcg3_winnum}
+\caption{NDCG@3 comparison of learning to rank methods}
+\label{fig:normalised_winning_number_ndcg3}
+\end{figure*}
 In contrast to the winning number comparison on LETOR 3.0, there will not be accuracy measurements for each algorithm on each dataset in our meta-analysis.
@@
 This Normalised Winning Number takes only those datasets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all datasets on which it has been evaluated.
-We will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as \\
+We will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as
 
@@
 \]
-
 From now on this adjusted version of Winning Number will be references to as \emph{Normalised Winning Number}. The mathematical definition of Normalised Winning Number is\\
@@
 
+\begin{figure*}
+\includegraphics[scale=0.37]{gfx/ndcg5_winnum}
+\caption{NDCG@5 comparison of learning to rank methods}
+\label{fig:normalised_winning_number_ndcg5}
+\end{figure*}
 \subsection{NDCG@3}
@@
 Table \ref{tab:raw_data_norm_winnum_ndcg3} in Appendix \ref{app:norm_winnum_ndcg3} provides the raw Normalised Winning Number data for the learning to rank methods for which NDCG@3 evaluation results were available.\\
-
-\begin{figure}[!h]
-\includegraphics[scale=0.285]{gfx/ndcg3_winnum}
-\caption{NDCG@3 comparison of learning to rank methods}
-\label{fig:normalised_winning_number_ndcg3}
-\end{figure}
 
@@
 
-\begin{figure}[!h]
-\includegraphics[scale=0.285]{gfx/ndcg5_winnum}
-\caption{NDCG@5 comparison of learning to rank methods}
-\label{fig:normalised_winning_number_ndcg5}
-\end{figure}
-
 LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a Normalised Winning Number of 1.0.
@@
 \subsection{NDCG@10}
+\begin{figure*}
+\includegraphics[scale=0.37]{gfx/ndcg10_winnum}
+\caption{NDCG@10 comparison of learning to rank methods}
+\label{fig:normalised_winning_number_ndcg10}
+\end{figure*}
+
 Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of learning to rank methods for the NDCG@10 metric.
 Table \ref{tab:raw_data_norm_winnum_ndcg10} in Appendix \ref{app:norm_winnum_ndcg10} provides the raw Normalised Winning Number data for the learning to rank methods.\\
-
-\begin{figure}[!h]
-\includegraphics[scale=0.285]{gfx/ndcg10_winnum}
-\caption{NDCG@10 comparison of learning to rank methods}
-\label{fig:normalised_winning_number_ndcg10}
-\end{figure}
 
@@
 
-\begin{figure}[!h]
-\includegraphics[scale=0.285]{gfx/map_winnum}
+\begin{figure*}
+\includegraphics[scale=0.37]{gfx/map_winnum}
 \caption{MAP comparison of learning to rank methods}
 \label{fig:normalised_winning_number_map}
-\end{figure}
+\end{figure*}
 
@@
 \begin{figure*}
-\includegraphics[scale=0.33]{gfx/combined_normalized_winnum}
+\includegraphics[scale=0.4235]{gfx/combined_normalized_winnum}
 \caption{Cross-benchmark comparison of learning to rank methods}

### 2014-05-19 cikm_paper/paper.tex
abstract gemaakt + tabelverwijzingen gefixt

@@
 \renewcommand{\dblfloatpagefraction}{2.0}
-\makeatletter
-\makeatother
 \begin{document}
@@
 \begin{abstract}
-
+Learning to rank is an increasingly important scientific field that comprises the use of machine learning for the ranking task.
+Learning to rank methods are generally evaluated on benchmark test collections.
+However, comparison of learning to rank methods based on evaluation results is hindered by a non-existence of a \emph{de facto} standard set of evaluation benchmark collections.
+This paper proposes a way to compare learning to rank methods based on sparse evaluation results data on a set of benchmark datasets.
+The comparison methodology comprises of two components: 1) Normalized Winning Number, which give insight on the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.
+Evaluation results on the LETOR 2.0, LETOR 3.0, Yahoo!
+Learning to Rank Challenge, Yandex Internet Mathematics 2009 contest, MSLR-web10/30k, WCL2R and AOL datasets are collected through a structured literature search.
+ListNet, SmoothRank, FenchelRank, FSMRank, LRUF and LARF turned out to be the best performing learning to rank methods in increasing order of Normalized Winning Number and decreasing order of Ideal Winning Number.
 \end{abstract}
@@
 One can find useful applications of ranking in many application domains outside web search as well.
-For example, ranking plays a vital role in automatic text summarisation, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}.
+For example, ranking plays a vital role in automatic text summarization, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}.
 In addition, the ranking task has been argued to be a better fit for recommender systems compared to the regression task on a continuous scale \cite{Adomavicius2005,McNee2006}.\\
@@
 Gomes et al.
-\cite{Gomes2013} analysed ranking accuracy of a set of models on both LETOR 3.0 and LETOR 4.0.
+\cite{Gomes2013} analyzed ranking accuracy of a set of models on both LETOR 3.0 and LETOR 4.0.
 Busa-Fekete et al.
@@
 \begin{table}[!h]
-\begin{tabular}{p{3.2cm}lp{1.8cm}}\toprule
+\begin{tabular}{p{2.8cm}lp{1.8cm}}\toprule
 Benchmark & Paper & \# of forward references \\
@@
 To collect measurements of learning to rank methods evaluated on these benchmarks, a Google Scholar search is performed on the name of the benchmark.
-Table~\ref{chap:benchmark_results} shows the results of this literature search.
+Table~\ref{tbl:ltr_benchmark_searches} shows the results of this literature search.
 
 \begin{table}[!h]
-\begin{tabular}{p{4.2cm}p{2.5cm}}\toprule
+\begin{tabular}{p{5.2cm}p{2.5cm}}\toprule
 Benchmark & Google scholar search results \\
@@
 \caption{Google scholar search results for learning to rank benchmarks}
-\label{tbl:ltr_benchmark_searchres}
+\label{tbl:ltr_benchmark_searches}
 \end{table}
-
-\subsection{Literature Selection}
-Table \ref{tab:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
-Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, LETOR 3.0 and LETOR 4.0 baselines respectively.\\
-
-Some studies with evaluation results found through in literature review were not usable for the meta-analysis.
-The following enumeration enumerates those properties that made one or more studies unusable for the meta-analysis.
-Between brackets are the studies that these properties apply to.
-
-\begin{enumerate}
-\item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011, Lin2012}
-\item The study focussed on a different learning to rank task (e.g.
-rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}
-\item The study used an altered version of a benchmark that contained additional features \cite{Bidoki2009, Ding2010}
-\item The study provides no exact data of the evaluation results (e.g.
-results are only in graphical form) \cite{Wang2008, Wang2010, Xu2010, Kuo2009, Li2008, Xia2008, Zhou2011, Wu2011, Zhu2009, Karimzadehgan2011, Swersky2012, Pan2011, Ni2008, Ciaramita2008, Stewart2012, Petterson2009, Agarwal2010, Chang2009, Qin2008c, Adams2011, Sculley2009, Huang2008, Alejo2010, Sun2011, He2010b, Benbouzid2012, Geng2012, Chen2012, Xu2012, Shivaswamy2011}
-\item The study reported evaluation results in a different metric than the metrics chosen for this meta-analysis \cite{Yu2009, Thuy2009, Pahikkala2009, Kersting2009, Mohan2011}
-\item The study reported a higher performance on baseline methods than official benchmark runs \cite{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}
-\item The study did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
-\end{enumerate}
-
 \begin{table*}[p]
@@
 
+\subsection{Literature Selection}
+Table \ref{tab:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
+Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, LETOR 3.0 and LETOR 4.0 baselines respectively.\\
+
+Some studies with evaluation results found through in literature review were not usable for the meta-analysis.
+The following enumeration enumerates those properties that made one or more studies unusable for the meta-analysis.
+Between brackets are the studies that these properties apply to.
+
+\begin{enumerate}
+\item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011, Lin2012}
+\item The study focuses on a different learning to rank task (e.g.
+rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}
+\item The study used an altered version of a benchmark that contained additional features \cite{Bidoki2009, Ding2010}
+\item The study provides no exact data of the evaluation results (e.g.
+results are only in graphical form) \cite{Wang2008, Wang2010, Xu2010, Kuo2009, Li2008, Xia2008, Zhou2011, Wu2011, Zhu2009, Karimzadehgan2011, Swersky2012, Pan2011, Ni2008, Ciaramita2008, Stewart2012, Petterson2009, Agarwal2010, Chang2009, Qin2008c, Adams2011, Sculley2009, Huang2008, Alejo2010, Sun2011, He2010b, Benbouzid2012, Geng2012, Chen2012, Xu2012, Shivaswamy2011}
+\item The study reported evaluation results in a different metric than the metrics chosen for this meta-analysis \cite{Yu2009, Thuy2009, Pahikkala2009, Kersting2009, Mohan2011}
+\item The study reported a higher performance on baseline methods than official benchmark runs \cite{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}
+\item The study did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
+\end{enumerate}
+
 \section{A Methodology for Comparing Learning to Rank Methods Cross-Benchmark}
@@
 \begin{figure*}
-\includegraphics[scale=0.37]{gfx/ndcg3_winnum}
+\includegraphics[scale=0.371]{gfx/ndcg3_winnum}
 \caption{NDCG@3 comparison of learning to rank methods}
@@
 In contrast to the winning number comparison on LETOR 3.0, there will not be accuracy measurements for each algorithm on each dataset in our meta-analysis.
-To compare algorithms based on a sparse set of evaluation measurements, a normalised version of the Winning Number metric will be used.
-This Normalised Winning Number takes only those datasets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all datasets on which it has been evaluated.
+To compare algorithms based on a sparse set of evaluation measurements, a normalized version of the Winning Number metric will be used.
+This Normalized Winning Number takes only those datasets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all datasets on which it has been evaluated.
 We will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as
@@
 \section{Results of Learning to Rank Comparison}
-The following subsections provide the performance of learning to rank methods in terms of Normalised Winning Number for NDCG@3, NDCG@5, NDCG@10 and MAP.
-Performance of the learning to rank methods is plotted with Normalised Winning Number on the vertical axis and the number of datasets on which the method has been evaluated on the horizontal axis.
+The following subsections provide the performance of learning to rank methods in terms of Normalized Winning Number for NDCG@3, NDCG@5, NDCG@10 and MAP.
+Performance of the learning to rank methods is plotted with Normalized Winning Number on the vertical axis and the number of datasets on which the method has been evaluated on the horizontal axis.
 The further to the right, the more certain we can be about the performance of the learning to rank method.
-The methods for which it holds that there is no other method that has 1) a higher Normalised Winning Number and 2) a higher number datasets evaluated, are identified as the best performing methods and are labelled with the name of the method.
+The methods for which it holds that there is no other method that has 1) a higher Normalized Winning Number and 2) a higher number datasets evaluated, are identified as the best performing methods and are labeled with the name of the method.
+Table \ref{tab:raw_data} in Appendix \ref{app:raw_data} provides the raw Normalized Winning Number data for the learning to rank methods at NDCG@3, @5, @10, MAP and their cross-metric weighted average.
 
 \begin{figure*}
-\includegraphics[scale=0.37]{gfx/ndcg5_winnum}
+\includegraphics[scale=0.371]{gfx/ndcg5_winnum}
 \caption{NDCG@5 comparison of learning to rank methods}
@@
 Figure \ref{fig:normalised_winning_number_ndcg3} shows the performance of learning to rank methods for the NDCG@3 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg3} in Appendix \ref{app:norm_winnum_ndcg3} provides the raw Normalised Winning Number data for the learning to rank methods for which NDCG@3 evaluation results were available.\\
-
-LambdaNeuralRank and CoList both acquired a perfect Normalised Winning Number score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
+LambdaNeuralRank and CoList both acquired a perfect Normalized Winning Number score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
 LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three datasets which in addition are more relevant datasets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 datasets (see Table \ref{tab:ltr_methods_used}).
-FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in Normalised Winning Number, but increasingly higher in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
+FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in Normalized Winning Number, but increasingly higher in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
 
@@
 Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of learning to rank methods for the NDCG@5 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg5} in Appendix \ref{app:norm_winnum_ndcg5} provides the raw Normalised Winning Number data for the learning to rank methods.\\
-
-LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a Normalised Winning Number of 1.0.
-LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of datasets, but score decreasingly well in terms of Normalised Winning Number.
+LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a Normalized Winning Number of 1.0.
+LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of datasets, but score decreasingly well in terms of Normalized Winning Number.
 These results are highly in agreement with the NDCG@3 comparison.
@@
 \begin{figure*}
-\includegraphics[scale=0.37]{gfx/ndcg10_winnum}
+\includegraphics[scale=0.371]{gfx/ndcg10_winnum}
 \caption{NDCG@10 comparison of learning to rank methods}
@@
 Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of learning to rank methods for the NDCG@10 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg10} in Appendix \ref{app:norm_winnum_ndcg10} provides the raw Normalised Winning Number data for the learning to rank methods.\\
-
-LambdaMART and LambdaNeuralRank score a Normalised Winning Number of 1.0 on the NDCG@10 comparison.
+LambdaMART and LambdaNeuralRank score a Normalized Winning Number of 1.0 on the NDCG@10 comparison.
 For LambdaNeuralRank these results are again based on AOL dataset measurements.
 LambdaMART showed the highest NDCG@10 performance for the MSLR-WEB10k dataset.
-The set of algorithms for which there is no other algorithm with both a higher Normalised Winning Number and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and NDCG@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
+The set of algorithms for which there is no other algorithm with both a higher Normalized Winning Number and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and NDCG@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
 SmoothRank and FSMRank were not present in this set for the NDCG@3 and NDCG@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_ndcg3} and \ref{tab:raw_data_norm_winnum_ndcg5} in the Appendices.
@@
 \subsection{MAP}
-Figure \ref{fig:normalised_winning_number_map} shows the performance of learning to rank methods for the MAP metric.
-Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw Normalised Winning Number data for the learning to rank methods.\\
-
 \begin{figure*}
-\includegraphics[scale=0.37]{gfx/map_winnum}
+\includegraphics[scale=0.371]{gfx/map_winnum}
 \caption{MAP comparison of learning to rank methods}
@@
 \end{figure*}
-
+Figure \ref{fig:normalised_winning_number_map} shows the performance of learning to rank methods for the MAP metric.
 Where comparisons on the NDCG metric at different cut-off points where highly in agreement in terms of the best performing algorithms, the comparison in terms of MAP shows different results.
-RankDE scores a Normalised Winning Number of 1.0 on one dataset, like LambdaNeuralRank did on for all NDCG-comparisons.
+RankDE scores a Normalized Winning Number of 1.0 on one dataset, like LambdaNeuralRank did on for all NDCG-comparisons.
 In contrast to LambdaNeuralRank, RankDE achieved this highest score on the LETOR 2.0 TD2003, a dataset on which many methods are evaluated.\\
 
-LARF and LRUF score very high Normalised Winning Number scores, but based on only relatively few datasets, just as in the NDCG comparisons.
+LARF and LRUF score very high Normalized Winning Number scores, but based on only relatively few datasets, just as in the NDCG comparisons.
 Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all NDCG comparisons, are not within the best performing methods in the MAP comparison.
-A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_ndcg3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_ndcg3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more datasets for MAP compared to NDCG, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalised Winning Number.
-SmoothRank performed a Normalised Winning Number of around 0.53 over 7 datasets, which is still good in both certainty and accuracy, but not among the top methods.
+A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_ndcg3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_ndcg3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more datasets for MAP compared to NDCG, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalized Winning Number.
+SmoothRank performed a Normalized Winning Number of around 0.53 over 7 datasets, which is still good in both certainty and accuracy, but not among the top methods.
 RE-QR is one of the best performers in the MAP comparison with a reasonable amount of benchmark evaluations.
 No reported NDCG performance was found in the literature study for RE-QR.
-There is a lot of certainty on the accuracy of RankBoost and RankSVM as both models are evaluated on the majority of datasets included in the comparison for the MAP metric, but given their Normalised Winning Number it can said that both methods are not within the top performing learning to rank methods.
+There is a lot of certainty on the accuracy of RankBoost and RankSVM as both models are evaluated on the majority of datasets included in the comparison for the MAP metric, but given their Normalized Winning Number it can said that both methods are not within the top performing learning to rank methods.
 
 \subsection{Cross-Metric}
-Figure \ref{fig:normalised_winning_number_all} shows the Normalised Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}.
-Table \ref{tab:raw_data_norm_winnum_all} in Appendix \ref{app:norm_winnum_all} provide the raw data plotted in Figure \ref{fig:normalised_winning_number_all}.\\
-
 \begin{figure*}
@@
 \end{figure*}
-
-The cross-metric comparison is based on the NDCG@3, NDCG@5, NDCG@10 and MAP comparisons combined, which justifies analysing the comparison more thoroughly.
+Figure \ref{fig:normalised_winning_number_all} shows the Normalized Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}.
+The cross-metric comparison is based on the NDCG@3, NDCG@5, NDCG@10 and MAP comparisons combined, which justifies analyzing the comparison more thoroughly.
 Figure \ref{fig:normalised_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
@@
 
-LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both Ideal Winning Number and Normalised Winning Number.
+LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both Ideal Winning Number and Normalized Winning Number.
 LRUF is the only method that achieved this in all NDCG comparisons, the MAP comparison as well as the cross-metric comparison.
@@
 
-C-CRF, DirectRank, FP-Rank, RankCSA, LambdaNeuralRank and VFLR all have near-perfect Normalised Winning Number measures, but have low Ideal Winning Number measures.
+C-CRF, DirectRank, FP-Rank, RankCSA, LambdaNeuralRank and VFLR all have near-perfect Normalized Winning Number measures, but have low Ideal Winning Number measures.
 Further evaluation runs of these methods on benchmark datasets that they are not yet evaluated on are desirable.
@@
 \section{Limitations}
-In the Normalised Winning Number calculation the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
+In the Normalized Winning Number calculation the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
 By calculating it in this way, we implicitly make the assumption that the learning to rank methods are (approximately) distributed uniformly over the benchmarks, such that the average learning to rank method tested are approximately equally hard for each data set.
@@
 Using evaluation results published by other researchers relies on the honesty of those researchers.
-Overly optimistic evaluation results published by other researchers do affect our Normalised Winning Number results.
+Overly optimistic evaluation results published by other researchers do affect our Normalized Winning Number results.
 Limiting the meta-analysis to those studies that report comparable results on one of the baseline methods of a benchmark set reduces this limitation but does not solve it completely.

### 2014-05-19 cikm_paper/paper.tex
conclusion + repeat caption

@@
 \section{Conclusions}
-We will now look back on our first research question using the knowledge gained with the cross-benchmark comparison.
-\begin{description}
-\item[RQ1] What are the best performing learning to rank algorithms in terms of accuracy on relevant benchmark datasets?\\
-\end{description}
 Although no closing arguments can be formulated on which learning to rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.\\
@@
 \cmidrule{2-3} \cmidrule{5-6} \cmidrule{8-9} \cmidrule{11-12} \cmidrule{14-16}
-Method & NWN & \#ds && NWN & \#ds && NWN & \#ds && NWN & \#ds && WM & IWN & NWN \\ \midrule
+Method & NWN & \#ds && NWN & \#ds && NWN & \#ds && NWN & \#ds && WM & IWN & NWN \\ \midrule\endhead
 AdaRank-MAP & 0.3486 & 12 && 0.3884 & 12 && 0.3648 & 13 && 0.3206 & 12 && 332 & 937 & 0.3543 \\

### 2014-05-19 cikm_paper/paper.tex
2e appendix

@@
 This paper proposes a way to compare learning to rank methods based on sparse evaluation results data on a set of benchmark datasets.
-The comparison methodology comprises of two components: 1) Normalized Winning Number, which give insight on the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.
+The comparison methodology comprises of two components: 1) Normalized Winning Number, which gives insight on the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.
 Evaluation results on the LETOR 2.0, LETOR 3.0, Yahoo!
@@
 \end{table}
-\begin{table*}[p]
-\begin{tabular}{lll|lll}\toprule
-Method & Described in & Evaluated in & Method & Described in & Evaluated in\\
-\midrule
-AdaRank-MAP & \cite{Xu2007} & L2, L3, L4 & Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} \\ 
-AdaRank-NDCG & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & ListMLE & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
-ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet & \cite{Cao2007} & L2, L3, L4 \\ 
-ApproxAP & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
-ApproxNDCG & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
-BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
-Best Single Feature &  & \cite{Gomes2013} & MHR & \cite{Qin2007} & L2 \\ 
-BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
-BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009, Volkovs2013} & NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
-BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c, Volkovs2013} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
-BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-MAP & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
-C-CRF & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
-CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & Q.D.KNN & \cite{Geng2008} & \cite{Wang2013} \\ 
-CCRank & \cite{Wang2011c} & \cite{Wang2011c} & RandomForest &  & \cite{Gomes2013} \\ 
-CoList & \cite{Gao2014} & \cite{Gao2014} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
-Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAggNDCG & \cite{Wang2013} & \cite{Wang2013} \\ 
-DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010} \\ 
-DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
-EnergyNDCG & \cite{Freno2011} & \cite{Freno2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\ 
-FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
-FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b, Laporte2013} & RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
-FocusedBoost & \cite{Niu2012} & \cite{Niu2012} & RankMGP & \cite{Lin2012} & \cite{Lin2012} \\ 
-FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
-FocusedSVM & \cite{Niu2012} & \cite{Niu2012} & RankRLS & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
-FP-Rank & \cite{Song2013} & \cite{Song2013} & RankSVM & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010} \\ 
-FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & RankSVM-Primal &  & L3, \cite{Lai2011} \\ 
-FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & RankSVM-Struct &  & L3, L4 \\
-FSM$^{SVM}$ & \cite{Lai2013c} & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
-GAS-E & \cite{Geng2007} & \cite{Lai2013c} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\
-GP & \cite{DeAlmeida2007} & \cite{Alcantara2010} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\  
-GPRank & \cite{Silva2009} & \cite{Torkestani2012} & Ridge Regression & \cite{Cossock2006} & L3 \\
-GRankRLS & \cite{Pahikkala2010} & \cite{Pahikkala2010} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
-GroupCE & \cite{Lin2011} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
-GroupMLE & \cite{Lin2010} & \cite{Lin2011} & SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010} \\
-IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
-IPRank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\
-KeepRank & \cite{Chen2009} & \cite{Chen2009} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
-Kernel-PCA RankBoost & \cite{Duh2008} & \cite{Duh2008, Sato2013} & SVD-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\
-KL-CRF & \cite{Volkovs2011} & \cite{Volkovs2011} & SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
-LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
-LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} & TGRank & \cite{Lai2013} & \cite{Lai2013} \\ 
-LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
-LambdaRank & \cite{Burges2006} &  & VFLR & \cite{Cai2012} & \cite{Cai2012} \\ 
-LARF & \cite{Torkestani2012} & \cite{Torkestani2012} &  &  &  \\
-\bottomrule
-\end{tabular}
-\caption{Learning to rank algorithms with measurements on benchmark datasets}
-\label{tab:ltr_methods_used}
-\end{table*}
 
 \subsection{Literature Selection}
-Table \ref{tab:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
+Table \ref{tab:ltr_methods_used} in Appendix \ref{app:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
 Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, LETOR 3.0 and LETOR 4.0 baselines respectively.\\
@@
 LambdaNeuralRank and CoList both acquired a perfect Normalized Winning Number score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
-LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three datasets which in addition are more relevant datasets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 datasets (see Table \ref{tab:ltr_methods_used}).
+LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three datasets which in addition are more relevant datasets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 datasets (see Table \ref{tab:ltr_methods_used} in Appendix \ref{app:ltr_methods_used}).
 FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in Normalized Winning Number, but increasingly higher in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
@@
 
-\section{Conclusions}
-Although no closing arguments can be formulated on which learning to rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.\\
+\section{Contributions}
+We proposed a way to compare learning to rank methods based on sparse evaluation results data on a set of benchmark datasets.
+The comparison methodology comprises of two components: 1) Normalized Winning Number, which provides insight in the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.\\
 
+Although no closing arguments can be formulated on which learning to rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.
 LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the learning to rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
@@
 \appendix
+\section{Learning to Rank Methods with Evaluation Results}
+\label{app:ltr_methods_used}
+\begin{table*}[!hp]
+\begin{tabular}{lll|lll}\toprule
+Method & Described in & Evaluated in & Method & Described in & Evaluated in\\
+\midrule
+AdaRank-MAP & \cite{Xu2007} & L2, L3, L4 & Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} \\ 
+AdaRank-NDCG & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & ListMLE & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
+ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet & \cite{Cao2007} & L2, L3, L4 \\ 
+ApproxAP & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
+ApproxNDCG & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
+BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
+Best Single Feature &  & \cite{Gomes2013} & MHR & \cite{Qin2007} & L2 \\ 
+BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
+BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009, Volkovs2013} & NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
+BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c, Volkovs2013} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
+BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-MAP & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
+C-CRF & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
+CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & Q.D.KNN & \cite{Geng2008} & \cite{Wang2013} \\ 
+CCRank & \cite{Wang2011c} & \cite{Wang2011c} & RandomForest &  & \cite{Gomes2013} \\ 
+CoList & \cite{Gao2014} & \cite{Gao2014} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
+Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAggNDCG & \cite{Wang2013} & \cite{Wang2013} \\ 
+DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010} \\ 
+DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
+EnergyNDCG & \cite{Freno2011} & \cite{Freno2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\ 
+FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b, Laporte2013} & RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FocusedBoost & \cite{Niu2012} & \cite{Niu2012} & RankMGP & \cite{Lin2012} & \cite{Lin2012} \\ 
+FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
+FocusedSVM & \cite{Niu2012} & \cite{Niu2012} & RankRLS & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
+FP-Rank & \cite{Song2013} & \cite{Song2013} & RankSVM & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010} \\ 
+FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & RankSVM-Primal &  & L3, \cite{Lai2011} \\ 
+FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & RankSVM-Struct &  & L3, L4 \\
+FSM$^{SVM}$ & \cite{Lai2013c} & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
+GAS-E & \cite{Geng2007} & \cite{Lai2013c} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\
+GP & \cite{DeAlmeida2007} & \cite{Alcantara2010} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\  
+GPRank & \cite{Silva2009} & \cite{Torkestani2012} & Ridge Regression & \cite{Cossock2006} & L3 \\
+GRankRLS & \cite{Pahikkala2010} & \cite{Pahikkala2010} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
+GroupCE & \cite{Lin2011} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
+GroupMLE & \cite{Lin2010} & \cite{Lin2011} & SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010} \\
+IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
+IPRank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\
+KeepRank & \cite{Chen2009} & \cite{Chen2009} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
+Kernel-PCA RankBoost & \cite{Duh2008} & \cite{Duh2008, Sato2013} & SVD-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\
+KL-CRF & \cite{Volkovs2011} & \cite{Volkovs2011} & SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
+LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
+LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} & TGRank & \cite{Lai2013} & \cite{Lai2013} \\ 
+LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
+LambdaRank & \cite{Burges2006} &  & VFLR & \cite{Cai2012} & \cite{Cai2012} \\ 
+LARF & \cite{Torkestani2012} & \cite{Torkestani2012} &  &  &  \\
+\bottomrule
+\end{tabular}
+\caption{Learning to rank algorithms with measurements on benchmark datasets}
+\label{tab:ltr_methods_used}
+\end{table*}
+
+\clearpage
 \section{Cross-Comparison Raw Data}
 \label{app:raw_data}
-
-\begin{longtable}[!h]{@{}lrrrrrrrrrrrrrrrr@{}}\toprule
+\begin{longtable}[!hp]{@{}lrrrrrrrrrrrrrrrr@{}}\toprule
 & \multicolumn{2}{c}{NDCG@3} & \phantom{a} 

### 2014-05-19 cikm_paper/paper.tex
draft versie paper

@@
 
+\begin{figure*}
+\includegraphics[scale=0.371]{gfx/ndcg3_winnum}
+\caption{NDCG@3 comparison of learning to rank methods}
+\label{fig:normalised_winning_number_ndcg3}
+\end{figure*}
 \section{A Methodology for Comparing Learning to Rank Methods Cross-Benchmark}
@@
 
-\begin{figure*}
-\includegraphics[scale=0.371]{gfx/ndcg3_winnum}
-\caption{NDCG@3 comparison of learning to rank methods}
-\label{fig:normalised_winning_number_ndcg3}
-\end{figure*}
 In contrast to the winning number comparison on LETOR 3.0, there will not be accuracy measurements for each algorithm on each dataset in our meta-analysis.
@@
 
+\begin{figure*}
+\includegraphics[scale=0.371]{gfx/ndcg5_winnum}
+\caption{NDCG@5 comparison of learning to rank methods}
+\label{fig:normalised_winning_number_ndcg5}
+\end{figure*}
 \section{Results of Learning to Rank Comparison}
@@
 
-\begin{figure*}
-\includegraphics[scale=0.371]{gfx/ndcg5_winnum}
-\caption{NDCG@5 comparison of learning to rank methods}
-\label{fig:normalised_winning_number_ndcg5}
-\end{figure*}
 \subsection{NDCG@3}
@@
 % produce the bibliography for the citations in your paper.
+\newpage
 \bibliographystyle{abbrv}

### 2014-05-19 cikm_paper/paper.tex
mooie rewrite :)

@@
 \renewcommand{\dblfloatpagefraction}{2.0}
+\usepackage{array}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 \begin{document}
@@
 Learning to rank is an increasingly important scientific field that comprises the use of machine learning for the ranking task.
-Learning to rank methods are generally evaluated on benchmark test collections.
-However, comparison of learning to rank methods based on evaluation results is hindered by a non-existence of a \emph{de facto} standard set of evaluation benchmark collections.
-This paper proposes a way to compare learning to rank methods based on sparse evaluation results data on a set of benchmark datasets.
-The comparison methodology comprises of two components: 1) Normalized Winning Number, which gives insight on the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.
-Evaluation results on the LETOR 2.0, LETOR 3.0, Yahoo!
-Learning to Rank Challenge, Yandex Internet Mathematics 2009 contest, MSLR-web10/30k, WCL2R and AOL datasets are collected through a structured literature search.
+New learning to rank methods are generally evaluated on benchmark test collections.
+However, comparison of learning to rank methods based on evaluation results is hindered by non-existence of a \emph{de facto} standard set of evaluation benchmark collections.
+In this paper we propose a way to compare learning to rank methods based on a sparse set of evaluation results on a set of benchmark datasets.
+Our comparison methodology consists of two components: 1) Normalized Winning Number, which gives insight in the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning its ranking accuracy.
+Evaluation results on a set of well known benchmark collections are collected through a structured literature search.
 ListNet, SmoothRank, FenchelRank, FSMRank, LRUF and LARF turned out to be the best performing learning to rank methods in increasing order of Normalized Winning Number and decreasing order of Ideal Winning Number.
@@
 
-\keywords{Learning to rank, Information retrieval, Benchmark datasets}
+\keywords{Learning to rank, Information retrieval, Evaluation metric}
 
@@
 Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
-The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective ranking models.
+increased amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective ranking models.
 Learning to rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\
 
-In recent years, several learning to rank benchmark datasets have been proposed with the aim of enabling comparison of the performance of different learning to rank methods.
+In recent years, several learning to rank benchmark datasets have been proposed with the aim of enabling comparison of different learning to rank methods in terms of ranking accuracy.
 Well-known benchmark datasets in the learning to rank field include the \emph{Yahoo!
-Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the \emph{Yandex Internet Mathematics 2009} contest \footnote{http://imat2009.yandex.ru/en}, the LETOR dataset \cite{Qin2010}, and the MSLR (Microsoft Learning to Rank) datasets \footnote{http://research.microsoft.com/en-us/projects/mslr/}.\\
-
-No agreement exists between authors in the learning to rank field on which benchmark set(s) to use for model evaluation.
-Comparing ranking performance of learning to rank methods is largely hindered by this lack of a \emph{de facto} standard way of benchmarking new models.\\
+Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the \emph{Yandex Internet Mathematics 2009} contest\footnote{http://imat2009.yandex.ru/en}, the LETOR dataset \cite{Qin2010}, and the MSLR (Microsoft Learning to Rank) datasets\footnote{http://research.microsoft.com/en-us/projects/mslr/}.
+There exists no agreement among authors in the learning to rank field on which benchmark collection(s) to use to evaluate a new model.
+Comparing ranking accuracy of learning to rank methods is largely hindered by this lack of a \emph{de facto} standard way of benchmarking new models.\\
 
 Gomes et al.
-\cite{Gomes2013} analyzed ranking accuracy of a set of models on both LETOR 3.0 and LETOR 4.0.
+\cite{Gomes2013} analyzed the ranking accuracy of a set of models on both LETOR 3.0 and 4.0.
 Busa-Fekete et al.
@@
 \label{sec:collecting_evaluation_results}
-Through a literature review we will collect evaluation results on the datasets/collections.
+We collect evaluation results on the datasets of benchmark collections through a structured literature search.
 The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
@@
 
-For the LETOR collections, the evaluation results of the baseline models will be used from LETOR 2.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor2.0/baseline.aspx}, LETOR 3.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and LETOR 4.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} as listed on the LETOR website.\\
+For the LETOR collections, the evaluation results of the baseline models will be used from LETOR 2.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor2.0/baseline.aspx}, 3.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and 4.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} as listed on the LETOR website.\\
 
-LETOR 1.0, LETOR 3.0, Yahoo!
+LETOR 1.0 and 3.0, Yahoo!
 Learning to Rank Challenge, WCL2R and AOL have accompanying papers which were published together with these benchmark collections.
-Users of those benchmark collections are encouraged to cite these papers.
+Authors publishing evaluation results on these benchmark collections are requested to cite these papers.
 Therefore, we collect evaluation measurements of learning to rank methods on these benchmark collections through forward literature search.
@@
 \begin{table}[!h]
-\begin{tabular}{p{2.8cm}lp{1.8cm}}\toprule
+\begin{tabular}{p{2.8cm}p{2.64cm}R{1.9cm}}\toprule
 Benchmark & Paper & \# of forward references \\
@@
 
-The LETOR 4.0, MSLR-web10/30k and Yandex Internet Mathematics Competition 2009 benchmark collections were not accompanied with a describing study.
-To collect measurements of learning to rank methods evaluated on these benchmarks, a Google Scholar search is performed on the name of the benchmark.
+The LETOR 4.0, MSLR-web10/30k and Yandex Internet Mathematics Competition 2009 benchmark collections are not accompanied by a paper.
+To collect evaluation results for learning to rank methods on these benchmarks, a Google Scholar search is performed on the name of the benchmark.
 Table~\ref{tbl:ltr_benchmark_searches} shows the results of this literature search.
@@
 \begin{table}[!h]
-\begin{tabular}{p{5.2cm}p{2.5cm}}\toprule
-Benchmark & Google scholar search results \\
+\begin{tabular}{p{5.2cm}R{2.5cm}}\toprule
+Search Query & Google scholar search results \\
 \midrule
-LETOR 4.0 & 75 results \\
-MSLR-web10k & 16 results \\
-MSLR-web30k & 15 results \\
-Yandex Internet Mathematics & 3 results \\
+"LETOR 4.0" & 75 \\
+"MSLR-web10k" & 16 \\
+"MSLR-web30k" & 15 \\
+"Yandex Internet Mathematics" & 3 \\
 \bottomrule
@@
 \subsection{Literature Selection}
-Table \ref{tab:ltr_methods_used} in Appendix \ref{app:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
-Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, LETOR 3.0 and LETOR 4.0 baselines respectively.\\
+Table \ref{tab:ltr_methods_used} in Appendix \ref{app:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found through described procedure.
+Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, 3.0 and 4.0 baselines respectively.\\
 
-Some studies with evaluation results found through in literature review were not usable for the meta-analysis.
+Some studies with evaluation results found through the literature search procedure were not usable for the meta-analysis.
 The following enumeration enumerates those properties that made one or more studies unusable for the meta-analysis.
@@
 \section{A Methodology for Comparing Learning to Rank Methods Cross-Benchmark}
-The LETOR 3.0 paper \cite{Qin2010} states that it may differ between datasets what the most accurate ranking methods are.
-To evaluate the overall performance of learning to rank methods over the multiple datasets in the LETOR 3.0 collections, Qin et al.
-\cite{Qin2010} proposed a measure called \emph{winning number} as the number of other algorithms that an algorithm can beat over the set of datasets.
-Formally the winning number measure is defined as\\
+Qin et al.
+\cite{Qin2010} state in the LETOR 3.0 paper that it may differ between datasets what the most accurate ranking methods are.
+They propose a measure they call \emph{Winning Number} to evaluate the overall performance of learning to rank methods over the multiple datasets in the LETOR 3.0 collection.
+Winning Number is defined as the number of other algorithms that an algorithm can beat over the set of datasets, or more formally\\
 
@@
 
-In contrast to the winning number comparison on LETOR 3.0, there will not be accuracy measurements for each algorithm on each dataset in our meta-analysis.
-To compare algorithms based on a sparse set of evaluation measurements, a normalized version of the Winning Number metric will be used.
+In contrast to the winning number comparison on LETOR 3.0, there will not be evaluation results for each algorithm on each dataset in our meta-analysis.
+We propose a normalized version of the Winning Number metric to compare algorithms based on a sparse set of evaluation measurements.
 This Normalized Winning Number takes only those datasets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all datasets on which it has been evaluated.
@@
 
-NDCG@3, NDCG@5, NDCG@10 and MAP are chosen as metrics on which the meta-analysis will be performed.
+NDCG@$\{3,5,10\}$ and MAP are chosen as metrics on which the meta-analysis will be performed.
 These metrics seem to be the most frequently used evaluation metrics in most of the used benchmark datasets.
@@
 \section{Results of Learning to Rank Comparison}
-The following subsections provide the performance of learning to rank methods in terms of Normalized Winning Number for NDCG@3, NDCG@5, NDCG@10 and MAP.
+The following subsections provide the performance of learning to rank methods in terms of Normalized Winning Number for NDCG@$\{3,5,10\}$ and MAP.
 Performance of the learning to rank methods is plotted with Normalized Winning Number on the vertical axis and the number of datasets on which the method has been evaluated on the horizontal axis.
@@
 The methods for which it holds that there is no other method that has 1) a higher Normalized Winning Number and 2) a higher number datasets evaluated, are identified as the best performing methods and are labeled with the name of the method.
-Table \ref{tab:raw_data} in Appendix \ref{app:raw_data} provides the raw Normalized Winning Number data for the learning to rank methods at NDCG@3, @5, @10, MAP and their cross-metric weighted average.
+Table \ref{tab:raw_data} in Appendix \ref{app:raw_data} provides the raw Normalized Winning Number data for the learning to rank methods at NDCG@$\{3,5,10\}$ and MAP and their cross-metric weighted average.
 
@@
 These results are highly in agreement with the NDCG@3 comparison.
-The only modification compared to the NDCG@3 comparison being that OWPC did show to be a method for which there were no methods performing better on both axes in the NDCG@5 comparison, but not in the NDCG@3 comparison.
+The only modification compared to the NDCG@3 comparison being that OWPC did show to be a method for which there were no methods performing better on both axes in the NDCG@5 comparison, but not in the @3 comparison.
 Like in the NDCG@3 comparison, SmoothRank and ListNet can be regarded as most reliable results because the evaluation measurements for these methods are based on LETOR official baselines.
@@
 LambdaMART showed the highest NDCG@10 performance for the MSLR-WEB10k dataset.
-The set of algorithms for which there is no other algorithm with both a higher Normalized Winning Number and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and NDCG@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
-SmoothRank and FSMRank were not present in this set for the NDCG@3 and NDCG@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_ndcg3} and \ref{tab:raw_data_norm_winnum_ndcg5} in the Appendices.
-DCMP is not in the set in contrast with the NDCG@3 and NDCG@5 comparison.
+The set of algorithms for which there is no other algorithm with both a higher Normalized Winning Number and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and @5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
+SmoothRank and FSMRank were not present in this set for the NDCG@3 and @5 comparison, but were close by, as can be seen in Table \ref{tab:raw_data} in the Appendix \ref{app:raw_data}.
+DCMP is not in the set in contrast with the NDCG@3 and @5 comparison.
 
@@
 Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all NDCG comparisons, are not within the best performing methods in the MAP comparison.
-A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_ndcg3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_ndcg3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more datasets for MAP compared to NDCG, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalized Winning Number.
+A deeper look in the raw data Table \ref{tab:raw_data} in the Appendix \ref{app:raw_data} shows that LAC-MR-OR is evaluated on many more datasets for MAP compared to NDCG, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalized Winning Number.
 SmoothRank performed a Normalized Winning Number of around 0.53 over 7 datasets, which is still good in both certainty and accuracy, but not among the top methods.
@@
 Figure \ref{fig:normalised_winning_number_all} shows the Normalized Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}.
-The cross-metric comparison is based on the NDCG@3, NDCG@5, NDCG@10 and MAP comparisons combined, which justifies analyzing the comparison more thoroughly.
+The cross-metric comparison is based on the NDCG@$\{3,5,10\}$ and MAP comparisons combined, which justifies analyzing the comparison more thoroughly.
 Figure \ref{fig:normalised_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
@@
 We proposed a way to compare learning to rank methods based on sparse evaluation results data on a set of benchmark datasets.
-The comparison methodology comprises of two components: 1) Normalized Winning Number, which provides insight in the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.\\
+Our comparison methodology comprises of two components: 1) Normalized Winning Number, which provides insight in the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.\\
 
-Although no closing arguments can be formulated on which learning to rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.
+Based on our structured search for evaluation results on well-known benchmarks collections, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.
+However, no closing arguments can be formulated on which learning to rank methods are most accurate.
 LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the learning to rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
-From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.
-For more definite conclusions on the relative performance of these five methods, more evaluation runs on are desirable for the methods on the left side on the list on benchmark datasets that these methods have not yet been evaluated on.\\
+From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.\\
+
+More evaluation runs on are needed for the methods on the left side on the list on benchmark datasets that these methods have not yet been evaluated on for more definite conclusions on which learning to rank methods are most accurate.
+Our work contributes in this sense by identifying promising learning to rank methods that researchers could focus on in performing additional evaluation runs.\\
 

### 2014-05-19 cikm_paper/paper.tex
punten Djoerd

@@
 Our comparison methodology consists of two components: 1) Normalized Winning Number, which gives insight in the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning its ranking accuracy.
-Evaluation results on a set of well known benchmark collections are collected through a structured literature search.
+Evaluation results on a set of well-known benchmark collections are collected through a structured literature search.
 ListNet, SmoothRank, FenchelRank, FSMRank, LRUF and LARF turned out to be the best performing learning to rank methods in increasing order of Normalized Winning Number and decreasing order of Ideal Winning Number.
@@
 Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
-increased amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective ranking models.
+Increased amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective ranking models.
 Learning to rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\
@@
 \caption{NDCG@3 comparison of learning to rank methods}
-\label{fig:normalised_winning_number_ndcg3}
+\label{fig:normalized_winning_number_ndcg3}
 \end{figure*}
@@
 \]
-From now on this adjusted version of Winning Number will be references to as \emph{Normalised Winning Number}. The mathematical definition of Normalised Winning Number is\\
+From now on this adjusted version of Winning Number will be references to as \emph{Normalized Winning Number}. The mathematical definition of Normalized Winning Number is\\
 
-$\text{Normalised Winning Number}_i(M) = \frac{\text{Winning Number}_i(M)}{\text{Ideal Winning Number}_i(M)}$\\
+$\text{Normalized Winning Number}_i(M) = \frac{\text{Winning Number}_i(M)}{\text{Ideal Winning Number}_i(M)}$\\
 
@@
 
-NDCG@$\{3,5,10\}$ and MAP are chosen as metrics on which the meta-analysis will be performed.
-These metrics seem to be the most frequently used evaluation metrics in most of the used benchmark datasets.
-An exception is the Yahoo!
-Learning to Rank Challenge datasets on which mainly ERR is used as main evaluation metric.
+NDCG@$\{3,5,10\}$ and MAP are the most frequently used evaluation metrics in the used benchmark collections, therefore we will limit our meta-analysis to evaluation results reported in one of these four metrics.
+The Yahoo!
+Learning to Rank Challenge benchmark is an exception, on this benchmark ERR is the most used evaluation metric.
 The lack of use of the ERR metric in other benchmarks makes it unsuitable for a cross-benchmark comparison.
 By making the decision to include NDCG at three cut-off points and only a single MAP entry, we implicitly attain a higher weight for NDCG compared to MAP on an analysis that combines all measurements on the four metrics.
-This implicit weighting is arbitrary and therefore undesirable, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.
+This implicit could be regarded as arbitrary, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.\\
 
+Note that another implicit weighting lies in the paper dimension.
+Hence, the higher number of evaluation results specified in a paper, the higher the influence of this paper on the outcome of the analysis.
+This implicit weighted is not harmful to the validity of our comparison, as papers with a large number of evaluation results are not expected to be less reliable than papers with fewer evaluation results.\\
+\\
 \begin{figure*}
@@
 \caption{NDCG@5 comparison of learning to rank methods}
-\label{fig:normalised_winning_number_ndcg5}
+\label{fig:normalized_winning_number_ndcg5}
 \end{figure*}
@@
 \subsection{NDCG@3}
-Figure \ref{fig:normalised_winning_number_ndcg3} shows the performance of learning to rank methods for the NDCG@3 metric.
+Figure \ref{fig:normalized_winning_number_ndcg3} shows the performance of learning to rank methods for the NDCG@3 metric.
 LambdaNeuralRank and CoList both acquired a perfect Normalized Winning Number score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
@@
 \subsection{NDCG@5}
-Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of learning to rank methods for the NDCG@5 metric.
+Figure \ref{fig:normalized_winning_number_ndcg5} shows the performance of learning to rank methods for the NDCG@5 metric.
 LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a Normalized Winning Number of 1.0.
@@
 \caption{NDCG@10 comparison of learning to rank methods}
-\label{fig:normalised_winning_number_ndcg10}
+\label{fig:normalized_winning_number_ndcg10}
 \end{figure*}
 
-Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of learning to rank methods for the NDCG@10 metric.
+Figure \ref{fig:normalized_winning_number_ndcg10} shows the performance of learning to rank methods for the NDCG@10 metric.
 LambdaMART and LambdaNeuralRank score a Normalized Winning Number of 1.0 on the NDCG@10 comparison.
@@
 \caption{MAP comparison of learning to rank methods}
-\label{fig:normalised_winning_number_map}
+\label{fig:normalized_winning_number_map}
 \end{figure*}
-Figure \ref{fig:normalised_winning_number_map} shows the performance of learning to rank methods for the MAP metric.
+Figure \ref{fig:normalized_winning_number_map} shows the performance of learning to rank methods for the MAP metric.
 Where comparisons on the NDCG metric at different cut-off points where highly in agreement in terms of the best performing algorithms, the comparison in terms of MAP shows different results.
@@
 \caption{Cross-benchmark comparison of learning to rank methods}
-\label{fig:normalised_winning_number_all}
+\label{fig:normalized_winning_number_all}
 \end{figure*}
-Figure \ref{fig:normalised_winning_number_all} shows the Normalized Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}.
+Figure \ref{fig:normalized_winning_number_all} shows the Normalized Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}.
 The cross-metric comparison is based on the NDCG@$\{3,5,10\}$ and MAP comparisons combined, which justifies analyzing the comparison more thoroughly.
-Figure \ref{fig:normalised_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
+Figure \ref{fig:normalized_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
 In addition, Linear Regression and the ranking method of simply sorting on the best single feature are labelled as baselines.\\
@@
 By calculating it in this way, we implicitly make the assumption that the learning to rank methods are (approximately) distributed uniformly over the benchmarks, such that the average learning to rank method tested are approximately equally hard for each data set.
-It could be the case however that this assumption is false and that significantly more accurate learning to rank methods are evaluated on some datasets than on other datasets.
-\\
+It could be the case however that this assumption is false and that significantly more accurate learning to rank methods are evaluated on some datasets than on other datasets.\\
 
-A second limitation is that the datasets on which learning to rank methods have been evaluated can not always be regarded a random choice.
+A second limitation is that the datasets on which learning to rank methods have been evaluated cannot always be regarded a random choice.
 It might be the case that some researchers chose to publish results for exactly those benchmark datasets that showed the most positive results for their learning to rank method.\\
@@
 Limiting the meta-analysis to those studies that report comparable results on one of the baseline methods of a benchmark set reduces this limitation but does not solve it completely.
-By taking the Ideal Winning Number into account in Figure \ref{fig:normalised_winning_number_all} we further mitigate this limitation, as the Ideal Winning Number is loosely related with the number of studies that reported evaluation results for an algorithm.
+By taking the Ideal Winning Number into account in Figure \ref{fig:normalized_winning_number_all} we further mitigate this limitation, as the Ideal Winning Number is loosely related with the number of studies that reported evaluation results for an algorithm.
 
@@
 % produce the bibliography for the citations in your paper.
-\newpage
+\clearpage
 \bibliographystyle{abbrv}

### 2014-05-20 azure_grant_proposal/proposal.tex
research proposal

--- 2014-04-15 report/Chapters/Introduction.tex
@@
-\section{Motivation and Problem Statement}
+% This is "sig-alternate.tex" V2.0 May 2012
+% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
+%
+% This example file demonstrates the use of the 'sig-alternate.cls'
+% V2.5 LaTeX2e document class file. It is for those submitting
+% articles to ACM Conference Proceedings WHO DO NOT WISH TO
+% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
+% The 'sig-alternate.cls' file will produce a similar-looking,
+% albeit, 'tighter' paper resulting in, invariably, fewer pages.
+%
+% ----------------------------------------------------------------------------------------------------------------
+% This .tex file (and associated .cls V2.5) produces:
+%       1) The Permission Statement
+%       2) The Conference (location) Info information
+%       3) The Copyright Line with ACM data
+%       4) NO page numbers
+%
+% as against the acm_proc_article-sp.cls file which
+% DOES NOT produce 1) thru' 3) above.
+%
+% Using 'sig-alternate.cls' you have control, however, from within
+% the source .tex file, over both the CopyrightYear
+% (defaulted to 200X) and the ACM Copyright Data
+% (defaulted to X-XXXXX-XX-X/XX/XX).
+% e.g.
+% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
+% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
+%
+% ---------------------------------------------------------------------------------------------------------------
+% This .tex source is an example which *does* use
+% the .bib file (from which the .bbl file % is produced).
+% REMEMBER HOWEVER: After having produced the .bbl file,
+% and prior to final submission, you *NEED* to 'insert'
+% your .bbl file into your source .tex file so as to provide
+% ONE 'self-contained' source file.
+%
+% ================= IF YOU HAVE QUESTIONS =======================
+% Questions regarding the SIGS styles, SIGS policies and
+% procedures, Conferences etc. should be sent to
+% Adrienne Griscti (griscti@acm.org)
+%
+% Technical questions _only_ to
+% Gerald Murray (murray@hq.acm.org)
+% ===============================================================
+%
+% For tracking purposes - this is V2.0 - May 2012
+
+\documentclass{sig-alternate}
+\begin{document}
+
+%
+% --- Author Metadata here ---
+\conferenceinfo{}{}
+\CopyrightYear{2014} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
+%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
+% --- End of Author Metadata ---
+
+\title{Research Proposal: A Scalability Analysis of Learning to Rank Algorithms using Hadoop}
+%
+% You need the command \numberofauthors to handle the 'placement
+% and alignment' of the authors beneath the title.
+%
+% For aesthetic reasons, we recommend 'three authors at a time'
+% i.e. three 'name/affiliation blocks' be placed beneath the title.
+%
+% NOTE: You are NOT restricted in how many 'rows' of
+% "name/affiliations" may appear. We just ask that you restrict
+% the number of 'columns' to three.
+%
+% Because of the available 'opening page real-estate'
+% we ask you to refrain from putting more than six authors
+% (two rows with three columns) beneath the article title.
+% More than six makes the first-page appear very cluttered indeed.
+%
+% Use the \alignauthor commands to handle the names
+% and affiliations for an 'aesthetic maximum' of six authors.
+% Add names, affiliations, addresses for
+% the seventh etc. author(s) as the argument for the
+% \additionalauthors command.
+% These 'additional authors' will be output/set for you
+% without further effort on your part as the last section in
+% the body of your article BEFORE References or any Appendices.
+
+\numberofauthors{1} %  in this sample file, there are a *total*
+% of EIGHT authors. SIX appear on the 'first-page' (for formatting
+% reasons) and the remaining two appear in the \additionalauthors section.
+%
+\author{
+% You can go ahead and credit any number of authors here,
+% e.g. one 'row of three' or two rows (consisting of one row of three
+% and a second row of one, two or three).
+%
+% The command \alignauthor (no curly braces needed) should
+% precede each author name, affiliation/snail-mail address and
+% e-mail address. Additionally, tag each line of
+% affiliation/address with \affaddr, and tag the
+% e-mail address with \email.
+%
+% 1st. author
+\alignauthor
+Niek Tax\\
+       \affaddr{University of Twente}\\
+       \affaddr{P.O. Box 217}\\
+       \affaddr{7500 AE Enschede}\\
+       \affaddr{The Netherlands}
+       \affaddr{d.hiemstra@utwente.nl}
+}
+
+\maketitle
+
+% A category with the (minimum) three required fields
+\category{H.3.3}{Information Systems}{Information Storage and Retrieval}[Information Search and Retrieval]
+%
+\category{D.1.3}{Concurrent Programming}{Distributed Programming}
+
+\keywords{Learning to rank, MapReduce}
+\section{Introduction}
 Ranking is a core problem in the field of information retrieval.
@@
 
-\section{Thesis Overview}
-
-\begin{description}
-\item[Chapter II: Background]{introduces the reader to the basic principles and recent work in the fields of Learning-to-Rank.}
-\item[Chapter III: Related Work]{concisely describes existing work in the field of parallel Learning-to-Rank.}
-\item[Chapter IV: Benchmark Results]{sketches the accuracy of existing Learning-to-Rank methods on several benchmark datasets and describes the selection of Learning-to-Rank methods for the parallelisation experiments.}
-\item[Chapter V: Selected Learning-to-Rank Methods]{describes the algorithms and details of the Learning-to-Rank methods selected in Chapter IV.}
-\item[Chapter VI: Implementation]{describes implementation details of the Learning-to-Rank algorithms in the Hadoop framework.}
-\item[Chapter VII: Results \& Discussion]{presents and discusses speed-up results for the implemented Learning-to-Rank methods.}
-\item[Chapter VIII: Conclusion]{summarizes the results and answers our research questions based on the results.
-The limitations of our research as well as future research directions in the field are mentioned here.}
-\end{description}
+\bibliographystyle{abbrv}
+\bibliography{Bibliography}  % sigproc.bib is the name of the
+\end{document}

### 2014-05-20 azure_grant_proposal/proposal.tex
opmaak

@@
 % For tracking purposes - this is V2.0 - May 2012
-
 \documentclass{sig-alternate}
+\usepackage{array}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\usepackage{booktabs}
 \begin{document}
@@
 
+\begin{abstract}
+Learning to rank is an increasingly important scientific field that comprises the use of machine learning for the ranking task.
+This research proposal proposes a two-stage research project in the learning to rank field.
+In the first research stage we will conduct a comparison of learning to rank methods over multiple benchmark collections.
+This is relevant, as a lack of a \emph{de facto} standard benchmark collection for evaluation of new learning to rank methods hinders comparison.
+The second stage analyses the potential speed-up gain of parallel execution of the most accurate learning to rank methods identified in the first research step through execution on Hadoop MapReduce.
+Research on scalability of learning to rank methods is widely seen as relevant and often overlooked \cite{Chapelle2011b, Liu2007}.
+\end{abstract}
 % A category with the (minimum) three required fields
@@
 
-\keywords{Learning to rank, MapReduce}
+\keywords{Learning to Rank, MapReduce}
 \section{Introduction}
@@
 The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models.
-Learning-to-Rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\
+Learning to rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\
 
-In recent years several Learning-to-Rank benchmark datasets have been proposed that enable comparison of the performance of different Learning-to-Rank methods.
+In recent years several learning to rank benchmark datasets have been proposed that enable comparison of the performance of different Learning to rank methods.
 Well-known benchmark datasets include the \emph{Yahoo!
@@
 One of the concluding observations of the \emph{Yahoo!
-Learning to Rank Challenge} was that almost all work in the Learning-to-Rank field focuses on ranking accuracy, while efficiency and scalability of Learning-to-Rank algorithms is still an underexposed research area that is likely to become more important in the near future as training sets are becoming larger and larger \cite{Chapelle2011b}.
-Liu \cite{Liu2007} confirms the observation that efficiency and scalability of Learning-to-Rank methods has so far been an overlooked research area in his influential book on Learning-to-Rank.\\
+Learning to Rank Challenge} was that almost all work in the Learning to rank field focuses on ranking accuracy, while efficiency and scalability of learning to rank algorithms is still an underexposed research area that is likely to become more important in the near future as training sets are becoming larger and larger \cite{Chapelle2011b}.
+Liu \cite{Liu2007} confirms the observation that efficiency and scalability of learning to rank methods has so far been an overlooked research area in his influential book on learning to rank.\\
 
 Some research has been done in the area of parallel or distributed machine learning \cite{Chu2007,Chang2007}.
-However, almost none of these studies include the Learning-to-Rank sub-field of machine learning.
-The field of efficient Learning-to-Rank has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012}, since Liu \cite{Liu2007} first stated its growing importance back in 2007.
-Only several of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient Learning-to-Rank through the use of parallel programming paradigms.\\
+However, almost none of these studies include the learning to rank sub-field of machine learning.
+The field of efficient learning to rank has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012}, since Liu \cite{Liu2007} first stated its growing importance back in 2007.
+Only several of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient learning to rank through the use of parallel programming paradigms.\\
 
@@
 Since Google developed the MapReduce parallel programming framework back in 2004 it has grown to be the industry standard model for parallel programming.
-Lin \cite{Lin2013} observed that algorithms that are of iterative nature, which most Learning-to-Rank algorithms are, are not amenable to the MapReduce framework.
+Lin \cite{Lin2013} observed that algorithms that are of iterative nature, which most learning to rank algorithms are, are not amenable to the MapReduce framework.
 Lin argued that as a solution to the non-amenability of iterative algorithms to the MapReduce framework, iterative algorithms can often be replaced with non-iterative alternatives or can still be optimized in such a way that its performance in a MapReduce setting is good enough.\\
 
-The appearance of benchmark datasets gave insight in the performance of different Learning-to-Rank approaches, which resulted in increasing popularity of those methods that showed to perform well on one or more benchmark datasets.
-Up to now it remains unknown whether popular existing Learning-to-Rank methods scale well when they are used in a parallel manner using the MapReduce framework.
-This thesis aims to be an exploratory start in this little researched area of parallel Learning-to-Rank.
-A more extensive overview of my research goals and questions are described in section \ref{sec:goals}.\\
+The appearance of benchmark datasets gave insight in the performance of different learning to rank approaches, which resulted in increasing popularity of those methods that showed to perform well on one or more benchmark datasets.
+Up to now it remains unknown whether popular existing learning to rank methods scale well when they are used in a parallel manner using the MapReduce framework.
+This thesis aims to be an exploratory start in this little researched area of parallel learning to rank.\\
 
@@
 \label{sec:goals}
-The objective of this thesis is to explore the speed-up in execution time of Learning-to-Rank algorithms through parallelisation using the MapReduce framework.
-This work focuses on those Learning-to-Rank algorithms that have shown leading performance on relevant benchmark datasets.
-This thesis addresses the following research questions:
+Our research contains two research stages.
+A first stage is concerned with identifying the best performing learning to rank methods in terms of accuracy.
+Several benchmark collections for model evaluation exist in the learning to rank field.
+Well-known benchmark collections include the \emph{Yahoo!
+Learning to Rank Challenge} collection \cite{Chapelle2011a}, the \emph{Yandex Internet Mathematics 2009} contest\footnote{http://imat2009.yandex.ru/en}, the LETOR datasets \cite{Qin2010}, and the MSLR (Microsoft Learning to Rank) datasets\footnote{http://research.microsoft.com/en-us/projects/mslr/}.
+There exists no agreement among authors in the learning to rank field on which benchmark collection(s) to use to evaluate a new model.
+Comparing ranking accuracy of learning to rank methods is largely hindered by this lack of a \emph{de facto} standard way of benchmarking new models.
+Therefore, our first research question concerns the identification of the most accurate learning to rank methods when compared on their performance on several benchmark collections.
 \begin{description}
-\item[RQ1] What are the best performing Learning-to-Rank algorithms in terms of accuracy on relevant benchmark datasets?\\
-
-\item[RQ2] What is the speed-up of those Learning-to-Rank algorithms when executed using the MapReduce framework?\\
-Where the definition of \emph{relative speed-up} is used for speed-up \cite{Sun1991}:\\
+\item[RQ1] What are the best performing learning to rank algorithms in terms of accuracy on when compared over multiple benchmark collections?
+\end{description}
+The second stage of our research is concerned with exploration of the speed-up in execution time of learning to rank algorithms through parallelisation using the Hadoop MapReduce framework.
+This second stages focuses on those learning to rank algorithms that showed to be most accurate in our cross-benchmark comparison of learning to rank methods.
+\begin{description}
+\item[RQ2] What is the speed-up of those learning to rank algorithms when executed using the MapReduce framework?
+\end{description}
+\noindent Where the definition of \emph{relative speed-up} is used for speed-up \cite{Sun1991}:\\
 
@@
 
-\item[RQ3] Can those Learning-to-Rank algorithms be adjusted in such a way that the parallel execution speed-up increases without decreasing accuracy?
-\end{description}
-
-\section{Approach}
-A literature study will be performed to get insight in relevant existing techniques for large scale Learning-to-Rank. The literature study will be performed by using the following query:\\
-
-\emph{("learning to rank" \emph{OR} "learning-to-rank" \emph{OR} "machine learned ranking") \emph{AND} ("parallel" \emph{OR} "distributed")}\\
-
-and the following bibliographic databases:
-\begin{itemize}
-\item Scopus
-\item Web of Science
-\item Google Scholar
+\section{Methodology}
+The following subsections will describe the research methodology per stage of our two-stage research.
+\subsection{Cross-benchmark learning to rank comparison}
+We collect evaluation results on the datasets of benchmark collections through a structured literature search.
+The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
+\begin{itemize}\itemsep0em
+\item LETOR 2.0
+\item LETOR 3.0
+\item LETOR 4.0
+\item Yahoo! Learning to Rank Challenge
+\item Yandex Internet Mathematics 2009 contest
+\item MSLR-web10/30k
+\item WCL2R
+\item AOL
 \end{itemize}
 
-The query incorporates different ways of writing of Learning-to-Rank, with and without hyphens, and the synonymous term \emph{machine learned ranking} to increase search recall, i.e.
-to make sure that no relevant studies are missed.
-For the same reason the terms \emph{parallel} and \emph{distributed} are included in the search query.
-Even though \emph{parallel} and \emph{distributed} are not always synonymous is all definitions, we are interested in both approaches in non-sequential data processing.\\
+For the LETOR collections, the evaluation results of the baseline models will be used from LETOR 2.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor2.0/baseline.aspx}, 3.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and 4.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} as listed on the LETOR website.\\
 
-A one-level forward and backward reference search is used to find relevant papers missed so far.
-To handle the large volume of studies involved in the backward and forward reference search, relevance of the studies will be evaluated solely on the title of the study.\\
+LETOR 1.0 and 3.0, Yahoo!
+Learning to Rank Challenge, WCL2R and AOL have accompanying papers which were published together with these benchmark collections.
+Authors publishing evaluation results on these benchmark collections are requested to cite these papers.
+Therefore, we collect evaluation measurements of learning to rank methods on these benchmark collections through forward literature search.
+Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search.
+Google Scholar will be used to perform the forward reference search.
 
-To answer the first research question, I will implement Learning-to-Rank methods in the MapReduce framework and measure runtime as a factor of the number of cluster nodes used to complete the computation.\\
+\begin{table}[!h]
+\begin{tabular}{p{2.8cm}p{2.64cm}R{1.9cm}}\toprule
+Benchmark & Paper & \# of forward references \\
+\midrule
+LETOR 1.0 \& 2.0 & Liu et al. \cite{Liu2007b} & 307\\
+LETOR 3.0 & Qin et al. \cite{Qin2010} & 105\\
+Yahoo! Learning to Rank Challenge & Chapelle et al. \cite{Chapelle2011a} & 102\\
+AOL dataset & Pass et al. \cite{Pass2006} & 339\\
+WCL2R & Alc{\^a}ntara et al. \cite{Alcantara2010} & 2\\
+\bottomrule
+\end{tabular}
+\caption{Forward references of learning to rank benchmark papers}
+\label{tbl:ltr_benchmark_forref}
+\end{table}
 
-I will use the HDInsight cloud-based MapReduce implementation from Microsoft to implement the Learning-to-Rank algorithms.
-HDInsight is based on the popular open source MapReduce implementation Hadoop\footnote{http://hadoop.apache.org/}.
-The algorithms that we include in the measurements will be determined based on experimental results on the \emph{Yahoo!
-Learning to Rank Challenge} \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, the LETOR \cite{Qin2010} dataset and the LETOR successors MSLR-WEB10k and MSLR-WEB30k.
+The LETOR 4.0, MSLR-web10/30k and Yandex Internet Mathematics Competition 2009 benchmark collections are not accompanied by a paper.
+To collect evaluation results for learning to rank methods on these benchmarks, a Google Scholar search is performed on the name of the benchmark.
+Table~\ref{tbl:ltr_benchmark_searches} shows the results of this literature search.
+
+\begin{table}[!h]
+\begin{tabular}{p{5.2cm}R{2.5cm}}\toprule
+Search Query & Google scholar search results \\
+\midrule
+"LETOR 4.0" & 75 \\
+"MSLR-web10k" & 16 \\
+"MSLR-web30k" & 15 \\
+"Yandex Internet Mathematics" & 3 \\
+\bottomrule
+\end{tabular}
+\caption{Google scholar search results for learning to rank benchmarks}
+\label{tbl:ltr_benchmark_searches}
+\end{table}
+
+\subsection{Analysis of scalability on MapReduce}
+The learning to rank methods identified in the first research stage will be implemented in Pig to run them in parallel on a MapReduce platform.
+The Microsoft HDInsight cloud-based MapReduce implementation will be used to execute to run the experiments on.
+Microsoft HDInsight has the benefit of being able to scale to large numbers of MapReduce nodes with guaranteed availability.
 

### 2014-05-21 azure_grant_proposal/proposal.tex
afwerking azure grant proposal

@@
 % --- Author Metadata here ---
-\conferenceinfo{}{}
-\CopyrightYear{2014} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
+
 %\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
@@
 This is relevant, as a lack of a \emph{de facto} standard benchmark collection for evaluation of new learning to rank methods hinders comparison.
-The second stage analyses the potential speed-up gain of parallel execution of the most accurate learning to rank methods identified in the first research step through execution on Hadoop MapReduce.
+The second stage analyses the potential speed-up through parallel execution of the most accurate learning to rank methods identified in the first research step through execution on Hadoop MapReduce.
 Research on scalability of learning to rank methods is widely seen as relevant and often overlooked \cite{Chapelle2011b, Liu2007}.
 \end{abstract}
-% A category with the (minimum) three required fields
-\category{H.3.3}{Information Systems}{Information Storage and Retrieval}[Information Search and Retrieval]
-%
-\category{D.1.3}{Concurrent Programming}{Distributed Programming}
 
-\keywords{Learning to Rank, MapReduce}
 \section{Introduction}
@@
 Research in the field of ranking models has for a long time been based on manually designed ranking functions, such as the well-known BM25 model \cite{Robertson1994}.
-Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
 The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models.
@@
 The following subsections will describe the research methodology per stage of our two-stage research.
-\subsection{Cross-benchmark learning to rank comparison}
-We collect evaluation results on the datasets of benchmark collections through a structured literature search.
-The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
+\subsection{Cross-Benchmark Learning to Rank Comparison}
+\subsubsection{Gathering Evaluation Results}
+We will collect evaluation results on the datasets of benchmark collections through a structured literature search.
+The following list presents an overview of the benchmark collections to be taken into account in the meta-analysis:
 \begin{itemize}\itemsep0em
@@
 Authors publishing evaluation results on these benchmark collections are requested to cite these papers.
-Therefore, we collect evaluation measurements of learning to rank methods on these benchmark collections through forward literature search.
-Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search.
-Google Scholar will be used to perform the forward reference search.
-
-\begin{table}[!h]
-\begin{tabular}{p{2.8cm}p{2.64cm}R{1.9cm}}\toprule
-Benchmark & Paper & \# of forward references \\
-\midrule
-LETOR 1.0 \& 2.0 & Liu et al. \cite{Liu2007b} & 307\\
-LETOR 3.0 & Qin et al. \cite{Qin2010} & 105\\
-Yahoo! Learning to Rank Challenge & Chapelle et al. \cite{Chapelle2011a} & 102\\
-AOL dataset & Pass et al. \cite{Pass2006} & 339\\
-WCL2R & Alc{\^a}ntara et al. \cite{Alcantara2010} & 2\\
-\bottomrule
-\end{tabular}
-\caption{Forward references of learning to rank benchmark papers}
-\label{tbl:ltr_benchmark_forref}
-\end{table}
+Therefore, we will collect evaluation measurements of learning to rank methods on these benchmark collections through forward literature search.
 
 The LETOR 4.0, MSLR-web10/30k and Yandex Internet Mathematics Competition 2009 benchmark collections are not accompanied by a paper.
-To collect evaluation results for learning to rank methods on these benchmarks, a Google Scholar search is performed on the name of the benchmark.
-Table~\ref{tbl:ltr_benchmark_searches} shows the results of this literature search.
+To collect evaluation results for learning to rank methods on these benchmarks, a Google Scholar search will be performed on the name of the benchmark.
 
-\begin{table}[!h]
-\begin{tabular}{p{5.2cm}R{2.5cm}}\toprule
-Search Query & Google scholar search results \\
-\midrule
-"LETOR 4.0" & 75 \\
-"MSLR-web10k" & 16 \\
-"MSLR-web30k" & 15 \\
-"Yandex Internet Mathematics" & 3 \\
-\bottomrule
-\end{tabular}
-\caption{Google scholar search results for learning to rank benchmarks}
-\label{tbl:ltr_benchmark_searches}
-\end{table}
+\subsubsection{Methodology for Cross-Benchmark Comparison of Learning to Rank models}
+Qin et al.
+\cite{Qin2010} state in the LETOR 3.0 paper that it may differ between datasets what the most accurate ranking methods are.
+They propose a measure they call \emph{Winning Number} to evaluate the overall performance of learning to rank methods over the multiple datasets in the LETOR 3.0 collection.
+Winning Number is defined as the number of other algorithms that an algorithm can beat over the set of datasets, or more formally\\
 
-\subsection{Analysis of scalability on MapReduce}
+$\text{Winning Number}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$\\
+
+where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as NDCG or MAP), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that\\
+
+$I_{\{M_i(j)>M_k(j)\}} = \begin{cases}
+1 & \text{if } M_i(j) > M_k(j), \\
+0 & \text{otherwise}
+\end{cases}$\\
+
+In contrast to the winning number comparison on LETOR 3.0, there will not be evaluation results for each algorithm on each dataset in our meta-analysis.
+We propose a normalized version of the Winning Number metric to compare algorithms based on a sparse set of evaluation measurements.
+This Normalized Winning Number takes only those datasets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all datasets on which it has been evaluated.
+We will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as
+
+\[
+I_{M_i(j)>M_k(j)} = \begin{cases}
+1 & \parbox[t]{5cm}{if $M_i(j)$ and $M_k(j$ are both defined and $M_i(j) > M_k(j)$,}\\
+0 & \parbox[t]{5cm}{otherwise}
+\end{cases}
+\]
+From now on this adjusted version of Winning Number will be references to as \emph{Normalized Winning Number}. The mathematical definition of Normalized Winning Number is\\
+
+$\text{Normalized Winning Number}_i(M) = \frac{\text{Winning Number}_i(M)}{\text{Ideal Winning Number}_i(M)}$\\
+
+\noindent
+where Ideal Winning Number is defined as\\
+
+$\text{Ideal Winning Number}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m D_{\{M_i(j),M_k(j)\}}$\\
+
+where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as NDCG or MAP), and $D_{\{M_i(j),M_k(j)\}}$ is an evaluation definition function such that\\
+
+$D_{\{M_i(j),M_k(j)\}} = \begin{cases}
+1 & \text{if } M_i(j) \text{ and } M_k(j) \text{ are both defined}, \\
+0 & \text{otherwise}
+\end{cases}$\\
+
+\subsection{Analysis of Scalability on MapReduce}
 The learning to rank methods identified in the first research stage will be implemented in Pig to run them in parallel on a MapReduce platform.

### 2014-05-21 cikm_paper/paper.tex
de grote Djoerd en Sander wisseltruc
mooi begin feedback Djoerd CIKM paper

@@
 
-\title{A Cross-Benchmark Comparison of Learning to Rank Methods}
+\title{A Cross-Benchmark Comparison of 87 Learning to Rank Methods}
 %
@@
        \affaddr{The Netherlands}
-       \email{niek.tax@avanade.com}
+       \affaddr{niek.tax@avanade.com}
 % 2nd. author
+\alignauthor
+Sander Bockting\\
+       \affaddr{Avanade Netherlands B.V.}\\
+       \affaddr{Versterkerstraat 6}\\
+       \affaddr{1322 AP Almere}\\
+       \affaddr{The Netherlands}
+       \affaddr{sander.bockting@avanade.com}
+% 3rd. author
 \alignauthor
@@
        \affaddr{The Netherlands}
-       \email{d.hiemstra@utwente.nl}
-% 3rd. author
-\alignauthor
-Sander Bockting\\
-       \affaddr{Avanade Netherlands B.V.}\\
-       \affaddr{Versterkerstraat 6}\\
-       \affaddr{1322 AP Almere}\\
-       \affaddr{The Netherlands}
-       \email{sander.bockting@avanade.com}
+       \affaddr{d.hiemstra@utwente.nl}
 }
@@
 Our comparison methodology consists of two components: 1) Normalized Winning Number, which gives insight in the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning its ranking accuracy.
-Evaluation results on a set of well-known benchmark collections are collected through a structured literature search.
-ListNet, SmoothRank, FenchelRank, FSMRank, LRUF and LARF turned out to be the best performing learning to rank methods in increasing order of Normalized Winning Number and decreasing order of Ideal Winning Number.
+Evaluation results of 87 learning to rank methods on 20 well-known benchmark datasets are collected through a structured literature search.
+ListNet, SmoothRank, FenchelRank, FSMRank, LRUF and LARF are the best performing learning to rank methods in increasing order of Normalized Winning Number and decreasing order of Ideal Winning Number.
 \end{abstract}
@@
 For example, ranking plays a vital role in automatic text summarization, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}.
-In addition, the ranking task has been argued to be a better fit for recommender systems compared to the regression task on a continuous scale \cite{Adomavicius2005,McNee2006}.\\
+In addition, the ranking task has been argued to be a better fit for recommender systems compared to the regression task (concerning predictions on a continuous scale) \cite{Adomavicius2005,McNee2006}.\\
 
@@
 Well-known benchmark datasets in the learning to rank field include the \emph{Yahoo!
-Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the \emph{Yandex Internet Mathematics 2009} contest\footnote{http://imat2009.yandex.ru/en}, the LETOR dataset \cite{Qin2010}, and the MSLR (Microsoft Learning to Rank) datasets\footnote{http://research.microsoft.com/en-us/projects/mslr/}.
+Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the \emph{Yandex Internet Mathematics 2009} contest\footnote{http://imat2009.yandex.ru/en}, the LETOR datasets \cite{Qin2010}, and the MSLR (Microsoft Learning to Rank) datasets\footnote{http://research.microsoft.com/en-us/projects/mslr/}.
 There exists no agreement among authors in the learning to rank field on which benchmark collection(s) to use to evaluate a new model.
@@
 Busa-Fekete et al.
-\cite{Busa-Fekete2013} compared the accuracy of a small set of models over the LETOR 4.0 datasets, both MSLR datasets, both Yahoo!
+\cite{Busa-Fekete2013} compared the accuracy of a small set of models over the LETOR 4.0 datasets, both MSLR datasets, both the Yahoo!
 Learning to Rank Challenge datasets and one of the datasets from LETOR 3.0.
@@
 We collect evaluation results on the datasets of benchmark collections through a structured literature search.
-The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
-\begin{itemize}\itemsep0em
-\item LETOR 2.0
-\item LETOR 3.0
-\item LETOR 4.0
-\item Yahoo! Learning to Rank Challenge
-\item Yandex Internet Mathematics 2009 contest
-\item MSLR-web10/30k
-\item WCL2R
-\item AOL
-\end{itemize}
+Table \ref{tbl:ltr_benchmark_collections} presents an overview of the benchmark collections taken into account in the meta-analysis:
+
+\begin{table}[!h]
+\begin{tabular}{p{6.4cm}R{1.26cm}}\toprule
+Benchmark collection & \# of datasets \\
+\midrule
+AOL		  & 1\\
+LETOR 2.0 & 3\\
+LETOR 3.0 & 7\\
+LETOR 4.0 & 2\\
+MSLR	  & 2\\
+WCL2R	  & 2\\
+Yahoo! Learning to Rank Challenge 	     & 2\\
+Yandex Internet Mathematics 2009 contest & 1\\
+\textbf{Total} & \textbf{20}\\ 	
+\bottomrule
+\end{tabular}
+\caption{Included learning to rank evaluation benchmark collections}
+\label{tbl:ltr_benchmark_collections}
+\end{table}
 
@@
 Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search.
-Google Scholar will be used to perform the forward reference search.
+Google Scholar is used to perform the forward reference search.\\
 
@@
 \subsection{Literature Selection}
-Table \ref{tab:ltr_methods_used} in Appendix \ref{app:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found through described procedure.
+Table \ref{tab:ltr_methods_used} in Appendix \ref{app:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found through the described procedure.
 Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, 3.0 and 4.0 baselines respectively.\\
@@
 \includegraphics[scale=0.371]{gfx/ndcg3_winnum}
-\caption{NDCG@3 comparison of learning to rank methods}
+\caption{NDCG@3 comparison of 87 learning to rank methods}
 \label{fig:normalized_winning_number_ndcg3}
@@
 Qin et al.
-\cite{Qin2010} state in the LETOR 3.0 paper that it may differ between datasets what the most accurate ranking methods are.
+\cite{Qin2010} state that it may differ between datasets what the most accurate ranking methods are.
 They propose a measure they call \emph{Winning Number} to evaluate the overall performance of learning to rank methods over the multiple datasets in the LETOR 3.0 collection.
@@
 
-$\text{Winning Number}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$\\
+$\text{WN}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$\\
 
@@
 
-In contrast to the winning number comparison on LETOR 3.0, there will not be evaluation results for each algorithm on each dataset in our meta-analysis.
-We propose a normalized version of the Winning Number metric to compare algorithms based on a sparse set of evaluation measurements.
+The LETOR 3.0 was a comparison on a \emph{dense} set of evaluation results, in the sense that there were evaluation results available for all learning to rank algorithms on all datasets included in their comparison.
+The Winning Number evaluation metric relies on the denseness of the evaluation results set.
+In contrast to the LETOR 3.0 comparison, our set of evaluation results will be sparse.
+We propose a normalized version of the Winning Number metric to enable comparison of a sparse set of evaluation results.
 This Normalized Winning Number takes only those datasets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all datasets on which it has been evaluated.
@@
 \[
-I_{M_i(j)>M_k(j)} = \begin{cases}
-1 & \parbox[t]{5cm}{if $M_i(j)$ and $M_k(j$ are both defined and $M_i(j) > M_k(j)$,}\\
+I^{'}_{M_i(j)>M_k(j)} = \begin{cases}
+1 & \parbox[t]{5cm}{if $M_i(j)$ and $M_k(j)$ are both defined and $M_i(j) > M_k(j)$,}\\
 0 & \parbox[t]{5cm}{otherwise}
@@
 
-$\text{Normalized Winning Number}_i(M) = \frac{\text{Winning Number}_i(M)}{\text{Ideal Winning Number}_i(M)}$\\
+$\text{NWN}_i(M) = \frac{\text{WN}_i(M)}{\text{IWN}_i(M)}$\\
 
 \noindent
-where Ideal Winning Number is defined as\\
+where $IWN$ is the Ideal Winning Number, defined as\\
 
-$\text{Ideal Winning Number}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m D_{\{M_i(j),M_k(j)\}}$\\
+$\text{IWN}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m D_{\{M_i(j),M_k(j)\}}$\\
 
@@
 Learning to Rank Challenge benchmark is an exception, on this benchmark ERR is the most used evaluation metric.
-The lack of use of the ERR metric in other benchmarks makes it unsuitable for a cross-benchmark comparison.
-By making the decision to include NDCG at three cut-off points and only a single MAP entry, we implicitly attain a higher weight for NDCG compared to MAP on an analysis that combines all measurements on the four metrics.
-This implicit could be regarded as arbitrary, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.\\
+The lack of use of the ERR metric in other benchmarks makes it unsuitable for a cross-benchmark comparison.\\
 
-Note that another implicit weighting lies in the paper dimension.
-Hence, the higher number of evaluation results specified in a paper, the higher the influence of this paper on the outcome of the analysis.
-This implicit weighted is not harmful to the validity of our comparison, as papers with a large number of evaluation results are not expected to be less reliable than papers with fewer evaluation results.\\
-\\
 \begin{figure*}
 \includegraphics[scale=0.371]{gfx/ndcg5_winnum}
-\caption{NDCG@5 comparison of learning to rank methods}
+\caption{NDCG@5 comparison of 87 learning to rank methods}
 \label{fig:normalized_winning_number_ndcg5}
@@
 \includegraphics[scale=0.371]{gfx/ndcg10_winnum}
-\caption{NDCG@10 comparison of learning to rank methods}
+\caption{NDCG@10 comparison of 87 learning to rank methods}
 \label{fig:normalized_winning_number_ndcg10}
@@
 The set of algorithms for which there is no other algorithm with both a higher Normalized Winning Number and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and @5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
-SmoothRank and FSMRank were not present in this set for the NDCG@3 and @5 comparison, but were close by, as can be seen in Table \ref{tab:raw_data} in the Appendix \ref{app:raw_data}.
-DCMP is not in the set in contrast with the NDCG@3 and @5 comparison.
+In contrast to the NDCG@3 and @5 comparison, DCMP is not among the best algorithms.
 
@@
 \includegraphics[scale=0.371]{gfx/map_winnum}
-\caption{MAP comparison of learning to rank methods}
+\caption{MAP comparison of 87 learning to rank methods}
 \label{fig:normalized_winning_number_map}
@@
 LARF and LRUF score very high Normalized Winning Number scores, but based on only relatively few datasets, just as in the NDCG comparisons.
-Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all NDCG comparisons, are not within the best performing methods in the MAP comparison.
+Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all NDCG comparisons, are not among the best performing methods in the MAP comparison.
 A deeper look in the raw data Table \ref{tab:raw_data} in the Appendix \ref{app:raw_data} shows that LAC-MR-OR is evaluated on many more datasets for MAP compared to NDCG, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalized Winning Number.
@@
 RE-QR is one of the best performers in the MAP comparison with a reasonable amount of benchmark evaluations.
-No reported NDCG performance was found in the literature study for RE-QR.
+No reported NDCG performance was found in the literature search for RE-QR.
 There is a lot of certainty on the accuracy of RankBoost and RankSVM as both models are evaluated on the majority of datasets included in the comparison for the MAP metric, but given their Normalized Winning Number it can said that both methods are not within the top performing learning to rank methods.
@@
 \includegraphics[scale=0.4235]{gfx/combined_normalized_winnum}
-\caption{Cross-benchmark comparison of learning to rank methods}
+\caption{Cross-benchmark comparison of 87 learning to rank methods}
 \label{fig:normalized_winning_number_all}
@@
 \section{Limitations}
-In the Normalized Winning Number calculation the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
+In the Normalized Winning Number calculation, the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
 By calculating it in this way, we implicitly make the assumption that the learning to rank methods are (approximately) distributed uniformly over the benchmarks, such that the average learning to rank method tested are approximately equally hard for each data set.
-It could be the case however that this assumption is false and that significantly more accurate learning to rank methods are evaluated on some datasets than on other datasets.\\
+It could be the case however that this assumption is false and that the accurateness of the learning to rank methods on a dataset is not dataset independent.\\
 
@@
 
-Another limitation lies in the meta-analysis methodology.
-Using evaluation results published by other researchers relies on the honesty of those researchers.
-Overly optimistic evaluation results published by other researchers do affect our Normalized Winning Number results.
+Another limitation is that our comparison methodology relies on the correctness of the evaluation results found in the literature search step.
+This brings up a risk of overly optimistic evaluation results affecting our Normalized Winning Number results.
 Limiting the meta-analysis to those studies that report comparable results on one of the baseline methods of a benchmark set reduces this limitation but does not solve it completely.
-By taking the Ideal Winning Number into account in Figure \ref{fig:normalized_winning_number_all} we further mitigate this limitation, as the Ideal Winning Number is loosely related with the number of studies that reported evaluation results for an algorithm.
+By taking the Ideal Winning Number into account in Figure \ref{fig:normalized_winning_number_all} we further mitigate this limitation, as the Ideal Winning Number is loosely related with the number of studies that reported evaluation results for an algorithm.\\
+
+Our comparison regarded evaluation results on NDCG@{3,5,10} and MAP.
+By making the decision to include NDCG at three cut-off points and only a single MAP entry, we implicitly attain a higher weight for NDCG compared to MAP on an analysis that combines all measurements on the four metrics.
+This implicit weighting could be regarded as arbitrary, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.
+Note that another implicit weighting lies in the paper dimension.
+Hence, the higher number of evaluation results specified in a paper, the higher the influence of this paper on the outcome of the analysis.
+This implicit weighting is not harmful to the validity of our comparison, as papers with a large number of evaluation results are more valuable than papers with a few evaluation results.
+In addition, papers with a high number of evaluation results are not expected to be less reliable than papers with fewer evaluation results.\\
 
 \section{Contributions}
-We proposed a way to compare learning to rank methods based on sparse evaluation results data on a set of benchmark datasets.
+We proposed a new way of comparing learning to rank methods based on sparse evaluation results data on a set of benchmark datasets.
 Our comparison methodology comprises of two components: 1) Normalized Winning Number, which provides insight in the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.\\
 
-Based on our structured search for evaluation results on well-known benchmarks collections, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.
+Based on our literature search for evaluation results on well-known benchmarks collections, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.
 However, no closing arguments can be formulated on which learning to rank methods are most accurate.
@@
 
-More evaluation runs on are needed for the methods on the left side on the list on benchmark datasets that these methods have not yet been evaluated on for more definite conclusions on which learning to rank methods are most accurate.
-Our work contributes in this sense by identifying promising learning to rank methods that researchers could focus on in performing additional evaluation runs.\\
+More evaluation runs are needed for the methods on the left side of Figure \ref{fig:normalized_winning_number_all}.
+Our work contributes to this by identifying promising learning to rank methods that researchers could focus on in performing additional evaluation runs.\\
 
@@
 % produce the bibliography for the citations in your paper.
-\clearpage
 \bibliographystyle{abbrv}

### 2014-05-21 implementation/java/src/letor/ListNet.java
ListNet gepatched, andere Folds ge-regext, RankLib gedownload

@@ -1,14 +1,9 @@
 package letor;
 
-import com.google.common.collect.Iterables;
 import org.apache.pig.PigServer;
-import org.apache.pig.data.DataBag;
 import org.apache.pig.data.Tuple;
 
-import java.util.ArrayDeque;
-import java.util.ArrayList;
 import java.util.Iterator;
-import java.util.List;
 
 /**
  * MapReduce implementation of the ListNet algorithm
@@ -16,7 +11,7 @@ import java.util.List;
 
 public class ListNet {
     // Initialise hyper-parameters
-    private static final double   STEPSIZE   = 0.0001;
+    private static final double   STEPSIZE   = 0.01;
     private static final int      ITERATIONS = 100;
 
     public static void main(String[] args) throws Exception {
@@ -64,7 +59,7 @@ public class ListNet {
             Tuple lossGradientTuple = LOSS_GRADIENT.next();
             double loss = Double.parseDouble(lossGradientTuple.get(0).toString());
             for(int j=1; j<lossGradientTuple.size(); j++){
-                gradient[j-1] = Double.parseDouble(lossGradientTuple.get(i).toString());
+                gradient[j-1] = Double.parseDouble(lossGradientTuple.get(j).toString());
             }
             System.out.println("gradient: "+toParamString(gradient));
 

### 2014-05-22 cikm_paper/paper.tex
kleinere bibliography
bibliography opgeschoond
proceedings of the fourth conference -> proc 4th conf

@@
 
-\documentclass{sig-alternate-2013}
+\documentclass{sig-alternate}
 \usepackage{longtable}
@@
 \bibliographystyle{abbrv}
-\bibliography{Bibliography}  % sigproc.bib is the name of the Bibliography in this case
+{\small\bibliography{Bibliography}}  % sigproc.bib is the name of the Bibliography in this case
 % You must have a proper ".bib" file
@@
 \cmidrule{2-3} \cmidrule{5-6} \cmidrule{8-9} \cmidrule{11-12} \cmidrule{14-16}
-Method & NWN & \#ds && NWN & \#ds && NWN & \#ds && NWN & \#ds && WM & IWN & NWN \\ \midrule\endhead
+Method & NWN & \#ds && NWN & \#ds && NWN & \#ds && NWN & \#ds && WN & IWN & NWN \\ \midrule\endhead
 AdaRank-MAP & 0.3486 & 12 && 0.3884 & 12 && 0.3648 & 13 && 0.3206 & 12 && 332 & 937 & 0.3543 \\

### 2014-05-23 cikm_paper/paper.tex
squeezed bibliography
10 pagina's :):):)

@@
 \documentclass{sig-alternate}
+\usepackage{booktabs}
+\usepackage{multicol}
 \usepackage{longtable}
-\usepackage{booktabs}
 \usepackage{dblfloatfix}
 \usepackage{fixltx2e}
+\usepackage{enumitem}
+\usepackage{float}
 \renewcommand{\dblfloatpagefraction}{2.0}
@@
 \newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\usepackage{balance}
 \begin{document}
@@
 New learning to rank methods are generally evaluated on benchmark test collections.
-However, comparison of learning to rank methods based on evaluation results is hindered by non-existence of a \emph{de facto} standard set of evaluation benchmark collections.
+However, comparison of learning to rank methods based on evaluation results is hindered by non-existence of a standard set of evaluation benchmark collections.
 In this paper we propose a way to compare learning to rank methods based on a sparse set of evaluation results on a set of benchmark datasets.
@@
 Ranking has become a vital part of web search, where commercial search engines help users find their need in the extremely large collection of the World Wide Web.
-One can find useful applications of ranking in many application domains outside web search as well.
-For example, ranking plays a vital role in automatic text summarization, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}.
-In addition, the ranking task has been argued to be a better fit for recommender systems compared to the regression task (concerning predictions on a continuous scale) \cite{Adomavicius2005,McNee2006}.\\
+Among useful applications outside web search are automatic text summarization, machine translation, drug discovery and determining the ideal order of maintenance operations \cite{Rudin2009}.
+In addition, the ranking task has been argued to be a better fit for recommender systems compared to the regression task (concerning predictions on a continuous scale) \cite{McNee2006}.\\
 
-Research in the field of ranking models has for a long time been based on manually designed ranking functions, such as the well-known BM25 model \cite{Robertson1994}.
-Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
+Research in the field of ranking models has been based on manually designed ranking functions, such as the well-known BM25 model \cite{Robertson1994}.
 Increased amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective ranking models.
@@
 
-In recent years, several learning to rank benchmark datasets have been proposed with the aim of enabling comparison of different learning to rank methods in terms of ranking accuracy.
+In recent years, several learning to rank benchmark datasets have been proposed with the aim of enabling comparison of learning to rank methods in terms of ranking accuracy.
 Well-known benchmark datasets in the learning to rank field include the \emph{Yahoo!
 Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the \emph{Yandex Internet Mathematics 2009} contest\footnote{http://imat2009.yandex.ru/en}, the LETOR datasets \cite{Qin2010}, and the MSLR (Microsoft Learning to Rank) datasets\footnote{http://research.microsoft.com/en-us/projects/mslr/}.
-There exists no agreement among authors in the learning to rank field on which benchmark collection(s) to use to evaluate a new model.
-Comparing ranking accuracy of learning to rank methods is largely hindered by this lack of a \emph{de facto} standard way of benchmarking new models.\\
+There exists no agreement among authors in the learning to rank field on the benchmark collection(s) to use to evaluate a new model.
+Comparing ranking accuracy of learning to rank methods is largely hindered by this lack of a standard way of benchmarking.\\
 
@@
 We collect evaluation results on the datasets of benchmark collections through a structured literature search.
-Table \ref{tbl:ltr_benchmark_collections} presents an overview of the benchmark collections taken into account in the meta-analysis:
+Table \ref{tbl:ltr_benchmark_collections} presents an overview of the benchmark collections included in the meta-analysis.
 
 \begin{table}[!h]
+\scriptsize{
 \begin{tabular}{p{6.4cm}R{1.26cm}}\toprule
@@
 \caption{Included learning to rank evaluation benchmark collections}
-\label{tbl:ltr_benchmark_collections}
+\label{tbl:ltr_benchmark_collections}}
 \end{table}
@@
 LETOR 1.0 and 3.0, Yahoo!
-Learning to Rank Challenge, WCL2R and AOL have accompanying papers which were published together with these benchmark collections.
+Learning to Rank Challenge, WCL2R and AOL have accompanying papers that were released with the collection.
 Authors publishing evaluation results on these benchmark collections are requested to cite these papers.
-Therefore, we collect evaluation measurements of learning to rank methods on these benchmark collections through forward literature search.
-Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search.
-Google Scholar is used to perform the forward reference search.\\
-
+We collect evaluation measurements of learning to rank methods on these benchmark collections through forward literature search.
+Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search performed using Google Scholar.
 \begin{table}[!h]
+\scriptsize{
 \begin{tabular}{p{2.8cm}p{2.64cm}R{1.9cm}}\toprule
@@
 \caption{Forward references of learning to rank benchmark papers}
-\label{tbl:ltr_benchmark_forref}
+\label{tbl:ltr_benchmark_forref}}
 \end{table}
@@
 \begin{table}[!h]
+\scriptsize{
 \begin{tabular}{p{5.2cm}R{2.5cm}}\toprule
@@
 "MSLR-web30k" & 15 \\
-"Yandex Internet Mathematics" & 3 \\
+"Yandex Internet Mathematics" & 3\\
 \bottomrule
@@
 \caption{Google scholar search results for learning to rank benchmarks}
-\label{tbl:ltr_benchmark_searches}
+\label{tbl:ltr_benchmark_searches}}
 \end{table}
@@
 \subsection{Literature Selection}
-Table \ref{tab:ltr_methods_used} in Appendix \ref{app:ltr_methods_used} gives an overview of the learning to rank methods for which evaluation results were found through the described procedure.
+Table \ref{tab:ltr_methods_used} in the appendix gives an overview of the learning to rank methods for which evaluation results were found through the described procedure.
 Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, 3.0 and 4.0 baselines respectively.\\
@@
 
-\begin{enumerate}
+\begin{enumerate}[noitemsep,nolistsep]
 \item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011, Lin2012}
@@
 
-\begin{figure*}
-\includegraphics[scale=0.371]{gfx/ndcg3_winnum}
+\begin{figure}
+\centering
+\includegraphics[scale=0.19]{gfx/ndcg3_winnum}
 \caption{NDCG@3 comparison of 87 learning to rank methods}
 \label{fig:normalized_winning_number_ndcg3}
-\end{figure*}
+\end{figure}
 \section{A Methodology for Comparing Learning to Rank Methods Cross-Benchmark}
@@
 Winning Number is defined as the number of other algorithms that an algorithm can beat over the set of datasets, or more formally\\
-
 $\text{WN}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$\\
-
 where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as NDCG or MAP), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that\\
-
 $I_{\{M_i(j)>M_k(j)\}} = \begin{cases}
@@
 \end{cases}$\\
-
 The LETOR 3.0 was a comparison on a \emph{dense} set of evaluation results, in the sense that there were evaluation results available for all learning to rank algorithms on all datasets included in their comparison.
@@
 This Normalized Winning Number takes only those datasets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all datasets on which it has been evaluated.
-We will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as
-
-\[
-I^{'}_{M_i(j)>M_k(j)} = \begin{cases}
+We will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as\\
+$I^{'}_{M_i(j)>M_k(j)} = \begin{cases}
 1 & \parbox[t]{5cm}{if $M_i(j)$ and $M_k(j)$ are both defined and $M_i(j) > M_k(j)$,}\\
 0 & \parbox[t]{5cm}{otherwise}
-\end{cases}
-\]
-From now on this adjusted version of Winning Number will be references to as \emph{Normalized Winning Number}. The mathematical definition of Normalized Winning Number is\\
-
+\end{cases}$\\
+From now on this adjusted version of Winning Number will be references to as \emph{Normalized Winning Number (NWN)}. The mathematical definition of Normalized Winning Number is\\
 $\text{NWN}_i(M) = \frac{\text{WN}_i(M)}{\text{IWN}_i(M)}$\\
-
 \noindent
 where $IWN$ is the Ideal Winning Number, defined as\\
-
 $\text{IWN}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m D_{\{M_i(j),M_k(j)\}}$\\
-
 where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as NDCG or MAP), and $D_{\{M_i(j),M_k(j)\}}$ is an evaluation definition function such that\\
-
 $D_{\{M_i(j),M_k(j)\}} = \begin{cases}
@@
 \end{cases}$\\
+NDCG@$\{3,5,10\}$ and MAP are the most frequently used evaluation metrics in the used benchmark collections combined, therefore we will limit our meta-analysis to evaluation results reported in one of these four metrics.
 
-NDCG@$\{3,5,10\}$ and MAP are the most frequently used evaluation metrics in the used benchmark collections, therefore we will limit our meta-analysis to evaluation results reported in one of these four metrics.
-The Yahoo!
-Learning to Rank Challenge benchmark is an exception, on this benchmark ERR is the most used evaluation metric.
-The lack of use of the ERR metric in other benchmarks makes it unsuitable for a cross-benchmark comparison.\\
-
-\begin{figure*}
-\includegraphics[scale=0.371]{gfx/ndcg5_winnum}
+\begin{figure}
+\includegraphics[scale=0.19]{gfx/ndcg5_winnum}
 \caption{NDCG@5 comparison of 87 learning to rank methods}
 \label{fig:normalized_winning_number_ndcg5}
-\end{figure*}
+\end{figure}
 \section{Results of Learning to Rank Comparison}
-The following subsections provide the performance of learning to rank methods in terms of Normalized Winning Number for NDCG@$\{3,5,10\}$ and MAP.
-Performance of the learning to rank methods is plotted with Normalized Winning Number on the vertical axis and the number of datasets on which the method has been evaluated on the horizontal axis.
-The further to the right, the more certain we can be about the performance of the learning to rank method.
-The methods for which it holds that there is no other method that has 1) a higher Normalized Winning Number and 2) a higher number datasets evaluated, are identified as the best performing methods and are labeled with the name of the method.
-Table \ref{tab:raw_data} in Appendix \ref{app:raw_data} provides the raw Normalized Winning Number data for the learning to rank methods at NDCG@$\{3,5,10\}$ and MAP and their cross-metric weighted average.
+The following subsections provide the performance of learning to rank methods in terms of NWN for NDCG@$\{3,5,10\}$ and MAP.
+Performance of the learning to rank methods is plotted with NWN on the vertical axis and the number of datasets on which the method has been evaluated on the horizontal axis.
+Moving to the right, certainty on the performance of the method increases.
+The methods for which it holds that there is no other method that has 1) a higher NWN and 2) a higher number datasets evaluated, are identified as the best performing methods and are labeled.
+Table \ref{tab:raw_data} in the appendix provides raw NWN data for the learning to rank methods at NDCG@$\{3,5,10\}$ and MAP and their cross-metric weighted average.
 
 \subsection{NDCG@3}
-Figure \ref{fig:normalized_winning_number_ndcg3} shows the performance of learning to rank methods for the NDCG@3 metric.
-LambdaNeuralRank and CoList both acquired a perfect Normalized Winning Number score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
-LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three datasets which in addition are more relevant datasets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 datasets (see Table \ref{tab:ltr_methods_used} in Appendix \ref{app:ltr_methods_used}).
-FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in Normalized Winning Number, but increasingly higher in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
+Figure \ref{fig:normalized_winning_number_ndcg3} shows the NWN of learning to rank methods based on NDCG@3 results.
+LambdaNeuralRank and CoList both acquired a NWN score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
+LARF and LRUF scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which results in more certainty on these methods' performance because they are validated on three datasets that additionally are more relevant than AOL and Yahoo Set 2 (number of evaluation results for LETOR 3.0 are higher than those for AOL and Yahoo set 2).
+FenchelRank, OWPC, SmoothRank, DCMP and ListNet are ordered decreasingly by NWN and at the same time increasingly in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
 
@@
 \subsection{NDCG@5}
-Figure \ref{fig:normalized_winning_number_ndcg5} shows the performance of learning to rank methods for the NDCG@5 metric.
-LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a Normalized Winning Number of 1.0.
-LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of datasets, but score decreasingly well in terms of Normalized Winning Number.
+Figure \ref{fig:normalized_winning_number_ndcg5} shows the NWN of learning to rank methods based on NDCG@5 results.
+LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a NWN of 1.0.
+LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of datasets, but score decreasingly well in terms of NWN.
 These results are highly in agreement with the NDCG@3 comparison.
@@
 \subsection{NDCG@10}
-\begin{figure*}
-\includegraphics[scale=0.371]{gfx/ndcg10_winnum}
+\begin{figure}
+\centering
+\includegraphics[scale=0.19]{gfx/ndcg10_winnum}
 \caption{NDCG@10 comparison of 87 learning to rank methods}
 \label{fig:normalized_winning_number_ndcg10}
-\end{figure*}
+\end{figure}
 
-Figure \ref{fig:normalized_winning_number_ndcg10} shows the performance of learning to rank methods for the NDCG@10 metric.
-LambdaMART and LambdaNeuralRank score a Normalized Winning Number of 1.0 on the NDCG@10 comparison.
+Figure \ref{fig:normalized_winning_number_ndcg10} shows the NWN of learning to rank methods based on NDCG@10 results.
+LambdaMART and LambdaNeuralRank score a NWN of 1.0 on the NDCG@10 comparison.
 For LambdaNeuralRank these results are again based on AOL dataset measurements.
 LambdaMART showed the highest NDCG@10 performance for the MSLR-WEB10k dataset.
-The set of algorithms for which there is no other algorithm with both a higher Normalized Winning Number and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and @5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
+The set of algorithms for which there is no other algorithm with both a higher NWN and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and @5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
 In contrast to the NDCG@3 and @5 comparison, DCMP is not among the best algorithms.
@@
 \subsection{MAP}
-\begin{figure*}
-\includegraphics[scale=0.371]{gfx/map_winnum}
+\begin{figure}
+\centering
+\includegraphics[scale=0.19]{gfx/map_winnum}
 \caption{MAP comparison of 87 learning to rank methods}
 \label{fig:normalized_winning_number_map}
-\end{figure*}
-Figure \ref{fig:normalized_winning_number_map} shows the performance of learning to rank methods for the MAP metric.
-Where comparisons on the NDCG metric at different cut-off points where highly in agreement in terms of the best performing algorithms, the comparison in terms of MAP shows different results.
-RankDE scores a Normalized Winning Number of 1.0 on one dataset, like LambdaNeuralRank did on for all NDCG-comparisons.
-In contrast to LambdaNeuralRank, RankDE achieved this highest score on the LETOR 2.0 TD2003, a dataset on which many methods are evaluated.\\
+\end{figure}
+Figure \ref{fig:normalized_winning_number_map} shows the NWN of learning to rank methods based on MAP results.
+Comparisons on the NDCG metrics where highly in agreement on the best performing algorithms, MAP-based NWN results show different results.
+RankDE scores a NWN of 1.0 on one dataset, which is achieved by obtaining highest MAP-score on the LETOR 2.0 TD2003 which has many evaluation results are evaluated.\\
 
-LARF and LRUF score very high Normalized Winning Number scores, but based on only relatively few datasets, just as in the NDCG comparisons.
-Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all NDCG comparisons, are not among the best performing methods in the MAP comparison.
-A deeper look in the raw data Table \ref{tab:raw_data} in the Appendix \ref{app:raw_data} shows that LAC-MR-OR is evaluated on many more datasets for MAP compared to NDCG, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalized Winning Number.
-SmoothRank performed a Normalized Winning Number of around 0.53 over 7 datasets, which is still good in both certainty and accuracy, but not among the top methods.
+LARF and LRUF score very high NWN scores, but based on only few datasets, just as in the NDCG-based comparisons.
+Notable is that low performance of SmoothRank and ListNet, given that those methods were top performing methods in the NDCG-based comparisons.
+Table \ref{tab:raw_data} in the appendix shows that LAC-MR-OR is evaluated on more datasets on MAP than on NDCG, thereby LAC-MR-OR obtained equal certainty to ListNet with a higher NWN.
+SmoothRank performed a NWN of around 0.53 over 7 datasets, which is still good in both certainty and accuracy, but not among the top methods.
 RE-QR is one of the best performers in the MAP comparison with a reasonable amount of benchmark evaluations.
 No reported NDCG performance was found in the literature search for RE-QR.
-There is a lot of certainty on the accuracy of RankBoost and RankSVM as both models are evaluated on the majority of datasets included in the comparison for the MAP metric, but given their Normalized Winning Number it can said that both methods are not within the top performing learning to rank methods.
+There is a lot of certainty on the accuracy of RankBoost and RankSVM as both models are evaluated on the majority of datasets included in the comparison for the MAP metric, but given their NWN it can said that both methods are not within the top performing learning to rank methods.
 
 \subsection{Cross-Metric}
-\begin{figure*}
-\includegraphics[scale=0.4235]{gfx/combined_normalized_winnum}
+\begin{figure}
+\centering
+\includegraphics[scale=0.21]{gfx/combined_normalized_winnum}
 \caption{Cross-benchmark comparison of 87 learning to rank methods}
 \label{fig:normalized_winning_number_all}
-\end{figure*}
-Figure \ref{fig:normalized_winning_number_all} shows the Normalized Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}.
+\end{figure}
+Figure \ref{fig:normalized_winning_number_all} shows the NWN as function of IWN for the methods described in Table \ref{tab:ltr_methods_used}.
 The cross-metric comparison is based on the NDCG@$\{3,5,10\}$ and MAP comparisons combined, which justifies analyzing the comparison more thoroughly.
@@
 
-LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both Ideal Winning Number and Normalized Winning Number.
+LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both IWN and NWN.
 LRUF is the only method that achieved this in all NDCG comparisons, the MAP comparison as well as the cross-metric comparison.
@@
 LARF, IPRank and DCMP and several variants of RankSVM performed very well on the cross-metric comparison, with all having only one method in its top right quadrant.
-LARF also performed among the top methods on the NDCG and MAP comparisons and DCMP was a top performer in a few of the NDCG comparisons.\\
-
-C-CRF, DirectRank, FP-Rank, RankCSA, LambdaNeuralRank and VFLR all have near-perfect Normalized Winning Number measures, but have low Ideal Winning Number measures.
+LARF also performed among the top methods on the NDCG and MAP comparisons and DCMP was a top performer in a few of the NDCG comparisons.
+C-CRF, DirectRank, FP-Rank, RankCSA, LambdaNeuralRank and VFLR all have a near-perfect NWN value, but have a low IWN value.
 Further evaluation runs of these methods on benchmark datasets that they are not yet evaluated on are desirable.
@@
 \section{Limitations}
-In the Normalized Winning Number calculation, the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
+In the NWN calculation, the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
 By calculating it in this way, we implicitly make the assumption that the learning to rank methods are (approximately) distributed uniformly over the benchmarks, such that the average learning to rank method tested are approximately equally hard for each data set.
@@
 Another limitation is that our comparison methodology relies on the correctness of the evaluation results found in the literature search step.
-This brings up a risk of overly optimistic evaluation results affecting our Normalized Winning Number results.
+This brings up a risk of overly optimistic evaluation results affecting our NWN results.
 Limiting the meta-analysis to those studies that report comparable results on one of the baseline methods of a benchmark set reduces this limitation but does not solve it completely.
-By taking the Ideal Winning Number into account in Figure \ref{fig:normalized_winning_number_all} we further mitigate this limitation, as the Ideal Winning Number is loosely related with the number of studies that reported evaluation results for an algorithm.\\
+By taking IWN into account in Figure \ref{fig:normalized_winning_number_all} we further mitigate this limitation, as IWN is loosely related with the number of studies that reported evaluation results for an algorithm.\\
 
-Our comparison regarded evaluation results on NDCG@{3,5,10} and MAP.
+Our comparison regarded evaluation results on NDCG@$\{3,5,10\}$ and MAP.
 By making the decision to include NDCG at three cut-off points and only a single MAP entry, we implicitly attain a higher weight for NDCG compared to MAP on an analysis that combines all measurements on the four metrics.
@@
 This implicit weighting is not harmful to the validity of our comparison, as papers with a large number of evaluation results are more valuable than papers with a few evaluation results.
-In addition, papers with a high number of evaluation results are not expected to be less reliable than papers with fewer evaluation results.\\
+In addition, papers with a high number of evaluation results are not expected to be less reliable than papers with fewer evaluation results.
 
@@
 We proposed a new way of comparing learning to rank methods based on sparse evaluation results data on a set of benchmark datasets.
-Our comparison methodology comprises of two components: 1) Normalized Winning Number, which provides insight in the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.\\
+Our comparison methodology comprises of two components: 1) NWN, which provides insight in the ranking accuracy of the learning to rank method, and 2) IWN, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.\\
 
@@
 More evaluation runs are needed for the methods on the left side of Figure \ref{fig:normalized_winning_number_all}.
-Our work contributes to this by identifying promising learning to rank methods that researchers could focus on in performing additional evaluation runs.\\
-
+Our work contributes to this by identifying promising learning to rank methods that researchers could focus on in performing additional evaluation runs.
 %
@@
 % produce the bibliography for the citations in your paper.
+\balance
 \bibliographystyle{abbrv}
-{\small\bibliography{Bibliography}}  % sigproc.bib is the name of the Bibliography in this case
+{\scriptsize\bibliography{Bibliography}}
+\onecolumn
+% sigproc.bib is the name of the Bibliography in this case
 % You must have a proper ".bib" file
@@
 %APPENDICES are optional
-%\balancecolumns
-\onecolumn
-\clearpage
-\newpage
 \appendix
-\section{Learning to Rank Methods with Evaluation Results}
+
 \label{app:ltr_methods_used}
-\begin{table*}[!hp]
+\label{app:raw_data}
+\begin{table*}[!hpt]
+\scriptsize{
 \begin{tabular}{lll|lll}\toprule
@@
 \end{tabular}
-\caption{Learning to rank algorithms with measurements on benchmark datasets}
+\caption{Learning to rank algorithms with measurements on benchmark datasets}}
 \label{tab:ltr_methods_used}
@@
 
-\clearpage
-\section{Cross-Comparison Raw Data}
-\label{app:raw_data}
-\begin{longtable}[!hp]{@{}lrrrrrrrrrrrrrrrr@{}}\toprule
+\scriptsize{
+\begin{longtable}[!hpt]{@{}lrrrrrrrrrrrrrrrr@{}}\toprule
 & \multicolumn{2}{c}{NDCG@3} & \phantom{a} 
@@
 \end{longtable}
+}
+
 \end{document}

### 2014-05-26 cikm_paper/paper.tex
laatste puntjes op de i CIKM

@@
 Among useful applications outside web search are automatic text summarization, machine translation, drug discovery and determining the ideal order of maintenance operations \cite{Rudin2009}.
-In addition, the ranking task has been argued to be a better fit for recommender systems compared to the regression task (concerning predictions on a continuous scale) \cite{McNee2006}.\\
+In addition, the ranking task has been argued to be a better fit for recommender systems than the regression task \cite{McNee2006} (which concerns predictions on a continuous scale) that is currently frequently used in such systems.\\
 
@@
 Well-known benchmark datasets in the learning to rank field include the \emph{Yahoo!
-Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the \emph{Yandex Internet Mathematics 2009} contest\footnote{http://imat2009.yandex.ru/en}, the LETOR datasets \cite{Qin2010}, and the MSLR (Microsoft Learning to Rank) datasets\footnote{http://research.microsoft.com/en-us/projects/mslr/}.
+Learning to Rank Challenge} datasets \cite{Chapelle2011a}, the \emph{Yandex Internet Mathematics 2009} contest\footnote{http://imat2009.yandex.ru/en}, the LETOR datasets \cite{Qin2010}, and the MSLR (Microsoft Learning to Rank) datasets\footnote{http://research.microsoft.com/en-us/projects/mslr/}.
 There exists no agreement among authors in the learning to rank field on the benchmark collection(s) to use to evaluate a new model.
@@
 Winning Number is defined as the number of other algorithms that an algorithm can beat over the set of datasets, or more formally\\
+
 $\text{WN}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$\\
+
 where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as NDCG or MAP), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that\\
+
 $I_{\{M_i(j)>M_k(j)\}} = \begin{cases}
@@
 \end{cases}$\\
+
 The LETOR 3.0 was a comparison on a \emph{dense} set of evaluation results, in the sense that there were evaluation results available for all learning to rank algorithms on all datasets included in their comparison.
@@
 We will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as\\
+
 $I^{'}_{M_i(j)>M_k(j)} = \begin{cases}
@@
 \end{cases}$\\
+
 From now on this adjusted version of Winning Number will be references to as \emph{Normalized Winning Number (NWN)}. The mathematical definition of Normalized Winning Number is\\
+
 $\text{NWN}_i(M) = \frac{\text{WN}_i(M)}{\text{IWN}_i(M)}$\\
+
 \noindent
 where $IWN$ is the Ideal Winning Number, defined as\\
+
 $\text{IWN}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m D_{\{M_i(j),M_k(j)\}}$\\
+
 where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as NDCG or MAP), and $D_{\{M_i(j),M_k(j)\}}$ is an evaluation definition function such that\\
+
 $D_{\{M_i(j),M_k(j)\}} = \begin{cases}
@@
 \end{cases}$\\
+
 NDCG@$\{3,5,10\}$ and MAP are the most frequently used evaluation metrics in the used benchmark collections combined, therefore we will limit our meta-analysis to evaluation results reported in one of these four metrics.
@@
 \midrule
-AdaRank-MAP & \cite{Xu2007} & L2, L3, L4 & Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} \\ 
-AdaRank-NDCG & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & ListMLE & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
-ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet & \cite{Cao2007} & L2, L3, L4 \\ 
-ApproxAP & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
-ApproxNDCG & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
-BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
-Best Single Feature &  & \cite{Gomes2013} & MHR & \cite{Qin2007} & L2 \\ 
-BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} & MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
-BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009, Volkovs2013} & NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
-BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c, Volkovs2013} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
-BT & \cite{Zhou2008} & \cite{Zhou2008} & PERF-MAP & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
-C-CRF & \cite{Qin2008b} & \cite{Qin2008b} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
-CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & Q.D.KNN & \cite{Geng2008} & \cite{Wang2013} \\ 
-CCRank & \cite{Wang2011c} & \cite{Wang2011c} & RandomForest &  & \cite{Gomes2013} \\ 
-CoList & \cite{Gao2014} & \cite{Gao2014} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
-Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAggNDCG & \cite{Wang2013} & \cite{Wang2013} \\ 
-DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010} \\ 
-DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
+AdaRank-MAP & \cite{Xu2007} & L2, L3, L4 & ListMLE & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\
+AdaRank-NDCG & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & ListNet & \cite{Cao2007} & L2, L3, L4 \\ 
+ADMM & \cite{Duh2011} & \cite{Duh2011} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
+ApproxAP & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
+ApproxNDCG & \cite{Qin2010b} & \cite{Qin2010b} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
+BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MHR & \cite{Qin2007} & L2 \\ 
+Best Single Feature &  & \cite{Gomes2013} &  MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
+BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} &  NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
+BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009, Volkovs2013} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
+BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c, Volkovs2013} & PERF-MAP & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
+BT & \cite{Zhou2008} & \cite{Zhou2008} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
+C-CRF & \cite{Qin2008b} & \cite{Qin2008b} & Q.D.KNN & \cite{Geng2008} & \cite{Wang2013} \\ 
+CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & RandomForest &  & \cite{Gomes2013} \\ 
+CCRank & \cite{Wang2011c} & \cite{Wang2011c} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
+CoList & \cite{Gao2014} & \cite{Gao2014} & RankAggNDCG & \cite{Wang2013} & \cite{Wang2013} \\
+Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010} \\
+DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost (Kernel-PCA) & \cite{Duh2008} & \cite{Duh2008, Sato2013} \\
+DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankBoost (SVD) & \cite{Lin2009} & \cite{Lin2009} \\
 EnergyNDCG & \cite{Freno2011} & \cite{Freno2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\
-FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
-FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b, Laporte2013} & RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
-FocusedBoost & \cite{Niu2012} & \cite{Niu2012} & RankMGP & \cite{Lin2012} & \cite{Lin2012} \\ 
-FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
-FocusedSVM & \cite{Niu2012} & \cite{Niu2012} & RankRLS & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
-FP-Rank & \cite{Song2013} & \cite{Song2013} & RankSVM & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010} \\ 
-FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & RankSVM-Primal &  & L3, \cite{Lai2011} \\ 
-FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & RankSVM-Struct &  & L3, L4 \\
-FSM$^{SVM}$ & \cite{Lai2013c} & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
-GAS-E & \cite{Geng2007} & \cite{Lai2013c} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\
-GP & \cite{DeAlmeida2007} & \cite{Alcantara2010} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\  
-GPRank & \cite{Silva2009} & \cite{Torkestani2012} & Ridge Regression & \cite{Cossock2006} & L3 \\
-GRankRLS & \cite{Pahikkala2010} & \cite{Pahikkala2010} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
-GroupCE & \cite{Lin2011} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
-GroupMLE & \cite{Lin2010} & \cite{Lin2011} & SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010} \\
-IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
-IPRank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\
-KeepRank & \cite{Chen2009} & \cite{Chen2009} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
-Kernel-PCA RankBoost & \cite{Duh2008} & \cite{Duh2008, Sato2013} & SVD-RankBoost & \cite{Lin2009} & \cite{Lin2009} \\
-KL-CRF & \cite{Volkovs2011} & \cite{Volkovs2011} & SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
-LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
-LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} & TGRank & \cite{Lai2013} & \cite{Lai2013} \\ 
-LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
-LambdaRank & \cite{Burges2006} &  & VFLR & \cite{Cai2012} & \cite{Cai2012} \\ 
-LARF & \cite{Torkestani2012} & \cite{Torkestani2012} &  &  &  \\
+FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankCSA & \cite{He2010} & \cite{He2010} \\
+FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b, Laporte2013} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FocusedBoost & \cite{Niu2012} & \cite{Niu2012} &  RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankMGP & \cite{Lin2012} & \cite{Lin2012} \\ 
+FocusedSVM & \cite{Niu2012} & \cite{Niu2012} & RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
+FP-Rank & \cite{Song2013} & \cite{Song2013} & RankRLS & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
+FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & RankSVM & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010} \\ 
+FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & RankSVM-Primal &  & L3, \cite{Lai2011} \\ 
+FSM$^{SVM}$ & \cite{Lai2013c} & \cite{Lai2013c} & RankSVM-Struct &  & L3, L4 \\
+GAS-E & \cite{Geng2007} & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
+GP & \cite{DeAlmeida2007} & \cite{Alcantara2010} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\
+GPRank & \cite{Silva2009} & \cite{Torkestani2012} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\  
+GRankRLS & \cite{Pahikkala2010} & \cite{Pahikkala2010} & Ridge Regression & \cite{Cossock2006} & L3 \\
+GroupCE & \cite{Lin2011} & \cite{Lin2011} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
+GroupMLE & \cite{Lin2010} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
+IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010} \\
+IPRank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
+KeepRank & \cite{Chen2009} & \cite{Chen2009} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\
+KL-CRF & \cite{Volkovs2011} & \cite{Volkovs2011} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
+LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
+LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} &SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
+LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank & \cite{Lai2013} & \cite{Lai2013} \\ 
+LambdaRank & \cite{Burges2006} &  & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
+LARF & \cite{Torkestani2012} & \cite{Torkestani2012} &  VFLR & \cite{Cai2012} & \cite{Cai2012} \\
+Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} & & & \\
 \bottomrule

### 2014-05-26 cikm_paper/paper.tex
volgorde klopt weer :)

@@
 DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankBoost (SVD) & \cite{Lin2009} & \cite{Lin2009} \\
-EnergyNDCG & \cite{Freno2011} & \cite{Freno2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\
-FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankCSA & \cite{He2010} & \cite{He2010} \\
+EnergyNDCG & \cite{Freno2011} & \cite{Freno2011} & RankCSA & \cite{He2010} & \cite{He2010} \\
+FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\
 FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b, Laporte2013} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 

### 2014-05-26 cikm_paper/paper.tex
git correcties tabel

@@
 C-CRF & \cite{Qin2008b} & \cite{Qin2008b} & Q.D.KNN & \cite{Geng2008} & \cite{Wang2013} \\ 
-CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & RandomForest &  & \cite{Gomes2013} \\ 
+CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & RandomForest & \cite{Gomes2013} & \cite{Gomes2013} \\ 
 CCRank & \cite{Wang2011c} & \cite{Wang2011c} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
 CoList & \cite{Gao2014} & \cite{Gao2014} & RankAggNDCG & \cite{Wang2013} & \cite{Wang2013} \\
-Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010} \\
+Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010, Sato2013} \\
 DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost (Kernel-PCA) & \cite{Duh2008} & \cite{Duh2008, Sato2013} \\
@@
 FP-Rank & \cite{Song2013} & \cite{Song2013} & RankRLS & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
-FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & RankSVM & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010} \\ 
+FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & RankSVM & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010, Papini2012} \\ 
 FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & RankSVM-Primal &  & L3, \cite{Lai2011} \\ 
@@
 IPRank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
-KeepRank & \cite{Chen2009} & \cite{Chen2009} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011} \\
+KeepRank & \cite{Chen2009} & \cite{Chen2009} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011, Papini2012} \\
 KL-CRF & \cite{Volkovs2011} & \cite{Volkovs2011} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
-LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008} & SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
+LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008, Alcantara2010} & SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
 LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} &SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
 LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank & \cite{Lai2013} & \cite{Lai2013} \\ 
-LambdaRank & \cite{Burges2006} &  & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
+LambdaRank & \cite{Burges2006} & \cite{Papini2012, Tan2013} & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
 LARF & \cite{Torkestani2012} & \cite{Torkestani2012} &  VFLR & \cite{Cai2012} & \cite{Cai2012} \\

### 2014-05-27 cikm_paper/paper.tex
gecorrigeerde waarden
paper mooi

@@
 \subsection{Literature Selection}
-Table \ref{tab:ltr_methods_used} in the appendix gives an overview of the learning to rank methods for which evaluation results were found through the described procedure.
-Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, 3.0 and 4.0 baselines respectively.\\
+Table 4 in the appendix gives an overview of the learning to rank methods for which evaluation results were found through the described procedure.
+Occurrences of L2, L3 and L4 in Table 4 imply that these algorithms are evaluated as official LETOR 2.0, 3.0 and 4.0 baselines respectively.\\
 
@@
 \end{figure}
-Figure \ref{fig:normalized_winning_number_all} shows the NWN as function of IWN for the methods described in Table \ref{tab:ltr_methods_used}.
+Figure \ref{fig:normalized_winning_number_all} shows the NWN as function of IWN for the methods described in Table 4.
 The cross-metric comparison is based on the NDCG@$\{3,5,10\}$ and MAP comparisons combined, which justifies analyzing the comparison more thoroughly.
@@
 
-\label{app:ltr_methods_used}
-\label{app:raw_data}
 \begin{table*}[!hpt]
@@
 Method & NWN & \#ds && NWN & \#ds && NWN & \#ds && NWN & \#ds && WN & IWN & NWN \\ \midrule\endhead
-AdaRank-MAP & 0.3486 & 12 && 0.3884 & 12 && 0.3648 & 13 && 0.3206 & 12 && 332 & 937 & 0.3543 \\
-AdaRank-NDCG & 0.3073 & 12 && 0.3259 & 12 && 0.3158 & 16 && 0.2863 & 12 && 293 & 951 & 0.3081 \\
+AdaRank-MAP & 0.3529 & 12 && 0.3884 & 12 && 0.3648 & 13 && 0.3206 & 12 && 334 & 940 & 0.3553 \\
+AdaRank-NDCG & 0.3122 & 12 && 0.3259 & 12 && 0.3158 & 16 && 0.2863 & 12 && 295 & 954 & 0.3092 \\
 ADMM & - & - && - & - && 0.4444 & 1 && - & - && 4 & 9 & 0.4444 \\
 ApproxAP & - & - && - & - && - & - && 0.5000 & 2 && 33 & 66 & 0.5000 \\
-ApproxNDCG & 0.7949 & 1 && 0.7500 & 1 && 0.8611 & 1 && - & - && 92 & 115 & 0.8000 \\
-BagBoo & 0.8478 & 2 && 0.8400 & 1 && - & - && 0.6545 & 2 && 96 & 126 & 0.7619 \\
+ApproxNDCG & 0.8000 & 1 && 0.7500 & 1 && 0.8611 & 1 && - & - && 93 & 116 & 0.8017 \\
+BagBoo & 0.8333 & 2 && 0.8400 & 1 && - & - && 0.6545 & 2 && 97 & 128 & 0.7578 \\
 Best Single Feature & - & - && - & - && 0.1615 & 8 && - & - && 26 & 161 & 0.1615 \\
-BL-MART & 0.9524 & 2 && 0.7200 & 1 && - & - && 0.8036 & 3 && 83 & 102 & 0.8137 \\
-BoltzRank-Pair & 0.8333 & 4 && 0.8350 & 4 && - & - && 0.5804 & 5 && 254 & 348 & 0.7299 \\
-BoltzRank-Single & 0.7549 & 4 && 0.7184 & 4 && - & - && 0.4336 & 5 && 213 & 348 & 0.6121 \\
+BL-MART & 0.8776 & 3 && 0.7200 & 1 && - & - && 0.8036 & 3 && 106 & 130 & 0.8154 \\
+BoltzRank-Pair & 0.8286 & 4 && 0.8350 & 4 && - & - && 0.5804 & 5 && 256 & 351 & 0.7293 \\
+BoltzRank-Single & 0.7524 & 4 && 0.7184 & 4 && - & - && 0.4336 & 5 && 215 & 351 & 0.6125 \\
 BT & 0.7273 & 3 && 0.7879 & 3 && - & - && 0.7500 & 3 && 75 & 99 & 0.7576 \\
@@
 CCRank & - & - && - & - && - & - && 0.6154 & 2 && 24 & 39 & 0.6154 \\
-CoList & 1.0000 & 1 && - & - && 0.1667 & 1 && - & - && 2 & 7 & 0.2857 \\
+CoList & 1.0000 & 1 && 1.0000 & 1 && 0.1667 & 1 && - & - && 3 & 8 & 0.3750 \\
 Consistent-RankCosine & - & - && - & - && 0.7692 & 2 && - & - && 10 & 13 & 0.7692 \\
-DCMP & 0.5459 & 9 && 0.5079 & 9 && 0.5888 & 9 && - & - && 320 & 584 & 0.5479 \\
+DCMP & 0.5477 & 9 && 0.5079 & 9 && 0.5888 & 9 && - & - && 322 & 587 & 0.5486 \\
 DirectRank & - & - && - & - && 0.9231 & 2 && - & - && 12 & 13 & 0.9231 \\
-EnergyNDCG & 0.3636 & 2 && 0.3778 & 2 && 0.4146 & 2 && - & - && 50 & 130 & 0.3846 \\
-FBPCRank & 0.4146 & 3 && 0.5529 & 3 && - & - && - & - && 81 & 167 & 0.4850 \\
-FenchelRank & 0.7742 & 5 && 0.7500 & 5 && 0.7623 & 5 && 0.6418 & 5 && 368 & 504 & 0.7302 \\
-FocusedBoost & 0.3750 & 2 && 0.4545 & 2 && 0.6863 & 2 && - & - && 73 & 143 & 0.5105 \\
+EnergyNDCG & 0.3778 & 2 && 0.3778 & 2 && 0.4146 & 2 && - & - && 51 & 131 & 0.3893 \\
+FBPCRank & 0.4235 & 3 && 0.5529 & 3 && - & - && - & - && 83 & 170 & 0.4882 \\
+FenchelRank & 0.7760 & 5 && 0.7500 & 5 && 0.7623 & 5 && 0.6418 & 5 && 369 & 505 & 0.7307 \\
+FocusedBoost & 0.3753 & 2 && 0.4545 & 2 && 0.6863 & 2 && - & - && 73 & 143 & 0.5105 \\
 FocusedNet & 0.4583 & 2 && 0.6364 & 2 && 0.8627 & 2 && - & - && 94 & 143 & 0.6573 \\
-FocusedSVM & 0.2500 & 2 && 0.2727 & 2 && 0.6078 & 2 && - & - && 55 & 143 & 0.3846 \\
+FocusedSVM & 0.2371 & 2 && 0.2727 & 2 && 0.6078 & 2 && - & - && 55 & 143 & 0.3846 \\
 FP-Rank & - & - && 0.9000 & 1 && - & - && - & - && 18 & 20 & 0.9000 \\
-FRank & 0.3085 & 11 && 0.2849 & 10 && 0.3029 & 11 && 0.2623 & 11 && 242 & 839 & 0.2884 \\
-FSMRank & 0.8333 & 4 && 0.8776 & 4 && 0.8621 & 5 && 0.5789 & 7 && 365 & 481 & 0.7588 \\
-FSM$^{SVM}$ & 0.2292 & 2 && 0.4082 & 4 && 0.5426 & 4 && 0.3500 & 4 && 148 & 388 & 0.3814 \\
-GAS-E & 0.3750 & 4 && 0.4694 & 4 && 0.4574 & 4 && 0.4100 & 4 && 166 & 388 & 0.4278 \\
+FRank & 0.3137 & 11 && 0.2849 & 10 && 0.3029 & 11 && 0.2623 & 11 && 244 & 842 & 0.2898 \\
+FSMRank & 0.8351 & 4 && 0.8776 & 4 && 0.8621 & 5 && 0.5789 & 7 && 366 & 482 & 0.7593 \\
+FSM$^{SVM}$ & 0.2292 & 2 && 0.4082 & 4 && 0.5426 & 4 && 0.3500 & 4 && 149 & 389 & 0.3830 \\
+GAS-E & 0.3814 & 4 && 0.4694 & 4 && 0.4574 & 4 && 0.4100 & 4 && 167 & 389 & 0.4293 \\
 GP & - & - && - & - && 0.6667 & 2 && 0.5000 & 2 && 7 & 12 & 0.5833 \\
-GPRank & 0.8710 & 3 && 0.7253 & 3 && 0.6591 & 3 && 0.8173 & 3 && 290 & 376 & 0.7713 \\
+GPRank & 0.8750 & 3 && 0.7253 & 3 && 0.6591 & 3 && 0.8173 & 3 && 293 & 379 & 0.7731 \\
 GRankRLS & - & - && - & - && 0.2895 & 2 && - & - && 11 & 38 & 0.2895 \\
-GroupCE & 0.7312 & 3 && - & - && 0.7273 & 3 && 0.7212 & 3 && 207 & 285 & 0.7263 \\
-GroupMLE & 0.5269 & 3 && - & - && 0.6250 & 3 && 0.6538 & 3 && 172 & 285 & 0.6035 \\
-IntervalRank & 0.5897 & 1 && 0.3750 & 1 && - & - && 0.3158 & 1 && 50 & 117 & 0.4274 \\
-IPRank & 0.9355 & 3 && 0.8132 & 3 && 0.7955 & 3 && 0.8514 & 6 && 357 & 420 & 0.8500 \\
+GroupCE & 0.7292 & 3 && - & - && 0.7273 & 3 && 0.7212 & 3 && 209 & 288 & 0.7257 \\
+GroupMLE & 0.5208 & 3 && - & - && 0.6250 & 3 && 0.6538 & 3 && 173 & 288 & 0.6007 \\
+IntervalRank & 0.6000 & 1 && 0.3750 & 1 && - & - && 0.3158 & 1 && 51 & 118 & 0.4322 \\
+IPRank & 0.9375 & 3 && 0.8132 & 3 && 0.7955 & 3 && 0.8514 & 6 && 360 & 423 & 0.8511 \\
 KeepRank & - & - && - & - && - & - && 0.5385 & 3 && 56 & 104 & 0.5385 \\
@@
 LAC-MR-OR & - & - && - & - && 0.6667 & 2 && 0.7642 & 12 && 179 & 235 & 0.7617 \\
-LambdaMART & 0.5714 & 2 && - & - && 1.0000 & 1 && 0.6786 & 3 && 54 & 81 & 0.6667 \\
+LambdaMART & 0.4082 & 3 && - & - && 1.0000 & 1 && 0.6786 & 3 && 62 & 109 & 0.5688 \\
 LambdaNeuralRank & 1.0000 & 1 && 1.0000 & 1 && 1.0000 & 1 && - & - && 15 & 15 & 1.0000 \\
 LambdaRank & 0.2000 & 1 && 0.2000 & 1 && 0.5714 & 2 && - & - && 10 & 24 & 0.4167 \\
-LARF & 0.9892 & 3 && 0.9890 & 3 && 0.9886 & 3 && 0.9808 & 3 && 371 & 376 & 0.9867 \\
-Linear Regression & 0.0714 & 9 && 0.1099 & 9 && 0.0829 & 8 && 0.0650 & 8 && 63 & 761 & 0.0828 \\
-ListMLE & 0.0000 & 1 && - & - && 0.0213 & 4 && 0.00962 & 3 && 3 & 199 & 0.0151 \\
-ListNet & 0.4495 & 12 && 0.4911 & 12 && 0.5982 & 12 && 0.4504 & 12 && 460 & 928 & 0.496 \\
-ListReg & 0.7312 & 3 && 0.6923 & 3 && - & - && 0.4327 & 3 && 176 & 288 & 0.6111 \\
-LRUF & 0.9823 & 4 && 0.9817 & 4 && 0.9818 & 4 && 0.9680 & 4 && 447 & 457 & 0.9781 \\
+LARF & 0.9896 & 3 && 0.9890 & 3 && 0.9886 & 3 && 0.9808 & 3 && 374 & 379 & 0.9868 \\
+Linear Regression & 0.0754 & 9 && 0.1099 & 9 && 0.0829 & 8 && 0.0650 & 8 && 64 & 771 & 0.0830 \\
+ListMLE & 0.0000 & 2 && 0.0000 & 1 && 0.0213 & 4 && 0.00962 & 3 && 3 & 240 & 0.0125 \\
+ListNet & 0.4480 & 12 && 0.4911 & 12 && 0.5982 & 12 && 0.4504 & 12 && 461 & 931 & 0.4952 \\
+ListReg & 0.7292 & 3 && 0.6923 & 3 && - & - && 0.4327 & 3 && 178 & 291 & 0.6117 \\
+LRUF & 0.9828 & 4 && 0.9817 & 4 && 0.9818 & 4 && 0.9680 & 4 && 450 & 460 & 0.9783 \\
 MCP & - & - && - & - && - & - && 0.5714 & 2 && 40 & 70 & 0.5714 \\
@@
 MultiStageBoost & - & - && - & - && - & - && 0.1364 & 2 && 6 & 44 & 0.1364 \\
-NewLoss & 0.5161 & 3 && 0.4286 & 3 && 0.3977 & 3 && - & - && 122 & 272 & 0.4485 \\
-OWPC & 0.6500 & 6 && - & - && - & - && 0.6241 & 6 && 166 & 261 & 0.6360 \\
-PERF-MAP & 0.3894 & 4 && 0.2661 & 4 && 0.2000 & 4 && 0.7680 & 4 && 191 & 457 & 0.4179 \\
+NewLoss & 0.5208 & 3 && 0.4286 & 3 && 0.3977 & 3 && - & - && 124 & 275 & 0.4509 \\
+OWPC & 0.6475 & 6 && - & - && - & - && 0.6241 & 6 && 167 & 263 & 0.6350 \\
+PERF-MAP & 0.3966 & 4 && 0.2661 & 4 && 0.2000 & 4 && 0.7680 & 4 && 193 & 460 & 0.4196 \\
 PermuRank & - & - && - & - && - & - && 0.4091 & 3 && 18 & 44 & 0.4091 \\
@@
 RandomForest & - & - && - & - && 0.4224 & 8 && 0.4389 & 8 && 147 & 341 & 0.4311 \\
-Rank-PMBGP & - & - && 0.7692 & 1 && 0.2727 & 1 && 0.8750 & 1 && 24 & 40 & 0.6750 \\
+Rank-PMBGP & - & - && 0.7692 & 1 && 0.2727 & 1 && 0.8750 & 1 && 27 & 40 & 0.6750 \\
 RankAggNDCG & - & - && 0.5000 & 3 && 0.8784 & 3 && 0.7922 & 3 && 165 & 229 & 0.7205 \\
-RankBoost & 0.3211 & 12 && 0.2794 & 10 && 0.3936 & 17 && 0.3134 & 14 && 309 & 939 & 0.3291 \\
+RankBoost & 0.3303 & 12 && 0.2794 & 10 && 0.3936 & 17 && 0.3134 & 14 && 312 & 942 & 0.3312 \\
 RankBoost (Kernel-PCA) & - & - && 0.2857 & 3 && - & - && - & - && 26 & 91 & 0.2857 \\
@@
 RankDE & - & - && 0.5385 & 1 && 0.1818 & 1 && 1.0000 & 1 && 25 & 40 & 0.6250 \\
-RankELM (pairwise) & 0.6154 & 1 && 0.6500 & 1 && 0.6944 & 1 && 0.5143 & 2 && 111 & 185 & 0.6000 \\
-RankELM (pointwise) & 0.6923 & 1 && 0.7000 & 1 && 0.8056 & 1 && 0.5429 & 2 && 122 & 185 & 0.6595 \\
+RankELM (pairwise) & 0.6475 & 1 && 0.6500 & 1 && 0.6944 & 1 && 0.5143 & 2 && 112 & 186 & 0.6022 \\
+RankELM (pointwise) & 0.7000 & 1 && 0.7000 & 1 && 0.8056 & 1 && 0.5429 & 2 && 123 & 186 & 0.6613 \\
 RankMGP & - & - && - & - && - & - && 0.2222 & 1 && 4 & 18 & 0.2222 \\
-RankNet & 0.1887 & 1 && 0.2857 & 3 && 0.5915 & 5 && - & - && 66 & 173 & 0.3815 \\
+RankNet & 0.1887 & 3 && 0.2857 & 3 && 0.5915 & 5 && - & - && 66 & 173 & 0.3815 \\
 RankRLS & - & - && - & - && 0.3684 & 2 && - & - && 14 & 38 & 0.3684 \\
-RankSVM & 0.3010 & 12 && 0.3613 & 11 && 0.4496 & 17 && 0.3400 & 13 && 323 & 885 & 0.3650 \\
-RankSVM-Primal & 0.3920 & 8 && 0.4509 & 8 && 0.4591 & 7 && 0.3520 & 7 && 283 & 687 & 0.4119 \\
-RankSVM-Struct & 0.3520 & 9 && 0.4136 & 9 && 0.4467 & 9 && 0.3624 & 9 && 315 & 793 & 0.3928 \\
+RankSVM & 0.3014 & 12 && 0.3613 & 11 && 0.4496 & 17 && 0.3400 & 13 && 324 & 888 & 0.3649 \\
+RankSVM-Primal & 0.3911 & 8 && 0.4509 & 8 && 0.4591 & 7 && 0.3520 & 7 && 284 & 690 & 0.4116 \\
+RankSVM-Struct & 0.3518 & 9 && 0.4136 & 9 && 0.4467 & 9 && 0.3624 & 9 && 316 & 805 & 0.3925 \\
 RCP & - & - && 0.5758 & 3 && 0.7407 & 3 && 0.3636 & 3 && 55 & 104 & 0.5288 \\
 RE-QR & - & - && - & - && - & - && 0.8659 & 7 && 155 & 179 & 0.8659 \\
-REG-SHG-SDCG & 0.3846 & 1 && 0.4500 & 1 && - & - && 0.6579 & 1 && 58 & 117 & 0.4957 \\
-Ridge Regression & 0.4088 & 7 && 0.3333 & 7 && 0.3648 & 7 && 0.2905 & 7 && 226 & 650 & 0.3477 \\
-RSRank & 0.5729 & 4 && 0.5306 & 4 && 0.6277 & 4 && 0.6600 & 4 && 232 & 388 & 0.5979 \\
+REG-SHG-SDCG & 0.4000 & 1 && 0.4500 & 1 && - & - && 0.6579 & 1 && 59 & 118 & 0.5000 \\
+Ridge Regression & 0.4074 & 7 && 0.3333 & 7 && 0.3648 & 7 && 0.2905 & 7 && 227 & 653 & 0.3476 \\
+RSRank & 0.5773 & 4 && 0.5306 & 4 && 0.6277 & 4 && 0.6600 & 4 && 233 & 389 & 0.5990 \\
 SmoothGrad & - & - && - & - && 0.3846 & 2 && - & - && 5 & 13 & 0.3846 \\
-SmoothRank & 0.6038 & 7 && 0.6340 & 7 && 0.6415 & 7 && 0.5307 & 7 && 390 & 650 & 0.6000 \\
-SoftRank & 0.2308 & 1 && 0.2750 & 1 && 0.6111 & 1 && - & - && 42 & 115 & 0.3652 \\
-SortNet & 0.2500 & 2 && 0.5147 & 4 && 0.5667 & 4 && 0.5000 & 2 && 113 & 238 & 0.4748 \\
-SparseRank & 0.8224 & 4 && 0.8173 & 4 && 0.7944 & 4 && - & - && 258 & 318 & 0.8113 \\
-SVM$^{MAP}$ & 0.2893 & 7 && 0.3801 & 8 && 0.3591 & 8 && 0.3498 & 10 && 254 & 734 & 0.3460 \\
+SmoothRank & 0.6049 & 7 && 0.6340 & 7 && 0.6415 & 7 && 0.5307 & 7 && 392 & 653 & 0.6003 \\
+SoftRank & 0.2500 & 1 && 0.2750 & 1 && 0.6111 & 1 && - & - && 43 & 116 & 0.3707 \\
+SortNet & 0.2667 & 2 && 0.5147 & 4 && 0.5667 & 4 && 0.5000 & 2 && 114 & 239 & 0.4770 \\
+SparseRank & 0.8241 & 4 && 0.8173 & 4 && 0.7944 & 4 && - & - && 259 & 319 & 0.8119 \\
+SVM$^{MAP}$ & 0.2901 & 7 && 0.3801 & 8 && 0.3591 & 8 && 0.3498 & 10 && 255 & 737 & 0.3460 \\
 SwarmRank & - & - && 0.1538 & 1 && 0.0909 & 1 && 0.1250 & 1 && 5 & 40 & 0.1250 \\
-TGRank & 0.5417 & 4 && 0.6122 & 4 && 0.5000 & 4 && 0.4600 & 4 && 205 & 388 & 0.5284 \\
+TGRank & 0.5464 & 4 && 0.6122 & 4 && 0.5000 & 4 && 0.4600 & 4 && 206 & 389 & 0.5296 \\
 TM & 0.5909 & 3 && 0.7576 & 3 && - & - && 0.6136 & 3 && 65 & 99 & 0.6566 \\

### 2014-05-27 implementation/java/src/letor/parallel/ListNet.java

@@ -1,4 +1,4 @@
-package letor;
+package letor.parallel;
 
 import org.apache.pig.PigServer;
 import org.apache.pig.data.Tuple;

### 2014-05-27 implementation/java/src/letor/serial/ListNetWrapper.java

@@ -1,0 +1,21 @@
+package letor.serial;
+
+import ciir.umass.edu.learning.neuralnet.ListNet;
+import ciir.umass.edu.metric.NDCGScorer;
+
+import java.io.File;
+import java.io.FileReader;
+
+/**
+ * Created by niek.tax on 5/27/2014.
+ */
+public class ListNetWrapper {
+
+    public static void main(String[] args) {
+        ListNet listNet = new ListNet();
+        listNet.printParameters();
+        listNet.init();
+        File[] train = FileReader("input/ohsumed/Fold1/train.txt");
+        listNet.model();
+    }
+}

### 2014-07-17 report/Chapters/Implementation.tex
nieuwe lineare regressie voor SmoothRank

@@
-Describes implementation details of the Learning-to-Rank algorithm parallel implementations in HDInsight that are either Hadoop, Microsoft Azure, or HDInsight and are not part of the algorithm specification itself.
+This chapter describes implementation details of the Learning-to-Rank algorithm parallel implementations in HDInsight that are either Hadoop, Microsoft Azure, or HDInsight and are not part of the algorithm specification itself.
+The first section of this chapter will briefly discuss the HDInsight platform, the Hadoop ecosystem components offered by HDInsight and, in this regard, the Hadoop components used for implementation of the algorithms described in Chapter \ref{chap:ltr_methods}.
+
+\section{The HDInsight platform}
+TODO:
+ - No direct access to jobtracker
+ - C# API and access via Windows PowerShell scripts
+ - REST endpoints for Templeton/WebHcat / Ambari / HiveServer / HiveServer2 and Oozie
+ - Pig for ease of implementation
+ - Templeton and Oozie allow submission of Pig jobs
+ - Oozie requires Pig jobs to be stored on HDFS, Templeton allows running dynamically generated Pig from code.
+
+\section{Algorithm implementation}
+\subsection{ListNet}

### 2014-07-17 report/Chapters/LTRMethods.tex
nieuwe lineare regressie voor SmoothRank

@@
+\label{chap:ltr_methods}
 The learning algorithms of the well-performing Learning-to-Rank methods selected in Chapter \ref{chap:cross_comparison} are presented and explained in this Chapter.

### 2014-07-17 implementation/java/src/letor/parallel/ListNet.scala

@@ -1,0 +1,17 @@
+/*for (i <- 1 to ITERATIONS) {
+  val gradient = sparkContext.accumulator(spark.examples.Vector.zeros(dim))
+  val loss = sparkContext.accumulator(0.0)
+  for (q <- queries) {
+    val expRelScores = q.relScores.map(y => math.exp(beta*y.toDouble))
+    val ourScores = q.docFeatures.map(x => w dot x); val expOurScores = ourScores.map(z => math.exp(z))
+    val sumExpRelScores = expRelScores.reduce(_ + _); val sumExpOurScores = expOurScores.reduce(_ + _)
+    val P_y = expRelScores.map(y => y/sumExpRelScores); val P_z = expOurScores.map(z => z/sumExpOurScores)
+    var lossForAQuery = 0.0; var gradientForAQuery = spark.examples.Vector.zeros(dim)
+    for (j <- 0 to q.relScores.length-1) {
+      gradientForAQuery += (q.docFeatures(j) * (P_z(j) - P_y(j)))
+      lossForAQuery += -P_y(j) * math.log(P_z(j))
+    }
+    gradient += gradientForAQuery; loss += lossForAQuery
+  }
+  w -= gradient.value * stepSize
+}*/

### 2014-07-17 implementation/java/src/letor/parallel/ListNetCluster.java

@@ -4,10 +4,8 @@ import letor.parallel.util.AzurePigWrapper;
 import letor.parallel.util.DataSets;
 import letor.parallel.util.LtrUtils;
 import letor.parallel.util.Metadata;
-import org.apache.pig.data.Tuple;
 
 import java.util.ArrayList;
-import java.util.LinkedList;
 import java.util.List;
 
 /**
@@ -19,8 +17,8 @@ import java.util.List;
 public class ListNetCluster {
     // Initialise hyper-parameters
     private static final DataSets.DataSet DATASET = DataSets.DataSet.MSLR_WEB30K;
-    private static final double   STEPSIZE   = 0.0001; // MSLR-WEB10K: 0.0001, ohsumed: 0.01
-    private static final int      ITERATIONS = 10;
+    private static final double   STEPSIZE   = 0.00005; // MSLR-WEB10K: 0.0001, ohsumed: 0.01
+    private static final int      ITERATIONS = 20;
     private static final int      FOLDS      = 5;
     private static final int      k          = 10; // NDCG@k
 

### 2014-07-17 implementation/java/src/letor/parallel/util/LtrUtils.java

@@ -57,7 +57,7 @@ public class LtrUtils {
     }
 
     public static String getPigConfigString(long fileSize, int mappers, int reducers){
-        long splitByBytes = fileSize / (mappers-1); // always keep one mapper available for templetonjob
+        long splitByBytes = (long)((fileSize / (mappers-1)) * 1.35); // always keep one mapper available for templetonjob. 0.7 adjusts for standardisation and feature scaling, which shrink data set by +- 30%
         int usedReducers = reducers > 1 ? reducers - 1 : 1; // always keep one reducer available for templetonjob
         String configString =
                 "SET default_parallel "+usedReducers+";"+

### 2014-07-17 implementation/java/src/letor/serial/FoldRunHandler.java

@@ -5,33 +5,35 @@ import ciir.umass.edu.learning.*;
 import ciir.umass.edu.metric.MetricScorer;
 import ciir.umass.edu.metric.MetricScorerFactory;
 import ciir.umass.edu.utilities.SimpleMath;
+import letor.serial.parameterizedRankers.AbstractParameterizedRanker;
 import letor.serial.util.Measurement;
 
 import java.util.List;
 
 /**
- * RankLib evaluation wrapper.
- * Calculates average running time and evaluation results on a prefolded dataset.
- *
- * @author Niek Tax
+ * Handles iteration over folds
+ * Created by niek.tax on 7/15/2014.
  */
-public class ListNetWrapper {
-    // java -jar bin/RankLib.jar -train C:\Git-data\Afstuderen\implementation\java\input\ohsumed\Fold1\train.txt -ranker 7 -metric2t NDCG@10 -gmax 2 -validate C:\Git-data\Afstuderen\implementation\java\input\ohsumed\Fold1\train.txt -test C:\Git-data\Afstuderen\implementation\java\input\ohsumed\Fold1\test.txt
+public class FoldRunHandler {
     private MetricScorerFactory mFact = new MetricScorerFactory();
 
     private String path  = "C:\\Git-data\\Afstuderen\\implementation\\java\\input\\";
     private int folds    = 5;
 
-    private RANKER_TYPE rType  = RANKER_TYPE.LISTNET;
+    private Ranker ranker;
+
     private String trainMetric = "NDCG@10";
     private String testMetric  = trainMetric;
 
     private MetricScorer trainScorer = mFact.createScorer(trainMetric);
     private MetricScorer testScorer  = mFact.createScorer(testMetric);
 
-    public Measurement averageScore(String pathPostFix) {
-        path = path + pathPostFix+"\\";
-        System.out.println(path);
+    public FoldRunHandler(AbstractParameterizedRanker ranker, String pathPostFix){
+        this.ranker      = ranker.getParameterizedRanker();
+        path = path + pathPostFix +"\\";
+    }
+
+    public Measurement averageScore() {
         double scoreSum          = 0.0;
 
         Long startTime = System.nanoTime();
@@ -55,8 +57,7 @@ public class ListNetWrapper {
         List<RankList> test       = readInput(testFile);
 
         int[] features = getFeatureFromSampleVector(train);
-        RankerFactory rFact = new RankerFactory();
-        Ranker ranker = rFact.createRanker(rType, train, features);
[…]

### 2014-07-17 implementation/java/src/letor/serial/SmoothRank.scala

@@ -1,0 +1,79 @@
+package letor.serial;
+
+import ciir.umass.edu.learning.{RankList, Ranker}
+import org.apache.commons.math3.optim.nonlinear.scalar.gradient.NonLinearConjugateGradientOptimizer
+import org.codehaus.jet.regression.estimators.OLSMultipleLinearRegressionEstimator
+
+/*
+ * Implements SmoothRank
+ * @Author Niek Tax
+ */
+class SmoothRank extends Ranker{
+  //Parameters
+  val smoothFactor             = 0.01
+  val initIterations           = 10
+  val initEps                  = 0.01 // step size
+
+  //Local variables
+  var w:Array[Double] = null
+
+  override def init() {
+    PRINT("Initializing... ")
+    val samplesArray = samples.toArray
+    //PRINTLN("BEFORE olsInput")
+    //PRINTLN("samplesArray.length: "+samplesArray.length)
+    //val olsInput = samplesArray.flatMap(x => toOlsInput(x.asInstanceOf[RankList]))
+    //PRINTLN("AFTER olsInput")
+    //val ols = new OLSMultipleLinearRegression()
+    val ols2 = new OLSMultipleLinearRegressionEstimator()
+    samplesArray.foreach(x => addToOLS(x.asInstanceOf[RankList], ols2))
+
+    w = ols2.estimateRegressionParameters()
+    PRINTLN("w: "+w)
+
+    //ols.newSampleData(olsInput, olsInput.length/features.length, features.length-1)
+
+    //w = ols.estimateRegressionParameters()
+
+
+    //var cost = 0.0
+    //cost = Math.pow(2, olsInput.foreach(cost += _) )
+    //val q = samples.map(x:RankList => ((RankList) x).)
+
+    PRINTLN("[DONE]")
+  }
+
+  override def learn() {
+    PRINTLN("Called Learn")
+    // Calculate starting point by regression on gain function values
+    val optimizer = new NonLinearConjugateGradientOptimizer(NonLinearConjugateGradientOptimizer.Formula.POLAK_RIBIERE, null)
+  }
+
+  def toOlsInput(rl: RankList):Array[Double] = {
+    var olsList = List[Double]()
+    for(i <- 1 until rl.size()){
+      val dp = rl.get(i)
+      olsList = olsList :+ (Math.pow(2, dp.getLabel) - 1)
+      for(j <- 1 until features.length)
[…]

### 2014-07-17 implementation/java/src/letor/serial/WrappedRunner.java

@@ -1,0 +1,16 @@
+package letor.serial;
+
+import letor.serial.parameterizedRankers.SmoothRankHandler;
+import letor.serial.util.Measurement;
+
+/**
+ * Created by niek.tax on 7/15/2014.
+ */
+public class WrappedRunner {
+    public static void main(String[] args){
+        FoldRunHandler listNetWrapper = new FoldRunHandler(new SmoothRankHandler(), "ohsumed");
+        Measurement measurement = listNetWrapper.averageScore();
+        System.out.println("Time elapsed: "+measurement.getRunningTime());
+        System.out.println("NDCG@10:      "+measurement.getEvaluationResult()+" ms");
+    }
+}

### 2014-07-17 implementation/java/src/letor/serial/WrappedSmoothRankRunner.java

@@ -5,13 +5,11 @@ import letor.serial.util.Measurement;
 /**
  * Created by niek.tax on 5/28/2014.
  */
-public class WrappedListNetRunner {
+public class WrappedSmoothRankRunner {
     public static void main(String[] args){
-        /*
-        FoldRunHandler listNetWrapper = new FoldRunHandler();
+        /*ListNetWrapper listNetWrapper = new ListNetWrapper();
         Measurement measurement = listNetWrapper.averageScore("MSLR-WEB30K");
         System.out.println("Time elapsed: "+measurement.getRunningTime());
-        System.out.println("NDCG@10:      "+measurement.getEvaluationResult()+" ms");
-        */
+        System.out.println("NDCG@10:      "+measurement.getEvaluationResult()+" ms");*/
     }
 }

### 2014-07-17 implementation/java/src/letor/serial/parameterizedRankers/AbstractParameterizedRanker.java

@@ -1,0 +1,10 @@
+package letor.serial.parameterizedRankers;
+
+import ciir.umass.edu.learning.Ranker;
+
+/**
+ * Created by niek.tax on 7/15/2014.
+ */
+public interface AbstractParameterizedRanker {
+    public Ranker getParameterizedRanker();
+}

### 2014-07-17 implementation/java/src/letor/serial/parameterizedRankers/ListNetHandler.java

@@ -1,0 +1,20 @@
+package letor.serial.parameterizedRankers;
+
+import ciir.umass.edu.learning.Ranker;
+import ciir.umass.edu.learning.RankerFactory;
+import ciir.umass.edu.learning.neuralnet.ListNet;
+import letor.serial.FoldRunHandler;
+
+/**
+ * Created by niek.tax on 7/15/2014.
+ */
+public class ListNetHandler implements AbstractParameterizedRanker{
+    // PARAMETERS
+    private int nIteration = 20000;
+
+    public Ranker getParameterizedRanker() {
+        ListNet ranker = new ListNet();
+        ranker.nIteration = nIteration;
+        return ranker;
+    }
+}

### 2014-07-17 implementation/java/src/letor/serial/parameterizedRankers/SmoothRankHandler.java

@@ -1,0 +1,16 @@
+package letor.serial.parameterizedRankers;
+
+import ciir.umass.edu.learning.Ranker;
+import letor.serial.SmoothRank;
+
+/**
+ * Created by niek.tax on 7/15/2014.
+ */
+public class SmoothRankHandler implements AbstractParameterizedRanker{
+    // PARAMETERS
+
+    public Ranker getParameterizedRanker() {
+        SmoothRank ranker = new SmoothRank();
+        return ranker;
+    }
+}

### 2014-09-02 paper/CIKM/paper.tex
begin ECIR-versie paper + betere benaming datachunks

--- 2014-05-27 cikm_paper/paper.tex
@@
 
-\begin{figure}
-\centering
-\includegraphics[scale=0.19]{gfx/ndcg3_winnum}
-\caption{NDCG@3 comparison of 87 learning to rank methods}
-\label{fig:normalized_winning_number_ndcg3}
-\end{figure}
 \section{A Methodology for Comparing Learning to Rank Methods Cross-Benchmark}
@@
 
-\begin{figure}
-\includegraphics[scale=0.19]{gfx/ndcg5_winnum}
-\caption{NDCG@5 comparison of 87 learning to rank methods}
-\label{fig:normalized_winning_number_ndcg5}
-\end{figure}
 \section{Results of Learning to Rank Comparison}
@@
 \subsection{NDCG@3}
+\begin{figure}
+\centering
+\includegraphics[scale=0.19]{gfx/ndcg3_winnum}
+\caption{NDCG@3 comparison of 87 learning to rank methods}
+\label{fig:normalized_winning_number_ndcg3}
+\end{figure}
 Figure \ref{fig:normalized_winning_number_ndcg3} shows the NWN of learning to rank methods based on NDCG@3 results.
@@
 \subsection{NDCG@5}
+\begin{figure}
+\includegraphics[scale=0.19]{gfx/ndcg5_winnum}
+\caption{NDCG@5 comparison of 87 learning to rank methods}
+\label{fig:normalized_winning_number_ndcg5}
+\end{figure}
 Figure \ref{fig:normalized_winning_number_ndcg5} shows the NWN of learning to rank methods based on NDCG@5 results.

### 2014-09-02 paper/ECIR/paper.tex
begin ECIR-versie paper + betere benaming datachunks

--- 2014-09-02 paper/CIKM/paper.tex
@@
-% This is "sig-alternate.tex" V2.0 May 2012
-% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
-%
-% This example file demonstrates the use of the 'sig-alternate.cls'
-% V2.5 LaTeX2e document class file. It is for those submitting
-% articles to ACM Conference Proceedings WHO DO NOT WISH TO
-% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
-% The 'sig-alternate.cls' file will produce a similar-looking,
-% albeit, 'tighter' paper resulting in, invariably, fewer pages.
-%
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V2.5) produces:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) NO page numbers
-%
-% as against the acm_proc_article-sp.cls file which
-% DOES NOT produce 1) thru' 3) above.
-%
-% Using 'sig-alternate.cls' you have control, however, from within
-% the source .tex file, over both the CopyrightYear
-% (defaulted to 200X) and the ACM Copyright Data
-% (defaulted to X-XXXXX-XX-X/XX/XX).
-% e.g.
-% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
-% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
-%
-% ---------------------------------------------------------------------------------------------------------------
-% This .tex source is an example which *does* use
-% the .bib file (from which the .bbl file % is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission, you *NEED* to 'insert'
-% your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% ================= IF YOU HAVE QUESTIONS =======================
-% Questions regarding the SIGS styles, SIGS policies and
-% procedures, Conferences etc. should be sent to
-% Adrienne Griscti (griscti@acm.org)
-%
-% Technical questions _only_ to
-% Gerald Murray (murray@hq.acm.org)
-% ===============================================================
-%
-% For tracking purposes - this is V2.0 - May 2012
-
-\documentclass{sig-alternate}
+\documentclass{llncs}
+\usepackage{llncsdoc}
 \usepackage{booktabs}
 \usepackage{multicol}
+\usepackage{amsmath}
 \usepackage{longtable}
+\usepackage{graphicx}
 \usepackage{dblfloatfix}
@@
 
+\title{A Cross-Benchmark Comparison of 87 Learning to Rank Methods}
+
+\author{First Author\inst{1,2} \and Second Author\inst{1} \and Third Author\inst{2}}
+
+\authorrunning{First Author et al.} % abbreviated author list (for running head)
+
+%%%% list of authors for the TOC (use if author list has to be modified)
+\tocauthor{First Author, Second Author, Third Author}
+
+\institute{Blinded Institute 1, Adress Line 1, Address Line 2, County\\
+\email{first.author@institute.com},\\
+\and
+Blinded Institute 2, Address Line 1, Address Line 2, Country}
+
+%\author{Niek Tax\inst{1,2} \and Sander Bockting\inst{1} \and Djoerd Hiemstra\inst{2}}
 %
-% --- Author Metadata here ---
-\conferenceinfo{ACM International Conference on Information and Knowledge Management}{'14 Shanghai, China}
-\CopyrightYear{2014} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-% --- End of Author Metadata ---
-
-\title{A Cross-Benchmark Comparison of 87 Learning to Rank Methods}
+%\authorrunning{Niek Tax et al.} % abbreviated author list (for running head)
 %
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
+%%%% list of authors for the TOC (use if author list has to be modified)
+%\tocauthor{Niek Tax, Sander Bockting, Djoerd Hiemstra}
 %
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{3} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
-Niek Tax\titlenote{author is affiliated with the University of Twente, but this paper written during his time at Avanade Netherlands B.V.}\\
-       \affaddr{Avanade Netherlands B.V.}\\
-       \affaddr{Versterkerstraat 6}\\
-       \affaddr{1322 AP Almere}\\
-       \affaddr{The Netherlands}
-       \affaddr{niek.tax@avanade.com}
-% 2nd. author
-\alignauthor
-Sander Bockting\\
-       \affaddr{Avanade Netherlands B.V.}\\
-       \affaddr{Versterkerstraat 6}\\
-       \affaddr{1322 AP Almere}\\
-       \affaddr{The Netherlands}
-       \affaddr{sander.bockting@avanade.com}
-% 3rd. author
-\alignauthor
-Djoerd Hiemstra\\
-       \affaddr{University of Twente}\\
-       \affaddr{P.O. Box 217}\\
-       \affaddr{7500 AE Enschede}\\
-       \affaddr{The Netherlands}
-       \affaddr{d.hiemstra@utwente.nl}
-}
+%\institute{Avanade Netherlands B.V., Versterkerstraat 6, 1322 AP Almere, the Netherlands\\
+%\email{niek.tax@avanade.com},\\
+%\and
+%University of Twente, P.O. Box 217, 7500AE Enschede, the Netherlands}
 
@@
 \end{abstract}
-
-% A category with the (minimum) three required fields
-\category{H.3.3}{Information Systems}{Information Storage and Retrieval}[Information Search and Retrieval]
 
@@
 \begin{enumerate}[noitemsep,nolistsep]
-\item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011, Lin2012}
+\item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011,Lin2012}
 \item The study focuses on a different learning to rank task (e.g.
-rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}
-\item The study used an altered version of a benchmark that contained additional features \cite{Bidoki2009, Ding2010}
+rank aggregation or transfer ranking) \cite{De2011,De2010,Derhami2013,De2012,Chen2010,Ah-Pine2008,Wang2009c,De2013,Miao2013,Hoi2008,De2012b,Duh2011b,Argentini2012,Qin2010c,Volkovs2013,Desarkar2011,Pan2013,Lin2011b,Volkovs2012,Dammak2011}
+\item The study used an altered version of a benchmark that contained additional features \cite{Bidoki2009,Ding2010}
 \item The study provides no exact data of the evaluation results (e.g.
-results are only in graphical form) \cite{Wang2008, Wang2010, Xu2010, Kuo2009, Li2008, Xia2008, Zhou2011, Wu2011, Zhu2009, Karimzadehgan2011, Swersky2012, Pan2011, Ni2008, Ciaramita2008, Stewart2012, Petterson2009, Agarwal2010, Chang2009, Qin2008c, Adams2011, Sculley2009, Huang2008, Alejo2010, Sun2011, He2010b, Benbouzid2012, Geng2012, Chen2012, Xu2012, Shivaswamy2011}
-\item The study reported evaluation results in a different metric than the metrics chosen for this meta-analysis \cite{Yu2009, Thuy2009, Pahikkala2009, Kersting2009, Mohan2011}
-\item The study reported a higher performance on baseline methods than official benchmark runs \cite{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}
-\item The study did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
+results are only in graphical form) \cite{Wang2008,Wang2010,Xu2010,Kuo2009,Li2008,Xia2008,Zhou2011,Wu2011,Zhu2009,Karimzadehgan2011,Swersky2012,Pan2011,Ni2008,Ciaramita2008,Stewart2012,Petterson2009,Agarwal2010,Chang2009,Qin2008c,Adams2011,Sculley2009,Huang2008,Alejo2010,Sun2011,He2010b,Benbouzid2012,Geng2012,Chen2012,Xu2012,Shivaswamy2011}
+\item The study reported evaluation results in a different metric than the metrics chosen for this meta-analysis \cite{Yu2009,Thuy2009,Pahikkala2009,Kersting2009,Mohan2011}
+\item The study reported a higher performance on baseline methods than official benchmark runs \cite{Dubey2009,Banerjee2009,Peng2010b,Song2014,Bian2010,Bian2010b,Carvalho2008,Acharyya2012,Peng2010b,Tran2012,Asadi2013c}
+\item The study did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008,Wang2012b,Buffoni2011}.
 \end{enumerate}
@@
 \midrule
-AdaRank-MAP & \cite{Xu2007} & L2, L3, L4 & ListMLE & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\
+AdaRank-MAP & \cite{Xu2007} & L2, L3, L4 & ListMLE & \cite{Xia2008} & \cite{Lin2010,Lin2011,Gao2014} \\
 AdaRank-NDCG & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & ListNet & \cite{Cao2007} & L2, L3, L4 \\ 
@@
 BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} &  NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
-BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009, Volkovs2013} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
-BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c, Volkovs2013} & PERF-MAP & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
+BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009,Volkovs2013} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
+BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009,Ganjisaffar2011c,Volkovs2013} & PERF-MAP & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
 BT & \cite{Zhou2008} & \cite{Zhou2008} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
@@
 CoList & \cite{Gao2014} & \cite{Gao2014} & RankAggNDCG & \cite{Wang2013} & \cite{Wang2013} \\
-Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010, Sato2013} \\
-DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost (Kernel-PCA) & \cite{Duh2008} & \cite{Duh2008, Sato2013} \\
+Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013,Alcantara2010,Sato2013} \\
+DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost (Kernel-PCA) & \cite{Duh2008} & \cite{Duh2008,Sato2013} \\
 DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankBoost (SVD) & \cite{Lin2009} & \cite{Lin2009} \\
@@
 FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\
-FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b, Laporte2013} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
+FenchelRank & \cite{Lai2013} & \cite{Lai2013,Lai2013b,Laporte2013} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
 FocusedBoost & \cite{Niu2012} & \cite{Niu2012} &  RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
 FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankMGP & \cite{Lin2012} & \cite{Lin2012} \\ 
-FocusedSVM & \cite{Niu2012} & \cite{Niu2012} & RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
+FocusedSVM & \cite{Niu2012} & \cite{Niu2012} & RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013,Papini2012,Niu2012} \\ 
 FP-Rank & \cite{Song2013} & \cite{Song2013} & RankRLS & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
-FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & RankSVM & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010, Papini2012} \\ 
+FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & RankSVM & \cite{Herbrich1999,Joachims2002} & L2, L3, \cite{Busa-Fekete2013,Freno2011,He2010,Alcantara2010,Papini2012} \\ 
 FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & RankSVM-Primal &  & L3, \cite{Lai2011} \\ 
@@
 GroupMLE & \cite{Lin2010} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
-IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010} \\
-IPRank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
-KeepRank & \cite{Chen2009} & \cite{Chen2009} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011, Papini2012} \\
+IntervalRank & \cite{Moon2010} & \cite{Moon2010,Freno2011} & SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010} \\
+IPRank & \cite{Wang2009b} & \cite{Wang2009b,Torkestani2012} & SoftRank & \cite{Taylor2008,Guiver2008} & \cite{Qin2010b} \\ 
+KeepRank & \cite{Chen2009} & \cite{Chen2009} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011,Papini2012} \\
 KL-CRF & \cite{Volkovs2011} & \cite{Volkovs2011} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
-LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008, Alcantara2010} & SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
-LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} &SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
+LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008,Alcantara2010} & SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012,Xu2008,Niu2012} \\ 
+LambdaMART & \cite{Burges2010} & \cite{Asadi2013a,Ganjisaffar2011c} &SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
 LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank & \cite{Lai2013} & \cite{Lai2013} \\ 
-LambdaRank & \cite{Burges2006} & \cite{Papini2012, Tan2013} & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
+LambdaRank & \cite{Burges2006} & \cite{Papini2012,Tan2013} & TM & \cite{Zhou2008} & \cite{Zhou2008,Papini2012,Tan2013} \\ 
 LARF & \cite{Torkestani2012} & \cite{Torkestani2012} &  VFLR & \cite{Cai2012} & \cite{Cai2012} \\
-Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} & & & \\
+Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012,Volkovs2011} & & & \\
 \bottomrule

### 2014-09-02 implementation/java/src/letor/data/LargeDatasetCreator.java

@@ -1,5 +1,6 @@
 package letor.data;
 
+import com.google.common.base.Strings;
 import org.apache.commons.lang.StringUtils;
 
 import java.io.*;
@@ -15,8 +16,8 @@ import java.util.List;
  */
 public class LargeDatasetCreator {
     // Class parameters
-    private static int lowerDuplicationNumber = 11;
-    private static int upperDuplicationNumber = 20;
+    private static int lowerDuplicationNumber = 21;
+    private static int upperDuplicationNumber = 40;
 
     // Global variables
     private static String dataFolderPath = "C:/Git-data/Afstuderen/implementation/java/input";
@@ -88,7 +89,8 @@ public class LargeDatasetCreator {
         BufferedWriter[] writer = new BufferedWriter[writers];
         for(int i = 0; i < writers; i++){
             try {
-                writer[i] = new BufferedWriter(new FileWriter(outputPath + "/Fold1/train" + (lowerDuplicationNumber + i) + ".txt"));
+                String paddedNumberString = Strings.padStart(""+(lowerDuplicationNumber + i), 4, '0');
+                writer[i] = new BufferedWriter(new FileWriter(outputPath + "/Fold1/train" + paddedNumberString + ".txt"));
             }catch (IOException e){
                 e.printStackTrace();
             }

### 2014-09-02 implementation/java/src/letor/parallel/ListNetCluster.java

@@ -23,11 +23,11 @@ public class ListNetCluster {
     private static final int      k          = 10; // NDCG@k
 
     // Initialise paralellisation parameters
-    private static int  dataNodes                     = 4;
+    private static int  dataNodes                     = 8;
     private static int  availableMappers              = 4*dataNodes;
     private static int  availableReducers             = 2*dataNodes;
 
-    private static int duplicationNumber              = 4; // only used for CUSTOM
+    private static int duplicationNumber              = 9; // only used for CUSTOM
     private static Metadata metadata                  = DataSets.getMetaData(DATASET, duplicationNumber);
     private static long MAX_TRAIN_SIZE                = metadata.getMax_train_size(); // ohsumed: 5151958, MQ2007: 25820919, MQ2008: 5927007, MSLR-WEB10K: 838011150, MSLR-WEB30K:
     private static long MAX_VALI_SIZE                 = metadata.getMax_vali_size(); // ohsumed: 1764005, MQ2007:  8753466, MQ2008: 2237346, MSLR-WEB10K: 280714022, MSLR-WEB30K:
@@ -40,7 +40,7 @@ public class ListNetCluster {
 
     public static void main(String[] args) throws Exception {
         // Cluster configuration
-        String clusterName          = "ltr4";
+        String clusterName          = "ltr8";
         String clusterUser          = "admin";
         String clusterPassword      = "Qw!23456789";
         String storageAccount       = "ltrstorage";
@@ -72,10 +72,15 @@ public class ListNetCluster {
             // Standardize data and scale features
             pigLines.add(trainConfigString);
             pigLines.add("REGISTER wasb:///user/hdp/lib/*.jar;");
-            if(metadata.getName().equals("custom"))
-                pigLines.add("TRAIN = LOAD '" + pathPrefix + "/input/" + metadata.getName() + "/Fold" + fold + "/train[0-"+duplicationNumber+"].txt' USING PigStorage(' ');");
-            else
+            if(metadata.getName().equals("custom")){
+                int thousands = duplicationNumber /1000;
+                int hundreds  = (duplicationNumber % 1000) / 100;
+                int tens      = (duplicationNumber % 100) / 10;
+                int units     = duplicationNumber % 10;
[…]

### 2014-09-02 paper/ECIR/paper.tex
juiste bibstyle

@@
 \institute{Blinded Institute 1, Adress Line 1, Address Line 2, County\\
-\email{first.author@institute.com},\\
 \and
 Blinded Institute 2, Address Line 1, Address Line 2, Country}
-
 %\author{Niek Tax\inst{1,2} \and Sander Bockting\inst{1} \and Djoerd Hiemstra\inst{2}}
@@
 %University of Twente, P.O. Box 217, 7500AE Enschede, the Netherlands}
-
 \maketitle
@@
 \end{abstract}
-
 \keywords{Learning to rank, Information retrieval, Evaluation metric}
-
 \section{Introduction}
@@
 Table \ref{tbl:ltr_benchmark_collections} presents an overview of the benchmark collections included in the meta-analysis.
-
 \begin{table}[!h]
 \scriptsize{
+\caption{Included learning to rank evaluation benchmark collections}
+\label{tbl:ltr_benchmark_collections}
 \begin{tabular}{p{6.4cm}R{1.26cm}}\toprule
@@
 \end{tabular}
-\caption{Included learning to rank evaluation benchmark collections}
-\label{tbl:ltr_benchmark_collections}}
+}
 \end{table}
-
 For the LETOR collections, the evaluation results of the baseline models will be used from LETOR 2.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor2.0/baseline.aspx}, 3.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and 4.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} as listed on the LETOR website.\\
@@
 \scriptsize{
+\caption{Forward references of learning to rank benchmark papers}
+\label{tbl:ltr_benchmark_forref}
 \begin{tabular}{p{2.8cm}p{2.64cm}R{1.9cm}}\toprule
@@
 \end{tabular}
-\caption{Forward references of learning to rank benchmark papers}
-\label{tbl:ltr_benchmark_forref}}
+}
 \end{table}
-
 The LETOR 4.0, MSLR-web10/30k and Yandex Internet Mathematics Competition 2009 benchmark collections are not accompanied by a paper.
@@
 Table~\ref{tbl:ltr_benchmark_searches} shows the results of this literature search.
-
 \begin{table}[!h]
 \scriptsize{
-\begin{tabular}{p{5.2cm}R{2.5cm}}\toprule
+\caption{Google scholar search results for learning to rank benchmarks}
+\label{tbl:ltr_benchmark_searches}
+\begin{tabular}{p{5.2cm}R{2.5cm}}
+\toprule
 Search Query & Google scholar search results \\
@@
 \end{tabular}
-\caption{Google scholar search results for learning to rank benchmarks}
-\label{tbl:ltr_benchmark_searches}}
+}
 \end{table}
-
 \subsection{Literature Selection}
@@
 
-$\text{WN}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$\\
+$\mbox{WN}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$\\
 
@@
 $I_{\{M_i(j)>M_k(j)\}} = \begin{cases}
-1 & \text{if } M_i(j) > M_k(j), \\
-0 & \text{otherwise}
+1 & \mbox{if } M_i(j) > M_k(j), \\
+0 & \mbox{otherwise}
 \end{cases}$\\
@@
 \subsection{NDCG@3}
-\begin{figure}
+Figure \ref{fig:normalized_winning_number_ndcg3} shows the NWN of learning to rank methods based on NDCG@3 results.
+LambdaNeuralRank and CoList both acquired a NWN score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
+LARF and LRUF scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which results in more certainty on these methods' performance because they are validated on three datasets that additionally are more relevant than AOL and Yahoo Set 2 (number of evaluation results for LETOR 3.0 are higher than those for AOL and Yahoo set 2).
+FenchelRank, OWPC, SmoothRank, DCMP and ListNet are ordered decreasingly by NWN and at the same time increasingly in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.
+\begin{figure}[h]
 \centering
@@
 \end{figure}
-Figure \ref{fig:normalized_winning_number_ndcg3} shows the NWN of learning to rank methods based on NDCG@3 results.
-LambdaNeuralRank and CoList both acquired a NWN score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
-LARF and LRUF scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which results in more certainty on these methods' performance because they are validated on three datasets that additionally are more relevant than AOL and Yahoo Set 2 (number of evaluation results for LETOR 3.0 are higher than those for AOL and Yahoo set 2).
-FenchelRank, OWPC, SmoothRank, DCMP and ListNet are ordered decreasingly by NWN and at the same time increasingly in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
-
 LambdaNeuralRank, CoList, LARF, LRUF, OWPC and DCMP evaluation results are all based on one study, therefore are subjected to the risk of one overly optimistic study producing those results.
@@
 \subsection{NDCG@5}
-\begin{figure}
-\includegraphics[scale=0.19]{gfx/ndcg5_winnum}
-\caption{NDCG@5 comparison of 87 learning to rank methods}
-\label{fig:normalized_winning_number_ndcg5}
-\end{figure}
 Figure \ref{fig:normalized_winning_number_ndcg5} shows the NWN of learning to rank methods based on NDCG@5 results.
@@
 Like in the NDCG@3 comparison, SmoothRank and ListNet can be regarded as most reliable results because the evaluation measurements for these methods are based on LETOR official baselines.
+\begin{figure}[h]
+\includegraphics[scale=0.19]{gfx/ndcg5_winnum}
+\caption{NDCG@5 comparison of 87 learning to rank methods}
+\label{fig:normalized_winning_number_ndcg5}
+\end{figure}
 
 \subsection{NDCG@10}
+Figure \ref{fig:normalized_winning_number_ndcg10} shows the NWN of learning to rank methods based on NDCG@10 results.
+LambdaMART and LambdaNeuralRank score a NWN of 1.0 on the NDCG@10 comparison.
+For LambdaNeuralRank these results are again based on AOL dataset measurements.
+LambdaMART showed the highest NDCG@10 performance for the MSLR-WEB10k dataset.
+The set of algorithms for which there is no other algorithm with both a higher NWN and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and @5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
+In contrast to the NDCG@3 and @5 comparison, DCMP is not among the best algorithms.
 \begin{figure}
@@
 \end{figure}
-
-Figure \ref{fig:normalized_winning_number_ndcg10} shows the NWN of learning to rank methods based on NDCG@10 results.
-LambdaMART and LambdaNeuralRank score a NWN of 1.0 on the NDCG@10 comparison.
-For LambdaNeuralRank these results are again based on AOL dataset measurements.
-LambdaMART showed the highest NDCG@10 performance for the MSLR-WEB10k dataset.
-The set of algorithms for which there is no other algorithm with both a higher NWN and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and @5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
-In contrast to the NDCG@3 and @5 comparison, DCMP is not among the best algorithms.
-
 \subsection{MAP}
-\begin{figure}
+Figure \ref{fig:normalized_winning_number_map} shows the NWN of learning to rank methods based on MAP results.
+Comparisons on the NDCG metrics where highly in agreement on the best performing algorithms, MAP-based NWN results show different results.
+RankDE scores a NWN of 1.0 on one dataset, which is achieved by obtaining highest MAP-score on the LETOR 2.0 TD2003 which has many evaluation results are evaluated.
+\begin{figure}[h]
 \centering
@@
 \end{figure}
-Figure \ref{fig:normalized_winning_number_map} shows the NWN of learning to rank methods based on MAP results.
-Comparisons on the NDCG metrics where highly in agreement on the best performing algorithms, MAP-based NWN results show different results.
-RankDE scores a NWN of 1.0 on one dataset, which is achieved by obtaining highest MAP-score on the LETOR 2.0 TD2003 which has many evaluation results are evaluated.\\
-
 LARF and LRUF score very high NWN scores, but based on only few datasets, just as in the NDCG-based comparisons.
@@
 There is a lot of certainty on the accuracy of RankBoost and RankSVM as both models are evaluated on the majority of datasets included in the comparison for the MAP metric, but given their NWN it can said that both methods are not within the top performing learning to rank methods.
+\subsection{Cross-Metric}
 
-\subsection{Cross-Metric}
-\begin{figure}
+Figure \ref{fig:normalized_winning_number_all} shows the NWN as function of IWN for the methods described in Table 4.
+The cross-metric comparison is based on the NDCG@$\{3,5,10\}$ and MAP comparisons combined, which justifies analyzing the comparison more thoroughly.
+Figure \ref{fig:normalized_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
+In addition, Linear Regression and the ranking method of simply sorting on the best single feature are labelled as baselines.
+\begin{figure}[h]
 \centering
@@
 \end{figure}
-Figure \ref{fig:normalized_winning_number_all} shows the NWN as function of IWN for the methods described in Table 4.
-The cross-metric comparison is based on the NDCG@$\{3,5,10\}$ and MAP comparisons combined, which justifies analyzing the comparison more thoroughly.
-Figure \ref{fig:normalized_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
-In addition, Linear Regression and the ranking method of simply sorting on the best single feature are labelled as baselines.\\
-
 LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both IWN and NWN.
@@
 \balance
-\bibliographystyle{abbrv}
+\bibliographystyle{splncs03}
 {\scriptsize\bibliography{Bibliography}}
@@
 %APPENDICES are optional
-\appendix
-
+\section{Appendix}
 \begin{table*}[!hpt]
 \scriptsize{
+\caption{Learning to rank algorithms with measurements on benchmark datasets}
+\label{tab:ltr_methods_used}
 \begin{tabular}{lll|lll}\toprule
@@
 \end{tabular}
-\caption{Learning to rank algorithms with measurements on benchmark datasets}}
-\label{tab:ltr_methods_used}
+}
 \end{table*}
-
 \scriptsize{
-\begin{longtable}[!hpt]{@{}lrrrrrrrrrrrrrrrr@{}}\toprule
+\begin{longtable}[!hpt]{@{}lrrrrrrrrrrrrrrrr@{}}
+\toprule
 & \multicolumn{2}{c}{NDCG@3} & \phantom{a} 
@@
 \bottomrule
-\caption{Raw data of cross-benchmark comparison}
-\label{tab:raw_data}
+\caption{Raw data of cross-benchmark comparison} \label{tab:raw_data}\\
 \end{longtable}

### 2014-09-02 implementation/java/src/letor/data/LargeDatasetCreator.java

@@ -16,8 +16,8 @@ import java.util.List;
  */
 public class LargeDatasetCreator {
     // Class parameters
-    private static int lowerDuplicationNumber = 21;
-    private static int upperDuplicationNumber = 40;
+    private static int lowerDuplicationNumber = 41;
+    private static int upperDuplicationNumber = 60;
 
     // Global variables
     private static String dataFolderPath = "C:/Git-data/Afstuderen/implementation/java/input";

### 2014-09-02 paper/ECIR/paper.tex
ECIR paper :)

--- 2014-09-02 paper/ECIR/paper.tex
@@
 \usepackage{longtable}
+\usepackage{times}
 \usepackage{graphicx}
@@
 \usepackage{balance}
+
 \begin{document}
-
 \title{A Cross-Benchmark Comparison of 87 Learning to Rank Methods}
-
 \author{First Author\inst{1,2} \and Second Author\inst{1} \and Third Author\inst{2}}
-
 \authorrunning{First Author et al.} % abbreviated author list (for running head)
@@
 \tocauthor{First Author, Second Author, Third Author}
-
 \institute{Blinded Institute 1, Adress Line 1, Address Line 2, County\\
-\email{first.author@institute.com},\\
 \and
 Blinded Institute 2, Address Line 1, Address Line 2, Country}
-
 %\author{Niek Tax\inst{1,2} \and Sander Bockting\inst{1} \and Djoerd Hiemstra\inst{2}}
@@
 In addition, the ranking task has been argued to be a better fit for recommender systems than the regression task \cite{McNee2006} (which concerns predictions on a continuous scale) that is currently frequently used in such systems.\\
-
 Research in the field of ranking models has been based on manually designed ranking functions, such as the well-known BM25 model \cite{Robertson1994}.
@@
 Learning to rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\
-
 In recent years, several learning to rank benchmark datasets have been proposed with the aim of enabling comparison of learning to rank methods in terms of ranking accuracy.
@@
 Comparing ranking accuracy of learning to rank methods is largely hindered by this lack of a standard way of benchmarking.\\
-
 Gomes et al.
@@
 The paper will describe two stages in the meta-analysis process: 1) collection of evaluation results, and 2) comparison of learning to rank methods.
-
+\vspace{-0.08in}
 \section{Collecting Evaluation Results}
@@
 Table \ref{tbl:ltr_benchmark_collections} presents an overview of the benchmark collections included in the meta-analysis.
-
-\begin{table}[!h]
+\begin{table}
+\centering
 \scriptsize{
-\begin{tabular}{p{6.4cm}R{1.26cm}}\toprule
-Benchmark collection & \# of datasets \\
+\caption{Included learning to rank evaluation benchmark collections}
+\label{tbl:ltr_benchmark_collections}
+\begin{tabular}{p{4.2cm}R{1.26cm}|p{4.2cm}R{1.26cm}}\toprule
+Benchmark collection & \# of datasets & Benchmark collection& \# of datasets\\
 \midrule
-AOL		  & 1\\
-LETOR 2.0 & 3\\
-LETOR 3.0 & 7\\
-LETOR 4.0 & 2\\
-MSLR	  & 2\\
-WCL2R	  & 2\\
-Yahoo! Learning to Rank Challenge 	     & 2\\
-Yandex Internet Mathematics 2009 contest & 1\\
-\textbf{Total} & \textbf{20}\\ 	
+AOL		  & 1 & MSLR	  & 2\\
+LETOR 2.0 & 3 & WCL2R	  & 2\\
+LETOR 3.0 & 7 & Yahoo! Learning to Rank Challenge 	     & 2\\
+LETOR 4.0 & 2 &Yandex Internet Mathematics 2009 contest & 1\\
+\textbf{Total} & \textbf{20}&&\\ 	
 \bottomrule
 \end{tabular}
-\caption{Included learning to rank evaluation benchmark collections}
-\label{tbl:ltr_benchmark_collections}}
+}
 \end{table}
-
 For the LETOR collections, the evaluation results of the baseline models will be used from LETOR 2.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor2.0/baseline.aspx}, 3.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and 4.0\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} as listed on the LETOR website.\\
-
 LETOR 1.0 and 3.0, Yahoo!
@@
 Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search performed using Google Scholar.
-\begin{table}[!h]
+\begin{table}
+\centering
 \scriptsize{
-\begin{tabular}{p{2.8cm}p{2.64cm}R{1.9cm}}\toprule
-Benchmark & Paper & \# of forward references \\
+\caption{Forward references of learning to rank benchmark papers}
+\label{tbl:ltr_benchmark_forref}
+\begin{tabular}{p{2cm}p{1.8cm}R{1.5cm}|p{3cm}p{1.9cm}R{1.6cm}}\toprule
+Benchmark & Paper & \# of forward references & Benchmark & Paper & \# of forward references \\
 \midrule
-LETOR 1.0 \& 2.0 & Liu et al. \cite{Liu2007b} & 307\\
-LETOR 3.0 & Qin et al. \cite{Qin2010} & 105\\
-Yahoo! Learning to Rank Challenge & Chapelle et al. \cite{Chapelle2011a} & 102\\
-AOL dataset & Pass et al. \cite{Pass2006} & 339\\
-WCL2R & Alc{\^a}ntara et al. \cite{Alcantara2010} & 2\\
+LETOR 1.0 \& 2.0 & Liu et al. \cite{Liu2007b} & 307 & Yahoo! LtR Challenge & Chapelle et al. \cite{Chapelle2011a} & 102\\
+LETOR 3.0 & Qin et al. \cite{Qin2010} & 105 & WCL2R & Alc{\^a}ntara et al. \cite{Alcantara2010} & 2\\
+AOL dataset & Pass et al. \cite{Pass2006} & 339 &&\\
 \bottomrule
 \end{tabular}
-\caption{Forward references of learning to rank benchmark papers}
-\label{tbl:ltr_benchmark_forref}}
+}
 \end{table}
-
 The LETOR 4.0, MSLR-web10/30k and Yandex Internet Mathematics Competition 2009 benchmark collections are not accompanied by a paper.
 To collect evaluation results for learning to rank methods on these benchmarks, a Google Scholar search is performed on the name of the benchmark.
-Table~\ref{tbl:ltr_benchmark_searches} shows the results of this literature search.
-
-\begin{table}[!h]
-\scriptsize{
-\begin{tabular}{p{5.2cm}R{2.5cm}}\toprule
-Search Query & Google scholar search results \\
-\midrule
-"LETOR 4.0" & 75 \\
-"MSLR-web10k" & 16 \\
-"MSLR-web30k" & 15 \\
-"Yandex Internet Mathematics" & 3\\
-\bottomrule
-\end{tabular}
-\caption{Google scholar search results for learning to rank benchmarks}
-\label{tbl:ltr_benchmark_searches}}
-\end{table}
-
+The literature search queries returned the following numbers of results:
+\begin{description}[noitemsep,nolistsep]
+\item["LETOR 4.0":]75
+\item["MSLR-web10k":]16
+\item["MSLR-web30k":]15
+\item["Yandex Internet Mathematics":]3
+\end{description}
+\vspace{-0.08in}
 \subsection{Literature Selection}
@@
 Occurrences of L2, L3 and L4 in Table 4 imply that these algorithms are evaluated as official LETOR 2.0, 3.0 and 4.0 baselines respectively.\\
-
 Some studies with evaluation results found through the literature search procedure were not usable for the meta-analysis.
@@
 Between brackets are the studies that these properties apply to.
-
 \begin{enumerate}[noitemsep,nolistsep]
@@
 \end{enumerate}
-
+\vspace{-0.08in}
 \section{A Methodology for Comparing Learning to Rank Methods Cross-Benchmark}
@@
 Winning Number is defined as the number of other algorithms that an algorithm can beat over the set of datasets, or more formally\\
-
-$\text{WN}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$\\
-
+$\mbox{WN}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m I_{\{M_i(j)>M_k(j)\}}$\\
 where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as NDCG or MAP), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that\\
-
 $I_{\{M_i(j)>M_k(j)\}} = \begin{cases}
-1 & \text{if } M_i(j) > M_k(j), \\
-0 & \text{otherwise}
+1 & \mbox{if } M_i(j) > M_k(j), \\
+0 & \mbox{otherwise}
 \end{cases}$\\
-
 The LETOR 3.0 was a comparison on a \emph{dense} set of evaluation results, in the sense that there were evaluation results available for all learning to rank algorithms on all datasets included in their comparison.
@@
 We will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as\\
-
 $I^{'}_{M_i(j)>M_k(j)} = \begin{cases}
@@
 \end{cases}$\\
-
 From now on this adjusted version of Winning Number will be references to as \emph{Normalized Winning Number (NWN)}. The mathematical definition of Normalized Winning Number is\\
-
 $\text{NWN}_i(M) = \frac{\text{WN}_i(M)}{\text{IWN}_i(M)}$\\
-
 \noindent
 where $IWN$ is the Ideal Winning Number, defined as\\
-
 $\text{IWN}_i(M) = \sum\nolimits_{j=1}^n \sum\nolimits_{k=1}^m D_{\{M_i(j),M_k(j)\}}$\\
-
 where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as NDCG or MAP), and $D_{\{M_i(j),M_k(j)\}}$ is an evaluation definition function such that\\
-
 $D_{\{M_i(j),M_k(j)\}} = \begin{cases}
@@
 \end{cases}$\\
-
 NDCG@$\{3,5,10\}$ and MAP are the most frequently used evaluation metrics in the used benchmark collections combined, therefore we will limit our meta-analysis to evaluation results reported in one of these four metrics.
-
 \section{Results of Learning to Rank Comparison}
@@
 Table \ref{tab:raw_data} in the appendix provides raw NWN data for the learning to rank methods at NDCG@$\{3,5,10\}$ and MAP and their cross-metric weighted average.
-
 \subsection{NDCG@3}
-\begin{figure}
+Figure \ref{fig:normalized_winning_number_ndcg3} shows the NWN of learning to rank methods based on NDCG@3 results.
+LambdaNeuralRank and CoList both acquired a NWN score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
+LARF and LRUF scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which results in more certainty on these methods' performance because they are validated on three datasets that additionally are more relevant than AOL and Yahoo Set 2 (number of evaluation results for LETOR 3.0 are higher than those for AOL and Yahoo set 2).
+FenchelRank, OWPC, SmoothRank, DCMP and ListNet are ordered decreasingly by NWN and at the same time increasingly in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.
+\begin{figure}[h]
 \centering
@@
 \end{figure}
-Figure \ref{fig:normalized_winning_number_ndcg3} shows the NWN of learning to rank methods based on NDCG@3 results.
-LambdaNeuralRank and CoList both acquired a NWN score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
-LARF and LRUF scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which results in more certainty on these methods' performance because they are validated on three datasets that additionally are more relevant than AOL and Yahoo Set 2 (number of evaluation results for LETOR 3.0 are higher than those for AOL and Yahoo set 2).
-FenchelRank, OWPC, SmoothRank, DCMP and ListNet are ordered decreasingly by NWN and at the same time increasingly in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
-
+\vspace{-0.08in}
 LambdaNeuralRank, CoList, LARF, LRUF, OWPC and DCMP evaluation results are all based on one study, therefore are subjected to the risk of one overly optimistic study producing those results.
@@
 SmoothRank and ListNet have the most reliable evaluation result source, as they were official LETOR baseline runs.
-
 \subsection{NDCG@5}
-\begin{figure}
-\includegraphics[scale=0.19]{gfx/ndcg5_winnum}
-\caption{NDCG@5 comparison of 87 learning to rank methods}
-\label{fig:normalized_winning_number_ndcg5}
-\end{figure}
 Figure \ref{fig:normalized_winning_number_ndcg5} shows the NWN of learning to rank methods based on NDCG@5 results.
@@
 Like in the NDCG@3 comparison, SmoothRank and ListNet can be regarded as most reliable results because the evaluation measurements for these methods are based on LETOR official baselines.
-
+\vspace{-0.08in}
+\begin{figure}
+\centering
+\includegraphics[scale=0.19]{gfx/ndcg5_winnum}
+\caption{NDCG@5 comparison of 87 learning to rank methods}
+\label{fig:normalized_winning_number_ndcg5}
+\end{figure}
+\vspace{-0.2in}
 \subsection{NDCG@10}
+Figure \ref{fig:normalized_winning_number_ndcg10} shows the NWN of learning to rank methods based on NDCG@10 results.
+LambdaMART and LambdaNeuralRank score a NWN of 1.0 on the NDCG@10 comparison.
+For LambdaNeuralRank these results are again based on AOL dataset measurements.
+LambdaMART showed the highest NDCG@10 performance for the MSLR-WEB10k dataset.
+The set of algorithms for which there is no other algorithm with both a higher NWN and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and @5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
+In contrast to the NDCG@3 and @5 comparison, DCMP is not among the best algorithms.
+\vspace{-0.08in}
 \begin{figure}
@@
 \end{figure}
-
-Figure \ref{fig:normalized_winning_number_ndcg10} shows the NWN of learning to rank methods based on NDCG@10 results.
-LambdaMART and LambdaNeuralRank score a NWN of 1.0 on the NDCG@10 comparison.
-For LambdaNeuralRank these results are again based on AOL dataset measurements.
-LambdaMART showed the highest NDCG@10 performance for the MSLR-WEB10k dataset.
-The set of algorithms for which there is no other algorithm with both a higher NWN and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and @5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
-In contrast to the NDCG@3 and @5 comparison, DCMP is not among the best algorithms.
-
+\vspace{-0.1in}
 \subsection{MAP}
+Figure \ref{fig:normalized_winning_number_map} shows the NWN of learning to rank methods based on MAP results.
+Comparisons on the NDCG metrics where highly in agreement on the best performing algorithms, MAP-based NWN results show different results.
+RankDE scores a NWN of 1.0 on one dataset, which is achieved by obtaining highest MAP-score on the LETOR 2.0 TD2003 which has many evaluation results are evaluated.
 \begin{figure}
@@
 There is a lot of certainty on the accuracy of RankBoost and RankSVM as both models are evaluated on the majority of datasets included in the comparison for the MAP metric, but given their NWN it can said that both methods are not within the top performing learning to rank methods.
-
 \subsection{Cross-Metric}
+Figure \ref{fig:normalized_winning_number_all} shows the NWN as function of IWN for the methods described in Table 4.
+The cross-metric comparison is based on the NDCG@$\{3,5,10\}$ and MAP comparisons combined, which justifies analyzing the comparison more thoroughly.
+Figure \ref{fig:normalized_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
+In addition, Linear Regression and the ranking method of simply sorting on the best single feature are labelled as baselines.
 \begin{figure}
@@
 \end{figure}
-Figure \ref{fig:normalized_winning_number_all} shows the NWN as function of IWN for the methods described in Table 4.
-The cross-metric comparison is based on the NDCG@$\{3,5,10\}$ and MAP comparisons combined, which justifies analyzing the comparison more thoroughly.
-Figure \ref{fig:normalized_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
-In addition, Linear Regression and the ranking method of simply sorting on the best single feature are labelled as baselines.\\
-
+\vspace{-0.1in}
 LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both IWN and NWN.
@@
 This was to be expected, because the cross-metric comparison input data of three NDCG entries (@3, @5, and @10) enables it to have up to three times as many as many weight as the MAP comparison.\\
-
 LARF, IPRank and DCMP and several variants of RankSVM performed very well on the cross-metric comparison, with all having only one method in its top right quadrant.
@@
 Some of the DirectRank measurements could not be used because measurements on some datasets were only available in graphical form and not in raw data.\\
-
 LAC-MR-OR and RE-QR showed very good ranking accuracy in the MAP comparison on multiple datasets.
 Because LAC-MR-OR is only evaluated on two datasets for NDCG@10 and RE-QR is not evaluated for NDCG at all, LAC-MR-OR and RE-QR are not within the top performing methods in the cross-metric comparison.
-
 \section{Limitations}
@@
 It could be the case however that this assumption is false and that the accurateness of the learning to rank methods on a dataset is not dataset independent.\\
-
 A second limitation is that the datasets on which learning to rank methods have been evaluated cannot always be regarded a random choice.
 It might be the case that some researchers chose to publish results for exactly those benchmark datasets that showed the most positive results for their learning to rank method.\\
-
 Another limitation is that our comparison methodology relies on the correctness of the evaluation results found in the literature search step.
@@
 By taking IWN into account in Figure \ref{fig:normalized_winning_number_all} we further mitigate this limitation, as IWN is loosely related with the number of studies that reported evaluation results for an algorithm.\\
-
 Our comparison regarded evaluation results on NDCG@$\{3,5,10\}$ and MAP.
@@
 In addition, papers with a high number of evaluation results are not expected to be less reliable than papers with fewer evaluation results.
-
 \section{Contributions}
@@
 Our comparison methodology comprises of two components: 1) NWN, which provides insight in the ranking accuracy of the learning to rank method, and 2) IWN, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.\\
-
 Based on our literature search for evaluation results on well-known benchmarks collections, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.
@@
 From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.\\
-
 More evaluation runs are needed for the methods on the left side of Figure \ref{fig:normalized_winning_number_all}.
 Our work contributes to this by identifying promising learning to rank methods that researchers could focus on in performing additional evaluation runs.
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-\balance
-\bibliographystyle{abbrv}
-{\scriptsize\bibliography{Bibliography}}
-\onecolumn
-% sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%APPENDICES are optional
-\appendix
+\bibliographystyle{splncs03}
+\bibliography{Bibliography}
 
+\newpage
+\section{Appendix}
+\vspace{-0.3in}
 \begin{table*}[!hpt]
-\scriptsize{
+\centering
+{\fontsize{7.6}{6.9}\selectfont
+\caption{Learning to rank algorithms with measurements on benchmark datasets}
+\label{tab:ltr_methods_used}
 \begin{tabular}{lll|lll}\toprule
@@
 \end{tabular}
-\caption{Learning to rank algorithms with measurements on benchmark datasets}}
-\label{tab:ltr_methods_used}
+}
 \end{table*}
-
-\scriptsize{
-\begin{longtable}[!hpt]{@{}lrrrrrrrrrrrrrrrr@{}}\toprule
+\vspace{-0.3in}
+{\fontsize{7.6}{6.9}\selectfont
+\centering
+\begin{longtable}[!hpt]{@{}lrrrrrrrrrrrrrrrr@{}}
+\toprule
 & \multicolumn{2}{c}{NDCG@3} & \phantom{a} 

### 2014-09-02 implementation/java/src/letor/parallel/ListNetCluster.java

@@ -23,11 +23,11 @@ public class ListNetCluster {
     private static final int      k          = 10; // NDCG@k
 
     // Initialise paralellisation parameters
-    private static int  dataNodes                     = 8;
+    private static int  dataNodes                     = 24;
     private static int  availableMappers              = 4*dataNodes;
     private static int  availableReducers             = 2*dataNodes;
 
-    private static int duplicationNumber              = 9; // only used for CUSTOM
+    private static int duplicationNumber              = 49; // only used for CUSTOM
     private static Metadata metadata                  = DataSets.getMetaData(DATASET, duplicationNumber);
     private static long MAX_TRAIN_SIZE                = metadata.getMax_train_size(); // ohsumed: 5151958, MQ2007: 25820919, MQ2008: 5927007, MSLR-WEB10K: 838011150, MSLR-WEB30K:
     private static long MAX_VALI_SIZE                 = metadata.getMax_vali_size(); // ohsumed: 1764005, MQ2007:  8753466, MQ2008: 2237346, MSLR-WEB10K: 280714022, MSLR-WEB30K:
@@ -40,12 +40,12 @@ public class ListNetCluster {
 
     public static void main(String[] args) throws Exception {
         // Cluster configuration
-        String clusterName          = "ltr8";
+        String clusterName          = "ltr24";
         String clusterUser          = "admin";
         String clusterPassword      = "Qw!23456789";
-        String storageAccount       = "ltrstorage";
-        String containerName        = "ltrcontainer";
-        String storageAccountKey    = "s7nkBIp0I6BUXJgxoqFGFjhsUILBQP54jpKJLS2fI8i9KYC1I6UJP7JjjplfHEFvJyX1n7PVdRnkT27KyzvdjQ==";
+        String storageAccount       = "ltrmini";
+        String containerName        = "ltrsmall";
+        String storageAccountKey    = "0eUNzH04rVPGDOIVIXtSqwIlvmIQ/pKQgW55r9CcJSA9RWLVTcDfGzalgu33k9OrP6nXvj++IRRHpOhoJrDHMA==";
         AzurePigWrapper apw         = new AzurePigWrapper(clusterName, containerName, clusterUser, clusterPassword, storageAccount, storageAccountKey, availableReducers);
 
         ArrayList<String> pigLines  = new ArrayList<String>();

### 2014-09-03 report/Chapters/Implementation.tex
voortgang thesis + nieuwe metingen

@@
 
-\section{The HDInsight platform}
-TODO:
- - No direct access to jobtracker
- - C# API and access via Windows PowerShell scripts
- - REST endpoints for Templeton/WebHcat / Ambari / HiveServer / HiveServer2 and Oozie
- - Pig for ease of implementation
- - Templeton and Oozie allow submission of Pig jobs
- - Oozie requires Pig jobs to be stored on HDFS, Templeton allows running dynamically generated Pig from code.
+\section{Architecture}
+Ranking models almost without exception consist of a sequence of operations on the input data, that are in most cases repeatedly executed in a series of iterations.
+Apache Pig will be used to implement the sequence of operation on the input data, because the higher level of abstraction that Apache Pig offers makes it easier to focus solely on the data operations as compared to native Hadoop MapReduce.\\
 
-\section{Algorithm implementation}
+% TODO: Stuk over werkzaamheden framework uitbreiden: uitlezen resultaten (evaluatieresultaten + doorgeven tussenresultaten (die tweede wellicht later in het document pas introduceren))
+In a cross-validation evaluation experiment of a ranking model, the handling of which folds are loaded for training, evaluation and testing is a task that need to be taken care of regardless of the ranking model.
+Given that most ranking models are of an iterative nature, the task of iteration handling is also a task that need to be done for (most) models.
+Since these tasks need to be considered for (almost) all ranking models it seems natural to develop a framework that takes care of these tasks.
+The aim of this framework is to let implementations of ranking models focus solely on implementing the sequential steps of one iteration of the algorithm, while the framework arranges that 1) these sequential steps are performed iteratively and 2) these sequential steps are performed on the multiple training folds of data.
+This framework that handles folding and iterations will be implemented in Java and will work such that fold- and iteration dependent parts of the Pig code will be generated dynamically by the Java framework after which the Pig job will be sent to the cluster.\\
+
+\subsection{Microsoft Azure HDInsight}
+Microsoft offers a scalable and on-demand Hadoop service with HDInsight, that enables Hadoop services for those not able to make the required investments for their own cluster.
+The latest HDInsight of HDInsight available at the time of writing, HDInsight 3.1, runs the Hortonworks distribution of Hadoop version 2.4.
+HDInsight 3.1 offers multiple ways of submitting Hadoop jobs to a HDInsight cluster, described in Table \ref{tbl:hdinsight_endpoints}.\\
+\begin{table}
+\centering
+\begin{tabular}{p{4cm}p{2.8cm}p{5.5cm}}\toprule
+Job submission method & Type & Description \\
+\midrule
+Powershell & Powershell scripts & The Azure module for Windows PowerShell enables direct submission of Hadoop jobs through PowerShell cmdlets.\\
+C\# API & C\# API & A wrapper API is offered to submit Hadoop MapReduce jobs directly from C\# code.\\
+HiveServer/HiveServer2 & REST endpoint & HiveServer and its successor HiveServer 2 are, as its name suggests, REST endpoints that allow remote submission of Hive queries.\\
+Oozie & REST endpoint & Apache Oozie is a workflow scheduler to manage Apache Hadoop jobs.
+Oozie enables users to specify Directed Acyclical Graphs of action, where each action is specified in either MapReduce or Pig.\\
+WebHCat/Templeton & REST endpoint & WebHCat, formerly known as Templeton, is a REST API for HCatalog, a table and storage management layer for Hadoop.
+WebHCat allows users to use either Apache Pig, Apache MapReduce or Apache Hive for data processing.\\
+\bottomrule
+\end{tabular}
+\caption{HDInsight REST endpoints for job submission}
+\label{tbl:hdinsight_endpoints}
+\end{table}
+Oozie and WebHCat/Templeton are the two methods for job submission in Table \ref{tbl:hdinsight_endpoints} that both 1) support Apache Pig jobs, and 2) Can be used from within Java code.
+Below the necessary procedures for job submission from Oozie as well as from WebHCat will be sketched.\\
+
+\begin{table}
+\centering
+\begin{tabular}{p{5cm}p{5cm}}\toprule
+Job submission procedure for Apache Oozie & Job submission procedure for WebHCat/Templeton \\
+\midrule
+1. Let the framework build the Pig job dynamically. & 1. Let the framework build the Pig job dynamically.\\
+2. Encapsulate the Pig job in an Oozie workflow. & 2. Submit the Pig job through the WebHCat/Templeton REST API.\\
+3. Upload the Oozie workflow to HDFS storage. & \\
+4. Execute the Oozie workflow through the Oozie REST API. & \\
+\bottomrule
+\end{tabular}
+\caption{Comparison of Oozie and WebHCat job submission procedures}
+\label{tbl:oozie_templeton}
+\end{table}
+
+Table \ref{tbl:oozie_templeton} shows how Oozie job submission is more complex then for the case of dynamically generated jobs than WebHCat/Templeton.
+Oozie is more fitting for static Hadoop jobs that require a workflow consisting of a sequence of Pig and MapReduce jobs mixed together, but is a lesser fit for our situation.
+Templeton will be used for submission of the Pig jobs, as the HDInsight version that was available at the start of the implementation did not yet support WebHCat.
+
 \subsection{ListNet}

### 2014-09-03 implementation/java/src/letor/parallel/ListNetCluster.java

@@ -23,7 +23,7 @@ public class ListNetCluster {
     private static final int      k          = 10; // NDCG@k
 
     // Initialise paralellisation parameters
-    private static int  dataNodes                     = 24;
+    private static int  dataNodes                     = 32;
     private static int  availableMappers              = 4*dataNodes;
     private static int  availableReducers             = 2*dataNodes;
 
@@ -40,7 +40,7 @@ public class ListNetCluster {
 
     public static void main(String[] args) throws Exception {
         // Cluster configuration
-        String clusterName          = "ltr24";
+        String clusterName          = "ltr32";
         String clusterUser          = "admin";
         String clusterPassword      = "Qw!23456789";
         String storageAccount       = "ltrmini";
@@ -77,7 +77,11 @@ public class ListNetCluster {
                 int hundreds  = (duplicationNumber % 1000) / 100;
                 int tens      = (duplicationNumber % 100) / 10;
                 int units     = duplicationNumber % 10;
-                pigLines.add("TRAIN = LOAD '" + pathPrefix + "/input/" + metadata.getName() + "/Fold" + fold + "/train[0-"+thousands+"][0-"+hundreds+"][0-"+tens+"][0-"+units+"].txt' USING PigStorage(' ');");
+                String thousandsString = thousands > 0 ?  "[0-"+thousands+"]" : "0";
+                String hundredsString  = hundreds > 0 ?  "[0-"+hundreds+"]" : "0";
+                String tensString      = tens > 0 ?  "[0-"+tens+"]" : "0";
+                String unitsString     = units > 0 ? "[0-"+units+"]" : "0";
+                pigLines.add("TRAIN = LOAD '" + pathPrefix + "/input/" + metadata.getName() + "/Fold" + fold + "/train"+thousandsString+hundredsString+tensString+unitsString+".txt' USING PigStorage(' ');");
             } else {
                 pigLines.add("TRAIN = LOAD '" + pathPrefix + "/input/" + metadata.getName() + "/Fold" + fold + "/train.txt' USING PigStorage(' ');");
             }
@@ -104,7 +108,11 @@ public class ListNetCluster {
                 int hundreds  = (duplicationNumber % 1000) / 100;
                 int tens      = (duplicationNumber % 100) / 10;
                 int units     = duplicationNumber % 10;
[…]

### 2014-09-03 report/Chapters/Implementation.tex
preprocessing stap in thesis

@@
 HDInsight 3.1 offers multiple ways of submitting Hadoop jobs to a HDInsight cluster, described in Table \ref{tbl:hdinsight_endpoints}.\\
+
 \begin{table}
@@
 \end{table}
+
 Oozie and WebHCat/Templeton are the two methods for job submission in Table \ref{tbl:hdinsight_endpoints} that both 1) support Apache Pig jobs, and 2) Can be used from within Java code.
@@
 \subsection{ListNet}
+The following sections describe the Pig jobs that form the three independent parts of the ListNet ranking model: 1) Preprocessing, 2) Training (this includes evaluation over the validation set) and 3) Testing (evaluation over the test set).
+\subsubsection{Preprocessing: feature scaling}
+The preprocessing phase consists of two separate Pig jobs.
+The first Pig job determines the minimum and the maximum values per feature in the training set.
+The second Pig job rescales each feature of the train, validation and test datasets using the following formula for rescaling:
+\begin{equation}
+x^{'} = \frac{x-min(x)}{max(x)-min(x)}
+\end{equation}
+This rescaling procedure set the range of each feature to $[0,1]$.
+
+The first Pig job:
+\begin{lstlisting}
+REGISTER wasb:///user/hdp/lib/*.jar;
+TRAIN = LOAD '[path prefix]/input/[dataset name]/Fold[fold number]/train.txt' USING PigStorage(' ');
+TRAIN_STD = FOREACH TRAIN GENERATE flatten(udf.util.ToStandardForm(*));
+TRAIN_STD_BY_QUERY = GROUP TRAIN_STD BY $1 PARALLEL [available reducers];
+MIN_MAX = FOREACH TRAIN_STD_BY_QUERY GENERATE flatten(udf.util.GetMinMax(*));
+MIN_MAX_GRPD = GROUP MIN_MAX ALL;
+MIN_MAX_FIN = FOREACH MIN_MAX_GRPD GENERATE flatten(udf.util.CombineMinMax(*));
+STORE MIN_MAX_FIN INTO 'minmax[fold number]';
+\end{lstlisting}
+The second Pig job:
+\begin{lstlisting}
+REGISTER wasb:///user/hdp/lib/*.jar;
+TRAIN = LOAD '[path prefix]/input/[dataset name]/Fold[fold number]/train.txt' USING PigStorage(' ');
+VALIDATE = LOAD '[path prefix]/input/[dataset name]/Fold[fold number]/vali.txt' USING PigStorage(' ');
+TEST = LOAD '[path prefix]/input/[dataset name]/Fold[fold number]/test.txt' USING PigStorage(' ');
+TRAIN_STD = FOREACH TRAIN GENERATE flatten(udf.util.ToStandardForm(*));
+VALIDATE_STD = FOREACH VALIDATE GENERATE flatten(udf.util.ToStandardForm(*));
+TEST_STD = FOREACH TEST GENERATE flatten(udf.util.ToStandardForm(*));
+DEFINE ScaleFeatures udf.util.ScaleFeatures('"+LtrUtils.toParamString(minmaxList)+"');
+TRAIN_SCA = FOREACH TRAIN_STD GENERATE flatten(ScaleFeatures(*));
+VALIDATE_SCA = FOREACH VALIDATE_STD GENERATE flatten(ScaleFeatures(*));
+TEST_SCA = FOREACH TEST_STD GENERATE flatten(ScaleFeatures(*));
+STORE TRAIN_SCA INTO 'train_sca[fold number]' USING BinStorage();
+STORE VALIDATE_SCA INTO 'validate_sca[fold number]' USING BinStorage();
+STORE TEST_SCA INTO 'test_sca[fold number]' USING BinStorage();
+\end{lstlisting}
+\subsubsection{}

### 2014-09-03 report/Chapters/Implementation.tex
udf descriptions

@@
 \end{lstlisting}
+
+\begin{table}
+\centering
+\begin{tabular}{p{5cm}p{8cm}}\toprule
+UDF & Description \\
+\midrule
+udf.util.ToStandardForm() & Transforms the data set into the standard form of relevance label in first column followed by feature values. Strips data of any other columns, if present.\\
+udf.util.GetMinMax() & Extracts the minimum and maximum value per feature, for the documents of a single query.\\
+udf.util.CombineMinMax() & Combines outputs of the udf.util.GetMinMax() UDF for each query into globally minimum and maximum feature values.\\
+\bottomrule
+\end{tabular}
+\caption{Description of preprocessing User Defined Functions (1/2)}
+\label{tbl:preprocessing_udfs_1}
+\end{table}
+
 The second Pig job:
@@
 \end{lstlisting}
-\subsubsection{}
+
+\begin{table}
+\centering
+\begin{tabular}{p{5cm}p{8cm}}\toprule
+UDF & Description \\
+\midrule
+udf.util.ToStandardForm() & See Table \ref{tbl:preprocessing_udfs_1} for description.\\
+udf.util.ScaleFeatures() & Uses the minimum and maximum feature values with which it is parametrised perform the following rescaling transformation to the features: $x^{'} = \frac{x-min(x)}{max(x)-min(x)}$.\\
+\bottomrule
+\end{tabular}
+\caption{Description of preprocessing User Defined Functions (2/2)}
+\label{tbl:preprocessing_udfs_2}
+\end{table}

### 2014-09-08 report/Chapters/Implementation.tex
implementation voortgang

@@
 The following sections describe the Pig jobs that form the three independent parts of the ListNet ranking model: 1) Preprocessing, 2) Training (this includes evaluation over the validation set) and 3) Testing (evaluation over the test set).
-\subsubsection{Preprocessing: feature scaling}
+\subsubsection{Preprocessing}
 The preprocessing phase consists of two separate Pig jobs.
@@
 \begin{lstlisting}
-REGISTER wasb:///user/hdp/lib/*.jar;
+REGISTER [path prefix]/lib/*.jar;
 TRAIN = LOAD '[path prefix]/input/[dataset name]/Fold[fold number]/train.txt' USING PigStorage(' ');
@@
 TEST_STD = FOREACH TEST GENERATE flatten(udf.util.ToStandardForm(*));
-DEFINE ScaleFeatures udf.util.ScaleFeatures('"+LtrUtils.toParamString(minmaxList)+"');
+DEFINE ScaleFeatures udf.util.ScaleFeatures('[array with minimum and maximum feature values]');
 TRAIN_SCA = FOREACH TRAIN_STD GENERATE flatten(ScaleFeatures(*));
@@
 \end{table}
+
+\subsubsection{Training}
+The first Pig job:
+\begin{lstlisting}
+REGISTER [path prefix]/lib/*.jar;
+DEFINE QueryLossGradient udf.listnet.QueryLossGradient('[feature dimensionality of data set]');
+DEFINE ExpRelOurScores udf.listnet.ExpRelOurScores('[neural network weights & iteration number]');
+(1) TRAIN_SCA = LOAD 'train_sca[fold number]/*' USING BinStorage();
+(1) TR_BY_QUERY = GROUP TRAIN_SCA BY $1 PARALLEL [number of avaiable reducers];
+(1) TR_EXP_REL_SCORES = FOREACH TR_BY_QUERY GENERATE flatten(ExpRelOurScores(TRAIN_SCA));
+(1) STORE TR_EXP_REL_SCORES INTO 'tr_exp_rel_scores-f[fold number]' USING BinStorage();
+(*) TR_EXP_REL_SCORES = LOAD 'tr_exp_rel_scores-f[fold number]/*' USING BinStorage();
+(*) TR_EXP_REL_SCORES = FOREACH TR_EXP_REL_SCORES GENERATE flatten(ExpRelOurScores(*)) PARALLEL [number of available reducers];
+TR_QUERY_LOSS_GRADIENT = FOREACH TR_EXP_REL_SCORES GENERATE flatten(QueryLossGradient(*)) PARALLEL [number of available reducers];
+TR_QUERY_LOSS_GRADIENT_GRPD = GROUP TR_QUERY_LOSS_GRADIENT ALL;
+TR_LOSS_GRADIENT = FOREACH TR_QUERY_LOSS_GRADIENT_GRPD GENERATE flatten(udf.listnet.MultiSum(*));
+STORE TR_LOSS_GRADIENT INTO 'tr_loss_gradient-f[fold number]i[iteration number]';
+\end{lstlisting}
+
+\begin{table}
+\centering
+\begin{tabular}{p{5cm}p{8cm}}\toprule
+UDF & Description \\
+\midrule
+udf.listnet.QueryLossGradient() & Calculates the Cross Entropy loss for a query and calculates the gradients per feature based on this query.\\
+udf.listnet.ExpRelOurScores() & Calculates the predicted relevance label of a query based on the current model weights and transforms this following the transformation $x -> e^{x}$.
+In case the current iteration is the first iteration, the same transformation is applied to the ground truth relevance label.\\
+udf.listnet.MultiSum() & Calculates aggregated loss and feature gradients by summing the per-query losses and per-query feature gradients.\\
+\bottomrule
+\end{tabular}
+\caption{Description of training User Defined Functions (1/2)}
+\label{tbl:training_udfs_1}
+\end{table}
+
+The second Pig job of the training stage tests the performance of the model weights that were trained in the first Pig job on the validation data set.
+The second Pig job:
+\begin{lstlisting}
+REGISTER [path prefix]/lib/*.jar;
+DEFINE Ndcg udf.util.Ndcg('[neural network weights & NDCG cut-off parameter]');
+(1) VALIDATE_SCA = LOAD 'validate_sca[fold number]/*' USING BinStorage();
+(1) VA_BY_QUERY = GROUP VALIDATE_SCA BY $1 PARALLEL [number of available reducers];
+(1) STORE VA_BY_QUERY INTO 'va_by_query-f[fold number]' USING BinStorage();
+(*) VA_BY_QUERY = LOAD 'va_by_query-f[fold number]/*' USING BinStorage();
+NDCG = FOREACH VA_BY_QUERY GENERATE Ndcg(*);
+NDCG_GRPD = GROUP NDCG ALL;
+AVG_NDCG = FOREACH NDCG_GRPD GENERATE AVG(NDCG);
+STORE AVG_NDCG INTO 'avg_ndcg-f[fold number]i[iteration number]';
+\end{lstlisting}
+\begin{table}
+\centering
+\begin{tabular}{p{5cm}p{8cm}}\toprule
+UDF & Description \\
+\midrule
+udf.util.Ndcg() & Calculates \ac{nDCG}@k for a query.\\
+\bottomrule
+\end{tabular}
+\caption{Description of training User Defined Functions (2/2)}
+\label{tbl:training_udfs_2}
+\end{table}
+
+\subsubsection{Testing}

### 2014-09-08 report/Chapters/Implementation.tex
duidelijker

@@
 DEFINE ExpRelOurScores udf.listnet.ExpRelOurScores('[neural network weights & iteration number]');
-(1) TRAIN_SCA = LOAD 'train_sca[fold number]/*' USING BinStorage();
-(1) TR_BY_QUERY = GROUP TRAIN_SCA BY $1 PARALLEL [number of avaiable reducers];
-(1) TR_EXP_REL_SCORES = FOREACH TR_BY_QUERY GENERATE flatten(ExpRelOurScores(TRAIN_SCA));
-(1) STORE TR_EXP_REL_SCORES INTO 'tr_exp_rel_scores-f[fold number]' USING BinStorage();
-(*) TR_EXP_REL_SCORES = LOAD 'tr_exp_rel_scores-f[fold number]/*' USING BinStorage();
-(*) TR_EXP_REL_SCORES = FOREACH TR_EXP_REL_SCORES GENERATE flatten(ExpRelOurScores(*)) PARALLEL [number of available reducers];
+[FIRST TRAINING ITERATION:]
+	TRAIN_SCA = LOAD 'train_sca[fold number]/*' USING BinStorage();
+	TR_BY_QUERY = GROUP TRAIN_SCA BY $1 PARALLEL [number of avaiable reducers];
+	TR_EXP_REL_SCORES = FOREACH TR_BY_QUERY GENERATE flatten(ExpRelOurScores(TRAIN_SCA));
+	STORE TR_EXP_REL_SCORES INTO 'tr_exp_rel_scores-f[fold number]' USING BinStorage();
+[SUBSEQUENT TRAINING ITERATIONS:]
+	TR_EXP_REL_SCORES = LOAD 'tr_exp_rel_scores-f[fold number]/*' USING BinStorage();
+	TR_EXP_REL_SCORES = FOREACH TR_EXP_REL_SCORES GENERATE flatten(ExpRelOurScores(*)) PARALLEL [number of available reducers];
 TR_QUERY_LOSS_GRADIENT = FOREACH TR_EXP_REL_SCORES GENERATE flatten(QueryLossGradient(*)) PARALLEL [number of available reducers];
@@
 DEFINE Ndcg udf.util.Ndcg('[neural network weights & NDCG cut-off parameter]');
-(1) VALIDATE_SCA = LOAD 'validate_sca[fold number]/*' USING BinStorage();
-(1) VA_BY_QUERY = GROUP VALIDATE_SCA BY $1 PARALLEL [number of available reducers];
-(1) STORE VA_BY_QUERY INTO 'va_by_query-f[fold number]' USING BinStorage();
-(*) VA_BY_QUERY = LOAD 'va_by_query-f[fold number]/*' USING BinStorage();
+[FIRST TRAINING ITERATION:]
+	VALIDATE_SCA = LOAD 'validate_sca[fold number]/*' USING BinStorage();
+	VA_BY_QUERY = GROUP VALIDATE_SCA BY $1 PARALLEL [number of available reducers];
+	STORE VA_BY_QUERY INTO 'va_by_query-f[fold number]' USING BinStorage();
+[SUBSEQUENT TRAINING ITERATIONS:]
+	VA_BY_QUERY = LOAD 'va_by_query-f[fold number]/*' USING BinStorage();
 NDCG = FOREACH VA_BY_QUERY GENERATE Ndcg(*);
@@
 \subsubsection{Testing}
+Pig Job:
+\begin{lstlisting}
+REGISTER [path prefix]/lib/*.jar;
+TEST_SCA = LOAD 'test_sca[fold number]/*' USING BinStorage();
+TE_BY_QUERY = GROUP TEST_SCA BY $1 PARALLEL [number of available reducers];
+DEFINE Ndcg udf.util.Ndcg('[neural network weights & NDCG cut-off parameter]');
+NDCG = FOREACH TE_BY_QUERY GENERATE Ndcg(*);
+NDCG_GRPD = GROUP NDCG ALL;
+AVG_NDCG = FOREACH NDCG_GRPD GENERATE AVG(NDCG);
+STORE AVG_NDCG INTO 'avg_ndcg';
+\end{lstlisting}

### 2014-09-08 report/Chapters/Implementation.tex
verder aan implementation

@@
 \end{equation}
-This rescaling procedure set the range of each feature to $[0,1]$.
+This rescaling procedure sets the values of all features to be within range $[0,1]$.\\
 
@@
 \end{lstlisting}
+
+The minimum and maximum values per feature stored in MIN\_MAX\_FIN can now be read from \ac{HDFS} and used within the Java framework by using the Windows Azure Storage library.
+The second Pig job uses these minimum and maximum values by parametrising the \ac{UDF} that performs the rescaling transformation, udf.util.ScaleFeatures(), with the minimum and maximum feature values.\\
 

### 2014-09-08 report/Chapters/Implementation.tex
minipage voor listings

@@
 
-The first Pig job:
+The first Pig job:\\
+\begin{minipage}{\linewidth}
 \begin{lstlisting}
@@
 \end{lstlisting}
+\end{minipage}
 
@@
 
-The second Pig job:
+The second Pig job:\\
+\begin{minipage}{\linewidth}
 \begin{lstlisting}
@@
 \subsubsection{Training}
-The first Pig job:
+The training stage, like the preprocessing stage, consists of two separate Pig jobs.
+The first Pig job calculates the Cross Entropy loss on training data of the current model and calculates the gradients for the next model update.
+The second Pig job is an internal validation step that validates the model on the validation set by calculating \ac{nDCG}@k.\\
+
+The first Pig job:\\
+\begin{minipage}{\linewidth}
 \begin{lstlisting}
@@
 
-The second Pig job of the training stage tests the performance of the model weights that were trained in the first Pig job on the validation data set.
-The second Pig job:
+The second Pig job of the training stage validates the performance of the model weights that were trained in the first Pig job on the validation data set.\\
+
+The second Pig job:\\
+\begin{minipage}{\linewidth}
 \begin{lstlisting}
@@
 \end{lstlisting}
+\end{minipage}
+
 \begin{table}
@@
 \subsubsection{Testing}
-Pig Job:
+The testing stage tests the best model found in the training iterations, selected on validation set \ac{nDCG}@k (as calculated in the second Pig job of the training stage), by calculating the \ac{nDCG}@k of this model on the test set.\\
+
+Test stage Pig job:\\
+\begin{minipage}{\linewidth}
 \begin{lstlisting}
@@
 \end{lstlisting}
+\end{minipage}
+
+\subsection{SmoothRank}

### 2014-09-08 report/Chapters/Implementation.tex
raw data file + visualisation R file

@@
+\label{chap:implementation}
 This chapter describes implementation details of the Learning-to-Rank algorithm parallel implementations in HDInsight that are either Hadoop, Microsoft Azure, or HDInsight and are not part of the algorithm specification itself.

### 2014-09-08 report/Chapters/Results.tex
raw data file + visualisation R file

This chapter describes the results of the running time measurements of the cluster and baseline implementations as described in \ref{chap:implementation}.

### 2014-09-08 measurements/visualisation.R
visualisation

@@ -1,0 +1,10 @@
+setwd("C:/Git-data/Afstuderen/measurements")
+library(ggplot2)
+library(scales)
+measure = read.csv("raw_data.csv") # read csv file
+name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
+name <- ifelse(name=="Serial NA", "Serial", name)
+measure$serial.parallel <- name
+d <- ggplot(data=measure, aes(x=dataSize, y=trainTime, colour=serial.parallel))
+d <- d + geom_point() + geom_smooth()
+d

### 2014-09-08 measurements/visualisation.R
x axis log scale

@@ -6,5 +6,5 @@ name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
 name <- ifelse(name=="Serial NA", "Serial", name)
 measure$serial.parallel <- name
 d <- ggplot(data=measure, aes(x=dataSize, y=trainTime, colour=serial.parallel))
-d <- d + geom_point() + geom_smooth()
+d <- d + geom_point() + geom_smooth() + scale_x_log10()
 d

### 2014-09-08 measurements/visualisation.R
mooiere visualisatie

@@ -6,5 +6,11 @@ name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
 name <- ifelse(name=="Serial NA", "Serial", name)
 measure$serial.parallel <- name
 d <- ggplot(data=measure, aes(x=dataSize, y=trainTime, colour=serial.parallel))
-d <- d + geom_point() + geom_smooth() + scale_x_log10()
+d <- d + geom_point() + geom_smooth(fill=NA) + 
+	scale_x_log10() + theme_bw() + opts(
+		panel.grid.major = theme_blank(),
+		panel.grid.minor = theme_blank(),
+		panel.border     = theme_blank(),
+		panel.background = theme_blank()
+	)
 d

### 2014-09-08 measurements/visualisation.R
alleen axis

@@ -1,16 +1,19 @@
 setwd("C:/Git-data/Afstuderen/measurements")
 library(ggplot2)
 library(scales)
+source("http://egret.psychol.cam.ac.uk/statistics/R/extensions/rnc_ggplot2_border_themes.r")
 measure = read.csv("raw_data.csv") # read csv file
 name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
 name <- ifelse(name=="Serial NA", "Serial", name)
 measure$serial.parallel <- name
 d <- ggplot(data=measure, aes(x=dataSize, y=trainTime, colour=serial.parallel))
-d <- d + geom_point() + geom_smooth(fill=NA) + 
-	scale_x_log10() + theme_bw() + opts(
-		panel.grid.major = theme_blank(),
-		panel.grid.minor = theme_blank(),
-		panel.border     = theme_blank(),
-		panel.background = theme_blank()
+d <- d + geom_point(size=2) + stat_smooth(se=FALSE) + 
+	scale_x_log10() + theme_bw() + 
+	theme(
+		panel.grid.major = element_blank(),
+		panel.grid.minor = element_blank(),
+		panel.border     = element_blank(),
+		panel.background = element_blank(),
+		axis.line = element_line(color='black')
 	)
 d

### 2014-09-08 measurements/visualisation.R
autoscale + nettere fitfunctie

@@ -1,13 +1,14 @@
 setwd("C:/Git-data/Afstuderen/measurements")
 library(ggplot2)
+library(mgcv)
 library(scales)
-source("http://egret.psychol.cam.ac.uk/statistics/R/extensions/rnc_ggplot2_border_themes.r")
 measure = read.csv("raw_data.csv") # read csv file
 name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
 name <- ifelse(name=="Serial NA", "Serial", name)
 measure$serial.parallel <- name
+maxTime <- max(measure$trainTime)
 d <- ggplot(data=measure, aes(x=dataSize, y=trainTime, colour=serial.parallel))
-d <- d + geom_point(size=2) + stat_smooth(se=FALSE) + 
+d <- d + geom_point(size=2) + stat_smooth(se=FALSE, method = "loess", formula = y ~ exp(x)) + 
 	scale_x_log10() + theme_bw() + 
 	theme(
 		panel.grid.major = element_blank(),
@@ -15,5 +16,5 @@ d <- d + geom_point(size=2) + stat_smooth(se=FALSE) +
 		panel.border     = element_blank(),
 		panel.background = element_blank(),
 		axis.line = element_line(color='black')
-	)
-d
+	) + scale_y_continuous(expand = c(0.001, maxTime/100))
+ggplot_build(d)

### 2014-09-08 measurements/faceted_graph.R
faceted graph

@@ -1,0 +1,40 @@
+setwd("C:/Git-data/Afstuderen/measurements")
+library(ggplot2)
+library(mgcv)
+library(scales)
+measure = read.csv("raw_data.csv") # read csv file
+name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
+
+getName <- function(s){
+	return(ifelse(grepl("^Parallel (\\d)+$", s), 
+			paste(
+				c(
+					"Cluster: ", 
+					str_extract(s, "\\d+"), 
+					" data nodes"
+				),
+				collapse = ''
+			), 
+			"Serial"
+		)
+	)
+}
+measure$seconds <- measure$trainTime/60
+name <- unlist(lapply(name, getName))
+measure$serial.parallel <- name
+maxTime <- max(measure$trainTime)
+d <- ggplot(data=measure, aes(x=dataSize, y=seconds, colour=serial.parallel))
+d <- d + geom_point(size=2) + stat_smooth(level=0.99, method = "glm", formula = y ~ exp(x)) + 
+	scale_x_log10() + theme_bw() + 
+	theme(
+		panel.grid.major = element_blank(),
+		panel.grid.minor = element_blank(),
+		panel.border     = element_blank(),
+		panel.background = element_blank(),
+		axis.line = element_line(color='black')
+	) + scale_y_continuous(expand = c(0.001, maxTime/10000))+	
+	facet_wrap(~serial.parallel)+
+	xlab("Dataset size (in Byte)") +
+	ylab("Training iteration time (in Seconds)") +
+	labs(colour = "Execution mode") 
+ggplot_build(d)

### 2014-09-08 measurements/visualisation.R

@@ -1,20 +1,0 @@
-setwd("C:/Git-data/Afstuderen/measurements")
-library(ggplot2)
-library(mgcv)
-library(scales)
-measure = read.csv("raw_data.csv") # read csv file
-name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
-name <- ifelse(name=="Serial NA", "Serial", name)
-measure$serial.parallel <- name
-maxTime <- max(measure$trainTime)
-d <- ggplot(data=measure, aes(x=dataSize, y=trainTime, colour=serial.parallel))
-d <- d + geom_point(size=2) + stat_smooth(se=FALSE, method = "loess", formula = y ~ exp(x)) + 
-	scale_x_log10() + theme_bw() + 
-	theme(
-		panel.grid.major = element_blank(),
-		panel.grid.minor = element_blank(),
-		panel.border     = element_blank(),
-		panel.background = element_blank(),
-		axis.line = element_line(color='black')
-	) + scale_y_continuous(expand = c(0.001, maxTime/100))
-ggplot_build(d)

### 2014-09-08 measurements/single_graph.R
shape + colour distinction

@@ -24,7 +24,7 @@ name <- unlist(lapply(name, getName))
 measure$serial.parallel <- name
 maxTime <- max(measure$trainTime)
 d <- ggplot(data=measure, aes(x=dataSize, y=seconds, colour=serial.parallel, shape=serial.parallel))
-d <- d + geom_point(size=2) + stat_smooth(level=0.99, method = "glm", formula = y ~ exp(x)) + 
+d <- d + geom_point(size=2) + stat_smooth(se=FALSE, method = "glm", formula = y ~ exp(x)) + 
 	scale_x_log10() + theme_bw() + 
 	theme(
 		panel.grid.major = element_blank(),
@@ -33,7 +33,6 @@ d <- d + geom_point(size=2) + stat_smooth(level=0.99, method = "glm", formula = 
 		panel.background = element_blank(),
 		axis.line = element_line(color='black')
 	) + scale_y_continuous(expand = c(0.001, maxTime/10000))+	
-	facet_wrap(~serial.parallel)+
 	xlab("Dataset size (in Byte)") +
 	ylab("Training iteration time (in Seconds)") +
 	labs(colour = "Execution mode", shape = "Execution mode") 

### 2014-09-08 measurements/faceted_graph.R
import gefixt

@@ -1,6 +1,7 @@
 setwd("C:/Git-data/Afstuderen/measurements")
 library(ggplot2)
 library(mgcv)
+library(stringr)
 library(scales)
 measure = read.csv("raw_data.csv") # read csv file
 name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
@@ -19,10 +20,10 @@ getName <- function(s){
 		)
 	)
 }
-measure$seconds <- measure$trainTime/60
+
 name <- unlist(lapply(name, getName))
+measure$seconds <- measure$trainTime/1000
 measure$serial.parallel <- name
-maxTime <- max(measure$trainTime)
 d <- ggplot(data=measure, aes(x=dataSize, y=seconds, colour=serial.parallel, shape=serial.parallel))
 d <- d + geom_point(size=2) + stat_smooth(level=0.99, method = "glm", formula = y ~ exp(x)) + 
 	scale_x_log10() + theme_bw() + 
@@ -32,7 +33,7 @@ d <- d + geom_point(size=2) + stat_smooth(level=0.99, method = "glm", formula = 
 		panel.border     = element_blank(),
 		panel.background = element_blank(),
 		axis.line = element_line(color='black')
-	) + scale_y_continuous(expand = c(0.001, maxTime/10000))+	
+	) + scale_y_continuous(expand = c(0, 0))+	
 	facet_wrap(~serial.parallel)+
 	xlab("Dataset size (in Byte)") +
 	ylab("Training iteration time (in Seconds)") +

### 2014-09-08 measurements/single_graph.R

@@ -1,6 +1,7 @@
 setwd("C:/Git-data/Afstuderen/measurements")
 library(ggplot2)
 library(mgcv)
+library(stringr)
 library(scales)
 measure = read.csv("raw_data.csv") # read csv file
 name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
@@ -19,10 +20,10 @@ getName <- function(s){
 		)
 	)
 }
-measure$seconds <- measure$trainTime/60
+
 name <- unlist(lapply(name, getName))
+measure$seconds <- measure$trainTime/1000
 measure$serial.parallel <- name
-maxTime <- max(measure$trainTime)
 d <- ggplot(data=measure, aes(x=dataSize, y=seconds, colour=serial.parallel, shape=serial.parallel))
 d <- d + geom_point(size=2) + stat_smooth(se=FALSE, method = "glm", formula = y ~ exp(x)) + 
 	scale_x_log10() + theme_bw() + 
@@ -32,7 +33,7 @@ d <- d + geom_point(size=2) + stat_smooth(se=FALSE, method = "glm", formula = y 
 		panel.border     = element_blank(),
 		panel.background = element_blank(),
 		axis.line = element_line(color='black')
-	) + scale_y_continuous(expand = c(0.001, maxTime/10000))+	
+	) + scale_y_continuous(expand = c(0, 0))+	
 	xlab("Dataset size (in Byte)") +
 	ylab("Training iteration time (in Seconds)") +
 	labs(colour = "Execution mode", shape = "Execution mode") 

### 2014-09-08 measurements/faceted_graph.R
fixed order

@@ -11,7 +11,7 @@ getName <- function(s){
 			paste(
 				c(
 					"Cluster: ", 
-					str_extract(s, "\\d+"), 
+					sprintf("% 3d",as.numeric(str_extract(s, "\\d+"))), 
 					" data nodes"
 				),
 				collapse = ''

### 2014-09-09 report/Chapters/Results.tex
mooie plaatjes + in report

@@
 This chapter describes the results of the running time measurements of the cluster and baseline implementations as described in \ref{chap:implementation}.
+\section{ListNet}
+Figure
+\begin{figure}
+\centering
+\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.7]{gfx/time_single.pdf}
+\caption{Processing time of a single ListNet training iteration}
+\label{fig:listnet_train_time}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.7]{gfx/time_single_logx.pdf}
+\caption{Processing time of a single ListNet training iteration on a logarithmic scale}
+\label{fig:listnet_train_time_log}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.7]{gfx/throughput_single_logx.pdf}
+\caption{Throughput of a ListNet training iteration}
+\label{fig:listnet_throughput}
+\end{figure}

### 2014-09-09 measurements/throughput_faceted.R

@@ -25,16 +25,16 @@ name <- unlist(lapply(name, getName))
 measure$seconds <- measure$trainTime/1000
 measure$throughput <- measure$dataSize/measure$seconds
 measure$serial.parallel <- name
-d <- ggplot(data=measure, aes(x=dataSize, y=seconds, colour=serial.parallel, shape=serial.parallel))
-d <- d + geom_point(size=2) + stat_smooth(level=0.99, method = "glm", formula = y ~ exp(x)) + 
-	scale_x_log10() + theme_bw() + 
+d <- ggplot(data=measure, aes(x=dataSize, y=throughput, colour=serial.parallel, shape=serial.parallel))
+d <- d + geom_point(size=2) + geom_line() + 
+	theme_bw() + 
 	theme(
 		panel.grid.major = element_blank(),
 		panel.grid.minor = element_blank(),
 		panel.border     = element_blank(),
 		panel.background = element_blank(),
 		axis.line = element_line(color='black')
-	) + scale_y_continuous(expand = c(0, 0))+	
+	) + scale_y_continuous(expand = c(0, 10000000))+	
 	facet_wrap(~serial.parallel)+
 	xlab("Dataset size (in Byte)") +
 	ylab("Training iteration time (in Seconds)") +

### 2014-09-09 measurements/throughput_single.R

@@ -1,5 +1,6 @@
 setwd("C:/Git-data/Afstuderen/measurements")
 library(ggplot2)
+library(boot)
 library(mgcv)
 library(stringr)
 library(scales)
@@ -25,18 +26,17 @@ name <- unlist(lapply(name, getName))
 measure$seconds <- measure$trainTime/1000
 measure$throughput <- measure$dataSize/measure$seconds
 measure$serial.parallel <- name
-d <- ggplot(data=measure, aes(x=dataSize, y=seconds, colour=serial.parallel, shape=serial.parallel))
-d <- d + geom_point(size=2) + stat_smooth(level=0.99, method = "glm", formula = y ~ exp(x)) + 
-	scale_x_log10() + theme_bw() + 
+d <- ggplot(data=measure, aes(x=dataSize, y=throughput, colour=serial.parallel, shape=serial.parallel))
+d <- d + geom_point(size=2) + geom_line() + 
+	theme_bw() + 
 	theme(
 		panel.grid.major = element_blank(),
 		panel.grid.minor = element_blank(),
 		panel.border     = element_blank(),
 		panel.background = element_blank(),
 		axis.line = element_line(color='black')
-	) + scale_y_continuous(expand = c(0, 0))+	
-	facet_wrap(~serial.parallel)+
-	xlab("Dataset size (in Byte)") +
-	ylab("Training iteration time (in Seconds)") +
+	) + scale_x_continuous(expand = c(0.01, 0)) + scale_y_continuous(expand = c(0, 0)) +
+	xlab("Dataset Size (in Byte)") +
+	ylab("Throughput (in Byte/Second)") +
 	labs(colour = "Execution mode", shape = "Execution mode") 
 ggplot_build(d)

### 2014-09-09 measurements/throughput_single_logx.R

@@ -1,5 +1,6 @@
 setwd("C:/Git-data/Afstuderen/measurements")
 library(ggplot2)
+library(boot)
 library(mgcv)
 library(stringr)
 library(scales)
@@ -25,8 +26,8 @@ name <- unlist(lapply(name, getName))
 measure$seconds <- measure$trainTime/1000
 measure$throughput <- measure$dataSize/measure$seconds
 measure$serial.parallel <- name
-d <- ggplot(data=measure, aes(x=dataSize, y=seconds, colour=serial.parallel, shape=serial.parallel))
-d <- d + geom_point(size=2) + stat_smooth(level=0.99, method = "glm", formula = y ~ exp(x)) + 
+d <- ggplot(data=measure, aes(x=dataSize, y=throughput, colour=serial.parallel, shape=serial.parallel))
+d <- d + geom_point(size=2) + geom_line() + 
 	scale_x_log10() + theme_bw() + 
 	theme(
 		panel.grid.major = element_blank(),
@@ -34,9 +35,8 @@ d <- d + geom_point(size=2) + stat_smooth(level=0.99, method = "glm", formula = 
 		panel.border     = element_blank(),
 		panel.background = element_blank(),
 		axis.line = element_line(color='black')
-	) + scale_y_continuous(expand = c(0, 0))+	
-	facet_wrap(~serial.parallel)+
-	xlab("Dataset size (in Byte)") +
-	ylab("Training iteration time (in Seconds)") +
+	) + scale_y_continuous(expand = c(0, 0)) +
+	xlab("Dataset Size (in Byte)") +
+	ylab("Throughput (in Byte/Second)") +
 	labs(colour = "Execution mode", shape = "Execution mode") 
 ggplot_build(d)

### 2014-09-09 report/Chapters/Introduction.tex
voortgang thesis + nieuwe metingen
uitbreiding methodology

@@
 $S_N = \frac{\text{execution time using one core}}{\text{execution time using \emph{N} cores}}$
-
-\item[RQ3] Can those Learning-to-Rank algorithms be adjusted in such a way that the parallel execution speed-up increases without decreasing accuracy?
 \end{description}
@@
 \section{Approach}
+\subsection{Methodology for Research Question I}
 A literature study will be performed to get insight in relevant existing techniques for large scale Learning-to-Rank. The literature study will be performed by using the following query:\\
@@
 
-To answer the first research question, I will implement Learning-to-Rank methods in the MapReduce framework and measure runtime as a factor of the number of cluster nodes used to complete the computation.\\
+\subsection{Methodology for Research Question II}
+To find an answer to the second research question, the Learning-to-Rank methods determined in the first research question will be implemented in the MapReduce framework and training time will be measured as a factor of the number of cluster nodes used to perform the computation.
+The HDInsight cloud-based MapReduce platform from Microsoft will be used to run the Learning-to-Rank algorithms on.
+HDInsight is based on the popular open source MapReduce implementation Hadoop\footnote{http://hadoop.apache.org/}.\\
 
-I will use the HDInsight cloud-based MapReduce implementation from Microsoft to implement the Learning-to-Rank algorithms.
-HDInsight is based on the popular open source MapReduce implementation Hadoop\footnote{http://hadoop.apache.org/}.
-The algorithms that we include in the measurements will be determined based on experimental results on the \emph{Yahoo!
-Learning to Rank Challenge} \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, the LETOR \cite{Qin2010} dataset and the LETOR successors MSLR-WEB10k and MSLR-WEB30k.
+To research the speed-up's dependence on the amount of data being processed, the training computations will be performed on datasets of varying sizes.
+The well-known benchmark collections LETOR 3.0, LETOR 4.0 and MSLR-WEB30/40K offer a good initial set of datasets.
+Table \ref{tbl:initial_datasets} shows the
+% TODO: Waarom .Gov uit LETOR 3.0 niet?%
+\begin{table}[!h]
+\centering
+\begin{tabular}{p{3.4cm}p{3.4cm}r}\toprule
+Dataset & Collection & Size \\
+\midrule
+OHSUMED     & LETOR 3.0       &   4.55 MB\\
+MQ2008      & LETOR 4.0       &   5.93 MB\\
+MQ2007      & LETOR 4.0       &  25.52 MB\\
+MSLR-WEB10K & MSLR-WEB10K     & 938.01 MB\\
+MSLR-WEB30K & MSLR-WEB30K     &   2.62 GB\\
+\bottomrule
+\end{tabular}
+\caption{LETOR 3.0, LETOR 4.0 and MSLR30/40K datasets and sizes}
+\label{tbl:initial_datasets}
+\end{table}
 
+To test the how the computational performance of Learning to Rank algorithms both on cluster and on single-node computation scales to large quantities of data, larger datasets will be created by cloning the MSLR-WEB30K dataset such that the cloned queries will hold new distinct query ID's.
 \section{Thesis Overview}

### 2014-09-09 report/Chapters/Results.tex
uitbreiding methodology

@@
-This chapter describes the results of the running time measurements of the cluster and baseline implementations as described in \ref{chap:implementation}.
+This chapter describes and discusses the results of the running time measurements of the cluster and baseline implementations as described in \ref{chap:implementation}.
 \section{ListNet}
-Figure
+Figure \ref{fig:listnet_train_time} and Figure \ref{fig:listnet_train_time_log} (with a logarithmic data size axis) show the processing times of a single training iteration using the ListNet implementation included in RankLib library and the cluster implementation described in chapter \ref{chap:implementation} with different numbers of data nodes.
 \begin{figure}
@@
 \includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.7]{gfx/time_single_logx.pdf}
-\caption{Processing time of a single ListNet training iteration on a logarithmic scale}
+\caption{Processing time of a single ListNet training iteration on a logarithmic axis}
 \label{fig:listnet_train_time_log}
 \end{figure}
-
+Figures \ref{fig:listnet_train_time} and \ref{fig:listnet_train_time_log} also show that serial execution using the RankLib implementation of ListNet is several orders of magnitude faster than running ListNet on a cluster.
 \begin{figure}

### 2014-09-09 report/Chapters/Results.tex
begin resultatensectie

@@
 Figure \ref{fig:listnet_train_time} and Figure \ref{fig:listnet_train_time_log} (with a logarithmic data size axis) show the processing times of a single training iteration using the ListNet implementation included in RankLib library and the cluster implementation described in chapter \ref{chap:implementation} with different numbers of data nodes.
+The horizontal positions of the measurements are largely identical between execution method, as they are determined by the collection of datasets used.
+As previously described in the methodology section, the LETOR 3.0, LETOR 4.0 and MSLR-WEB10/30K datasets are used, supplemented with generated datasets that are multiplications of MSLR-WEB30K.
+Table \ref{tbl:recap_datasets} describes the datasets used in the experiments.
+
+\begin{table}[!h]
+\centering
+\begin{tabular}{p{3.4cm}p{3.4cm}r}\toprule
+Dataset & Collection & Size \\
+\midrule
+MINI		& GENERATED		  & 143.38 KB\\
+OHSUMED     & LETOR 3.0       &   4.55 MB\\
+MQ2008      & LETOR 4.0       &   5.93 MB\\
+MQ2007      & LETOR 4.0       &  25.52 MB\\
+MSLR-WEB10K & MSLR-WEB10K     & 938.01 MB\\
+MSLR-WEB30K & MSLR-WEB30K     &   2.62 GB\\
+CUSTOM-2	& GENERATED		  &   5.25 GB\\
+CUSTOM-5	& GENERATED		  &  13.12 GB\\
+CUSTOM-10	& GENERATED		  &  26.24 GB\\
+CUSTOM-20   & GENERATED       &  52.42 GB\\
+CUSTOM-50	& GENERATED		  & 131.21 GB\\
+CUSTOM-100	& GENERATED		  & 262.41 GB\\
+\bottomrule
+\end{tabular}
+\caption{Description of datasets used for running time experiments}
+\label{tbl:recap_datasets}
+\end{table}
+
 \begin{figure}

### 2014-09-10 report/Chapters/Results.tex
speedup graph

@@
 Figure \ref{fig:listnet_train_time} and Figure \ref{fig:listnet_train_time_log} (with a logarithmic data size axis) show the processing times of a single training iteration using the ListNet implementation included in RankLib library and the cluster implementation described in chapter \ref{chap:implementation} with different numbers of data nodes.
-The horizontal positions of the measurements are largely identical between execution method, as they are determined by the collection of datasets used.
+The horizontal positions of the measurements are identical between execution method, as they are determined by the collection of datasets used.
 As previously described in the methodology section, the LETOR 3.0, LETOR 4.0 and MSLR-WEB10/30K datasets are used, supplemented with generated datasets that are multiplications of MSLR-WEB30K.
@@
 
+Figures \ref{fig:listnet_train_time} and \ref{fig:listnet_train_time_log}
+
 \begin{figure}
@@
 \end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.7]{gfx/speedup_faceted.pdf}
+\caption{Processing time as a function of data nodes in a cluster}
+\label{fig:speedup_train_time}
+\end{figure}
+
 Figures \ref{fig:listnet_train_time} and \ref{fig:listnet_train_time_log} also show that serial execution using the RankLib implementation of ListNet is several orders of magnitude faster than running ListNet on a cluster.

### 2014-09-10 measurements/single_graph_logx.R

@@ -26,7 +26,7 @@ measure$seconds <- measure$trainTime/1000
 measure$serial.parallel <- name
 d <- ggplot(data=measure, aes(x=dataSize, y=seconds, colour=serial.parallel, shape=serial.parallel))
 d <- d + geom_point(size=2) + geom_line() + 
-	theme_bw() + 
+	scale_x_log10() + theme_bw() + 
 	theme(
 		panel.grid.major = element_blank(),
 		panel.grid.minor = element_blank(),

### 2014-09-10 measurements/speedup_graphs.R

@@ -1,0 +1,20 @@
+setwd("C:/Git-data/Afstuderen/measurements")
+library(ggplot2)
+library(mgcv)
+library(stringr)
+library(scales)
+measure = read.csv("raw_data.csv") # read csv file
+measure$seconds <- measure$trainTime/1000
+measure$dataset <- factor(measure$dataset, levels = c("MINI", "OHSUMED", "MQ2008", "MQ2007", "MSLR-WEB10K", "MSLR-WEB30K", "CUSTOM-2", "CUSTOM-5", "CUSTOM-10", "CUSTOM-20", "CUSTOM-50", "CUSTOM-100"))
+d <- ggplot(data=measure, aes(x=nodeCount, y=seconds))
+d <- d + geom_point(size=2) + geom_line() + theme_bw() + 
+	theme(
+		panel.grid.major = element_blank(),
+		panel.grid.minor = element_blank(),
+		panel.border     = element_blank(),
+		panel.background = element_blank(),
+		axis.line = element_line(color='black')
+	) + facet_wrap(~ dataset, scales="free")+
+	xlab("Nodes in cluster") +
+	ylab("Training iteration time (in seconds)")
+ggplot_build(d)

### 2014-09-10 report/Chapters/Results.tex
metingen-todo
results

@@
 
+Figure \ref{fig:speedup_train_time} shows that speed-up is sub-linear such that the training time converges to a constant unit of time somewhere within the range of 150-200 seconds.
+This time is likely to be caused by Hadoop job scheduling overhead, this presumption is strengthened by long waiting periods between the separate MapReduce jobs that form a training iteration.
+
 \begin{figure}

### 2014-09-10 report/Chapters/Results.tex
Amdahl's law erin verwerkt :)

--- 2014-09-10 report/Chapters/Results.tex
@@
 
+Figure \ref{fig:speedup_train_time} shows that speed-up is sub-linear such that the training time converges to a constant unit of time somewhere within the range of 150-200 seconds.
+This time is likely to be caused by Hadoop job scheduling overhead, this presumption is strengthened by long waiting periods between the separate MapReduce jobs that form a training iteration.\\
+
+Amdahl's law states that the speed-up of a program using parallel computing is limited by the time needed for the sequential fraction.
+A consequence of Amdahl's Law is that all parallel programs the have a non-parallelisable part have sub-linear spreed-up.
+%Leg uit dat dit overeenkomt met onze bevindingen: sub-linear speed-up en we hebben een constante factor.
+Leg ook uit dat deze constante tijd niet (nauwelijks) groeit bij toename van formaat van de dataset, waardoor verhoudingsgewijs het niet-paralleliseerbare deel van de berekening afneemt wanneer de grootte van de dataset toeneemt.
+Uitleggen aan de hand van Amdahl's law dat dit als gevolg heeft dat de maximaal haalbare speedup hierdoor hoger is bij grote datasets dan bij kleine datasets.
+
 \begin{figure}

### 2014-09-11 report/Chapters/Results.tex
amdahl's law + throughput

--- 2014-09-10 report/Chapters/Results.tex
@@
 
+Figure \ref{fig:speedup_train_time} shows that speed-up is sub-linear such that the training time converges to a constant unit of time somewhere within the range of 150-200 seconds.
+This time is likely to be caused by Hadoop job scheduling overhead, this presumption is strengthened by long waiting periods between the separate MapReduce jobs that form a training iteration.\\
+
+Amdahl's law states that the speed-up of a program using parallel computing is limited by the time needed for the sequential fraction of the program.
+A consequence of Amdahl's Law is that all parallel programs the have a non-parallelisable part have sub-linear speed-up.
+Behaviour in accordance with Amdahl's law can be seen in Figure \ref{fig:speedup_train_time}, where the speed-up is sub-linear as a result of the existence a non-parallelisable fraction of the program.\\
+
+Note however that Hadoop job scheduling overhead is independent of dataset size.
+Therefore, the non-parallelisable fraction of the program will be smaller when the to be processed dataset is larger, allowing larger speed-up values for larger datasets.
+From this observation we can derive that for \emph{"large enough"} datasets, the speed-up obtained by parallelising ListNet using Hadoop MapReduce is large enough for the parallelisation to be beneficial in terms of processing time compared to the RankLib single-machine version of ListNet, even when the size of the to be processed dataset would not have been memory-bounded in RankLib ListNet.
+Figure \ref{fig:listnet_throughput} supports this claim as the processing throughput increases both when a larger cluster is used and when more data is processed.\\
+
 \begin{figure}

### 2014-09-12 report/Chapters/Results.tex
nieuwer serial meting + hernoeming scripts

@@
 
-Figures \ref{fig:listnet_train_time} and \ref{fig:listnet_train_time_log}
+Figures \ref{fig:listnet_train_time} and \ref{fig:listnet_train_time_log} show that the single-machine RankLib implementation of ListNet is very fast compared to the Hadoop MapReduce cluster implementation of ListNet for datasets that fit in memory.
+RankLib performs much worse in cases where the dataset does not fit into memory, which is likely to be a result of swapping data to and from virtual memory.
 
-\begin{figure}
+\begin{figure}[!h]
 \centering
@@
 
-\begin{figure}
+\begin{figure}[!h]
 \centering
@@
 Figures \ref{fig:listnet_train_time} and \ref{fig:listnet_train_time_log} also show that serial execution using the RankLib implementation of ListNet is several orders of magnitude faster than running ListNet on a cluster.
-\begin{figure}
+\begin{figure}[!h]
 \centering

### 2014-09-12 implementation/java/src/letor/data/DataFileChecker.java

@@ -16,7 +16,7 @@ import java.util.TreeMap;
  */
 public class DataFileChecker {
     // Initialize parameter
-    private static final DataSets.DataSet DATASET = DataSets.DataSet.CUSTOM_10;
+    private static final DataSets.DataSet DATASET = DataSets.DataSet.CUSTOM;
 
     // Global variables
     private static String dataFolderPath = "C:/Git-data/Afstuderen/implementation/java/input/";

### 2014-09-12 implementation/java/src/letor/data/LargeDatasetCreator.java

@@ -16,8 +16,8 @@ import java.util.List;
  */
 public class LargeDatasetCreator {
     // Class parameters
-    private static int lowerDuplicationNumber = 41;
-    private static int upperDuplicationNumber = 60;
+    private static int lowerDuplicationNumber = 0;
+    private static int upperDuplicationNumber = 20;
 
     // Global variables
     private static String dataFolderPath = "C:/Git-data/Afstuderen/implementation/java/input";

### 2014-09-12 implementation/java/src/letor/parallel/util/DataSets.java

@@ -9,7 +9,7 @@ import scala.tools.cmd.Meta;
 public class DataSets {
     public static enum DataSet {
         OHSUMED, MQ2007, MQ2008, MSLR_WEB10K, MSLR_WEB30K,
-        CUSTOM, CUSTOM_2, CUSTOM_5, CUSTOM_10, MINI
+        CUSTOM, MINI
     }
 
     public static Metadata getMetaData(DataSet ds, int duplicationNumber){

### 2014-09-12 implementation/java/src/letor/serial/FoldRunHandler.java

@@ -21,8 +21,9 @@ import java.util.List;
 public class FoldRunHandler {
     private MetricScorerFactory mFact = new MetricScorerFactory();
 
-    private String path  = "C:\\Git-data\\Afstuderen\\implementation\\java\\input\\";
-    private int folds    = 1;
+    private String path             = "C:\\Git-data\\Afstuderen\\implementation\\java\\input\\";
+    private int folds               = 1;
+    private int duplicationNumber   = 0;
 
     private Ranker ranker;
 
@@ -37,13 +38,24 @@ public class FoldRunHandler {
     private Long trainTime;
     private Long endTime;
 
-    public FoldRunHandler(AbstractParameterizedRanker ranker, DataSets.DataSet dataSet, int folds, int iterations){
+    public FoldRunHandler(AbstractParameterizedRanker ranker, DataSets.DataSet dataSet, int duplicationNumber, int folds, int iterations){
         this.ranker      = ranker.getParameterizedRanker();
+        this.folds       = folds;
+
         if(ranker instanceof ListNetHandler){
             ((ListNet)this.ranker).nIteration = iterations;
         }
-        this.path        = path + DataSets.getMetaData(dataSet).getName() +"\\";
-        this.folds       = folds;
+
+        if(dataSet.equals(DataSets.DataSet.CUSTOM)) {
+            this.duplicationNumber = duplicationNumber;
+            this.path        = path + DataSets.getMetaData(DataSets.DataSet.CUSTOM, duplicationNumber).getName() +"\\";
+        }else{
+            this.path        = path + DataSets.getMetaData(dataSet).getName() +"\\";
+        }
+    }
+
+    public FoldRunHandler(AbstractParameterizedRanker ranker, DataSets.DataSet dataSet, int folds, int iterations){
+        this(ranker, dataSet, 0, folds, iterations);
     }
 
     public Measurement averageScore() {
@@ -65,7 +77,7 @@ public class FoldRunHandler {
     }
 
     public double evaluate(String trainFile, String validationFile, String testFile){
-        List<RankList> train      = readInput(trainFile);
+        List<RankList> train      = readTrainInput(trainFile, duplicationNumber);
         List<RankList> validation = readInput(validationFile);
         List<RankList> test       = readInput(testFile);
 
@@ -88,8 +100,22 @@ public class FoldRunHandler {
         return samples;
     }
 
-    public int[] getFeatureFromSampleVector(List<RankList> samples)
-    {
[…]

### 2014-09-12 implementation/java/src/letor/serial/WrappedRunner.java

@@ -9,8 +9,16 @@ import letor.serial.util.Measurement;
  * Created by niek.tax on 7/15/2014.
  */
 public class WrappedRunner {
+
     public static void main(String[] args){
-        FoldRunHandler listNetWrapper = new FoldRunHandler(new ListNetHandler(), DataSets.DataSet.MSLR_WEB30K, 1, 1); // Method, dataset, folds, iterations
+        DataSets.DataSet    dataset             = DataSets.DataSet.CUSTOM;
+        Integer             duplicationNumber   = 7; // Only relevant in case of dataset Custom
+
+        FoldRunHandler listNetWrapper = null;
+        if(dataset == DataSets.DataSet.CUSTOM)
+            listNetWrapper = new FoldRunHandler(new ListNetHandler(), DataSets.DataSet.CUSTOM, duplicationNumber,  1, 1); // Method, dataset, folds, iterations
+        else
+            listNetWrapper = new FoldRunHandler(new ListNetHandler(), dataset, 1, 1); // Method, dataset, folds, iterations
         Measurement measurement = listNetWrapper.averageScore();
         System.out.println("Preprocessing: "+measurement.getPreprocessingTime());
         System.out.println("Train:         "+measurement.getTrainingTime());

### 2014-09-12 measurements/speedup_faceted.R

@@ -6,6 +6,7 @@ library(scales)
 measure = read.csv("raw_data.csv") # read csv file
 measure$seconds <- measure$trainTime/1000
 measure$dataset <- factor(measure$dataset, levels = c("MINI", "OHSUMED", "MQ2008", "MQ2007", "MSLR-WEB10K", "MSLR-WEB30K", "CUSTOM-2", "CUSTOM-5", "CUSTOM-10", "CUSTOM-20", "CUSTOM-50", "CUSTOM-100"))
+measure <- measure[!is.na(measure$dataset),]
 d <- ggplot(data=measure, aes(x=nodeCount, y=seconds))
 d <- d + geom_point(size=2) + geom_line() + theme_bw() + 
 	theme(

### 2014-09-12 report/Chapters/Results.tex
bespreking throughput graph

@@
 \section{ListNet}
-Figure \ref{fig:listnet_train_time} and Figure \ref{fig:listnet_train_time_log} (with a logarithmic data size axis) show the processing times of a single training iteration using the ListNet implementation included in RankLib library and the cluster implementation described in chapter \ref{chap:implementation} with different numbers of data nodes.
+Figure \ref{fig:listnet_train_time} and Figure \ref{fig:listnet_train_time_log} (which presents the same data with a logarithmic data size axis) show the processing times of a single training iteration using the ListNet implementation included in RankLib library and the cluster implementation described in chapter \ref{chap:implementation} with different numbers of data nodes.
 The horizontal positions of the measurements are identical between execution method, as they are determined by the collection of datasets used.
@@
 Figures \ref{fig:listnet_train_time} and \ref{fig:listnet_train_time_log} show that the single-machine RankLib implementation of ListNet is very fast compared to the Hadoop MapReduce cluster implementation of ListNet for datasets that fit in memory.
-RankLib performs much worse in cases where the dataset does not fit into memory, which is likely to be a result of swapping data to and from virtual memory.
+As soon as the amount of data processed with RankLib ListNet approaches the physical memory limit of the machine used for the computation (16 GB for the machine used in our experiments), RankLib ListNet processing time start to increase very rapidly.
+This increase is likely to be a result of swapping data to and from virtual memory.
+The largest dataset processed with RankLib ListNet is CUSTOM-8, which is 20.99 GB in size and therewith larger than the 16 GB physical memory limit.
+Experiments with RankLib ListNet on CUSTOM-10 were attempted, but were stopped after not completing within 12 hours.\\
 
@@
 \centering
-\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.7]{gfx/time_single.pdf}
+\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.8]{gfx/time_single.pdf}
 \caption{Processing time of a single ListNet training iteration}
@@
 \centering
-\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.7]{gfx/time_single_logx.pdf}
+\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.8]{gfx/time_single_logx.pdf}
 \caption{Processing time of a single ListNet training iteration on a logarithmic axis}
@@
 Therefore, the non-parallelisable fraction of the program will be smaller when the to be processed dataset is larger, allowing larger speed-up values for larger datasets.
-From this observation we can derive that for \emph{"large enough"} datasets, the speed-up obtained by parallelising ListNet using Hadoop MapReduce is large enough for the parallelisation to be beneficial in terms of processing time compared to the RankLib single-machine version of ListNet, even when the size of the to be processed dataset would not have been memory-bounded in RankLib ListNet.
-Figure \ref{fig:listnet_throughput} supports this claim as the processing throughput increases both when a larger cluster is used and when more data is processed.\\
+From this observation we can derive that for \emph{"large enough"} datasets, the speed-up obtained by parallelising ListNet using Hadoop MapReduce is large enough for the parallelisation to be beneficial in terms of processing time compared to the RankLib single-machine version of ListNet, even when the size of the to be processed dataset would not have been memory-bounded in RankLib ListNet.\\
+
+Figure \ref{fig:listnet_throughput} shows the throughput in bytes per second.
+Our observation that RankLib ListNet performs very slow for data sets that do not fit in physical memory and virtual memory is needed is very notable in this graph.
+Furthermore, this graph shows how both increase in number in data nodes in a cluster and increase in input data size result in an increase in throughput.
+The sub-linear growth of throughput as a factor of input data size can again be explained as a result of the non-parallelisable Hadoop job scheduling part of the operation.\\
 
@@
 
-Figures \ref{fig:listnet_train_time} and \ref{fig:listnet_train_time_log} also show that serial execution using the RankLib implementation of ListNet is several orders of magnitude faster than running ListNet on a cluster.
 \begin{figure}[!h]
 \centering
-\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.7]{gfx/throughput_single_logx.pdf}
+\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.8]{gfx/throughput_single_logx.pdf}
 \caption{Throughput of a ListNet training iteration}

### 2014-09-15 paper/IPM/paper.tex
IP&M draft

--- 2014-09-02 paper/CIKM/paper.tex
@@
-% This is "sig-alternate.tex" V2.0 May 2012
-% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
-%
-% This example file demonstrates the use of the 'sig-alternate.cls'
-% V2.5 LaTeX2e document class file. It is for those submitting
-% articles to ACM Conference Proceedings WHO DO NOT WISH TO
-% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
-% The 'sig-alternate.cls' file will produce a similar-looking,
-% albeit, 'tighter' paper resulting in, invariably, fewer pages.
-%
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V2.5) produces:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) NO page numbers
-%
-% as against the acm_proc_article-sp.cls file which
-% DOES NOT produce 1) thru' 3) above.
-%
-% Using 'sig-alternate.cls' you have control, however, from within
-% the source .tex file, over both the CopyrightYear
-% (defaulted to 200X) and the ACM Copyright Data
-% (defaulted to X-XXXXX-XX-X/XX/XX).
-% e.g.
-% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
-% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
-%
-% ---------------------------------------------------------------------------------------------------------------
-% This .tex source is an example which *does* use
-% the .bib file (from which the .bbl file % is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission, you *NEED* to 'insert'
-% your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% ================= IF YOU HAVE QUESTIONS =======================
-% Questions regarding the SIGS styles, SIGS policies and
-% procedures, Conferences etc. should be sent to
-% Adrienne Griscti (griscti@acm.org)
-%
-% Technical questions _only_ to
-% Gerald Murray (murray@hq.acm.org)
-% ===============================================================
-%
-% For tracking purposes - this is V2.0 - May 2012
-
-\documentclass{sig-alternate}
+\documentclass[preprint]{elsarticle}
 \usepackage{booktabs}
@@
 \usepackage{enumitem}
+\usepackage{amsmath}
 \usepackage{float}
@@
 \usepackage{balance}
+
+\newcommand\bibsection{%
+  \section*{\bibname\markright{\MakeUppercase{\bibname}}}}
+
 \begin{document}
+\begin{frontmatter}
+\title{A Cross-Benchmark Comparison of 87 Learning to Rank Methods}
 
-%
-% --- Author Metadata here ---
-\conferenceinfo{ACM International Conference on Information and Knowledge Management}{'14 Shanghai, China}
-\CopyrightYear{2014} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-% --- End of Author Metadata ---
+\author[avanade,ut]{Niek Tax\corref{cor}\fnref{fn}}
+\ead{niek.tax@avanade.com}
 
-\title{A Cross-Benchmark Comparison of 87 Learning to Rank Methods}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
+\author[avanade]{Sander Bockting}
+\ead{sander.bockting@avanade.com}
 
-\numberofauthors{3} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
-Niek Tax\titlenote{author is affiliated with the University of Twente, but this paper written during his time at Avanade Netherlands B.V.}\\
-       \affaddr{Avanade Netherlands B.V.}\\
-       \affaddr{Versterkerstraat 6}\\
-       \affaddr{1322 AP Almere}\\
-       \affaddr{The Netherlands}
-       \affaddr{niek.tax@avanade.com}
-% 2nd. author
-\alignauthor
-Sander Bockting\\
-       \affaddr{Avanade Netherlands B.V.}\\
-       \affaddr{Versterkerstraat 6}\\
-       \affaddr{1322 AP Almere}\\
-       \affaddr{The Netherlands}
-       \affaddr{sander.bockting@avanade.com}
-% 3rd. author
-\alignauthor
-Djoerd Hiemstra\\
-       \affaddr{University of Twente}\\
-       \affaddr{P.O. Box 217}\\
-       \affaddr{7500 AE Enschede}\\
-       \affaddr{The Netherlands}
-       \affaddr{d.hiemstra@utwente.nl}
-}
+\author[ut]{Djoerd Hiemstra}
+\ead{d.hiemstra@utwente.nl}
 
-\maketitle
+\cortext[cor]{Corresponding author}
+\fntext[fn]{Author is affiliated with the University of Twente, but this paper written during his time at Avanade Netherlands B.V.}
+
+\address[avanade]{Avanade Netherlands B.V., Versterkerstraat 6,
+1322AP Almere, the Netherlands}
+\address[ut]{University of Twente, P.O. Box 217, 7500AE Enschede,
+the Netherlands}
 \begin{abstract}
@@
 \end{abstract}
-
-% A category with the (minimum) three required fields
-\category{H.3.3}{Information Systems}{Information Storage and Retrieval}[Information Search and Retrieval]
-
-\keywords{Learning to rank, Information retrieval, Evaluation metric}
+\begin{keyword}
+Learning to rank \sep Information retrieval \sep Evaluation metric
+\end{keyword}
+\end{frontmatter}
 
@@
 Ranking has become a vital part of web search, where commercial search engines help users find their need in the extremely large collection of the World Wide Web.
-Among useful applications outside web search are automatic text summarization, machine translation, drug discovery and determining the ideal order of maintenance operations \cite{Rudin2009}.
+Among useful applications outside web search are automatic text summarisation, machine translation, drug discovery and determining the ideal order of maintenance operations \cite{Rudin2009}.
 In addition, the ranking task has been argued to be a better fit for recommender systems than the regression task \cite{McNee2006} (which concerns predictions on a continuous scale) that is currently frequently used in such systems.\\
@@
 \begin{table}[!h]
-\scriptsize{
-\begin{tabular}{p{6.4cm}R{1.26cm}}\toprule
+\centering
+\begin{tabular}{p{8cm}R{3cm}}\toprule
 Benchmark collection & \# of datasets \\
@@
 \caption{Included learning to rank evaluation benchmark collections}
-\label{tbl:ltr_benchmark_collections}}
+\label{tbl:ltr_benchmark_collections}
 \end{table}
@@
 \begin{table}[!h]
-\scriptsize{
-\begin{tabular}{p{2.8cm}p{2.64cm}R{1.9cm}}\toprule
+\centering
+\begin{tabular}{p{6cm}p{3cm}R{2cm}}\toprule
 Benchmark & Paper & \# of forward references \\
@@
 \caption{Forward references of learning to rank benchmark papers}
-\label{tbl:ltr_benchmark_forref}}
+\label{tbl:ltr_benchmark_forref}
 \end{table}
@@
 \begin{table}[!h]
-\scriptsize{
-\begin{tabular}{p{5.2cm}R{2.5cm}}\toprule
+\centering
+\begin{tabular}{p{8cm}R{3cm}}\toprule
 Search Query & Google scholar search results \\
@@
 \caption{Google scholar search results for learning to rank benchmarks}
-\label{tbl:ltr_benchmark_searches}}
+\label{tbl:ltr_benchmark_searches}
 \end{table}
@@
 \subsection{Literature Selection}
-Table 4 in the appendix gives an overview of the learning to rank methods for which evaluation results were found through the described procedure.
-Occurrences of L2, L3 and L4 in Table 4 imply that these algorithms are evaluated as official LETOR 2.0, 3.0 and 4.0 baselines respectively.\\
+Table \ref{tab:ltr_methods_used} in the appendix gives an overview of the learning to rank methods for which evaluation results were found through the described procedure.
+Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, 3.0 and 4.0 baselines respectively.\\
 
@@
 \end{figure}
-Figure \ref{fig:normalized_winning_number_all} shows the NWN as function of IWN for the methods described in Table 4.
-The cross-metric comparison is based on the NDCG@$\{3,5,10\}$ and MAP comparisons combined, which justifies analyzing the comparison more thoroughly.
+Figure \ref{fig:normalized_winning_number_all} shows the NWN as function of IWN for the methods described in Table \ref{tab:ltr_methods_used}.
+The cross-metric comparison is based on the NDCG@$\{3,5,10\}$ and MAP comparisons combined, which justifies analysing the comparison more thoroughly.
 Figure \ref{fig:normalized_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
@@
 Our work contributes to this by identifying promising learning to rank methods that researchers could focus on in performing additional evaluation runs.
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-\balance
-\bibliographystyle{abbrv}
-{\scriptsize\bibliography{Bibliography}}
-\onecolumn
-% sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%APPENDICES are optional
+
+\bibliographystyle{apa}
+\bibliography{Bibliography}
+
 \appendix
+\section{Meta-analysis raw data}
 
 \begin{table*}[!hpt]
-\scriptsize{
+\tiny{
 \begin{tabular}{lll|lll}\toprule
@@
 
-\scriptsize{
+\tiny{
 \begin{longtable}[!hpt]{@{}lrrrrrrrrrrrrrrrr@{}}\toprule

### 2014-09-15 paper/IPM/paper.tex
title page
apa + png->jpg
2->1 column

@@
-\documentclass[preprint]{elsarticle}
+\documentclass[english, authoryear, preprint]{elsarticle}
 \usepackage{booktabs}
@@
 \usepackage{dblfloatfix}
+\usepackage{apalike}
 \usepackage{fixltx2e}
@@
 \title{A Cross-Benchmark Comparison of 87 Learning to Rank Methods}
-
 \author[avanade,ut]{Niek Tax\corref{cor}\fnref{fn}}
 \ead{niek.tax@avanade.com}
-
 \author[avanade]{Sander Bockting}
 \ead{sander.bockting@avanade.com}
-
 \author[ut]{Djoerd Hiemstra}
 \ead{d.hiemstra@utwente.nl}
-
-\cortext[cor]{Corresponding author}
+\cortext[cor]{Corresponding Author: Niek Tax, Avanade Netherlands B.V., Versterkerstraat 6,
+1322AP Almere, the Netherlands; Email, niek.tax@avanade.com; Phone, +31634085760}
 \fntext[fn]{Author is affiliated with the University of Twente, but this paper written during his time at Avanade Netherlands B.V.}
-
 \address[avanade]{Avanade Netherlands B.V., Versterkerstraat 6,
@@
 \centering
-\includegraphics[scale=0.19]{gfx/ndcg3_winnum}
+\includegraphics[scale=0.26]{gfx/ndcg3_winnum}
 \caption{NDCG@3 comparison of 87 learning to rank methods}
@@
 \begin{figure}
-\includegraphics[scale=0.19]{gfx/ndcg5_winnum}
+\centering
+\includegraphics[scale=0.26]{gfx/ndcg5_winnum}
 \caption{NDCG@5 comparison of 87 learning to rank methods}
@@
 \centering
-\includegraphics[scale=0.19]{gfx/ndcg10_winnum}
+\includegraphics[scale=0.26]{gfx/ndcg10_winnum}
 \caption{NDCG@10 comparison of 87 learning to rank methods}
@@
 \centering
-\includegraphics[scale=0.19]{gfx/map_winnum}
+\includegraphics[scale=0.26]{gfx/map_winnum}
 \caption{MAP comparison of 87 learning to rank methods}
@@
 \centering
-\includegraphics[scale=0.21]{gfx/combined_normalized_winnum}
+\includegraphics[scale=0.28]{gfx/combined_normalized_winnum}
 \caption{Cross-benchmark comparison of 87 learning to rank methods}
@@
 
-\bibliographystyle{apa}
+\bibliographystyle{apalike}
 \bibliography{Bibliography}
@@
 \tiny{
-\begin{tabular}{lll|lll}\toprule
-Method & Described in & Evaluated in & Method & Described in & Evaluated in\\
+\begin{tabular}{lll}\toprule
+Method & Described in & Evaluated in\\
 \midrule
-AdaRank-MAP & \cite{Xu2007} & L2, L3, L4 & ListMLE & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\
-AdaRank-NDCG & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & ListNet & \cite{Cao2007} & L2, L3, L4 \\ 
-ADMM & \cite{Duh2011} & \cite{Duh2011} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
-ApproxAP & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
-ApproxNDCG & \cite{Qin2010b} & \cite{Qin2010b} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
-BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MHR & \cite{Qin2007} & L2 \\ 
-Best Single Feature &  & \cite{Gomes2013} &  MultiStageBoost & \cite{Kao2013} & \cite{Kao2013} \\ 
-BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} &  NewLoss & \cite{Peng2010} & \cite{Peng2010} \\ 
-BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009, Volkovs2013} & OWPC & \cite{Usunier2009} & \cite{Usunier2009} \\ 
-BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c, Volkovs2013} & PERF-MAP & \cite{Pan2011} & \cite{Torkestani2012b} \\ 
-BT & \cite{Zhou2008} & \cite{Zhou2008} & PermuRank & \cite{Xu2008} & \cite{Xu2008} \\ 
-C-CRF & \cite{Qin2008b} & \cite{Qin2008b} & Q.D.KNN & \cite{Geng2008} & \cite{Wang2013} \\ 
-CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} & RandomForest & \cite{Gomes2013} & \cite{Gomes2013} \\ 
-CCRank & \cite{Wang2011c} & \cite{Wang2011c} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
-CoList & \cite{Gao2014} & \cite{Gao2014} & RankAggNDCG & \cite{Wang2013} & \cite{Wang2013} \\
-Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010, Sato2013} \\
-DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost (Kernel-PCA) & \cite{Duh2008} & \cite{Duh2008, Sato2013} \\
-DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankBoost (SVD) & \cite{Lin2009} & \cite{Lin2009} \\
-EnergyNDCG & \cite{Freno2011} & \cite{Freno2011} & RankCSA & \cite{He2010} & \cite{He2010} \\
-FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\
-FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b, Laporte2013} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
-FocusedBoost & \cite{Niu2012} & \cite{Niu2012} &  RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
-FocusedNet & \cite{Niu2012} & \cite{Niu2012} & RankMGP & \cite{Lin2012} & \cite{Lin2012} \\ 
-FocusedSVM & \cite{Niu2012} & \cite{Niu2012} & RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013, Papini2012, Niu2012} \\ 
-FP-Rank & \cite{Song2013} & \cite{Song2013} & RankRLS & \cite{Pahikkala2009} & \cite{Pahikkala2010} \\ 
-FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} & RankSVM & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010, Papini2012} \\ 
-FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} & RankSVM-Primal &  & L3, \cite{Lai2011} \\ 
-FSM$^{SVM}$ & \cite{Lai2013c} & \cite{Lai2013c} & RankSVM-Struct &  & L3, L4 \\
-GAS-E & \cite{Geng2007} & \cite{Lai2013c} & RCP & \cite{Elsas2008} & \cite{Elsas2008} \\ 
-GP & \cite{DeAlmeida2007} & \cite{Alcantara2010} & RE-QR & \cite{Veloso2010} & \cite{Veloso2010} \\
-GPRank & \cite{Silva2009} & \cite{Torkestani2012} & REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009} \\  
-GRankRLS & \cite{Pahikkala2010} & \cite{Pahikkala2010} & Ridge Regression & \cite{Cossock2006} & L3 \\
-GroupCE & \cite{Lin2011} & \cite{Lin2011} & RSRank & \cite{Sun2009} & \cite{Lai2013} \\ 
-GroupMLE & \cite{Lin2010} & \cite{Lin2011} & SmoothGrad & \cite{Le2007} & \cite{Tan2013} \\ 
-IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} & SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010} \\
-IPRank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} & SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b} \\ 
-KeepRank & \cite{Chen2009} & \cite{Chen2009} & SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011, Papini2012} \\
-KL-CRF & \cite{Volkovs2011} & \cite{Volkovs2011} & SparseRank & \cite{Lai2013b} & \cite{Lai2013b} \\ 
-LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008, Alcantara2010} & SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012} \\ 
-LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} &SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013} \\ 
-LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} & TGRank & \cite{Lai2013} & \cite{Lai2013} \\ 
-LambdaRank & \cite{Burges2006} & \cite{Papini2012, Tan2013} & TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013} \\ 
-LARF & \cite{Torkestani2012} & \cite{Torkestani2012} &  VFLR & \cite{Cai2012} & \cite{Cai2012} \\
-Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} & & & \\
+AdaRank-MAP & \cite{Xu2007} & L2, L3, L4 \\
+AdaRank-NDCG & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} \\ 
+ADMM & \cite{Duh2011} & \cite{Duh2011} \\ 
+ApproxAP & \cite{Qin2010b} & \cite{Qin2010b} \\ 
+ApproxNDCG & \cite{Qin2010b} & \cite{Qin2010b} \\ 
+BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} \\ 
+Best Single Feature &  & \cite{Gomes2013} \\ 
+BL-MART & \cite{Ganjisaffar2011c} & \cite{Ganjisaffar2011c} \\ 
+BoltzRank-Single & \cite{Volkovs2009} & \cite{Volkovs2009, Volkovs2013} \\ 
+BoltzRank-Pair & \cite{Volkovs2009} & \cite{Volkovs2009, Ganjisaffar2011c, Volkovs2013} \\ 
+BT & \cite{Zhou2008} & \cite{Zhou2008} \\ 
+C-CRF & \cite{Qin2008b} & \cite{Qin2008b} \\ 
+CA & \cite{Metzler2007} & \cite{Busa-Fekete2013,Tan2013} \\ 
+CCRank & \cite{Wang2011c} & \cite{Wang2011c} \\ 
+CoList & \cite{Gao2014} & \cite{Gao2014} \\
+Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} \\
+DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  \\
+DirectRank & \cite{Tan2013} & \cite{Tan2013} \\
+EnergyNDCG & \cite{Freno2011} & \cite{Freno2011} \\
+FBPCRank & \cite{Lai2011} & \cite{Lai2011} \\
+FenchelRank & \cite{Lai2013} & \cite{Lai2013, Lai2013b, Laporte2013} \\ 
+FocusedBoost & \cite{Niu2012} & \cite{Niu2012} \\ 
+FocusedNet & \cite{Niu2012} & \cite{Niu2012} \\ 
+FocusedSVM & \cite{Niu2012} & \cite{Niu2012} \\ 
+FP-Rank & \cite{Song2013} & \cite{Song2013} \\ 
+FRank & \cite{Tsai2007} & L2, L3, \cite{Wang2012} \\ 
+FSMRank & \cite{Lai2013c} & \cite{Lai2013c,Laporte2013} \\ 
+FSM$^{SVM}$ & \cite{Lai2013c} & \cite{Lai2013c} \\
+GAS-E & \cite{Geng2007} & \cite{Lai2013c} \\ 
+GP & \cite{DeAlmeida2007} & \cite{Alcantara2010} \\
+GPRank & \cite{Silva2009} & \cite{Torkestani2012} \\  
+GRankRLS & \cite{Pahikkala2010} & \cite{Pahikkala2010} \\
+GroupCE & \cite{Lin2011} & \cite{Lin2011} \\ 
+GroupMLE & \cite{Lin2010} & \cite{Lin2011} \\ 
+IntervalRank & \cite{Moon2010} & \cite{Moon2010, Freno2011} \\
+IPRank & \cite{Wang2009b} & \cite{Wang2009b, Torkestani2012} \\ 
+KeepRank & \cite{Chen2009} & \cite{Chen2009} \\
+KL-CRF & \cite{Volkovs2011} & \cite{Volkovs2011} \\ 
+LAC-MR-OR & \cite{Veloso2008} & \cite{Veloso2008, Alcantara2010} \\ 
+LambdaMART & \cite{Burges2010} & \cite{Asadi2013a, Ganjisaffar2011c} \\ 
+LambdaNeuralRank & \cite{Papini2012} & \cite{Papini2012} \\ 
+LambdaRank & \cite{Burges2006} & \cite{Papini2012, Tan2013} \\ 
+LARF & \cite{Torkestani2012} & \cite{Torkestani2012} \\
+Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} \\
+ListMLE & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\
+ListNet & \cite{Cao2007} & L2, L3, L4 \\
+ListReg & \cite{Wu2011} & \cite{Wu2011}\\
+LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b}\\
+MCP & \cite{Laporte2013} & \cite{Laporte2013}\\
+MHR & \cite{Qin2007} & L2\\
+MultiStageBoost & \cite{Kao2013} & \cite{Kao2013}\\
+NewLoss & \cite{Peng2010} & \cite{Peng2010}\\
+OWPC & \cite{Usunier2009} & \cite{Usunier2009}\\
+PERF-MAP & \cite{Pan2011} & \cite{Torkestani2012b}\\
+PermuRank & \cite{Xu2008} & \cite{Xu2008}\\
+Q.D.KNN & \cite{Geng2008} & \cite{Wang2013}\\
+RandomForest & \cite{Gomes2013} & \cite{Gomes2013}\\
+Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013}\\
+RankAggNDCG & \cite{Wang2013} & \cite{Wang2013}\\
+RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010, Sato2013}\\
+RankBoost (Kernel-PCA) & \cite{Duh2008} & \cite{Duh2008, Sato2013}\\
+RankBoost (SVD) & \cite{Lin2009} & \cite{Lin2009}\\
+RankCSA & \cite{He2010} & \cite{He2010}\\
+RankDE & \cite{Bollegala2011} & \cite{Sato2013}\\
+RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013}\\
+RankELM (pointwise) & \cite{Zong2013} & \cite{Zong2013}\\
+RankMGP & \cite{Lin2012} & \cite{Lin2012}\\
+RankNet & \cite{Burges2005} & \cite{Busa-Fekete2013, Papini2012, Niu2012}\\
+RankRLS & \cite{Pahikkala2009} & \cite{Pahikkala2010}\\
+RankSVM & \cite{Herbrich1999, Joachims2002} & L2, L3, \cite{Busa-Fekete2013, Freno2011, He2010, Alcantara2010, Papini2012}\\
+RankSVM-Struct &  & L3, L4\\
+RankSVM-Primal &  & L3, \cite{Lai2011}\\
+RCP & \cite{Elsas2008} & \cite{Elsas2008}\\
+RE-QR & \cite{Veloso2010} & \cite{Veloso2010}\\
+REG-SHF-SDCG & \cite{Wu2009} & \cite{Wu2009}\\
+Ridge Regression & \cite{Cossock2006} & L3\\
+RSRank & \cite{Sun2009} & \cite{Lai2013}\\
+SmoothGrad & \cite{Le2007} & \cite{Tan2013}\\
+SmoothRank & \cite{Chapelle2010} & L3, \cite{Chapelle2010}\\
+SoftRank & \cite{Taylor2008, Guiver2008} & \cite{Qin2010b}\\
+SortNet & \cite{Rigutini2008} & \cite{Rigutini2008,Freno2011, Papini2012}\\
+SparseRank & \cite{Lai2013b} & \cite{Lai2013b}\\
+SVM$^{MAP}$ & \cite{Yue2007} & L3, \cite{Wang2012, Xu2008, Niu2012}\\
+SwarmRank & \cite{Diaz-Aviles2009} & \cite{Sato2013}\\
+TGRank & \cite{Lai2013} & \cite{Lai2013}\\
+TM & \cite{Zhou2008} & \cite{Zhou2008, Papini2012, Tan2013}\\
+VFLR & \cite{Cai2012} & \cite{Cai2012}\\
 \bottomrule

### 2014-09-15 paper/IPM/paper.tex
data sources table opmaak

@@
 Ranking has become a vital part of web search, where commercial search engines help users find their need in the extremely large collection of the World Wide Web.
-Among useful applications outside web search are automatic text summarisation, machine translation, drug discovery and determining the ideal order of maintenance operations \cite{Rudin2009}.
+Among useful applications outside web search are automatic text summarisation, machine translation, drug discovery and determining the ideal order of maintenance operations \citep{Rudin2009}.
 In addition, the ranking task has been argued to be a better fit for recommender systems than the regression task \cite{McNee2006} (which concerns predictions on a continuous scale) that is currently frequently used in such systems.\\
@@
 \tiny{
-\begin{tabular}{lll}\toprule
+\begin{tabular}{lp{3cm}p{6cm}}\toprule
 Method & Described in & Evaluated in\\

### 2014-09-15 paper/IPM/paper.tex
in-text citations gefixt naar APA + chinese-namen-terror gefixt

@@
 Among useful applications outside web search are automatic text summarisation, machine translation, drug discovery and determining the ideal order of maintenance operations \citep{Rudin2009}.
-In addition, the ranking task has been argued to be a better fit for recommender systems than the regression task \cite{McNee2006} (which concerns predictions on a continuous scale) that is currently frequently used in such systems.\\
+In addition, \cite{McNee2006} found the ranking task to be a better fit for recommender systems than the regression task (continuous scale predictions), which is currently still frequently used within such systems.\\
 
-Research in the field of ranking models has been based on manually designed ranking functions, such as the well-known BM25 model \cite{Robertson1994}.
+Research in the field of ranking models has been based on manually designed ranking functions, such as the well-known BM25 model \citep{Robertson1994}.
 Increased amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective ranking models.
@@
 Well-known benchmark datasets in the learning to rank field include the \emph{Yahoo!
-Learning to Rank Challenge} datasets \cite{Chapelle2011a}, the \emph{Yandex Internet Mathematics 2009} contest\footnote{http://imat2009.yandex.ru/en}, the LETOR datasets \cite{Qin2010}, and the MSLR (Microsoft Learning to Rank) datasets\footnote{http://research.microsoft.com/en-us/projects/mslr/}.
+Learning to Rank Challenge} datasets \citep{Chapelle2011a}, the \emph{Yandex Internet Mathematics 2009} contest\footnote{http://imat2009.yandex.ru/en}, the LETOR datasets \citep{Qin2010}, and the MSLR (Microsoft Learning to Rank) datasets\footnote{http://research.microsoft.com/en-us/projects/mslr/}.
 There exists no agreement among authors in the learning to rank field on the benchmark collection(s) to use to evaluate a new model.
@@
 
-Gomes et al.
 \cite{Gomes2013} analyzed the ranking accuracy of a set of models on both LETOR 3.0 and 4.0.
-Busa-Fekete et al.
 \cite{Busa-Fekete2013} compared the accuracy of a small set of models over the LETOR 4.0 datasets, both MSLR datasets, both the Yahoo!
@@
 We collect evaluation measurements of learning to rank methods on these benchmark collections through forward literature search.
-Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search performed using Google Scholar.
+Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search performed using Google Scholar.\\
 \begin{table}[!h]
@@
 \midrule
-LETOR 1.0 \& 2.0 & Liu et al. \cite{Liu2007b} & 307\\
-LETOR 3.0 & Qin et al. \cite{Qin2010} & 105\\
-Yahoo! Learning to Rank Challenge & Chapelle et al. \cite{Chapelle2011a} & 102\\
-AOL dataset & Pass et al. \cite{Pass2006} & 339\\
-WCL2R & Alc{\^a}ntara et al. \cite{Alcantara2010} & 2\\
+LETOR 1.0 \& 2.0 & \cite{Liu2007b} & 307\\
+LETOR 3.0 & \cite{Qin2010} & 105\\
+Yahoo! Learning to Rank Challenge & \cite{Chapelle2011a} & 102\\
+AOL dataset & \cite{Pass2006} & 339\\
+WCL2R & \cite{Alcantara2010} & 2\\
 \bottomrule
@@
 \begin{enumerate}[noitemsep,nolistsep]
-\item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011, Lin2012}
+\item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \citep{Geng2011, Lin2012}
 \item The study focuses on a different learning to rank task (e.g.
-rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}
-\item The study used an altered version of a benchmark that contained additional features \cite{Bidoki2009, Ding2010}
+rank aggregation or transfer ranking) \citep{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}
+\item The study used an altered version of a benchmark that contained additional features \citep{Bidoki2009, Ding2010}
 \item The study provides no exact data of the evaluation results (e.g.
-results are only in graphical form) \cite{Wang2008, Wang2010, Xu2010, Kuo2009, Li2008, Xia2008, Zhou2011, Wu2011, Zhu2009, Karimzadehgan2011, Swersky2012, Pan2011, Ni2008, Ciaramita2008, Stewart2012, Petterson2009, Agarwal2010, Chang2009, Qin2008c, Adams2011, Sculley2009, Huang2008, Alejo2010, Sun2011, He2010b, Benbouzid2012, Geng2012, Chen2012, Xu2012, Shivaswamy2011}
-\item The study reported evaluation results in a different metric than the metrics chosen for this meta-analysis \cite{Yu2009, Thuy2009, Pahikkala2009, Kersting2009, Mohan2011}
-\item The study reported a higher performance on baseline methods than official benchmark runs \cite{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}
-\item The study did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
+results are only in graphical form) \citep{Wang2008, Wang2010, Xu2010, Kuo2009, Li2008, Xia2008, Zhou2011, Wu2011, Zhu2009, Karimzadehgan2011, Swersky2012, Pan2011, Ni2008, Ciaramita2008, Stewart2012, Petterson2009, Agarwal2010, Chang2009, Qin2008c, Adams2011, Sculley2009, Huang2008, Alejo2010, Sun2011, He2010b, Benbouzid2012, Geng2012, Chen2012, Xu2012, Shivaswamy2011}
+\item The study reported evaluation results in a different metric than the metrics chosen for this meta-analysis \citep{Yu2009, Thuy2009, Pahikkala2009, Kersting2009, Mohan2011}
+\item The study reported a higher performance on baseline methods than official benchmark runs \citep{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}
+\item The study did not report any baseline performance that allowed us to check validity of the results \citep{Chakrabarti2008, Wang2012b, Buffoni2011}.
 \end{enumerate}
@@
 \section{A Methodology for Comparing Learning to Rank Methods Cross-Benchmark}
-Qin et al.
 \cite{Qin2010} state that it may differ between datasets what the most accurate ranking methods are.

### 2014-09-16 paper/IPM/paper.tex
appendix op 1 pagina gefit
highlights + code cleaning SmoothRank

@@
 1322AP Almere, the Netherlands; Email, niek.tax@avanade.com; Phone, +31634085760}
-\fntext[fn]{Author is affiliated with the University of Twente, but this paper written during his time at Avanade Netherlands B.V.}
+\fntext[fn]{Author is affiliated with the University of Twente, but this paper was written during his stay at Avanade Netherlands B.V.}
 \address[avanade]{Avanade Netherlands B.V., Versterkerstraat 6,
@@
 \bibliography{Bibliography}
+\newpage
+\appendix
+\section{Meta-analysis Ranking Methods \& Data Sources}
 
-\appendix
-\section{Meta-analysis raw data}
-
-\begin{table*}[!hpt]
-\tiny{
-\begin{tabular}{lp{3cm}p{6cm}}\toprule
+{\fontsize{5.5pt}{5.4pt}\selectfont
+\begin{longtable}{p{2.671cm}p{2.812cm}p{7.7cm}}\toprule
 Method & Described in & Evaluated in\\
@@
 \bottomrule
-\end{tabular}
-\caption{Learning to rank algorithms with measurements on benchmark datasets}}
+\caption{Learning to rank algorithms with measurements on benchmark datasets}
 \label{tab:ltr_methods_used}
-\end{table*}
+\end{longtable}
+}
 
-\tiny{
-\begin{longtable}[!hpt]{@{}lrrrrrrrrrrrrrrrr@{}}\toprule
+\newpage
+\section{Meta-analysis Raw Data}
+{\fontsize{5.5pt}{5.4pt}\selectfont
+\begin{longtable}{@{}lrrrrrrrrrrrrrrrr@{}}\toprule
 & \multicolumn{2}{c}{NDCG@3} & \phantom{a} 

### 2014-09-16 implementation/java/src/letor/serial/SmoothRank.scala

@@ -2,8 +2,6 @@ package letor.serial;
 
 import ciir.umass.edu.learning.{RankList, Ranker, DataPoint}
 import ciir.umass.edu.utilities.Sorter
-import org.ejml.data.{DenseMatrix64F, ReshapeMatrix64F}
-import org.ejml.ops.CommonOps
 import scala.collection.JavaConverters._
 
 /*
@@ -109,20 +107,6 @@ class SmoothRank extends Ranker{
     val wEvaluated = (0 until rl.size).map(x => eval(rl.get(x)))
     val idx = Sorter.sort(wEvaluated.toArray, false)
     new RankList(rl, idx)
-  }
-
-  def addToPseudoSolver(rl:RankList, a:ReshapeMatrix64F, b:ReshapeMatrix64F) = {
-    var aRl = new DenseMatrix64F(rl.size, features.length)
-    var bRl = new DenseMatrix64F(rl.size, 1)
-    for(i <- 0 until rl.size){
-      var xElem = List[Double]()
-      val dp = rl.get(i)
-      for(j <- 1 to features.length)
-        aRl.set(i, j-1, dp.getFeatureValue(j).toDouble)
-      bRl.set(i, 0, Math.pow(2, dp.getLabel)-1 )
-    }
-    CommonOps.insert(aRl,a,0,0)
-    CommonOps.insert(bRl,b,0,0)
   }
 
   def scaleFeatures(samples:java.util.List[RankList], numFeatures:Int):Array[RankList] = {

### 2014-09-16 implementation/java/src/letor/serial/WrappedRunner.java

@@ -1,6 +1,7 @@
 package letor.serial;
 
 import letor.parallel.util.DataSets;
+import letor.serial.parameterizedRankers.AbstractParameterizedRanker;
 import letor.serial.parameterizedRankers.ListNetHandler;
 import letor.serial.parameterizedRankers.SmoothRankHandler;
 import letor.serial.util.Measurement;
@@ -11,14 +12,15 @@ import letor.serial.util.Measurement;
 public class WrappedRunner {
 
     public static void main(String[] args){
-        DataSets.DataSet    dataset             = DataSets.DataSet.CUSTOM;
-        Integer             duplicationNumber   = 7; // Only relevant in case of dataset Custom
+        AbstractParameterizedRanker handler             = new SmoothRankHandler();
+        DataSets.DataSet            dataset             = DataSets.DataSet.CUSTOM;
+        Integer                     duplicationNumber   = 4; // Only relevant in case of dataset Custom
 
         FoldRunHandler listNetWrapper = null;
         if(dataset == DataSets.DataSet.CUSTOM)
-            listNetWrapper = new FoldRunHandler(new ListNetHandler(), DataSets.DataSet.CUSTOM, duplicationNumber,  1, 1); // Method, dataset, folds, iterations
+            listNetWrapper = new FoldRunHandler(handler, DataSets.DataSet.CUSTOM, duplicationNumber,  1, 1); // Method, dataset, folds, iterations
         else
-            listNetWrapper = new FoldRunHandler(new ListNetHandler(), dataset, 1, 1); // Method, dataset, folds, iterations
+            listNetWrapper = new FoldRunHandler(handler, dataset, 1, 1); // Method, dataset, folds, iterations
         Measurement measurement = listNetWrapper.averageScore();
         System.out.println("Preprocessing: "+measurement.getPreprocessingTime());
         System.out.println("Train:         "+measurement.getTrainingTime());

### 2014-09-16 implementation/java/src/letor/serial/WrappedRunner.java
refactor

@@ -16,12 +16,12 @@ public class WrappedRunner {
         DataSets.DataSet            dataset             = DataSets.DataSet.CUSTOM;
         Integer                     duplicationNumber   = 4; // Only relevant in case of dataset Custom
 
-        FoldRunHandler listNetWrapper = null;
+        FoldRunHandler ltrWrapper = null;
         if(dataset == DataSets.DataSet.CUSTOM)
-            listNetWrapper = new FoldRunHandler(handler, DataSets.DataSet.CUSTOM, duplicationNumber,  1, 1); // Method, dataset, folds, iterations
+            ltrWrapper = new FoldRunHandler(handler, DataSets.DataSet.CUSTOM, duplicationNumber,  1, 1); // Method, dataset, folds, iterations
         else
-            listNetWrapper = new FoldRunHandler(handler, dataset, 1, 1); // Method, dataset, folds, iterations
-        Measurement measurement = listNetWrapper.averageScore();
+            ltrWrapper = new FoldRunHandler(handler, dataset, 1, 1); // Method, dataset, folds, iterations
+        Measurement measurement = ltrWrapper.averageScore();
         System.out.println("Preprocessing: "+measurement.getPreprocessingTime());
         System.out.println("Train:         "+measurement.getTrainingTime());
         System.out.println("Test:          "+measurement.getTestTime());

### 2014-09-16 measurements/faceted_graph.R
overbodig script

@@ -1,42 +1,0 @@
-setwd("C:/Git-data/Afstuderen/measurements")
-library(ggplot2)
-library(mgcv)
-library(stringr)
-library(scales)
-measure = read.csv("raw_data.csv") # read csv file
-name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
-
-getName <- function(s){
-	return(ifelse(grepl("^Parallel (\\d)+$", s), 
-			paste(
-				c(
-					"Cluster: ", 
-					sprintf("% 3d",as.numeric(str_extract(s, "\\d+"))), 
-					" data nodes"
-				),
-				collapse = ''
-			), 
-			"Serial"
-		)
-	)
-}
-
-name <- unlist(lapply(name, getName))
-measure$seconds <- measure$trainTime/1000
-measure$throughput <- measure$dataSize/measure$seconds
-measure$serial.parallel <- name
-d <- ggplot(data=measure, aes(x=dataSize, y=seconds, colour=serial.parallel, shape=serial.parallel))
-d <- d + geom_point(size=2) + stat_smooth(level=0.99, method = "glm", formula = y ~ exp(x)) + 
-	scale_x_log10() + theme_bw() + 
-	theme(
-		panel.grid.major = element_blank(),
-		panel.grid.minor = element_blank(),
-		panel.border     = element_blank(),
-		panel.background = element_blank(),
-		axis.line = element_line(color='black')
-	) + scale_y_continuous(expand = c(0, 0))+	
-	facet_wrap(~serial.parallel)+
-	xlab("Dataset size (in Byte)") +
-	ylab("Training iteration time (in Seconds)") +
-	labs(colour = "Execution mode", shape = "Execution mode") 
-ggplot_build(d)

### 2014-09-16 report/Chapters/Implementation.tex
uitbreiding storage

--- 2014-09-08 report/Chapters/Implementation.tex
@@
-This chapter describes implementation details of the Learning-to-Rank algorithm parallel implementations in HDInsight that are either Hadoop, Microsoft Azure, or HDInsight and are not part of the algorithm specification itself.
+\label{chap:implementation}
 The first section of this chapter will briefly discuss the HDInsight platform, the Hadoop ecosystem components offered by HDInsight and, in this regard, the Hadoop components used for implementation of the algorithms described in Chapter \ref{chap:ltr_methods}.
+The second section of this chapter describes a Java framework that handles the joint components needed for MapReduce Learning to Rank computation that are independent of the Learning to Rank model.
+The subsequent sections describe the implementation details of specific Learning to Rank models.
 
 \section{Architecture}
-Ranking models almost without exception consist of a sequence of operations on the input data, that are in most cases repeatedly executed in a series of iterations.
-Apache Pig will be used to implement the sequence of operation on the input data, because the higher level of abstraction that Apache Pig offers makes it easier to focus solely on the data operations as compared to native Hadoop MapReduce.\\
+Ranking algorithms consist of a sequence of operations on the input data which are often, but not always, of iterative nature.
+Apache Pig \cite{Olston2008} will be used to implement the sequence of operation on the input data, because the higher level of abstraction that Apache Pig offers makes it easier to focus solely on the data operations as compared to native Hadoop MapReduce.\\
 
-% TODO: Stuk over werkzaamheden framework uitbreiden: uitlezen resultaten (evaluatieresultaten + doorgeven tussenresultaten (die tweede wellicht later in het document pas introduceren))
-In a cross-validation evaluation experiment of a ranking model, the handling of which folds are loaded for training, evaluation and testing is a task that need to be taken care of regardless of the ranking model.
-Given that most ranking models are of an iterative nature, the task of iteration handling is also a task that need to be done for (most) models.
-Since these tasks need to be considered for (almost) all ranking models it seems natural to develop a framework that takes care of these tasks.
-The aim of this framework is to let implementations of ranking models focus solely on implementing the sequential steps of one iteration of the algorithm, while the framework arranges that 1) these sequential steps are performed iteratively and 2) these sequential steps are performed on the multiple training folds of data.
-This framework that handles folding and iterations will be implemented in Java and will work such that fold- and iteration dependent parts of the Pig code will be generated dynamically by the Java framework after which the Pig job will be sent to the cluster.\\
+\subsection{Storing Input Data for HDInsight Processing}
+Azure HDInsight supports both the traditional \ac{HDFS} as described in \cite{Shvachko2010}, as well as Microsoft's own \ac{WASB} storage solution.
+Blob storage decouples the storage from the HDInsight Hadoop cluster; it enables safe deletion of a HDInsight cluster without losing data, as data is not solely stored on the cluster itself, but also on a separate storage that is not cluster-dependent.
+Azure \ac{WASB} storage allows the user to select one of their data centre for storage.
+\ac{WASB} storage in the West Europe region (located in Amsterdam) will be used for storage, as this data centre is located close to where the experiments are executed.
 
-\subsection{Microsoft Azure HDInsight}
+\subsection{Microsoft Azure HDInsight Job submission}
 Microsoft offers a scalable and on-demand Hadoop service with HDInsight, that enables Hadoop services for those not able to make the required investments for their own cluster.
@@
 \centering
-\begin{tabular}{p{4cm}p{2.8cm}p{5.5cm}}\toprule
+\begin{tabular}{p{4.2cm}p{2.6cm}p{5.5cm}}\toprule
 Job submission method & Type & Description \\
@@
 C\# API & C\# API & A wrapper API is offered to submit Hadoop MapReduce jobs directly from C\# code.\\
-HiveServer/HiveServer2 & REST endpoint & HiveServer and its successor HiveServer 2 are, as its name suggests, REST endpoints that allow remote submission of Hive queries.\\
-Oozie & REST endpoint & Apache Oozie is a workflow scheduler to manage Apache Hadoop jobs.
+HiveServer/HiveServer2 & REST endpoint & Apache Hive \cite{Thusoo2009} is an open-source data warehousing solution on top of Hadoop, that supports processing of a SQL-like declarative language called HiveQL.
+HiveServer and its successor HiveServer 2 are REST endpoints that allow remote submission of HiveQL queries.\\
+Oozie & REST endpoint & Apache Oozie \cite{Islam2012} is a workflow scheduler to manage Apache Hadoop jobs.
 Oozie enables users to specify Directed Acyclical Graphs of action, where each action is specified in either MapReduce or Pig.\\
@@
 Below the necessary procedures for job submission from Oozie as well as from WebHCat will be sketched.\\
+
+% TODO: Stuk over werkzaamheden framework uitbreiden: uitlezen resultaten (evaluatieresultaten + doorgeven tussenresultaten (die tweede wellicht later in het document pas introduceren))
+In a cross-validation evaluation experiment of a ranking model, the handling of which folds are loaded for training, evaluation and testing is a task that need to be taken care of regardless of the ranking model.
+Given that most ranking models are of an iterative nature, the task of iteration handling is also a task that need to be done for (most) models.
+Since these tasks need to be considered for (almost) all ranking models it seems natural to develop a framework that takes care of these tasks.
+The aim of this framework is to let implementations of ranking models focus solely on implementing the sequential steps of one iteration of the algorithm, while the framework arranges that 1) these sequential steps are performed iteratively and 2) these sequential steps are performed on the multiple training folds of data.
+This framework that handles folding and iterations will be implemented in Java and will work such that fold- and iteration dependent parts of the Pig code will be generated dynamically by the Java framework after which the Pig job will be sent to the cluster.\\
 

### 2014-09-16 report/Chapters/Implementation.tex
uitbreiding implementation

@@
 Ranking algorithms consist of a sequence of operations on the input data which are often, but not always, of iterative nature.
-Apache Pig \cite{Olston2008} will be used to implement the sequence of operation on the input data, because the higher level of abstraction that Apache Pig offers makes it easier to focus solely on the data operations as compared to native Hadoop MapReduce.\\
+Apache Pig \cite{Olston2008} will be used to implement the sequence of operation on the input data.
+Pig Latin, the data processing language that runs on Apache Pig, designed based on the observation that the traditional MapReduce paradigm is too low-level and rigid, and holds the middle between the declarative style of SQL and the procedural style of MapReduce.
+The Apache Pig system translates Pig Latin into MapReduce plans that are executed over Hadoop.
+The choice to implement the Learning to Rank algorithms in Pig Latin allow more focus on the data operations and less focus on low-level implementation details compared to native Hadoop MapReduce.
+Furthermore, it allows us to rely on Apache Pig to create efficient MapReduce plans out of the Pig Latin code which lowers the implementation-dependent factor of the experiments.\\
 
 \subsection{Storing Input Data for HDInsight Processing}
-Azure HDInsight supports both the traditional \ac{HDFS} as described in \cite{Shvachko2010}, as well as Microsoft's own \ac{WASB} storage solution.
+Azure HDInsight supports both the traditional \ac{HDFS} as described by Shvacko et al.
+\cite{Shvachko2010}, as well as Microsoft's own \ac{WASB} storage solution.
 Blob storage decouples the storage from the HDInsight Hadoop cluster; it enables safe deletion of a HDInsight cluster without losing data, as data is not solely stored on the cluster itself, but also on a separate storage that is not cluster-dependent.
@@
 % TODO: Stuk over werkzaamheden framework uitbreiden: uitlezen resultaten (evaluatieresultaten + doorgeven tussenresultaten (die tweede wellicht later in het document pas introduceren))
-In a cross-validation evaluation experiment of a ranking model, the handling of which folds are loaded for training, evaluation and testing is a task that need to be taken care of regardless of the ranking model.
-Given that most ranking models are of an iterative nature, the task of iteration handling is also a task that need to be done for (most) models.
-Since these tasks need to be considered for (almost) all ranking models it seems natural to develop a framework that takes care of these tasks.
+Fold handling  in a cross-validation experiment is the process of loading the correct data folds for training, validation and testing in the multiple rounds that the cross-validation experiment consists of.
+Fold handling is a task that needs to be taken care of regardless of the ranking model that is being evaluated.
+Given that most ranking models are of an iterative nature, the task of iteration handling is also a task that need to be done for most models.
+Iteration handling and fold handling are procedures in which the same steps are repeated for each iteration or cross-validation round respectively.
+Pig Latin, however, has no support for loops which are needed to perform iteration handling and fold handling.
+Since iteration handling and fold handling are tasks that need to be addressed for each ranking model and that cannot be solved with Pig, we create a framework that takes care of both iteration and fold handling and generates the Pig Latin code for the current iteration and fold.
 The aim of this framework is to let implementations of ranking models focus solely on implementing the sequential steps of one iteration of the algorithm, while the framework arranges that 1) these sequential steps are performed iteratively and 2) these sequential steps are performed on the multiple training folds of data.
 This framework that handles folding and iterations will be implemented in Java and will work such that fold- and iteration dependent parts of the Pig code will be generated dynamically by the Java framework after which the Pig job will be sent to the cluster.\\
+
+Another responsibility of the framework is to handle communication between multiple Pig jobs within a single iteration of a Learning to Rank algorithm.
+The framework enables reading the result of a Pig job from \ac{WASB} storage which can then be used as parameter within a subsequent Pig job.
+An example of this is a Pig implementation of gradient descent, where a first Pig job might calculate gradients and writes them to storage, after which the framework assists in reading the gradients from storage and allows them to be used as input parameters of a second Pig job that calculates new feature weights based on its gradient parameter and a predetermined step size.\\
 

### 2014-09-16 implementation/java/src/letor/serial/WrappedRunner.java

@@ -14,7 +14,7 @@ public class WrappedRunner {
     public static void main(String[] args){
         AbstractParameterizedRanker handler             = new SmoothRankHandler();
         DataSets.DataSet            dataset             = DataSets.DataSet.CUSTOM;
-        Integer                     duplicationNumber   = 4; // Only relevant in case of dataset Custom
+        Integer                     duplicationNumber   = 1; // Only relevant in case of dataset Custom
 
         FoldRunHandler ltrWrapper = null;
         if(dataset == DataSets.DataSet.CUSTOM)

### 2014-09-16 report/Chapters/Implementation.tex
uitbreiding implementation

@@
 
-\subsection{Microsoft Azure HDInsight Job submission}
+\subsection{The HDInsight Platform}
 Microsoft offers a scalable and on-demand Hadoop service with HDInsight, that enables Hadoop services for those not able to make the required investments for their own cluster.
 The latest HDInsight of HDInsight available at the time of writing, HDInsight 3.1, runs the Hortonworks distribution of Hadoop version 2.4.
-HDInsight 3.1 offers multiple ways of submitting Hadoop jobs to a HDInsight cluster, described in Table \ref{tbl:hdinsight_endpoints}.\\
+While early versions of Hadoop were merely an open source implementation of MapReduce, Hadoop offers a wide variety of programming models since the introduction of Hadoop YARN in Hadoop 2.0 \cite{Vavilapalli2013}.
+Newly supported programming models since Hadoop 2.0 include Dryad \cite{Isard2007}, Giraph \cite{Avery2011}, MPI, REEF \cite{Chun2013}, Spark \cite{Zaharia2010}, and Storm \cite{Aniello2013}.
+Even tough listed programming models are now supported by Hadoop and even though some of these programming models have recently increased in popularity, they still lack the critical mass as the data processing framework of choice as Lin argued back in 2012 \cite{Lin2013}.
+Therefore, even though some of these programming models might be a better fit for iterative algorithms, we use Hadoop MapReduce of the programming model to implement the Learning to Rank algorithms.\\
+
+HDInsight 3.1 offers multiple ways of submitting Hadoop jobs to a HDInsight cluster, described in Table \ref{tbl:hdinsight_endpoints}.
+Oozie and WebHCat/Templeton are the two methods for job submission in Table \ref{tbl:hdinsight_endpoints} that both 1) support Apache Pig jobs, and 2) Can be used from within Java code.
+Table \ref{tbl:oozie_templeton} shows the necessary procedures for job submission from Oozie as well as from WebHCat will be sketched.
+Table \ref{tbl:oozie_templeton} shows Oozie job submission to be more complex for the case of dynamically generated jobs than WebHCat/Templeton.
+Oozie is more fitting for static Hadoop jobs that require a workflow consisting of a sequence of Pig and MapReduce jobs mixed together, but is a lesser fit for our situation.
+Templeton will be used for submission of the Pig jobs, as the HDInsight version that was available at the start of the implementation did not yet support WebHCat.\\
+
+\begin{table}
+\centering
+\begin{tabular}{p{5cm}p{5cm}}\toprule
+Job submission procedure for Apache Oozie & Job submission procedure for WebHCat/Templeton \\
+\midrule
+1. Let the framework build the Pig job dynamically. & 1. Let the framework build the Pig job dynamically.\\
+2. Encapsulate the Pig job in an Oozie workflow. & 2. Submit the Pig job through the WebHCat/Templeton REST API.\\
+3. Upload the Oozie workflow to HDFS storage. & \\
+4. Execute the Oozie workflow through the Oozie REST API. & \\
+\bottomrule
+\end{tabular}
+\caption{Comparison of Oozie and WebHCat job submission procedures}
+\label{tbl:oozie_templeton}
+\end{table}
 
@@
 
-Oozie and WebHCat/Templeton are the two methods for job submission in Table \ref{tbl:hdinsight_endpoints} that both 1) support Apache Pig jobs, and 2) Can be used from within Java code.
-Below the necessary procedures for job submission from Oozie as well as from WebHCat will be sketched.\\
-
-% TODO: Stuk over werkzaamheden framework uitbreiden: uitlezen resultaten (evaluatieresultaten + doorgeven tussenresultaten (die tweede wellicht later in het document pas introduceren))
+\subsection{Handling repetitive tasks}
 Fold handling  in a cross-validation experiment is the process of loading the correct data folds for training, validation and testing in the multiple rounds that the cross-validation experiment consists of.
@@
 Iteration handling and fold handling are procedures in which the same steps are repeated for each iteration or cross-validation round respectively.
-Pig Latin, however, has no support for loops which are needed to perform iteration handling and fold handling.
-Since iteration handling and fold handling are tasks that need to be addressed for each ranking model and that cannot be solved with Pig, we create a framework that takes care of both iteration and fold handling and generates the Pig Latin code for the current iteration and fold.
-The aim of this framework is to let implementations of ranking models focus solely on implementing the sequential steps of one iteration of the algorithm, while the framework arranges that 1) these sequential steps are performed iteratively and 2) these sequential steps are performed on the multiple training folds of data.
-This framework that handles folding and iterations will be implemented in Java and will work such that fold- and iteration dependent parts of the Pig code will be generated dynamically by the Java framework after which the Pig job will be sent to the cluster.\\
+Pig Latin, in contrast to more procedural languages that generate MapReduce job like Sawzall \cite{Pike2005}, has no support for loops which are needed to perform iteration handling and fold handling.
+Since iteration handling and fold handling are tasks that need to be addressed for each ranking model and that cannot be solved with Pig, we create a framework that takes care of both iteration and fold handling and generates the Pig Latin code for the current iteration and fold.\\
 
-Another responsibility of the framework is to handle communication between multiple Pig jobs within a single iteration of a Learning to Rank algorithm.
+Handling communication between Pig jobs within a single iteration of a Learning to Rank algorithm is another problem that needs to be addressed.
+A Pig Job writes its result to \ac{WASB} storage, while this result is in most cases needed by a subsequent Pig job as parameter.
 The framework enables reading the result of a Pig job from \ac{WASB} storage which can then be used as parameter within a subsequent Pig job.
@@
 
-\begin{table}
-\centering
-\begin{tabular}{p{5cm}p{5cm}}\toprule
-Job submission procedure for Apache Oozie & Job submission procedure for WebHCat/Templeton \\
-\midrule
-1. Let the framework build the Pig job dynamically. & 1. Let the framework build the Pig job dynamically.\\
-2. Encapsulate the Pig job in an Oozie workflow. & 2. Submit the Pig job through the WebHCat/Templeton REST API.\\
-3. Upload the Oozie workflow to HDFS storage. & \\
-4. Execute the Oozie workflow through the Oozie REST API. & \\
-\bottomrule
-\end{tabular}
-\caption{Comparison of Oozie and WebHCat job submission procedures}
-\label{tbl:oozie_templeton}
-\end{table}
+The aim of this framework is to let implementations of ranking models focus solely on implementing the sequential steps of one iteration of the algorithm, while the framework handles that 1) these sequential steps are performed iteratively, 2) these sequential steps are performed on the multiple training folds of data and 3) data can be passed from one Pig job within the algorithm to another Pig job.
+This framework that handles folding and iterations will be implemented in Java and will work such that fold- and iteration dependent parts of the Pig code will be generated dynamically by the Java framework after which the Pig job will be sent to the cluster.\\
 
-Table \ref{tbl:oozie_templeton} shows how Oozie job submission is more complex then for the case of dynamically generated jobs than WebHCat/Templeton.
-Oozie is more fitting for static Hadoop jobs that require a workflow consisting of a sequence of Pig and MapReduce jobs mixed together, but is a lesser fit for our situation.
-Templeton will be used for submission of the Pig jobs, as the HDInsight version that was available at the start of the implementation did not yet support WebHCat.
+\subsection{Framework description}
+TODO
 
-\subsection{ListNet}
+\section{ListNet}
 The following sections describe the Pig jobs that form the three independent parts of the ListNet ranking model: 1) Preprocessing, 2) Training (this includes evaluation over the validation set) and 3) Testing (evaluation over the test set).
-\subsubsection{Preprocessing}
+\subsection{Preprocessing}
 The preprocessing phase consists of two separate Pig jobs.
@@
 \end{lstlisting}
-\end{minipage}
+\end{minipage}\\
 
@@
 \centering
-\begin{tabular}{p{5cm}p{8cm}}\toprule
+\begin{tabular}{p{6cm}p{7cm}}\toprule
 UDF & Description \\
@@
 
-\subsubsection{Training}
+\subsection{Training}
 The training stage, like the preprocessing stage, consists of two separate Pig jobs.
@@
 
-\subsubsection{Testing}
+\subsection{Testing}
 The testing stage tests the best model found in the training iterations, selected on validation set \ac{nDCG}@k (as calculated in the second Pig job of the training stage), by calculating the \ac{nDCG}@k of this model on the test set.\\
@@
 \end{lstlisting}
-\end{minipage}
+\end{minipage}\\
 
-\subsection{SmoothRank}
+\section{SmoothRank}
+\subsection{Preprocessing}
+\subsection{Training}
+\subsection{Testing}

### 2014-09-16 report/Chapters/Introduction.tex
uitbreiding implementation

@@
 Lin \cite{Lin2013} observed that algorithms that are of iterative nature, which most Learning-to-Rank algorithms are, are not amenable to the MapReduce framework.
-Lin argued that as a solution to the non-amenability of iterative algorithms to the MapReduce framework, iterative algorithms can often be replaced with non-iterative alternatives or can still be optimized in such a way that its performance in a MapReduce setting is good enough.\\
+Lin argued that as a solution to the non-amenability of iterative algorithms to the MapReduce framework, iterative algorithms can often be replaced with non-iterative alternatives or can still be optimized in such a way that its performance in a MapReduce setting is good enough.
+Alternative programming models are argued against by Lin, as they lack the critical mass as the data processing framework of choice.\\
 

### 2014-09-16 implementation/java/src/letor/serial/WrappedRunner.java

@@ -13,7 +13,7 @@ public class WrappedRunner {
 
     public static void main(String[] args){
         AbstractParameterizedRanker handler             = new SmoothRankHandler();
-        DataSets.DataSet            dataset             = DataSets.DataSet.CUSTOM;
+        DataSets.DataSet            dataset             = DataSets.DataSet.MINI;
         Integer                     duplicationNumber   = 1; // Only relevant in case of dataset Custom
 
         FoldRunHandler ltrWrapper = null;

### 2014-09-17 report/Chapters/Implementation.tex
eerste aanpassingen

@@
 Ranking algorithms consist of a sequence of operations on the input data which are often, but not always, of iterative nature.
-Apache Pig \cite{Olston2008} will be used to implement the sequence of operation on the input data.
-Pig Latin, the data processing language that runs on Apache Pig, designed based on the observation that the traditional MapReduce paradigm is too low-level and rigid, and holds the middle between the declarative style of SQL and the procedural style of MapReduce.
+Apache Pig \cite{Olston2008} will be used to implement the sequence of operations on the input data.
+Pig Latin, the data processing language that runs on Apache Pig, was designed based on the observation that the traditional MapReduce paradigm is too low-level and rigid, and holds the middle between the declarative style of SQL and the procedural style of MapReduce.
 The Apache Pig system translates Pig Latin into MapReduce plans that are executed over Hadoop.
-The choice to implement the Learning to Rank algorithms in Pig Latin allow more focus on the data operations and less focus on low-level implementation details compared to native Hadoop MapReduce.
-Furthermore, it allows us to rely on Apache Pig to create efficient MapReduce plans out of the Pig Latin code which lowers the implementation-dependent factor of the experiments.\\
+The choice to implement the Learning to Rank algorithms in Pig Latin allows for more focus on the data operations and less focus on low-level implementation details, as compared to native Hadoop MapReduce.
+Furthermore, it allows us to rely on Apache Pig to create efficient MapReduce plans out of the Pig Latin code and therefore lowers the implementation-dependent factor of the experiments.\\
 
-\subsection{Storing Input Data for HDInsight Processing}
+\subsection{The HDInsight Platform
 Azure HDInsight supports both the traditional \ac{HDFS} as described by Shvacko et al.
-\cite{Shvachko2010}, as well as Microsoft's own \ac{WASB} storage solution.
+\cite{Shvachko2010}, as well as Microsoft's own storage solution \ac{WASB}.
 Blob storage decouples the storage from the HDInsight Hadoop cluster; it enables safe deletion of a HDInsight cluster without losing data, as data is not solely stored on the cluster itself, but also on a separate storage that is not cluster-dependent.
-Azure \ac{WASB} storage allows the user to select one of their data centre for storage.
+Azure \ac{WASB} storage allows the user to select one of Microsoft's data centres for storage.
 \ac{WASB} storage in the West Europe region (located in Amsterdam) will be used for storage, as this data centre is located close to where the experiments are executed.
-
-\subsection{The HDInsight Platform}
-Microsoft offers a scalable and on-demand Hadoop service with HDInsight, that enables Hadoop services for those not able to make the required investments for their own cluster.
-The latest HDInsight of HDInsight available at the time of writing, HDInsight 3.1, runs the Hortonworks distribution of Hadoop version 2.4.
-While early versions of Hadoop were merely an open source implementation of MapReduce, Hadoop offers a wide variety of programming models since the introduction of Hadoop YARN in Hadoop 2.0 \cite{Vavilapalli2013}.
+}
+Microsoft offers a scalable and on-demand Hadoop service with HDInsight, which enables Hadoop services for those not able to make the required investments for their own cluster.
+The latest HDInsight version available at the time of writing, HDInsight 3.1, runs the Hortonworks distribution of Hadoop version 2.4.
+While early versions of Hadoop were merely an open source implementation of MapReduce, newer versions since Hadoop 2.0 offer support for variety of programming models with the introduction of Hadoop YARN \cite{Vavilapalli2013}.
 Newly supported programming models since Hadoop 2.0 include Dryad \cite{Isard2007}, Giraph \cite{Avery2011}, MPI, REEF \cite{Chun2013}, Spark \cite{Zaharia2010}, and Storm \cite{Aniello2013}.
-Even tough listed programming models are now supported by Hadoop and even though some of these programming models have recently increased in popularity, they still lack the critical mass as the data processing framework of choice as Lin argued back in 2012 \cite{Lin2013}.
+Even though these programming models are now supported by Hadoop and even though some of these programming models have recently increased in popularity, they still lack the critical mass as the data processing framework of choice as Lin argued back in 2012 \cite{Lin2013}.
 Therefore, even though some of these programming models might be a better fit for iterative algorithms, we use Hadoop MapReduce of the programming model to implement the Learning to Rank algorithms.\\
@@
 Templeton will be used for submission of the Pig jobs, as the HDInsight version that was available at the start of the implementation did not yet support WebHCat.\\
-
-\begin{table}
-\centering
-\begin{tabular}{p{5cm}p{5cm}}\toprule
-Job submission procedure for Apache Oozie & Job submission procedure for WebHCat/Templeton \\
-\midrule
-1. Let the framework build the Pig job dynamically. & 1. Let the framework build the Pig job dynamically.\\
-2. Encapsulate the Pig job in an Oozie workflow. & 2. Submit the Pig job through the WebHCat/Templeton REST API.\\
-3. Upload the Oozie workflow to HDFS storage. & \\
-4. Execute the Oozie workflow through the Oozie REST API. & \\
-\bottomrule
-\end{tabular}
-\caption{Comparison of Oozie and WebHCat job submission procedures}
-\label{tbl:oozie_templeton}
-\end{table}
 
@@
 \label{tbl:hdinsight_endpoints}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{p{5cm}p{5cm}}\toprule
+Job submission procedure for Apache Oozie & Job submission procedure for WebHCat/Templeton \\
+\midrule
+1. Let the framework build the Pig job dynamically. & 1. Let the framework build the Pig job dynamically.\\
+2. Encapsulate the Pig job in an Oozie workflow. & 2. Submit the Pig job through the WebHCat/Templeton REST API.\\
+3. Upload the Oozie workflow to HDFS storage. & \\
+4. Execute the Oozie workflow through the Oozie REST API. & \\
+\bottomrule
+\end{tabular}
+\caption{Comparison of Oozie and WebHCat job submission procedures}
+\label{tbl:oozie_templeton}
 \end{table}

### 2014-09-17 implementation/java/src/letor/serial/WrappedRunner.java

@@ -13,7 +13,7 @@ public class WrappedRunner {
 
     public static void main(String[] args){
         AbstractParameterizedRanker handler             = new SmoothRankHandler();
-        DataSets.DataSet            dataset             = DataSets.DataSet.MINI;
+        DataSets.DataSet            dataset             = DataSets.DataSet.MSLR_WEB10K;
         Integer                     duplicationNumber   = 1; // Only relevant in case of dataset Custom
 
         FoldRunHandler ltrWrapper = null;

### 2014-09-17 report/Chapters/Implementation.tex
nieuwe iteratie implementation

@@
 
-\subsection{The HDInsight Platform
+\subsection{The HDInsight Platform}
 Azure HDInsight supports both the traditional \ac{HDFS} as described by Shvacko et al.
@@
 Azure \ac{WASB} storage allows the user to select one of Microsoft's data centres for storage.
-\ac{WASB} storage in the West Europe region (located in Amsterdam) will be used for storage, as this data centre is located close to where the experiments are executed.
-}
+\ac{WASB} storage in the West Europe region (located in Amsterdam) will be used for storage, as this data centre is located close to where the experiments are executed.\\
+
 Microsoft offers a scalable and on-demand Hadoop service with HDInsight, which enables Hadoop services for those not able to make the required investments for their own cluster.
@@
 Newly supported programming models since Hadoop 2.0 include Dryad \cite{Isard2007}, Giraph \cite{Avery2011}, MPI, REEF \cite{Chun2013}, Spark \cite{Zaharia2010}, and Storm \cite{Aniello2013}.
-Even though these programming models are now supported by Hadoop and even though some of these programming models have recently increased in popularity, they still lack the critical mass as the data processing framework of choice as Lin argued back in 2012 \cite{Lin2013}.
+Even though these programming models are now supported by Hadoop and even though some of these programming models have recently increased in popularity, they still lack the critical mass as the data processing framework of choice that MapReduce has as Lin argued back in 2012 \cite{Lin2013}(see section \ref{sec:motivation}).
 Therefore, even though some of these programming models might be a better fit for iterative algorithms, we use Hadoop MapReduce of the programming model to implement the Learning to Rank algorithms.\\
 
-HDInsight 3.1 offers multiple ways of submitting Hadoop jobs to a HDInsight cluster, described in Table \ref{tbl:hdinsight_endpoints}.
-Oozie and WebHCat/Templeton are the two methods for job submission in Table \ref{tbl:hdinsight_endpoints} that both 1) support Apache Pig jobs, and 2) Can be used from within Java code.
-Table \ref{tbl:oozie_templeton} shows the necessary procedures for job submission from Oozie as well as from WebHCat will be sketched.
-Table \ref{tbl:oozie_templeton} shows Oozie job submission to be more complex for the case of dynamically generated jobs than WebHCat/Templeton.
-Oozie is more fitting for static Hadoop jobs that require a workflow consisting of a sequence of Pig and MapReduce jobs mixed together, but is a lesser fit for our situation.
-Templeton will be used for submission of the Pig jobs, as the HDInsight version that was available at the start of the implementation did not yet support WebHCat.\\
+HDInsight 3.1 offers multiple ways of submitting Hadoop jobs to a HDInsight cluster, which are described in Table \ref{tbl:hdinsight_endpoints}.
+Similar to \ac{WASB} storage, one of the Microsoft data centres can be chosen to host the Hadoop cluster when a HDInsight cluster is created.
+To guarantee proximity between storage and cluster, the West Europe data centre in Amsterdam will be used as cluster location.\\
 
@@
 
+\subsection{Framework description}
+Fold handling in the context of a cross-validation experiment is the process of loading the correct data folds for training, validation and testing in the multiple rounds that form the cross-validation experiment.
+Fold handling is a task that needs to be taken care of independent of the ranking model that is being evaluated.
+Given that most ranking models are of an iterative nature, the task of iteration handling is also a task that need to be performed independent of the ranking model.
+Iteration handling and fold handling are procedures in which the same steps are repeated for each iteration or cross-validation round respectively.
+Pig Latin, in contrast to more procedural MapReduce-job-generating languages like Sawzall \cite{Pike2005}, has no support for loops, which are needed to perform iteration handling and fold handling.
+Since iteration handling and fold handling are tasks that need to be addressed for each ranking model and that cannot be solved with Pig, we create a framework that takes care of both iteration and fold handling and generates the Pig Latin code for the current iteration and fold.\\
+
+Learning to Rank algorithms are likely to consist of multiple loops over the input data per iteration of the algorithm, and consequently multiple Pig jobs will be needed per iteration of the algorithm.
+An example of this is a Pig implementation of gradient descent, where a first Pig job might calculate gradients and writes them to storage, after which the framework assists in reading the gradients from storage and allows them to be used as input parameters of a second Pig job that calculates new feature weights based on its gradient parameter and a predetermined step size.
+Handling communication between Pig jobs within a single iteration of a Learning to Rank algorithm is not trivial.
+A Pig Job, after completion, writes its result to \ac{WASB} storage, while this result is needed by a subsequent Pig job as parameter.
+The framework enables reading the result of a Pig job from \ac{WASB} storage which can then be used as parameter within a subsequent Pig job.
+\\
+
+The aim of this framework is to let implementations of ranking models focus solely on implementing the sequential steps of one iteration of the algorithm, while the framework handles that 1) these sequential steps are performed iteratively, 2) these sequential steps are performed on the multiple training folds of data and 3) data can be passed from one Pig job within the algorithm to another Pig job.
+This framework that handles folding and iterations will be implemented in Java and will work such that fold- and iteration dependent parts of the Pig code will be generated dynamically by the Java framework after which the Pig job will be sent to the cluster.\\
+
+Oozie and WebHCat/Templeton are the two methods for job submission in Table \ref{tbl:hdinsight_endpoints} that both 1) support Apache Pig jobs, and 2) Can be used from within Java code.
+Table \ref{tbl:oozie_templeton} shows the necessary procedures for job submission from Oozie as well as from WebHCat will be sketched.
+Table \ref{tbl:oozie_templeton} shows Oozie job submission to be more complex for the case of dynamically generated jobs than WebHCat/Templeton.
+Oozie is more fitting for static Hadoop jobs that require a workflow consisting of a sequence of Pig and MapReduce jobs mixed together, but is a lesser fit for our situation.
+Templeton will be used for submission of the Pig jobs, as the HDInsight version that was available at the start of the implementation did not yet support WebHCat.\\
+
 \begin{table}
@@
 \end{table}
-
-\subsection{Handling repetitive tasks}
-Fold handling  in a cross-validation experiment is the process of loading the correct data folds for training, validation and testing in the multiple rounds that the cross-validation experiment consists of.
-Fold handling is a task that needs to be taken care of regardless of the ranking model that is being evaluated.
-Given that most ranking models are of an iterative nature, the task of iteration handling is also a task that need to be done for most models.
-Iteration handling and fold handling are procedures in which the same steps are repeated for each iteration or cross-validation round respectively.
-Pig Latin, in contrast to more procedural languages that generate MapReduce job like Sawzall \cite{Pike2005}, has no support for loops which are needed to perform iteration handling and fold handling.
-Since iteration handling and fold handling are tasks that need to be addressed for each ranking model and that cannot be solved with Pig, we create a framework that takes care of both iteration and fold handling and generates the Pig Latin code for the current iteration and fold.\\
-
-Handling communication between Pig jobs within a single iteration of a Learning to Rank algorithm is another problem that needs to be addressed.
-A Pig Job writes its result to \ac{WASB} storage, while this result is in most cases needed by a subsequent Pig job as parameter.
-The framework enables reading the result of a Pig job from \ac{WASB} storage which can then be used as parameter within a subsequent Pig job.
-An example of this is a Pig implementation of gradient descent, where a first Pig job might calculate gradients and writes them to storage, after which the framework assists in reading the gradients from storage and allows them to be used as input parameters of a second Pig job that calculates new feature weights based on its gradient parameter and a predetermined step size.\\
-
-The aim of this framework is to let implementations of ranking models focus solely on implementing the sequential steps of one iteration of the algorithm, while the framework handles that 1) these sequential steps are performed iteratively, 2) these sequential steps are performed on the multiple training folds of data and 3) data can be passed from one Pig job within the algorithm to another Pig job.
-This framework that handles folding and iterations will be implemented in Java and will work such that fold- and iteration dependent parts of the Pig code will be generated dynamically by the Java framework after which the Pig job will be sent to the cluster.\\
-
-\subsection{Framework description}
-TODO
 

### 2014-09-17 report/Chapters/Implementation.tex
wat refactoring + werkverdeling in thesis

@@
 While early versions of Hadoop were merely an open source implementation of MapReduce, newer versions since Hadoop 2.0 offer support for variety of programming models with the introduction of Hadoop YARN \cite{Vavilapalli2013}.
+HDInsight Hadoop data nodes are regular Large A3 Microsoft Azure virtual machines with four core processors and 7GB RAM.
 Newly supported programming models since Hadoop 2.0 include Dryad \cite{Isard2007}, Giraph \cite{Avery2011}, MPI, REEF \cite{Chun2013}, Spark \cite{Zaharia2010}, and Storm \cite{Aniello2013}.
@@
 \end{table}
+
+We create a utility class HDInsightWrapper, which abstracts away connection setup, authorisation and response parsing from the tasks of submitting jobs to the cluster and retrieving data from \ac{WASB} storage.
+The HDInsightWrapper class is parametrised with connection details like the cluster name, cluster user, cluster password, storage account, storage container and storage key, and has two methods:
+
+\begin{description}
+\item[void runPig(String pigCode)]{Receives a Pig script in String format as input and handles the submission of this script as MapReduce job on the cluster.
+Polls the Templeton REST API with progress requests every 50 milliseconds until the MapReduce job has completed.
+Blocks until the MapReduce job has completed.}
+\item[String runPig(String pigCode, String outputDir)]{Identical to runPig(String pigCode), but reads the result of the computation from location 'outputDir' after completion of the MapReduce job and outputs the result as String value.}
+\end{description}
+
+Setting the degree of parallelisation is crucial to obtain optimal cluster utilisation.
+HDInsight cluster machines offer four mapper slots and two reducer slots per data node.
+Cluster utilisation is optimal when the computational work is evenly distributed over exactly the number of available mappers in the map phase, and evenly distributed over exactly the number of reducers in the reduce phase.
+The number of mappers used can be controlled by setting MapReduce configuration parameters mapreduce.input.fileinputformat.split.maxsize, mapreduce.input.fileinputformat.split.minsize, and the Pig configuration parameter pig.maxCombinedSplitSize to the data size in bytes of the input training data divided by the number of available mappers.
+To control the number of used reducers, Pig offers two options: 1) the default\_parallel parameters, which can set a default number of reducers used throughout all MapReduce jobs that the Pig script consists of, and 2) the "PARALLEL" clause, which sets the number of reducers at operator level (overrides default\_parallel if set).
+The framework uses listed configuratoin parameters to control the numbers of mappers used and uses the "PARALLEL" clause to set the number of used reducers.
 

### 2014-09-17 implementation/java/src/letor/parallel/AzurePigWrapperTest.java

@@ -1,9 +1,9 @@
 package letor.parallel;
 
-import letor.parallel.util.AzurePigWrapper;
+import letor.parallel.util.HDInsightWrapper;
 
 /**
- * Test class for AzurePigWrapper
+ * Test class for HDInsightWrapper
  *
  * Created by niek.tax on 6/11/2014.
  */
@@ -21,13 +21,13 @@ public class AzurePigWrapperTest {
         String storageAccount       = "ltrstorage";
         String storageAccountKey    = "igtZD3Jih9lsvxoIcxCury1GDqS7Z4DQ0Ci7xVphY9p/6rnwaHG5qZFBKXjt0wOgwNwVqno5sitAy/eucuPGMA==";
 
-        AzurePigWrapper wrapper = new AzurePigWrapper(clusterName, clusterUser, clusterPassword, storageAccount, storageAccountKey, 1);
+        HDInsightWrapper wrapper = new HDInsightWrapper(clusterName, clusterUser, clusterPassword, storageAccount, storageAccountKey, 1);
 
         String pigLine = "TRAIN = LOAD '" + pathPrefix + "/input/" + DATASET + "/Fold" + fold + "/train.txt' USING PigStorage(' '); " +
                          "TR = FOREACH TRAIN GENERATE $1; "+
                          "STORE TR INTO '"+outDir+"';";
 
-        String result = wrapper.azureRunPig(pigLine, outDir);
+        String result = wrapper.runPig(pigLine, outDir);
 
         System.out.println("RESULT: "+result);
     }

### 2014-09-17 implementation/java/src/letor/parallel/ListNetCluster.java

@@ -1,6 +1,6 @@
 package letor.parallel;
 
-import letor.parallel.util.AzurePigWrapper;
+import letor.parallel.util.HDInsightWrapper;
 import letor.parallel.util.DataSets;
 import letor.parallel.util.LtrUtils;
 import letor.parallel.util.Metadata;
@@ -10,7 +10,7 @@ import java.util.List;
 
 /**
  * MapReduce (Hadoop) implementation of the ListNet algorithm
- * Makes use of AzurePigWrapper
+ * Makes use of HDInsightWrapper
  *
  */
 
@@ -27,7 +27,7 @@ public class ListNetCluster {
     private static int  availableMappers              = 4*dataNodes;
     private static int  availableReducers             = 2*dataNodes;
 
-    private static int duplicationNumber              = 49; // only used for CUSTOM
+    private static int  duplicationNumber             = 49; // only used for CUSTOM
     private static Metadata metadata                  = DataSets.getMetaData(DATASET, duplicationNumber);
     private static long MAX_TRAIN_SIZE                = metadata.getMax_train_size(); // ohsumed: 5151958, MQ2007: 25820919, MQ2008: 5927007, MSLR-WEB10K: 838011150, MSLR-WEB30K:
     private static long MAX_VALI_SIZE                 = metadata.getMax_vali_size(); // ohsumed: 1764005, MQ2007:  8753466, MQ2008: 2237346, MSLR-WEB10K: 280714022, MSLR-WEB30K:
@@ -46,7 +46,7 @@ public class ListNetCluster {
         String storageAccount       = "ltrstorage";
         String containerName        = "ltrcontainer";
         String storageAccountKey    = "rabkwaQc5z1PwYgVI27ri32OQwmkVXZV8ZnyavGL4+JgTsUo1YLmkm7YUBXUCzXQ1JJjfIMG7Y8UWWE1JgLK/A==";
-        AzurePigWrapper apw         = new AzurePigWrapper(clusterName, containerName, clusterUser, clusterPassword, storageAccount, storageAccountKey, availableReducers);
+        HDInsightWrapper hdw        = new HDInsightWrapper(clusterName, containerName, clusterUser, clusterPassword, storageAccount, storageAccountKey, availableReducers);
 
         ArrayList<String> pigLines  = new ArrayList<String>();
 
@@ -93,7 +93,7 @@ public class ListNetCluster {
             pigLines.add("STORE MIN_MAX_FIN INTO 'minmax"+f+"';");
[…]

### 2014-09-17 implementation/java/src/letor/parallel/util/HDInsightWrapper.java

@@ -25,7 +25,7 @@ import java.util.concurrent.*;
  *
  * @author Niek Tax
  */
-public class AzurePigWrapper {
+public class HDInsightWrapper {
     String clusterName;
     String clusterUser;
     String clusterPassword;
@@ -46,7 +46,7 @@ public class AzurePigWrapper {
 
     int availableReducers;
 
-    public AzurePigWrapper(String clusterName, String containerName, String clusterUser, String clusterPassword, String storageAccount, String storageAccountKey, int availableReducers){
+    public HDInsightWrapper(String clusterName, String containerName, String clusterUser, String clusterPassword, String storageAccount, String storageAccountKey, int availableReducers){
         this.clusterName        = clusterName;
         this.containerName      = containerName;
         this.clusterUser        = clusterUser;
@@ -67,11 +67,11 @@ public class AzurePigWrapper {
         this.retrieveStatusURL   = azureClusterURL+"/queue/";
     }
 
-    public AzurePigWrapper(String clusterName, String clusterUser, String clusterPassword, String storageAccount, String storageAccountKey, int availableReducers){
+    public HDInsightWrapper(String clusterName, String clusterUser, String clusterPassword, String storageAccount, String storageAccountKey, int availableReducers){
         this(clusterName, clusterName, clusterUser, clusterPassword, storageAccount, storageAccountKey, availableReducers);
     }
 
-    public void azureRunPig(String pigLine) throws Exception{
+    public void runPig(String pigLine) throws Exception{
         // STEP 1: Post pig job
         HttpResponse response = null;
         int          attempt  = 0;
@@ -156,23 +156,23 @@ public class AzurePigWrapper {
         }
     }
 
-    public String azureRunPig(String pigLine, String tmpDir) throws Exception{
+    public String runPig(String pigLine, String outputDir) throws Exception{
         //STEP 1 & 2: Post pig job and poll server until job finished
-        azureRunPig(pigLine);
+        runPig(pigLine);
 
         System.out.println("M/R job done, starting data retrieval");
 
         //STEP 3: Retrieve answer
-        return retrieveData(tmpDir);
+        return retrieveData(outputDir);
     }
 
-    private String retrieveData(String tmpDir) throws Exception{
+    private String retrieveData(String fileDir) throws Exception{
         boolean succeeded = false;
         int     attempts = 0;
         String  answer = null;
         while(!succeeded && attempts < 3) {
             try {
-                answer = retrieveDataAttempt(tmpDir);
[…]

### 2014-09-17 report/Chapters/Implementation.tex
impl

@@
 To control the number of used reducers, Pig offers two options: 1) the default\_parallel parameters, which can set a default number of reducers used throughout all MapReduce jobs that the Pig script consists of, and 2) the "PARALLEL" clause, which sets the number of reducers at operator level (overrides default\_parallel if set).
-The framework uses listed configuratoin parameters to control the numbers of mappers used and uses the "PARALLEL" clause to set the number of used reducers.
+The framework uses listed configuration parameters to control the numbers of mappers used and uses the "PARALLEL" clause to set the number of used reducers.
+Pig code to set the number of mappers and reducers are dynamically set by the framework such that the optimal number of mappers and reducers are used.
 

### 2014-09-17 report/Chapters/Implementation.tex
ranking accuracy measurements for validation

@@
 \section{ListNet}
-The following sections describe the Pig jobs that form the three independent parts of the ListNet ranking model: 1) Preprocessing, 2) Training (this includes evaluation over the validation set) and 3) Testing (evaluation over the test set).
+The following sections describe the Pig jobs that jointly form the three independent parts of the ListNet ranking model: 1) Preprocessing, 2) Training (this includes evaluation over the validation set) and 3) Testing (evaluation over the test set).
 \subsection{Preprocessing}
@@
 x^{'} = \frac{x-min(x)}{max(x)-min(x)}
+\label{eq:rescaling}
 \end{equation}
@@
 \end{minipage}\\
-
-The minimum and maximum values per feature stored in MIN\_MAX\_FIN can now be read from \ac{HDFS} and used within the Java framework by using the Windows Azure Storage library.
-The second Pig job uses these minimum and maximum values by parametrising the \ac{UDF} that performs the rescaling transformation, udf.util.ScaleFeatures(), with the minimum and maximum feature values.\\
+All expressions between square brackets in the code snippet above are dynamically set by the framework.
+The minimum and maximum values per feature stored in MIN\_MAX\_FIN are read from storage by the framework after completion of the MapReduce job.
+The framework passes these minimum and maximum values to the second Pig job by passing an array of minimum and maximum values to the constructor of the udf.util.ScaleFeatures() \ac{UDF}, that performs the rescaling operation of Equation \ref{eq:rescaling}s.\\
 
@@
 \end{tabular}
-\caption{Description of preprocessing User Defined Functions (1/2)}
+\caption{Description of preprocessing phase User Defined Functions (Pig job 1)}
 \label{tbl:preprocessing_udfs_1}
@@
 \end{tabular}
-\caption{Description of preprocessing User Defined Functions (2/2)}
+\caption{Description of preprocessing phase User Defined Functions (Pig job 2)}
 \label{tbl:preprocessing_udfs_2}
@@
 \subsection{Training}
+The Pig jobs that form the ListNet training phase form an implementation of the theoretical description of ListNet as described in section \ref{sec:ltrmethods_listnet}.
 The training stage, like the preprocessing stage, consists of two separate Pig jobs.
@@
 
+The actual neural network weights, passed by the framework as constructor parameter in the script above, are administered in Java code.
+The per-feature gradients calculated by the Pig job above are read from storage after completion and used to update the neural network by, for each feature, adding the feature gradient multiplied with a step size parameter to the previous feature weight.
+
 \begin{table}
@@
 udf.listnet.QueryLossGradient() & Calculates the Cross Entropy loss for a query and calculates the gradients per feature based on this query.\\
-udf.listnet.ExpRelOurScores() & Calculates the predicted relevance label of a query based on the current model weights and transforms this following the transformation $x -> e^{x}$.
+udf.listnet.ExpRelOurScores() & Calculates the predicted relevance label of a query based on the current model weights and transforms this following the transformation $x \rightarrow e^{x}$.
 In case the current iteration is the first iteration, the same transformation is applied to the ground truth relevance label.\\
@@
 \end{tabular}
-\caption{Description of training User Defined Functions (1/2)}
+\caption{Description of training phase User Defined Functions (Pig job 1)}
 \label{tbl:training_udfs_1}
@@
 \end{tabular}
-\caption{Description of training User Defined Functions (2/2)}
+\caption{Description of training phase User Defined Functions (Pig job 2)}
 \label{tbl:training_udfs_2}

### 2014-09-17 report/Chapters/LTRMethods.tex
ranking accuracy measurements for validation

@@
 \section{ListNet}
+\label{sec:ltrmethods_listnet}
 ListNet \cite{Cao2007} is a listwise ranking function whose loss function is not directly related to information retrieval evaluation metrics.

### 2014-09-17 report/Chapters/Results.tex
ranking accuracy measurements for validation

@@
 This chapter describes and discusses the results of the running time measurements of the cluster and baseline implementations as described in \ref{chap:implementation}.
+In addition, ranking accuracy results are included to validate the correctness of the Learning to Rank algorithm implementations.
 \section{ListNet}
+\subsection{Ranking accuracy}
+\begin{table}[!h]
+\centering
+\begin{tabular}{p{6cm}p{1.6cm}p{1.4cm}r}\toprule
+Dataset & RankLib & Cluster & Official evaluation \\
+\midrule
+OHSUMED (LETOR 3.0 version)     & 0.29          &   TODO & 0.3793 \\
+MQ2008      					& 0.3865        &   TODO & 0.4689 \\
+MQ2007      					& 0.2873        &   TODO & 0.4767 \\
+MSLR-WEB10K 					& 0.1744     	&   TODO & - \\
+MSLR-WEB30K 					& 0.1837     	&   TODO & - \\
+\bottomrule
+\end{tabular}
+\caption{\acs{NDCG}@10 performance on the testing set on the first fold}
+\label{tbl:accuracy_comparison}
+\end{table}
+
+\subsection{Speedup}
 Figure \ref{fig:listnet_train_time} and Figure \ref{fig:listnet_train_time_log} (which presents the same data with a logarithmic data size axis) show the processing times of a single training iteration using the ListNet implementation included in RankLib library and the cluster implementation described in chapter \ref{chap:implementation} with different numbers of data nodes.
@@
 As previously described in the methodology section, the LETOR 3.0, LETOR 4.0 and MSLR-WEB10/30K datasets are used, supplemented with generated datasets that are multiplications of MSLR-WEB30K.
-Table \ref{tbl:recap_datasets} describes the datasets used in the experiments.
+Table \ref{tbl:recap_datasets} describes the datasets used in the experiments.\\
 

### 2014-09-17 implementation/java/src/letor/serial/WrappedRunner.java

@@ -12,20 +12,22 @@ import letor.serial.util.Measurement;
 public class WrappedRunner {
 
     public static void main(String[] args){
-        AbstractParameterizedRanker handler             = new SmoothRankHandler();
-        DataSets.DataSet            dataset             = DataSets.DataSet.MSLR_WEB10K;
+        Integer                     folds               = 1;
+        Integer                     iterations          = 5;
+        AbstractParameterizedRanker handler             = new ListNetHandler();
+        DataSets.DataSet            dataset             = DataSets.DataSet.MSLR_WEB30K;
         Integer                     duplicationNumber   = 1; // Only relevant in case of dataset Custom
 
         FoldRunHandler ltrWrapper = null;
         if(dataset == DataSets.DataSet.CUSTOM)
-            ltrWrapper = new FoldRunHandler(handler, DataSets.DataSet.CUSTOM, duplicationNumber,  1, 1); // Method, dataset, folds, iterations
+            ltrWrapper = new FoldRunHandler(handler, DataSets.DataSet.CUSTOM, duplicationNumber,  folds, iterations);
         else
-            ltrWrapper = new FoldRunHandler(handler, dataset, 1, 1); // Method, dataset, folds, iterations
+            ltrWrapper = new FoldRunHandler(handler, dataset, folds, iterations);
         Measurement measurement = ltrWrapper.averageScore();
-        System.out.println("Preprocessing: "+measurement.getPreprocessingTime());
-        System.out.println("Train:         "+measurement.getTrainingTime());
-        System.out.println("Test:          "+measurement.getTestTime());
-        System.out.println("Total time:    "+measurement.getTotalTime());
-        System.out.println("NDCG@10:       "+measurement.getEvaluationResult()+" ms");
+        System.out.println("Preprocessing: "+measurement.getPreprocessingTime()+" ms");
+        System.out.println("Train:         "+measurement.getTrainingTime()+" ms");
+        System.out.println("Test:          "+measurement.getTestTime()+" ms");
+        System.out.println("Total time:    "+measurement.getTotalTime()+" ms");
+        System.out.println("NDCG@10:       "+measurement.getEvaluationResult());
     }
 }

### 2014-09-18 report/Chapters/Introduction.tex
nieuwe iteratie implementation
2e iteratie resutls

@@
 \section{Motivation and Problem Statement}
+\label{sec:motivation}
 Ranking is a core problem in the field of information retrieval.
@@
 \subsection{Methodology for Research Question II}
+\label{ssec:rq2_methodology}
 To find an answer to the second research question, the Learning-to-Rank methods determined in the first research question will be implemented in the MapReduce framework and training time will be measured as a factor of the number of cluster nodes used to perform the computation.

### 2014-09-18 report/Chapters/Results.tex
2e iteratie resutls

@@
 \end{tabular}
-\caption{\acs{NDCG}@10 performance on the testing set on the first fold}
+\caption{\acs{nDCG}@10 performance on the testing set on the first fold}
 \label{tbl:accuracy_comparison}
@@
 \subsection{Speedup}
-Figure \ref{fig:listnet_train_time} and Figure \ref{fig:listnet_train_time_log} (which presents the same data with a logarithmic data size axis) show the processing times of a single training iteration using the ListNet implementation included in RankLib library and the cluster implementation described in chapter \ref{chap:implementation} with different numbers of data nodes.
-The horizontal positions of the measurements are identical between execution method, as they are determined by the collection of datasets used.
-As previously described in the methodology section, the LETOR 3.0, LETOR 4.0 and MSLR-WEB10/30K datasets are used, supplemented with generated datasets that are multiplications of MSLR-WEB30K.
-Table \ref{tbl:recap_datasets} describes the datasets used in the experiments.\\
+Figure \ref{fig:listnet_train_time} (linear data size axis) and Figure \ref{fig:listnet_train_time_log} (logarithmic data size axis) show the processing times of a single iteration of the ListNet training algorithm, using the ListNet implementation included in the RankLib library as well as the cluster implementation described in chapter \ref{chap:implementation} with different numbers of data nodes.
+The horizontal positions of the measurements are identical between execution method, as they are originate from the data sizes of the datasets used.
+Measurements were performed on a set of datasets consisting of the LETOR 3.0 OHSUMED dataset, LETOR 4.0 datasets and the MSLR-WEB10/30K datasets are used, supplemented with generated datasets that are multiplications of MSLR-WEB30K (as described in section \ref{ssec:rq2_methodology}).
+Table \ref{tbl:recap_datasets} describes the datasets used in the experiments and their single training fold data sizes.\\
 

### 2014-09-18 report/Chapters/Results.tex
eerste verbeteriteratie

@@
 \begin{tabular}{p{3.4cm}p{3.4cm}r}\toprule
-Dataset & Collection & Size \\
+Dataset & Collection & Single fold training size \\
 \midrule
@@
 Figures \ref{fig:listnet_train_time} and \ref{fig:listnet_train_time_log} show that the single-machine RankLib implementation of ListNet is very fast compared to the Hadoop MapReduce cluster implementation of ListNet for datasets that fit in memory.
-As soon as the amount of data processed with RankLib ListNet approaches the physical memory limit of the machine used for the computation (16 GB for the machine used in our experiments), RankLib ListNet processing time start to increase very rapidly.
-This increase is likely to be a result of swapping data to and from virtual memory.
-The largest dataset processed with RankLib ListNet is CUSTOM-8, which is 20.99 GB in size and therewith larger than the 16 GB physical memory limit.
-Experiments with RankLib ListNet on CUSTOM-10 were attempted, but were stopped after not completing within 12 hours.\\
+As soon as the amount of data processed with RankLib ListNet approaches the physical memory limit of the machine that is used for computation (16 GB for our single-machine measurements), RankLib ListNet processing time start to increase exponentially.
+This increase is likely to be a result of the machine needing to use the on-disk sections of virtual memory and the swapping that takes place as a result thereof.
+As an exception to the datasets described in Table \ref{tbl:recap_datasets}, the largest dataset processed with RankLib ListNet is CUSTOM-8, which is 20.99 GB in size and thereby larger than the 16 GB physical memory limit.
+Experiments with RankLib ListNet on CUSTOM-10 were attempted, but were stopped after not completing within 12 hours.
+In section \ref{sec:goals} we defined speed-up as Sun and Gustafson \emph{relative speed-up} metric \cite{Sun1991} (as stated in section \ref{sec:goals}), which defines speed-up as the number of times that the execution of the fastest single machine solution is lower than the execution time with $N$ machines.
+Note that speed-up, following this definition, is not computable for larger datasets, as it is not computationally feasible to perform computation on datasets larger than CUSTOM-8 with a single machine.
+As an alternative, Figure \ref{fig:speedup_train_time} shows visualises speed-up by plotting cluster size against training time.\\
 
@@
 
-Figure \ref{fig:speedup_train_time} shows that speed-up is sub-linear such that the training time converges to a constant unit of time somewhere within the range of 150-200 seconds.
+Figure \ref{fig:speedup_train_time} shows that speed-up is sub-linear, from which we can deduct that the training time converges to a constant unit of time.
+Based on our measurements on the small datasets MINI, OHSUMED, MQ2008 and MQ2007, this constant time seems to be within the range of 150 to 200 seconds.
 This time is likely to be caused by Hadoop job scheduling overhead, this presumption is strengthened by long waiting periods between the separate MapReduce jobs that form a training iteration.\\
@@
 Amdahl's law states that the speed-up of a program using parallel computing is limited by the time needed for the sequential fraction of the program.
-A consequence of Amdahl's Law is that all parallel programs the have a non-parallelisable part have sub-linear speed-up.
+A consequence of Amdahl's Law is that all parallel programs that have a non-parallelisable part have sub-linear speed-up.
 Behaviour in accordance with Amdahl's law can be seen in Figure \ref{fig:speedup_train_time}, where the speed-up is sub-linear as a result of the existence a non-parallelisable fraction of the program.\\

### 2014-09-18 report/Chapters/Results.tex
foutje gefixt

@@
 Experiments with RankLib ListNet on CUSTOM-10 were attempted, but were stopped after not completing within 12 hours.
-In section \ref{sec:goals} we defined speed-up as Sun and Gustafson \emph{relative speed-up} metric \cite{Sun1991} (as stated in section \ref{sec:goals}), which defines speed-up as the number of times that the execution of the fastest single machine solution is lower than the execution time with $N$ machines.
+In section \ref{sec:goals} we defined speed-up as Sun and Gustafson \emph{relative speed-up} metric \cite{Sun1991}, which defines speed-up as the number of times that the execution of the fastest single machine solution is lower than the execution time with $N$ machines.
 Note that speed-up, following this definition, is not computable for larger datasets, as it is not computationally feasible to perform computation on datasets larger than CUSTOM-8 with a single machine.

### 2014-09-19 report/Chapters/Results.tex
NDCG measurements + correcties implementatie

@@
 \midrule
-OHSUMED (LETOR 3.0 version)     & 0.29          &   TODO & 0.3793 \\
-MQ2008      					& 0.3865        &   TODO & 0.4689 \\
-MQ2007      					& 0.2873        &   TODO & 0.4767 \\
-MSLR-WEB10K 					& 0.1744     	&   TODO & - \\
-MSLR-WEB30K 					& 0.1837     	&   TODO & - \\
+OHSUMED (LETOR 3.0 version)     & 0.1144        &   0.3952 & 0.3793 \\
+MQ2008      					& 0.3539        &   0.4969 & 0.4689 \\
+MQ2007      					& 0.2581        &   0.5385 & 0.4767 \\
+MSLR-WEB10K 					& 0.1744     	&   0.4023 & - \\
+MSLR-WEB30K 					& 0.1837     	&   0.4020 & - \\
 \bottomrule

### 2014-09-19 implementation/java/src/letor/parallel/ListNetCluster.java

@@ -7,6 +7,7 @@ import letor.parallel.util.Metadata;
 
 import java.util.ArrayList;
 import java.util.List;
+import java.util.Random;
 
 /**
  * MapReduce (Hadoop) implementation of the ListNet algorithm
@@ -16,14 +17,14 @@ import java.util.List;
 
 public class ListNetCluster {
     // Initialise hyper-parameters
-    private static final DataSets.DataSet DATASET = DataSets.DataSet.MSLR_WEB10K;
-    private static final double   STEPSIZE   = 0.00001; // MSLR-WEB10K: 0.0001, ohsumed: 0.01
-    private static final int      ITERATIONS = 2;
+    private static final DataSets.DataSet DATASET = DataSets.DataSet.OHSUMED;
+    private static final double   STEPSIZE   = 0.0001; // MSLR-WEB10K: 0.0001, ohsumed: 0.01
+    private static final int      ITERATIONS = 5;
     private static final int      FOLDS      = 1;
     private static final int      k          = 10; // NDCG@k
 
     // Initialise paralellisation parameters
-    private static int  dataNodes                     = 2;
+    private static int  dataNodes                     = 1;
     private static int  availableMappers              = 4*dataNodes;
     private static int  availableReducers             = 2*dataNodes;
 
@@ -40,12 +41,12 @@ public class ListNetCluster {
 
     public static void main(String[] args) throws Exception {
         // Cluster configuration
-        String clusterName          = "ltr2";
+        String clusterName          = "ltr1";
         String clusterUser          = "admin";
         String clusterPassword      = "Qw!23456789";
-        String storageAccount       = "ltrstorage";
-        String containerName        = "ltrcontainer";
-        String storageAccountKey    = "rabkwaQc5z1PwYgVI27ri32OQwmkVXZV8ZnyavGL4+JgTsUo1YLmkm7YUBXUCzXQ1JJjfIMG7Y8UWWE1JgLK/A==";
+        String storageAccount       = "ltrorlando";
+        String containerName        = "ltr1";
+        String storageAccountKey    = "/Qv1ci7kewXJG6OMCZ9L+z+4S0PaRfN0g6LjLgAQmqiqIWje3DiXJwL04L64hXvzFb0gajuIgzVzkWyjkzYobw==";
[…]

### 2014-09-19 implementation/java/src/letor/serial/WrappedListNetRunner.java

@@ -1,17 +1,0 @@
-package letor.serial;
-
-import letor.serial.util.Measurement;
-
-/**
- * Created by niek.tax on 5/28/2014.
- */
-public class WrappedListNetRunner {
-    public static void main(String[] args){
-        /*
-        FoldRunHandler listNetWrapper = new FoldRunHandler();
-        Measurement measurement = listNetWrapper.averageScore("MSLR-WEB30K");
-        System.out.println("Time elapsed: "+measurement.getRunningTime());
-        System.out.println("NDCG@10:      "+measurement.getEvaluationResult()+" ms");
-        */
-    }
-}

### 2014-09-19 implementation/java/src/letor/serial/WrappedRunner.java

@@ -15,7 +15,7 @@ public class WrappedRunner {
         Integer                     folds               = 1;
         Integer                     iterations          = 5;
         AbstractParameterizedRanker handler             = new ListNetHandler();
-        DataSets.DataSet            dataset             = DataSets.DataSet.MSLR_WEB30K;
+        DataSets.DataSet            dataset             = DataSets.DataSet.OHSUMED;
         Integer                     duplicationNumber   = 1; // Only relevant in case of dataset Custom
 
         FoldRunHandler ltrWrapper = null;

### 2014-09-19 implementation/java/src/udf/listnet/QueryLossGradient.java

@@ -38,22 +38,35 @@ public class QueryLossGradient extends EvalFunc<Tuple> {
         // Create output Tuple
         Tuple value = TupleFactory.getInstance().newTuple();
 
+        double sumRel = 0.0;
+        double sumOur = 0.0;
         Iterator<Tuple> docIterator = documentBag.iterator();
         while(docIterator.hasNext()) {
             Tuple docTuple = docIterator.next();
-            double rel = Double.parseDouble(docTuple.get(docTuple.size()-2).toString()); // P_y(j)
-            double our = Double.parseDouble(docTuple.get(docTuple.size()-1).toString()); // P_z(j)
-            for(int i=0; i<DIM; i++){
-                // gradientForQuery += (q.docFeatures(j) * (P_z(j) - P_y(j)));
-                gradientForQuery[i] += Double.parseDouble(docTuple.get(i+2).toString()) * (our - rel); // q.docFeatures(j). +2 to skip relevance label and qid
-            }
+            sumRel += Double.parseDouble(docTuple.get(docTuple.size() - 2).toString()); // P_y(j)
+            sumOur += Double.parseDouble(docTuple.get(docTuple.size() - 1).toString()); // P_z(j)
+        }
 
-            // lossForQuery += -P_y(j) * math.log(P_z(j)));
-            lossForQuery += -rel * SimpleMath.logBase2(our);
+        if(sumRel>0) {
+            docIterator = documentBag.iterator();
+            while (docIterator.hasNext()) {
+                Tuple docTuple = docIterator.next();
+                double rel = Double.parseDouble(docTuple.get(docTuple.size() - 2).toString()) / sumRel; // P_y(j)
+                double our = sumOur >0 ? Double.parseDouble(docTuple.get(docTuple.size() - 1).toString()) / sumOur : 0; // P_z(j)
+                for (int i = 0; i < DIM; i++) {
+                    // gradientForQuery += (q.docFeatures(j) * (P_z(j) - P_y(j)));
+                    gradientForQuery[i] += Double.parseDouble(docTuple.get(i + 2).toString()) * (our - rel); // q.docFeatures(j). +2 to skip relevance label and qid
+                }
+
+                // lossForQuery += -P_y(j) * math.log(P_z(j)));
+                if(our == 0)
+                    our = 0.00001;
+                lossForQuery += -rel * SimpleMath.logBase2(our);
+            }
         }
 
         // Fill output tuple
-        value.append(lossForQuery);
[…]

### 2014-09-19 implementation/java/src/udf/util/Ndcg.java

@@ -33,10 +33,12 @@ public class Ndcg extends EvalFunc<Double>{
     }
 
     public Double exec(Tuple input) throws IOException {
-        Integer[] top_k_rels     = new Integer[k];
-        Double[] top_k_preds = new Double[k];
+        Integer[]       top_k_rels    = new Integer[k];
+        PredRelPair[]   top_k_preds   = new PredRelPair[k];
+
         Arrays.fill(top_k_rels, 0);
-        Arrays.fill(top_k_preds, 0.0);
+        for(int i=0; i<k; i++)
+            top_k_preds[i] = new PredRelPair(0.0,0);
 
         Iterator docBagIterator = ((DataBag) input.get(1)).iterator();
         while(docBagIterator.hasNext()){
@@ -73,18 +75,23 @@ public class Ndcg extends EvalFunc<Double>{
             min_i  = 0;
             for(int i=0;i<k;i++){
                 if(i>0)
-                    min_i = top_k_preds[min_i] > top_k_preds[i] ? i : min_i;
-                if(!higher && pred > top_k_preds[i])
+                    min_i = top_k_preds[min_i].pred > top_k_preds[i].pred ? i : min_i;
+                if(!higher && pred > top_k_preds[i].pred)
                     higher = true;
             }
-            if(higher)
-                top_k_preds[min_i] = pred;
+            if(higher) {
+                top_k_preds[min_i] = new PredRelPair(pred, rel);
+            }
         }
         // Calculate ideal NDCG@k
 
         // docList is now ordered by relevance
         double idealDCG = getDCG(top_k_rels);
-        double DCG      = getDCG(top_k_preds);
+
+        Integer[] predrels = new Integer[k];
+        for(int i=0; i<k; i++)
+            predrels[i] = top_k_preds[i].rel;
+        double DCG      = getDCG(predrels);
         return (idealDCG>0) ? DCG/idealDCG : 0.0;
     }
 
@@ -100,15 +107,12 @@ public class Ndcg extends EvalFunc<Double>{
         return dcg;
     }
 
-    private static double getDCG(Double[] l){
-        Arrays.sort(l, Collections.reverseOrder());
-        double dcg = 0.0;
-
-        int i = 1;
-        for(double e:l) {
-            dcg += (Math.pow(2.0, e) - 1.0) / SimpleMath.logBase2(i + 1);
-            i++;
+    private class PredRelPair {
+        public final Double pred;
+        public final Integer rel;
[…]

### 2014-09-19 implementation/java/src/letor/parallel/ListNetCluster.java
ndcg validation measurements

@@ -17,14 +17,14 @@ import java.util.Random;
 
 public class ListNetCluster {
     // Initialise hyper-parameters
-    private static final DataSets.DataSet DATASET = DataSets.DataSet.OHSUMED;
+    private static final DataSets.DataSet DATASET = DataSets.DataSet.MQ2007;
     private static final double   STEPSIZE   = 0.0001; // MSLR-WEB10K: 0.0001, ohsumed: 0.01
     private static final int      ITERATIONS = 5;
     private static final int      FOLDS      = 1;
     private static final int      k          = 10; // NDCG@k
 
     // Initialise paralellisation parameters
-    private static int  dataNodes                     = 1;
+    private static int  dataNodes                     = 4;
     private static int  availableMappers              = 4*dataNodes;
     private static int  availableReducers             = 2*dataNodes;
 
@@ -41,7 +41,7 @@ public class ListNetCluster {
 
     public static void main(String[] args) throws Exception {
         // Cluster configuration
-        String clusterName          = "ltr1";
+        String clusterName          = "ltr4";
         String clusterUser          = "admin";
         String clusterPassword      = "Qw!23456789";
         String storageAccount       = "ltrorlando";

### 2014-09-19 implementation/java/src/letor/serial/WrappedRunner.java

@@ -15,7 +15,7 @@ public class WrappedRunner {
         Integer                     folds               = 1;
         Integer                     iterations          = 5;
         AbstractParameterizedRanker handler             = new ListNetHandler();
-        DataSets.DataSet            dataset             = DataSets.DataSet.OHSUMED;
+        DataSets.DataSet            dataset             = DataSets.DataSet.MSLR_WEB30K;
         Integer                     duplicationNumber   = 1; // Only relevant in case of dataset Custom
 
         FoldRunHandler ltrWrapper = null;

### 2014-09-22 implementation/java/src/letor/parallel/ListNetCluster.java
nieuwe metingen in thesis

@@ -17,7 +17,7 @@ import java.util.Random;
 
 public class ListNetCluster {
     // Initialise hyper-parameters
-    private static final DataSets.DataSet DATASET = DataSets.DataSet.MQ2007;
+    private static final DataSets.DataSet DATASET = DataSets.DataSet.MSLR_WEB30K;
     private static final double   STEPSIZE   = 0.0001; // MSLR-WEB10K: 0.0001, ohsumed: 0.01
     private static final int      ITERATIONS = 5;
     private static final int      FOLDS      = 1;

### 2014-09-22 report/Chapters/Results.tex
nieuwe metingen in thesis
ndcg validation stukje geschreven

--- 2014-09-18 report/Chapters/Results.tex
@@
 \section{ListNet}
+
 \subsection{Ranking accuracy}
+To validate out ListNet Cluster implementation we run experiments on the LETOR 3.0 OHSUMED and the MQ2007 and MQ2008 datasets from LETOR 4.0 and compare their performance in terms of \ac{nDCG}@10.
+Hadoop job scheduling overhead is large when processing relatively small datasets (as we will explain in more detail in section \ref{ssec:speedup}), therefore \ac{nDCG}@10 experiments are limited to five training iterations and are run only on one of the five folding combinations.
+Because of the stochastic nature of the ListNet ranking procedure, which is a result of random weight initialisation prior to the first training iteration, the \ac{nDCG}@10 performance is not guaranteed to be identical between multiple runs on the same input data.
+To account for the non-determinism of ListNet, we repeat each experiment five times and report the average \ac{nDCG}@10 as well as the standard deviation in \ac{nDCG}@10.0
 \begin{table}[!h]
 \centering
-\begin{tabular}{p{6cm}p{1.6cm}p{1.4cm}r}\toprule
+\begin{tabular}{p{3.5cm}p{2.6cm}p{2.6cm}r}\toprule
 Dataset & RankLib & Cluster & Official evaluation \\
 \midrule
-OHSUMED (LETOR 3.0 version)     & 0.29          &   TODO & 0.3793 \\
-MQ2008      					& 0.3865        &   TODO & 0.4689 \\
-MQ2007      					& 0.2873        &   TODO & 0.4767 \\
-MSLR-WEB10K 					& 0.1744     	&   TODO & - \\
-MSLR-WEB30K 					& 0.1837     	&   TODO & - \\
+OHSUMED (LETOR 3.0 version) & 0.1916 $\pm$ 0.0634 & 0.3395 $\pm$ 0.0952 & 0.3793 \\
+MQ2008      				& 0.3889 $\pm$ 0.0600 & 0.4280 $\pm$ 0.0998 & 0.4689 \\
+MQ2007      				& 0.2918 $\pm$ 0.0546 & 0.4554 $\pm$ 0.0085 & 0.4767 \\
 \bottomrule
@@
 \subsection{Speedup}
+\label{ssec:speedup}
 Figure \ref{fig:listnet_train_time} (linear data size axis) and Figure \ref{fig:listnet_train_time_log} (logarithmic data size axis) show the processing times of a single iteration of the ListNet training algorithm, using the ListNet implementation included in the RankLib library as well as the cluster implementation described in chapter \ref{chap:implementation} with different numbers of data nodes.

### 2014-09-22 report/Chapters/BenchmarkResults.tex
stukje conclusie

@@
 The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 \cite{Liu2007b} to solve the absence of a experimental platform for Learning-to-Rank at that time.
-LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 in December 2008 and LETOR 4.0 in July 2009.
+LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 \cite{Qin2010} in December 2008 and LETOR 4.0 \cite{Qin2013} in July 2009.
 

### 2014-09-22 report/Chapters/Conclusions.tex
stukje conclusie

Running ListNet on a Hadoop cluster using the MapReduce computing model comes with its own cost in the form of a job scheduling overhead in the range of 150-200 seconds per training iteration. However, with a normalisation preprocessing iteration we can greatly reduce the number of training iterations needed for converge. ListNet on a single machine does not scale well to data sizes larger than the physical memory size. To process data sets in the 100+GB range with the ListNet training algorithm, Hadoop MapReduce is a large improvement compared to a single machine.\\

No generalisations can be drawn from this results to other Learning to Rank algorithms. Although we can extend our findings on job scheduling overhead and scaling benefit from data larger than 100+GB to the gradient descent procedure that is used in ListNet as well as in many other Learning to Rank algorithms and in many learning algorithms in general. Other Learning to Rank algorithms using the gradient descent procedure might not scale well when the MapReduce computing model is used, but any bad scaling behaviour on MapReduce of Learning to Rank algorithms will not be because of gradient descent.

% TODO: how does this generalise to other Learning to Rank algorithms %

### 2014-09-22 report/Chapters/Future.tex
stukje conclusie

\subsection{Computing models}
We explored the possibilities of Hadoop MapReduce for computation

\subsection{Learning to Rank algorithms}

### 2014-09-22 report/Chapters/Results.tex
stukje conclusie

@@
 \subsection{Ranking accuracy}
-To validate out ListNet Cluster implementation we run experiments on the LETOR 3.0 OHSUMED and the MQ2007 and MQ2008 datasets from LETOR 4.0 and compare their performance in terms of \ac{nDCG}@10.
-Hadoop job scheduling overhead is large when processing relatively small datasets (as we will explain in more detail in section \ref{ssec:speedup}), therefore \ac{nDCG}@10 experiments are limited to five training iterations and are run only on one of the five folding combinations.
+To validate our ListNet cluster implementation we run experiments on the LETOR 3.0 OHSUMED and the LETOR 4.0 datasets MQ2007 and MQ2008 and compare their performance with the RankLib implementation of ListNet and with the official evaluation results in terms of \ac{nDCG}@10.
+Hadoop job scheduling overhead is large when processing relatively small datasets (as we will explain in more detail in section \ref{ssec:speedup}), therefore \ac{nDCG}@10 experiments are limited to five training iterations and are run only on one of the five folds.
 Because of the stochastic nature of the ListNet ranking procedure, which is a result of random weight initialisation prior to the first training iteration, the \ac{nDCG}@10 performance is not guaranteed to be identical between multiple runs on the same input data.
-To account for the non-determinism of ListNet, we repeat each experiment five times and report the average \ac{nDCG}@10 as well as the standard deviation in \ac{nDCG}@10.0
-\begin{table}[!h]
+To account for this non-determinism of ListNet, we repeat each experiment five times and report the average as well as the standard deviation of the resulting rankings on the testing fold in terms of \ac{nDCG}@10.\\
+
+Table \ref{tbl:accuracy_comparison} shows the mean and the standard deviation of the ranking accuracy in \ac{nDCG}@10 on the first data fold for our own cluster implementation of ListNet, the RankLib implementation of ListNet, as well as the official LETOR evaluation results of ListNet as reported on the LETOR 3.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and LETOR 4.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} websites.
+Note that the official LETOR evaluation results are structurally higher compared to the results obtained with RankLib and our cluster implementation, which can be attributed to the fact that the official evaluation runs use as many ListNet iterations as needed for convergence on the validation fold, while we limited our experiment to five iterations.
+The official evaluation results of the LETOR benchmarks do not report the standard deviation of their experimental results and it is not mentioned in the LETOR papers \cite{Qin2010, Qin2013} how many times their experiments were repeated.
+For the ListNet step size hyperparameter we choose a value of $0.0001$, which is the default value step size value in RankLib, for both the cluster and the RankLib runs.
+It is not clear what step size was used in the official evaluation runs.
+We expect, given that LETOR aims to compare ranking methods in terms of their ranking potential, that the official LETOR benchmark results optimises the step size hyperparameter on the validation set and reported the performance on the optimal step size found with this procedure.
+Using this procedure one is expected to find a higher ranking accuracy than one would expect to find using a default step size value.
+Hyperparameter optimisation on the validation set would however not be feasible for our cluster implementation of ListNet given the Hadoop job scheduling overhead of the cluster.\\
+
+A remarkable finding is that the cluster implementation shows better ranking accuracy than the RankLib version of ListNet after five iterations.
+We attribute this to the normalisation preprocessing step that we incorporated into our ListNet implementation that transforms each feature into a [0,1]-range.
+RankLib does not perform this normalisation preprocessing step.
+It has been shown that the gradient descent optimisation procedure, which is used in ListNet, converges faster on normalised data \cite{Ng1999}.
+
+\begin{table}
 \centering
@@
 
-\begin{table}[!h]
+\begin{table}
 \centering
@@
 
-\begin{figure}[!h]
+\begin{figure}
 \centering

### 2014-09-22 implementation/java/src/letor/parallel/ListNetCluster.java

@@ -17,7 +17,7 @@ import java.util.Random;
 
 public class ListNetCluster {
     // Initialise hyper-parameters
-    private static final DataSets.DataSet DATASET = DataSets.DataSet.MSLR_WEB30K;
+    private static final DataSets.DataSet DATASET = DataSets.DataSet.MQ2007;
     private static final double   STEPSIZE   = 0.0001; // MSLR-WEB10K: 0.0001, ohsumed: 0.01
     private static final int      ITERATIONS = 5;
     private static final int      FOLDS      = 1;
@@ -127,12 +127,12 @@ public class ListNetCluster {
             pigLines.add("VALIDATE_STD = FOREACH VALIDATE GENERATE flatten(udf.util.ToStandardForm(*));");
             pigLines.add("TEST_STD = FOREACH TEST GENERATE flatten(udf.util.ToStandardForm(*));");
             pigLines.add("DEFINE ScaleFeatures udf.util.ScaleFeatures('"+LtrUtils.toParamString(minmaxList)+"');");
-            pigLines.add("TRAIN_SCA = FOREACH TRAIN_STD GENERATE flatten(ScaleFeatures(*));");
-            pigLines.add("VALIDATE_SCA = FOREACH VALIDATE_STD GENERATE flatten(ScaleFeatures(*));");
-            pigLines.add("TEST_SCA = FOREACH TEST_STD GENERATE flatten(ScaleFeatures(*));");
-            pigLines.add("STORE TRAIN_SCA INTO 'train_sca"+f+"' USING BinStorage();");
-            pigLines.add("STORE VALIDATE_SCA INTO 'validate_sca"+f+"' USING BinStorage();");
-            pigLines.add("STORE TEST_SCA INTO 'test_sca"+f+"' USING BinStorage();");
+            //pigLines.add("TRAIN_SCA = FOREACH TRAIN_STD GENERATE flatten(ScaleFeatures(*));");
+            //pigLines.add("VALIDATE_SCA = FOREACH VALIDATE_STD GENERATE flatten(ScaleFeatures(*));");
+            //pigLines.add("TEST_SCA = FOREACH TEST_STD GENERATE flatten(ScaleFeatures(*));");
+            pigLines.add("STORE TRAIN_STD INTO 'train_sca"+f+"' USING BinStorage();");
+            pigLines.add("STORE VALIDATE_STD INTO 'validate_sca"+f+"' USING BinStorage();");
[…]

### 2014-09-22 report/Chapters/Conclusions.tex
conclusiestukje + nieuwe titel :)

@@
-Running ListNet on a Hadoop cluster using the MapReduce computing model comes with its own cost in the form of a job scheduling overhead in the range of 150-200 seconds per training iteration.
+Using our experimental results we will reflect on our research questions stated in the \ref{sec:goals} section of this thesis. We formulated the following research questions:
+\begin{description}
+\item[RQ1] What are the best performing Learning-to-Rank algorithms in terms of accuracy on relevant benchmark datasets?
+\end{description}
+To answer this research question we proposed a new way of comparing learning to rank methods based on sparse evaluation results data on a set of benchmark datasets.
+Our comparison methodology comprises of two components: 1) NWN, which provides insight in the ranking accuracy of the learning to rank method, and 2) IWN, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.
+Based on our literature search for evaluation results on well-known benchmarks collections, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.
+However, no closing arguments can be formulated on which learning to rank methods are most accurate.
+LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the learning to rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
+From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.
+More evaluation runs are needed for the methods on the left side of Figure \ref{fig:normalised_winning_number_all}.
+Our work contributes to this by identifying promising learning to rank methods that researchers could focus on in performing additional evaluation runs.
+
+\begin{description}
+\item[RQ2] What is the speed-up of those Learning-to-Rank algorithms when executed using the MapReduce framework?\\
+Where the definition of \emph{relative speed-up} is used for speed-up \cite{Sun1991}:\\
+
+$S_N = \frac{\text{execution time using one core}}{\text{execution time using \emph{N} cores}}$
+\end{description}
+To answer this research question we took the approach to implementing the list of algorithms found in answering the first research question starting with the Learning to Rank method with the highest certainty on ranking accuracy, ListNet.
+We found that running ListNet on a Hadoop cluster using the MapReduce computing model comes with its own cost in the form of a job scheduling overhead in the range of 150-200 seconds per training iteration.
 However, with a normalisation preprocessing iteration we can greatly reduce the number of training iterations needed for converge.
@@
 To process data sets in the 100+GB range with the ListNet training algorithm, Hadoop MapReduce is a large improvement compared to a single machine.\\
-
+% TODO: iets zeggen over waarom de oorspronkelijk speed-up metric niet zo geschikt is hier %
 No generalisations can be drawn from this results to other Learning to Rank algorithms.
@@
 Other Learning to Rank algorithms using the gradient descent procedure might not scale well when the MapReduce computing model is used, but any bad scaling behaviour on MapReduce of Learning to Rank algorithms will not be because of gradient descent.
-
-% TODO: how does this generalise to other Learning to Rank algorithms %

### 2014-09-22 report/FrontBackmatter/Preface.tex
herstructurering frontbackmatter

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Preface}{preface}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Preface}
Dear reader,\\

Thank you for taking interest in this thesis, which I have written as final project as part of the Master programme in Computer Science at the University of Twente. This research was conducted at Avanade Netherlands B.V. under the primary supervision of Sander Boackting Msc. I would like to use this page to express my gratitude to everyone who supported my throughout this project in any way.\\

Many thanks go to Dr. ir. Djoerd Hiemstra of the University of Twente and to Sander Bockting Msc. of Avanade Netherlands BV for their great supervision throughout the project.\\

In addition I would like to thank all other graduate interns at Avanade as well as all the Avanade employees for the great talks at the coffee machine, during the Friday afternoon drink, or elsewhere. In particular I owe many thanks to Michel Barneveld for giving me insight in potential applications of Learning-to-Rank within Avanade.\\

-- Niek Tax

\endgroup

### 2014-09-23 report/Chapters/Background.tex
algorithm line numbers + SmoothRank algorithm
naar refactoring werk

@@
 %************************************************
-This chapter is written with the goal to provide the reader with a subtle introduction in the Learning-to-Rank field.
+This chapter is written with the goal to provide the reader with a subtle introduction in the Learning to Rank field.
 Models and theories explained in this chapter can be regarded as prior knowledge needed to understand the subsequent chapters of this thesis.
-\section{A basic introduction to Learning-to-Rank}
-Different definitions of Learning-to-Rank exist.
-In general, all ranking methods that use machine learning technologies to solve the problem of ranking are called Learning-to-Rank methods.
+\section{A basic introduction to Learning to Rank}
+Different definitions of Learning to Rank exist.
+In general, all ranking methods that use machine learning technologies to solve the problem of ranking are called Learning to Rank methods.
 Figure \ref{fig:discriminative_training} describes the general process of machine learning.
@@
 \includegraphics[scale=0.26]{gfx/descriminative_training}
-\caption{Machine learning framework for Learning-to-Rank, obtained from Liu \cite{Liu2007}}
+\caption{Machine learning framework for Learning to Rank, obtained from Liu \cite{Liu2007}}
 \label{fig:discriminative_training}
 \end{figure}\\
-Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning-to-Rank method when it is \emph{feature based} and uses \emph{discriminative training}, which are itself defined as follows:
+Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning to Rank method when it is \emph{feature based} and uses \emph{discriminative training}, which are itself defined as follows:
 \begin{description}
@@
 \item[Discriminative Training]{means that the learning process can be well described by the four components of discriminative learning.
-That is, a Learning-to-Rank method has its own \emph{input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function}, like the machine learning process described by Figure \ref{fig:discriminative_training}.
+That is, a Learning to Rank method has its own \emph{input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function}, like the machine learning process described by Figure \ref{fig:discriminative_training}.
 \emph{Input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function} are hereby defined as follows:
@@
 \includegraphics[scale=0.25]{gfx/ltr_framework}
-\caption{A typical Learning-to-Rank setting, obtained from Liu \cite{Liu2007}}
+\caption{A typical Learning to Rank setting, obtained from Liu \cite{Liu2007}}
 \label{fig:ltr_framework}
@@
 These three approaches are in literature respectively called the pointwise approach, the pairwise approach and the listwise approach.
-These three approaches to Learning-to-Rank will be described in more detail in section \ref{sec:ltr_approaches}.
+These three approaches to Learning to Rank will be described in more detail in section \ref{sec:ltr_approaches}.
 
@@
 \subsubsection{Normalized Discounted Cumulative Gain}
-\ac{nDCG} normalizes the \ac{DCG} metric to a value in the [0,1] interval by dividing by the \ac{DCG} value of the optimal rank.
+\ac{NDCG} normalizes the \ac{DCG} metric to a value in the [0,1] interval by dividing by the \ac{DCG} value of the optimal rank.
 This optimal rank is obtained by sorting documents on relevance for a given query.
-The definition of \ac{nDCG} can be written down mathematically as\\
+The definition of \ac{NDCG} can be written down mathematically as\\
 
@@
 
-Table \ref{tab:example_calculation_nDCG} shows an example calculation for \ac{nDCG} for both the J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} and Burges et al.
+Table \ref{tab:example_calculation_NDCG} shows an example calculation for \ac{NDCG} for both the J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} and Burges et al.
 \cite{Burges2005} version of \ac{DCG}.\\
@@
  &  &  &  &  &  &  &  &  &  &  &  \\
- & \multicolumn{10}{c}{J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} \ac{nDCG} = $\frac{29.17}{29.96} = 0.97$} &  \\  
- & \multicolumn{10}{c}{Burges \cite{Burges2005} et al. \ac{nDCG} = $\frac{732.7}{788.0} = 0.93$} &  \\ 
+ & \multicolumn{10}{c}{J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} \ac{NDCG} = $\frac{29.17}{29.96} = 0.97$} &  \\  
+ & \multicolumn{10}{c}{Burges \cite{Burges2005} et al. \ac{NDCG} = $\frac{732.7}{788.0} = 0.93$} &  \\ 
 \end{tabular}
-\caption{Example calculation of \acs{nDCG}}
-\label{tab:example_calculation_nDCG}
+\caption{Example calculation of \acs{NDCG}}
+\label{tab:example_calculation_NDCG}
 \end{table}
 
-One limitation of \ac{nDCG} that it does not penalise for missing documents in the result set.
-\acl{nDCG} is often used with a fixed set size for the result set to mitigate the missing document problem.
-\ac{nDCG} with a fixed set size is often called \ac{nDCG}@k, where $k$ represents the set size.
+One limitation of \ac{NDCG} that it does not penalise for missing documents in the result set.
+\acl{NDCG} is often used with a fixed set size for the result set to mitigate the missing document problem.
+\ac{NDCG} with a fixed set size is often called \ac{NDCG}@k, where $k$ represents the set size.
 
 \subsection{Expected Reciprocal Rank}
-\ac{ERR} \cite{Chapelle2009} was designed based on the observation that \ac{nDCG} is based on the false assumption that the usefulness of a document at rank $i$ is independent of the usefulness of the documents at rank less than $i$.
+\ac{ERR} \cite{Chapelle2009} was designed based on the observation that \ac{NDCG} is based on the false assumption that the usefulness of a document at rank $i$ is independent of the usefulness of the documents at rank less than $i$.
 \ac{ERR} is based on the reasoning that users are likely to stop exploring the result list once they have found a document that satisfied their information need.
@@
 
+\LinesNumbered
 \begin{algorithm}[H]
@@
 
-\section{Approaches to Learning-to-Rank}
+\section{Approaches to Learning to Rank}
 \label{sec:ltr_approaches}
@@
 The pointwise approach can be seen as the most straightforward way of using machine learning for ranking.
-Pointwise Learning-to-Rank methods directly apply machine learning methods to the ranking problem by observing each document in isolation.
+Pointwise Learning to Rank methods directly apply machine learning methods to the ranking problem by observing each document in isolation.
 They can be subdivided in the following approaches:
@@
 \subsection{Pairwise Approach}
-Pointwise Learning-to-Rank methods have the drawback that they optimise real valued expected relevance, while evaluation metrics like \ac{nDCG} and \ac{ERR} are only impacted by a change in expected relevance when that change impacts a pairwise preference.
+Pointwise Learning to Rank methods have the drawback that they optimise real valued expected relevance, while evaluation metrics like \ac{NDCG} and \ac{ERR} are only impacted by a change in expected relevance when that change impacts a pairwise preference.
 The pairwise approach solves this drawback of the pointwise approach by regarding ranking as pairwise classification.\\
@@
 
-Well-known pairwise Learning-to-Rank algorithms include FRank \cite{Tsai2007}, GBRank \cite{Zheng2007}, LambdaRank \cite{Burges2006}, RankBoost \cite{Freund2003}, RankNet \cite{Burges2005}, Ranking \acs{SVM} \cite{Herbrich1999b,Joachims2002}, and SortNet \cite{Rigutini2008}.
+Well-known pairwise Learning to Rank algorithms include FRank \cite{Tsai2007}, GBRank \cite{Zheng2007}, LambdaRank \cite{Burges2006}, RankBoost \cite{Freund2003}, RankNet \cite{Burges2005}, Ranking \acs{SVM} \cite{Herbrich1999b,Joachims2002}, and SortNet \cite{Rigutini2008}.
 \subsection{Listwise Approach}
 Listwise ranking optimises the actual evaluation metric.
-The learner learns to predict an actual ranking itself without using an intermediate step like in pointwise or pairwise Learning-to-Rank.
+The learner learns to predict an actual ranking itself without using an intermediate step like in pointwise or pairwise Learning to Rank.
 The main challenge in this approach is that most evaluation metrics are not differentiable.
-\ac{MAP}, \ac{ERR} and \ac{nDCG} are non-differentiable, non-convex and discontinuous functions, what makes them very hard to optimize.\\
+\ac{MAP}, \ac{ERR} and \ac{NDCG} are non-differentiable, non-convex and discontinuous functions, what makes them very hard to optimize.\\
 
-Although the properties of \ac{MAP}, \ac{ERR} and \ac{nDCG} are not ideal for direct optimisation, some listwise approaches do focus on direct metric optimisation \cite{Yue2007, Taylor2008, Chapelle2010}.
-Most listwise approaches work around optimisation of the non-differentiable, non-convex and discontinuous metrics by optimising surrogate cost functions that mimic the behaviour of \ac{MAP}, \ac{ERR} or \ac{nDCG}, but have nicer properties for optimisation.\\
+Although the properties of \ac{MAP}, \ac{ERR} and \ac{NDCG} are not ideal for direct optimisation, some listwise approaches do focus on direct metric optimisation \cite{Yue2007, Taylor2008, Chapelle2010}.
+Most listwise approaches work around optimisation of the non-differentiable, non-convex and discontinuous metrics by optimising surrogate cost functions that mimic the behaviour of \ac{MAP}, \ac{ERR} or \ac{NDCG}, but have nicer properties for optimisation.\\
 

### 2014-09-23 report/Chapters/BenchmarkResults.tex
naar refactoring werk

@@
 This chapter describes benchmark characteristics, e.g.
-collection size and features, of Learning-to-Rank benchmark collections and datasets and gives an overview of the performance of baseline algorithms on these collections and datasets.\\
+collection size and features, of Learning to Rank benchmark collections and datasets and gives an overview of the performance of baseline algorithms on these collections and datasets.\\
 
-The accuracies of the Learning-to-Rank methods described in the following sections can only be compared within the benchmark and not between benchmarks for the following reasons:
+The accuracies of the Learning to Rank methods described in the following sections can only be compared within the benchmark and not between benchmarks for the following reasons:
 \begin{enumerate}
@@
 Yahoo!
-published a subset of their own commercial training data and launched a Learning-to-Rank competition based on this data.
+published a subset of their own commercial training data and launched a Learning to Rank competition based on this data.
 The Yahoo!
-Learning to Rank Challenge \cite{Chapelle2011a} is a public Learning-to-Rank competition which took place from March to May 2010, with the goal to promote the datasets and encourage the research community to develop new Learning-to-Rank algorithms.\\
+Learning to Rank Challenge \cite{Chapelle2011a} is a public Learning to Rank competition which took place from March to May 2010, with the goal to promote the datasets and encourage the research community to develop new Learning to Rank algorithms.\\
 
 The Yahoo!
-Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning-to-Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country.
-For this experiment, I will only look at the standard Learning-to-Rank dataset, because transfer learning is a separate research area that is not included in this thesis.\\
+Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning to Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country.
+For this experiment, I will only look at the standard Learning to Rank dataset, because transfer learning is a separate research area that is not included in this thesis.\\
 \begin{table}[!h]
@@
 Both \ac{nDCG} and \ac{ERR} are measured as performance metrics, but the final standings of the challenge were based on the \ac{ERR} values.
-Model validation on the Learning-to-Rank methods participating in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}.
+Model validation on the Learning to Rank methods participating in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}.
 Competitors could train on the training set and get immediate feedback on their performance on the validation set.
@@
 \section{LETOR}
-The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 \cite{Liu2007b} to solve the absence of a experimental platform for Learning-to-Rank at that time.
+The LETOR benchmark set was first released by Microsoft Research Asia in April 2007 \cite{Liu2007b} to solve the absence of a experimental platform for Learning to Rank at that time.
 LETOR has been updated several times since: LETOR 2.0 was released at the end of 2007, LETOR 3.0 \cite{Qin2010} in December 2008 and LETOR 4.0 \cite{Qin2013} in July 2009.
@@
 
-LETOR 2.0 only uses the topic distillation queries. Baseline Learning-to-Rank algorithms evaluated on LETOR 2.0 by the organisation are:
+LETOR 2.0 only uses the topic distillation queries. Baseline Learning to Rank algorithms evaluated on LETOR 2.0 by the organisation are:
 \begin{description}
@@
 The \emph{winning number} metric is defined as the number of other algorithms that it can beat over all of the seven datasets (six .gov sets + OHSUMED).
-The LETOR organisation implemented and evaluated several well-known Learning-to-Rank algorithms themselves and in addition gathers and publications and results of new algorithms evaluated on the LETOR benchmark.
-The baseline Learning-to-Rank algorithms evaluated by the LETOR team are:
+The LETOR organisation implemented and evaluated several well-known Learning to Rank algorithms themselves and in addition gathers and publications and results of new algorithms evaluated on the LETOR benchmark.
+The baseline Learning to Rank algorithms evaluated by the LETOR team are:
 \begin{description}
@@
 \includegraphics[scale=0.30]{gfx/ndcg_winning_number}
-\caption{Comparison across the seven datasets in LETOR by \acs{nDCG}, obtained from Qin et al. \cite{Qin2010}}
+\caption{Comparison of ranking accuracy across the seven datasets in LETOR by \acs{NDCG}, obtained from Qin et al. \cite{Qin2010}}
 \label{fig:ndcg_winning_number}
@@
 
-Since LETOR is arguably the most well-known benchmark collection in the field, it is conceivable that creators of new Learning-to-Rank methods evaluate their new method on the LETOR collection to show how well their new method works.
+Since LETOR is arguably the most well-known benchmark collection in the field, it is conceivable that creators of new Learning to Rank methods evaluate their new method on the LETOR collection to show how well their new method works.
 
@@
 \section{Other datasets}
-LETOR and the Yahoo! Learning to Rank Challenge datasets are the most used benchmark collections in the field. Several other benchmark collections for the Learning-to-Rank task have been proposed.
+LETOR and the Yahoo! Learning to Rank Challenge datasets are the most used benchmark collections in the field. Several other benchmark collections for the Learning to Rank task have been proposed.
 
@@
 Pass et al.
-\cite{Pass2006} describe an America Online Learning-to-Rank dataset which they published in 2006.
+\cite{Pass2006} describe an America Online Learning to Rank dataset which they published in 2006.
 This dataset is unique in that it is the only large commercial web search dataset that contains user session and user behaviour information.
@@
 The removal of the data was too late, as the data was already downloadable from several mirror sites\footnote{http://gregsadetsky.com/aol-data/}.
-This controversial AOL dataset is to date still used for Learning-to-Rank research.
+This controversial AOL dataset is to date still used for Learning to Rank research.
 No official baseline evaluation is provided by the AOL team.
@@
 \subsection{Yandex Internet Mathematics contest}
-Russian commercial web search engine Yandex dedicated their yearly Internet Mathematics contest to the task of Learning-to-Rank in the 2009 edition of the contest.
+Russian commercial web search engine Yandex dedicated their yearly Internet Mathematics contest to the task of Learning to Rank in the 2009 edition of the contest.
 Features (245 in total) are only numbered and their semantics are not revealed, equivalent to the Yahoo!

### 2014-09-23 report/Chapters/CrossBenchmarkComparison.tex
naar refactoring werk

@@
 \label{chap:cross_comparison}
-As we have seen in Chapter~\ref{chap:benchmark_results}, the evaluation of Learning-to-Rank methods is spread over several benchmark datasets.
-However, as the Learning-to-Rank methods evaluated differs between benchmarks, no single benchmark comparison can be regarded as a conclusive argument on which Learning-to-Rank method is most accurate.\\
+As we have seen in Chapter~\ref{chap:benchmark_results}, the evaluation of Learning to Rank methods is spread over several benchmark datasets.
+However, as the Learning to Rank methods evaluated differs between benchmarks, no single benchmark comparison can be regarded as a conclusive argument on which Learning to Rank method is most accurate.\\
 
-Several studies make a small start in considering Learning-to-Rank methods performance over multiple benchmark datasets.
+Several studies make a small start in considering Learning to Rank methods performance over multiple benchmark datasets.
 Gomes et al.
@@
 To my knowledge, no structured meta-analysis on ranking accuracy has been conducted where evaluation results on several benchmark collections are taken into account.
-With a meta-analysis we will compare the performance of Learning-to-Rank methods across the Learning-to-Rank benchmark datasets described in foregoing sections.
+With a meta-analysis we will compare the performance of Learning to Rank methods across the Learning to Rank benchmark datasets described in foregoing sections.
 
@@
 Users of those benchmark collections are encouraged to cite these papers.
-Therefore, we collect evaluation measurements of Learning-to-Rank methods on these benchmark collections through forward literature search.
+Therefore, we collect evaluation measurements of Learning to Rank methods on these benchmark collections through forward literature search.
 Table~\ref{tbl:ltr_benchmark_forref} presents an overview of the results of this forward literature search.
@@
 \end{tabular}
-\caption{Forward references of Learning-to-Rank benchmark papers}
+\caption{Forward references of Learning to Rank benchmark papers}
 \label{tbl:ltr_benchmark_forref}
@@
 The LETOR 4.0, MSLR-web10/30k and Yandex Internet Mathematics Competition 2009 benchmark collections were not accompanied with a describing study.
-To collect measurements of Learning-to-Rank methods evaluated on these benchmarks, a Google Scholar search is performed on the name of the benchmark.
+To collect measurements of Learning to Rank methods evaluated on these benchmarks, a Google Scholar search is performed on the name of the benchmark.
 Table~\ref{chap:benchmark_results} shows the results of this literature search.
@@
 \end{tabular}
-\caption{Google scholar search results for Learning-to-Rank benchmarks}
+\caption{Google scholar search results for Learning to Rank benchmarks}
 \label{tbl:ltr_benchmark_searchres}
@@
 The LETOR 3.0 paper \cite{Qin2010} states that it may differ between datasets what the most accurate ranking methods are.
-To evaluate the overall performance of Learning-to-Rank methods over the multiple datasets in the LETOR 3.0 collections, Qin et al.
+To evaluate the overall performance of Learning to Rank methods over the multiple datasets in the LETOR 3.0 collections, Qin et al.
 \cite{Qin2010} proposed a measure called \emph{winning number} as the number of other algorithms that an algorithm can beat over the set of datasets.
@@
 
-where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{nDCG} or \ac{MAP}), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that\\
+where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{NDCG} or \ac{MAP}), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that\\
 
@@
 
-where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{nDCG} or \ac{MAP}), and $D_{\{M_i(j),M_k(j)\}}$ is an evaluation definition function such that\\
+where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{NDCG} or \ac{MAP}), and $D_{\{M_i(j),M_k(j)\}}$ is an evaluation definition function such that\\
 
@@
 
-\ac{nDCG}@3, \ac{nDCG}@5, \ac{nDCG}@10 and \ac{MAP} are chosen as metrics on which the meta-analysis will be performed.
+\ac{NDCG}@3, \ac{NDCG}@5, \ac{NDCG}@10 and \ac{MAP} are chosen as metrics on which the meta-analysis will be performed.
 These metrics seem to be the most frequently used evaluation metrics in most of the used benchmark datasets.
@@
 The lack of use of the \ac{ERR}-metric in other benchmarks makes it unsuitable for a cross-benchmark comparison.
-By making the decision to include \ac{nDCG} at three cut-off points and only a single \ac{MAP} entry, we implicitly attain a higher weight for \ac{nDCG} compared to \ac{MAP} on an analysis that combines all measurements on the four metrics.
+By making the decision to include \ac{NDCG} at three cut-off points and only a single \ac{MAP} entry, we implicitly attain a higher weight for \ac{NDCG} compared to \ac{MAP} on an analysis that combines all measurements on the four metrics.
 This implicit weighting is arbitrary and therefore undesirable, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.
@@
 \section{Evaluation Results Found in Literature}
-Table \ref{tab:ltr_methods_used} gives an overview of the Learning-to-Rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
+Table \ref{tab:ltr_methods_used} gives an overview of the Learning to Rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
 Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, LETOR 3.0 and LETOR 4.0 baselines respectively.\\
@@
 \item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011, Lin2012}
-\item The study focussed on a different Learning-to-Rank task (e.g.
+\item The study focussed on a different Learning to Rank task (e.g.
 rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}
@@
 AdaRank-\acs{MAP} & \cite{Xu2007} & L2, L3, L4 & Linear Regression & \cite{Cossock2006} & L3, \cite{Wang2012, Volkovs2011} \\ 
-AdaRank-\acs{nDCG} & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
+AdaRank-\acs{NDCG} & \cite{Xu2007} & L2, L3, L4,  \cite{Busa-Fekete2013,Tan2013} & List\acs{MLE} & \cite{Xia2008} & \cite{Lin2010, Lin2011, Gao2014} \\ 
 ADMM & \cite{Duh2011} & \cite{Duh2011} & ListNet & \cite{Cao2007} & L2, L3, L4 \\ 
 Approx\acs{AP} & \cite{Qin2010b} & \cite{Qin2010b} & ListReg & \cite{Wu2011} & \cite{Wu2011} \\ 
-Approx\acs{nDCG} & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
+Approx\acs{NDCG} & \cite{Qin2010b} & \cite{Qin2010b} & LRUF & \cite{Torkestani2012b} & \cite{Torkestani2012b} \\ 
 BagBoo & \cite{Pavlov2010} & \cite{Ganjisaffar2011c} & MCP & \cite{Laporte2013} & \cite{Laporte2013} \\ 
@@
 CoList & \cite{Gao2014} & \cite{Gao2014} & Rank-PMBGP & \cite{Sato2013} & \cite{Sato2013} \\ 
-Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAgg\acs{nDCG} & \cite{Wang2013} & \cite{Wang2013} \\ 
+Consistent-RankCosine & \cite{Ravikumar2011} & \cite{Tan2013} & RankAgg\acs{NDCG} & \cite{Wang2013} & \cite{Wang2013} \\ 
 DCMP & \cite{Renjifo2012}  & \cite{Renjifo2012}  & RankBoost & \cite{Freund2003} & L2, L3, L4, \cite{Busa-Fekete2013, Alcantara2010} \\ 
 DirectRank & \cite{Tan2013} & \cite{Tan2013} & RankCSA & \cite{He2010} & \cite{He2010} \\ 
-Energy\acs{nDCG} & \cite{Freno2011} & \cite{Freno2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\ 
+Energy\acs{NDCG} & \cite{Freno2011} & \cite{Freno2011} & RankDE & \cite{Bollegala2011} & \cite{Sato2013} \\ 
 FBPCRank & \cite{Lai2011} & \cite{Lai2011} & RankELM (pairwise) & \cite{Zong2013} & \cite{Zong2013} \\ 
@@
 \end{tabular}}
-\caption{Learning-to-Rank algorithms with measurements on benchmark datasets}
+\caption{Learning to Rank algorithms with measurements on benchmark datasets}
 \label{tab:ltr_methods_used}
@@
 \section{Results \& Discussion}
-The following subsections provide the performance of Learning-to-Rank methods in terms of Normalised Winning Number for \ac{nDCG}@3, \ac{nDCG}@5, \ac{nDCG}@10 and \ac{MAP}.
-Performance of the Learning-to-Rank methods is plotted with Normalised Winning Number on the vertical axis and the number of datasets on which the method has been evaluated on the horizontal axis.
-The further to the right, the more certain we can be about the performance of the Learning-to-Rank method.
+The following subsections provide the performance of Learning to Rank methods in terms of Normalised Winning Number for \ac{NDCG}@3, \ac{NDCG}@5, \ac{NDCG}@10 and \ac{MAP}.
+Performance of the Learning to Rank methods is plotted with Normalised Winning Number on the vertical axis and the number of datasets on which the method has been evaluated on the horizontal axis.
+The further to the right, the more certain we can be about the performance of the Learning to Rank method.
 The methods for which it holds that there is no other method that has 1) a higher Normalised Winning Number and 2) a higher number datasets evaluated, are identified as the best performing methods and are labelled with the name of the method.
@@
 \subsection{NDCG@3}
-Figure \ref{fig:normalised_winning_number_ndcg3} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@3 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg3} in Appendix \ref{app:norm_winnum_ndcg3} provides the raw Normalised Winning Number data for the Learning-to-Rank methods for which \ac{nDCG}@3 evaluation results were available.\\
+Figure \ref{fig:normalised_winning_number_NDCG3} shows the performance of Learning to Rank methods for the \ac{nDCG}@3 metric.
+Table \ref{tab:raw_data_norm_winnum_NDCG3} in Appendix \ref{app:norm_winnum_NDCG3} provides the raw Normalised Winning Number data for the Learning to Rank methods for which \ac{NDCG}@3 evaluation results were available.\\
 
@@
 \includegraphics[scale=0.285]{gfx/ndcg3_winnum}
-\caption{\acs{nDCG}@3 comparison of Learning-to-Rank methods}
-\label{fig:normalised_winning_number_ndcg3}
+\caption{\acs{NDCG}@3 comparison of Learning to Rank methods}
+\label{fig:normalised_winning_number_NDCG3}
 \end{figure}
@@
 \subsection{NDCG@5}
-Figure \ref{fig:normalised_winning_number_ndcg5} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@5 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg5} in Appendix \ref{app:norm_winnum_ndcg5} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.\\
+Figure \ref{fig:normalised_winning_number_NDCG5} shows the performance of Learning to Rank methods for the \ac{NDCG}@5 metric.
+Table \ref{tab:raw_data_norm_winnum_NDCG5} in Appendix \ref{app:norm_winnum_NDCG5} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\
 
@@
 \includegraphics[scale=0.285]{gfx/ndcg5_winnum}
-\caption{\acs{nDCG}@5 comparison of Learning-to-Rank methods}
-\label{fig:normalised_winning_number_ndcg5}
+\caption{\acs{NDCG}@5 comparison of Learning to Rank methods}
+\label{fig:normalised_winning_number_NDCG5}
 \end{figure}
@@
 LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of datasets, but score decreasingly well in terms of Normalised Winning Number.
-These results are highly in agreement with the \ac{nDCG}@3 comparison.
-The only modification compared to the \ac{nDCG}@3 comparison being that OWPC did show to be a method for which there were no methods performing better on both axes in the \ac{nDCG}@5 comparison, but not in the \ac{nDCG}@3 comparison.
-Like in the \ac{nDCG}@3 comparison, SmoothRank and ListNet can be regarded as most reliable results because the evaluation measurements for these methods are based on LETOR official baselines.
+These results are highly in agreement with the \ac{NDCG}@3 comparison.
+The only modification compared to the \ac{NDCG}@3 comparison being that OWPC did show to be a method for which there were no methods performing better on both axes in the \ac{NDCG}@5 comparison, but not in the \ac{NDCG}@3 comparison.
+Like in the \ac{NDCG}@3 comparison, SmoothRank and ListNet can be regarded as most reliable results because the evaluation measurements for these methods are based on LETOR official baselines.
 
 \subsection{NDCG@10}
-Figure \ref{fig:normalised_winning_number_ndcg10} shows the performance of Learning-to-Rank methods for the \ac{nDCG}@10 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg10} in Appendix \ref{app:norm_winnum_ndcg10} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.\\
+Figure \ref{fig:normalised_winning_number_NDCG10} shows the performance of Learning to Rank methods for the \ac{NDCG}@10 metric.
+Table \ref{tab:raw_data_norm_winnum_NDCG10} in Appendix \ref{app:norm_winnum_NDCG10} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\
 
@@
 \includegraphics[scale=0.285]{gfx/ndcg10_winnum}
-\caption{\acs{nDCG}@10 comparison of Learning-to-Rank methods}
-\label{fig:normalised_winning_number_ndcg10}
+\caption{\acs{NDCG}@10 comparison of Learning to Rank methods}
+\label{fig:normalised_winning_number_NDCG10}
 \end{figure}
 
-LambdaMART and LambdaNeuralRank score a Normalised Winning Number of 1.0 on the \ac{nDCG}@10 comparison.
+LambdaMART and LambdaNeuralRank score a Normalised Winning Number of 1.0 on the \ac{NDCG}@10 comparison.
 For LambdaNeuralRank these results are again based on AOL dataset measurements.
-LambdaMART showed the highest \ac{nDCG}@10 performance for the MSLR-WEB10k dataset.
-The set of algorithms for which there is no other algorithm with both a higher Normalised Winning Number and number of datasets evaluated on is partly in agreement with those for the \ac{nDCG}@3 and \ac{nDCG}@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
-SmoothRank and FSMRank were not present in this set for the \ac{nDCG}@3 and \ac{nDCG}@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_ndcg3} and \ref{tab:raw_data_norm_winnum_ndcg5} in the Appendices.
-DCMP is not in the set in contrast with the \ac{nDCG}@3 and \ac{nDCG}@5 comparison.
+LambdaMART showed the highest \ac{NDCG}@10 performance for the MSLR-WEB10k dataset.
+The set of algorithms for which there is no other algorithm with both a higher Normalised Winning Number and number of datasets evaluated on is partly in agreement with those for the \ac{NDCG}@3 and \ac{NDCG}@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
+SmoothRank and FSMRank were not present in this set for the \ac{NDCG}@3 and \ac{NDCG}@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_NDCG3} and \ref{tab:raw_data_norm_winnum_NDCG5} in the Appendices.
+DCMP is not in the set in contrast with the \ac{NDCG}@3 and \ac{NDCG}@5 comparison.
 
 \subsection{MAP}
-Figure \ref{fig:normalised_winning_number_map} shows the performance of Learning-to-Rank methods for the \ac{MAP} metric.
-Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw Normalised Winning Number data for the Learning-to-Rank methods.\\
+Figure \ref{fig:normalised_winning_number_map} shows the performance of Learning to Rank methods for the \ac{MAP} metric.
+Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\
 
@@
 \includegraphics[scale=0.285]{gfx/map_winnum}
-\caption{\acs{MAP} comparison of Learning-to-Rank methods}
+\caption{\acs{MAP} comparison of Learning to Rank methods}
 \label{fig:normalised_winning_number_map}
@@
 
-Where comparisons on the \ac{nDCG}-metric at different cut-off points where highly in agreement in terms of the best performing algorithms, the comparison in terms of \ac{MAP} shows different results.
-RankDE scores a Normalised Winning Number of 1.0 on one dataset, like LambdaNeuralRank did on for all \ac{nDCG}-comparisons.
+Where comparisons on the \ac{NDCG}-metric at different cut-off points where highly in agreement in terms of the best performing algorithms, the comparison in terms of \ac{MAP} shows different results.
+RankDE scores a Normalised Winning Number of 1.0 on one dataset, like LambdaNeuralRank did on for all \ac{NDCG}-comparisons.
 In contrast to LambdaNeuralRank, RankDE achieved this highest score on the LETOR 2.0 TD2003, a dataset on which many methods are evaluated.\\
 
-LARF and LRUF score very high Normalised Winning Number scores, but based on only relatively few datasets, just as in the \ac{nDCG}-comparisons.
-Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all \ac{nDCG}-comparisons, are not within the best performing methods in the \ac{MAP}-comparison.
-A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_ndcg3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_ndcg3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more datasets for \ac{MAP} compared to \ac{nDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalised Winning Number.
+LARF and LRUF score very high Normalised Winning Number scores, but based on only relatively few datasets, just as in the \ac{NDCG}-comparisons.
+Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all \ac{NDCG}-comparisons, are not within the best performing methods in the \ac{MAP}-comparison.
+A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_NDCG3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_NDCG3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more datasets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalised Winning Number.
 SmoothRank performed a Normalised Winning Number of around 0.53 over 7 datasets, which is still good in both certainty and accuracy, but not among the top methods.
 RE-QR is one of the best performers in the \ac{MAP}-comparison with a reasonable amount of benchmark evaluations.
-No reported \ac{nDCG} performance was found in the literature study for RE-QR.
-There is a lot of certainty on the accuracy of RankBoost and Rank\ac{SVM} as both models are evaluated on the majority of datasets included in the comparison for the \ac{MAP}-metric, but given their Normalised Winning Number it can said that both methods are not within the top performing Learning-to-Rank methods.
+No reported \ac{NDCG} performance was found in the literature study for RE-QR.
+There is a lot of certainty on the accuracy of RankBoost and Rank\ac{SVM} as both models are evaluated on the majority of datasets included in the comparison for the \ac{MAP}-metric, but given their Normalised Winning Number it can said that both methods are not within the top performing Learning to Rank methods.
 
@@
 \includegraphics[scale=0.33]{gfx/combined_normalized_winnum}
-\caption{Cross-benchmark comparison of Learning-to-Rank methods}
+\caption{Cross-benchmark comparison of Learning to Rank methods}
 \label{fig:normalised_winning_number_all}
@@
 
-The cross-metric comparison is based on the \ac{nDCG}@3, \ac{nDCG}@5, \ac{nDCG}@10 and \ac{MAP} comparisons combined, which justifies analysing the comparison more thoroughly.
+The cross-metric comparison is based on the \ac{NDCG}@3, \ac{NDCG}@5, \ac{NDCG}@10 and \ac{MAP} comparisons combined, which justifies analysing the comparison more thoroughly.
 Figure \ref{fig:normalised_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
@@
 LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both Ideal Winning Number and Normalised Winning Number.
-LRUF is the only method that achieved this in all \ac{nDCG} comparisons, the \ac{MAP} comparison as well as the cross-metric comparison.
-With FenchelRank, FSMRank, SmoothRank and ListNet being among the top performing methods in all \ac{nDCG} comparisons as well as in the cross-metric comparison, it can be concluded that the cross-metric results are highly defined by the \ac{nDCG} performance as opposed to the \ac{MAP} performance.
-This was to be expected, because the cross-metric comparison input data of three \ac{nDCG} entries (@3, @5, and @10) enables it to have up to three times as many as many weight as the \ac{MAP} comparison.\\
+LRUF is the only method that achieved this in all \ac{NDCG} comparisons, the \ac{MAP} comparison as well as the cross-metric comparison.
+With FenchelRank, FSMRank, SmoothRank and ListNet being among the top performing methods in all \ac{NDCG} comparisons as well as in the cross-metric comparison, it can be concluded that the cross-metric results are highly defined by the \ac{NDCG} performance as opposed to the \ac{MAP} performance.
+This was to be expected, because the cross-metric comparison input data of three \ac{NDCG} entries (@3, @5, and @10) enables it to have up to three times as many as many weight as the \ac{MAP} comparison.\\
 
 LARF, \ac{IP}Rank and DCMP and several variants of Rank\ac{SVM} performed very well on the cross-metric comparison, with all having only one method in its top right quadrant.
-LARF also performed among the top methods on the \ac{nDCG} and \ac{MAP} comparisons and DCMP was a top performer in a few of the \ac{nDCG} comparisons.\\
+LARF also performed among the top methods on the \ac{NDCG} and \ac{MAP} comparisons and DCMP was a top performer in a few of the \ac{NDCG} comparisons.\\
 
@@
 LAC-MR-OR and RE-QR showed very good ranking accuracy in the \ac{MAP} comparison on multiple datasets.
-Because LAC-MR-OR is only evaluated on two datasets for \ac{nDCG}@10 and RE-QR is not evaluated for \ac{nDCG} at all, LAC-MR-OR and RE-QR are not within the top performing methods in the cross-metric comparison.
+Because LAC-MR-OR is only evaluated on two datasets for \ac{NDCG}@10 and RE-QR is not evaluated for \ac{NDCG} at all, LAC-MR-OR and RE-QR are not within the top performing methods in the cross-metric comparison.
 
@@
 In the Normalised Winning Number calculation the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
-By calculating it in this way, we implicitly make the assumption that the Learning-to-Rank methods are (approximately) distributed uniformly over the benchmarks, such that the average Learning-to-Rank method tested are approximately equally hard for each data set.
-It could be the case however that this assumption is false and that significantly more accurate Learning-to-Rank methods are evaluated on some datasets than on other datasets.
+By calculating it in this way, we implicitly make the assumption that the Learning to Rank methods are (approximately) distributed uniformly over the benchmarks, such that the average Learning to Rank method tested are approximately equally hard for each data set.
+It could be the case however that this assumption is false and that significantly more accurate Learning to Rank methods are evaluated on some datasets than on other datasets.
 \\
 
-A second limitation is that the datasets on which Learning-to-Rank methods have been evaluated can not always be regarded a random choice.
-It might be the case that some researchers chose to publish results for exactly those benchmark datasets that showed the most positive results for their Learning-to-Rank method.\\
+A second limitation is that the datasets on which Learning to Rank methods have been evaluated can not always be regarded a random choice.
+It might be the case that some researchers chose to publish results for exactly those benchmark datasets that showed the most positive results for their Learning to Rank method.\\
 
@@
 \begin{description}
-\item[RQ1] What are the best performing Learning-to-Rank algorithms in terms of accuracy on relevant benchmark datasets?\\
+\item[RQ1] What are the best performing Learning to Rank algorithms in terms of accuracy on relevant benchmark datasets?\\
 \end{description}
-Although no closing arguments can be formulated on which Learning-to-Rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.\\
+Although no closing arguments can be formulated on which Learning to Rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.\\
 
-LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the Learning-to-Rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
+LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the Learning to Rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
 From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.
@@
 
-In the following chapters of this thesis, concerning parallel execution of the Learning-to-Rank training phase, the scope will be limited to the five methods that turned out to be superior Learning-to-Rank methods in terms of ranking accuracy and certainty about this ranking accuracy: LRUF, FSMRank, FenchelRank, SmoothRank and ListNet.
-Although it can not be concluded that these methods are inarguably the most accurate Learning-to-Rank methods, a strong presumption has been raised that these five Learning-to-Rank are accurate ranking methods.
+In the following chapters of this thesis, concerning parallel execution of the Learning to Rank training phase, the scope will be limited to the five methods that turned out to be superior Learning to Rank methods in terms of ranking accuracy and certainty about this ranking accuracy: LRUF, FSMRank, FenchelRank, SmoothRank and ListNet.
+Although it can not be concluded that these methods are inarguably the most accurate Learning to Rank methods, a strong presumption has been raised that these five Learning to Rank are accurate ranking methods.

### 2014-09-23 report/Chapters/Implementation.tex
naar refactoring werk

@@
 The first Pig job calculates the Cross Entropy loss on training data of the current model and calculates the gradients for the next model update.
-The second Pig job is an internal validation step that validates the model on the validation set by calculating \ac{nDCG}@k.\\
+The second Pig job is an internal validation step that validates the model on the validation set by calculating \ac{NDCG}@k.\\
 
@@
 \midrule
-udf.util.Ndcg() & Calculates \ac{nDCG}@k for a query.\\
+udf.util.Ndcg() & Calculates \ac{NDCG}@k for a query.\\
 \bottomrule
@@
 \subsection{Testing}
-The testing stage tests the best model found in the training iterations, selected on validation set \ac{nDCG}@k (as calculated in the second Pig job of the training stage), by calculating the \ac{nDCG}@k of this model on the test set.\\
+The testing stage tests the best model found in the training iterations, selected on validation set \ac{NDCG}@k (as calculated in the second Pig job of the training stage), by calculating the \ac{NDCG}@k of this model on the test set.\\
 

### 2014-09-23 report/Chapters/Introduction.tex
naar refactoring werk

@@
 The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models.
-Learning-to-Rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\
+Learning to Rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\
 
-In recent years several Learning-to-Rank benchmark datasets have been proposed that enable comparison of the performance of different Learning-to-Rank methods.
+In recent years several Learning to Rank benchmark datasets have been proposed that enable comparison of the performance of different Learning to Rank methods.
 Well-known benchmark datasets include the \emph{Yahoo!
@@
 One of the concluding observations of the \emph{Yahoo!
-Learning to Rank Challenge} was that almost all work in the Learning-to-Rank field focuses on ranking accuracy, while efficiency and scalability of Learning-to-Rank algorithms is still an underexposed research area that is likely to become more important in the near future as training sets are becoming larger and larger \cite{Chapelle2011b}.
-Liu \cite{Liu2007} confirms the observation that efficiency and scalability of Learning-to-Rank methods has so far been an overlooked research area in his influential book on Learning-to-Rank.\\
+Learning to Rank Challenge} was that almost all work in the Learning to Rank field focuses on ranking accuracy, while efficiency and scalability of Learning to Rank algorithms is still an underexposed research area that is likely to become more important in the near future as training sets are becoming larger and larger \cite{Chapelle2011b}.
+Liu \cite{Liu2007} confirms the observation that efficiency and scalability of Learning to Rank methods has so far been an overlooked research area in his influential book on Learning to Rank.\\
 
 Some research has been done in the area of parallel or distributed machine learning \cite{Chu2007,Chang2007}.
-However, almost none of these studies include the Learning-to-Rank sub-field of machine learning.
-The field of efficient Learning-to-Rank has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012}, since Liu \cite{Liu2007} first stated its growing importance back in 2007.
-Only several of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient Learning-to-Rank through the use of parallel programming paradigms.\\
+However, almost none of these studies include the Learning to Rank sub-field of machine learning.
+The field of efficient Learning to Rank has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012}, since Liu \cite{Liu2007} first stated its growing importance back in 2007.
+Only several of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient Learning to Rank through the use of parallel programming paradigms.\\
 
@@
 Since Google developed the MapReduce parallel programming framework back in 2004 it has grown to be the industry standard model for parallel programming.
-Lin \cite{Lin2013} observed that algorithms that are of iterative nature, which most Learning-to-Rank algorithms are, are not amenable to the MapReduce framework.
+Lin \cite{Lin2013} observed that algorithms that are of iterative nature, which most Learning to Rank algorithms are, are not amenable to the MapReduce framework.
 Lin argued that as a solution to the non-amenability of iterative algorithms to the MapReduce framework, iterative algorithms can often be replaced with non-iterative alternatives or can still be optimized in such a way that its performance in a MapReduce setting is good enough.
@@
 
-The appearance of benchmark datasets gave insight in the performance of different Learning-to-Rank approaches, which resulted in increasing popularity of those methods that showed to perform well on one or more benchmark datasets.
-Up to now it remains unknown whether popular existing Learning-to-Rank methods scale well when they are used in a parallel manner using the MapReduce framework.
-This thesis aims to be an exploratory start in this little researched area of parallel Learning-to-Rank.
+The appearance of benchmark datasets gave insight in the performance of different Learning to Rank approaches, which resulted in increasing popularity of those methods that showed to perform well on one or more benchmark datasets.
+Up to now it remains unknown whether popular existing Learning to Rank methods scale well when they are used in a parallel manner using the MapReduce framework.
+This thesis aims to be an exploratory start in this little researched area of parallel Learning to Rank.
 A more extensive overview of my research goals and questions are described in section \ref{sec:goals}.\\
@@
 \label{sec:goals}
-The objective of this thesis is to explore the speed-up in execution time of Learning-to-Rank algorithms through parallelisation using the MapReduce framework.
-This work focuses on those Learning-to-Rank algorithms that have shown leading performance on relevant benchmark datasets.
+The objective of this thesis is to explore the speed-up in execution time of Learning to Rank algorithms through parallelisation using the MapReduce framework.
+This work focuses on those Learning to Rank algorithms that have shown leading performance on relevant benchmark datasets.
 This thesis addresses the following research questions:
 \begin{description}
-\item[RQ1] What are the best performing Learning-to-Rank algorithms in terms of accuracy on relevant benchmark datasets?\\
+\item[RQ1] What are the best performing Learning to Rank algorithms in terms of accuracy on relevant benchmark datasets?\\
 
-\item[RQ2] What is the speed-up of those Learning-to-Rank algorithms when executed using the MapReduce framework?\\
+\item[RQ2] What is the speed-up of those Learning to Rank algorithms when executed using the MapReduce framework?\\
 Where the definition of \emph{relative speed-up} is used for speed-up \cite{Sun1991}:\\
@@
 \subsection{Methodology for Research Question I}
-A literature study will be performed to get insight in relevant existing techniques for large scale Learning-to-Rank. The literature study will be performed by using the following query:\\
+A literature study will be performed to get insight in relevant existing techniques for large scale Learning to Rank. The literature study will be performed by using the following query:\\
 
@@
 
-The query incorporates different ways of writing of Learning-to-Rank, with and without hyphens, and the synonymous term \emph{machine learned ranking} to increase search recall, i.e.
+The query incorporates different ways of writing of Learning to Rank, with and without hyphens, and the synonymous term \emph{machine learned ranking} to increase search recall, i.e.
 to make sure that no relevant studies are missed.
@@
 \label{ssec:rq2_methodology}
-To find an answer to the second research question, the Learning-to-Rank methods determined in the first research question will be implemented in the MapReduce framework and training time will be measured as a factor of the number of cluster nodes used to perform the computation.
-The HDInsight cloud-based MapReduce platform from Microsoft will be used to run the Learning-to-Rank algorithms on.
+To find an answer to the second research question, the Learning to Rank methods determined in the first research question will be implemented in the MapReduce framework and training time will be measured as a factor of the number of cluster nodes used to perform the computation.
+The HDInsight cloud-based MapReduce platform from Microsoft will be used to run the Learning to Rank algorithms on.
 HDInsight is based on the popular open source MapReduce implementation Hadoop\footnote{http://hadoop.apache.org/}.\\
@@
 \begin{description}
-\item[Chapter II: Background]{introduces the reader to the basic principles and recent work in the fields of Learning-to-Rank.}
-\item[Chapter III: Related Work]{concisely describes existing work in the field of parallel Learning-to-Rank.}
-\item[Chapter IV: Benchmark Results]{sketches the accuracy of existing Learning-to-Rank methods on several benchmark datasets and describes the selection of Learning-to-Rank methods for the parallelisation experiments.}
-\item[Chapter V: Selected Learning-to-Rank Methods]{describes the algorithms and details of the Learning-to-Rank methods selected in Chapter IV.}
-\item[Chapter VI: Implementation]{describes implementation details of the Learning-to-Rank algorithms in the Hadoop framework.}
-\item[Chapter VII: Results \& Discussion]{presents and discusses speed-up results for the implemented Learning-to-Rank methods.}
+\item[Chapter II: Background]{introduces the reader to the basic principles and recent work in the fields of Learning to Rank.}
+\item[Chapter III: Related Work]{concisely describes existing work in the field of parallel Learning to Rank.}
+\item[Chapter IV: Benchmark Results]{sketches the accuracy of existing Learning to Rank methods on several benchmark datasets and describes the selection of Learning to Rank methods for the parallelisation experiments.}
+\item[Chapter V: Selected Learning to Rank Methods]{describes the algorithms and details of the Learning to Rank methods selected in Chapter IV.}
+\item[Chapter VI: Implementation]{describes implementation details of the Learning to Rank algorithms in the Hadoop framework.}
+\item[Chapter VII: Results \& Discussion]{presents and discusses speed-up results for the implemented Learning to Rank methods.}
 \item[Chapter VIII: Conclusion]{summarizes the results and answers our research questions based on the results.

### 2014-09-23 report/Chapters/LTRMethods.tex
naar refactoring werk

@@
 \label{chap:ltr_methods}
-The learning algorithms of the well-performing Learning-to-Rank methods selected in Chapter \ref{chap:cross_comparison} are presented and explained in this Chapter.
-The Learning-to-Rank methods will be discussed in order of an increasing degree of certainty and decreasing ranking accuracy, as concluded in Chapter \ref{chap:cross_comparison}.
+The learning algorithms of the well-performing Learning to Rank methods selected in Chapter \ref{chap:cross_comparison} are presented and explained in this Chapter.
+The Learning to Rank methods will be discussed in order of an increasing degree of certainty and decreasing ranking accuracy, as concluded in Chapter \ref{chap:cross_comparison}.
 
@@
 SmoothRank enables direct optimisation of the non-convex and discontinuous evaluation measure by smoothing, that is, approximating the rank position.
-In this section, I apply the SmoothRank approximation method to the \ac{nDCG} evaluation measure for illustration, but the same procedure can be applied to \ac{MAP} or any other information retrieval measure.
+In this section, I apply the SmoothRank approximation method to the \ac{NDCG} evaluation measure for illustration, but the same procedure can be applied to \ac{MAP} or any other information retrieval measure.
 The smoothing function used in SmoothRank is based on the softmax activation function \cite{Bridle1990}, which is often used in neural networks.
@@
 Chapelle et al.
-\cite{Chapelle2010} apply this softmax function to the \ac{nDCG} formula and hereby introduce a soft version of indicator variable $h_{i,j}$ through the computation shown in Equation \ref{eq:soft_ndcg}.
+\cite{Chapelle2010} apply this softmax function to the \ac{NDCG} formula and hereby introduce a soft version of indicator variable $h_{i,j}$ through the computation shown in Equation \ref{eq:soft_NDCG}.
 
@@
 h_{i,j} = e^{-\frac{(f(x_i))-(f(x_{d(j)}))^2}{\sigma}}\Big/\sum\nolimits_{k=1}^{m}e^{-\frac{(f(x_k))-f(x_{d(j)})^2}{\sigma}}
-\label{eq:soft_ndcg}
+\label{eq:soft_NDCG}
 \end{equation}
 
-It can be shown that the derivative of the smoothed \ac{nDCG} version shown in Equation \ref{eq:soft_ndcg} and the smoothed versions of other \ac{IR} metrics can be calculated in $\mathcal{O}(m^2)$, which enable fast gradient descent optimisation.
+It can be shown that the derivative of the smoothed \ac{NDCG} version shown in Equation \ref{eq:soft_NDCG} and the smoothed versions of other \ac{IR} metrics can be calculated in $\mathcal{O}(m^2)$, which enable fast gradient descent optimisation.
 The optimisation step in SmoothRank uses the nonlinear conjugate gradient method  with Polak-Ribiere update \cite{Shewchuk1994}, which is a type of gradient descent method.
@@
 
-The choice of the smoothing parameter $\sigma$ in Equation \ref{eq:soft_ndcg} is important, because a too small value makes the function more non-smooth and therefore harder to optimise, while a too large value results in a optimisation function with optima that substantially differ from the true optimal rankings.
+The choice of the smoothing parameter $\sigma$ in Equation \ref{eq:soft_NDCG} is important, because a too small value makes the function more non-smooth and therefore harder to optimise, while a too large value results in a optimisation function with optima that substantially differ from the true optimal rankings.
 To deal with the problem of choosing the smoothing parameter, SmoothRank uses an annealing method where the optimisation procedure starts with a large $\sigma$ and iteratively reduces it by dividing it by two at each iteration.
@@
 \section{FenchelRank}
-FenchelRank \cite{Lai2013} is a ranking method that addresses the sparse Learning-to-Rank problem, which is the problem of learning a ranking function with only a few non-zero coefficients with respect to the input features.
+FenchelRank \cite{Lai2013} is a ranking method that addresses the sparse Learning to Rank problem, which is the problem of learning a ranking function with only a few non-zero coefficients with respect to the input features.
 FenchelRank is based on the theory of Fenchel Duality \cite{Rifkin2007} and uses a generic algorithm framework proposed by Shalev-Shwartz and Singer \cite{Shalev-Shwartz2010}.
@@
 Lai et al.
-\cite{Lai2013c} observed that existing feature selection methods in Learning-to-Rank all follow a two-stage paradigm consisting of a first stage of selecting a subset of features from the original features, followed by a second stage where a ranking model learnt based on the selected features.
+\cite{Lai2013c} observed that existing feature selection methods in Learning to Rank all follow a two-stage paradigm consisting of a first stage of selecting a subset of features from the original features, followed by a second stage where a ranking model learnt based on the selected features.
 Lai et al.
@@
 \section{LRUF}
-LRUF \cite{Torkestani2012b} is Learning-to-Rank method based on the theory of learning automata.
+LRUF \cite{Torkestani2012b} is Learning to Rank method based on the theory of learning automata.
 Learning automata are adaptive decision-making units that learn the optimal set of actions through repeated interactions with its environment.

### 2014-09-23 report/Chapters/RelatedWork.tex
naar refactoring werk

@@
 
-Research in scaling up the training phase of Learning-to-Rank models can be categorised according to the approach in scaling up.
-Figure \ref{fig:related_work_categories} illustrates the categories of scalable training approaches in Learning-to-Rank.
+Research in scaling up the training phase of Learning to Rank models can be categorised according to the approach in scaling up.
+Figure \ref{fig:related_work_categories} illustrates the categories of scalable training approaches in Learning to Rank.
 The numbers in Figure \ref{fig:related_work_categories} correspond to the sections that describe the related work belonging to these categories.
@@
 \includegraphics{gfx/related_work_categories}
-\caption{Categories in research on large scale training of  Learning-to-Rank models}
+\caption{Categorisation of research on large scale training of  Learning to Rank models}
 \label{fig:related_work_categories}
@@
 
-\section{Low computational complexity Learning-to-Rank}
+\section{Low computational complexity Learning to Rank}
 \label{sec:related_work_low_complexity}
-One approach for handling large volumes of training data for Learning-to-Rank is through design of low time complexity Learning-to-Rank methods.
+One approach for handling large volumes of training data for Learning to Rank is through design of low time complexity Learning to Rank methods.
 Pahikkala et al.
@@
 Lee and Lin \cite{Lee2014} observed that linear kernel Rank\ac{SVM}s are inferior in accuracy compared to nonlinear kernel Rank\ac{SVM}s and \ac{GBDT}s and are mainly useful to quickly produce a baseline model.
-Details of the lower time complexity version of the linear kernel Rank\ac{SVM} will not be discussed as it is shown to be an inferior Learning-to-Rank method in terms of accuracy.\\
+Details of the lower time complexity version of the linear kernel Rank\ac{SVM} will not be discussed as it is shown to be an inferior Learning to Rank method in terms of accuracy.\\
 
-Learning-to-Rank methods that are specifically designed for their low computational complexity, like RankRLS and the linear kernel Rank\ac{SVM} methods described in this section, are generally not among the top achieving models in terms of accuracy.
+Learning to Rank methods that are specifically designed for their low computational complexity, like RankRLS and the linear kernel Rank\ac{SVM} methods described in this section, are generally not among the top achieving models in terms of accuracy.
 From results on benchmarks and competitions it can be observed that the best generalisation accuracy are often more complex ones.
-This makes low time complexity models as a solution for large scale Learning-to-Rank less applicable and increases the relevance of the search for efficient training of more complex Learning-to-Rank models.\\
+This makes low time complexity models as a solution for large scale Learning to Rank less applicable and increases the relevance of the search for efficient training of more complex Learning to Rank models.\\
 
-\section{Distributed hyperparameter tuning of Learning-to-Rank models}
-Hyperparameter optimisation is the task of selecting the combination of hyperparameters such that the Learning-to-Rank model shows optimal generalisation accuracy.
+\section{Distributed hyperparameter tuning of Learning to Rank models}
+Hyperparameter optimisation is the task of selecting the combination of hyperparameters such that the Learning to Rank model shows optimal generalisation accuracy.
 Ganjisaffar et al.
-\cite{Ganjisaffar2011, Ganjisaffar2011b} observed that long training times are often a result of hyperparameter optimisation, because it results in training multiple Learning-to-Rank models.
+\cite{Ganjisaffar2011, Ganjisaffar2011b} observed that long training times are often a result of hyperparameter optimisation, because it results in training multiple Learning to Rank models.
 Grid search is the \emph{de facto} standard of hyperparameter optimisation and is simply an exhaustive search through a manually specified subset of hyperparameter combinations.
 The authors show how to perform parallel grid search on MapReduce clusters, which is easy because grid search is an embarrassingly parallel method as hyperparameter combinations are mutually independent.
-They apply their grid search on MapReduce approach in a Learning-to-Rank setting to train a LambdaMART \cite{Wu2008} ranking model, which uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
+They apply their grid search on MapReduce approach in a Learning to Rank setting to train a LambdaMART \cite{Wu2008} ranking model, which uses the Gradient Boosting \cite{Friedman2002} ensemble method combined with regression tree weak learners.
 Experiments showed that the solution scales linearly in number of hyperparameter combinations.
@@
 
-Notice that methods described in this section train multiple Learning-to-Rank models at the same time to find the optimal set of parameters for a model, but that the Learning-to-Rank models itself are still trained sequentially.
-In the next sections we will present literature focusing on training Learning-to-Rank models in such a way that steps in this training process can be executed simultaneously.\\
+Notice that methods described in this section train multiple Learning to Rank models at the same time to find the optimal set of parameters for a model, but that the Learning to Rank models itself are still trained sequentially.
+In the next sections we will present literature focusing on training Learning to Rank models in such a way that steps in this training process can be executed simultaneously.\\
 
-\section{Hardware accelerated Learning-to-Rank}
+\section{Hardware accelerated Learning to Rank}
 Hardware accelerators are special purpose processors designed to speed up compute-intensive tasks.
@@
 In general, \ac{FPGA}s provide better performance while \ac{GPU}s tend to be easier to program \cite{Che2008}.
-Some research has been done in parallelising Learning-to-Rank methods using hardware accelerators.
+Some research has been done in parallelising Learning to Rank methods using hardware accelerators.
 
-\subsection{FPGA-based parallel Learning-to-Rank}
+\subsection{FPGA-based parallel Learning to Rank}
 Yan et al.
-\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run, the Neural Network based LambdaRank Learning-to-Rank algorithm.
+\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run, the Neural Network based LambdaRank Learning to Rank algorithm.
 This architecture achieved a 29.3X speed-up compared to the software implementation, when evaluated on data from a commercial search engine.
-The exploration of \ac{FPGA} for Learning-to-Rank showed additional benefits other than the speed-up originally aimed for.
+The exploration of \ac{FPGA} for Learning to Rank showed additional benefits other than the speed-up originally aimed for.
 In their latest publication \cite{Yan2012} the \ac{FPGA} based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.\\
@@
 Xu et al.
-\cite{Xu2009} state that RankBoost is a Learning-to-Rank method that is not widely used in practice because of its long training time.
+\cite{Xu2009} state that RankBoost is a Learning to Rank method that is not widely used in practice because of its long training time.
 Experiments on MSN search engine data showed the implementation on a \ac{FPGA} with \ac{SIMD} architecture to be 170.6x faster than the original software implementation \cite{Xu2007b}.
@@
 
-\subsection{GPGPU for parallel Learning-to-Rank}
+\subsection{GPGPU for parallel Learning to Rank}
 Wang et al.
@@
 \cite{DeSousa2012} proposed a \ac{GPGPU} approach to improve both training time and query evaluation through \ac{GPU} use.
-An association rule based Learning-to-Rank approach, proposed by Veloso et al.
+An association rule based Learning to Rank approach, proposed by Veloso et al.
 \cite{Veloso2008}, has been implemented using the \ac{GPU} in such a way that the set of rules van be computed simultaneously for each document.
@@
 
-\section{Parallel execution of Learning-to-Rank algorithm steps}
-Some research focused on parallelising the steps Learning-to-Rank algorithms that can be characterised as strong learners.
+\section{Parallel execution of Learning to Rank algorithm steps}
+Some research focused on parallelising the steps Learning to Rank algorithms that can be characterised as strong learners.
 
 Tyree et al.
-\cite{Tyree2011} described a way of parallelising \ac{GBDT} models for Learning-to-Rank where the boosting step is still executed sequentially, but instead the construction of the regression trees themselves is parallelised.
+\cite{Tyree2011} described a way of parallelising \ac{GBDT} models for Learning to Rank where the boosting step is still executed sequentially, but instead the construction of the regression trees themselves is parallelised.
 The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}, which are build layer-by-layer.
@@
 Shukla et al.
-\cite{Shukla2012} explored the parallelisation of the well-known ListNet Learning-to-Rank method using Spark.
+\cite{Shukla2012} explored the parallelisation of the well-known ListNet Learning to Rank method using Spark.
 Spark is a parallel computing model that is designed for cyclic data flows which makes it more suitable for iterative algorithms.
@@
 \section{Parallelisable search heuristics for Listwise ranking}
-Direct minimisation of ranking metrics is a hard problem due to the non-continuous, non-differentiable and non-convex nature of the \ac{nDCG}, \ac{ERR} and \ac{MAP} evaluation metrics.
+Direct minimisation of ranking metrics is a hard problem due to the non-continuous, non-differentiable and non-convex nature of the \ac{NDCG}, \ac{ERR} and \ac{MAP} evaluation metrics.
 This optimisation problem is generally addressed either by replacing the ranking metric with a convex surrogate, or by heuristic optimisation methods such as Simulated Annealing or a \ac{EA}.
@@
 \cite{Wang2011a,Wang2011b} proposed a parallel evolutionary algorithm based on \ac{CC}, a type of evolutionary algorithm.
-The \ac{CC} algorithm is capable of directly optimizing non-differentiable functions, as \ac{nDCG}, in contrary to many optimization algorithms.
+The \ac{CC} algorithm is capable of directly optimizing non-differentiable functions, as \ac{NDCG}, in contrary to many optimization algorithms.
 the divide-and-conquer nature of the \ac{CC} algorithm enables parallelisation.
@@
 Two reasons have been identified for not achieving linear scale-up with CCRank: 1) parallel execution is suspended after each generation to perform combination in order to produce the candidate solution, 2) Combination has to wait until all parallel tasks have finished, which may spend different running time.
-\subsection{nDCG-Annealing}
+\subsection{NDCG-Annealing}
 Karimzadeghan et al.
 \cite{Karimzadehgan2011} proposed a method using Simulated Annealing along with the Simplex method for its parameter search.
-This method directly optimises the often non-differentiable Learning-to-Rank evaluation metrics like \ac{nDCG} and \ac{MAP}.
+This method directly optimises the often non-differentiable Learning to Rank evaluation metrics like \ac{NDCG} and \ac{MAP}.
 The authors successfully parallelised their method in the MapReduce paradigm using Hadoop.
@@
 \subsection{Alternating Direction Method of Multipliers}
-Duh et al. \cite{Duh2011} proposed the use of \ac{ADMM} for the Learning-to-Rank task. \ac{ADMM} is a general optimization method that solves problems of the form
+Duh et al. \cite{Duh2011} proposed the use of \ac{ADMM} for the Learning to Rank task. \ac{ADMM} is a general optimization method that solves problems of the form
 \begin{alignat*}{2}
@@
 
-Another \ac{ADMM} Learning-to-Rank based approach was proposed by Boyd et al.
+Another \ac{ADMM} Learning to Rank based approach was proposed by Boyd et al.
 \cite{Boyd2012}.
-They implemented an \ac{ADMM} based Learning-to-Rank method in Pregel \cite{Malewicz2010}, a parallel computation framework for graph computations.
+They implemented an \ac{ADMM} based Learning to Rank method in Pregel \cite{Malewicz2010}, a parallel computation framework for graph computations.
 No experimental results on parallelisation speed-up were reported on this Pregel based approach.
@@
 Acharyya et al.
-\cite{Acharyya2012, Acharyya2013} proposed a Learning-to-Rank method searches for an order preserving transformation (\emph{monotone retargeting}) of the target scores that is easier for a regressor to fit.
+\cite{Acharyya2012, Acharyya2013} proposed a Learning to Rank method searches for an order preserving transformation (\emph{monotone retargeting}) of the target scores that is easier for a regressor to fit.
 This approach is based on the observation that it is not necessary to fit scores exactly, since the evaluation is dependent on the order and not on the pointwise predictions themselves.\\
@@
 Acharyya et al.
-\cite{Acharyya2012, Acharyya2013} defined a parallel algorithm that optimises a Bregman divergence function as surrogate of \ac{nDCG} that is claimed to be well suited for implementation of a \ac{GPGPU}.
+\cite{Acharyya2012, Acharyya2013} defined a parallel algorithm that optimises a Bregman divergence function as surrogate of \ac{NDCG} that is claimed to be well suited for implementation of a \ac{GPGPU}.
 No experiments on speed-up have been performed.
-\subsection{Parallel robust Learning-to-Rank}
+\subsection{Parallel robust Learning to Rank}
 Robust learning \cite{Huber1981} is defined as the task to lean a model in the presence of outliers.
 Yun et al.
-described a \cite{Yun2014} robust Learning-to-Rank model called RoBiRank that has the additional advantage that it is executable in parallel.
+described a \cite{Yun2014} robust Learning to Rank model called RoBiRank that has the additional advantage that it is executable in parallel.
 RoBiRank uses parameter expansion to linearise a surrogate loss function, after which the elements of the linear combination can be divided over the available nodes.
@@
 Long et al.
-\cite{Long2012} described special case of their pairwise cross-domain factor Learning-to-Rank method using distributed optimization of \ac{SGD} based on Hadoop MapReduce.
+\cite{Long2012} described special case of their pairwise cross-domain factor Learning to Rank method using distributed optimization of \ac{SGD} based on Hadoop MapReduce.
 Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce based method described by  Zinkevich et al.
@@
 
-\section{Ensemble learning for parallel Learning-to-Rank}
+\section{Ensemble learning for parallel Learning to Rank}
 Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure called boosting \cite{Schapire1990}.
@@
 Ensemble methods can be used to parallelise learning methods by training the weak models in the ensemble on different nodes.
-Parallelisation of the training phase of Learning-to-Rank models is often achieved through ensemble learning, mainly boosting, and combined with decision trees, jointly called \acf{GBDT}.
-\ac{GBDT}s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
+Parallelisation of the training phase of Learning to Rank models is often achieved through ensemble learning, mainly boosting, and combined with decision trees, jointly called \acf{GBDT}.
+\ac{GBDT}s are shown to be able to achieve good accuracy in a Learning to Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
 
@@
 \subsection{Randomisation}
-Most ensemble learning based Learning-to-Rank methods are based on boosting.
-Although boosting have shown to be an effective method in Learning-to-Rank, it has the drawback that the models in the ensemble are not completely mutually independent, which makes the process harder to parallelise.
+Most ensemble learning based Learning to Rank methods are based on boosting.
+Although boosting have shown to be an effective method in Learning to Rank, it has the drawback that the models in the ensemble are not completely mutually independent, which makes the process harder to parallelise.
 Geurts and Louppe \cite{Geurts2011} proposed a tree based ensemble method in which the trees are built using random subsets of the available features.

### 2014-09-23 report/Chapters/Results.tex
naar refactoring werk

@@
 \subsection{Ranking accuracy}
-To validate our ListNet cluster implementation we run experiments on the LETOR 3.0 OHSUMED and the LETOR 4.0 datasets MQ2007 and MQ2008 and compare their performance with the RankLib implementation of ListNet and with the official evaluation results in terms of \ac{nDCG}@10.
-Hadoop job scheduling overhead is large when processing relatively small datasets (as we will explain in more detail in section \ref{ssec:speedup}), therefore \ac{nDCG}@10 experiments are limited to five training iterations and are run only on one of the five folds.
-Because of the stochastic nature of the ListNet ranking procedure, which is a result of random weight initialisation prior to the first training iteration, the \ac{nDCG}@10 performance is not guaranteed to be identical between multiple runs on the same input data.
-To account for this non-determinism of ListNet, we repeat each experiment five times and report the average as well as the standard deviation of the resulting rankings on the testing fold in terms of \ac{nDCG}@10.\\
+To validate our ListNet cluster implementation we run experiments on the LETOR 3.0 OHSUMED and the LETOR 4.0 datasets MQ2007 and MQ2008 and compare their performance with the RankLib implementation of ListNet and with the official evaluation results in terms of \ac{NDCG}@10.
+Hadoop job scheduling overhead is large when processing relatively small datasets (as we will explain in more detail in section \ref{ssec:speedup}), therefore \ac{NDCG}@10 experiments are limited to five training iterations and are run only on one of the five folds.
+Because of the stochastic nature of the ListNet ranking procedure, which is a result of random weight initialisation prior to the first training iteration, the \ac{NDCG}@10 performance is not guaranteed to be identical between multiple runs on the same input data.
+To account for this non-determinism of ListNet, we repeat each experiment five times and report the average as well as the standard deviation of the resulting rankings on the testing fold in terms of \ac{NDCG}@10.\\
 
-Table \ref{tbl:accuracy_comparison} shows the mean and the standard deviation of the ranking accuracy in \ac{nDCG}@10 on the first data fold for our own cluster implementation of ListNet, the RankLib implementation of ListNet, as well as the official LETOR evaluation results of ListNet as reported on the LETOR 3.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and LETOR 4.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} websites.
+Table \ref{tbl:accuracy_comparison} shows the mean and the standard deviation of the ranking accuracy in \ac{NDCG}@10 on the first data fold for our own cluster implementation of ListNet, the RankLib implementation of ListNet, as well as the official LETOR evaluation results of ListNet as reported on the LETOR 3.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and LETOR 4.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} websites.
 Note that the official LETOR evaluation results are structurally higher compared to the results obtained with RankLib and our cluster implementation, which can be attributed to the fact that the official evaluation runs use as many ListNet iterations as needed for convergence on the validation fold, while we limited our experiment to five iterations.
@@
 \end{tabular}
-\caption{\acs{nDCG}@10 performance on the testing set on the first fold}
+\caption{\acs{NDCG}@10 performance on the testing set on the first fold}
 \label{tbl:accuracy_comparison}
@@
 \end{figure}
+
+\section{SmoothRank}

### 2014-09-23 report/Chapters/Results.tex
throughput -> processing speed

@@
 \includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.8]{gfx/time_single_logx.pdf}
-\caption{Processing time of a single ListNet training iteration on a logarithmic axis}
+\caption{Processing time of a single ListNet training iteration on a logarithmic data size axis}
 \label{fig:listnet_train_time_log}
@@
 
-Figure \ref{fig:listnet_throughput} shows the throughput in bytes per second.
+Figure \ref{fig:listnet_processing_speed} shows the processing speed in bytes per second.
 Our observation that RankLib ListNet performs very slow for data sets that do not fit in physical memory and virtual memory is needed is very notable in this graph.
-Furthermore, this graph shows how both increase in number in data nodes in a cluster and increase in input data size result in an increase in throughput.
-The sub-linear growth of throughput as a factor of input data size can again be explained as a result of the non-parallelisable Hadoop job scheduling part of the operation.\\
+Furthermore, this graph shows how both increase in number in data nodes in a cluster and increase in input data size result in an increase in processing speed.
+The sub-linear growth of processing speed as a factor of input data size can again be explained as a result of the non-parallelisable Hadoop job scheduling part of the operation.\\
 
@@
 \includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.7]{gfx/speedup_faceted.pdf}
-\caption{Processing time as a function of data nodes in a cluster}
+\caption{Processing time of a single ListNet training iteration as a function of the number of data nodes in a cluster}
 \label{fig:speedup_train_time}
@@
 \centering
-\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.8]{gfx/throughput_single_logx.pdf}
-\caption{Throughput of a ListNet training iteration}
-\label{fig:listnet_throughput}
+\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.8]{gfx/processing_speed_single_logx.pdf}
+\caption{Processing speed of a a single ListNet training iteration on various data sets}
+\label{fig:listnet_processing_speed}
 \end{figure}

### 2014-09-23 report/Chapters/AppendixA.tex
naar refactoring werk
dataset -> data set

@@
-\chapter{Appendix A: LETOR Feature set}
+\chapter{LETOR Feature set}
 \label{chap:appendix_a}
@@
 }
-\caption{Features of the LETOR OHSUMED dataset, obtained from Qin et al \cite{Qin2010}}
+\caption{Features of the LETOR OHSUMED data set, obtained from Qin et al \cite{Qin2010}}
 \label{tbl:features_ohsumed}
@@
 }
-\caption{Features of the LETOR .GOV dataset, obtained from Qin et al \cite{Qin2010}}
+\caption{Features of the LETOR .GOV data set, obtained from Qin et al \cite{Qin2010}}
 \label{tbl:features_gov}

### 2014-09-23 report/Chapters/BenchmarkResults.tex
dataset -> data set

@@
 \label{chap:benchmark_results}
-Benchmark datasets enable fair comparisons of ranking models over a fixed set of documents and queries.
-The approach of using fixed benchmark datasets to compare the performance of Information Retrieval systems under equal circumstances has been the standard evaluation methodology in the Information Retrieval field since the release of the Cranfield collection \cite{Cleverdon1966}.\\
+Benchmark data sets enable fair comparisons of ranking models over a fixed set of documents and queries.
+The approach of using fixed benchmark data sets to compare the performance of Information Retrieval systems under equal circumstances has been the standard evaluation methodology in the Information Retrieval field since the release of the Cranfield collection \cite{Cleverdon1966}.\\
 
 This chapter describes benchmark characteristics, e.g.
-collection size and features, of Learning to Rank benchmark collections and datasets and gives an overview of the performance of baseline algorithms on these collections and datasets.\\
+collection size and features, of Learning to Rank benchmark collections and data sets and gives an overview of the performance of baseline algorithms on these collections and data sets.\\
 
@@
 \section{Yahoo! Learning to Rank Challenge}
-Yahoo's observation that all existing benchmark datasets were too small to draw reliable conclusions prompted Yahoo to release two internal datasets from Yahoo!
+Yahoo's observation that all existing benchmark data sets were too small to draw reliable conclusions prompted Yahoo to release two internal data sets from Yahoo!
 search.
-Datasets used at commercial search engines are many times larger than available benchmark datasets.
+Data sets used at commercial search engines are many times larger than available benchmark data sets.
 Yahoo!
@@
 The Yahoo!
-Learning to Rank Challenge \cite{Chapelle2011a} is a public Learning to Rank competition which took place from March to May 2010, with the goal to promote the datasets and encourage the research community to develop new Learning to Rank algorithms.\\
+Learning to Rank Challenge \cite{Chapelle2011a} is a public Learning to Rank competition which took place from March to May 2010, with the goal to promote the data sets and encourage the research community to develop new Learning to Rank algorithms.\\
 
 The Yahoo!
-Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning to Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country.
-For this experiment, I will only look at the standard Learning to Rank dataset, because transfer learning is a separate research area that is not included in this thesis.\\
+Learning to Rank Challenge consists of two tracks that uses the two data sets respectively: a standard Learning to Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country.
+For this experiment, I will only look at the standard Learning to Rank data set, because transfer learning is a separate research area that is not included in this thesis.\\
 \begin{table}[!h]
@@
 \end{tabular}
-\caption{Yahoo! Learning to Rank Challenge dataset characteristics, as described in the challenge overview paper \cite{Chapelle2011a}}
+\caption{Yahoo! Learning to Rank Challenge data set characteristics, as described in the overview paper \cite{Chapelle2011a}}
 \label{tab:yahoo_characteristics}
@@
 The test set performance is used to create the final standings and is only measured after the competition has ended to avoid overfitting on the test set.
-The large number of documents, queries and features compared to other benchmark datasets makes the Yahoo!
-Learning to Rank Challenge dataset interesting.
+The large number of documents, queries and features compared to other benchmark data sets makes the Yahoo!
+Learning to Rank Challenge data set interesting.
 Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo.
@@
 \subsection{LETOR 2.0}
-LETOR 2.0 consists of three datasets: the OHSUMED dataset and two datasets of the on the .gov collection.
+LETOR 2.0 consists of three data sets: the OHSUMED data set and two data sets of the on the .gov collection.
 The OHSUMED collection is a subset of the medical publication database MEDLINE and contains medical publications from 270 journals that were published between 1987 and 1991.
@@
 \end{tabular}
-\caption{\acs{nDCG}@10 results of baseline methods on LETOR 2.0}
+\caption{\acs{nDCG}@10 results of the baseline methods on LETOR 2.0}
 \label{tab:letor2.0_results}
 \end{table}
-Table \ref{tab:letor2.0_results} shows the performance of the baseline methods on the datasets in the LETOR 2.0 collection.
-ListNet can be regarded as the winner of the comparison as it outperformed the other methods on two of the three datasets and ranked third on the third dataset.
-\acs{MHR} was only evaluated on the OHSUMED dataset, on which it performed second best, after ListNet.
+Table \ref{tab:letor2.0_results} shows the performance of the baseline methods on the data sets in the LETOR 2.0 collection.
+ListNet can be regarded as the winner of the comparison as it outperformed the other methods on two of the three data sets and ranked third on the third data set.
+\acs{MHR} was only evaluated on the OHSUMED data set, on which it performed second best, after ListNet.
 
@@
 With query sets available from both the \ac{TREC} 2003 and 2004 conferences there are six query sets in total: TD2003, TD2004, NP2003, NP2004, HP2003 and HP2004.
-It is noteworthy that the OHSUMED, TD2003 and TD2004 datasets of LETOR 2.0 and LETOR 3.0 are not identical and therefore not comparable due to differences in sampling method.\\
+It is noteworthy that the OHSUMED, TD2003 and TD2004 data sets of LETOR 2.0 and LETOR 3.0 are not identical and therefore not comparable due to differences in sampling method.\\
 
 The evaluation metrics used are \ac{nDCG} and \ac{MAP}.
-The \emph{winning number} metric is defined as the number of other algorithms that it can beat over all of the seven datasets (six .gov sets + OHSUMED).
+The \emph{winning number} metric is defined as the number of other algorithms that it can beat over all of the seven data sets (six .gov sets + OHSUMED).
 The LETOR organisation implemented and evaluated several well-known Learning to Rank algorithms themselves and in addition gathers and publications and results of new algorithms evaluated on the LETOR benchmark.
@@
 \includegraphics[scale=0.30]{gfx/ndcg_winning_number}
-\caption{Comparison of ranking accuracy across the seven datasets in LETOR by \acs{NDCG}, obtained from Qin et al. \cite{Qin2010}}
+\caption{Comparison of ranking accuracy across the seven data sets in LETOR by \acs{NDCG}, obtained from Qin et al. \cite{Qin2010}}
 \label{fig:ndcg_winning_number}
@@
 \includegraphics[scale=0.30]{gfx/map_winning_number}
-\caption{Comparison across the seven datasets in LETOR by \acs{MAP}, obtained from Qin et al. \cite{Qin2010}}
+\caption{Comparison across the seven data sets in LETOR by \acs{MAP}, obtained from Qin et al. \cite{Qin2010}}
 \label{fig:map_winning_number}
@@
 \cite{Qin2010}, as listed on the LETOR website\footnotemark[2] by comparing their performance with the ListNet baseline.
-We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven datasets in \ac{nDCG} value.
+We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven data sets in \ac{nDCG} value.
 Note that this does not necessarily imply that these methods would have scored a higher \ac{nDCG} winning number than ListNet.
-Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR datasets in terms of \ac{nDCG}@10 and \ac{MAP}.
+Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR data sets in terms of \ac{nDCG}@10 and \ac{MAP}.
 \begin{table}[!h]
@@
 OHSUMED & 0.4436 & 0.4504 & 0.4523 & 0.4568 & 0.441 \\ 
-\# winning datasets & 2 & 2 & 1 & 3 & - \\ 
+\# winning data sets & 2 & 2 & 1 & 3 & - \\ 
 \end{tabular}
@@
 \subsection{LETOR 4.0}
-The LETOR 4.0 benchmark collection\footnote{http://http://research.microsoft.com/en-us/um/beijing/projects/letor/} consists of the Gov-2 document collection and two query datasets from the Million Query Track at \ac{TREC} 2007 (MQ2007) and \ac{TREC} 2008 (MQ2008).
+The LETOR 4.0 benchmark collection\footnote{http://http://research.microsoft.com/en-us/um/beijing/projects/letor/} consists of the Gov-2 document collection and two query data sets from the Million Query Track at \ac{TREC} 2007 (MQ2007) and \ac{TREC} 2008 (MQ2008).
 LETOR 4.0 consists of a semi-supervised ranking task, a rank aggregation task and a listwise ranking task next to the supervised ranking task.
 Table \ref{tab:letor4_characteristics} shows the collection characteristics in number of queries, documents and features for LETOR 4.0.
-Evaluation on this collection is performed using a five-fold cross-validation, where partitioning of the data into folds was performed beforehand by the creators of the MQ2007 and MQ2008 datasets.
-Documents in the dataset were
+Evaluation on this collection is performed using a five-fold cross-validation, where partitioning of the data into folds was performed beforehand by the creators of the MQ2007 and MQ2008 data sets.
+Documents in the data set were
 
@@
 \begin{tabular}{l|ll}
- Dataset & MQ2007 & MQ2008 \\ 
+ Data set & MQ2007 & MQ2008 \\ 
  \hline
@@
 \subsubsection{Results}
-Rank\ac{SVM}-Struct, ListNet, AdaRank-\ac{nDCG}, AdaRank-\ac{MAP} and RankBoost were used as baseline models on the LETOR 4.0 dataset and were implemented and evaluated by the publishers of LETOR 4.0.
+Rank\ac{SVM}-Struct, ListNet, AdaRank-\ac{nDCG}, AdaRank-\ac{MAP} and RankBoost were used as baseline models on the LETOR 4.0 data set and were implemented and evaluated by the publishers of LETOR 4.0.
 Table \ref{tab:letor4_baseline_results} shows the performance of those baseline models on the LETOR 4.0 benchmark collection.\\
@@
 \begin{tabular}{l|lll}
-Model & Dataset & Mean \ac{nDCG} & \ac{nDCG}@10 \\ 
+Model & Data set & Mean \ac{nDCG} & \ac{nDCG}@10 \\ 
 \hline
@@
 \end{tabular}
-\caption{Comparison of LETOR 4.0 baseline models}
+\caption{Comparison of the LETOR 4.0 baseline models}
 \label{tab:letor4_baseline_results}
@@
 
-BoostedTree model \cite{Kocsis2013} showed an \ac{nDCG} of 0.5071 on the MQ2007 dataset and thereby beat all baseline models.
+BoostedTree model \cite{Kocsis2013} showed an \ac{nDCG} of 0.5071 on the MQ2007 data set and thereby beat all baseline models.
 
-\section{Other datasets}
-LETOR and the Yahoo! Learning to Rank Challenge datasets are the most used benchmark collections in the field. Several other benchmark collections for the Learning to Rank task have been proposed.
+\section{Other data sets}
+LETOR and the Yahoo! Learning to Rank Challenge data sets are the most used benchmark collections in the field. Several other benchmark collections for the Learning to Rank task have been proposed.
 
 \subsection{MSLR-WEB10/30k}
-MSLR-WEB30k and the smaller subset MSLR WEB10k are two large datasets with 30,000 and 10,000 queries respectively.
+MSLR-WEB30k and the smaller subset MSLR WEB10k are two large data sets with 30,000 and 10,000 queries respectively.
 The data was obtained from the Microsoft Bing\footnote{http://www.bing.com/} commercial web search engine.
 A feature list and feature descriptions are available.
-The dataset only includes features that are commonly used in the research community, as proprietary features were excluded from the dataset.
-Microsoft published these datasets as unofficial successor to LETOR in June 2010, but no baseline evaluations were described by the MSLR-WEB10/30k team.
-The Microsoft Research website from which the dataset is obtainable\footnote{http://research.microsoft.com/en-us/projects/mslr/download.aspx} also offers an evaluation script for \ac{nDCG} and \ac{MAP}.
+The data set only includes features that are commonly used in the research community, as proprietary features were excluded from the data set.
+Microsoft published these data sets as unofficial successor to LETOR in June 2010, but no baseline evaluations were described by the MSLR-WEB10/30k team.
+The Microsoft Research website from which the data set is obtainable\footnote{http://research.microsoft.com/en-us/projects/mslr/download.aspx} also offers an evaluation script for \ac{nDCG} and \ac{MAP}.
 The presence of an official evaluation script enables fair comparison of evaluation results from other researchers comparable.
@@
 \subsection{WCL2R}
-The WCL2R collection, released by Alc{\^a}ntara et al \cite{Alcantara2010}, contains of two datasets from the Chilean search engine TodoCL\footnote{www.todocl.cl}.
-Both datasets contain approximately 3 million documents.
+The WCL2R collection, released by Alc{\^a}ntara et al \cite{Alcantara2010}, contains of two data sets from the Chilean search engine TodoCL\footnote{www.todocl.cl}.
+Both data sets contain approximately 3 million documents.
 WCL2R is the only benchmark collection known that provides click-through data on user level.
@@
 \end{tabular}
-\caption{\acs{nDCG} results of baseline methods on the WCL2R collection, obtained from \cite{Alcantara2010}}
+\caption{\acs{nDCG} results of the baseline methods on the WCL2R collection, obtained from \cite{Alcantara2010}}
 \label{tab:results_WCL2R}
@@
 Pass et al.
-\cite{Pass2006} describe an America Online Learning to Rank dataset which they published in 2006.
-This dataset is unique in that it is the only large commercial web search dataset that contains user session and user behaviour information.
-The dataset contains 20 million search keywords for over 650,000 users over a three month time period.
-AOL later realised the publication of the dataset to be a mistake, as personally identifiable information turned out to be presented in some of the queries\footnote{http://select.nytimes.com/gst/abstract.html?res=F10612FC345B0C7A8CDDA10894DE404482}.
+\cite{Pass2006} describe an America Online Learning to Rank data set which they published in 2006.
+This data set is unique in that it is the only large commercial web search data set that contains user session and user behaviour information.
+The data set contains 20 million search keywords for over 650,000 users over a three month time period.
+AOL later realised the publication of the data set to be a mistake, as personally identifiable information turned out to be presented in some of the queries\footnote{http://select.nytimes.com/gst/abstract.html?res=F10612FC345B0C7A8CDDA10894DE404482}.
 AOL acknowledged the mistake of publishing the data and removed the data from its website.
 The removal of the data was too late, as the data was already downloadable from several mirror sites\footnote{http://gregsadetsky.com/aol-data/}.
-This controversial AOL dataset is to date still used for Learning to Rank research.
+This controversial AOL data set is to date still used for Learning to Rank research.
 No official baseline evaluation is provided by the AOL team.

### 2014-09-23 report/Chapters/CrossBenchmarkComparison.tex
dataset -> data set

@@
 \label{chap:cross_comparison}
-As we have seen in Chapter~\ref{chap:benchmark_results}, the evaluation of Learning to Rank methods is spread over several benchmark datasets.
+As we have seen in Chapter~\ref{chap:benchmark_results}, the evaluation of Learning to Rank methods is spread over several benchmark data sets.
 However, as the Learning to Rank methods evaluated differs between benchmarks, no single benchmark comparison can be regarded as a conclusive argument on which Learning to Rank method is most accurate.\\
 
-Several studies make a small start in considering Learning to Rank methods performance over multiple benchmark datasets.
+Several studies make a small start in considering Learning to Rank methods performance over multiple benchmark data sets.
 Gomes et al.
@@
 Busa-Fekete et al.
-\cite{Busa-Fekete2013} compared the accuracy of a small set of models over the LETOR 4.0 datasets, both MSLR datasets, both Yahoo!
-Learning to Rank Challenge datasets and the OHSUMED dataset from LETOR 3.0.
+\cite{Busa-Fekete2013} compared the accuracy of a small set of models over the LETOR 4.0 data sets, both MSLR data sets, both Yahoo!
+Learning to Rank Challenge data sets and the OHSUMED dataset from LETOR 3.0.
 To my knowledge, no structured meta-analysis on ranking accuracy has been conducted where evaluation results on several benchmark collections are taken into account.
-With a meta-analysis we will compare the performance of Learning to Rank methods across the Learning to Rank benchmark datasets described in foregoing sections.
+With a meta-analysis we will compare the performance of Learning to Rank methods across the Learning to Rank benchmark data sets described in foregoing sections.
 
@@
 \label{sec:collecting_evaluation_results}
-With a literature review we will collect evaluation results on the datasets/collections. The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
+With a literature review we will collect evaluation results on the data sets / collections.
+The following list presents an overview of the benchmark collections taken into account in the meta-analysis:
 \begin{itemize}
@@
 \begin{tabular}{l|l}
-Benchmark & Google scholar search results \\
+Benchmark & Google Scholar search results \\
 \hline
@@
 \end{tabular}
-\caption{Google scholar search results for Learning to Rank benchmarks}
+\caption{Google Scholar search results statistics for Learning to Rank benchmarks}
 \label{tbl:ltr_benchmark_searchres}
@@
 \section{Comparison Methodology}
-The LETOR 3.0 paper \cite{Qin2010} states that it may differ between datasets what the most accurate ranking methods are.
-To evaluate the overall performance of Learning to Rank methods over the multiple datasets in the LETOR 3.0 collections, Qin et al.
-\cite{Qin2010} proposed a measure called \emph{winning number} as the number of other algorithms that an algorithm can beat over the set of datasets.
+The LETOR 3.0 paper \cite{Qin2010} states that it may differ between data sets what the most accurate ranking methods are.
+To evaluate the overall performance of Learning to Rank methods over the multiple data sets in the LETOR 3.0 collections, Qin et al.
+\cite{Qin2010} proposed a measure called \emph{winning number} as the number of other algorithms that an algorithm can beat over the set of data sets.
 Formally the winning number measure is defined as\\
@@
 
-where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{NDCG} or \ac{MAP}), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that\\
+where $j$ is the index of a dataset, $n$ the number of data sets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{NDCG} or \ac{MAP}), and $I_{\{M_i(j)>M_k(j)\}}$ is an indicator function such that\\
 
@@
 To compare algorithms based on a sparse set of evaluation measurements, a normalised version of the Winning Number metric will be used.
-This Normalised Winning Number takes only those datasets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all datasets on which it has been evaluated.
-We will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as \\
+This Normalised Winning Number takes only those data sets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all data sets on which it has been evaluated.
+We will redefine the indicator function $I$ in order to only take into account those data sets that an algorithm is evaluated on, as \\
 
@@
 
-where $j$ is the index of a dataset, $n$ the number of datasets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{NDCG} or \ac{MAP}), and $D_{\{M_i(j),M_k(j)\}}$ is an evaluation definition function such that\\
+where $j$ is the index of a dataset, $n$ the number of data sets in the comparison, $i$ and $k$ are indices of an algorithm, $M_i(j)$ is the performance of the $i$-th algorithm on the $j$-th dataset, $M$ is a ranking measure (such as \ac{NDCG} or \ac{MAP}), and $D_{\{M_i(j),M_k(j)\}}$ is an evaluation definition function such that\\
 
@@
 \ac{NDCG}@3, \ac{NDCG}@5, \ac{NDCG}@10 and \ac{MAP} are chosen as metrics on which the meta-analysis will be performed.
-These metrics seem to be the most frequently used evaluation metrics in most of the used benchmark datasets.
+These metrics seem to be the most frequently used evaluation metrics in most of the used benchmark data sets.
 An exception is the Yahoo!
-Learning to Rank Challenge datasets on which mainly \ac{ERR} is used as main evaluation metric.
+Learning to Rank Challenge data sets on which mainly \ac{ERR} is used as main evaluation metric.
 The lack of use of the \ac{ERR}-metric in other benchmarks makes it unsuitable for a cross-benchmark comparison.
@@
 \section{Evaluation Results Found in Literature}
-Table \ref{tab:ltr_methods_used} gives an overview of the Learning to Rank methods for which evaluation results were found for one or more of the benchmark datasets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
+Table \ref{tab:ltr_methods_used} gives an overview of the Learning to Rank methods for which evaluation results were found for one or more of the benchmark data sets listed in section~\ref{sec:collecting_evaluation_results} through the literature review process also described in section~\ref{sec:collecting_evaluation_results}.
 Occurrences of L2, L3 and L4 in Table \ref{tab:ltr_methods_used} imply that these algorithms are evaluated as official LETOR 2.0, LETOR 3.0 and LETOR 4.0 baselines respectively.\\
@@
 \end{tabular}}
-\caption{Learning to Rank algorithms with measurements on benchmark datasets}
+\caption{An overview of Learning to Rank algorithms and their occurrence in evaluation experiments on benchmark data sets}
 \label{tab:ltr_methods_used}
@@
 The following subsections provide the performance of Learning to Rank methods in terms of Normalised Winning Number for \ac{NDCG}@3, \ac{NDCG}@5, \ac{NDCG}@10 and \ac{MAP}.
-Performance of the Learning to Rank methods is plotted with Normalised Winning Number on the vertical axis and the number of datasets on which the method has been evaluated on the horizontal axis.
+Performance of the Learning to Rank methods is plotted with Normalised Winning Number on the vertical axis and the number of data sets on which the method has been evaluated on the horizontal axis.
 The further to the right, the more certain we can be about the performance of the Learning to Rank method.
-The methods for which it holds that there is no other method that has 1) a higher Normalised Winning Number and 2) a higher number datasets evaluated, are identified as the best performing methods and are labelled with the name of the method.
+The methods for which it holds that there is no other method that has 1) a higher Normalised Winning Number and 2) a higher number data sets evaluated, are identified as the best performing methods and are labelled with the name of the method.
 
@@
 LambdaNeuralRank and CoList both acquired a perfect Normalised Winning Number score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
-LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three datasets which in addition are more relevant datasets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 datasets (see Table \ref{tab:ltr_methods_used}).
-FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in Normalised Winning Number, but increasingly higher in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
+LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 data sets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three data sets which in addition are more relevant data sets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 data sets (see Table \ref{tab:ltr_methods_used}).
+FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in Normalised Winning Number, but increasingly higher in number of data sets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
 
@@
 LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a Normalised Winning Number of 1.0.
-LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of datasets, but score decreasingly well in terms of Normalised Winning Number.
+LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of data sets, but score decreasingly well in terms of Normalised Winning Number.
 These results are highly in agreement with the \ac{NDCG}@3 comparison.
@@
 LambdaMART showed the highest \ac{NDCG}@10 performance for the MSLR-WEB10k dataset.
-The set of algorithms for which there is no other algorithm with both a higher Normalised Winning Number and number of datasets evaluated on is partly in agreement with those for the \ac{NDCG}@3 and \ac{NDCG}@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
+The set of algorithms for which there is no other algorithm with both a higher Normalised Winning Number and number of data sets evaluated on is partly in agreement with those for the \ac{NDCG}@3 and \ac{NDCG}@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
 SmoothRank and FSMRank were not present in this set for the \ac{NDCG}@3 and \ac{NDCG}@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_NDCG3} and \ref{tab:raw_data_norm_winnum_NDCG5} in the Appendices.
@@
 
-LARF and LRUF score very high Normalised Winning Number scores, but based on only relatively few datasets, just as in the \ac{NDCG}-comparisons.
+LARF and LRUF score very high Normalised Winning Number scores, but based on only relatively few data sets, just as in the \ac{NDCG}-comparisons.
 Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all \ac{NDCG}-comparisons, are not within the best performing methods in the \ac{MAP}-comparison.
-A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_NDCG3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_NDCG3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more datasets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalised Winning Number.
-SmoothRank performed a Normalised Winning Number of around 0.53 over 7 datasets, which is still good in both certainty and accuracy, but not among the top methods.
+A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_NDCG3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_NDCG3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more data sets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalised Winning Number.
+SmoothRank performed a Normalised Winning Number of around 0.53 over 7 data sets, which is still good in both certainty and accuracy, but not among the top methods.
 RE-QR is one of the best performers in the \ac{MAP}-comparison with a reasonable amount of benchmark evaluations.
 No reported \ac{NDCG} performance was found in the literature study for RE-QR.
-There is a lot of certainty on the accuracy of RankBoost and Rank\ac{SVM} as both models are evaluated on the majority of datasets included in the comparison for the \ac{MAP}-metric, but given their Normalised Winning Number it can said that both methods are not within the top performing Learning to Rank methods.
+There is a lot of certainty on the accuracy of RankBoost and Rank\ac{SVM} as both models are evaluated on the majority of data sets included in the comparison for the \ac{MAP}-metric, but given their Normalised Winning Number it can said that both methods are not within the top performing Learning to Rank methods.
 
@@
 C-CRF, DirectRank, FP-Rank, RankCSA, LambdaNeuralRank and VFLR all have near-perfect Normalised Winning Number measures, but have low Ideal Winning Number measures.
-Further evaluation runs of these methods on benchmark datasets that they are not yet evaluated on are desirable.
-The DirectRank paper \cite{Tan2013} shows that the method  is evaluated on more datasets than the number of datasets that we included evaluation results for in this meta-analysis.
-Some of the DirectRank measurements could not be used because measurements on some datasets were only available in graphical form and not in raw data.\\
+Further evaluation runs of these methods on benchmark data sets that they are not yet evaluated on are desirable.
+The DirectRank paper \cite{Tan2013} shows that the method  is evaluated on more data sets than the number of data sets that we included evaluation results for in this meta-analysis.
+Some of the DirectRank measurements could not be used because measurements on some data sets were only available in graphical form and not in raw data.\\
 
-LAC-MR-OR and RE-QR showed very good ranking accuracy in the \ac{MAP} comparison on multiple datasets.
-Because LAC-MR-OR is only evaluated on two datasets for \ac{NDCG}@10 and RE-QR is not evaluated for \ac{NDCG} at all, LAC-MR-OR and RE-QR are not within the top performing methods in the cross-metric comparison.
+LAC-MR-OR and RE-QR showed very good ranking accuracy in the \ac{MAP} comparison on multiple data sets.
+Because LAC-MR-OR is only evaluated on two data sets for \ac{NDCG}@10 and RE-QR is not evaluated for \ac{NDCG} at all, LAC-MR-OR and RE-QR are not within the top performing methods in the cross-metric comparison.
 
@@
 By calculating it in this way, we implicitly make the assumption that the Learning to Rank methods are (approximately) distributed uniformly over the benchmarks, such that the average Learning to Rank method tested are approximately equally hard for each data set.
-It could be the case however that this assumption is false and that significantly more accurate Learning to Rank methods are evaluated on some datasets than on other datasets.
+It could be the case however that this assumption is false and that significantly more accurate Learning to Rank methods are evaluated on some data sets than on other data sets.
 \\
 
-A second limitation is that the datasets on which Learning to Rank methods have been evaluated can not always be regarded a random choice.
-It might be the case that some researchers chose to publish results for exactly those benchmark datasets that showed the most positive results for their Learning to Rank method.\\
+A second limitation is that the data sets on which Learning to Rank methods have been evaluated can not always be regarded a random choice.
+It might be the case that some researchers chose to publish results for exactly those benchmark data sets that showed the most positive results for their Learning to Rank method.\\
 
@@
 \begin{description}
-\item[RQ1] What are the best performing Learning to Rank algorithms in terms of accuracy on relevant benchmark datasets?\\
+\item[RQ1] What are the best performing Learning to Rank algorithms in terms of accuracy on relevant benchmark data sets?\\
 \end{description}
@@
 From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.
-For more definite conclusions on the relative performance of these five methods, more evaluation runs on are desirable for the methods on the left side on the list on benchmark datasets that these methods have not yet been evaluated on.\\
+For more definite conclusions on the relative performance of these five methods, more evaluation runs on are desirable for the methods on the left side on the list on benchmark data sets that these methods have not yet been evaluated on.\\
 

### 2014-09-23 report/Chapters/Implementation.tex
dataset -> data set

@@
 The first Pig job determines the minimum and the maximum values per feature in the training set.
-The second Pig job rescales each feature of the train, validation and test datasets using the following formula for rescaling:
+The second Pig job rescales each feature of the train, validation and test data sets using the following formula for rescaling:
 \begin{equation}
@@
 REGISTER [path prefix]/lib/*.jar;
-TRAIN = LOAD '[path prefix]/input/[dataset name]/Fold[fold number]/train.txt' USING PigStorage(' ');
+TRAIN = LOAD '[path prefix]/input/[data set name]/Fold[fold number]/train.txt' USING PigStorage(' ');
 TRAIN_STD = FOREACH TRAIN GENERATE flatten(udf.util.ToStandardForm(*));
@@
 REGISTER [path prefix]/lib/*.jar;
-TRAIN = LOAD '[path prefix]/input/[dataset name]/Fold[fold number]/train.txt' USING PigStorage(' ');
-VALIDATE = LOAD '[path prefix]/input/[dataset name]/Fold[fold number]/vali.txt' USING PigStorage(' ');
-TEST = LOAD '[path prefix]/input/[dataset name]/Fold[fold number]/test.txt' USING PigStorage(' ');
+TRAIN = LOAD '[path prefix]/input/[data set name]/Fold[fold number]/train.txt' USING PigStorage(' ');
+VALIDATE = LOAD '[path prefix]/input/[data set name]/Fold[fold number]/vali.txt' USING PigStorage(' ');
+TEST = LOAD '[path prefix]/input/[data set name]/Fold[fold number]/test.txt' USING PigStorage(' ');
 TRAIN_STD = FOREACH TRAIN GENERATE flatten(udf.util.ToStandardForm(*));

### 2014-09-23 report/Chapters/Introduction.tex
verbeteringen
dataset -> data set

@@
 
-In recent years several Learning to Rank benchmark datasets have been proposed that enable comparison of the performance of different Learning to Rank methods.
-Well-known benchmark datasets include the \emph{Yahoo!
-Learning to Rank Challenge} dataset \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR dataset \cite{Qin2010} that was published by Microsoft Research.
+In recent years several Learning to Rank benchmark data sets have been proposed that enable comparison of the performance of different Learning to Rank methods.
+Well-known benchmark data sets include the \emph{Yahoo!
+Learning to Rank Challenge} data set \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR data set \cite{Qin2010} that was published by Microsoft Research.
 One of the concluding observations of the \emph{Yahoo!
@@
 
-The appearance of benchmark datasets gave insight in the performance of different Learning to Rank approaches, which resulted in increasing popularity of those methods that showed to perform well on one or more benchmark datasets.
+The appearance of benchmark data sets gave insight in the performance of different Learning to Rank approaches, which resulted in increasing popularity of those methods that showed to perform well on one or more benchmark data sets.
 Up to now it remains unknown whether popular existing Learning to Rank methods scale well when they are used in a parallel manner using the MapReduce framework.
@@
 The objective of this thesis is to explore the speed-up in execution time of Learning to Rank algorithms through parallelisation using the MapReduce framework.
-This work focuses on those Learning to Rank algorithms that have shown leading performance on relevant benchmark datasets.
+This work focuses on those Learning to Rank algorithms that have shown leading performance on relevant benchmark data sets.
 This thesis addresses the following research questions:
 \begin{description}
-\item[RQ1] What are the best performing Learning to Rank algorithms in terms of accuracy on relevant benchmark datasets?\\
+\item[RQ1] What are the best performing Learning to Rank algorithms in terms of accuracy on relevant benchmark data sets?\\
 
@@
 
-To research the speed-up's dependence on the amount of data being processed, the training computations will be performed on datasets of varying sizes.
-The well-known benchmark collections LETOR 3.0, LETOR 4.0 and MSLR-WEB30/40K offer a good initial set of datasets.
+To research the speed-up's dependence on the amount of data being processed, the training computations will be performed on data sets of varying sizes.
+The well-known benchmark collections LETOR 3.0, LETOR 4.0 and MSLR-WEB30/40K offer a good initial set of data sets.
 Table \ref{tbl:initial_datasets} shows the
@@
 \begin{tabular}{p{3.4cm}p{3.4cm}r}\toprule
-Dataset & Collection & Size \\
+Data set & Collection & Size \\
 \midrule
@@
 \end{tabular}
-\caption{LETOR 3.0, LETOR 4.0 and MSLR30/40K datasets and sizes}
+\caption{The LETOR 3.0, LETOR 4.0 and MSLR30/40K data sets and their data sizes}
 \label{tbl:initial_datasets}
@@
 
-To test the how the computational performance of Learning to Rank algorithms both on cluster and on single-node computation scales to large quantities of data, larger datasets will be created by cloning the MSLR-WEB30K dataset such that the cloned queries will hold new distinct query ID's.
+To test the how the computational performance of Learning to Rank algorithms both on cluster and on single-node computation scales to large quantities of data, larger data sets will be created by cloning the MSLR-WEB30K data set such that the cloned queries will hold new distinct query ID's.
 \section{Thesis Overview}
@@
 \item[Chapter III: Related Work]{concisely describes existing work in the field of parallel Learning to Rank.}
-\item[Chapter IV: Benchmark Results]{sketches the accuracy of existing Learning to Rank methods on several benchmark datasets and describes the selection of Learning to Rank methods for the parallelisation experiments.}
+\item[Chapter IV: Benchmark Results]{sketches the accuracy of existing Learning to Rank methods on several benchmark data sets and describes the selection of Learning to Rank methods for the parallelisation experiments.}
 \item[Chapter V: Selected Learning to Rank Methods]{describes the algorithms and details of the Learning to Rank methods selected in Chapter IV.}

### 2014-09-23 report/Chapters/RelatedWork.tex
dataset -> data set

@@
 \cite{Veloso2008}, has been implemented using the \ac{GPU} in such a way that the set of rules van be computed simultaneously for each document.
-A speed-up of 127X in query processing time is reported based on evaluation on the LETOR dataset.
+A speed-up of 127X in query processing time is reported based on evaluation on the LETOR data set.
 The speed-up achieved at learning the ranking function was unfortunately not stated.\\
@@
 the divide-and-conquer nature of the \ac{CC} algorithm enables parallelisation.
-CCRank showed an increase in both accuracy and efficiency on the LETOR 4.0 benchmark dataset compared to its baselines.
+CCRank showed an increase in both accuracy and efficiency on the LETOR 4.0 benchmark data set compared to its baselines.
 However, the increased efficiency was achieved through speed-up and not scale-up.
@@
 The authors successfully parallelised their method in the MapReduce paradigm using Hadoop.
-The approach showed to be effective on both the LETOR 3.0 dataset and their own dataset with contextual advertising data.
+The approach showed to be effective on both the LETOR 3.0 data set and their own data set with contextual advertising data.
 Unfortunately their work does not directly report on the speed-up obtained by parallelising  with Hadoop, but it is mentioned that further work needs to be done to effectively leverage parallel execution.\\
@@
 In contract to BagBoo, BL-MART is limited in the number of trees.
-An experiment on the TD2004 dataset resulted in 4500 trees using BL-MART while 1.1 million trees were created with the BagBoo model.
+An experiment on the TD2004 data set resulted in 4500 trees using BL-MART while 1.1 million trees were created with the BagBoo model.
 
@@
 The authors called this method \ac{ET} and originally proposed this method for classification and regression settings instead of ranking settings \cite{Geurts2006}.
-The \ac{ET} algorithm is similar to the well-known Random Forest algorithm, but with two differences: 1) in \ac{ET} each tree is build on the complete dataset instead of random subsets of the data and 2) \ac{ET} sets the discretisation thresholds that define the splits in the tree randomly, instead of based on the sample.
+The \ac{ET} algorithm is similar to the well-known Random Forest algorithm, but with two differences: 1) in \ac{ET} each tree is build on the complete data set instead of random subsets of the data and 2) \ac{ET} sets the discretisation thresholds that define the splits in the tree randomly, instead of based on the sample.
 The randomisations in the process make the models in the ensemble mutually independent, which makes it trivial to parallelise by training each tree on a different node.

### 2014-09-23 report/Chapters/Results.tex
dataset -> data set

@@
 \end{tabular}
-\caption{\acs{NDCG}@10 performance on the testing set on the first fold}
+\caption{\acs{NDCG}@10 performance on the test set of the first fold}
 \label{tbl:accuracy_comparison}
@@
 Figure \ref{fig:listnet_train_time} (linear data size axis) and Figure \ref{fig:listnet_train_time_log} (logarithmic data size axis) show the processing times of a single iteration of the ListNet training algorithm, using the ListNet implementation included in the RankLib library as well as the cluster implementation described in chapter \ref{chap:implementation} with different numbers of data nodes.
-The horizontal positions of the measurements are identical between execution method, as they are originate from the data sizes of the datasets used.
-Measurements were performed on a set of datasets consisting of the LETOR 3.0 OHSUMED dataset, LETOR 4.0 datasets and the MSLR-WEB10/30K datasets are used, supplemented with generated datasets that are multiplications of MSLR-WEB30K (as described in section \ref{ssec:rq2_methodology}).
-Table \ref{tbl:recap_datasets} describes the datasets used in the experiments and their single training fold data sizes.\\
+The horizontal positions of the measurements are identical between execution method, as they are originate from the data sizes of the data sets used.
+Measurements were performed on a set of data sets consisting of the LETOR 3.0 OHSUMED data set, LETOR 4.0 data sets and the MSLR-WEB10/30K data sets are used, supplemented with generated data sets that are multiplications of MSLR-WEB30K (as described in section \ref{ssec:rq2_methodology}).
+Table \ref{tbl:recap_datasets} describes the data sets used in the experiments and their single training fold data sizes.\\
 
@@
 \begin{tabular}{p{3.4cm}p{3.4cm}r}\toprule
-Dataset & Collection & Single fold training size \\
+Data set & Collection & Single fold training size \\
 \midrule
@@
 \end{tabular}
-\caption{Description of datasets used for running time experiments}
+\caption{Description of data sets used for running time experiments}
 \label{tbl:recap_datasets}
@@
 
-Figures \ref{fig:listnet_train_time} and \ref{fig:listnet_train_time_log} show that the single-machine RankLib implementation of ListNet is very fast compared to the Hadoop MapReduce cluster implementation of ListNet for datasets that fit in memory.
+Figures \ref{fig:listnet_train_time} and \ref{fig:listnet_train_time_log} show that the single-machine RankLib implementation of ListNet is very fast compared to the Hadoop MapReduce cluster implementation of ListNet for data sets that fit in memory.
 As soon as the amount of data processed with RankLib ListNet approaches the physical memory limit of the machine that is used for computation (16 GB for our single-machine measurements), RankLib ListNet processing time start to increase exponentially.
 This increase is likely to be a result of the machine needing to use the on-disk sections of virtual memory and the swapping that takes place as a result thereof.
-As an exception to the datasets described in Table \ref{tbl:recap_datasets}, the largest dataset processed with RankLib ListNet is CUSTOM-8, which is 20.99 GB in size and thereby larger than the 16 GB physical memory limit.
+As an exception to the data sets described in Table \ref{tbl:recap_datasets}, the largest data set processed with RankLib ListNet is CUSTOM-8, which is 20.99 GB in size and thereby larger than the 16 GB physical memory limit.
 Experiments with RankLib ListNet on CUSTOM-10 were attempted, but were stopped after not completing within 12 hours.
 In section \ref{sec:goals} we defined speed-up as Sun and Gustafson \emph{relative speed-up} metric \cite{Sun1991}, which defines speed-up as the number of times that the execution of the fastest single machine solution is lower than the execution time with $N$ machines.
-Note that speed-up, following this definition, is not computable for larger datasets, as it is not computationally feasible to perform computation on datasets larger than CUSTOM-8 with a single machine.
+Note that speed-up, following this definition, is not computable for larger data sets, as it is not computationally feasible to perform computation on data sets larger than CUSTOM-8 with a single machine.
 As an alternative, Figure \ref{fig:speedup_train_time} shows visualises speed-up by plotting cluster size against training time.\\
@@
 Figure \ref{fig:speedup_train_time} shows that speed-up is sub-linear, from which we can deduct that the training time converges to a constant unit of time.
-Based on our measurements on the small datasets MINI, OHSUMED, MQ2008 and MQ2007, this constant time seems to be within the range of 150 to 200 seconds.
+Based on our measurements on the small data sets MINI, OHSUMED, MQ2008 and MQ2007, this constant time seems to be within the range of 150 to 200 seconds.
 This time is likely to be caused by Hadoop job scheduling overhead, this presumption is strengthened by long waiting periods between the separate MapReduce jobs that form a training iteration.\\
@@
 
-Note however that Hadoop job scheduling overhead is independent of dataset size.
-Therefore, the non-parallelisable fraction of the program will be smaller when the to be processed dataset is larger, allowing larger speed-up values for larger datasets.
-From this observation we can derive that for \emph{"large enough"} datasets, the speed-up obtained by parallelising ListNet using Hadoop MapReduce is large enough for the parallelisation to be beneficial in terms of processing time compared to the RankLib single-machine version of ListNet, even when the size of the to be processed dataset would not have been memory-bounded in RankLib ListNet.\\
+Note however that Hadoop job scheduling overhead is independent of data set size.
+Therefore, the non-parallelisable fraction of the program will be smaller when the to be processed data set is larger, allowing larger speed-up values for larger data sets.
+From this observation we can derive that for \emph{"large enough"} data sets, the speed-up obtained by parallelising ListNet using Hadoop MapReduce is large enough for the parallelisation to be beneficial in terms of processing time compared to the RankLib single-machine version of ListNet, even when the size of the to be processed data set would not have been memory-bounded in RankLib ListNet.\\
 

### 2014-09-23 report/Chapters/AppendixA.tex
afwerking

@@
 }
-\caption{Features of the LETOR OHSUMED data set, obtained from Qin et al \cite{Qin2010}}
+\caption{Features of the LETOR 3.0 OHSUMED data set, obtained from Qin et al \cite{Qin2010}}
 \label{tbl:features_ohsumed}
@@
 }
-\caption{Features of the LETOR .GOV data set, obtained from Qin et al \cite{Qin2010}}
+\caption{Features of the LETOR 3.0 .gov data set, obtained from Qin et al \cite{Qin2010}}
 \label{tbl:features_gov}

### 2014-09-23 report/Chapters/BenchmarkResults.tex
afwerking

@@
 The LETOR 3.0 benchmark collection \cite{Qin2010} as released in 2007 contained two data sets: the OHSUMED collection and the .gov collection.
-Tables \ref{tbl:features_gov} and \ref{tbl:features_ohsumed} in Appendix \ref{chap:appendix_a} provide the descriptions of the features of the .gov and the OHSUMED collections for LETOR 3.0 respectively.
+Tables \ref{tbl:features_ohsumed} and \ref{tbl:features_gov} in Appendix \ref{chap:appendix_a} provide the descriptions of the features of the OHSUMED and the .gov collections for LETOR 3.0 respectively.
 Where LETOR 2.0 only used topic distillation (TD) queries on the .gov corpus, LETOR 3.0 also uses named page finding (NP) and home page finding (HP) queries.

### 2014-09-23 report/Chapters/Implementation.tex
begin feedback Jurjen

@@
 The first section of this chapter will briefly discuss the HDInsight platform, the Hadoop ecosystem components offered by HDInsight and, in this regard, the Hadoop components used for implementation of the algorithms described in Chapter \ref{chap:ltr_methods}.
-The second section of this chapter describes a Java framework that handles the joint components needed for MapReduce Learning to Rank computation that are independent of the Learning to Rank model.
-The subsequent sections describe the implementation details of specific Learning to Rank models.
+The second section of this chapter describes a Java framework that handles the joint components needed for MapReduce Learning to Rank computation, independent of the Learning to Rank model.
+The subsequent sections report the implementation details of specific Learning to Rank models.
 
 \section{Architecture}
-Ranking algorithms consist of a sequence of operations on the input data which are often, but not always, of iterative nature.
-Apache Pig \cite{Olston2008} will be used to implement the sequence of operations on the input data.
-Pig Latin, the data processing language that runs on Apache Pig, was designed based on the observation that the traditional MapReduce paradigm is too low-level and rigid, and holds the middle between the declarative style of SQL and the procedural style of MapReduce.
-The Apache Pig system translates Pig Latin into MapReduce plans that are executed over Hadoop.
+Ranking algorithms consist of a sequence of operations on the input data which are often of iterative nature.
+Apache Pig \cite{Olston2008} is used to implement the sequence of operations on the input data.
+Pig Latin is the data processing language that runs on Apache Pig.
+It was designed based on the observation that the traditional MapReduce paradigm is too low-level and rigid, and holds the middle between the declarative style of SQL and the procedural style of MapReduce.
+The Apache Pig system translates Pig Latin into MapReduce plans that are executed on Hadoop.
 The choice to implement the Learning to Rank algorithms in Pig Latin allows for more focus on the data operations and less focus on low-level implementation details, as compared to native Hadoop MapReduce.
@@
 \subsection{The HDInsight Platform}
-Azure HDInsight supports both the traditional \ac{HDFS} as described by Shvacko et al.
+Azure HDInsight supports the traditional \ac{HDFS} as described by Shvacko et al.
 \cite{Shvachko2010}, as well as Microsoft's own storage solution \ac{WASB}.
@@
 Azure \ac{WASB} storage allows the user to select one of Microsoft's data centres for storage.
-\ac{WASB} storage in the West Europe region (located in Amsterdam) will be used for storage, as this data centre is located close to where the experiments are executed.\\
+\ac{WASB} storage in the West Europe region (located in Amsterdam) is used for storage, as this data centre is located close to where the experiments are executed.\\
 
@@
 Newly supported programming models since Hadoop 2.0 include Dryad \cite{Isard2007}, Giraph \cite{Avery2011}, MPI, REEF \cite{Chun2013}, Spark \cite{Zaharia2010}, and Storm \cite{Aniello2013}.
-Even though these programming models are now supported by Hadoop and even though some of these programming models have recently increased in popularity, they still lack the critical mass as the data processing framework of choice that MapReduce has as Lin argued back in 2012 \cite{Lin2013}(see section \ref{sec:motivation}).
+Even though these programming models are now supported by Hadoop and some of these programming models have recently increased in popularity, they still lack the critical mass as the data processing framework of choice that MapReduce has as Lin argued back in 2012 \cite{Lin2013}(see section \ref{sec:motivation}).
 Therefore, even though some of these programming models might be a better fit for iterative algorithms, we use Hadoop MapReduce of the programming model to implement the Learning to Rank algorithms.\\
@@
 Similar to \ac{WASB} storage, one of the Microsoft data centres can be chosen to host the Hadoop cluster when a HDInsight cluster is created.
-To guarantee proximity between storage and cluster, the West Europe data centre in Amsterdam will be used as cluster location.\\
+To guarantee proximity between storage and cluster, the West Europe data centre in Amsterdam is used as cluster location.\\
 

### 2014-09-23 report/Chapters/Background.tex
verbeteringen
begin feedback Jurjen
nieuwe captions algorithms

@@
 where the product sequence part of the formula represents the chance that the user will stop at position $r$.\\
-An algorithm to compute \ac{ERR} is shown in Listing \ref{alg:err}.
+The algorithm to compute \ac{ERR} is shown in Listing \ref{alg:err}.
 The algorithm requires relevance grades $g_i$, $1 \le i \le n$ and mapping function $R$ that maps relevance grades to probability of relevance.
@@
  Output $ERR$
- \caption{Algorithm to compute the \acs{ERR} metric, obtained from \cite{Chapelle2009}}
+ \caption{The algorithm for computation of the \acs{ERR} metric, obtained from \cite{Chapelle2009}}
  \label{alg:err}
@@
 Table \ref{tab:example_calculation_AP} provides an example calculation of average precision where de documents at positions 1, 5, 6 and 8 in the ranking are relevant.
+The total number of available relevant documents in the document set \emph{R} is assumed to be seven.
 \ac{MAP} is the average \ac{AP} for a set of queries.\\
@@
 \hline
- &  &  &  &  &  &  &  &  &  & /\#Relevant docs & = & 7 \\ 
+ &  &  &  &  &  &  &  &  &  & \# of relevant docs & = & 7 \\ 
  &  &  &  &  &  &  &  &  &  & AP@10 & = & 0.34 \\ 
 \end{tabular}
-\caption{Average Precision example calculation. The number of relevant documents \emph{R} is assumed to be seven.}
+\caption{Average Precision example calculation.}
 \label{tab:example_calculation_AP}
@@
 Well-known algorithms that belong to the listwise approach include AdaRank \cite{Xu2007}, BoltzRank \cite{Volkovs2009}, ListMLE \cite{Xia2008}, ListNet \cite{Cao2007}, RankCosine \cite{Qin2008}, SmoothRank \cite{Chapelle2010}, SoftRank \cite{Taylor2008}, \acs{SVM}$^{map}$ \cite{Yue2007}.
+
+\section{Cross-validation experiments}
+TODO
+
+\section{An introduction to the MapReduce programming model}
+TODO

### 2014-09-23 report/Chapters/Implementation.tex
nieuwe captions algorithms

@@
 Fold handling is a task that needs to be taken care of independent of the ranking model that is being evaluated.
-Given that most ranking models are of an iterative nature, the task of iteration handling is also a task that need to be performed independent of the ranking model.
+Given that most ranking models are of iterative nature, the task of iteration handling is also a task that need to be performed independent of the ranking model.
 Iteration handling and fold handling are procedures in which the same steps are repeated for each iteration or cross-validation round respectively.
@@
 The aim of this framework is to let implementations of ranking models focus solely on implementing the sequential steps of one iteration of the algorithm, while the framework handles that 1) these sequential steps are performed iteratively, 2) these sequential steps are performed on the multiple training folds of data and 3) data can be passed from one Pig job within the algorithm to another Pig job.
-This framework that handles folding and iterations will be implemented in Java and will work such that fold- and iteration dependent parts of the Pig code will be generated dynamically by the Java framework after which the Pig job will be sent to the cluster.\\
+Our framework that handles folding and iterations will be implemented in Java and will work such that fold- and iteration dependent parts of the Pig code will be generated dynamically by the Java framework after which the Pig job will be sent to the cluster.\\
 
@@
 The minimum and maximum values per feature stored in MIN\_MAX\_FIN are read from storage by the framework after completion of the MapReduce job.
-The framework passes these minimum and maximum values to the second Pig job by passing an array of minimum and maximum values to the constructor of the udf.util.ScaleFeatures() \ac{UDF}, that performs the rescaling operation of Equation \ref{eq:rescaling}s.\\
+The framework passes these minimum and maximum values to the second Pig job by passing an array of minimum and maximum values to the constructor of the udf.util.ScaleFeatures() \ac{UDF}, that performs the rescaling operation of Equation \ref{eq:rescaling}.\\
 

### 2014-09-23 report/Chapters/LTRMethods.tex
nieuwe captions algorithms

@@
  Output Neural Network model $\omega$.
- \caption{Learning algorithm of ListNet, obtained from \cite{Cao2007}}
+ \caption{The ListNet learning algorithm, obtained from \cite{Cao2007}}
  \label{alg:listnet}
@@
  }
- \caption{Learning algorithm of SmoothRank, obtained from \cite{Chapelle2010}}
+ \caption{The SmoothRank learning algorithm, obtained from \cite{Chapelle2010}}
  \label{alg:smoothrank}
@@
  return $w_{T}$ as ranking predictor for $w$
- \caption{Learning algorithm of FenchelRank, obtained from \cite{Lai2013}}
+ \caption{The FenchelRank learning algorithm, obtained from \cite{Lai2013}}
  \label{alg:fenchelrank}
@@
 
- \caption{Learning algorithm of FSMRank, obtained from \cite{Lai2013c}}
+ \caption{The FSMRank learning algorithm, obtained from \cite{Lai2013c}}
  \label{alg:fsmrank}
@@
  Output $\underline{R_i}$
- \caption{LRUF algorithm, obtained from \cite{Torkestani2012b}}
+ \caption{The LRUF algorithm, obtained from \cite{Torkestani2012b}}
  \label{alg:lruf}

### 2014-09-23 report/Chapters/Results.tex
nieuwe captions algorithms

@@
-This chapter describes and discusses the results of the running time measurements of the cluster and baseline implementations as described in \ref{chap:implementation}.
+This chapter describes and discusses the results of the execution time measurements of the cluster and baseline implementations as described in \ref{chap:implementation}.
 In addition, ranking accuracy results are included to validate the correctness of the Learning to Rank algorithm implementations.
@@
 \subsection{Ranking accuracy}
-To validate our ListNet cluster implementation we run experiments on the LETOR 3.0 OHSUMED and the LETOR 4.0 datasets MQ2007 and MQ2008 and compare their performance with the RankLib implementation of ListNet and with the official evaluation results in terms of \ac{NDCG}@10.
+To validate our ListNet cluster implementation we run experiments on the LETOR 3.0 OHSUMED, the LETOR 4.0 datasets MQ2007 and MQ2008 and compare their performance with the RankLib implementation of ListNet and with the official evaluation results in terms of \ac{NDCG}@10.
 Hadoop job scheduling overhead is large when processing relatively small datasets (as we will explain in more detail in section \ref{ssec:speedup}), therefore \ac{NDCG}@10 experiments are limited to five training iterations and are run only on one of the five folds.
@@
 Note that the official LETOR evaluation results are structurally higher compared to the results obtained with RankLib and our cluster implementation, which can be attributed to the fact that the official evaluation runs use as many ListNet iterations as needed for convergence on the validation fold, while we limited our experiment to five iterations.
-The official evaluation results of the LETOR benchmarks do not report the standard deviation of their experimental results and it is not mentioned in the LETOR papers \cite{Qin2010, Qin2013} how many times their experiments were repeated.
+The official evaluation results of the LETOR benchmarks do not report the standard deviation of their experimental results and it is not mentioned in the LETOR 3.0 and 4.0 papers \cite{Qin2010, Qin2013} how many times their experiments were repeated.
 For the ListNet step size hyperparameter we choose a value of $0.0001$, which is the default value step size value in RankLib, for both the cluster and the RankLib runs.

### 2014-09-23 report/Chapters/Introduction.tex
genoeg voor vandaag

@@
 The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a given query.
-Ranking has become a vital part of web search, where commercial search engines help users find their need in the extremely large collection of the World Wide Web.
+Ranking has become a vital part of web search, where commercial search engines help users find their need in the extremely large document collection of the World Wide Web.
 One can find useful applications of ranking in many application domains outside web search as well.

### 2014-09-24 report/Chapters/Introduction.tex
toevoegingen intro

@@
 One can find useful applications of ranking in many application domains outside web search as well.
-For example, it plays a vital role in amongst others: automatic document summarisation, machine translation, drug discovery and in determining the ideal order of maintenance operations \cite{Rudin2009}.
+For example, it plays a vital role in automatic document summarisation, where it can be used to rank sentences in a document according to their contribution to a summary of that document
+\cite{Bravo-Marquez2012}.
+Learning to Rank also plays a role in the fields of machine translation \cite{Hopkins2011}, automatic drug discovery \cite{Agarwal2010b} and it is used to determine the ideal order in a sequence of maintenance operations \cite{Rudin2009}.
 In addition, the ranking task has been argued to be more fitting to recommender systems than regression-based rating prediction on a continuous scale \cite{Adomavicius2005,McNee2006}.\\

### 2014-09-24 report/Chapters/Introduction.tex
aanscherping intro

@@
 The ranking task in information retrieval entails the ranking of candidate documents according to their relevance for a given query.
-Ranking has become a vital part of web search, where commercial search engines help users find their need in the extremely large document collection of the World Wide Web.
+Ranking has become a vital part of web search, where commercial search engines help users find their need in the extremely large document collection of the World Wide Web.\\
+
 One can find useful applications of ranking in many application domains outside web search as well.
@@
 \cite{Bravo-Marquez2012}.
-Learning to Rank also plays a role in the fields of machine translation \cite{Hopkins2011}, automatic drug discovery \cite{Agarwal2010b} and it is used to determine the ideal order in a sequence of maintenance operations \cite{Rudin2009}.
+Learning to Rank also plays a role in the fields of machine translation \cite{Hopkins2011}, automatic drug discovery \cite{Agarwal2010b}, the prediction of chemical reactions in the field of chemistry \cite{Kayala2011}, and it is used to determine the ideal order in a sequence of maintenance operations \cite{Rudin2009}.
 In addition, the ranking task has been argued to be more fitting to recommender systems than regression-based rating prediction on a continuous scale \cite{Adomavicius2005,McNee2006}.\\

### 2014-09-24 report/Chapters/Introduction.tex
betere zin

@@
 Learning to Rank also plays a role in the fields of machine translation \cite{Hopkins2011}, automatic drug discovery \cite{Agarwal2010b}, the prediction of chemical reactions in the field of chemistry \cite{Kayala2011}, and it is used to determine the ideal order in a sequence of maintenance operations \cite{Rudin2009}.
-In addition, the ranking task has been argued to be more fitting to recommender systems than regression-based rating prediction on a continuous scale \cite{Adomavicius2005,McNee2006}.\\
+In addition, Learning to Rank has been found to be a better fit as an underlying technique compared to continuous scale regression-based prediction for applications in recommender systems \cite{Adomavicius2005,McNee2006}, like those found in Netflix or Amazon.\\
 

### 2014-09-24 report/Chapters/Introduction.tex
bbeter

@@
 Up to now it remains unknown whether popular existing Learning to Rank methods scale well when they are used in a parallel manner using the MapReduce framework.
-This thesis aims to be an exploratory start in this little researched area of parallel Learning to Rank.
-A more extensive overview of my research goals and questions are described in section \ref{sec:goals}.\\
+This thesis aims to be an exploratory start in this little researched area of parallel Learning to Rank.\\
 

### 2014-09-24 report/Chapters/Introduction.tex
intro herschreven :)

@@
 
-Research in the field of ranking models has for a long time been based on manually designed ranking functions, such as the well-known BM25 model \cite{Robertson1994}.
 Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
+This started a transformation of the Information Retrieval field from a focus on the binary classification task of labelling documents as either \emph{relevant} or \emph{not relevant} into a \emph{ranked retrieval} task that aims at ranking the documents from most to least relevant.
+Research in the field of ranking models has long been based on manually designed ranking functions, such as the well-known BM25 model \cite{Robertson1994}, that simply rank documents based on the appearance of the query terms in these documents.
 The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models.
@@
 Well-known benchmark data sets include the \emph{Yahoo!
-Learning to Rank Challenge} data set \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR data set \cite{Qin2010} that was published by Microsoft Research.
+Learning to Rank Challenge} data set \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR data sets \cite{Qin2010} that are published by Microsoft Research.\\
+
 One of the concluding observations of the \emph{Yahoo!
-Learning to Rank Challenge} was that almost all work in the Learning to Rank field focuses on ranking accuracy, while efficiency and scalability of Learning to Rank algorithms is still an underexposed research area that is likely to become more important in the near future as training sets are becoming larger and larger \cite{Chapelle2011b}.
-Liu \cite{Liu2007} confirms the observation that efficiency and scalability of Learning to Rank methods has so far been an overlooked research area in his influential book on Learning to Rank.\\
+Learning to Rank Challenge} was that almost all work in the Learning to Rank field focuses on ranking accuracy.
+Meanwhile, efficiency and scalability of Learning to Rank algorithms is still an underexposed research area that is likely to become more important in the near future as available data sets are rapidly increasing in size \cite{Chapelle2011b}.
+Liu \cite{Liu2007}, one of the members of the LETOR team at Microsoft, confirms the observation that efficiency and scalability of Learning to Rank methods has so far been an overlooked research area in his influential book on Learning to Rank.\\
 
@@
 The field of efficient Learning to Rank has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012}, since Liu \cite{Liu2007} first stated its growing importance back in 2007.
-Only several of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient Learning to Rank through the use of parallel programming paradigms.\\
+Only a few of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient Learning to Rank through the use of parallel programming paradigms.\\
 
-MapReduce \cite{Dean2004} is a parallel programming framework that is inspired by the \emph{Map} and \emph{Reduce} functions commonly used in functional programming.
-Since Google developed the MapReduce parallel programming framework back in 2004 it has grown to be the industry standard model for parallel programming.
+MapReduce \cite{Dean2004} is a parallel computing model that is inspired by the \emph{Map} and \emph{Reduce} functions that are commonly used in the field of functional programming.
+Since Google developed the MapReduce parallel programming framework back in 2004, it has grown to be the industry standard model for parallel programming.
+The release of Hadoop, an open-source version of MapReduce system that was already in use at Google, contributed greatly to MapReduce becoming the industry standard way of doing parallel computation.\\
+
 Lin \cite{Lin2013} observed that algorithms that are of iterative nature, which most Learning to Rank algorithms are, are not amenable to the MapReduce framework.
-Lin argued that as a solution to the non-amenability of iterative algorithms to the MapReduce framework, iterative algorithms can often be replaced with non-iterative alternatives or can still be optimized in such a way that its performance in a MapReduce setting is good enough.
-Alternative programming models are argued against by Lin, as they lack the critical mass as the data processing framework of choice.\\
+Lin argued that as a solution to the non-amenability of iterative algorithms to the MapReduce framework, iterative algorithms can often be replaced with non-iterative alternatives or by iterative alternatives that need fewer iterations, in such a way that its performance in a MapReduce setting is good enough.
+Alternative programming models are argued against by Lin, as they lack the critical mass as the data processing framework of choice and are as a result not worth their integration costs.\\
 
-The appearance of benchmark data sets gave insight in the performance of different Learning to Rank approaches, which resulted in increasing popularity of those methods that showed to perform well on one or more benchmark data sets.
-Up to now it remains unknown whether popular existing Learning to Rank methods scale well when they are used in a parallel manner using the MapReduce framework.
+The appearance of benchmark data sets for Learning to Rank gave insight in the ranking accuracy of different Learning to Rank methods.
+As observed by Liu \cite{Liu2007} and the \emph{Yahoo!
+Learning to Rank Challenge team} \cite{Chapelle2011b}, scalability of these Learning to Rank methods to large chunks of data is still an underexposed area of research.
+Up to now it remains unknown whether the Learning to Rank methods that perform well in terms of ranking accuracy also perform well in terms of scalability when they are used in a parallel manner using the MapReduce framework.
 This thesis aims to be an exploratory start in this little researched area of parallel Learning to Rank.\\

### 2014-09-24 report/Chapters/Introduction.tex
beter

@@
 \label{sec:goals}
-The objective of this thesis is to explore the speed-up in execution time of Learning to Rank algorithms through parallelisation using the MapReduce framework.
-This work focuses on those Learning to Rank algorithms that have shown leading performance on relevant benchmark data sets.
+The objective of this thesis is to explore the speed-up in execution time of Learning to Rank algorithms through parallelisation using the MapReduce computational model.
+The set of Learning to Rank models described in literature is of such size that it is infeasible to conduct an exhaustive experiment on all Learning to Rank models.
+Therefore we set the scope of our scalability experiment to include those Learning to Rank algorithms that have shown leading performance on relevant benchmark data sets.
 This thesis addresses the following research questions:
 \begin{description}
-\item[RQ1] What are the best performing Learning to Rank algorithms in terms of accuracy on relevant benchmark data sets?\\
+\item[RQ1] What are the best performing Learning to Rank algorithms in terms of ranking accuracy on relevant benchmark data sets?\\
 

### 2014-09-24 report/Chapters/Introduction.tex
research goals

@@
 \begin{description}
-\item[RQ1] What are the best performing Learning to Rank algorithms in terms of ranking accuracy on relevant benchmark data sets?\\
+\item[RQ1] What are the best performing Learning to Rank algorithms in terms of ranking accuracy on relevant benchmark data sets?
+\end{description}
+There are several existing metrics that can be used to express the accuracy of a ranking, which we will explore in section \ref{sec:how_to_evaluate_a_ranking}.
+\begin{description}
+\item[RQ2] What is the speed-up of those Learning to Rank algorithms when executed using the MapReduce framework?
+\end{description}
+With multiple existing definitions of speed-up, we will use the speed-up definition known as \emph{relative speed-up} \cite{Sun1991}, which is formulated as follows:\\
 
-\item[RQ2] What is the speed-up of those Learning to Rank algorithms when executed using the MapReduce framework?\\
-Where the definition of \emph{relative speed-up} is used for speed-up \cite{Sun1991}:\\
+$S_N = \frac{\text{execution time using one core}}{\text{execution time using \emph{N} cores}}$\\
 
-$S_N = \frac{\text{execution time using one core}}{\text{execution time using \emph{N} cores}}$
-\end{description}
+The single core execution time in this formula should be measured by used the fastest known single-machine implementation of the algorithm.
 

### 2014-09-24 report/Chapters/BenchmarkResults.tex
nDCG -> NDCG fix

@@
 \item Differences in feature sets between data sets detract from fair comparison
-\item Although the \ac{nDCG} definition is unambiguous, Busa-Fekete et al. \cite{Busa-Fekete2012} found that \ac{nDCG} evaluation tools of benchmark data sets produced different scores
+\item Although the \ac{NDCG} definition is unambiguous, Busa-Fekete et al. \cite{Busa-Fekete2012} found that \ac{NDCG} evaluation tools of benchmark data sets produced different scores
 \end{enumerate}
@@
 \end{table}\\
-Both \ac{nDCG} and \ac{ERR} are measured as performance metrics, but the final standings of the challenge were based on the \ac{ERR} values.
+Both \ac{NDCG} and \ac{ERR} are measured as performance metrics, but the final standings of the challenge were based on the \ac{ERR} values.
 Model validation on the Learning to Rank methods participating in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}.
@@
 
-Although the winning \ac{GBDT} models performs very well in terms of \ac{nDCG} their high complexity makes them unsuitable to use in production.
+Although the winning \ac{GBDT} models performs very well in terms of \ac{NDCG} their high complexity makes them unsuitable to use in production.
 The winning models take weeks to train and are very slow during query evaluation.
@@
 	\item AdaRank-MAP \cite{Xu2007}
-	\item AdaRank-nDCG \cite{Xu2007}
+	\item AdaRank-NDCG \cite{Xu2007}
 	\end{itemize}}
@@
 AdaRank-\acs{MAP} & 0.4385 & 0.1940 & 0.4063 \\ 
-AdaRank-\acs{nDCG} & 0.4369 & 0.2702 & 0.3878 \\ 
+AdaRank-\acs{NDCG} & 0.4369 & 0.2702 & 0.3878 \\ 
 \acs{MHR} & 0.4423 &  &  \\ 
 \end{tabular}
-\caption{\acs{nDCG}@10 results of the baseline methods on LETOR 2.0}
+\caption{\acs{NDCG}@10 results of the baseline methods on LETOR 2.0}
 \label{tab:letor2.0_results}
@@
 
-The evaluation metrics used are \ac{nDCG} and \ac{MAP}.
+The evaluation metrics used are \ac{NDCG} and \ac{MAP}.
 The \emph{winning number} metric is defined as the number of other algorithms that it can beat over all of the seven data sets (six .gov sets + OHSUMED).
@@
 	\item AdaRank-MAP \cite{Xu2007}
-	\item AdaRank-nDCG \cite{Xu2007}
+	\item AdaRank-NDCG \cite{Xu2007}
 	\item \ac{SVM}$^{MAP}$ \cite{Yue2007} 
@@
 The LETOR paper \cite{Qin2010} describes the performance of the LETOR baseline methods.
-Figures \ref{fig:ndcg_winning_number} and \ref{fig:map_winning_number} show ListNet to be the best performing baseline in terms of winning number on both \ac{nDCG} and \ac{MAP}.
+Figures \ref{fig:ndcg_winning_number} and \ref{fig:map_winning_number} show ListNet to be the best performing baseline in terms of winning number on both \ac{NDCG} and \ac{MAP}.
 The LETOR website lists\footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} a few algorithms that have since been evaluated on the LETOR benchmark collection.\\
@@
 \cite{Qin2010}, as listed on the LETOR website\footnotemark[2] by comparing their performance with the ListNet baseline.
-We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven data sets in \ac{nDCG} value.
-Note that this does not necessarily imply that these methods would have scored a higher \ac{nDCG} winning number than ListNet.
-Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR data sets in terms of \ac{nDCG}@10 and \ac{MAP}.
+We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven data sets in \ac{NDCG} value.
+Note that this does not necessarily imply that these methods would have scored a higher \ac{NDCG} winning number than ListNet.
+Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR data sets in terms of \ac{NDCG}@10 and \ac{MAP}.
 \begin{table}[!h]
 \begin{tabular}{l|ll}
- & \ac{nDCG}@10 & \ac{MAP} \\ 
+ & \ac{NDCG}@10 & \ac{MAP} \\ 
  \hline
@@
 \end{tabular}
-\caption{\acs{nDCG}@10 comparison of algorithms recently evaluated on LETOR 3.0 with the ListNet baselines}
+\caption{\acs{NDCG}@10 comparison of algorithms recently evaluated on LETOR 3.0 with the ListNet baselines}
 \label{tbl:LETOR_recently_added}
@@
 \subsubsection{Results}
-Rank\ac{SVM}-Struct, ListNet, AdaRank-\ac{nDCG}, AdaRank-\ac{MAP} and RankBoost were used as baseline models on the LETOR 4.0 data set and were implemented and evaluated by the publishers of LETOR 4.0.
+Rank\ac{SVM}-Struct, ListNet, AdaRank-\ac{NDCG}, AdaRank-\ac{MAP} and RankBoost were used as baseline models on the LETOR 4.0 data set and were implemented and evaluated by the publishers of LETOR 4.0.
 Table \ref{tab:letor4_baseline_results} shows the performance of those baseline models on the LETOR 4.0 benchmark collection.\\
@@
 \begin{tabular}{l|lll}
-Model & Data set & Mean \ac{nDCG} & \ac{nDCG}@10 \\ 
+Model & Data set & Mean \ac{NDCG} & \ac{NDCG}@10 \\ 
 \hline
@@
 \hline
-AdaRank-\ac{nDCG} & MQ2007 & 0.4914 & 0.4369 \\ 
+AdaRank-\ac{NDCG} & MQ2007 & 0.4914 & 0.4369 \\ 
  & MQ2008 & 0.4950 & 0.2307 \\ 
@@
 
-BoostedTree model \cite{Kocsis2013} showed an \ac{nDCG} of 0.5071 on the MQ2007 data set and thereby beat all baseline models.
+BoostedTree model \cite{Kocsis2013} showed an \ac{NDCG} of 0.5071 on the MQ2007 data set and thereby beat all baseline models.
 
@@
 Microsoft published these data sets as unofficial successor to LETOR in June 2010, but no baseline evaluations were described by the MSLR-WEB10/30k team.
-The Microsoft Research website from which the data set is obtainable\footnote{http://research.microsoft.com/en-us/projects/mslr/download.aspx} also offers an evaluation script for \ac{nDCG} and \ac{MAP}.
+The Microsoft Research website from which the data set is obtainable\footnote{http://research.microsoft.com/en-us/projects/mslr/download.aspx} also offers an evaluation script for \ac{NDCG} and \ac{MAP}.
 The presence of an official evaluation script enables fair comparison of evaluation results from other researchers comparable.
@@
 \end{tabular}
-\caption{\acs{nDCG} results of the baseline methods on the WCL2R collection, obtained from \cite{Alcantara2010}}
+\caption{\acs{NDCG} results of the baseline methods on the WCL2R collection, obtained from \cite{Alcantara2010}}
 \label{tab:results_WCL2R}

### 2014-09-25 report/Chapters/Introduction.tex
resarch questions stuk beter ingeleid

@@
 The set of Learning to Rank models described in literature is of such size that it is infeasible to conduct an exhaustive experiment on all Learning to Rank models.
-Therefore we set the scope of our scalability experiment to include those Learning to Rank algorithms that have shown leading performance on relevant benchmark data sets.
-This thesis addresses the following research questions:
+Therefore, we set the scope of our scalability experiment to include those Learning to Rank algorithms that have shown leading performance on relevant benchmark data sets.\\
+
+The existence of multiple benchmark data sets for Learning to Rank makes it non-trivial to determine the best Learning to Rank methods in terms of ranking accuracy.
+Given two ranking methods, there might be non-agreement between evaluation results on different benchmarks on which ranking method is more accurate.
+Furthermore, given two benchmark data sets, the sets of Learning to Rank methods that are evaluated on these benchmark data sets might not be identical.
+To be able to scope our experiment on the speed-up of the most accurate Learning to Rank methods to the most accurate ranking methods we first need to identify what the most accurate Learning to Rank algorithms are.
+This brings us the the first research question: \\
+
 \begin{description}
@@
 \end{description}
+
 There are several existing metrics that can be used to express the accuracy of a ranking, which we will explore in section \ref{sec:how_to_evaluate_a_ranking}.
+
 \begin{description}

### 2014-09-25 report/Chapters/Introduction.tex
bigskips om RQs heen

@@
 \end{description}
-
-There are several existing metrics that can be used to express the accuracy of a ranking, which we will explore in section \ref{sec:how_to_evaluate_a_ranking}.
+\bigskip
+Ranking accuracy is an ambiguous concept, as there exists several metrics that can be used to express the accuracy of a ranking.
+We will explore several metrics for ranking accuracy in section \ref{sec:how_to_evaluate_a_ranking}.\\
 
@@
 \end{description}
+\bigskip
 With multiple existing definitions of speed-up, we will use the speed-up definition known as \emph{relative speed-up} \cite{Sun1991}, which is formulated as follows:\\

### 2014-09-25 report/Chapters/Introduction.tex
research questions hoofdstuk mooi

@@
 
+After determining the most accurate ranking methods, we perform speed-up experiment on distributed MapReduce implementations of those algorithms.
+We formulate this in the following research question: \\
+
 \begin{description}
@@
 
-The single core execution time in this formula should be measured by used the fastest known single-machine implementation of the algorithm.
+The single core execution time in this formula is defined as the time that the fastest known single-machine implementation of the algorithm takes to perform the execution.
 

### 2014-09-25 report/Chapters/Introduction.tex
methodology uitgebreid

@@
 \section{Approach}
-\subsection{Methodology for Research Question I}
+We will describe our research methodology on a per Research Question basis.
+Prior to describing the methodologies for answering the Research Questions, we will describe the characteristics of our search for related work.
+
+\subsection{Literature Study Methodology}
 A literature study will be performed to get insight in relevant existing techniques for large scale Learning to Rank. The literature study will be performed by using the following query:\\
@@
 For the same reason the terms \emph{parallel} and \emph{distributed} are included in the search query.
-Even though \emph{parallel} and \emph{distributed} are not always synonymous is all definitions, we are interested in both approaches in non-sequential data processing.\\
+Even though \emph{parallel} and \emph{distributed} are not always synonymous, we are interested in both approaches in non-sequential data processing.\\
 
 A one-level forward and backward reference search is used to find relevant papers missed so far.
-To handle the large volume of studies involved in the backward and forward reference search, relevance of the studies will be evaluated solely on the title of the study.\\
+To handle the large volume of studies involved in the backward and forward reference search, relevance of the studies will be evaluated solely on the title of the study.
+
+\subsection{Methodology for Research Question I}
+To answer our first research question we will identify the Learning to Rank benchmark data sets that are used in literature to report the ranking accuracy of new Learning to Rank methods.
+These benchmark data sets will be identified by observing the data sets used in the papers found in the previously described literature study.
+Based on the benchmark data sets found, a literature search for papers will be performed and a cross-benchmark comparison method will be formulated.
+This literature search and cross-benchmark comparison procedure will be described in detail in section \ref{chap:cross_comparison}.
 
@@
 
-To research the speed-up's dependence on the amount of data being processed, the training computations will be performed on data sets of varying sizes.
-The well-known benchmark collections LETOR 3.0, LETOR 4.0 and MSLR-WEB30/40K offer a good initial set of data sets.
-Table \ref{tbl:initial_datasets} shows the
-% TODO: Waarom .Gov uit LETOR 3.0 niet?%
+To research the speed-up's dependence on the amount of data that is being processed, the training computations will be performed on data sets of varying sizes.
+We use the well-known benchmark collections LETOR 3.0, LETOR 4.0 and MSLR-WEB30/40K as a starting set of data sets for our experiments.
+Table \ref{tbl:initial_datasets} shows the data sizes of these data sets.
+The data sizes reported are not the total on-disk sizes of the data sets, but instead the size of the largest training set of all data folds (for an explanation of the concept of data folds, see \ref{sec:cross_validation}).
+
 \begin{table}[!h]
@@
 
+MSLR-WEB30K is the largest in data size of the benchmark data sets used in practice, but 2.62GB is still relatively small for MapReduce data processing.
 To test the how the computational performance of Learning to Rank algorithms both on cluster and on single-node computation scales to large quantities of data, larger data sets will be created by cloning the MSLR-WEB30K data set such that the cloned queries will hold new distinct query ID's.

### 2014-09-25 report/Chapters/Introduction.tex
thesis overview klopt weer

@@
 \begin{description}
-\item[Chapter II: Background]{introduces the reader to the basic principles and recent work in the fields of Learning to Rank.}
-\item[Chapter III: Related Work]{concisely describes existing work in the field of parallel Learning to Rank.}
-\item[Chapter IV: Benchmark Results]{sketches the accuracy of existing Learning to Rank methods on several benchmark data sets and describes the selection of Learning to Rank methods for the parallelisation experiments.}
-\item[Chapter V: Selected Learning to Rank Methods]{describes the algorithms and details of the Learning to Rank methods selected in Chapter IV.}
-\item[Chapter VI: Implementation]{describes implementation details of the Learning to Rank algorithms in the Hadoop framework.}
-\item[Chapter VII: Results \& Discussion]{presents and discusses speed-up results for the implemented Learning to Rank methods.}
-\item[Chapter VIII: Conclusion]{summarizes the results and answers our research questions based on the results.
+\item[Chapter II: Background]{introduces the reader to the basic principles and recent work in the fields of Learning to Rank and the MapReduce computing model.
+Aims at bringing those unfamiliar with those topics up to speed to be able to understand the succeeding chapters of this thesis.}
+\item[Chapter III: Related Work]{concisely describes existing work in the field of parallel and distributed Learning to Rank.}
+\item[Chapter IV: Benchmark Data Sets]{described the characteristics of the existing benchmark data sets in the Learning to Rank field.}
+\item[Chapter V: Cross Benchmark Comparison]{described the methodology of a comparison of ranking accuracy of Learning to Rank methods across benchmark data sets and described the results of this comparison.}
+\item[Chapter VI: Selected Learning to Rank Methods]{describes the algorithms and details of the Learning to Rank methods selected in Chapter V.}
+\item[Chapter VII: Implementation]{describes implementation details of the Learning to Rank algorithms in the Hadoop framework.}
+\item[Chapter VIII: MapReduce Experiments]{presents and discusses speed-up results for the implemented Learning to Rank methods.}
+\item[Chapter IX: Conclusions]{summarizes the results and answers our research questions based on the results.
 The limitations of our research as well as future research directions in the field are mentioned here.}
+\item[Chapter X: Future Work]{describes several directions of research worthy follow-up research based on our findings.}
 \end{description}

### 2014-09-25 report/Chapters/Background.tex
research goals
methodology uitgebreid
tweaks aan background section

@@
 %************************************************
-This chapter is written with the goal to provide the reader with a subtle introduction in the Learning to Rank field.
+This chapter is written with the goal to provide the reader with a subtle introduction in the Learning to Rank field and the MapReduce computational model.
 Models and theories explained in this chapter can be regarded as prior knowledge needed to understand the subsequent chapters of this thesis.
@@
 \end{figure}\\
-Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning to Rank method when it is \emph{feature based} and uses \emph{discriminative training}, which are itself defined as follows:
+Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning to Rank method when it is \emph{feature based} and uses \emph{discriminative training}, in which the concepts \emph{feature based} and \emph{discriminative training} are itself defined as:
 \begin{description}
@@
 \end{enumerate}
-These three approaches are in literature respectively called the pointwise approach, the pairwise approach and the listwise approach.
+These three approaches are in literature respectively labelled the pointwise approach, the pairwise approach and the listwise approach.
 These three approaches to Learning to Rank will be described in more detail in section \ref{sec:ltr_approaches}.
@@
 \section{How to evaluate a ranking}
+\label{sec:how_to_evaluate_a_ranking}
 Evaluation metrics have long been studied in the field of information retrieval.
@@
 Different benchmarking settings have used different evaluation metrics.
-Metrics introduced in this section will be used to express the evaluation results in chapter \ref{chap:benchmark_results} of this thesis.
+Metrics introduced in this section will be used in chapters \ref{chap:benchmark_results} and \ref{chap:cross_comparison} of this thesis to compare Learning to Rank methods in terms of ranking accuracy.
 \subsection{Normalized Discounted Cumulative Gain}
-Cumulative gain, or its predecessor discounted cumulative gain and normalized discounted cumulative gain, is one of the most widely used measures for effectiveness of ranking methods.
+Cumulative gain, or its successor discounted cumulative gain and normalized discounted cumulative gain, is one of the most widely used measures for effectiveness of ranking methods.
 \subsubsection{Discounted Cumulative Gain}
@@
  &  &  &  &  &  &  &  &  &  &  &  \\
- & \multicolumn{10}{c}{J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} \ac{NDCG} = $\frac{29.17}{29.96} = 0.97$} &  \\  
- & \multicolumn{10}{c}{Burges \cite{Burges2005} et al. \ac{NDCG} = $\frac{732.7}{788.0} = 0.93$} &  \\ 
+ & \multicolumn{10}{c}{Burges \cite{Burges2005} et al. \ac{NDCG} = $\frac{732.7}{788.0} = 0.9298$} &  \\ 
+  & \multicolumn{10}{c}{J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} \ac{NDCG} = $\frac{29.17}{29.96} = 0.9736$} &  \\
 \end{tabular}
@@
 
-%\subsection{Concluding notes on evaluation measures}
-%TODO: Ingaan op verbanden/correlaties tussen metrics.
-
 \section{Approaches to Learning to Rank}
@@
 \section{Cross-validation experiments}
+\label{sec:cross_validation}
 TODO

### 2014-09-25 report/Chapters/BenchmarkResults.tex
chapter 4 feedback verwerkt

@@
 \item Differences in feature sets between data sets detract from fair comparison
-\item Although the \ac{NDCG} definition is unambiguous, Busa-Fekete et al. \cite{Busa-Fekete2012} found that \ac{NDCG} evaluation tools of benchmark data sets produced different scores
+\item Although the \ac{NDCG} definition is unambiguous (note: there are two existing versions of \ac{NDCG}: J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} and Burges et al.
+\cite{Burges2005}, but both are in itself defined unambiguously), Busa-Fekete et al.
+\cite{Busa-Fekete2012} found that \ac{NDCG} evaluation tools of different benchmark data sets produced different scores
 \end{enumerate}
+% TODO: hoe zit dit? Komt de vondst van Busa-Fekete door het Burges and Jarvelin verschil? %
 
@@
 AdaRank-\acs{NDCG} & 0.4369 & 0.2702 & 0.3878 \\ 
-\acs{MHR} & 0.4423 &  &  \\ 
+\acs{MHR} & 0.4423 & - & - \\ 
 \end{tabular}
@@
 \subsection{WCL2R}
-The WCL2R collection, released by Alc{\^a}ntara et al \cite{Alcantara2010}, contains of two data sets from the Chilean search engine TodoCL\footnote{www.todocl.cl}.
+The WCL2R collection, released by Alc{\^a}ntara et al.
+\cite{Alcantara2010}, contains of two data sets originating from the Chilean search engine TodoCL\footnote{www.todocl.cl}.
 Both data sets contain approximately 3 million documents.

### 2014-09-25 report/Chapters/BenchmarkResults.tex
I -> we

@@
 Learning to Rank Challenge consists of two tracks that uses the two data sets respectively: a standard Learning to Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country.
-For this experiment, I will only look at the standard Learning to Rank data set, because transfer learning is a separate research area that is not included in this thesis.\\
+For this experiment, we will only look at the standard Learning to Rank data set, because transfer learning is a separate research area that is not included in this thesis.\\
 \begin{table}[!h]
@@
 
-I will describe the performance of the algorithms that were evaluated by the LETOR team on LETOR 3.0 after publication of the LETOR paper by Qin et al.
+We will describe the performance of the algorithms that were evaluated by the LETOR team on LETOR 3.0 after publication of the LETOR paper by Qin et al.
 \cite{Qin2010}, as listed on the LETOR website\footnotemark[2] by comparing their performance with the ListNet baseline.

### 2014-09-25 report/Chapters/CrossBenchmarkComparison.tex
I -> we

@@
 Learning to Rank Challenge data sets and the OHSUMED dataset from LETOR 3.0.
-To my knowledge, no structured meta-analysis on ranking accuracy has been conducted where evaluation results on several benchmark collections are taken into account.
+To our knowledge, no structured meta-analysis on ranking accuracy has been conducted where evaluation results on several benchmark collections are taken into account.
 With a meta-analysis we will compare the performance of Learning to Rank methods across the Learning to Rank benchmark data sets described in foregoing sections.
@@
 \begin{description}
-\item[RQ1] What are the best performing Learning to Rank algorithms in terms of accuracy on relevant benchmark data sets?\\
+\item[RQ1] What are the best performing Learning to Rank algorithms in terms of ranking accuracy on relevant benchmark data sets?\\
 \end{description}

### 2014-09-25 report/Chapters/LTRMethods.tex
I -> we

@@
 SmoothRank enables direct optimisation of the non-convex and discontinuous evaluation measure by smoothing, that is, approximating the rank position.
-In this section, I apply the SmoothRank approximation method to the \ac{NDCG} evaluation measure for illustration, but the same procedure can be applied to \ac{MAP} or any other information retrieval measure.
+In this section, we apply the SmoothRank approximation method to the \ac{NDCG} evaluation measure for illustration, but the same procedure can be applied to \ac{MAP} or any other information retrieval measure.
 The smoothing function used in SmoothRank is based on the softmax activation function \cite{Bridle1990}, which is often used in neural networks.

### 2014-09-25 report/Chapters/RelatedWork.tex
I -> we

@@
 An abstract-based manual filtering step is applied where those results are filtered that use the terms \emph{parallel} or \emph{distributed} in context to \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}.
-As a last step I will filter out studies based on the whole document that only focus on efficient query evaluation and not on parallel or distributed learning of ranking functions, as those studies are likely to meet listed search terms.\\
+As a last step we will filter out studies based on the whole document that only focus on efficient query evaluation and not on parallel or distributed learning of ranking functions, as those studies are likely to meet listed search terms.\\
 

### 2014-09-25 report/Chapters/CrossBenchmarkComparison.tex
hele verslag doorgelopen

@@
 \begin{enumerate}
-\item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011, Lin2012}
+\item A different evaluation methodology was used in the study compared to what was used in other studies using the same benchmark \cite{Geng2011, Lin2012}.
 \item The study focussed on a different Learning to Rank task (e.g.
-rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}
-\item The study used an altered version of a benchmark that contained additional features \cite{Bidoki2009, Ding2010}
+rank aggregation or transfer ranking) \cite{De2011, De2010, Derhami2013, De2012, Chen2010, Ah-Pine2008, Wang2009c, De2013, Miao2013, Hoi2008, De2012b, Duh2011b, Argentini2012, Qin2010c, Volkovs2013, Desarkar2011, Pan2013, Lin2011b, Volkovs2012, Dammak2011}.
+\item The study used an altered version of a benchmark that contained additional features \cite{Bidoki2009, Ding2010}.
 \item The study provides no exact data of the evaluation results (e.g.
-results are only in graphical form) \cite{Wang2008, Wang2010, Xu2010, Kuo2009, Li2008, Xia2008, Zhou2011, Wu2011, Zhu2009, Karimzadehgan2011, Swersky2012, Pan2011, Ni2008, Ciaramita2008, Stewart2012, Petterson2009, Agarwal2010, Chang2009, Qin2008c, Adams2011, Sculley2009, Huang2008, Alejo2010, Sun2011, He2010b, Benbouzid2012, Geng2012, Chen2012, Xu2012, Shivaswamy2011}
-\item The study reported evaluation results in a different metric than the metrics chosen for this meta-analysis \cite{Yu2009, Thuy2009, Pahikkala2009, Kersting2009, Mohan2011}
-\item The study reported a higher performance on baseline methods than official benchmark runs \cite{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}
+results are only in graphical form) \cite{Wang2008, Wang2010, Xu2010, Kuo2009, Li2008, Xia2008, Zhou2011, Wu2011, Zhu2009, Karimzadehgan2011, Swersky2012, Pan2011, Ni2008, Ciaramita2008, Stewart2012, Petterson2009, Agarwal2010, Chang2009, Qin2008c, Adams2011, Sculley2009, Huang2008, Alejo2010, Sun2011, He2010b, Benbouzid2012, Geng2012, Chen2012, Xu2012, Shivaswamy2011}.
+\item The study reported evaluation results in a different metric than the metrics chosen for this meta-analysis \cite{Yu2009, Thuy2009, Pahikkala2009, Kersting2009, Mohan2011}.
+\item The study reported a higher performance on baseline methods than official benchmark runs \cite{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}.
+Such cases are not necessarily caused by malicious intent, but might be caused by a mix-up of the two existing versions of the \ac{NDCG} metric (J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} and Burges et al.
+\cite{Burges2005}).
 \item The study did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.

### 2014-09-25 report/Chapters/Implementation.tex
hele verslag doorgelopen

@@
 Newly supported programming models since Hadoop 2.0 include Dryad \cite{Isard2007}, Giraph \cite{Avery2011}, MPI, REEF \cite{Chun2013}, Spark \cite{Zaharia2010}, and Storm \cite{Aniello2013}.
-Even though these programming models are now supported by Hadoop and some of these programming models have recently increased in popularity, they still lack the critical mass as the data processing framework of choice that MapReduce has as Lin argued back in 2012 \cite{Lin2013}(see section \ref{sec:motivation}).
+Even though these programming models are now supported by Hadoop and some of these programming models have recently increased in popularity, they still lack the critical mass as the data processing framework of choice that MapReduce has as Lin argued back in 2012 \cite{Lin2013} (see section \ref{sec:motivation}).
 Therefore, even though some of these programming models might be a better fit for iterative algorithms, we use Hadoop MapReduce of the programming model to implement the Learning to Rank algorithms.\\

### 2014-09-25 report/Chapters/Background.tex
cross validation

@@
 \label{sec:cross_validation}
-TODO
+A cross-validation experiment \cite{kohavi1995}, sometimes called rotation estimation, is an experimental set-up for model evaluation where the data is split into $k$ chunks of approximately equal size, called \emph{folds}.
+One of the folds is used as validation set and one of the folds is used as test set, the rest of the $k-2$ folds are used as training data.
+This procedure is repeated $k$ times, such that each fold is once used for validation, once as test set, and $k-2$ times as training data.
+The performance can be measured in any model evaluation metric, and is averaged over the model performances on each of the folds.
+The goal of cross-validation is to define a data set to test the model in the training phase, in order to limit the problem of overfitting.\\
+
+Cross-validation is one of the most frequently used model evaluations methods in the field of Machine Learning, including the Learning to Rank subfield.
+Often, folds in a cross-validation are created in a \emph{stratified} manner, meaning that the folds are created in such a way that the distributions of the target variable are approximately identical between the folds.
 

### 2014-09-25 report/Chapters/Background.tex
mapreduce kort uitgelegd

@@
 \section{An introduction to the MapReduce programming model}
-TODO
+MapReduce \cite{Dean2004} is a programming model invented at Google, where users specify a \emph{map} function that processes a key/value pair to generate a set of intermediate key/value pairs, and a \emph{reduce} function that merges all intermediate values associated with the same intermediate key.
+This model draws its inspiration from the field of functional programming, where \emph{map} and \emph{reduce} are commonly used functions.\\
+
+This combination of the \emph{map} and \emph{reduce} functions allows for parallel computation.
+In the \emph{map} phase parallel computation can be performed by simply splitting the input data after a certain number of bytes, where each worker nodes performs the user-specified \emph{map}-function on its share of the data.
+Before the \emph{reduce} phase these intermediate answers of the different worker nodes are transformed in such a way that they are grouped by key value, this is called the shuffle-phase.
+After the shuffle-phase, the user-defined \emph{reduce}-function is applied to each group of key/value pairs in the reduce phase.
+Since the key/value pairs are already grouped by key in the shuffle phase, this \emph{reduce}-function can be applied to a group of key/value pairs on any of the worker nodes.\\

### 2014-09-26 report/FrontBackmatter/Titlepage.tex
Bosschaart's voorkantje

%*******************************************************
% Titlepage
%*******************************************************
\begin{titlepage}
\title{Master's thesis Computer Science\\ (Chair: Databases, Track: Information System Engineering)}
\course{Scaling Learning to Rank to Big Data}
\subtitle{A study concerning parallelisation of Learning to Rank algorithms using the MapReduce computing model}
\university{University of Twente}
\faculty{Faculty of Electrical Engineering, Mathematics and Computer Science (EEMCS)}
\author{Niek Tax (s0166197)}

\committee{Dr. ir. Djoerd Hiemstra (UT/Databases)\\
Sander Bockting, Msc (Avanade Netherlands B.V.)\\
}	
\date{\today}
\num{}

\supervisors{Dr. ir. Djoerd Hiemstra (UT/Databases)\\
Sander Bockting (Avanade Netherlands B.V.)}    
\end{titlepage}
\maketitle

### 2014-09-26 report/preamble.tex
Bosschaart's voorkantje

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\theoremstyle{definition}
\newtheorem{defin}[thm]{Definition}

\lstset{ %
basicstyle=\footnotesize,       % the size of the fonts that are used for the code
numbers=left,                   % where to put the line-numbers
numberstyle=\footnotesize,      % the size of the fonts that are used for the line-numbers
stepnumber=2,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
frame=single,			% adds a frame around the code
tabsize=2,			% sets default tabsize to 2 spaces
captionpos=b,			% sets the caption-position to bottom
breaklines=true,		% sets automatic line breaking
breakatwhitespace=false,	% sets if automatic breaks should only happen at whitespace
escapeinside={\%*}{*)}          % if you want to add a comment within your code
}
%\renewcommand{\bibname}{Recommended Reading}
\DeclareGraphicsExtensions{.pdf,.png}
\setcounter{secnumdepth}{3}
\dominitoc

### 2014-09-29 report/Chapters/Conclusions.tex
conclusie een stukje mooier

@@
-Using our experimental results we will reflect on our research questions stated in the \ref{sec:goals} section of this thesis. We formulated the following research questions:
+Using our experimental results we will now reflect on our research questions stated in the \ref{sec:goals} section of this thesis. We formulated the following research questions:
 \begin{description}
-\item[RQ1] What are the best performing Learning-to-Rank algorithms in terms of accuracy on relevant benchmark datasets?
+\item[RQ1] What are the best performing Learning to Rank algorithms in terms of ranking accuracy on relevant benchmark data sets?
 \end{description}
@@
 \begin{description}
-\item[RQ2] What is the speed-up of those Learning-to-Rank algorithms when executed using the MapReduce framework?\\
+\item[RQ2] What is the speed-up of those Learning to Rank algorithms when executed using the MapReduce framework?
+\end{description}
+\bigskip
 Where the definition of \emph{relative speed-up} is used for speed-up \cite{Sun1991}:\\
 
-$S_N = \frac{\text{execution time using one core}}{\text{execution time using \emph{N} cores}}$
-\end{description}
-To answer this research question we took the approach to implementing the list of algorithms found in answering the first research question starting with the Learning to Rank method with the highest certainty on ranking accuracy, ListNet.
+$S_N = \frac{\text{execution time using one core}}{\text{execution time using \emph{N} cores}}$\\
+
+To answer this research question, we took the approach of implementing the list of algorithms found in answering the first research question, starting with the Learning to Rank method with the highest certainty on ranking accuracy, ListNet.\\
+
 We found that running ListNet on a Hadoop cluster using the MapReduce computing model comes with its own cost in the form of a job scheduling overhead in the range of 150-200 seconds per training iteration.
-However, with a normalisation preprocessing iteration we can greatly reduce the number of training iterations needed for converge.
+This makes Hadoop very inefficient for the processing of small data sets, where the Hadoop overhead tend to make up a large share of the total processing time.
+For small data sets where the constant 150-200 seconds job scheduling overhead is a large fraction of the total processing time, single-machine computation, which does not have this job scheduling overhead, is found to be faster than Hadoop MapReduce computation.
+For large data sets, where 150-200 seconds overhead per iteration is small compared to the total time that it would take to process the data, Hadoop MapReduce can provide a speed-up to the training process.
 ListNet on a single machine does not scale well to data sizes larger than the physical memory size.
-To process data sets in the 100+GB range with the ListNet training algorithm, Hadoop MapReduce is a large improvement compared to a single machine.\\
-% TODO: iets zeggen over waarom de oorspronkelijk speed-up metric niet zo geschikt is hier %
-No generalisations can be drawn from this results to other Learning to Rank algorithms.
-Although we can extend our findings on job scheduling overhead and scaling benefit from data larger than 100+GB to the gradient descent procedure that is used in ListNet as well as in many other Learning to Rank algorithms and in many learning algorithms in general.
-Other Learning to Rank algorithms using the gradient descent procedure might not scale well when the MapReduce computing model is used, but any bad scaling behaviour on MapReduce of Learning to Rank algorithms will not be because of gradient descent.
+To process large data sets with the ListNet training algorithm, Hadoop MapReduce is a large improvement compared to a single machine.\\
+
+Moreover, we found that the addition of a normalisation preprocessing step to the ListNet procedure can greatly improve the ranking accuracy of the ListNet training procedure after the first five iterations.
+This suggests that the addition of a normalisation preprocessing step can reduce the number of training iterations needed for convergence.
+Lin stated in his essay \cite{Lin2013} that MapReduce is often good enough for tasks that are not-amenable to the MapReduce model.
+Lin motivates this statement in the context of iterative algorithms with the observation that these iterative algorithms can often be optimised in such a way that less iterations are needed for convergence.
+Our preprocessing step can improve the convergence rate of the ListNet training iteration, and therefore fits into Lin's point of view.\\
+
+No generalisations can be drawn from these results to the scalability on MapReduce of other Learning to Rank algorithms.
+However, we can extend our findings on job scheduling overhead and scaling benefit on large data sets to the gradient descent procedure that is used in ListNet as well as in many other Learning to Rank algorithms and in many learning algorithms in general.
+Other Learning to Rank algorithms using the gradient descent procedure might not scale well when the MapReduce computing model is used, but any bad scaling behaviour on MapReduce of Learning to Rank algorithms will not be caused by bad scaling of the gradient descent procedure.

### 2014-09-29 report/Chapters/Future.tex
naar refactoring werk
future work puntje

@@
-\subsection{Computing models}
-We explored the possibilities of Hadoop MapReduce for computation
+\section{Computing models}
+We explored the possibilities of Hadoop MapReduce for distributed computation of Learning to Rank training algorithms.
+Hadoop, since the introduction of Hadoop YARN in Hadoop 2.0, offers integration with other distributed computing models, including Spark \cite{}, Storm \cite{} and MPI \cite{}.
+Of these distributed programming models, Spark is particularly promising, since Shukla et al.
+\cite{Shukla2012} already showed that it is able to speed-up the ListNet training procedure with much lower job scheduling overhead than that we found for MapReduce.
 
-\subsection{Learning to Rank algorithms}
+\section{Learning to Rank algorithms}

### 2014-09-29 report/Chapters/CrossBenchmarkComparison.tex
nog een nDCG gevonden
crossbenchmark comparison verbeterd vanuit IPM versie

@@
 To compare algorithms based on a sparse set of evaluation measurements, a normalised version of the Winning Number metric will be used.
-This Normalised Winning Number takes only those data sets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all data sets on which it has been evaluated.
+This \ac{NWN} takes only those data sets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all data sets on which it has been evaluated.
 We will redefine the indicator function $I$ in order to only take into account those data sets that an algorithm is evaluated on, as \\
@@
 
-From now on this adjusted version of Winning Number will be references to as \emph{Normalised Winning Number}. The mathematical definition of Normalised Winning Number is\\
+From now on this adjusted version of Winning Number will be references to as \ac{NWN}. The mathematical definition of \ac{NWN} is\\
 
@@
 \noindent
-where Ideal Winning Number is defined as\\
+where \ac{IWN} is defined as\\
 
@@
 \section{Results \& Discussion}
-The following subsections provide the performance of Learning to Rank methods in terms of Normalised Winning Number for \ac{NDCG}@3, \ac{NDCG}@5, \ac{NDCG}@10 and \ac{MAP}.
-Performance of the Learning to Rank methods is plotted with Normalised Winning Number on the vertical axis and the number of data sets on which the method has been evaluated on the horizontal axis.
+The following subsections provide the performance of Learning to Rank methods in terms of \ac{NWN} for \ac{NDCG}@3, \ac{NDCG}@5, \ac{NDCG}@10 and \ac{MAP}.
+Performance of the Learning to Rank methods is plotted with \ac{NWN} on the vertical axis and the number of data sets on which the method has been evaluated on the horizontal axis.
 The further to the right, the more certain we can be about the performance of the Learning to Rank method.
-The methods for which it holds that there is no other method that has 1) a higher Normalised Winning Number and 2) a higher number data sets evaluated, are identified as the best performing methods and are labelled with the name of the method.
+The methods for which it holds that there is no other method that has 1) a higher \ac{NWN} and 2) a higher number data sets evaluated, are identified as the best performing methods and are labelled with the name of the method.
 
 \subsection{NDCG@3}
-Figure \ref{fig:normalised_winning_number_NDCG3} shows the performance of Learning to Rank methods for the \ac{nDCG}@3 metric.
-Table \ref{tab:raw_data_norm_winnum_NDCG3} in Appendix \ref{app:norm_winnum_NDCG3} provides the raw Normalised Winning Number data for the Learning to Rank methods for which \ac{NDCG}@3 evaluation results were available.\\
+Figure \ref{fig:normalised_winning_number_NDCG3} shows the performance of Learning to Rank methods for the \ac{NDCG}@3 metric.
+Table \ref{tab:raw_data_norm_winnum_NDCG3} in Appendix \ref{app:norm_winnum_NDCG3} provides the raw \ac{NWN} data for the Learning to Rank methods for which \ac{NDCG}@3 evaluation results were available.\\
 
@@
 
-LambdaNeuralRank and CoList both acquired a perfect Normalised Winning Number score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
+LambdaNeuralRank and CoList both acquired a perfect \ac{NWN} score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
 LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 data sets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three data sets which in addition are more relevant data sets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 data sets (see Table \ref{tab:ltr_methods_used}).
-FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in Normalised Winning Number, but increasingly higher in number of data sets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
+FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in \ac{NWN}, but increasingly higher in number of data sets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
 
@@
 
-LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a Normalised Winning Number of 1.0.
-LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of data sets, but score decreasingly well in terms of Normalised Winning Number.
+LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a \ac{NWN} of 1.0.
+LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of data sets, but score decreasingly well in terms of \ac{NWN}.
 These results are highly in agreement with the \ac{NDCG}@3 comparison.
@@
 
-LambdaMART and LambdaNeuralRank score a Normalised Winning Number of 1.0 on the \ac{NDCG}@10 comparison.
+LambdaMART and LambdaNeuralRank score a \ac{NWN} of 1.0 on the \ac{NDCG}@10 comparison.
 For LambdaNeuralRank these results are again based on AOL dataset measurements.
 LambdaMART showed the highest \ac{NDCG}@10 performance for the MSLR-WEB10k dataset.
-The set of algorithms for which there is no other algorithm with both a higher Normalised Winning Number and number of data sets evaluated on is partly in agreement with those for the \ac{NDCG}@3 and \ac{NDCG}@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
+The set of algorithms for which there is no other algorithm with both a higher \ac{NWN} and number of data sets evaluated on is partly in agreement with those for the \ac{NDCG}@3 and \ac{NDCG}@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
 SmoothRank and FSMRank were not present in this set for the \ac{NDCG}@3 and \ac{NDCG}@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_NDCG3} and \ref{tab:raw_data_norm_winnum_NDCG5} in the Appendices.
@@
 Figure \ref{fig:normalised_winning_number_map} shows the performance of Learning to Rank methods for the \ac{MAP} metric.
-Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\
+Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw \ac{NWN} data for the Learning to Rank methods.\\
 
@@
 Where comparisons on the \ac{NDCG}-metric at different cut-off points where highly in agreement in terms of the best performing algorithms, the comparison in terms of \ac{MAP} shows different results.
-RankDE scores a Normalised Winning Number of 1.0 on one dataset, like LambdaNeuralRank did on for all \ac{NDCG}-comparisons.
+RankDE scores a \ac{NWN} of 1.0 on one dataset, like LambdaNeuralRank did on for all \ac{NDCG}-comparisons.
 In contrast to LambdaNeuralRank, RankDE achieved this highest score on the LETOR 2.0 TD2003, a dataset on which many methods are evaluated.\\
 
-LARF and LRUF score very high Normalised Winning Number scores, but based on only relatively few data sets, just as in the \ac{NDCG}-comparisons.
+LARF and LRUF score very high \ac{NWN} scores, but based on only relatively few data sets, just as in the \ac{NDCG}-comparisons.
 Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all \ac{NDCG}-comparisons, are not within the best performing methods in the \ac{MAP}-comparison.
-A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_NDCG3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_NDCG3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more data sets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher Normalised Winning Number.
-SmoothRank performed a Normalised Winning Number of around 0.53 over 7 data sets, which is still good in both certainty and accuracy, but not among the top methods.
+A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_NDCG3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_NDCG3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more data sets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher \ac{NWN}.
+SmoothRank performed a \ac{NWN} of around 0.53 over 7 data sets, which is still good in both certainty and accuracy, but not among the top methods.
 RE-QR is one of the best performers in the \ac{MAP}-comparison with a reasonable amount of benchmark evaluations.
 No reported \ac{NDCG} performance was found in the literature study for RE-QR.
-There is a lot of certainty on the accuracy of RankBoost and Rank\ac{SVM} as both models are evaluated on the majority of data sets included in the comparison for the \ac{MAP}-metric, but given their Normalised Winning Number it can said that both methods are not within the top performing Learning to Rank methods.
+There is a lot of certainty on the accuracy of RankBoost and Rank\ac{SVM} as both models are evaluated on the majority of data sets included in the comparison for the \ac{MAP}-metric, but given their \ac{NWN} it can said that both methods are not within the top performing Learning to Rank methods.
 
 \subsection{Cross-metric}
-Figure \ref{fig:normalised_winning_number_all} shows the Normalised Winning Number as function of Ideal Winning Number for the methods described in Table \ref{tab:ltr_methods_used}.
+Figure \ref{fig:normalised_winning_number_all} shows the \ac{NWN} as function of \ac{IWN} for the methods described in Table \ref{tab:ltr_methods_used}.
 Table \ref{tab:raw_data_norm_winnum_all} in Appendix \ref{app:norm_winnum_all} provide the raw data plotted in Figure \ref{fig:normalised_winning_number_all}.\\
@@
 
-LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both Ideal Winning Number and Normalised Winning Number.
+LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both \ac{IWN} and \ac{NWN}.
 LRUF is the only method that achieved this in all \ac{NDCG} comparisons, the \ac{MAP} comparison as well as the cross-metric comparison.
@@
 
-C-CRF, DirectRank, FP-Rank, RankCSA, LambdaNeuralRank and VFLR all have near-perfect Normalised Winning Number measures, but have low Ideal Winning Number measures.
+C-CRF, DirectRank, FP-Rank, RankCSA, LambdaNeuralRank and VFLR all have near-perfect \ac{NWN} measures, but have low \ac{IWN} measures.
 Further evaluation runs of these methods on benchmark data sets that they are not yet evaluated on are desirable.
@@
 \section{Limitations}
-In the Normalised Winning Number calculation the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
+In the \ac{NWN} calculation the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
 By calculating it in this way, we implicitly make the assumption that the Learning to Rank methods are (approximately) distributed uniformly over the benchmarks, such that the average Learning to Rank method tested are approximately equally hard for each data set.
@@
 
-Another limitation lies in the meta-analysis methodology.
-Using evaluation results published by other researchers relies on the honesty of those researchers.
-Overly optimistic evaluation results published by other researchers do affect our Normalised Winning Number results.
+Another limitation is that our comparison methodology relies on the correctness of the evaluation results found in the literature search step.
+This brings up a risk of overly optimistic evaluation results affecting our NWN results.
 Limiting the meta-analysis to those studies that report comparable results on one of the baseline methods of a benchmark set reduces this limitation but does not solve it completely.
-By taking the Ideal Winning Number into account in Figure \ref{fig:normalised_winning_number_all} we further mitigate this limitation, as the Ideal Winning Number is loosely related with the number of studies that reported evaluation results for an algorithm.
+By taking IWN into account in Figure \ref{fig:normalised_winning_number_all} we further mitigate this limitation, as IWN is loosely related with the number of studies that reported evaluation results for an algorithm.\\
+
+Our comparison regarded evaluation results on \ac{NDCG}@{3, 5, 10} and \ac{MAP}.
+By making the decision to include \ac{NDCG} at three cut-off points and only a single \ac{MAP} entry, we implicitly attain a higher weight for \ac{NDCG} compared to \ac{MAP} on an analysis that combines all measurements on the four metrics.
+This implicit weighting could be regarded as arbitrary, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.
+Note that another implicit weighting lies in the paper dimension.
+Hence, the higher number of evaluation results specified in a paper, the higher the influence of this paper on the outcome of the analysis.
+This implicit weighting is not harmful to the validity of our comparison, as papers with a large number of evaluation results are more valuable than papers with a few evaluation results.
+In addition, papers with a high number of evaluation results are not expected to be less reliable than papers with fewer evaluation results.
 
 \section{Conclusions}
-We will now look back on our first research question using the knowledge gained with the cross-benchmark comparison.
+We proposed a new way of comparing learning to rank methods based on sparse evaluation results data on a set of benchmark datasets.
+Our comparison methodology comprises of two components: 1) \ac{NWN}, which provides insight in the ranking accuracy of the learning to rank method, and 2) \ac{IWN}, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.\\
+
+Based on this this new comparison approach for a set of sparse evaluation results, we will now look back on the first research question of the thesis.
 \begin{description}
@@
 
+Based on our literature search for evaluation results on well-known benchmarks collections, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.
+However, no closing arguments can be formulated on which learning to rank methods are most accurate.
+LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the learning to rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
+From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.\\
+
 LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the Learning to Rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
@@
 
+More evaluation runs are needed for the methods on the left side of Figure \ref{fig:normalised_winning_number_all}.
+Our work contributes to this by identifying promising learning to rank methods that researchers could focus on in performing additional evaluation runs.\\
+
 In the following chapters of this thesis, concerning parallel execution of the Learning to Rank training phase, the scope will be limited to the five methods that turned out to be superior Learning to Rank methods in terms of ranking accuracy and certainty about this ranking accuracy: LRUF, FSMRank, FenchelRank, SmoothRank and ListNet.
-Although it can not be concluded that these methods are inarguably the most accurate Learning to Rank methods, a strong presumption has been raised that these five Learning to Rank are accurate ranking methods.
+Although it can not be concluded that these methods are inarguably the most accurate Learning to Rank methods, a strong presumption has been raised that these five Learning to Rank are accurate ranking methods.\\

### 2014-09-29 report/Chapters/Future.tex
crossbenchmark comparison verbeterd vanuit IPM versie

@@
 We explored the possibilities of Hadoop MapReduce for distributed computation of Learning to Rank training algorithms.
-Hadoop, since the introduction of Hadoop YARN in Hadoop 2.0, offers integration with other distributed computing models, including Spark \cite{}, Storm \cite{} and MPI \cite{}.
+Hadoop, since the introduction of Hadoop YARN in Hadoop 2.0, offers integration with other distributed computing models, including Dryad \cite{Isard2007}, Spark \cite{Zaharia2010}, Storm \cite{Aniello2013} and MPI.
 Of these distributed programming models, Spark is particularly promising, since Shukla et al.
 \cite{Shukla2012} already showed that it is able to speed-up the ListNet training procedure with much lower job scheduling overhead than that we found for MapReduce.
+For the newly supported programming models in Hadoop, it holds that they lack the critical mass as the distributed programming model of choice.
+This lack of a critical mass results in higher integration costs and less available support.
+The integration of new programming model into the Hadoop YARN framework is a step in the good direction to alleviate the integration cost of these programming models, but does not yet completely eliminate the problem.
+Cloud-based Hadoop services like Microsoft HDInsight and Amazon Elastic MapReduce.
 

### 2014-09-29 report/Chapters/CrossBenchmarkComparison.tex
mooie dingen

@@
 Figure \ref{fig:normalised_winning_number_NDCG3} shows the performance of Learning to Rank methods for the \ac{NDCG}@3 metric.
-Table \ref{tab:raw_data_norm_winnum_NDCG3} in Appendix \ref{app:norm_winnum_NDCG3} provides the raw \ac{NWN} data for the Learning to Rank methods for which \ac{NDCG}@3 evaluation results were available.\\
+Table \ref{tab:raw_data_norm_winnum_ndcg3} in Appendix \ref{app:norm_winnum_ndcg3} provides the raw \ac{NWN} data for the Learning to Rank methods for which \ac{NDCG}@3 evaluation results were available.\\
 
@@
 Figure \ref{fig:normalised_winning_number_NDCG5} shows the performance of Learning to Rank methods for the \ac{NDCG}@5 metric.
-Table \ref{tab:raw_data_norm_winnum_NDCG5} in Appendix \ref{app:norm_winnum_NDCG5} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\
+Table \ref{tab:raw_data_norm_winnum_ndcg5} in Appendix \ref{app:norm_winnum_ndcg5} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\
 
@@
 Figure \ref{fig:normalised_winning_number_NDCG10} shows the performance of Learning to Rank methods for the \ac{NDCG}@10 metric.
-Table \ref{tab:raw_data_norm_winnum_NDCG10} in Appendix \ref{app:norm_winnum_NDCG10} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\
+Table \ref{tab:raw_data_norm_winnum_ndcg10} in Appendix \ref{app:norm_winnum_ndcg10} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\
 
@@
 The set of algorithms for which there is no other algorithm with both a higher \ac{NWN} and number of data sets evaluated on is partly in agreement with those for the \ac{NDCG}@3 and \ac{NDCG}@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
-SmoothRank and FSMRank were not present in this set for the \ac{NDCG}@3 and \ac{NDCG}@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_NDCG3} and \ref{tab:raw_data_norm_winnum_NDCG5} in the Appendices.
+SmoothRank and FSMRank were not present in this set for the \ac{NDCG}@3 and \ac{NDCG}@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_ndcg3} and \ref{tab:raw_data_norm_winnum_ndcg5} in the Appendices.
 DCMP is not in the set in contrast with the \ac{NDCG}@3 and \ac{NDCG}@5 comparison.
@@
 Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all \ac{NDCG}-comparisons, are not within the best performing methods in the \ac{MAP}-comparison.
-A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_NDCG3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_NDCG3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more data sets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher \ac{NWN}.
+A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_ndcg3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_ndcg3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more data sets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher \ac{NWN}.
 SmoothRank performed a \ac{NWN} of around 0.53 over 7 data sets, which is still good in both certainty and accuracy, but not among the top methods.
@@
 
-Our comparison regarded evaluation results on \ac{NDCG}@{3, 5, 10} and \ac{MAP}.
+Our comparison regarded evaluation results on \ac{NDCG}@$\{3,5,10\}$ and \ac{MAP}.
 By making the decision to include \ac{NDCG} at three cut-off points and only a single \ac{MAP} entry, we implicitly attain a higher weight for \ac{NDCG} compared to \ac{MAP} on an analysis that combines all measurements on the four metrics.

### 2014-09-29 report/Chapters/Future.tex
mooie dingen

@@
-\section{Computing models}
+\section{Distributed Computing Models}
 We explored the possibilities of Hadoop MapReduce for distributed computation of Learning to Rank training algorithms.
@@
 
-\section{Learning to Rank algorithms}
+\section{Learning to Rank Algorithms}
+As an answer to the first research question of this thesis, we found five Learning to Rank methods for it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
+These algorithms were, from highest ranking accuracy / lowest certainty to lowest ranking accuracy / highest certainty: ListNet, SmoothRank, FenchelRank, FSMRank and LRUF.
+In this thesis we explored the speed-up characteristics of ListNet on Hadoop MapReduce.
+The speed-up characteristics of SmoothRank, FenchelRank, FSMRank and LRUF are using the MapReduce computing model are still to be explored.
+
+\section{Optimisation Algorithms}

### 2014-09-29 report/FrontBackmatter/Titlepage.tex
voorkant
Dolf erbij

@@
 
-\committee{Dr. ir. Djoerd Hiemstra (UT/Databases)\\
+\committee{
+Dr. ir. Djoerd Hiemstra (University of Twente / Databases)\\
+Ir. Dolf Trieschnigg (University of Twente / Human Media Interaction)\\
 Sander Bockting, Msc (Avanade Netherlands B.V.)\\
@@
 
-\supervisors{Dr. ir. Djoerd Hiemstra (UT/Databases)\\
-Sander Bockting (Avanade Netherlands B.V.)}    
+\supervisors{
+Dr. ir. Djoerd Hiemstra (University of Twente / Databases)\\
+Sander Bockting (Avanade Netherlands B.V.)
+}    
 \end{titlepage}
 \maketitle
+\clearpage

### 2014-09-29 report/FrontBackmatter/Preface.tex
preface klaar

@@
 This research was conducted at Avanade Netherlands B.V.
-under the primary supervision of Sander Boackting Msc.
+under the primary supervision of Sander Bockting Msc at Avanade Netherlands and Dr.
+ir.
+Djoerd Hiemstra at the University of Twente.
 I would like to use this page to express my gratitude to everyone who supported my throughout this project in any way.\\
 
-Many thanks go to Dr. ir. Djoerd Hiemstra of the University of Twente and to Sander Bockting Msc. of Avanade Netherlands BV for their great supervision throughout the project.\\
+Many thanks go to Dr.
+ir.
+Djoerd Hiemstra of the University of Twente and to Sander Bockting Msc.
+of Avanade Netherlands B.V.
+for their great supervision throughout the project.
+Even though we kept face-to-face meetings to a very minimum, you both provided me with very insightful and valuable feedback either in those meetings or per e-mail.\\
 
-In addition I would like to thank all other graduate interns at Avanade as well as all the Avanade employees for the great talks at the coffee machine, during the Friday afternoon drink, or elsewhere.
-In particular I owe many thanks to Michel Barneveld for giving me insight in potential applications of Learning-to-Rank within Avanade.\\
+In addition I would like to thank all fellow graduate interns at Avanade as well as all the Avanade employees for the great talks at the coffee machine, during the Friday afternoon drink, or elsewhere.
+In particular I would like to mention fellow graduate interns Fayaz Kalan, Casper Veldhuijzen, Peter Mein, and (again) Jurjen Nienhuis for the very good time that we had together at the office as well as during the numerous drinks and diners that we had together outside office hours.\\
+
+I finish this section by thanking everyone that helped improving the quality of my work by providing me with valuable feedback.
+I would like to thank my former fellow boards members at study association Inter-\emph{Actief} Rick van Galen and Jurien Wagenaar, who provided me with feedback in the early stages of the process.
+In particular I would like to thank fellow graduate intern Jurjen Nienhuis, for the numerous mutual feedback sessions that we held, which most certainly helped raising the quality of this thesis to a higher level.\\
 

### 2014-09-29 report/FrontBackmatter/Titlepage.tex
preface klaar

@@
 Dr. ir. Djoerd Hiemstra (University of Twente / Databases)\\
-Ir. Dolf Trieschnigg (University of Twente / Human Media Interaction)\\
+Dr. ir. Dolf Trieschnigg (University of Twente / Human Media Interaction)\\
 Sander Bockting, Msc (Avanade Netherlands B.V.)\\
@@
 Dr. ir. Djoerd Hiemstra (University of Twente / Databases)\\
-Sander Bockting (Avanade Netherlands B.V.)
+Sander Bockting, Msc (Avanade Netherlands B.V.)
 }    

### 2014-09-29 report/FrontBackmatter/Preface.tex
preface is mooi zo :)

@@
 This research was conducted at Avanade Netherlands B.V.
-under the primary supervision of Sander Bockting Msc at Avanade Netherlands and Dr.
+under the primary supervision of Sander Bockting Msc of Avanade and Dr.
 ir.
-Djoerd Hiemstra at the University of Twente.
-I would like to use this page to express my gratitude to everyone who supported my throughout this project in any way.\\
+Djoerd Hiemstra from the University of Twente.
+I would like to use this page to express my gratitude to everyone who supported me throughout this project in any way.\\
 
@@
 ir.
-Djoerd Hiemstra of the University of Twente and to Sander Bockting Msc.
+Djoerd Hiemstra from the University of Twente and to Sander Bockting Msc.
 of Avanade Netherlands B.V.
@@
 In addition I would like to thank all fellow graduate interns at Avanade as well as all the Avanade employees for the great talks at the coffee machine, during the Friday afternoon drink, or elsewhere.
-In particular I would like to mention fellow graduate interns Fayaz Kalan, Casper Veldhuijzen, Peter Mein, and (again) Jurjen Nienhuis for the very good time that we had together at the office as well as during the numerous drinks and diners that we had together outside office hours.\\
+In particular I would like to mention fellow graduate interns Fayaz Kallan, Casper Veldhuijzen, Peter Mein, and Jurjen Nienhuis for the very good time that we had together at the office as well as during the numerous drinks and dinners that we had together outside office hours.\\
 
 I finish this section by thanking everyone that helped improving the quality of my work by providing me with valuable feedback.
-I would like to thank my former fellow boards members at study association Inter-\emph{Actief} Rick van Galen and Jurien Wagenaar, who provided me with feedback in the early stages of the process.
-In particular I would like to thank fellow graduate intern Jurjen Nienhuis, for the numerous mutual feedback sessions that we held, which most certainly helped raising the quality of this thesis to a higher level.\\
+I would like to thank my good friends and former fellow boards members at study association I.C.T.S.V.
+Inter-\emph{Actief} Rick van Galen and Jurri\"{e}n Wagenaar, who provided me with feedback in the early stages of the process.
+In particular I would like to thank Jurjen Nienhuis (yes, again), for the numerous mutual feedback sessions that we held, which most certainly helped raising the quality of this thesis to a higher level.\\
+\bigskip
 

### 2014-09-29 report/FrontBackmatter/Preface.tex
consistentie in Sander's titel
Dolf er ook in

@@
 ir.
-Djoerd Hiemstra from the University of Twente and to Sander Bockting Msc.
-of Avanade Netherlands B.V.
+Djoerd Hiemstra from the University of Twente and to Sander Bockting Msc of Avanade Netherlands B.V.
 for their great supervision throughout the project.
-Even though we kept face-to-face meetings to a very minimum, you both provided me with very insightful and valuable feedback either in those meetings or per e-mail.\\
+Even though we kept face-to-face meetings to a very minimum, you both provided me with very insightful and valuable feedback either in those meetings or per e-mail.
+Also, I would like to thank Dr.
+ir.
+Dolf Trieschnigg, together with Djoerd and Sander a member of the assessment committee of this graduation project, for still being available for the role as second assessor of this work, while having joined the process rather late.\\
 

### 2014-09-30 report/Chapters/CrossBenchmarkComparison.tex
verwijzingen aangepast aan nieuwe appendixstructuur

@@
 Figure \ref{fig:normalised_winning_number_NDCG3} shows the performance of Learning to Rank methods for the \ac{NDCG}@3 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg3} in Appendix \ref{app:norm_winnum_ndcg3} provides the raw \ac{NWN} data for the Learning to Rank methods for which \ac{NDCG}@3 evaluation results were available.\\
+Table \ref{tab:raw_data_norm_winnum_ndcg35} in Appendix \ref{app:norm_winnum_ndcg35} provides the raw \ac{NWN} data for the Learning to Rank methods for which \ac{NDCG}@3 evaluation results were available.\\
 
@@
 Figure \ref{fig:normalised_winning_number_NDCG5} shows the performance of Learning to Rank methods for the \ac{NDCG}@5 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg5} in Appendix \ref{app:norm_winnum_ndcg5} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\
+Table \ref{tab:raw_data_norm_winnum_ndcg35} in Appendix \ref{app:norm_winnum_ndcg35} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\
 
@@
 Figure \ref{fig:normalised_winning_number_NDCG10} shows the performance of Learning to Rank methods for the \ac{NDCG}@10 metric.
-Table \ref{tab:raw_data_norm_winnum_ndcg10} in Appendix \ref{app:norm_winnum_ndcg10} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\
+Table \ref{tab:raw_data_norm_winnum_ndcg10map} in Appendix \ref{app:norm_winnum_ndcg10map} provides the raw Normalised Winning Number data for the Learning to Rank methods.\\
 
@@
 The set of algorithms for which there is no other algorithm with both a higher \ac{NWN} and number of data sets evaluated on is partly in agreement with those for the \ac{NDCG}@3 and \ac{NDCG}@5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
-SmoothRank and FSMRank were not present in this set for the \ac{NDCG}@3 and \ac{NDCG}@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_ndcg3} and \ref{tab:raw_data_norm_winnum_ndcg5} in the Appendices.
+SmoothRank and FSMRank were not present in this set for the \ac{NDCG}@3 and \ac{NDCG}@5 comparison, but were close by, as can be seen in Tables \ref{tab:raw_data_norm_winnum_ndcg35} in Appendix \ref{app:norm_winnum_ndcg35}.
 DCMP is not in the set in contrast with the \ac{NDCG}@3 and \ac{NDCG}@5 comparison.
@@
 Figure \ref{fig:normalised_winning_number_map} shows the performance of Learning to Rank methods for the \ac{MAP} metric.
-Table \ref{tab:raw_data_norm_winnum_map} in Appendix \ref{app:norm_winnum_map} provides the raw \ac{NWN} data for the Learning to Rank methods.\\
+Table \ref{tab:raw_data_norm_winnum_ndcg10map} in Appendix \ref{app:norm_winnum_ndcg10map} provides the raw \ac{NWN} data for the Learning to Rank methods.\\
 
@@
 Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all \ac{NDCG}-comparisons, are not within the best performing methods in the \ac{MAP}-comparison.
-A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_ndcg3} to \ref{tab:raw_data_norm_winnum_map} in Appendices \ref{app:norm_winnum_ndcg3} to \ref{app:norm_winnum_map} shows that LAC-MR-OR is evaluated on many more data sets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher \ac{NWN}.
+A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_ndcg35} and \ref{tab:raw_data_norm_winnum_ndcg10map} in Appendices \ref{app:norm_winnum_ndcg35} and \ref{app:norm_winnum_ndcg10map} respectively shows that LAC-MR-OR is evaluated on many more data sets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher \ac{NWN}.
 SmoothRank performed a \ac{NWN} of around 0.53 over 7 data sets, which is still good in both certainty and accuracy, but not among the top methods.

### 2014-09-30 report/FrontBackmatter/Titlepage.tex
voorkantje fix
generalisatie van titlepage

--- 2014-09-26 report/FrontBackmatter/Titlepage.tex
@@
 \begin{titlepage}
-\title{Master's thesis Computer Science\\ (Chair: Databases, Track: Information System Engineering)}
-\course{Scaling Learning to Rank to Big Data}
-\subtitle{A study concerning parallelisation of Learning to Rank algorithms using the MapReduce computing model}
-\university{University of Twente}
-\faculty{Faculty of Electrical Engineering, Mathematics and Computer Science (EEMCS)}
-\author{Niek Tax (s0166197)}
+\title{Master's thesis Computer Science\\ (Chair: \myDepartment, Track: \myTrack)}
+\course{\myTitle}
+\subtitle{\mySubtitle}
+\university{\myUni}
+\faculty{\myFaculty}
+\author{\myName (\myStudNum)}
 
-\committee{Dr. ir. Djoerd Hiemstra (UT/Databases)\\
-Sander Bockting, Msc (Avanade Netherlands B.V.)\\
+\committee{
+\myProf (\myProfAffil)\\
+\myOtherProf (\myOtherProfAffil)\\
+\mySupervisor (\mySupervisorAffil)\\
 }	
@@
 
-\supervisors{Dr. ir. Djoerd Hiemstra (UT/Databases)\\
-Sander Bockting (Avanade Netherlands B.V.)}    
+\supervisors{
+\myProf (\myProfAffil)\\
+\mySupervisor (\mySupervisorAffil)
+}    
 \end{titlepage}
 \maketitle
+\clearpage

### 2014-09-30 report/Chapters/Background.tex
kleine fixes
abstract + toevoeging results + conclusion

@@
 \end{enumerate}
+\smallskip
 These three approaches are in literature respectively labelled the pointwise approach, the pairwise approach and the listwise approach.
@@
 A cross-validation experiment \cite{kohavi1995}, sometimes called rotation estimation, is an experimental set-up for model evaluation where the data is split into $k$ chunks of approximately equal size, called \emph{folds}.
-One of the folds is used as validation set and one of the folds is used as test set, the rest of the $k-2$ folds are used as training data.
-This procedure is repeated $k$ times, such that each fold is once used for validation, once as test set, and $k-2$ times as training data.
+One of the folds is used as validation set, one of the folds is used as test set, and the rest of the $k - 2$ folds are used as training data.
+This procedure is repeated $k$ times, such that each fold is once used for validation, once as test set, and $k - 2$ times as training data.
 The performance can be measured in any model evaluation metric, and is averaged over the model performances on each of the folds.

### 2014-09-30 report/Chapters/Conclusions.tex
abstract + toevoeging results + conclusion

@@
 To answer this research question we proposed a new way of comparing learning to rank methods based on sparse evaluation results data on a set of benchmark datasets.
-Our comparison methodology comprises of two components: 1) NWN, which provides insight in the ranking accuracy of the learning to rank method, and 2) IWN, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.
+Our comparison methodology comprises of two components: 1) \ac{NWN}, which provides insight in the ranking accuracy of the learning to rank method, and 2) \ac{IWN}, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.
 Based on our literature search for evaluation results on well-known benchmarks collections, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.
@@
 
+Most importantly, we found the training time of our cluster version of ListNet to grow better than linearly in terms of data size increase.
+This shows that the cluster implementation of ListNet can be used to scale the ListNet training procedure to arbitrarily large data sets, given that enough data nodes are available for computation.\\
+
 No generalisations can be drawn from these results to the scalability on MapReduce of other Learning to Rank algorithms.

### 2014-09-30 report/Chapters/CrossBenchmarkComparison.tex
abstract + toevoeging results + conclusion

@@
 
-LARF, \ac{IP}Rank and DCMP and several variants of Rank\ac{SVM} performed very well on the cross-metric comparison, with all having only one method in its top right quadrant.
+LARF, \acs{IP}Rank and DCMP and several variants of Rank\ac{SVM} performed very well on the cross-metric comparison, with all having only one method in its top right quadrant.
 LARF also performed among the top methods on the \ac{NDCG} and \ac{MAP} comparisons and DCMP was a top performer in a few of the \ac{NDCG} comparisons.\\
@@
 Another limitation is that our comparison methodology relies on the correctness of the evaluation results found in the literature search step.
-This brings up a risk of overly optimistic evaluation results affecting our NWN results.
+This brings up a risk of overly optimistic evaluation results affecting our \ac{NWN} results.
 Limiting the meta-analysis to those studies that report comparable results on one of the baseline methods of a benchmark set reduces this limitation but does not solve it completely.
-By taking IWN into account in Figure \ref{fig:normalised_winning_number_all} we further mitigate this limitation, as IWN is loosely related with the number of studies that reported evaluation results for an algorithm.\\
+By taking \ac{IWN} into account in Figure \ref{fig:normalised_winning_number_all} we further mitigate this limitation, as \ac{IWN} is loosely related with the number of studies that reported evaluation results for an algorithm.\\
 

### 2014-09-30 report/Chapters/Future.tex
abstract + toevoeging results + conclusion

@@
+Potential future areas of research that follow from this thesis can be categorised into three categories. These categories of potential future work are described in the sections below.
+
 \section{Distributed Computing Models}
 We explored the possibilities of Hadoop MapReduce for distributed computation of Learning to Rank training algorithms.
-Hadoop, since the introduction of Hadoop YARN in Hadoop 2.0, offers integration with other distributed computing models, including Dryad \cite{Isard2007}, Spark \cite{Zaharia2010}, Storm \cite{Aniello2013} and MPI.
+Hadoop, since the introduction of Hadoop YARN in Hadoop 2.0, offers integration with other distributed computing models, including Dryad \cite{Isard2007}, Spark \cite{Zaharia2010}, Storm \cite{Aniello2013} and \ac{MPI}.
 Of these distributed programming models, Spark is particularly promising, since Shukla et al.
@@
 \section{Optimisation Algorithms}
+Lin \cite{Lin2013} stated that the gradient descent optimisation procedure can often be replaced with other optimisation procedures with faster convergence properties (often at the cost of being more computationally expensive per iteration), in order to make iterative machine learning algorithms perform better on MapReduce.
+An example of an optimisation procedures that fits this higher convergence rate at the cost of higher per-iteration computational costs is L-BFGS \cite{Liu1989}, or any other quasi-Newton optimisation method.
+Testing the effect that replacing the optimisation method of Learning to Rank methods has on the speed-up that can be achieved by parallelising the methods with the MapReduce model is still to be determined.
+Note however that replacing the optimisation method in a Learning to Rank algorithm basically turns it into a different, new, Learning to Rank method, as different optimisation methods might not be equivalent in how well they are able to find a good set of model parameters.
+One should therefore also explore the ranking accuracy characteristics when an existing Learning to Rank method is evaluated in combination with an optimisation algorithm that is not prescribed to be a part of the Learning to Rank method.

### 2014-09-30 report/Chapters/Implementation.tex
kleine fixes
abstract + toevoeging results + conclusion

@@
 HDInsight Hadoop data nodes are regular Large A3 Microsoft Azure virtual machines with four core processors and 7GB RAM.
-Newly supported programming models since Hadoop 2.0 include Dryad \cite{Isard2007}, Giraph \cite{Avery2011}, MPI, REEF \cite{Chun2013}, Spark \cite{Zaharia2010}, and Storm \cite{Aniello2013}.
+Newly supported programming models since Hadoop 2.0 include Dryad \cite{Isard2007}, Giraph \cite{Avery2011}, \ac{MPI}, REEF \cite{Chun2013}, Spark \cite{Zaharia2010}, and Storm \cite{Aniello2013}.
 Even though these programming models are now supported by Hadoop and some of these programming models have recently increased in popularity, they still lack the critical mass as the data processing framework of choice that MapReduce has as Lin argued back in 2012 \cite{Lin2013} (see section \ref{sec:motivation}).
@@
 \subsection{Preprocessing}
+\label{ssec:preprocessing}
 The preprocessing phase consists of two separate Pig jobs.
@@
 \end{minipage}\\
+
 All expressions between square brackets in the code snippet above are dynamically set by the framework.

### 2014-09-30 report/Chapters/RelatedWork.tex
kleine fixes
abstract + toevoeging results + conclusion

@@
 \cite{Burges2011} described their Yahoo!
-Learning to Rank Challenge submission that was built by performing an extensive hyperparameter search on a 122-node MPI cluster, running Microsoft HPC Server 2008.
+Learning to Rank Challenge submission that was built by performing an extensive hyperparameter search on a 122-node \ac{MPI} cluster, running Microsoft HPC Server 2008.
 The hyperparameter optimisation was performed on a linear combination ensemble of eight LambdaMART models, two LambdaRank models and two MART models using a logistic regression cost.
@@
 Some research focused on parallelising the steps Learning to Rank algorithms that can be characterised as strong learners.
-
 Tyree et al.
@@
 Ye et al.
-\cite{Ye2009} described how to implement the stochastic \ac{GBDT} model in a parallel manner using both MPI and Hadoop.
+\cite{Ye2009} described how to implement the stochastic \ac{GBDT} model in a parallel manner using both \ac{MPI} and Hadoop.
 Stochastic \ac{GBDT} differs from regular \ac{GBDT} models by using stochastic gradient boosting instead of regular gradient boosting.
@@
 Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
-The MPI approach proved to be successful and obtained near linear speed-ups.\\
+The \ac{MPI} approach proved to be successful and obtained near linear speed-ups.\\
 

### 2014-09-30 report/Chapters/Results.tex
abstract + toevoeging results + conclusion

@@
 
-Table \ref{tbl:accuracy_comparison} shows the mean and the standard deviation of the ranking accuracy in \ac{NDCG}@10 on the first data fold for our own cluster implementation of ListNet, the RankLib implementation of ListNet, as well as the official LETOR evaluation results of ListNet as reported on the LETOR 3.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and LETOR 4.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} websites.
+Table \ref{tbl:accuracy_comparison} shows the mean and the standard deviation of the ranking accuracy in \ac{NDCG}@10 on the first data fold for our own cluster implementation (with and without the preprocessing step described in section \ref{ssec:preprocessing}) of ListNet, the RankLib implementation of ListNet, as well as the official LETOR evaluation results of ListNet as reported on the LETOR 3.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and LETOR 4.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} websites.
 Note that the official LETOR evaluation results are structurally higher compared to the results obtained with RankLib and our cluster implementation, which can be attributed to the fact that the official evaluation runs use as many ListNet iterations as needed for convergence on the validation fold, while we limited our experiment to five iterations.
@@
 
-A remarkable finding is that the cluster implementation shows better ranking accuracy than the RankLib version of ListNet after five iterations.
-We attribute this to the normalisation preprocessing step that we incorporated into our ListNet implementation that transforms each feature into a [0,1]-range.
-RankLib does not perform this normalisation preprocessing step.
-It has been shown that the gradient descent optimisation procedure, which is used in ListNet, converges faster on normalised data \cite{Ng1999}.
+The ranking accuracy of the ListNet cluster implementation without the normalisation preprocessing step seems very comparable to the ranking accuracy obtained with the RankLib ListNet implementation, which suggests that the ListNet ranking method is indeed implemented correctly.
+Note however that the standard deviations of the measured ranking accuracies are rather high, which leaves us unable to conclude that the ranking accuracy of RankLib ListNet and our cluster implementation are indeed equivalent in achieved ranking accuracy up to high precision.\\
+
+A remarkable finding is that the cluster implementation with the normalisation preprocessing procedure shows better ranking accuracy than the RankLib version of ListNet after five iterations.
+It seems to be the case that the normalisation procedure enables ListNet to converge faster.
+It has been shown in literature, that the gradient descent optimisation procedure, which is also being used in ListNet, converges faster on normalised data \cite{Ng1999}.
 
@@
 \centering
-\begin{tabular}{p{3.5cm}p{2.6cm}p{2.6cm}r}\toprule
-Dataset & RankLib & Cluster & Official evaluation \\
+\begin{tabular}{p{3.4cm}p{2.6cm}p{2.6cm}p{2.6cm}}\toprule
+ &  OHSUMED (LETOR 3.0 version) & MQ2008 & MQ2007 \\
 \midrule
-OHSUMED (LETOR 3.0 version) & 0.1916 $\pm$ 0.0634 & 0.3395 $\pm$ 0.0952 & 0.3793 \\
-MQ2008      				& 0.3889 $\pm$ 0.0600 & 0.4280 $\pm$ 0.0998 & 0.4689 \\
-MQ2007      				& 0.2918 $\pm$ 0.0546 & 0.4554 $\pm$ 0.0085 & 0.4767 \\
+RankLib                             & 0.1916 $\pm$ 0.0634 & 0.3889 $\pm$ 0.0600 & 0.2918 $\pm$ 0.0546 \\
+Cluster \newline (no preprocessing)	& 0.2212 $\pm$ 0.0564 & 0.3496 $\pm$ 0.0788 & 0.2846 $\pm$ 0.0891\\
+Cluster \newline (preprocessing)    & 0.3395 $\pm$ 0.0952 & 0.4280 $\pm$ 0.0998 & 0.4554 $\pm$ 0.0085\\
+Official evaluation                 & 0.3793 & 0.4689 & 0.4767 \\
 \bottomrule
@@
 
-Figure \ref{fig:speedup_train_time} shows that speed-up is sub-linear, from which we can deduct that the training time converges to a constant unit of time.
+Figure \ref{fig:speedup_train_time} shows that the speed-up achieved as function of the number of processing data nodes is sub-linear, from which we can deduct that the training time converges to a constant unit of time.
 Based on our measurements on the small data sets MINI, OHSUMED, MQ2008 and MQ2007, this constant time seems to be within the range of 150 to 200 seconds.
@@
 
+A result of that the Hadoop job scheduling overhead does not grow when the input data grows, is that the speed-up is linear or potentially better than linear as a function of data size.
+This can be seen in Figure \ref{fig:listnet_train_time}.
+In this Figure the cluster run lines are linear or, in case of the 24 and 32 data nodes cluster lines, slightly better than linear.
+The lines of the clusters of 8 data nodes or smaller have high variance, this can be attributed to the fact that these clusters were mostly tested on very small datasets, leaving the measurement very dependent on the variance of the job scheduling overhead.
+
 \begin{figure}

### 2014-09-30 report/FrontBackmatter/Abstract.tex
abstract + toevoeging results + conclusion

%*******************************************************
% Abstract
%*******************************************************
%\renewcommand{\abstractname}{Abstract}
\pdfbookmark[1]{Abstract}{Abstract}
\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Abstract}
Learning to rank is an increasingly important scientific field that comprises the use of machine learning for the ranking task. New learning to rank methods are generally evaluated in terms of ranking accuracy on benchmark test collections. However, comparison of learning to rank methods based on evaluation results is hindered by non-existence of a standard set of evaluation benchmark collections. Furthermore, little research is done in the field of scalability of the training procedure of Learning to Rank methods, to prepare us for input data sets that are getting larger and larger. This paper concerns both the comparison of Learning to Rank methods using a sparse set of evaluation results on benchmark data sets, as well as the speed-up that can be achieved by parallelising Learning to Rank methods using MapReduce.\\

In the first part of this thesis we propose a way to compare learning to rank methods based on a sparse set of evaluation results on a set of benchmark datasets. Our comparison methodology consists of two components: 1) Normalized Winning Number, which gives insight in the ranking accuracy of the learning to rank method, and 2) Ideal Winning Number, which gives insight in the degree of certainty concerning its ranking accuracy. Evaluation results of 87 learning to rank methods on 20 well-known benchmark datasets are collected through a structured literature search. ListNet, SmoothRank, FenchelRank, FSMRank, LRUF and LARF were found to be the best performing learning to rank methods in increasing order of Normalized Winning Number and decreasing order of Ideal Winning Number.\\

In the second part of this thesis we analysed the speed-up of the ListNet training algorithm when implemented in the MapReduce computing model. We found that running ListNet on MapReduce comes with a job scheduling overhead in the range of 150-200 seconds per training iteration. This makes MapReduce very inefficient to process small data sets with ListNet, compared to a single-machine implementation of the algorithm. The MapReduce implementation of ListNet was found to be able to offer large improvements in processing time for data sets that are larger than the physical memory of the single machine otherwise available for computation. In addition we showed that ListNet tends to converge faster when a normalisation preprocessing procedure is used. The training time of our cluster version of ListNet was found to grow better than linearly in terms of data size increase. This shows that the cluster implementation of ListNet can be used to scale the ListNet training procedure to arbitrarily large data sets, given that enough data nodes are available for computation.

\endgroup

\vfill

### 2014-10-14 implementation/java/src/letor/serial/WrappedRunner.java
begin presentatie

@@ -15,7 +15,7 @@ public class WrappedRunner {
         Integer                     folds               = 1;
         Integer                     iterations          = 5;
         AbstractParameterizedRanker handler             = new SmoothRankHandler();
-        DataSets.DataSet            dataset             = DataSets.DataSet.MQ2007;
+        DataSets.DataSet            dataset             = DataSets.DataSet.MSLR_WEB30K;
         Integer                     duplicationNumber   = 1; // Only relevant in case of dataset Custom
 
         FoldRunHandler ltrWrapper = null;

### 2014-10-15 implementation/java/src/letor/serial/WrappedRunner.java
verder aan presentatie

@@ -15,7 +15,7 @@ public class WrappedRunner {
         Integer                     folds               = 1;
         Integer                     iterations          = 5;
         AbstractParameterizedRanker handler             = new SmoothRankHandler();
-        DataSets.DataSet            dataset             = DataSets.DataSet.MSLR_WEB30K;
+        DataSets.DataSet            dataset             = DataSets.DataSet.MQ2008;
         Integer                     duplicationNumber   = 1; // Only relevant in case of dataset Custom
 
         FoldRunHandler ltrWrapper = null;

### 2014-10-23 report/Chapters/Background.tex
feedback Sander pagina 19

@@
 %************************************************
-This chapter is written with the goal to provide the reader with a subtle introduction in the Learning to Rank field and the MapReduce computational model.
+This chapter is written with the goal to provide the reader with a brief introduction in the Learning to Rank field and the MapReduce computational model.
 Models and theories explained in this chapter can be regarded as prior knowledge needed to understand the subsequent chapters of this thesis.
@@
 Figure \ref{fig:discriminative_training} describes the general process of machine learning.
-It shows training elements from an input space that are mapped to an output space using a model such that the difference between the actual labels of the training elements and the labels predicted with with the model are minimal in terms of a loss function.
+Input space $X$ consists of input objects $x$.
+A hypothesis $h$ defines a mapping of input objects from $X$ into the output space $Y$, resulting in prediction $\hat{y}$.
+The loss of an hypothesis is the difference between the predictions made by the hypothesis and the correct values mapped from the input space into the output space, called the \emph{ground truth labels}.
+The task of machine learning is to find the best fitting hypothesis $h$ from the set of all possible hypotheses $H$, called the hypothesis space.\\
+
 \begin{figure}[!h]
@@
 \label{fig:discriminative_training}
-\end{figure}\\
+\end{figure}
+
 Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning to Rank method when it is \emph{feature based} and uses \emph{discriminative training}, in which the concepts \emph{feature based} and \emph{discriminative training} are itself defined as:

### 2014-10-23 report/Chapters/Background.tex
feedback Sander pagina 20

@@
 The functions operate on the feature vectors of the input object, and make predictions according to the format of the output space.}
-	\item[Loss Function]{in order to learn the optimal hypothesis, a training set is usually used, which contains a number of objects and their ground truth labels, sampled from the product of the input and output spaces.}
+	\item[Loss Function]{in order to learn the optimal hypothesis, a training set is usually used, which contains a number of objects and their ground truth labels, sampled from the product of the input and output spaces.
+A loss function calculates the difference between the predictions $\hat{y}$ and the ground truth labels on a given set of data.
 	\end{description}

### 2014-10-23 report/Chapters/BenchmarkResults.tex
kleine fixes
feedback Sander pagina 59

@@
 Note that this does not necessarily imply that these methods would have scored a higher \ac{NDCG} winning number than ListNet.
-Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR data sets in terms of \ac{NDCG}@10 and \ac{MAP}.
+Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR data sets in terms of \ac{NDCG}@10 and \ac{MAP}.\\
 \begin{table}[!h]
@@
 \end{tabular}
-\caption{\acs{NDCG} results of the baseline methods on the WCL2R collection, obtained from \cite{Alcantara2010}}
+\caption{\acs{NDCG} results of the baseline methods on the WCL2R collection, obtained from Alc\^{a}ntara et al. \cite{Alcantara2010}}
 \label{tab:results_WCL2R}

### 2014-10-23 report/Chapters/LTRMethods.tex
opmaakpuntje
kleine fixes
feedback Sander pagina 59

@@
  Output Neural Network model $\omega$.
- \caption{The ListNet learning algorithm, obtained from \cite{Cao2007}}
+ \caption{The ListNet learning algorithm, obtained from Cao et al. \cite{Cao2007}}
  \label{alg:listnet}
@@
 Algorithm \ref{alg:smoothrank} shows the algorithm that summarises all steps of the SmoothRank method.\\
+
 \LinesNumbered
@@
  }
- \caption{The SmoothRank learning algorithm, obtained from \cite{Chapelle2010}}
+ \caption{The SmoothRank learning algorithm, obtained from Chapelle and Wu \cite{Chapelle2010}}
  \label{alg:smoothrank}
@@
  return $w_{T}$ as ranking predictor for $w$
- \caption{The FenchelRank learning algorithm, obtained from \cite{Lai2013}}
+ \caption{The FenchelRank learning algorithm, obtained from Lai et al. \cite{Lai2013}}
  \label{alg:fenchelrank}
@@
 Loss function $f$ can in theory be any convex loss function, but Lai et al.
-\cite{Lai2013c} used a squared hinge loss function for their evaluation measurements.
+\cite{Lai2013c} used a squared hinge loss function for their evaluation measurements.\\
 
@@
 
- \caption{The FSMRank learning algorithm, obtained from \cite{Lai2013c}}
+ \caption{The FSMRank learning algorithm, obtained from Lai et al. \cite{Lai2013c}}
  \label{alg:fsmrank}
@@
  Output $\underline{R_i}$
- \caption{The LRUF algorithm, obtained from \cite{Torkestani2012b}}
+ \caption{The LRUF algorithm, obtained from Torkestani \cite{Torkestani2012b}}
  \label{alg:lruf}

### 2014-10-23 report/Chapters/LTRMethods.tex
kleine opmaakfix

@@
 \end{equation}
-\noindent in which $N_i$ refers to the size of feedback set $\underline{f_i}$, $r_i^j$ is the rank of document $d_i^j$ in $R_i$ and, again, $a$ denotes the learning rate.\\
-
-Initially, before any feedback, the action probability factors are initialised with equal probability.
+\noindent in which $N_i$ refers to the size of feedback set $\underline{f_i}$, $r_i^j$ is the rank of document $d_i^j$ in $R_i$ and, again, $a$ denotes the learning rate.
+Initially, before any feedback, the action probability factors are initialised with equal probability.\\
 
@@
 new documents are indexed or old documents are removed, LRUF has an Increase Action-Set Size (IAS) and a Reduce Action-Set Size (RAS) procedure defined to adapt to the new document set without needing a complete retraining of the model.
-Because this thesis focusses on model accuracy and scalability the IAS and RAS procedure of LRUF will not be explained in further detail.
+Because this thesis focusses on model accuracy and scalability the IAS and RAS procedure of LRUF will not be explained in further detail.\\
 

### 2014-10-23 report/Chapters/Implementation.tex
feedback Sander pagina 70

@@
 
+\begin{sloppypar}
 Setting the degree of parallelisation is crucial to obtain optimal cluster utilisation.
@@
 Cluster utilisation is optimal when the computational work is evenly distributed over exactly the number of available mappers in the map phase, and evenly distributed over exactly the number of reducers in the reduce phase.
-The number of mappers used can be controlled by setting MapReduce configuration parameters mapreduce.input.fileinputformat.split.maxsize, mapreduce.input.fileinputformat.split.minsize, and the Pig configuration parameter pig.maxCombinedSplitSize to the data size in bytes of the input training data divided by the number of available mappers.
+The number of mappers used can be controlled by setting MapReduce configuration parameters \emph{mapreduce.input.fileinputformat.split.maxsize}, \emph{mapreduce.input.fileinputformat.split.minsize}, and the Pig configuration parameter \emph{pig.maxCombinedSplitSize} to the data size in bytes of the input training data divided by the number of available mappers.
 To control the number of used reducers, Pig offers two options: 1) the default\_parallel parameters, which can set a default number of reducers used throughout all MapReduce jobs that the Pig script consists of, and 2) the "PARALLEL" clause, which sets the number of reducers at operator level (overrides default\_parallel if set).
@@
 Pig code to set the number of mappers and reducers are dynamically set by the framework such that the optimal number of mappers and reducers are used.
+\end{sloppypar}
 

### 2014-10-23 report/FrontBackmatter/Abstract.tex
toevoeging op verzoek van Djoerd

@@
 Evaluation results of 87 learning to rank methods on 20 well-known benchmark datasets are collected through a structured literature search.
-ListNet, SmoothRank, FenchelRank, FSMRank, LRUF and LARF were found to be the best performing learning to rank methods in increasing order of Normalized Winning Number and decreasing order of Ideal Winning Number.\\
+ListNet, SmoothRank, FenchelRank, FSMRank, LRUF and LARF were found to be the best performing learning to rank methods in increasing order of Normalized Winning Number and decreasing order of Ideal Winning Number.
+Of these ranking algorithms, FenchelRank and FSMRank are pairwise ranking algorithms and the others are listwise ranking algorithms.\\
 
@@
 This makes MapReduce very inefficient to process small data sets with ListNet, compared to a single-machine implementation of the algorithm.
-The MapReduce implementation of ListNet was found to be able to offer large improvements in processing time for data sets that are larger than the physical memory of the single machine otherwise available for computation.
+The MapReduce implementation of ListNet was found to be able to offer improvements in processing time for data sets that are larger than the physical memory of the single machine otherwise available for computation.
 In addition we showed that ListNet tends to converge faster when a normalisation preprocessing procedure is used.

### 2014-10-23 report/Chapters/Introduction.tex
feedback Djoerd tm introduction verwerkt

--- 2014-09-25 report/Chapters/Introduction.tex
@@
 Well-known benchmark data sets include the \emph{Yahoo!
-Learning to Rank Challenge} data set \cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, and the LETOR data sets \cite{Qin2010} that are published by Microsoft Research.\\
+Learning to Rank Challenge} data set \cite{Chapelle2011a}, the \emph{Yandex Internet Mathematics competition}\footnote{http://imat-relpred.yandex.ru/en/}, and the \emph{LETOR} data sets \cite{Qin2010} that are published by Microsoft Research.\\
 
-One of the concluding observations of the \emph{Yahoo!
-Learning to Rank Challenge} was that almost all work in the Learning to Rank field focuses on ranking accuracy.
+One of the concluding observations of the Yahoo!
+Learning to Rank Challenge was that almost all work in the Learning to Rank field focuses on ranking accuracy.
 Meanwhile, efficiency and scalability of Learning to Rank algorithms is still an underexposed research area that is likely to become more important in the near future as available data sets are rapidly increasing in size \cite{Chapelle2011b}.
@@
 The appearance of benchmark data sets for Learning to Rank gave insight in the ranking accuracy of different Learning to Rank methods.
-As observed by Liu \cite{Liu2007} and the \emph{Yahoo!
-Learning to Rank Challenge team} \cite{Chapelle2011b}, scalability of these Learning to Rank methods to large chunks of data is still an underexposed area of research.
+As observed by Liu \cite{Liu2007} and the Yahoo!
+Learning to Rank Challenge team \cite{Chapelle2011b}, scalability of these Learning to Rank methods to large chunks of data is still an underexposed area of research.
 Up to now it remains unknown whether the Learning to Rank methods that perform well in terms of ranking accuracy also perform well in terms of scalability when they are used in a parallel manner using the MapReduce framework.
@@
 The objective of this thesis is to explore the speed-up in execution time of Learning to Rank algorithms through parallelisation using the MapReduce computational model.
-The set of Learning to Rank models described in literature is of such size that it is infeasible to conduct an exhaustive experiment on all Learning to Rank models.
+The set of Learning to Rank models described in literature is of such size that it is infeasible to conduct exhaustive experiments on all Learning to Rank models.
 Therefore, we set the scope of our scalability experiment to include those Learning to Rank algorithms that have shown leading performance on relevant benchmark data sets.\\
@@
 Furthermore, given two benchmark data sets, the sets of Learning to Rank methods that are evaluated on these benchmark data sets might not be identical.
-To be able to scope our experiment on the speed-up of the most accurate Learning to Rank methods to the most accurate ranking methods we first need to identify what the most accurate Learning to Rank algorithms are.
+To be able to scope our experiment on the speed-up of the most accurate Learning to Rank methods we first need to identify what the most accurate Learning to Rank algorithms are.
 This brings us the the first research question: \\
@@
 To test the how the computational performance of Learning to Rank algorithms both on cluster and on single-node computation scales to large quantities of data, larger data sets will be created by cloning the MSLR-WEB30K data set such that the cloned queries will hold new distinct query ID's.
+
 \section{Thesis Overview}
-
 \begin{description}
-\item[Chapter II: Background]{introduces the reader to the basic principles and recent work in the fields of Learning to Rank.}
-\item[Chapter III: Related Work]{concisely describes existing work in the field of parallel Learning to Rank.}
-\item[Chapter IV: Benchmark Results]{sketches the accuracy of existing Learning to Rank methods on several benchmark data sets and describes the selection of Learning to Rank methods for the parallelisation experiments.}
-\item[Chapter V: Selected Learning to Rank Methods]{describes the algorithms and details of the Learning to Rank methods selected in Chapter IV.}
-\item[Chapter VI: Implementation]{describes implementation details of the Learning to Rank algorithms in the Hadoop framework.}
-\item[Chapter VII: Results \& Discussion]{presents and discusses speed-up results for the implemented Learning to Rank methods.}
-\item[Chapter VIII: Conclusion]{summarizes the results and answers our research questions based on the results.
+\item[Chapter 2: Background]{introduces the reader to the basic principles and recent work in the fields of Learning to Rank and the MapReduce computing model.
+Aims at bringing those unfamiliar with those topics up to speed to be able to understand the succeeding chapters of this thesis.}
+\item[Chapter 3: Related Work]{concisely describes existing work in the field of parallel and distributed Learning to Rank.}
+\item[Chapter 4: Benchmark Data Sets]{described the characteristics of the existing benchmark data sets in the Learning to Rank field.}
+\item[Chapter 5: Cross Benchmark Comparison]{described the methodology of a comparison of ranking accuracy of Learning to Rank methods across benchmark data sets and described the results of this comparison.}
+\item[Chapter 6: Selected Learning to Rank Methods]{describes the algorithms and details of the Learning to Rank methods selected in Chapter V.}
+\item[Chapter 7: Implementation]{describes implementation details of the Learning to Rank algorithms in the Hadoop framework.}
+\item[Chapter 8: MapReduce Experiments]{presents and discusses speed-up results for the implemented Learning to Rank methods.}
+\item[Chapter 9: Conclusions]{summarizes the results and answers our research questions based on the results.
 The limitations of our research as well as future research directions in the field are mentioned here.}
+\item[Chapter 10: Future Work]{describes several directions of research worthy follow-up research based on our findings.}
 \end{description}

### 2014-10-23 report/FrontBackmatter/Abstract.tex
feedback Djoerd tm introduction verwerkt

@@
 The MapReduce implementation of ListNet was found to be able to offer improvements in processing time for data sets that are larger than the physical memory of the single machine otherwise available for computation.
-In addition we showed that ListNet tends to converge faster when a normalisation preprocessing procedure is used.
-The training time of our cluster version of ListNet was found to grow better than linearly in terms of data size increase.
+In addition we showed that ListNet tends to converge faster when a normalisation preprocessing procedure is applied to the input data.
+The training time of our cluster version of ListNet was found to grow better sublinearly in terms of data size increase.
 This shows that the cluster implementation of ListNet can be used to scale the ListNet training procedure to arbitrarily large data sets, given that enough data nodes are available for computation.

### 2014-10-23 report/Chapters/Background.tex
feedback Sander pagina 59
feedback Djoerd tm pagina 19 verwerkt

@@
 \item[Feature Based]{means that all documents under investigation are represented by feature vectors.
-Those feature vectors reflect the relevance of the documents to the query, or the importance of the document in itself.}
+Those feature vectors can be used to predict the relevance of the documents to the query, or the importance of the document in itself.}
 \item[Discriminative Training]{means that the learning process can be well described by the four components of discriminative learning.
@@
 	\item[Loss Function]{in order to learn the optimal hypothesis, a training set is usually used, which contains a number of objects and their ground truth labels, sampled from the product of the input and output spaces.
-A loss function calculates the difference between the predictions $\hat{y}$ and the ground truth labels on a given set of data.
+A loss function calculates the difference between the predictions $\hat{y}$ and the ground truth labels on a given set of data.}
 	\end{description}
@@
 Figure \ref{fig:ltr_framework} shows how the machine learning process as described in Figure \ref{fig:discriminative_training} typically takes place in a ranking scenario.
-A set of queries $q_i$ with $n > i > 1$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgements of those documents $y_i$ are used together to train a model $h$.
+A set of queries $q_i$ with $1 \le i \le n$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgements of those documents $y_i$ are used together to train a model $h$.
 Model $h$ can after training be used to predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is minimal in terms of a loss function.

### 2014-10-23 report/Chapters/Background.tex
a

@@
 Figure \ref{fig:ltr_framework} shows how the machine learning process as described in Figure \ref{fig:discriminative_training} typically takes place in a ranking scenario.
-A set of queries $q_i$ with $1 \le i \le n$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgements of those documents $y_i$ are used together to train a model $h$.
+A set of queries $q_i$ with $1 \le i \le n$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevance judgements of those documents $y_i$ are used together to train a model $h$.
 Model $h$ can after training be used to predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is minimal in terms of a loss function.

### 2014-10-24 report/Chapters/Background.tex
betere verklaring voor NDCG cutoff

--- 2014-10-23 report/Chapters/Background.tex
@@
 Figure \ref{fig:ltr_framework} shows how the machine learning process as described in Figure \ref{fig:discriminative_training} typically takes place in a ranking scenario.
-A set of queries $q_i$ with $1 \le i \le n$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgements of those documents $y_i$ are used together to train a model $h$.
-Model $h$ can after training be used to predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is minimal in terms of a loss function.
+Let $q_i$ with $1 \le i \le n$ be a set of queries of size $n$.
+Let $x_j^i$ with $1 \le j \le m$ be the sets of documents of size $m$ that are associated with query $i$, in which each document is represented by a feature vector.
+The queries, the associated documents and the relevance judgements $y_i$ are jointly used to train a model $h$.
+Model $h$ can after training be used to predict a ranking of the documents for a given query, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is minimal in terms of a certain loss function.
 \begin{figure}[!h]
-\includegraphics[scale=0.25]{gfx/ltr_framework}
+\centering
+\includegraphics[scale=0.3]{gfx/ltr_framework}
 \caption{A typical Learning to Rank setting, obtained from Liu \cite{Liu2007}}
@@
 
-The predictions and the loss function might either be defined for:
-\begin{enumerate}
-\item the relevance of a single document
-\item the classification of the most relevant document out of a document-pair
-\item the ranking of documents directly
-\end{enumerate}
+Learning to Rank algorithms can be divided into three groups: the pointwise approach, the pairwise approach and the listwise approach.
+The approaches are explained in more detail in section \ref{sec:ltr_approaches}.
+The main difference between the three approaches is in the way in which they define the input space and the output space.
+\begin{description}
+\item[Pointwise] the relevance of each associated document
+\item[Pairwise] the classification of the most relevant document out for each pair of documents in the set of associated documents
+\item[Listwise] the relevance ranking of the associated documents
+\end{description}
 \smallskip
-These three approaches are in literature respectively labelled the pointwise approach, the pairwise approach and the listwise approach.
-These three approaches to Learning to Rank will be described in more detail in section \ref{sec:ltr_approaches}.
 
@@
 \subsection{Normalized Discounted Cumulative Gain}
-Cumulative gain, or its successor discounted cumulative gain and normalized discounted cumulative gain, is one of the most widely used measures for effectiveness of ranking methods.
+Cumulative gain and its successors discounted cumulative gain and normalized discounted cumulative gain are arguably the most widely used measures for effectiveness of ranking methods.
 \subsubsection{Discounted Cumulative Gain}
-There are two definitions of \ac{DCG} used in practice. \ac{DCG} at position $p$ was originally defined by J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} as\\
+There are two definitions of \ac{DCG} used in practice. \ac{DCG} for a predicted ranking of length $p$ was originally defined by J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} as\\
 
-$DCG_p = \sum\nolimits_{i=1}^p \frac{rel_i-1}{log_2(i+1)}$\\
+$DCG_{JK} = \sum\nolimits_{i=1}^p \frac{rel_i-1}{log_2(i+1)}$\\
 
@@
 
-$DCG_p = \sum\nolimits_{i=1}^p \frac{2^{rel_i-1}}{log_2(i+1)}$\\
+$DCG_{B} = \sum\nolimits_{i=1}^p \frac{2^{rel_i-1}}{log_2(i+1)}$\\
 
@@
 
-$nDCG_p = \frac{DCG_p}{IDCG_p}$\\
+$NDCG = \frac{DCG}{IDCG}$\\
 
-Table \ref{tab:example_calculation_NDCG} shows an example calculation for \ac{NDCG} for both the J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} and Burges et al.
+Often it is the case that queries in the data set differ in the number of documents that are associated with them.
+For queries with a large number of associated documents it might not always be needed to rank the complete set of associated documents, since the lower sections of this ranking might never be visited.
+\acl{NDCG} is often used with a fixed set size for the result set to mitigate this problem.
+\ac{NDCG} with a fixed set size is often called \ac{NDCG}@k, where $k$ represents the set size.\\
+
+Table \ref{tab:example_calculation_NDCG} shows an example calculation for \ac{NDCG}@k with $k=10$ for both the J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} and Burges et al.
 \cite{Burges2005} version of \ac{DCG}.\\
@@
 \hline
-rel(i) & 10 & 7 & 6 & 8 & 9 & 5 & 1 & 3 & 2 & 4 &  \\
+$rel_i$ & 10 & 7 & 6 & 8 & 9 & 5 & 1 & 3 & 2 & 4 &  \\
 \hline
@@
  &  &  &  &  &  &  &  &  &  &  &  \\
- & \multicolumn{10}{c}{Burges \cite{Burges2005} et al. \ac{NDCG} = $\frac{732.7}{788.0} = 0.9298$} &  \\ 
-  & \multicolumn{10}{c}{J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} \ac{NDCG} = $\frac{29.17}{29.96} = 0.9736$} &  \\
+  & \multicolumn{10}{c}{$NDCG_{B}@10$ = $\frac{732.7}{788.0} = 0.9298$} &  \\ 
+  & \multicolumn{10}{c}{$NDCG_{JK}@10$ = $\frac{29.17}{29.96} = 0.9736$} &  \\
 \end{tabular}
-\caption{Example calculation of \acs{NDCG}}
+\caption{Example calculation of \acs{NDCG}@10}
 \label{tab:example_calculation_NDCG}
 \end{table}
-
-One limitation of \ac{NDCG} that it does not penalise for missing documents in the result set.
-\acl{NDCG} is often used with a fixed set size for the result set to mitigate the missing document problem.
-\ac{NDCG} with a fixed set size is often called \ac{NDCG}@k, where $k$ represents the set size.
 

### 2014-10-24 report/Chapters/Background.tex
ERR verhaal klopt nu beter

@@
 Often it is the case that queries in the data set differ in the number of documents that are associated with them.
-For queries with a large number of associated documents it might not always be needed to rank the complete set of associated documents, since the lower sections of this ranking might never be visited.
+For queries with a large number of associated documents it might not always be needed to rank the complete set of associated documents, since the lower sections of this ranking might never be examined.
 \acl{NDCG} is often used with a fixed set size for the result set to mitigate this problem.
@@
 \ac{ERR} \cite{Chapelle2009} was designed based on the observation that \ac{NDCG} is based on the false assumption that the usefulness of a document at rank $i$ is independent of the usefulness of the documents at rank less than $i$.
-\ac{ERR} is based on the reasoning that users are likely to stop exploring the result list once they have found a document that satisfied their information need.
+\ac{ERR} is based on the reasoning that a user examines search results from top to bottom and at each position has a certain probability of being satisfied in his information need, at which point he stops examining the remainder of the list.
 The \ac{ERR} metric is defined as the expected reciprocal length of time that the user will take to find a relevant document.
@@
 
-$ERR = \sum\nolimits_{r=1}^n \frac{1}{r} \prod\nolimits_{i=1}^{r-1}(1-R_i)R_i$\\
+$ERR = \sum\nolimits_{r=1}^n \frac{1}{r} \prod\nolimits_{i=1}^{r-1}(1-R_i)R_r$\\
 
-where the product sequence part of the formula represents the chance that the user will stop at position $r$.\\
+where the product sequence part of the formula represents the chance that the user will stop at position $r$.
+$R_i$ in this formula represents the probability of the user being satisfied in his information need at position $i$ in the ranking.\\
+
 The algorithm to compute \ac{ERR} is shown in Listing \ref{alg:err}.
@@
  \For{$r\leftarrow 1$ \KwTo $n$}{
-	$R \leftarrow R(g[r])$\\
+	$R \leftarrow \mathcal{R}(rel_r)$\\
 	$ERR \leftarrow ERR + p * \frac{R}{r}$\\

### 2014-10-24 report/Chapters/Background.tex
ERR fixed

@@
 where the product sequence part of the formula represents the chance that the user will stop at position $r$.
-$R_i$ in this formula represents the probability of the user being satisfied in his information need at position $i$ in the ranking.\\
+$R_i$ in this formula represents the probability of the user being satisfied in his information need after assessing the document at position $i$ in the ranking.\\
 
 The algorithm to compute \ac{ERR} is shown in Listing \ref{alg:err}.
-The algorithm requires relevance grades $g_i$, $1 \le i \le n$ and mapping function $R$ that maps relevance grades to probability of relevance.
+The algorithm requires relevance grades $g_i$, $1 \le i \le n$ and mapping function $R$ that maps relevance grades to probability of relevance.\\
 
@@
  Output $ERR$
- \caption{The algorithm for computation of the \acs{ERR} metric, obtained from \cite{Chapelle2009}}
+ \caption{The algorithm for computation of the \acs{ERR} metric, obtained from Chapelle et al. \cite{Chapelle2009}}
  \label{alg:err}
 \end{algorithm}
+\vspace{0.5cm}
+In this algorithm $\mathcal{R}$ is a mapping from relevance grades to the probability of the document satisfying the information need of the user.
+Chapelle et al.
+\cite{Chapelle2009} state that there are different ways to define this mapping, but they describe one possible mapping that is based on the Burges version \cite{Burges2005} of the gain function for \ac{DCG}:\\
+
+$\mathcal{R}(r) = \frac{2^{r}-1}{2^{max\_rel}}$\\
+
+\noindent where $max\_rel$ is the maximum relevance value present in the data set.
 

### 2014-10-24 report/Chapters/Background.tex
MAP gefixt
reduce + fold

@@
 
-$AP = \frac{\sum\nolimits_{k=1}^{n}Precision(k)*relevance(k)}{\text{number of relevant docs}}$\\
+$AP(q) = \frac{\sum\nolimits_{k=1}^{n}Precision(k)*rel_k}{\text{number of relevant docs}}$\\
 
-With $k$ being the positions in the result set between 1 and $n$.
+where $n$ is the number of documents in query $q$.
+Since \ac{AP} is a binary relevance judgement metric, $rel_k$ is either 1 (relevant) or 0 (not relevant).
 Table \ref{tab:example_calculation_AP} provides an example calculation of average precision where de documents at positions 1, 5, 6 and 8 in the ranking are relevant.
@@
 MapReduce \cite{Dean2004} is a programming model invented at Google, where users specify a \emph{map} function that processes a key/value pair to generate a set of intermediate key/value pairs, and a \emph{reduce} function that merges all intermediate values associated with the same intermediate key.
-This model draws its inspiration from the field of functional programming, where \emph{map} and \emph{reduce} are commonly used functions.\\
+This model draws its inspiration from the field of functional programming, where \emph{map} and \emph{reduce} (in some functional languages called \emph{fold}) are commonly used functions.\\
 

### 2014-10-24 report/Chapters/RelatedWork.tex
related work fixes Djoerd

@@
 Most ensemble learning based Learning to Rank methods are based on boosting.
-Although boosting have shown to be an effective method in Learning to Rank, it has the drawback that the models in the ensemble are not completely mutually independent, which makes the process harder to parallelise.
-Geurts and Louppe \cite{Geurts2011} proposed a tree based ensemble method in which the trees are built using random subsets of the available features.
+Although boosting has shown to be an effective method in Learning to Rank, it has the drawback that the models in the ensemble are not completely mutually independent, which makes the process harder to parallelise.
+Geurts and Louppe \cite{Geurts2011} proposed a tree-based ensemble method in which the trees are built using random subsets of the available features.
 The authors called this method \ac{ET} and originally proposed this method for classification and regression settings instead of ranking settings \cite{Geurts2006}.

### 2014-10-24 report/Chapters/BenchmarkResults.tex
Yahoo! consistency

@@
 \section{Yahoo! Learning to Rank Challenge}
-Yahoo's observation that all existing benchmark data sets were too small to draw reliable conclusions prompted Yahoo to release two internal data sets from Yahoo!
+Yahoo!'s observation that all existing benchmark data sets were too small to draw reliable conclusions prompted Yahoo!
+to release two internal data sets from Yahoo!
 search.
@@
 Learning to Rank Challenge data set interesting.
-Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo.
+Yahoo!
+did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo!
+data collection and features used at Yahoo!.
 Instead high level descriptions of feature categories are provided.

### 2014-10-24 report/Chapters/CrossBenchmarkComparison.tex
feedback Sander pagina 59
Yahoo! consistency

@@
 Some studies with evaluation results found through in literature review were not usable for the meta-analysis.
-The following enumeration enumerates those properties that made one or more studies unusable for the meta-analysis.
+The following enumeration lists those properties that made one or more studies unusable for the meta-analysis.
 Between brackets are the studies that these properties apply to.
@@
 
-LambdaNeuralRank and CoList both acquired a perfect \ac{NWN} score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
-LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 data sets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three data sets which in addition are more relevant data sets than AOL and Yahoo Set 2 because there are more evaluation results available for the LETOR 3.0 data sets (see Table \ref{tab:ltr_methods_used}).
+LambdaNeuralRank and CoList both acquired a perfect \ac{NWN} score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo!
+Set 2.
+LARF and LRUF both scored very high scores of near 1.0 on three of the LETOR 3.0 data sets, which can be said to have a higher degree of certainty on the methods' performance because they are validated on three data sets which in addition are more relevant data sets than AOL and Yahoo!
+Set 2 because there are more evaluation results available for the LETOR 3.0 data sets (see Table \ref{tab:ltr_methods_used}).
 FenchelRank, OWPC, SmoothRank, DCMP and ListNet are in that order increasingly lower in \ac{NWN}, but increasingly higher in number of data sets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\

### 2014-10-27 report/Chapters/RelatedWork.tex
betere omschrijving evolutionary algorithms in related work

@@
 Shukla et al.
-\cite{Shukla2012} explored the parallelisation of the well-known ListNet Learning to Rank method using Spark.
-Spark is a parallel computing model that is designed for cyclic data flows which makes it more suitable for iterative algorithms.
+\cite{Shukla2012} explored the parallelisation of the well-known ListNet Learning to Rank method using Spark, which is a parallel computing model that is designed for cyclic data flows which makes it more suitable for iterative algorithms.
 Spark is incorporated into Hadoop since Hadoop 2.0.
@@
 Wang et al.
-\cite{Wang2009b} proposed a \ac{IP} solution to direct ranking metric optimisation and observed that all \ac{EA}s are generally easy to implement such that it can be carried out on different computers distributively.
+\cite{Wang2009b} proposed a \ac{IP} solution to direct ranking metric optimisation.
+\ac{IP} \cite{Musilek2006} is, like \ac{GP} \cite{Koza1992}, a paradigm in the field of evolutionary computing, but where \ac{GP} draws its inspiration from the principles of biological evolution, \ac{IP} draws its inspiration from the principles of the adaptive immune system.
+Wang et al.
+\cite{Wang2009b} observed that all \ac{EA}s, including \ac{GP} and \ac{IP} are generally easy to implement in a distributed manner.
 However, no statements on the possible speed-up of a distributed implementation of the \ac{IP} solution has been made and no speed-up experiments have been conducted.
@@
 Wang et al.
-\cite{Wang2011a,Wang2011b} proposed a parallel evolutionary algorithm based on \ac{CC}, a type of evolutionary algorithm.
+\cite{Wang2011a,Wang2011b} proposed a parallel evolutionary algorithm based on \ac{CC} \cite{Potter2000}, which is, like \ac{GP} and \ac{IP}, another paradigm in the field of evolutionary computing.
 The \ac{CC} algorithm is capable of directly optimizing non-differentiable functions, as \ac{NDCG}, in contrary to many optimization algorithms.

### 2014-10-28 implementation/java/src/letor/parallel/ListNetCluster.java
fix van de code + 64 nodes op CUSTOM-50 meting

@@ -17,14 +17,14 @@ import java.util.Random;
 
 public class ListNetCluster {
     // Initialise hyper-parameters
-    private static final DataSets.DataSet DATASET = DataSets.DataSet.MQ2007;
+    private static final DataSets.DataSet DATASET = DataSets.DataSet.CUSTOM;
     private static final double   STEPSIZE   = 0.0001; // MSLR-WEB10K: 0.0001, ohsumed: 0.01
-    private static final int      ITERATIONS = 5;
+    private static final int      ITERATIONS = 2;
     private static final int      FOLDS      = 1;
     private static final int      k          = 10; // NDCG@k
 
     // Initialise paralellisation parameters
-    private static int  dataNodes                     = 4;
+    private static int  dataNodes                     = 64;
     private static int  availableMappers              = 4*dataNodes;
     private static int  availableReducers             = 2*dataNodes;
 
@@ -41,12 +41,12 @@ public class ListNetCluster {
 
     public static void main(String[] args) throws Exception {
         // Cluster configuration
-        String clusterName          = "ltr4";
-        String clusterUser          = "admin";
+        String clusterName          = "ltr64";
+        String clusterUser          = "ltradmin";
         String clusterPassword      = "Qw!23456789";
-        String storageAccount       = "ltrorlando";
-        String containerName        = "ltr1";
-        String storageAccountKey    = "/Qv1ci7kewXJG6OMCZ9L+z+4S0PaRfN0g6LjLgAQmqiqIWje3DiXJwL04L64hXvzFb0gajuIgzVzkWyjkzYobw==";
+        String storageAccount       = "ltrmini";
+        String containerName        = "ltrsmall";
+        String storageAccountKey    = "BhtOmUzMsrWBmSdyzEGgX5zPbeBQrJNdtOypfskEXLwvkZh5lcrcf2JO023oepUhee67vt9+XrMXoHYw7Yjflg==";
         HDInsightWrapper hdw        = new HDInsightWrapper(clusterName, containerName, clusterUser, clusterPassword, storageAccount, storageAccountKey, availableReducers);
 
         ArrayList<String> pigLines  = new ArrayList<String>();
@@ -127,12 +127,12 @@ public class ListNetCluster {
[…]

### 2014-11-04 report/Chapters/CrossBenchmarkComparison.tex
kleine stabiliteitstweak clustercommunicatie + betere uitleg ListNet

@@
 No reported \ac{NDCG} performance was found in the literature study for RE-QR.
-There is a lot of certainty on the accuracy of RankBoost and Rank\ac{SVM} as both models are evaluated on the majority of data sets included in the comparison for the \ac{MAP}-metric, but given their \ac{NWN} it can said that both methods are not within the top performing Learning to Rank methods.
+There is a lot of certainty on the accuracy of RankBoost and Rank\acs{SVM} as both models are evaluated on the majority of data sets included in the comparison for the \ac{MAP}-metric, but given their \ac{NWN} it can said that both methods are not within the top performing Learning to Rank methods.
 

### 2014-11-04 report/Chapters/LTRMethods.tex
kleine stabiliteitstweak clustercommunicatie + betere uitleg ListNet

@@
 Probability distributions on permutations have been a research topic within the field of probability theory which has been extensively researched.
-ListNet uses the Plackett-Luce model \cite{Plackett1975,Luce1959}, which is one of the most well-known permutation probability distributions.
-The Plackett-Luce model defines the probability of all possible permutations $\pi$, given all document ranking scores $S$, as shown in Equation \ref{eq:plackett_luce}.
+ListNet is based on the Plackett-Luce model \cite{Plackett1975,Luce1959}, which is a permutation probability distribution over permutations that is often used in the field of econometrics.
+The Plackett-Luce model defines a probability over permutations $\pi$ from the set of all possible permutations $\Omega$, given all document ranking scores $s$.
+The Plackett-Luce model defines the probability of a permutation $\pi$ given the list as scores $s$ as shown in Equation \ref{eq:plackett_luce}.
 \begin{equation}
-P(\pi|S) = \prod\limits_{j=1}^{m}\frac{\phi(s_{\pi^{-1}(j)})}{\sum\nolimits_{u=1}^{m}\phi(s_{\pi^{-1}(u)})}
+P_s(\pi) = \prod\limits_{j=1}^{n}\frac{\phi(s_{\pi(j)})}{\sum\nolimits_{k=j}^{n}\phi(s_{\pi(k)})}
 \label{eq:plackett_luce}
 \end{equation}
-ListNet uses Gradient Descent to optimise a neural network such that its Cross Entropy loss compared to the Plackett-Luce distribution over the training data relevance labels is minimal.
+where $\pi$ is a permutation on the $n$ objects, $\phi$ is an monotonically increasing and strictly positive function.\\
+
+However, the total number of permutations of a list of $n$ documents is $n!$, therefore the calculations of the probabilities of all possible permutations quickly becomes intractable.
+To cope with this problem, the authors of ListNet propose the calculation of a \emph{top one probability} as alternative.
+The top one probability of an object $j$ equals the sum of the permutations probabilities of permutations in which object $j$ is ranked as the top object.
+It seems to be the case that the $n!$ permutation probabilities still need to be calculated for this, but we can calculate the top one probability efficiently using Equation \ref{eq:top_one}.
+\begin{equation}
+P_s(j) = \frac{\phi{s_j}}{\sum\nolimits_{k=1}^{n}\phi(s_k)}
+\label{eq:top_one}
+\end{equation}
+
+ListNet uses Gradient Descent to optimise a neural network such that its Cross Entropy loss compared to the top one over the training data relevance labels is minimal.
+For the monotonically increasing and strictly positive function $\phi$ is chosen to be the exponential function.
 Note that some sources, including Liu \cite{Liu2007}, describe ListNet as using \ac{KL divergence} as loss function.
@@
 \begin{equation}
-H(p,q) = H(p) + D_{KL}(p||q)
+L(y^{(i)},z^{(i)}) = \sum\limits_{j=1}^{n}P_{y^{(i)}}(j)\log(P_{z^{(i)}}(j))
 \label{eq:cross_entropy}
@@
 
-\noindent Equation \ref{eq:gradient_descent} describes the gradient descent step to minimise loss function $L(y^{(i)},z^{(i)}(f_\omega))$ with respect to parameter $\omega$. 
-\begin{equation}
-\Delta\omega = \frac{\partial L(y^{(i)},z^{(i)}(f_\omega))}{\partial \omega} = - \sum_{\forall g \in \mathscr{G}_k}\limits\frac{\partial P_{z^{(i)}(f_\omega)}(g)}{\partial \omega}\frac{P_{y^{(i)}}(g)}{P_{z^{(i)}(f_\omega)}(g)}
+\noindent Equation \ref{eq:gradient_descent} describes the gradient descent step to minimise listwise loss function $L(y^{(i)},z^{(i)}(f_\omega))$ with respect to parameter $\omega$. 
+\begin{multline}
+\Delta\omega = \frac{\partial L(y^{(i)},z^{(i)}(f_\omega))}{\partial \omega} = - \sum\limits_{j=1}^{n^{(i)}}P_{y^{(i)}}(x_{j}^{(i)})\frac{\partial f_{\omega}(x_{j}^{(i)})}{\partial\omega} \\ 
++ \frac{1}{\sum\nolimits_{j=1}^{n^{(i)}}exp(f_{\omega}(x_{j}^{(i)}))} \sum\limits_{j=1}^{n^{(i)}}exp(f_{\omega}(x_{j}^{(i)})) \frac{\partial f_{\omega(x_{j}^{(i)})}}{\partial\omega}
 \label{eq:gradient_descent}
-\end{equation}
+\end{multline}
 

### 2014-11-04 report/Chapters/RelatedWork.tex
kleine stabiliteitstweak clustercommunicatie + betere uitleg ListNet

@@
 \subsection{Stacked generalisation}
-% TODO: Kijken naar deze subsection, is erg zwak %
-A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable.
-\ac{DSN} is based on the stacked generalization (stacking) ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack.
-Each layer in the stack learns has the task of learning the weights of the inputs of that layer.
-The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.
+A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}.
+\ac{DSN}s are based on the stacked generalization (also known as stacking) ensemble learning method \cite{Wolpert1992}, which is a multi-layer approach where multiple learning models are stacked on top of each other in such a way that the outputs of the models are used as the input features for the models in the next layer of the stack.
+Stacked generalisation models are easy to compute in parallel, as the models within the same layer can be trained independently.
 Deng et al.
-\cite{Deng2013} % TODO: beschrijven Deng et al.%\\
+\cite{Deng2013} used a deep architecture of stacked generalisation (\ac{DSN}) to make pointwise relevance label predictions such that the \ac{MSE} of the predicted relevance labels as compared to the ground truth relevance labels are optimised.
+No speed-up results of parallel computation are reported, although it is stated repeatedly that the method is scalable and easy to compute in parallel
 

### 2014-11-04 implementation/java/src/letor/data/LargeDatasetCreator.java

@@ -16,8 +16,8 @@ import java.util.List;
  */
 public class LargeDatasetCreator {
     // Class parameters
-    private static int lowerDuplicationNumber = 0;
-    private static int upperDuplicationNumber = 20;
+    private static int lowerDuplicationNumber = 97;
+    private static int upperDuplicationNumber = 99;
 
     // Global variables
     private static String dataFolderPath = "C:/Git-data/Afstuderen/implementation/java/input";

### 2014-11-04 implementation/java/src/letor/parallel/ListNetCluster.java

@@ -24,11 +24,11 @@ public class ListNetCluster {
     private static final int      k          = 10; // NDCG@k
 
     // Initialise paralellisation parameters
-    private static int  dataNodes                     = 64;
+    private static int  dataNodes                     = 19;
     private static int  availableMappers              = 4*dataNodes;
     private static int  availableReducers             = 2*dataNodes;
 
-    private static int  duplicationNumber             = 49; // only used for CUSTOM
+    private static int  duplicationNumber             = 99; // only used for CUSTOM
     private static Metadata metadata                  = DataSets.getMetaData(DATASET, duplicationNumber);
     private static long MAX_TRAIN_SIZE                = metadata.getMax_train_size(); // ohsumed: 5151958, MQ2007: 25820919, MQ2008: 5927007, MSLR-WEB10K: 838011150, MSLR-WEB30K:
     private static long MAX_VALI_SIZE                 = metadata.getMax_vali_size(); // ohsumed: 1764005, MQ2007:  8753466, MQ2008: 2237346, MSLR-WEB10K: 280714022, MSLR-WEB30K:
@@ -42,7 +42,7 @@ public class ListNetCluster {
     public static void main(String[] args) throws Exception {
         // Cluster configuration
         String clusterName          = "ltr64";
-        String clusterUser          = "ltradmin";
+        String clusterUser          = "admin";
         String clusterPassword      = "Qw!23456789";
         String storageAccount       = "ltrmini";
         String containerName        = "ltrsmall";

### 2014-11-04 implementation/java/src/letor/parallel/util/HDInsightWrapper.java

@@ -124,9 +124,10 @@ public class HDInsightWrapper {
             };
             Future<HttpResponse> future = executor.submit(task);
             try {
-                response = future.get(10, TimeUnit.SECONDS);
+                response = future.get(12, TimeUnit.SECONDS);
             }catch (TimeoutException e){
                 System.err.println("Time out, no response obtained from cluster");
+                Thread.sleep(1000);
                 continue;
             }
 

### 2014-11-04 report/Chapters/LTRMethods.tex
ListNet-tekst nu een stuk beter :)

@@
 \label{sec:ltrmethods_listnet}
-ListNet \cite{Cao2007} is a listwise ranking function whose loss function is not directly related to information retrieval evaluation metrics.
+ListNet \cite{Cao2007} is a listwise ranking function whose loss function is not directly related to an information retrieval evaluation metric.
 ListNet's loss function is defined using a probability distribution on permutations.
 Probability distributions on permutations have been a research topic within the field of probability theory which has been extensively researched.
-ListNet is based on the Plackett-Luce model \cite{Plackett1975,Luce1959}, which is a permutation probability distribution over permutations that is often used in the field of econometrics.
+ListNet is based on the Plackett-Luce model \cite{Plackett1975,Luce1959}, which is a permutation probability distribution over permutations that is well-known in the field of econometrics.
 The Plackett-Luce model defines a probability over permutations $\pi$ from the set of all possible permutations $\Omega$, given all document ranking scores $s$.
@@
 \end{equation}
-where $\pi$ is a permutation on the $n$ objects, $\phi$ is an monotonically increasing and strictly positive function.\\
+where $\pi$ is a permutation on the $n$ objects, $\phi$ is a monotonically increasing and strictly positive function, and $s_{\pi(j)}$ is the score of the object at position $j$ of permutation $\pi$.\\
 
-However, the total number of permutations of a list of $n$ documents is $n!$, therefore the calculations of the probabilities of all possible permutations quickly becomes intractable.
-To cope with this problem, the authors of ListNet propose the calculation of a \emph{top one probability} as alternative.
+However, the total number of permutations of a list of $n$ documents is $n!$, therefore, calculating the probabilities of all possible permutations quickly becomes intractable.
+To cope with this problem, the authors of ListNet propose the calculation of a \emph{top one probability} as an alternative to the actual permutation probabilities.
 The top one probability of an object $j$ equals the sum of the permutations probabilities of permutations in which object $j$ is ranked as the top object.
-It seems to be the case that the $n!$ permutation probabilities still need to be calculated for this, but we can calculate the top one probability efficiently using Equation \ref{eq:top_one}.
+At first sight it seems to be the case that the $n!$ permutation probabilities still need to be calculated for this, but we can calculate the top one probability efficiently using Equation \ref{eq:top_one}.
 \begin{equation}
@@
 For the monotonically increasing and strictly positive function $\phi$ is chosen to be the exponential function.
+With this choice for the exponential function, we can rewrite the abstract version top one probability definition of Equation \ref{eq:top_one} with the more specific Equation \ref{eq:top_one_2}.\\
+\begin{equation}
+P_s(j) = \frac{\phi{s_j}}{\sum\nolimits_{k=1}^{n}\phi(s_k)} = \frac{exp(s_j)}{\sum\nolimits_{k=1}^{n}exp(s_k)}
+\label{eq:top_one_2}
+\end{equation}
+
 Note that some sources, including Liu \cite{Liu2007}, describe ListNet as using \ac{KL divergence} as loss function.
-\ac{KL divergence} and Cross Entropy are however identical up to an additive constant when comparing distribution $q$ against a fixed reference distribution $p$.
-This follows from the definition of Cross Entropy Loss, which is shown in Equation \ref{eq:cross_entropy}.
+This difference in definition is not relevant however, as \ac{KL divergence} and Cross Entropy are identical up to an additive constant when comparing distribution $q$ against a fixed reference distribution $p$.
+The listwise Cross Entropy loss function $L(y^{(i)},z^{(i)})$ is defined in Equation \ref{eq:cross_entropy}.
 \begin{equation}
@@
 \noindent Equation \ref{eq:gradient_descent} describes the gradient descent step to minimise listwise loss function $L(y^{(i)},z^{(i)}(f_\omega))$ with respect to parameter $\omega$.
+
 \begin{multline}
@@
 \end{multline}
+\\
+The ListNet training algorithm is similar to RankNet, with the difference that ListNet uses a listwise loss function where RankNet uses a pairwise loss function.
+Algorithm \ref{alg:listnet} shows the pseudo-code of the ListNet training algorithm.\\
 
-\noindent Algorithm \ref{alg:listnet} shows the pseudo-code of the ListNet training phase.\\
 \LinesNumbered

### 2014-11-04 implementation/java/src/letor/parallel/ListNetCluster.java

@@ -24,11 +24,11 @@ public class ListNetCluster {
     private static final int      k          = 10; // NDCG@k
 
     // Initialise paralellisation parameters
-    private static int  dataNodes                     = 19;
+    private static int  dataNodes                     = 64;
     private static int  availableMappers              = 4*dataNodes;
     private static int  availableReducers             = 2*dataNodes;
 
-    private static int  duplicationNumber             = 99; // only used for CUSTOM
+    private static int  duplicationNumber             = 19; // only used for CUSTOM
     private static Metadata metadata                  = DataSets.getMetaData(DATASET, duplicationNumber);
     private static long MAX_TRAIN_SIZE                = metadata.getMax_train_size(); // ohsumed: 5151958, MQ2007: 25820919, MQ2008: 5927007, MSLR-WEB10K: 838011150, MSLR-WEB30K:
     private static long MAX_VALI_SIZE                 = metadata.getMax_vali_size(); // ohsumed: 1764005, MQ2007:  8753466, MQ2008: 2237346, MSLR-WEB10K: 280714022, MSLR-WEB30K:

### 2014-11-05 report/Chapters/LTRMethods.tex
kleine fix comment Djoerd

@@
  \For{$t\leftarrow 1$ \KwTo $T$}{
- 	//check if the early stopping criterion is satisfied\\
+ 	\emph{//check if the early stopping criterion is satisfied}\\
  	\If{$||g_{t}||_{\infty} + \langle d_{t}, -Kw_{t} \rangle \le \epsilon$}{
- 	//here $d_{t} = \nabla f^{*}(-Kw_{t}) = \frac{\partial f^{*}(-Kw)}{\partial(Kw)}|w=w_{t}$, \\
- 	// $\langle x,y \rangle$ represents the inner products of vectors $x$ and $y$, and\\
- 	// $g_{t} = d_{t}^{T}K$\\
+ 	\emph{//here $d_{t} = \nabla f^{*}(-Kw_{t}) = \frac{\partial f^{*}(-Kw)}{\partial(Kw)}|w=w_{t}$,} \\
+ 	\emph{// $\langle x,y \rangle$ represents the inner products of vectors $x$ and $y$, and}\\
+ 	\emph{// $g_{t} = d_{t}^{T}K$}\\
  	return $w_{t}$ as ranking predictor $w$
  	}
- //greedily choose a feature to update\\
+ \emph{//greedily choose a feature to update}\\
  Choose $j_{t} = \argmax_{j}|(g_t)_j)|$\\
- //compute an appropriate step size\\
+ \emph{//compute an appropriate step size}\\
  Let $\mu_t = \argmax_{0 \le \mu_{t} \le 1} D((1-\mu_{t})w_{t} + \mu_{t}\sign((g_{t})_{j_{t}})e^{j_{t}})$\\
- //update the model with the chosen feature and step size\\
+ \emph{//update the model with the chosen feature and step size}\\
  Update $w_{t+1}=(1-\mu_{t})w_{t} + \mu_{t}\sign((g_{t})j_{t})e^{j_{t}}$
@@
  	\While{true}{
- 		//projection step\\
+ 		\emph{//projection step}\\
  		$w_t= \argmin_{w_t:w_t\in \mathbb{R}_{+}^{2d}} Q(w_t,z_t)$\\
@@
  	\If{$\frac{|F(w_t)-F(w_{t-1})|}{|F(w_{t-1})|} \le \epsilon_s$}{
- 		//early stopping criterion\\
+ 		\emph{//early stopping criterion}\\
  		break

### 2014-11-13 report/Chapters/Results.tex
verder aan presentatie
presentatie Avanade af + eerste feedbackpuntjes Dolf

@@
 To validate our ListNet cluster implementation we run experiments on the LETOR 3.0 OHSUMED, the LETOR 4.0 datasets MQ2007 and MQ2008 and compare their performance with the RankLib implementation of ListNet and with the official evaluation results in terms of \ac{NDCG}@10.
-Hadoop job scheduling overhead is large when processing relatively small datasets (as we will explain in more detail in section \ref{ssec:speedup}), therefore \ac{NDCG}@10 experiments are limited to five training iterations and are run only on one of the five folds.
+Hadoop job scheduling overhead is relatively large when processing relatively small datasets (as we will explain in more detail in section \ref{ssec:speedup}), therefore \ac{NDCG}@10 experiments are limited to five training iterations and are run only on one of the five folds.
 Because of the stochastic nature of the ListNet ranking procedure, which is a result of random weight initialisation prior to the first training iteration, the \ac{NDCG}@10 performance is not guaranteed to be identical between multiple runs on the same input data.
-To account for this non-determinism of ListNet, we repeat each experiment five times and report the average as well as the standard deviation of the resulting rankings on the testing fold in terms of \ac{NDCG}@10.\\
+To account for this non-determinism in ListNet, we repeat each experiment five times and report the average as well as the standard deviation of the resulting rankings on the testing fold in terms of \ac{NDCG}@10.\\
 
@@
 \centering
-\begin{tabular}{p{3.4cm}p{2.6cm}p{2.6cm}p{2.6cm}}\toprule
- &  OHSUMED (LETOR 3.0 version) & MQ2008 & MQ2007 \\
+\begin{tabular}{p{3.4cm}p{3.0cm}p{2.6cm}p{2.6cm}}\toprule
+ &  OHSUMED  (LETOR 3.0) & MQ2008 & MQ2007 \\
 \midrule

### 2014-11-13 report/Chapters/Introduction.tex
feedbackpuntjes Dolf

@@
 
-Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
+In the context of Learning to Rank applied to information retrieval, Luhn \cite{Luhn1957} was the first to propose a model that assigned relevance scores to documents given a query back in 1957.
 This started a transformation of the Information Retrieval field from a focus on the binary classification task of labelling documents as either \emph{relevant} or \emph{not relevant} into a \emph{ranked retrieval} task that aims at ranking the documents from most to least relevant.
 Research in the field of ranking models has long been based on manually designed ranking functions, such as the well-known BM25 model \cite{Robertson1994}, that simply rank documents based on the appearance of the query terms in these documents.
-The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective models.
+The increasing amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective and more accurate ranking models.
 Learning to Rank is the relatively new research area that covers the use of machine learning models for the ranking task.\\

### 2014-11-13 report/FrontBackmatter/Abstract.tex
feedbackpuntjes Dolf

@@
 \chapter*{Abstract}
-Learning to rank is an increasingly important scientific field that comprises the use of machine learning for the ranking task.
+Learning to rank is an increasingly important task within the scientific fields of machine learning and information retrieval, that comprises the use of machine learning for the ranking task.
 New learning to rank methods are generally evaluated in terms of ranking accuracy on benchmark test collections.
@@
 Furthermore, little research is done in the field of scalability of the training procedure of Learning to Rank methods, to prepare us for input data sets that are getting larger and larger.
-This paper concerns both the comparison of Learning to Rank methods using a sparse set of evaluation results on benchmark data sets, as well as the speed-up that can be achieved by parallelising Learning to Rank methods using MapReduce.\\
+This thesis concerns both the comparison of Learning to Rank methods using a sparse set of evaluation results on benchmark data sets, as well as the speed-up that can be achieved by parallelising Learning to Rank methods using MapReduce.\\
 
@@
 
-In the second part of this thesis we analysed the speed-up of the ListNet training algorithm when implemented in the MapReduce computing model.
+In the second part of this thesis we analyse the speed-up of the ListNet training algorithm when implemented in the MapReduce computing model.
 We found that running ListNet on MapReduce comes with a job scheduling overhead in the range of 150-200 seconds per training iteration.

### 2014-11-13 report/Chapters/Introduction.tex
nieuw puntje dolf

@@
 
-Some research has been done in the area of parallel or distributed machine learning \cite{Chu2007,Chang2007}.
-However, almost none of these studies include the Learning to Rank sub-field of machine learning.
-The field of efficient Learning to Rank has been getting some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012}, since Liu \cite{Liu2007} first stated its growing importance back in 2007.
+Some research has been done in the area of parallel or distributed machine learning \cite{Chu2007,Chang2007}, with the aim to speed-up machine learning computation or to increase the size of the data sets that can be processed with machine learning techniques.
+However, almost none of these parallel or distributed machine learning studies target the Learning to Rank sub-field of machine learning.
+The field of efficient Learning to Rank has received some attention lately \cite{Asadi2013a,Asadi2013b,Busa-Fekete2012,Sousa2012,Shukla2012}, since Liu \cite{Liu2007} first stated its growing importance back in 2007.
 Only a few of these studies \cite{Sousa2012,Shukla2012} have explored the possibilities of efficient Learning to Rank through the use of parallel programming paradigms.\\

### 2014-11-13 report/Chapters/Introduction.tex
nieuw puntje Dolf

@@
 \label{sec:goals}
-The objective of this thesis is to explore the speed-up in execution time of Learning to Rank algorithms through parallelisation using the MapReduce computational model.
 The set of Learning to Rank models described in literature is of such size that it is infeasible to conduct exhaustive experiments on all Learning to Rank models.
@@
 Given two ranking methods, there might be non-agreement between evaluation results on different benchmarks on which ranking method is more accurate.
-Furthermore, given two benchmark data sets, the sets of Learning to Rank methods that are evaluated on these benchmark data sets might not be identical.
-To be able to scope our experiment on the speed-up of the most accurate Learning to Rank methods we first need to identify what the most accurate Learning to Rank algorithms are.
-This brings us the the first research question: \\
+Furthermore, given two benchmark data sets, the sets of Learning to Rank methods that are evaluated on these benchmark data sets might not be identical.\\
+
+The objective of this thesis is twofold.
+firstly, we aim to provide insight in the most accurate ranking methods while taking into account evaluation results on multiple benchmark data sets.
+Secondly, we use this insight to scope an experiment on the speed-up of the most accurate Learning to Rank methods to explore the speed-up in execution time of Learning to Rank algorithms through parallelisation using the MapReduce computational model.
+The first part of the objective of this thesis brings us to the first research question: \\
 
@@
 \bigskip
-Ranking accuracy is an ambiguous concept, as there exists several metrics that can be used to express the accuracy of a ranking.
+Ranking accuracy is an ambiguous concept, as several several metrics exist that can be used to express the accuracy of a ranking.
 We will explore several metrics for ranking accuracy in section \ref{sec:how_to_evaluate_a_ranking}.\\

### 2014-11-13 report/Chapters/Introduction.tex
weer dingen van Dolf :)

@@
 
-To research the speed-up's dependence on the amount of data that is being processed, the training computations will be performed on data sets of varying sizes.
+To research the speed-up's dependence on the amount of processed data, the training computations will be performed on data sets of varying sizes.
 We use the well-known benchmark collections LETOR 3.0, LETOR 4.0 and MSLR-WEB30/40K as a starting set of data sets for our experiments.
@@
 \begin{description}
-\item[Chapter 2: Background]{introduces the reader to the basic principles and recent work in the fields of Learning to Rank and the MapReduce computing model.
-Aims at bringing those unfamiliar with those topics up to speed to be able to understand the succeeding chapters of this thesis.}
+\item[Chapter 2: Background]{introduces the basic principles and recent work in the fields of Learning to Rank and the MapReduce computing model.}
 \item[Chapter 3: Related Work]{concisely describes existing work in the field of parallel and distributed Learning to Rank.}
-\item[Chapter 4: Benchmark Data Sets]{described the characteristics of the existing benchmark data sets in the Learning to Rank field.}
-\item[Chapter 5: Cross Benchmark Comparison]{described the methodology of a comparison of ranking accuracy of Learning to Rank methods across benchmark data sets and described the results of this comparison.}
+\item[Chapter 4: Benchmark Data Sets]{describes the characteristics of the existing benchmark data sets in the Learning to Rank field.}
+\item[Chapter 5: Cross-Benchmark Comparison]{describes the methodology of a comparison of ranking accuracy of Learning to Rank methods across benchmark data sets and describes the results of this comparison.}
 \item[Chapter 6: Selected Learning to Rank Methods]{describes the algorithms and details of the Learning to Rank methods selected in Chapter V.}

### 2014-11-13 report/Chapters/Background.tex
dolf :)

@@
-%************************************************
-This chapter is written with the goal to provide the reader with a brief introduction in the Learning to Rank field and the MapReduce computational model.
-Models and theories explained in this chapter can be regarded as prior knowledge needed to understand the subsequent chapters of this thesis.
+This chapter provides an introduction to Learning to Rank and MapReduce.
+Knowledge about the models and theories explained in this chapter is required to understand the subsequent chapters of this thesis.
 \section{A basic introduction to Learning to Rank}

### 2014-11-13 report/Chapters/Background.tex
puntjes Dolf

@@
 \begin{figure}[!h]
-\includegraphics[scale=0.26]{gfx/descriminative_training}
+\centering
+\includegraphics[scale=0.36]{gfx/descriminative_training}
 \caption{Machine learning framework for Learning to Rank, obtained from Liu \cite{Liu2007}}
@@
 
-Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning to Rank method when it is \emph{feature based} and uses \emph{discriminative training}, in which the concepts \emph{feature based} and \emph{discriminative training} are itself defined as:
+Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning to Rank method when it is \emph{feature based} and uses \emph{discriminative training}, in which the concepts \emph{feature-based} and \emph{discriminative training} are themselves defined as:
 \begin{description}
-\item[Feature Based]{means that all documents under investigation are represented by feature vectors.
-Those feature vectors can be used to predict the relevance of the documents to the query, or the importance of the document in itself.}
+\item[Feature-Based]{means that all objects under investigation are represented by feature vectors.
+In a Learning to Rank for Information Retrieval case, this means that the feature vectors can be used to predict the relevance of the documents to the query, or the importance of the document itself.}
 \item[Discriminative Training]{means that the learning process can be well described by the four components of discriminative learning.
@@
 	\begin{description}
-	\item[Input Space]{contains the objects under investigation. Usually objects are represented by feature vectors, extracted according to different applications.}
+	\item[Input Space]{contains the objects under investigation. Usually objects are represented by feature vectors, extracted from the objects themselves.}
 	\item[Output Space]{contains the learning target with respect to the input objects.}
@@
 \centering
-\includegraphics[scale=0.3]{gfx/ltr_framework}
+\includegraphics[scale=0.36]{gfx/ltr_framework}
 \caption{A typical Learning to Rank setting, obtained from Liu \cite{Liu2007}}
@@
 Cumulative gain and its successors discounted cumulative gain and normalized discounted cumulative gain are arguably the most widely used measures for effectiveness of ranking methods.
+Cumulative Gain, without discounting factor and normalisation step, is defined as\\
+
+$CG_{k} = \sum\nolimits_{i=1}^k rel_i$
+
 \subsubsection{Discounted Cumulative Gain}

### 2014-11-13 report/Chapters/Background.tex
meer Dolf

@@
 \subsection{Mean Average Precision}
-\ac{AP} \cite{Zhu2004} is an often used binary relevance judgement based metric that can be seen as a trade-off between precision and recall that is defined as\\
+\ac{AP} \cite{Zhu2004} is an often used binary relevance-judgement-based metric that can be seen as a trade-off between precision and recall that is defined as\\
 
@@
 $MAP = \frac{\sum\nolimits_{q=1}^{Q}AP(q)}{Q}$\\
+
+In this formula $Q$ is the number queries.
 
@@
 \subsection{Pairwise Approach}
-Pointwise Learning to Rank methods have the drawback that they optimise real valued expected relevance, while evaluation metrics like \ac{NDCG} and \ac{ERR} are only impacted by a change in expected relevance when that change impacts a pairwise preference.
+Pointwise Learning to Rank methods have the drawback that they optimise real-valued expected relevance, while evaluation metrics like \ac{NDCG} and \ac{ERR} are only impacted by a change in expected relevance when that change impacts a pairwise preference.
 The pairwise approach solves this drawback of the pointwise approach by regarding ranking as pairwise classification.\\

### 2014-11-13 report/Chapters/Background.tex
:)

@@
 \label{sec:cross_validation}
-A cross-validation experiment \cite{kohavi1995}, sometimes called rotation estimation, is an experimental set-up for model evaluation where the data is split into $k$ chunks of approximately equal size, called \emph{folds}.
+A cross-validation experiment \cite{kohavi1995}, sometimes called rotation estimation, is an experimental set-up for evaluation where the data is split into $k$ chunks of approximately equal size, called \emph{folds}.
 One of the folds is used as validation set, one of the folds is used as test set, and the rest of the $k - 2$ folds are used as training data.

### 2014-11-13 report/Chapters/AppendixA.tex
nDCG gevonden + uitlijning gefixt
uitlijning gefixt
based -> -based

@@
 \scalebox{0.87}{
-\begin{tabular}{p{0.29cm}|p{6.88cm}||p{0.29cm}|p{6cm}}
+\begin{tabular}{p{0.29cm}|p{6.5cm}||p{0.29cm}|p{6cm}}
  \textbf{ID} & \textbf{Feature Description} & \textbf{ID} & \textbf{Feature Description}\\ 
@@
 \scalebox{0.87}{
-\begin{tabular}{p{0.29cm}|p{7.26cm}||p{0.29cm}|p{9.55cm}}
+\begin{tabular}{p{0.29cm}|p{6.50cm}||p{0.29cm}|p{8cm}}
  \textbf{ID} & \textbf{Feature Description} & \textbf{ID} & \textbf{Feature Description}\\ 
@@
 
-6 & \acs{IDF} of body 				& 41& Sitemap based term propagation\\
-7 & \acs{IDF} of anchor 				& 42& Sitemap based score propagation\\
-8 & \acs{IDF} of title 				& 43& Hyperlink based score propagation: weighted in-link\\
-9 & \acs{IDF} of \acs{URL} 			& 44& Hyperlink based score propagation: weighted out-link\\
-10& \acs{IDF} of whole document 		& 45& Hyperlink based score propagation: uniform out-link\\
+6 & \acs{IDF} of body 				& 41& Sitemap-based term propagation\\
+7 & \acs{IDF} of anchor 				& 42& Sitemap-based score propagation\\
+8 & \acs{IDF} of title 				& 43& Hyperlink-based score propagation: weighted in-link\\
+9 & \acs{IDF} of \acs{URL} 			& 44& Hyperlink-based score propagation: weighted out-link\\
+10& \acs{IDF} of whole document 		& 45& Hyperlink-based score propagation: uniform out-link\\
 
-11& \acs{TF-IDF} of body 			& 46& Hyperlink based feature propagation: weighted in-link\\
-12& \acs{TF-IDF} of anchor 			& 47& Hyperlink based feature propagation: weighted out-link\\
-13& \acs{TF-IDF} of title 			& 48& Hyperlink based feature propagation: uniform out-link\\
+11& \acs{TF-IDF} of body 			& 46& Hyperlink-based feature propagation: weighted in-link\\
+12& \acs{TF-IDF} of anchor 			& 47& Hyperlink-based feature propagation: weighted out-link\\
+13& \acs{TF-IDF} of title 			& 48& Hyperlink-based feature propagation: uniform out-link\\
 14& \acs{TF-IDF} of \acs{URL}			& 49& HITS authority\\

### 2014-11-13 report/Chapters/RelatedWork.tex
based -> -based

@@
 \section{Literature study characteristics}
+The literature study described in this section is performed with the aim of getting insight in relevant existing techniques for large scale Learning to Rank.
 The literature research is performed by using the bibliographic databases Scopus and Web of Science with the following search query:\\
@@
 
-On Google Scholar, the defined search query resulted in 3300 documents. Because it infeasible to evaluate all 3300 studies we focus on the first 300 search results.\\
+On Google Scholar, the defined search query resulted in 3300 documents. Because it infeasible to evaluate all 3300 studies we focus on the first 300 search results as ranked by Google Scholar.\\
 
@@
 Experiments showed that the solution scales linearly in number of hyperparameter combinations.
-However, the risk of overfitting grows overwhelmingly as the number of hyperparameter combinations grow, even when validation sets grow large.\\
+However, the risk of overfitting grows as the number of hyperparameter combinations grow, even when validation sets grow large.\\
 
@@
 Hardware accelerators are special purpose processors designed to speed up compute-intensive tasks.
-An \ac{FPGA} and a \ac{GPU} are two different types of hardware that can achieve better performance on some tasks though parallel computing.
+A \ac{FPGA} and a \ac{GPU} are two different types of hardware that can achieve better performance on some tasks though parallel computing.
 In general, \ac{FPGA}s provide better performance while \ac{GPU}s tend to be easier to program \cite{Che2008}.
@@
 Yan et al.
-\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run, the Neural Network based LambdaRank Learning to Rank algorithm.
+\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run, the Neural-Network-based LambdaRank Learning to Rank algorithm.
 This architecture achieved a 29.3X speed-up compared to the software implementation, when evaluated on data from a commercial search engine.
 The exploration of \ac{FPGA} for Learning to Rank showed additional benefits other than the speed-up originally aimed for.
-In their latest publication \cite{Yan2012} the \ac{FPGA} based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.\\
+In their latest publication \cite{Yan2012} the \acs{FPGA}-based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.\\
 
 Xu et al.
-\cite{Xu2007b,Xu2009} designed an \ac{FPGA}-based accelerator to reduce the training time of the RankBoost algorithm \cite{Freund2003}, a pairwise ranking function based on Freund and Schapire's AdaBoost ensemble learning method \cite{Freund1997}.
+\cite{Xu2007b,Xu2009} designed an \acs{FPGA}-based accelerator to reduce the training time of the RankBoost algorithm \cite{Freund2003}, a pairwise ranking function based on Freund and Schapire's AdaBoost ensemble learning method \cite{Freund1997}.
 Xu et al.
@@
 \cite{DeSousa2012} proposed a \ac{GPGPU} approach to improve both training time and query evaluation through \ac{GPU} use.
-An association rule based Learning to Rank approach, proposed by Veloso et al.
+An association-rule-based Learning to Rank approach, proposed by Veloso et al.
 \cite{Veloso2008}, has been implemented using the \ac{GPU} in such a way that the set of rules van be computed simultaneously for each document.
@@
 \cite{Tyree2011} described a way of parallelising \ac{GBDT} models for Learning to Rank where the boosting step is still executed sequentially, but instead the construction of the regression trees themselves is parallelised.
-The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}, which are build layer-by-layer.
+The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}, which are built layer-by-layer.
 The calculations needed for building each new layer in the tree are divided among the nodes, using a master-worker paradigm.
@@
 Wang et al.
-\cite{Wang2011a,Wang2011b} proposed a parallel evolutionary algorithm based on \ac{CC} \cite{Potter2000}, which is, like \ac{GP} and \ac{IP}, another paradigm in the field of evolutionary computing.
+\cite{Wang2011a,Wang2011b} proposed a parallel evolutionary-algorithm-based on \ac{CC} \cite{Potter2000}, which is, like \ac{GP} and \ac{IP}, another paradigm in the field of evolutionary computing.
 The \ac{CC} algorithm is capable of directly optimizing non-differentiable functions, as \ac{NDCG}, in contrary to many optimization algorithms.
@@
 
-Another \ac{ADMM} Learning to Rank based approach was proposed by Boyd et al.
+Another \ac{ADMM}-based Learning to Rank approach was proposed by Boyd et al.
 \cite{Boyd2012}.
-They implemented an \ac{ADMM} based Learning to Rank method in Pregel \cite{Malewicz2010}, a parallel computation framework for graph computations.
-No experimental results on parallelisation speed-up were reported on this Pregel based approach.
+They implemented an \ac{ADMM}-based Learning to Rank method in Pregel \cite{Malewicz2010}, a parallel computation framework for graph computations.
+No experimental results on parallelisation speed-up were reported on this Pregel-based approach.
 \subsection{Bregman Divergences and Monotone Retargeting}
 Acharyya et al.
-\cite{Acharyya2012, Acharyya2013} proposed a Learning to Rank method searches for an order preserving transformation (\emph{monotone retargeting}) of the target scores that is easier for a regressor to fit.
+\cite{Acharyya2012, Acharyya2013} proposed a Learning to Rank method that makes use of an order preserving transformation (\emph{monotone retargeting}) of the target scores that is easier for a regressor to fit.
 This approach is based on the observation that it is not necessary to fit scores exactly, since the evaluation is dependent on the order and not on the pointwise predictions themselves.\\
@@
 Long et al.
-\cite{Long2012} described special case of their pairwise cross-domain factor Learning to Rank method using distributed optimization of \ac{SGD} based on Hadoop MapReduce.
-Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce based method described by  Zinkevich et al.
+\cite{Long2012} described a special case of their pairwise cross-domain factor Learning to Rank method using distributed optimization of \acs{SGD}-based on Hadoop MapReduce.
+Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce-based method described by  Zinkevich et al.
 \cite{Zinkevich2010} has been used.
@@
 \subsection{Randomisation}
-Most ensemble learning based Learning to Rank methods are based on boosting.
+Most ensemble-learning-based Learning to Rank methods are based on boosting.
 Although boosting has shown to be an effective method in Learning to Rank, it has the drawback that the models in the ensemble are not completely mutually independent, which makes the process harder to parallelise.

### 2014-11-13 report/Chapters/RelatedWork.tex
meer Dolf

@@
 Parallelisation of the \ac{SGD} optimization algorithm was performed using the MapReduce-based method described by  Zinkevich et al.
-\cite{Zinkevich2010} has been used.
+\cite{Zinkevich2010}.
 Real world data from Yahoo!
@@
 Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure called boosting \cite{Schapire1990}.
-The invention of the boosting method resulted in an increasing focus in machine learning research on methods to combine multiple weak models, which are called ensemble learning methods.
+The invention of the boosting method resulted in an increasing focus on ensemble methods, which are methods of combining multiple weak models.
 Well-known ensemble methods include gradient boosting \cite{Friedman2001}, bagging \cite{Breiman1996}, AdaBoost \cite{Freund1997} and stacked generalisation (often called stacking) \cite{Wolpert1992}.\\
@@
 Stochastic gradient boosting is a minor alteration to regular gradient boosting that, at each iteration of the algorithm, trains a base learner on a sub-sample of the training set that is drawn at random without replacement \cite{Friedman2002}.
-Experiments showed the Hadoop implementation to resulted in too expensive communication cost to be useful.
+Experiments showed the Hadoop implementation to result in too high communication cost to be useful.
 Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm.
@@
 
-Svore and Burges \cite{Svore2010,Svore2012} designed two approaches to train the LambdaMART \cite{Wu2008} model in a distributed manner.
+Svore and Burges \cite{Svore2010,Svore2012} designed two approaches to train the \mbox{LambdaMART} \cite{Wu2008} model in a distributed manner.
 Their first approach is applicable in case the whole training set fits in main memory on every node, in that case the tree split computations of the trees in the LambdaMART tree-ensemble are split instead of the data.
@@
 Yandex researchers Pavlov et al.
-\cite{Pavlov2010} were the first to proposed a boosting-wrapped-in-bagging approach, which they called BagBoo.
+\cite{Pavlov2010} were the first to propose a boosting-wrapped-in-bagging approach, which they called BagBoo.
 The boosting step itself is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.\\
@@
 Deng et al.
-\cite{Deng2013} used a deep architecture of stacked generalisation (\ac{DSN}) to make pointwise relevance label predictions such that the \ac{MSE} of the predicted relevance labels as compared to the ground truth relevance labels are optimised.
-No speed-up results of parallel computation are reported, although it is stated repeatedly that the method is scalable and easy to compute in parallel
+\cite{Deng2013} used a deep architecture of stacked generalisation (\ac{DSN}) to make pointwise predictions of relevance labels such that the \ac{MSE} of those predictions compared to the ground truth relevance labels are minimised.
+No speed-up results of parallel computation are reported, although it is stated repeatedly that the method is scalable and easy to compute in parallel.
 

### 2014-11-13 report/Chapters/BenchmarkResults.tex
kleine stabiliteitstweak clustercommunicatie + betere uitleg ListNet
based -> -based
plaatjes beter

@@
 Benchmark data sets enable fair comparisons of ranking models over a fixed set of documents and queries.
-The approach of using fixed benchmark data sets to compare the performance of Information Retrieval systems under equal circumstances has been the standard evaluation methodology in the Information Retrieval field since the release of the Cranfield collection \cite{Cleverdon1966}.\\
+The approach of using fixed benchmark data sets to compare the performance of Information Retrieval systems under equal circumstances has been the standard evaluation methodology in the Information Retrieval field since the release of the Cranfield collection \cite{Cleverdon1966} in 1966.\\
 
 This chapter describes benchmark characteristics, e.g.
-collection size and features, of Learning to Rank benchmark collections and data sets and gives an overview of the performance of baseline algorithms on these collections and data sets.\\
+collection size and features, of Learning to Rank benchmark data sets and benchmark collections (sets consisting of multiple benchmark data sets) and gives an overview of the performance of baseline algorithms on these collections and data sets.\\
 
@@
 Gottschalk and Vogel used a combination of RandomForest models and \ac{GBDT} models.
-Pavlov and Brunk used a regression based model using the BagBoo \cite{Pavlov2010} ensemble technique, which combines bagging and boosting.
+Pavlov and Brunk used a regression-based model using the BagBoo \cite{Pavlov2010} ensemble technique, which combines bagging and boosting.
 Sorokina used a similar combination of bagging and boosting that is called Additive Groves \cite{Sorokina2007}.\\
@@
 	\begin{itemize}
-	\item Rank\ac{SVM} \cite{Herbrich1999,Joachims2002}
+	\item Rank\acs{SVM} \cite{Herbrich1999,Joachims2002}
 	\item RankBoost \cite{Freund2003}

### 2014-11-13 report/Chapters/Conclusions.tex
feedback Dolf door :)

@@
 Our comparison methodology comprises of two components: 1) \ac{NWN}, which provides insight in the ranking accuracy of the learning to rank method, and 2) \ac{IWN}, which gives insight in the degree of certainty concerning the performance of the ranking accuracy.
-Based on our literature search for evaluation results on well-known benchmarks collections, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.
+Based on our literature search for evaluation results on well-known benchmarks collections, insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.
 However, no closing arguments can be formulated on which learning to rank methods are most accurate.
@@
 From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.
-More evaluation runs are needed for the methods on the left side of Figure \ref{fig:normalised_winning_number_all}.
+More evaluation runs are needed to increase the certainty of the ranking accuracy of the methods that were found to have low \ac{IWN} values.
 Our work contributes to this by identifying promising learning to rank methods that researchers could focus on in performing additional evaluation runs.
@@
 We found that running ListNet on a Hadoop cluster using the MapReduce computing model comes with its own cost in the form of a job scheduling overhead in the range of 150-200 seconds per training iteration.
-This makes Hadoop very inefficient for the processing of small data sets, where the Hadoop overhead tend to make up a large share of the total processing time.
+This makes Hadoop very inefficient for the processing of small data sets, where the Hadoop overhead tends to make up a large share of the total processing time.
 For small data sets where the constant 150-200 seconds job scheduling overhead is a large fraction of the total processing time, single-machine computation, which does not have this job scheduling overhead, is found to be faster than Hadoop MapReduce computation.

### 2014-11-13 report/Chapters/CrossBenchmarkComparison.tex
plaatjes beter
feedback Dolf door :)

@@
 
-From now on this adjusted version of Winning Number will be references to as \ac{NWN}. The mathematical definition of \ac{NWN} is\\
+From now on this adjusted version of Winning Number will be referred to as \ac{NWN}. The mathematical definition of \ac{NWN} is\\
 
@@
 \begin{figure}[!h]
-\includegraphics[scale=0.285]{gfx/ndcg3_winnum}
+\centering
+\includegraphics[scale=0.32]{gfx/ndcg3_winnum}
 \caption{\acs{NDCG}@3 comparison of Learning to Rank methods}
@@
 \begin{figure}[!h]
-\includegraphics[scale=0.285]{gfx/ndcg5_winnum}
+\centering
+\includegraphics[scale=0.35]{gfx/ndcg5_winnum}
 \caption{\acs{NDCG}@5 comparison of Learning to Rank methods}
@@
 \begin{figure}[!h]
-\includegraphics[scale=0.285]{gfx/ndcg10_winnum}
+\centering
+\includegraphics[scale=0.30]{gfx/ndcg10_winnum}
 \caption{\acs{NDCG}@10 comparison of Learning to Rank methods}
@@
 \begin{figure}[!h]
-\includegraphics[scale=0.285]{gfx/map_winnum}
+\centering
+\includegraphics[scale=0.30]{gfx/map_winnum}
 \caption{\acs{MAP} comparison of Learning to Rank methods}
@@
 \begin{figure}[!h]
-\includegraphics[scale=0.33]{gfx/combined_normalized_winnum}
+\centering
+\includegraphics[scale=0.38]{gfx/combined_normalized_winnum}
 \caption{Cross-benchmark comparison of Learning to Rank methods}
@@
 LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both \ac{IWN} and \ac{NWN}.
+In other words, we can conclude these methods to be pareto-optimal in the trade-off between certainty and ranking accuracy.
 LRUF is the only method that achieved this in all \ac{NDCG} comparisons, the \ac{MAP} comparison as well as the cross-metric comparison.
@@
 \section{Limitations}
-In the \ac{NWN} calculation the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
+In the \ac{NWN} calculation, the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
 By calculating it in this way, we implicitly make the assumption that the Learning to Rank methods are (approximately) distributed uniformly over the benchmarks, such that the average Learning to Rank method tested are approximately equally hard for each data set.
-It could be the case however that this assumption is false and that significantly more accurate Learning to Rank methods are evaluated on some data sets than on other data sets.
-\\
+This assumption goes wrong in the case that, for a given data set $s$, the evaluation results are on average obtained from significantly better Learning to Rank methods compared to another data sets.
+As a result, the $IWN$ of Learning to Rank methods evaluated on $s$ will be undervalued.\\
 
-A second limitation is that the data sets on which Learning to Rank methods have been evaluated can not always be regarded a random choice.
-It might be the case that some researchers chose to publish results for exactly those benchmark data sets that showed the most positive results for their Learning to Rank method.\\
+A second limitation is that the data sets on which Learning to Rank methods have been evaluated cannot always be regarded a random choice.
+It might be the case that some researchers specifically targeted benchmark data sets that showed the most positive results for the Learning to Rank method.
+A result of this publication bias in the evaluation results will be that the Learning to Rank method obtains an artificially high \ac{NWN}.
+High \ac{IWN} scores are achieved by having evaluation results on multiple data sets, which decreases the possibilities for a publication bias.\\
 
@@
 
-Based on our literature search for evaluation results on well-known benchmarks collections, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.
-However, no closing arguments can be formulated on which learning to rank methods are most accurate.
-LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the learning to rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
+Based on our literature search, LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were found to be the learning to rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy (e.g.
+pareto-optimal in the trade-off between certainty and ranking accuracy).
 From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.\\
 
-LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the Learning to Rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
-From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.
-For more definite conclusions on the relative performance of these five methods, more evaluation runs on are desirable for the methods on the left side on the list on benchmark data sets that these methods have not yet been evaluated on.\\
-
-More evaluation runs are needed for the methods on the left side of Figure \ref{fig:normalised_winning_number_all}.
+For more definite conclusions on the relative performance of these five methods, more evaluation runs on are desirable for the methods on the left side on the list on benchmark data sets that these methods have not yet been evaluated on.
+More evaluation runs are especially needed for the methods on the left side of Figure \ref{fig:normalised_winning_number_all}.
 Our work contributes to this by identifying promising learning to rank methods that researchers could focus on in performing additional evaluation runs.\\
@@
 In the following chapters of this thesis, concerning parallel execution of the Learning to Rank training phase, the scope will be limited to the five methods that turned out to be superior Learning to Rank methods in terms of ranking accuracy and certainty about this ranking accuracy: LRUF, FSMRank, FenchelRank, SmoothRank and ListNet.
-Although it can not be concluded that these methods are inarguably the most accurate Learning to Rank methods, a strong presumption has been raised that these five Learning to Rank are accurate ranking methods.\\
+Although it cannot be concluded that these methods are inarguably the most accurate Learning to Rank methods, a strong presumption has been raised that these five Learning to Rank are accurate ranking methods.\\

### 2014-11-13 report/Chapters/Future.tex
feedback Dolf door :)

@@
-Potential future areas of research that follow from this thesis can be categorised into three categories. These categories of potential future work are described in the sections below.
+Follow-up research from this research can be categorised into three categories. These categories of potential future work are described in the sections below.
 
@@
 \section{Optimisation Algorithms}
-Lin \cite{Lin2013} stated that the gradient descent optimisation procedure can often be replaced with other optimisation procedures with faster convergence properties (often at the cost of being more computationally expensive per iteration), in order to make iterative machine learning algorithms perform better on MapReduce.
+The gradient descent optimisation procedure can often be replaced with other optimisation procedures with faster convergence properties, often at the cost of being more computationally expensive per iteration.
+Lin \cite{Lin2013} hypothesised that the replacement of gradient descent by optimisation methods with faster convergence properties can especially be beneficial in a MapReduce setting, as the MapReduce job scheduling overhead leads to a high constant factor in iteration time.
 An example of an optimisation procedures that fits this higher convergence rate at the cost of higher per-iteration computational costs is L-BFGS \cite{Liu1989}, or any other quasi-Newton optimisation method.

### 2014-11-13 report/Chapters/LTRMethods.tex
presentatie Avanade af + eerste feedbackpuntjes Dolf
feedback Dolf door :)

@@
 \begin{equation}
-P_s(j) = \frac{\phi{s_j}}{\sum\nolimits_{k=1}^{n}\phi(s_k)} = \frac{exp(s_j)}{\sum\nolimits_{k=1}^{n}exp(s_k)}
+P_s(j) = \frac{\phi(s_j)}{\sum\nolimits_{k=1}^{n}\phi(s_k)} = \frac{exp(s_j)}{\sum\nolimits_{k=1}^{n}exp(s_k)}
 \label{eq:top_one_2}
@@
 The starting point in SmoothRank is set to either the solution of a simple linear regression, or alternatively to the solution of Rank\acs{SVM}.
-Since this starting point is expected to already be a good solution, a regulariser term is added to the SmoothRank objective function to prevent the solution from deviating to much from the starting point.
+Since this starting point is expected to already be a good solution, a regulariser term is added to the SmoothRank objective function to prevent the solution from deviating too much from the starting point.
 The regularised smooth objective function is formulated as $\lambda||w-w_0||_2$ where $\lambda$ is a hyper-parameter tuned on the validation set, $w_0$ is the starting point solution, and $||x||_2$ is the $\ell2$-norm regularisation function defined as $||x||_2 = \sqrt{\sum\nolimits_ix_i^2}$.\\
@@
 In case the document set of the search engine changes, i.e.
-new documents are indexed or old documents are removed, LRUF has an Increase Action-Set Size (IAS) and a Reduce Action-Set Size (RAS) procedure defined to adapt to the new document set without needing a complete retraining of the model.
+new documents are indexed or old documents are removed, LRUF has an Increase Action-Set Size (IAS) and a Reduce Action-Set Size (RAS) procedure to adapt to the new document set without needing a complete retraining of the model.
 Because this thesis focusses on model accuracy and scalability the IAS and RAS procedure of LRUF will not be explained in further detail.\\
 
-Torkestani \cite{Torkestani2012b} does not provide pseudo code specifically for the training phase of the LRUF algorithm.
+Torkestani \cite{Torkestani2012b} does not provide pseudo-code specifically for the training phase of the LRUF algorithm.
 Instead an algorithm including both ranking and automaton update is presented, and included in Algorithm \ref{alg:lruf}.

### 2014-11-13 report/Chapters/Results.tex
feedback Dolf door :)

@@
 Note that speed-up, following this definition, is not computable for larger data sets, as it is not computationally feasible to perform computation on data sets larger than CUSTOM-8 with a single machine.
-As an alternative, Figure \ref{fig:speedup_train_time} shows visualises speed-up by plotting cluster size against training time.\\
+As an alternative, Figure \ref{fig:speedup_train_time} visualises the speed-up by plotting the size of the MapReduce cluster against training time.\\
 
@@
 Figure \ref{fig:listnet_processing_speed} shows the processing speed in bytes per second.
-Our observation that RankLib ListNet performs very slow for data sets that do not fit in physical memory and virtual memory is needed is very notable in this graph.
-Furthermore, this graph shows how both increase in number in data nodes in a cluster and increase in input data size result in an increase in processing speed.
+Our observation that RankLib ListNet processes very slow for data sets that do not fit in physical memory and virtual memory is needed is very notable in this graph.
+Furthermore, this graph shows how both increase in number of data nodes in a cluster and increase in input data size result in an increase in processing speed.
 The sub-linear growth of processing speed as a factor of input data size can again be explained as a result of the non-parallelisable Hadoop job scheduling part of the operation.\\

### 2014-11-18 implementation/RankLib/src/ciir/umass/edu/experimenter/NormalizationExperimenter.java
SItools + mslr-metingen

@@ -27,14 +27,12 @@ public class NormalizationExperimenter {
 
         Set<DataSets.DataSet>       datasets            = new HashSet<DataSets.DataSet>();
         // Add all LETOR 3.0 data sets
-        datasets.add(DataSets.DataSet.OHSUMED);
-        datasets.add(DataSets.DataSet.TD2003);datasets.add(DataSets.DataSet.TD2004);
-        datasets.add(DataSets.DataSet.NP2003);datasets.add(DataSets.DataSet.NP2004);
-        datasets.add(DataSets.DataSet.HP2003);datasets.add(DataSets.DataSet.HP2004);
+        datasets.add(DataSets.DataSet.MSLR_WEB30K);
+        datasets.add(DataSets.DataSet.MSLR_WEB10K);
 
         BufferedWriter out = null;
         try {
-            out = new BufferedWriter(new FileWriter("normalization_measurements.csv"));
+            out = new BufferedWriter(new FileWriter("normalization_measurements_mslr_web.csv"));
         } catch (IOException e) {
             e.printStackTrace();
         }
@@ -47,7 +45,7 @@ public class NormalizationExperimenter {
                 if(iterations % 2 != 0)
                     continue;
 
-                double learningRate = 0.01;
+                double learningRate = 0.0001;
 
                 if (dataset.equals(DataSets.DataSet.NP2003) || dataset.equals(DataSets.DataSet.NP2004))
                     learningRate = 0.001;

### 2014-11-18 implementation/RankLib/src/ciir/umass/edu/experimenter/WrappedRunner.java

@@ -16,7 +16,7 @@ public class WrappedRunner {
         Integer                     iterations          = 1000;
         Integer                     k                   = 10;
         AbstractParameterizedRanker handler             = new ListNetHandler();
-        DataSets.DataSet            dataset             = DataSets.DataSet.TD2004;
+        DataSets.DataSet            dataset             = DataSets.DataSet.MSLR_WEB10K;
         Integer                     duplicationNumber   = 1; // Only relevant in case of dataset Custom
 
         FoldRunHandler ltrWrapper = null;

### 2014-11-18 implementation/RankLib/src/ciir/umass/edu/metric/NDCGScorer.java

@@ -32,6 +32,8 @@ public class NDCGScorer extends MetricScorer {
         gainMap.put(0,0.0);
         gainMap.put(1,1.0);
         gainMap.put(2,3.0);
+        gainMap.put(3,7.0);
+        gainMap.put(4,15.0);
 	}
 	public NDCGScorer(int k)
 	{
@@ -151,7 +153,9 @@ public class NDCGScorer extends MetricScorer {
 		for(int i=1;i<=size;i++)
 		{
             if(i==1)
-			    dcg += gainMap.get(rel.get(idx[i-1]));
+			    dcg += gainMap.get(
+                        rel.get(idx[i-1])
+                );
             else
                 dcg += gainMap.get(rel.get(idx[i-1]))/SimpleMath.logBase2(i);
 		}

### 2014-11-18 measurements/processing_speed_single.R

@@ -3,6 +3,7 @@ library(ggplot2)
 library(boot)
 library(mgcv)
 library(stringr)
+library(sitools)
 library(scales)
 measure = read.csv("raw_data.csv") # read csv file
 name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
@@ -35,7 +36,7 @@ d <- d + geom_point(size=2) + geom_line() +
 		panel.border     = element_blank(),
 		panel.background = element_blank(),
 		axis.line = element_line(color='black')
-	) + scale_x_continuous(expand = c(0.01, 0)) + scale_y_continuous(expand = c(0, 0)) +
+	) + scale_x_continuous(labels=f2si, expand = c(0.01, 0)) + scale_y_continuous(labels=f2si, expand = c(0, 0)) +
 	xlab("Dataset Size (in Byte)") +
 	ylab("Throughput (in Byte/Second)") +
 	labs(colour = "Execution mode", shape = "Execution mode") 

### 2014-11-18 measurements/processing_speed_single_logx.R

@@ -3,6 +3,7 @@ library(ggplot2)
 library(boot)
 library(mgcv)
 library(stringr)
+library(sitools)
 library(scales)
 measure = read.csv("raw_data.csv") # read csv file
 name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
@@ -28,14 +29,14 @@ measure$throughput <- measure$dataSize/measure$seconds
 measure$serial.parallel <- name
 d <- ggplot(data=measure, aes(x=dataSize, y=throughput, colour=serial.parallel, shape=serial.parallel))
 d <- d + geom_point(size=2) + geom_line() + 
-	scale_x_log10() + theme_bw() + 
+	scale_x_log10(labels=f2si) + theme_bw() + 
 	theme(
 		panel.grid.major = element_blank(),
 		panel.grid.minor = element_blank(),
 		panel.border     = element_blank(),
 		panel.background = element_blank(),
 		axis.line = element_line(color='black')
-	) + scale_y_continuous(expand = c(0, 0)) +
+	) + scale_y_continuous(labels=f2si,expand = c(0, 0)) +
 	xlab("Dataset Size (in Byte)") +
 	ylab("Throughput (in Byte/Second)") +
 	labs(colour = "Execution mode", shape = "Execution mode") 

### 2014-11-18 measurements/processing_speeed_faceted.R

@@ -1,6 +1,7 @@
 setwd("C:/Git-data/Afstuderen/measurements")
 library(ggplot2)
 library(mgcv)
+library(sitools)
 library(stringr)
 library(scales)
 measure = read.csv("raw_data.csv") # read csv file
@@ -34,7 +35,7 @@ d <- d + geom_point(size=2) + geom_line() +
 		panel.border     = element_blank(),
 		panel.background = element_blank(),
 		axis.line = element_line(color='black')
-	) + scale_y_continuous(expand = c(0, 10000000))+	
+	) + scale_y_continuous(expand = c(0, 10000000))+
 	facet_wrap(~serial.parallel)+
 	xlab("Dataset size (in Byte)") +
 	ylab("Training iteration time (in Seconds)") +

### 2014-11-18 measurements/time_single.R

@@ -3,6 +3,7 @@ library(ggplot2)
 library(mgcv)
 library(stringr)
 library(scales)
+library(sitools)
 measure = read.csv("raw_data.csv") # read csv file
 name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
 
@@ -33,7 +34,9 @@ d <- d + geom_point(size=2) + geom_line() +
 		panel.border     = element_blank(),
 		panel.background = element_blank(),
 		axis.line = element_line(color='black')
-	) + scale_y_continuous(expand = c(0, 0))+	
+	) + 
+	scale_x_continuous(labels=f2si)+
+	scale_y_continuous(expand = c(0, 0))+	
 	xlab("Dataset size (in Byte)") +
 	ylab("Training iteration time (in Seconds)") +
 	labs(colour = "Execution mode", shape = "Execution mode") 

### 2014-11-18 measurements/time_single_logx.R

@@ -2,6 +2,7 @@ setwd("C:/Git-data/Afstuderen/measurements")
 library(ggplot2)
 library(mgcv)
 library(stringr)
+library(sitools)
 library(scales)
 measure = read.csv("raw_data.csv") # read csv file
 name <- paste(measure$serial.parallel, measure$nodeCount, sep = " ")
@@ -26,14 +27,15 @@ measure$seconds <- measure$trainTime/1000
 measure$serial.parallel <- name
 d <- ggplot(data=measure, aes(x=dataSize, y=seconds, colour=serial.parallel, shape=serial.parallel))
 d <- d + geom_point(size=2) + geom_line() + 
-	scale_x_log10() + theme_bw() + 
+	scale_x_log10(labels=f2si) + theme_bw() + 
 	theme(
 		panel.grid.major = element_blank(),
 		panel.grid.minor = element_blank(),
 		panel.border     = element_blank(),
 		panel.background = element_blank(),
 		axis.line = element_line(color='black')
-	) + scale_y_continuous(expand = c(0, 0))+	
+	) + 
+	scale_y_continuous(expand = c(0, 0))+	
 	xlab("Dataset size (in Byte)") +
 	ylab("Training iteration time (in Seconds)") +
 	labs(colour = "Execution mode", shape = "Execution mode") 

### 2014-11-19 report/Chapters/LTRMethods.tex
smoothrank-uitleg verbeterd + nieuwe metingen

@@
 SmoothRank enables direct optimisation of the non-convex and discontinuous evaluation measure by smoothing, that is, approximating the rank position.
-In this section, we apply the SmoothRank approximation method to the \ac{NDCG} evaluation measure for illustration, but the same procedure can be applied to \ac{MAP} or any other information retrieval measure.
-The smoothing function used in SmoothRank is based on the softmax activation function \cite{Bridle1990}, which is often used in neural networks.
-The softmax function is formally defined as shown in Equation \ref{eq:softmax_algorithm}.
+In this section, we apply the SmoothRank approximation method to the \ac{NDCG} evaluation measure for illustration, but the same procedure can be applied to \ac{ERR}, \ac{MAP}, or any other information retrieval measure.
+The smoothing function used in SmoothRank is based on the softmax function \cite{Bridle1990} that is often used as an activation function for the neurons in neural networks.
+The softmax function as a smoothing framework for ranking evaluation measures in given in Equation \ref{eq:softmax_algorithm}.
 \begin{equation}
-p_i = \frac{e^{f_i/\sigma}}{\sum\nolimits_{j=1}^{m}e^{f_j/\sigma}}
+p_i = \frac{e^{f_i/\sigma}}{\sum\nolimits_{j=1}^{n}e^{f_j/\sigma}}
 \label{eq:softmax_algorithm}
@@
 
-where $\sigma$ is the smoothing parameter.
+where $p_i$ is a probability over documents $i \in \{1,...,n\}$ given the outputs $f_1,...,f_n$ of the evaluation function on the document relevance labels, with $\sigma$ being a smoothing parameter.
 Chapelle et al.
@@
 \begin{equation}
-h_{i,j} = e^{-\frac{(f(x_i))-(f(x_{d(j)}))^2}{\sigma}}\Big/\sum\nolimits_{k=1}^{m}e^{-\frac{(f(x_k))-f(x_{d(j)})^2}{\sigma}}
+h_{i,j} = e^{-\frac{(f(x_i))-(f(x_{d(j)}))^2}{\sigma}}\Big/\sum\nolimits_{k=1}^{n}e^{-\frac{(f(x_k))-f(x_{d(j)})^2}{\sigma}}
 \label{eq:soft_NDCG}
@@
 
-It can be shown that the derivative of the smoothed \ac{NDCG} version shown in Equation \ref{eq:soft_NDCG} and the smoothed versions of other \ac{IR} metrics can be calculated in $\mathcal{O}(m^2)$, which enable fast gradient descent optimisation.
+It can be shown that the derivative of the smoothed \ac{NDCG} version shown in Equation \ref{eq:soft_NDCG} and the smoothed versions of other \ac{IR} metrics can be calculated in $\mathcal{O}(n^2)$, which enable fast gradient descent optimisation.
 The optimisation step in SmoothRank uses the nonlinear conjugate gradient method  with Polak-Ribiere update \cite{Shewchuk1994}, which is a type of gradient descent method.

### 2014-11-19 report/Chapters/LTRMethods.tex
smoothrank beter

@@
 
-where $p_i$ is a probability over documents $i \in \{1,...,n\}$ given the outputs $f_1,...,f_n$ of the evaluation function on the document relevance labels, with $\sigma$ being a smoothing parameter.
+where $p_i$ is a probability over documents $i \in \{1,...,n\}$ given the outputs $f_1,...,f_n$ of the ranking function on the documents, with $\sigma$ being a smoothing parameter.
 Chapelle et al.
-\cite{Chapelle2010} apply this softmax function to the \ac{NDCG} formula and hereby introduce a soft version of indicator variable $h_{i,j}$ through the computation shown in Equation \ref{eq:soft_NDCG}.
+\cite{Chapelle2010} apply this softmax function to the \ac{NDCG} formula and hereby introduce a soft version $h_{i,j}$ of the indicator variable $\mathbb{1}_{i,j}$ ("is document $i$ ranked at the $j$-th position?") through the computation shown in Equation \ref{eq:soft_NDCG}.
 
@@
 It can be shown that the derivative of the smoothed \ac{NDCG} version shown in Equation \ref{eq:soft_NDCG} and the smoothed versions of other \ac{IR} metrics can be calculated in $\mathcal{O}(n^2)$, which enable fast gradient descent optimisation.
-The optimisation step in SmoothRank uses the nonlinear conjugate gradient method  with Polak-Ribiere update \cite{Shewchuk1994}, which is a type of gradient descent method.
+The optimisation step in SmoothRank uses the non-linear conjugate gradient method  with Polak-Ribiere update \cite{Shewchuk1994}.
 This optimisation method is prone to local optima, which is alleviated by adding a pre-calculated, better than naive, starting point and by adding a regularisation term.\\
@@
 The choice of the smoothing parameter $\sigma$ in Equation \ref{eq:soft_NDCG} is important, because a too small value makes the function more non-smooth and therefore harder to optimise, while a too large value results in a optimisation function with optima that substantially differ from the true optimal rankings.
-To deal with the problem of choosing the smoothing parameter, SmoothRank uses an annealing method where the optimisation procedure starts with a large $\sigma$ and iteratively reduces it by dividing it by two at each iteration.
+To deal with the problem of choosing the smoothing parameter, SmoothRank uses an annealing procedure where optimisation starts with a large $\sigma$ in the first iteration and iteratively reduces it by dividing it by two at each iteration.
 Algorithm \ref{alg:smoothrank} shows the algorithm that summarises all steps of the SmoothRank method.\\

### 2014-11-19 report/Chapters/LTRMethods.tex
eindelijk is ListNet een helder verhaal :):) + nieuwe metingen

@@
 \begin{equation}
-P_s(j) = \frac{\phi{s_j}}{\sum\nolimits_{k=1}^{n}\phi(s_k)}
+P_s(j) = \frac{\phi(s_j)}{\sum\nolimits_{k=1}^{n}\phi(s_k)}
 \label{eq:top_one}
@@
 \begin{equation}
-L(y^{(i)},z^{(i)}) = \sum\limits_{j=1}^{n}P_{y^{(i)}}(j)\log(P_{z^{(i)}}(j))
+L(y^{(i)},z^{(i)}) = -\sum\limits_{j=1}^{n}P_{y^{(i)}}(j)\log(P_{z^{(i)}}(j))
 \label{eq:cross_entropy}
 \end{equation}
-where $H(p)$ is the entropy of $p$ and $D_{KL}(p||q)$ is the \ac{KL divergence} of $q$ from $p$.\\
+where $P_s(j)$ is the top one probability of object $j$ given relevance score $s$. This relevance score $s$ is either the true relevance score $y^{(i)}$ or the predicted relevance score $z^{(i)}$.\\
 
-\noindent Equation \ref{eq:gradient_descent} describes the gradient descent step to minimise listwise loss function $L(y^{(i)},z^{(i)}(f_\omega))$ with respect to parameter $\omega$.
+\noindent Equation \ref{eq:gradient_descent} describes the gradient descent step to minimise listwise loss function $L(y^{(i)},z^{(i)}(f_\omega))$ with respect to the model parameters $\omega$.
+In this equation, $z^{(i)}(f_\omega)$ represents the predicted relevance scores of object $i$ when a neural network $f$ with weights $\omega$ is used.
 
@@
 \begin{equation}
-\min_x(f(x)-g(x)) = \max_p(f_*(p)-f^*(p))
+\min_x(f(x)-g(x)) = \max_p(g_*(p)-f^*(p))
 \label{eq:fenchel_duality_theorem}
 \end{equation}
-\noindent where $f^*$ is the convex, and $f_*$ the concave conjugate of $f$.\\
+\noindent where $f^*$ is the convex conjugate of $f$ and $g_*$ is the concave conjugate of $g$.\\
 
-To ease applying Fenchel's duality theorem (Equation 7) to the $\ell_1$-regularised pairwise loss function (Equation \ref{eq:pairwise_l1_loss}), Lai et al.
+To ease applying Fenchel's duality theorem (Equation \ref{eq:fenchel_duality_theorem}) to the $\ell_1$-regularised pairwise loss function (Equation \ref{eq:pairwise_l1_loss}), Lai et al.
 define $D(w)$ as $D(w) = -G(w)$.
-Equation \ref{eq:fenchel_dual} shown the resulting Fenchel dual of the $\ell_1$-regularised pairwise loss function, which is the loss function that is used in FenchelRank.
+Equation \ref{eq:fenchel_dual} shows the resulting Fenchel dual of the $\ell_1$-regularised pairwise loss function, which is the loss function that is used in FenchelRank.
 \begin{equation}

### 2014-11-19 report/Chapters/LTRMethods.tex
betere uitleg FSMRank

@@
 \noindent Equation \ref{eq:gradient_descent} describes the gradient descent step to minimise listwise loss function $L(y^{(i)},z^{(i)}(f_\omega))$ with respect to the model parameters $\omega$.
-In this equation, $z^{(i)}(f_\omega)$ represents the predicted relevance scores of object $i$ when a neural network $f$ with weights $\omega$ is used.
+In this equation, $z^{(i)}(f_\omega)$ represents the predicted relevance scores of object $i$ when a neural network $f$ with weight vector $\omega$ is used.
 
@@
  \KwIn{number of iterations $T$ and learning rate $\eta$}
- Initialize parameter $\omega$\\
+ Initialize weight vector $\omega$\\
  \For{$t\leftarrow 1$ \KwTo $T$}{
@@
 \begin{equation}
-p_i = \frac{e^{f_i/\sigma}}{\sum\nolimits_{j=1}^{n}e^{f_j/\sigma}}
+p_i = \frac{e^{f(i)/\sigma}}{\sum\nolimits_{j=1}^{n}e^{f(j)/\sigma}}
 \label{eq:softmax_algorithm}
@@
 Chapelle et al.
-\cite{Chapelle2010} apply this softmax function to the \ac{NDCG} formula and hereby introduce a soft version $h_{i,j}$ of the indicator variable $\mathbb{1}_{i,j}$ ("is document $i$ ranked at the $j$-th position?") through the computation shown in Equation \ref{eq:soft_NDCG}.
+\cite{Chapelle2010} apply this softmax function to the \ac{NDCG} formula and hereby introduce a soft version $h_{i,j}$ of the indicator variable $\mathbb{1}_{i,j}$ that represents a function that outputs the boolean representation of the answer to the questions: "is document $i$ ranked at the $j$-th position?".
+$h_{i,j}$ is calculated through the formula shown in Equation \ref{eq:soft_NDCG}.
 
@@
 \end{equation}
+
+where $d(j)$ is a function that returns the index of the document that was ranked at position $j$ by ranking function $f$.\\
 
@@
 The choice of the smoothing parameter $\sigma$ in Equation \ref{eq:soft_NDCG} is important, because a too small value makes the function more non-smooth and therefore harder to optimise, while a too large value results in a optimisation function with optima that substantially differ from the true optimal rankings.
-To deal with the problem of choosing the smoothing parameter, SmoothRank uses an annealing procedure where optimisation starts with a large $\sigma$ in the first iteration and iteratively reduces it by dividing it by two at each iteration.
-Algorithm \ref{alg:smoothrank} shows the algorithm that summarises all steps of the SmoothRank method.\\
+To deal with the problem of choosing the smoothing parameter, SmoothRank uses an annealing procedure where optimisation starts with a large $\sigma$ in the first iteration and iteratively reduces it by dividing it by two at each iteration.\\
+
+Algorithm \ref{alg:smoothrank} shows the algorithm that summarises all steps of the SmoothRank method.
+In this algorithm, $A_q(w,\sigma)$ is the smoothed version of the \ac{NDCG} metric, calculated for ranking function $f$ and using smoothing parameter $\sigma$, based on the smoothed indicator function $h_{i,j}$ as defined in Equation \ref{eq:soft_NDCG}.
+This smoothed version of the Burges \ac{NDCG} is defined in Equation \ref{eq:smooth_NDCG}.
+\begin{equation}
+A_q(f,\sigma) = \sum\nolimits_{i=1}^{n} \sum\nolimits_{j=1}^{n} \frac{2^{rel_i}-1}{\log_2(j+1)} h_{i,j}
+\label{eq:smooth_NDCG}
+\end{equation}
 
@@
 \noindent where $I_{C}(w)$ is a function that is 0 if condition $C$ is satisfied, and $\infty$ otherwise.
+$w$ is the weight vector with the weights of the features.
+$r$ is a hyper-parameter of the model that represents the radius of the $\ell_1$ ball.
 $m$ is the dimension of the data, $p$ is the number of comparable object pairs, $K$ is a matrix in $\mathbb{R}^{p \times m}$ that contains pairwise information.
@@
  \For{$t\leftarrow 1$ \KwTo $T$}{
- 	\emph{//check if the early stopping criterion is satisfied}\\
+ 	\emph{// check if the early stopping criterion is satisfied}\\
  	\If{$||g_{t}||_{\infty} + \langle d_{t}, -Kw_{t} \rangle \le \epsilon$}{
- 	\emph{//here $d_{t} = \nabla f^{*}(-Kw_{t}) = \frac{\partial f^{*}(-Kw)}{\partial(Kw)}|w=w_{t}$,} \\
+ 	\emph{// here $d_{t} = \nabla f^{*}(-Kw_{t}) = \frac{\partial f^{*}(-Kw)}{\partial(Kw)}|w=w_{t}$,} \\
  	\emph{// $\langle x,y \rangle$ represents the inner products of vectors $x$ and $y$, and}\\
@@
  	}
- \emph{//greedily choose a feature to update}\\
+ \emph{// greedily choose a feature to update}\\
  Choose $j_{t} = \argmax_{j}|(g_t)_j)|$\\
- \emph{//compute an appropriate step size}\\
+ \emph{// compute an appropriate step size}\\
  Let $\mu_t = \argmax_{0 \le \mu_{t} \le 1} D((1-\mu_{t})w_{t} + \mu_{t}\sign((g_{t})_{j_{t}})e^{j_{t}})$\\
- \emph{//update the model with the chosen feature and step size}\\
+ \emph{// update the model with the chosen feature and step size}\\
  Update $w_{t+1}=(1-\mu_{t})w_{t} + \mu_{t}\sign((g_{t})j_{t})e^{j_{t}}$
@@
 
-Let $S = {(q_k, \hat{X}^{(q_k), Y^{(q_k)}})}_{k=1}^n$ be a training set with queries $q_k$, corresponding retrieval objects $\hat{X}^{(q_k)}$, and corresponding relevance labels $Y^{(q_k)}$.
+Let $S = {(q_k, x^{(q_k)}, y^{(q_k)})}_{k=1}^n$ be a training set with queries $q_k$, corresponding retrieval objects $x^{(q_k)}$, and corresponding relevance labels $y^{(q_k)}$.
 Let $||x||_1$ be the $l_1$-norm regularisation function that is defined as $||x||_1 = \sum\nolimits_i |x_i|$.
+FSMRank defines a normalisation preprocessing step following the formula $x' = \frac{x - min(x)}{max(x)-min(x)}$ to scale the input features to a $[0,1]$-range.
 Let $\oslash$ be the element-wise division operator.
@@
 \begin{equation}
-\min_{\hat{w}} \frac{\lambda_1}{2} \hat{w}^T \hat{A}\hat{w} + \lambda_2 ||\hat{w}\oslash\hat{s}||_1 + f(\hat{w}, (\hat{X}^{(q_k)}, Y^{(q_k)})_{k=1}^n)
+\min_{\hat{w}} \frac{\lambda_1}{2} \hat{w}^T \hat{A}\hat{w} + \lambda_2 ||\hat{w}\oslash\hat{s}||_1 + f(\hat{w}, (x^{(q_k)}, y^{(q_k)})_{k=1}^n)
 \label{eq:fsmrank_optimisation_function}
@@
 The well-known Pearson correlation coefficient is used to calculate similarity matrix $\hat{A}$, based on the values for the features in the training data.
-$\lambda_1$ is a hyper-parameter of the model and can be used to set the weight of the similarity penalty term.\\
+$\lambda_1$ is a hyper-parameter of the model and can be used to set the weight of the similarity penalty term.
+$\hat{w}$ is the weight vector of the features.\\
 
 The second term of Equation \ref{eq:fsmrank_optimisation_function} contains a $\ell_1$-norm regularisation term to select the effective features from the feature set.
-$\lambda_2$ is a hyper-parameter of the model and can be used to set the weight of the regularisation term.\\
+$\lambda_2$ is a hyper-parameter of the model and can be used to set the weight of the regularisation term.
+The $\hat{s}$ term is a vector containing importance scores of the features.
+It is not defined which metric is used to calculate these importance scores, but the range of the importance scores is said to be [0,1].\\
 
@@
 As stated, FSMRank uses an  extended version of Nesterov's accelerated gradient method.
-This optimisation method can handle optimisation problems in the form of $\min_w \l(w) + r(w)$ where $l(w)$ is a convex function with Lipschitz gradient and $r(w)$ is convex, but non-smooth.
+This optimisation method can handle optimisation problems in the form of $\min_w l(w) + r(w)$ where $l(w)$ is a convex function with Lipschitz gradient and $r(w)$ is convex, but non-smooth.
 The optimization function of Equation \ref{eq:fsmrank_optimisation_function} is reformulated to match this form as presented in Equation \ref{eq:fsmrank_optimisation_reformulated}.
@@
 \noindent where $Q(w_t,z_t)$ is a combination of the non-smooth part $r(w_t)$ and a quadratic approximation of the smooth part $l(w_t)$.
-$\lambda_1$, $\lambda_2$ and Lipschitz constant $L_0$ are input parameters of the algorithm and can be optimised through cross-validation or on a validation set.\\
+The $l(z_t)$ function is defined as $l(z_t) = \lambda_1 / 2z_t^{T}Az_t+f(z_t)$.
+$\lambda_1$, $\lambda_2$ (both used in function $Q$) and Lipschitz constant $L_0$ are input parameters of the algorithm and can be optimised through cross-validation or on a validation set.\\
 
@@
 \begin{algorithm}[H]
- \KwData{training set $S = {(q_k,\hat{X}^{(q_k)},Y^(q_k)}_{k=1}^{n}$}
+ \KwData{training set $S = {(q_k,x^{(q_k)},y^(q_k)}_{k=1}^{n}$}
  \KwIn{$\lambda_1, \lambda_2, T \text{ and } L_0$}

### 2014-11-19 implementation/RankLib/src/ciir/umass/edu/experimenter/FoldRunHandler.java

@@ -162,18 +162,22 @@ public class FoldRunHandler {
         return features;
     }
 
-    private List<RankList> normalize(List<RankList> rl, float[] minmax){
-        for(int i=0; i<rl.size(); i++){
-            for(int j=0; j<rl.get(i).size(); j++){
-                DataPoint dp = rl.get(i).get(j);
+    private List<RankList> normalize(List<RankList> rls, float[] minmax){
+        for(int i=0; i<rls.size(); i++){
+            RankList rl = rls.get(i);
+            for(int j=0; j<rls.get(i).size(); j++){
+                DataPoint dp = rls.get(i).get(j);
                 for(int f=0; f < features.length; f++){
                     if(minmax[(2*f)+1] > 0)
                         dp.setFeatureValue(f, dp.getFeatureValue(f) - minmax[2*f] / (minmax[(2*f)+1] - minmax[2*f]));
                     else
                         dp.setFeatureValue(f, 0.0f);
+                    rl.remove(j);
+                    rl.add(dp);
                 }
+                rls.set(i, rl);
             }
         }
-        return rl;
+        return rls;
     }
 }

### 2014-11-19 implementation/RankLib/src/ciir/umass/edu/experimenter/NormalizationExperimenter.java

@@ -27,7 +27,7 @@ public class NormalizationExperimenter {
 
         Set<DataSets.DataSet>       datasets            = new HashSet<DataSets.DataSet>();
         // Add all LETOR 3.0 data sets
-        datasets.add(DataSets.DataSet.MSLR_WEB30K);
+        //datasets.add(DataSets.DataSet.MSLR_WEB30K);
         datasets.add(DataSets.DataSet.MSLR_WEB10K);
 
         BufferedWriter out = null;
@@ -42,9 +42,6 @@ public class NormalizationExperimenter {
         for (int iterations = 1; iterations < maxIterations; iterations++) {
             for(DataSets.DataSet dataset: datasets) {
 
-                if(iterations % 2 != 0)
-                    continue;
-
                 double learningRate = 0.0001;
 
                 if (dataset.equals(DataSets.DataSet.NP2003) || dataset.equals(DataSets.DataSet.NP2004))
@@ -53,6 +50,13 @@ public class NormalizationExperimenter {
                     if(dataset.equals(DataSets.DataSet.HP2004))
                         learningRate = 0.00004;
 
+                // With normalization
+                System.out.println("DATASET: "+dataset+", LR: "+learningRate+", ITERATIONS: "+iterations+", NORMALISED");
+                FoldRunHandler ltrWrapper2 = new FoldRunHandler(handler, dataset, folds, iterations, k, true, learningRate);
+                Measurement measurement2 = ltrWrapper2.averageScore();
+                String[] values2 = {dataset.name(), ""+iterations, "normalized", ""+measurement2.getEvaluationResult()};
+                writer.writeNext(values2);
+
                 // Without normalization
                 System.out.println("DATASET: "+dataset+", LR: "+learningRate+", ITERATIONS: "+iterations+", UNNORMALISED");
                 FoldRunHandler ltrWrapper = new FoldRunHandler(handler, dataset, folds, iterations, k, false, learningRate);
@@ -60,12 +64,6 @@ public class NormalizationExperimenter {
                 String[] values = {dataset.name(), ""+iterations, "unnormalized", ""+measurement.getEvaluationResult()};
                 writer.writeNext(values);
 
-                // With normalization
-                System.out.println("DATASET: "+dataset+", LR: "+learningRate+", ITERATIONS: "+iterations+", NORMALISED");
-                FoldRunHandler ltrWrapper2 = new FoldRunHandler(handler, dataset, folds, iterations, k, true, learningRate);
[…]

### 2014-11-19 report/Chapters/LTRMethods.tex
ltr methods mooier :)

@@
  	\While{true}{
- 		\emph{//projection step}\\
+ 		\emph{// projection step}\\
  		$w_t= \argmin_{w_t:w_t\in \mathbb{R}_{+}^{2d}} Q(w_t,z_t)$\\
@@
  	\If{$\frac{|F(w_t)-F(w_{t-1})|}{|F(w_{t-1})|} \le \epsilon_s$}{
- 		\emph{//early stopping criterion}\\
+ 		\emph{// early stopping criterion}\\
  		break
@@
 Variable structure learning automata can be defined as a triple $<\beta,\alpha,L>$, with $\beta$ being the set of inputs, $\alpha$ the set of actions.
-$L$, the learning algorithm ,is a recurrence relation that modifies the action probability vector of the actions in $\alpha$.\\
+$L$, the learning algorithm, is a recurrence relation that modifies the action probability vector of the actions in $\alpha$.\\
 
-Three well-known learning algorithms $L$ used in variable structure learning automata are \emph{linear reward-penalty} ($L_{R-P}$), \emph{linear reward-$\epsilon$-penalty} ($L_{R-\epsilon P}$) and \emph{linear reward-inaction} ($L_{R-I}$).
-LRUF uses the $L_{R-I}$ learning algorithm, which updates the action probability vector following Equation \ref{eq:lruf_reward_update} when the selected action $a_i(k)$ is rewarded by the environment.
+Three well-known learning algorithms for variable structure learning automata that can be used as $L$ are \emph{linear reward-penalty} ($L_{R-P}$), \emph{linear reward-$\epsilon$-penalty} ($L_{R-\epsilon P}$) and \emph{linear reward-inaction} ($L_{R-I}$).
+LRUF uses the $L_{R-I}$ learning algorithm, which updates the action probability vector following Equation \ref{eq:lruf_reward_update} at the moment that the selected action $a_i(k)$ is performed and rewarded by the environment.
 \begin{equation}
@@
 It has been shown that a combination of a variable action-set learning automaton with the $L_{R-I}$ learning algorithm is absolutely expedient and $\epsilon$-optimal \cite{Thathachar1987}, which means that it is guaranteed to approximate the optimal solution to some value $\epsilon$ and each update step is guaranteed not to decrease performance of the model.
-A variable-action set learning automata has a finite set of $r$ actions $\alpha={\alpha_1,...,\alpha_r}$.
+A variable-action set learning automaton has a finite set of $r$ actions $\alpha={\alpha_1,...,\alpha_r}$.
 $A$ is defined as the power set of $\alpha$, $A={A_1,..,A_m}$ with $m = 2^{r}-1$.
+$A(k)$ is the subset of all action that can be chosen by the learning automaton at each instant k.
 $\psi(k)$ is a probability distribution over $A$ such that $\psi(k)=p(A(k)=A_i|A_i\in A, 1\le i \le 2^r-1)$.
@@
 Initial training of the algorithm can be performed by using the training data relevance labels in the feedback set of the algorithm.
+$T_i$ in this algorithm is a dynamic relevance threshold initially set to zero.
+$n_i$ is the number of documents to rank using $R_i$.
 

### 2014-11-20 report/Chapters/Conclusions.tex
kleine fixes thesis

@@
 
-To answer this research question, we took the approach of implementing the list of algorithms found in answering the first research question, starting with the Learning to Rank method with the highest certainty on ranking accuracy, ListNet.\\
+To answer this research question, we implemented the ListNet algorithm that was found to have the highest certainty in ranking accuracy.\\
 

### 2014-11-20 report/Chapters/Future.tex
kleine fixes thesis

@@
 Follow-up research from this research can be categorised into three categories. These categories of potential future work are described in the sections below.
-
-\section{Distributed Computing Models}
-We explored the possibilities of Hadoop MapReduce for distributed computation of Learning to Rank training algorithms.
-Hadoop, since the introduction of Hadoop YARN in Hadoop 2.0, offers integration with other distributed computing models, including Dryad \cite{Isard2007}, Spark \cite{Zaharia2010}, Storm \cite{Aniello2013} and \ac{MPI}.
-Of these distributed programming models, Spark is particularly promising, since Shukla et al.
-\cite{Shukla2012} already showed that it is able to speed-up the ListNet training procedure with much lower job scheduling overhead than that we found for MapReduce.
-For the newly supported programming models in Hadoop, it holds that they lack the critical mass as the distributed programming model of choice.
-This lack of a critical mass results in higher integration costs and less available support.
-The integration of new programming model into the Hadoop YARN framework is a step in the good direction to alleviate the integration cost of these programming models, but does not yet completely eliminate the problem.
-Cloud-based Hadoop services like Microsoft HDInsight and Amazon Elastic MapReduce.
 
@@
 In this thesis we explored the speed-up characteristics of ListNet on Hadoop MapReduce.
-The speed-up characteristics of SmoothRank, FenchelRank, FSMRank and LRUF are using the MapReduce computing model are still to be explored.
+The speed-up characteristics of SmoothRank, FenchelRank, FSMRank and LRUF when implemented using the MapReduce computing model are still to be explored.
 
@@
 One should therefore also explore the ranking accuracy characteristics when an existing Learning to Rank method is evaluated in combination with an optimisation algorithm that is not prescribed to be a part of the Learning to Rank method.
+
+\section{Distributed Computing Models}
+We explored the possibilities of Hadoop MapReduce for distributed computation of Learning to Rank training algorithms.
+Hadoop, since the introduction of Hadoop YARN in Hadoop 2.0, offers integration with other distributed computing models, including but not limited to Dryad \cite{Isard2007}, Spark \cite{Zaharia2010}, Storm \cite{Aniello2013} and \ac{MPI}.
+Of these distributed programming models, Spark is particularly promising, since Shukla et al.
+\cite{Shukla2012} already showed that it is able to speed-up the ListNet training procedure with much lower job scheduling overhead than that we found for MapReduce.
+For the newly supported programming models in Hadoop, it holds that they lack the critical mass as the distributed programming model of choice.
+This lack of a critical mass results in higher integration costs and less available support.
+The integration of new programming model into the Hadoop YARN framework is a step in the good direction to alleviate the integration cost of these programming models, but does not yet completely eliminate the problem.
+The rise of cloud-based Hadoop services like Microsoft HDInsight and Amazon Elastic MapReduce and their increasing support for programming models other than MapReduce further alleviates these integration cost, as it takes away the need for high expenses on hardware.

### 2014-11-20 report/Chapters/Results.tex
laatste punten Dolf :):):)
kleine fixes thesis

@@
 The horizontal positions of the measurements are identical between execution method, as they are originate from the data sizes of the data sets used.
-Measurements were performed on a set of data sets consisting of the LETOR 3.0 OHSUMED data set, LETOR 4.0 data sets and the MSLR-WEB10/30K data sets are used, supplemented with generated data sets that are multiplications of MSLR-WEB30K (as described in section \ref{ssec:rq2_methodology}).
+Measurements were performed on a set of data sets consisting of the LETOR 3.0 OHSUMED data set, LETOR 4.0 data sets and the MSLR-WEB10/30K data sets are used, supplemented with generated data sets that are duplications of MSLR-WEB30K (as described in section \ref{ssec:rq2_methodology}).
 Table \ref{tbl:recap_datasets} describes the data sets used in the experiments and their single training fold data sizes.\\
@@
 
+The measurements of the clusters of 8 data nodes or smaller in Figure \ref{fig:listnet_train_time} have high variance, this can be attributed to the fact that these clusters were mostly tested on very small datasets, leaving the measurement very dependent on the variance of the job scheduling overhead.
+When we look at the more stable measurements, the ones on the clusters of 24 data nodes or more, we see that the processing time of a ListNet training iteration increases either linearly of very slightly sub-linearly as a function of the data size.\\
+
 \begin{figure}
@@
 Our observation that RankLib ListNet processes very slow for data sets that do not fit in physical memory and virtual memory is needed is very notable in this graph.
-Furthermore, this graph shows how both increase in number of data nodes in a cluster and increase in input data size result in an increase in processing speed.
-The sub-linear growth of processing speed as a factor of input data size can again be explained as a result of the non-parallelisable Hadoop job scheduling part of the operation.\\
-
-A result of that the Hadoop job scheduling overhead does not grow when the input data grows, is that the speed-up is linear or potentially better than linear as a function of data size.
-This can be seen in Figure \ref{fig:listnet_train_time}.
-In this Figure the cluster run lines are linear or, in case of the 24 and 32 data nodes cluster lines, slightly better than linear.
-The lines of the clusters of 8 data nodes or smaller have high variance, this can be attributed to the fact that these clusters were mostly tested on very small datasets, leaving the measurement very dependent on the variance of the job scheduling overhead.
+This graph shows how both an increase in number of data nodes in a cluster and an increase in input data size result in an increase in processing speed.
+The super-linear increase of processing speed as a function of data size that can be seen in the 2, 4 and 8 data-node-cluster-lines can be explained as a result of the non-parallelisable Hadoop job scheduling part of the operation.
+For the lines originating from the 24, 32 and 64 data node clusters, we see a slowing growth of the processing speed as a function of data size.
+Note that the processing speed of the clusters is bounded by their CPU processing speed and will therefore not increase indefinitely as a function of data size.\\
 
@@
 \end{figure}
-
-\section{SmoothRank}

### 2014-11-20 implementation/RankLib/src/ciir/umass/edu/experimenter/NormalizationExperimenter.java

@@ -27,8 +27,8 @@ public class NormalizationExperimenter {
 
         Set<DataSets.DataSet>       datasets            = new HashSet<DataSets.DataSet>();
         // Add all LETOR 3.0 data sets
-        //datasets.add(DataSets.DataSet.MSLR_WEB30K);
-        datasets.add(DataSets.DataSet.MSLR_WEB10K);
+        datasets.add(DataSets.DataSet.MSLR_WEB30K);
+        //datasets.add(DataSets.DataSet.MSLR_WEB10K);
 
         BufferedWriter out = null;
         try {

### 2014-11-20 report/Chapters/Background.tex
based -> -based
listings -> algorithm

@@
 \begin{description}
-\item[Feature-Based]{means that all objects under investigation are represented by feature vectors.
+\item[Feature-based]{means that all objects under investigation are represented by feature vectors.
 In a Learning to Rank for Information Retrieval case, this means that the feature vectors can be used to predict the relevance of the documents to the query, or the importance of the document itself.}
@@
 
-The algorithm to compute \ac{ERR} is shown in Listing \ref{alg:err}.
+The algorithm to compute \ac{ERR} is shown in Algorithm \ref{alg:err}.
 The algorithm requires relevance grades $g_i$, $1 \le i \le n$ and mapping function $R$ that maps relevance grades to probability of relevance.\\

### 2014-11-20 report/Chapters/Implementation.tex
laatste punten Dolf :):):)
listings -> algorithm

@@
 The preprocessing phase consists of two separate Pig jobs.
-The first Pig job determines the minimum and the maximum values per feature in the training set.
-The second Pig job rescales each feature of the train, validation and test data sets using the following formula for rescaling:
+The first Pig job (Algorithm \ref{alg:pig_normalisation_1}) determines the minimum and the maximum values per feature in the training set.
+The second Pig job (Algorithm \ref{alg:pig_normalisation_2}) rescales each feature of the train, validation and test data sets using the following formula for rescaling:
 \begin{equation}
@@
 
-The first Pig job:\\
-\begin{minipage}{\linewidth}
-\begin{lstlisting}
-REGISTER [path prefix]/lib/*.jar;
-TRAIN = LOAD '[path prefix]/input/[data set name]/Fold[fold number]/train.txt' USING PigStorage(' ');
-TRAIN_STD = FOREACH TRAIN GENERATE flatten(udf.util.ToStandardForm(*));
-TRAIN_STD_BY_QUERY = GROUP TRAIN_STD BY $1 PARALLEL [available reducers];
-MIN_MAX = FOREACH TRAIN_STD_BY_QUERY GENERATE flatten(udf.util.GetMinMax(*));
-MIN_MAX_GRPD = GROUP MIN_MAX ALL;
-MIN_MAX_FIN = FOREACH MIN_MAX_GRPD GENERATE flatten(udf.util.CombineMinMax(*));
-STORE MIN_MAX_FIN INTO 'minmax[fold number]';
-\end{lstlisting}
-\end{minipage}\\
+\begin{algorithm}
+REGISTER [path prefix]/lib/*.jar;\\
+TRAIN = LOAD '[path prefix]/input/[data set name]/Fold[fold number]/train.txt' USING PigStorage(' ');\\
+TRAIN\_STD = FOREACH TRAIN GENERATE flatten(udf.util.ToStandardForm(*));\\
+TRAIN\_STD\_BY\_QUERY = GROUP TRAIN\_STD BY \$1 PARALLEL [available reducers];\\
+MIN\_MAX = FOREACH TRAIN\_STD\_BY\_QUERY GENERATE flatten(udf.util.GetMinMax(*));\\
+MIN\_MAX\_GRPD = GROUP MIN\_MAX ALL;\\
+MIN\_MAX\_FIN = FOREACH MIN\_MAX\_GRPD GENERATE flatten(udf.util.CombineMinMax(*));\\
+STORE MIN\_MAX\_FIN INTO 'minmax[fold number]';\\
+\caption{The first Pig job of the normalisation preprocessing procedure}
+\label{alg:pig_normalisation_1}
+\end{algorithm}
 
@@
 
-The second Pig job:\\
-\begin{minipage}{\linewidth}
-\begin{lstlisting}
-REGISTER [path prefix]/lib/*.jar;
-TRAIN = LOAD '[path prefix]/input/[data set name]/Fold[fold number]/train.txt' USING PigStorage(' ');
-VALIDATE = LOAD '[path prefix]/input/[data set name]/Fold[fold number]/vali.txt' USING PigStorage(' ');
-TEST = LOAD '[path prefix]/input/[data set name]/Fold[fold number]/test.txt' USING PigStorage(' ');
-TRAIN_STD = FOREACH TRAIN GENERATE flatten(udf.util.ToStandardForm(*));
-VALIDATE_STD = FOREACH VALIDATE GENERATE flatten(udf.util.ToStandardForm(*));
-TEST_STD = FOREACH TEST GENERATE flatten(udf.util.ToStandardForm(*));
-DEFINE ScaleFeatures udf.util.ScaleFeatures('[array with minimum and maximum feature values]');
-TRAIN_SCA = FOREACH TRAIN_STD GENERATE flatten(ScaleFeatures(*));
-VALIDATE_SCA = FOREACH VALIDATE_STD GENERATE flatten(ScaleFeatures(*));
-TEST_SCA = FOREACH TEST_STD GENERATE flatten(ScaleFeatures(*));
-STORE TRAIN_SCA INTO 'train_sca[fold number]' USING BinStorage();
-STORE VALIDATE_SCA INTO 'validate_sca[fold number]' USING BinStorage();
-STORE TEST_SCA INTO 'test_sca[fold number]' USING BinStorage();
-\end{lstlisting}
-\end{minipage}\\
+\begin{algorithm}
+REGISTER [path prefix]/lib/*.jar;\\
+TRAIN = LOAD '[path prefix]/input/[data set name]/Fold[fold number]/train.txt' USING PigStorage(' ');\\
+VALIDATE = LOAD '[path prefix]/input/[data set name]/Fold[fold number]/vali.txt' USING PigStorage(' ');\\
+TEST = LOAD '[path prefix]/input/[data set name]/Fold[fold number]/test.txt' USING PigStorage(' ');\\
+TRAIN\_STD = FOREACH TRAIN GENERATE flatten(udf.util.ToStandardForm(*));\\
+VALIDATE\_STD = FOREACH VALIDATE GENERATE flatten(udf.util.ToStandardForm(*));\\
+TEST\_STD = FOREACH TEST GENERATE flatten(udf.util.ToStandardForm(*));\\
+DEFINE ScaleFeatures udf.util.ScaleFeatures('[array with minimum and maximum feature values]');\\
+TRAIN\_SCA = FOREACH TRAIN\_STD GENERATE flatten(ScaleFeatures(*));\\
+VALIDATE\_SCA = FOREACH VALIDATE\_STD GENERATE flatten(ScaleFeatures(*));\\
+TEST\_SCA = FOREACH TEST\_STD GENERATE flatten(ScaleFeatures(*));\\
+STORE TRAIN\_SCA INTO 'train\_sca[fold number]' USING BinStorage();\\
+STORE VALIDATE\_SCA INTO 'validate\_sca[fold number]' USING BinStorage();\\
+STORE TEST\_SCA INTO 'test\_sca[fold number]' USING BinStorage();\\
+\caption{The second Pig job of the normalisation preprocessing procedure}
+\label{alg:pig_normalisation_2}
+\end{algorithm}
 
@@
 The training stage, like the preprocessing stage, consists of two separate Pig jobs.
-The first Pig job calculates the Cross Entropy loss on training data of the current model and calculates the gradients for the next model update.
-The second Pig job is an internal validation step that validates the model on the validation set by calculating \ac{NDCG}@k.\\
+The first Pig job, displayed in Algorithm \ref{alg:pig_listnet_1}, calculates the Cross Entropy loss on training data of the current model and calculates the gradients for the next model update.
+The second Pig job (Algorithm \ref{alg:pig_listnet_2}) is an internal validation step that validates the model on the validation set by calculating \ac{NDCG}@k.\\
 
-The first Pig job:\\
-\begin{minipage}{\linewidth}
-\begin{lstlisting}
-REGISTER [path prefix]/lib/*.jar;
-DEFINE QueryLossGradient udf.listnet.QueryLossGradient('[feature dimensionality of data set]');
-DEFINE ExpRelOurScores udf.listnet.ExpRelOurScores('[neural network weights & iteration number]');
-[FIRST TRAINING ITERATION:]
-	TRAIN_SCA = LOAD 'train_sca[fold number]/*' USING BinStorage();
-	TR_BY_QUERY = GROUP TRAIN_SCA BY $1 PARALLEL [number of avaiable reducers];
-	TR_EXP_REL_SCORES = FOREACH TR_BY_QUERY GENERATE flatten(ExpRelOurScores(TRAIN_SCA));
-	STORE TR_EXP_REL_SCORES INTO 'tr_exp_rel_scores-f[fold number]' USING BinStorage();
-[SUBSEQUENT TRAINING ITERATIONS:]
-	TR_EXP_REL_SCORES = LOAD 'tr_exp_rel_scores-f[fold number]/*' USING BinStorage();
-	TR_EXP_REL_SCORES = FOREACH TR_EXP_REL_SCORES GENERATE flatten(ExpRelOurScores(*)) PARALLEL [number of available reducers];
-TR_QUERY_LOSS_GRADIENT = FOREACH TR_EXP_REL_SCORES GENERATE flatten(QueryLossGradient(*)) PARALLEL [number of available reducers];
-TR_QUERY_LOSS_GRADIENT_GRPD = GROUP TR_QUERY_LOSS_GRADIENT ALL;
-TR_LOSS_GRADIENT = FOREACH TR_QUERY_LOSS_GRADIENT_GRPD GENERATE flatten(udf.listnet.MultiSum(*));
-STORE TR_LOSS_GRADIENT INTO 'tr_loss_gradient-f[fold number]i[iteration number]';
-\end{lstlisting}
-\end{minipage}\\
+\begin{algorithm}
+REGISTER [path prefix]/lib/*.jar;\\
+DEFINE QueryLossGradient udf.listnet.QueryLossGradient('[feature dimensionality of data set]');\\
+DEFINE ExpRelOurScores udf.listnet.ExpRelOurScores('[neural network weights \& iteration number]');\\
+$[$FIRST TRAINING ITERATION:$]$\\
+\Indp
+	TRAIN\_SCA = LOAD 'train\_sca[fold number]/*' USING BinStorage();\\
+	TR\_BY\_QUERY = GROUP TRAIN\_SCA BY \$1 PARALLEL [number of avaiable reducers];\\
+	TR\_EXP\_REL\_SCORES = FOREACH TR\_BY\_QUERY GENERATE flatten(ExpRelOurScores(TRAIN\_SCA));\\
+	STORE TR\_EXP\_REL\_SCORES INTO 'tr\_exp\_rel\_scores-f[fold number]' USING BinStorage();\\
+\Indm 
+$[$SUBSEQUENT TRAINING ITERATIONS:$]$\\
+\Indp
+TR\_EXP\_REL\_SCORES = LOAD 'tr\_exp\_rel\_scores-f[fold number]/*' USING BinStorage();\\
+TR\_EXP\_REL\_SCORES = FOREACH TR\_EXP\_REL\_SCORES GENERATE flatten(ExpRelOurScores(*)) PARALLEL [number of available reducers];\\
+\Indm
+TR\_QUERY\_LOSS\_GRADIENT = FOREACH TR\_EXP\_REL\_SCORES GENERATE flatten(QueryLossGradient(*)) PARALLEL [number of available reducers];\\
+TR\_QUERY\_LOSS\_GRADIENT\_GRPD = GROUP TR\_QUERY\_LOSS\_GRADIENT ALL;\\
+TR\_LOSS\_GRADIENT = FOREACH TR\_QUERY\_LOSS\_GRADIENT\_GRPD GENERATE flatten(udf.listnet.MultiSum(*));\\
+STORE TR\_LOSS\_GRADIENT INTO 'tr\_loss\_gradient-f[fold number]i[iteration number]';\\
+\caption{The second Pig job of the normalisation preprocessing procedure}
+\label{alg:pig_listnet_1}
+\end{algorithm}
 
@@
 
-The second Pig job of the training stage validates the performance of the model weights that were trained in the first Pig job on the validation data set.\\
+The second Pig job of the training stage, shown in Algorithm \ref{alg:pig_listnet_2}, validates the performance of the model weights that were trained in the first Pig job on the validation data set.\\
 
-The second Pig job:\\
-\begin{minipage}{\linewidth}
-\begin{lstlisting}
-REGISTER [path prefix]/lib/*.jar;
-DEFINE Ndcg udf.util.Ndcg('[neural network weights & NDCG cut-off parameter]');
-[FIRST TRAINING ITERATION:]
-	VALIDATE_SCA = LOAD 'validate_sca[fold number]/*' USING BinStorage();
-	VA_BY_QUERY = GROUP VALIDATE_SCA BY $1 PARALLEL [number of available reducers];
-	STORE VA_BY_QUERY INTO 'va_by_query-f[fold number]' USING BinStorage();
-[SUBSEQUENT TRAINING ITERATIONS:]
-	VA_BY_QUERY = LOAD 'va_by_query-f[fold number]/*' USING BinStorage();
-NDCG = FOREACH VA_BY_QUERY GENERATE Ndcg(*);
-NDCG_GRPD = GROUP NDCG ALL;
-AVG_NDCG = FOREACH NDCG_GRPD GENERATE AVG(NDCG);
-STORE AVG_NDCG INTO 'avg_ndcg-f[fold number]i[iteration number]';
-\end{lstlisting}
-\end{minipage}\\
+\begin{algorithm}
+REGISTER [path prefix]/lib/*.jar;\\
+DEFINE Ndcg udf.util.Ndcg('[neural network weights \& NDCG cut-off parameter]');\\
+$[$FIRST TRAINING ITERATION:$]$\\
+\Indp
+	VALIDATE\_SCA = LOAD 'validate\_sca[fold number]/*' USING BinStorage();\\
+	VA\_BY\_QUERY = GROUP VALIDATE\_SCA BY \$1 PARALLEL [number of available reducers];\\
+	STORE VA\_BY\_QUERY INTO 'va\_by\_query-f[fold number]' USING BinStorage();\\
+\Indm 
+$[$SUBSEQUENT TRAINING ITERATIONS:$]$\\
+\Indp
+VA\_BY\_QUERY = LOAD 'va\_by\_query-f[fold number]/*' USING BinStorage();\\
+\Indm 
+NDCG = FOREACH VA\_BY\_QUERY GENERATE Ndcg(*);\\
+NDCG\_GRPD = GROUP NDCG ALL;\\
+AVG\_NDCG = FOREACH NDCG\_GRPD GENERATE AVG(NDCG);\\
+STORE AVG\_NDCG INTO 'avg\_ndcg-f[fold number]i[iteration number]';\\
+\caption{The second Pig job of the normalisation preprocessing procedure}
+\label{alg:pig_listnet_2}
+\end{algorithm}
 
@@
 \end{minipage}\\
-
-\section{SmoothRank}
-\subsection{Preprocessing}
-\subsection{Training}
-\subsection{Testing}

### 2014-11-20 report/Chapters/Implementation.tex
algorithms :)

@@
 STORE TR\_LOSS\_GRADIENT INTO 'tr\_loss\_gradient-f[fold number]i[iteration number]';\\
-\caption{The second Pig job of the normalisation preprocessing procedure}
+\caption{The first Pig job of the ListNet training procedure}
 \label{alg:pig_listnet_1}
@@
 STORE AVG\_NDCG INTO 'avg\_ndcg-f[fold number]i[iteration number]';\\
-\caption{The second Pig job of the normalisation preprocessing procedure}
+\caption{The second Pig job of the ListNet training procedure}
 \label{alg:pig_listnet_2}
@@
 \subsection{Testing}
-The testing stage tests the best model found in the training iterations, selected on validation set \ac{NDCG}@k (as calculated in the second Pig job of the training stage), by calculating the \ac{NDCG}@k of this model on the test set.\\
+The testing stage tests the best model found in the training iterations, selected on validation set \ac{NDCG}@k (as calculated in the second Pig job of the training stage), by calculating the \ac{NDCG}@k of this model on the test set.
+Algorithm \ref{alg:pig_test} describes the Pig code for model evaluation on the test data set.\\
 
-Test stage Pig job:\\
-\begin{minipage}{\linewidth}
-\begin{lstlisting}
-REGISTER [path prefix]/lib/*.jar;
-TEST_SCA = LOAD 'test_sca[fold number]/*' USING BinStorage();
-TE_BY_QUERY = GROUP TEST_SCA BY $1 PARALLEL [number of available reducers];
-DEFINE Ndcg udf.util.Ndcg('[neural network weights & NDCG cut-off parameter]');
-NDCG = FOREACH TE_BY_QUERY GENERATE Ndcg(*);
-NDCG_GRPD = GROUP NDCG ALL;
-AVG_NDCG = FOREACH NDCG_GRPD GENERATE AVG(NDCG);
-STORE AVG_NDCG INTO 'avg_ndcg';
-\end{lstlisting}
-\end{minipage}\\
+\begin{algorithm}
+REGISTER [path prefix]/lib/*.jar;\\
+TEST\_SCA = LOAD 'test\_sca[fold number]/*' USING BinStorage();\\
+TE\_BY\_QUERY = GROUP TEST\_SCA BY \$1 PARALLEL [number of available reducers];\\
+DEFINE Ndcg udf.util.Ndcg('[neural network weights \& NDCG cut-off parameter]');\\
+NDCG = FOREACH TE\_BY\_QUERY GENERATE Ndcg(*);\\
+NDCG\_GRPD = GROUP NDCG ALL;\\
+AVG\_NDCG = FOREACH NDCG\_GRPD GENERATE AVG(NDCG);\\
+STORE AVG\_NDCG INTO 'avg\_ndcg';
+\caption{The Pig job for model evaluation}
+\label{alg:pig_test}
+\end{algorithm}

### 2014-11-20 report/Chapters/BenchmarkResults.tex
lessen van Dolf :)

@@
 \cite{Pavlov2010} in their paper claim to have won the Yandex Internet Mathematics contest 2009 with their BagBoo method.
+
+\section{Conclusions}
+\label{chap:cross_comparison}
+We have seen in the preceding sections that the evaluation of Learning to Rank methods is spread over several benchmark data sets.
+However, as the Learning to Rank methods evaluated differs between benchmarks, no single benchmark comparison can be regarded as a conclusive argument on which Learning to Rank method is most accurate.
+In the next chapter we will describe a meta-analysis based on a structured literature research that compares the ranking accuracy of Learning to Rank methods over multiple benchmark sets.\\

### 2014-11-20 report/Chapters/CrossBenchmarkComparison.tex
lessen van Dolf :)

@@
-\label{chap:cross_comparison}
-As we have seen in Chapter~\ref{chap:benchmark_results}, the evaluation of Learning to Rank methods is spread over several benchmark data sets.
-However, as the Learning to Rank methods evaluated differs between benchmarks, no single benchmark comparison can be regarded as a conclusive argument on which Learning to Rank method is most accurate.\\
-
 Several studies make a small start in considering Learning to Rank methods performance over multiple benchmark data sets.

### 2014-11-20 report/Chapters/RelatedWork.tex
lessen van Dolf :)

@@
 Learning to Rank Challenge.\\
+
+\section{Conclusions}
+We have seen that a wide range of techniques for parallel and distributed computing are used in the field of parallel or distributed Learning to Rank.
+Only a small fraction of related work focusses on the MapReduce programming model, while MapReduce is the data processing framework of choice in industry \cite{Lin2013}.
+Furthermore we see that most related work that does involve MapReduce focusses either on distributed hyper-parameter optimisation \cite{Ganjisaffar2011} or on model optimisation techniques \cite{Long2012,Zinkevich2010}, while only one study describes a MapReduce implementation of a Learning to Rank algorithm itself \cite{Karimzadehgan2011}.
+This thesis provides an exploration of using MapReduce for parallel computation of Learning to Rank model training.

### 2014-11-20 report/Chapters/Background.tex
rechtvaardiging twee versies NDCG

@@
 \subsection{Normalized Discounted Cumulative Gain}
+\label{sec:ndcg}
 Cumulative gain and its successors discounted cumulative gain and normalized discounted cumulative gain are arguably the most widely used measures for effectiveness of ranking methods.

### 2014-11-20 report/Chapters/CrossBenchmarkComparison.tex
rechtvaardiging twee versies NDCG

@@
 By making the decision to include \ac{NDCG} at three cut-off points and only a single \ac{MAP} entry, we implicitly attain a higher weight for \ac{NDCG} compared to \ac{MAP} on an analysis that combines all measurements on the four metrics.
-This implicit weighting is arbitrary and therefore undesirable, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.
+This implicit weighting is arbitrary and therefore undesirable, but the number of algorithm evaluation results gained by this makes it a pragmatic approach.\\
+
+We described in section \ref{sec:ndcg} that there are two definitions of the \ac{NDCG}-metric that vary slightly: one defined by J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} and the other defined by Burges et al.
+\cite{Burges2005}.
+The official benchmarks evaluation runs of the different benchmark sets differ in the \ac{NDCG}-definition that is used.
+For example, on the official LETOR evaluation runs the definition by J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} is used, where the official evaluation runs on WCL2R use the definition by Burges et al.
+\cite{Burges2005}.
+For fair comparison, it is of importance to only take into account evaluation results that report \ac{NDCG} scores using the same \ac{NDCG}-definition as the official evaluation runs of the benchmark, therefore we filter out evaluation results that do not fulfil this property.
+Note that this approach results in the \ac{IWN} and \ac{NWN} calculations taking into account evaluation results in both definitions of \ac{NDCG} (although only one definition is used per data set).
+This is not a problem, since both definitions of \ac{NDCG} are highly correlated and vary only slightly.
 
@@
 \item The study reported evaluation results in a different metric than the metrics chosen for this meta-analysis \cite{Yu2009, Thuy2009, Pahikkala2009, Kersting2009, Mohan2011}.
-\item The study reported a higher performance on baseline methods than official benchmark runs \cite{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}.
+\item The study reported a higher performance on baseline methods than the official benchmark results \cite{Dubey2009, Banerjee2009, Peng2010b, Song2014, Bian2010, Bian2010b, Carvalho2008, Acharyya2012, Peng2010b, Tran2012, Asadi2013c}.
 Such cases are not necessarily caused by malicious intent, but might be caused by a mix-up of the two existing versions of the \ac{NDCG} metric (J{\"a}rvelin and Kek{\"a}l{\"a}inen \cite{Jarvelin2002} and Burges et al.
 \cite{Burges2005}).
-\item The study did not report any baseline performance that allowed us to check validity of the results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
+\item The study did not report any results on the baseline methods to compare to the official benchmark results \cite{Chakrabarti2008, Wang2012b, Buffoni2011}.
 \end{enumerate}

### 2014-11-20 report/Chapters/Results.tex
mooiere uitlijning figuren

@@
 \centering
-\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.8]{gfx/time_single_logx.pdf}
+\includegraphics[trim=3cm 5cm 0cm 5cm, scale=0.8]{gfx/time_single_logx.pdf}
 \caption{Processing time of a single ListNet training iteration on a logarithmic data size axis}
@@
 \centering
-\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.7]{gfx/speedup_faceted.pdf}
+\includegraphics[trim=5cm 5cm 0cm 5cm, scale=0.7]{gfx/speedup_faceted.pdf}
 \caption{Processing time of a single ListNet training iteration as a function of the number of data nodes in a cluster}
@@
 \includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.8]{gfx/processing_speed_single_logx.pdf}
-\caption{Processing speed of a a single ListNet training iteration on various data sets}
+\caption{Processing speed of a single ListNet training iteration on various data sets}
 \label{fig:listnet_processing_speed}

### 2014-11-21 report/Chapters/Conclusions.tex
laatste werk

@@
 
-Moreover, we found that the addition of a normalisation preprocessing step to the ListNet procedure can greatly improve the ranking accuracy of the ListNet training procedure after the first five iterations.
-This suggests that the addition of a normalisation preprocessing step can reduce the number of training iterations needed for convergence.
+Moreover, we found that the addition of a normalisation preprocessing step to a data set that is not yet normalised can greatly improve the convergence of the ranking accuracy of the ListNet training procedure.
 Lin stated in his essay \cite{Lin2013} that MapReduce is often good enough for tasks that are not-amenable to the MapReduce model.
@@
 
-Most importantly, we found the training time of our cluster version of ListNet to grow better than linearly in terms of data size increase.
+Most importantly, we found the training time of our cluster version of ListNet to grow linearly in terms of data size increase.
 This shows that the cluster implementation of ListNet can be used to scale the ListNet training procedure to arbitrarily large data sets, given that enough data nodes are available for computation.\\

### 2014-11-21 report/Chapters/Results.tex
laatste werk

@@
 The official evaluation results of the LETOR benchmarks do not report the standard deviation of their experimental results and it is not mentioned in the LETOR 3.0 and 4.0 papers \cite{Qin2010, Qin2013} how many times their experiments were repeated.
-For the ListNet step size hyperparameter we choose a value of $0.0001$, which is the default value step size value in RankLib, for both the cluster and the RankLib runs.
+For the ListNet step size hyperparameter we choose a value of $1*10^{-4}$, which is the default value step size value in RankLib, for both the cluster and the RankLib runs.
 It is not clear what step size was used in the official evaluation runs.
@@
 It seems to be the case that the normalisation procedure enables ListNet to converge faster.
-It has been shown in literature, that the gradient descent optimisation procedure, which is also being used in ListNet, converges faster on normalised data \cite{Ng1999}.
+It has been shown in literature that some neural network training algorithms converge faster on normalised data \cite{Ng1999}, however, this has not been researched in a Learning to Rank setting.\\
 
@@
 \end{table}
+
+Based on our own implementation of the normalisation preprocessing step that we added to the RankLib code, we further explored the effects of normalisation on the convergence rate of ListNet.
+We conducted experiments on the LETOR 3.0 datasets HP2003, NP2003 and TD2003 and on the MSLR-WEB10k data set.
+The LETOR 3.0 datasets are in published form already normalised on query-level, meaning that the $max(x)$ and $min(x)$ terms in the normalisation formula $x^{'} = \frac{x-min(x)}{max(x)-min(x)}$ are calculated per query instead of on the whole data set.
+The MSLR-WEB10k data set is not published in a normalised way.
+Looking at Figures \ref{fig:convergence_hp2003}, \ref{fig:convergence_np2003} and \ref{fig:convergence_td2003} that represent the convergence on HP2003, NP2003 and TD2003 respectively, we see that there is no difference in convergence properties between ListNet applied on globally normalised and query-level normalised data.
+Figure \ref{fig:convergence_mslrweb10k} shows that the ListNet convergence increases when normalisation preprocessing is applied to the unnormalised data set MSLR-WEB10k.
+
+\begin{figure}
+\centering
+\includegraphics[trim=5cm 5cm 5cm 5cm, scale=0.85]{gfx/convergence_hp2003.pdf}
+\caption{Convergence of ListNet on query-level and globally normalised versions of HP2003}
+\label{fig:convergence_hp2003}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[trim=5cm 5cm 5cm 5cm, scale=0.85]{gfx/convergence_np2003.pdf}
+\caption{Convergence of ListNet on query-level and globally normalised versions of NP2003}
+\label{fig:convergence_np2003}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[trim=5cm 5cm 5cm 5cm, scale=0.85]{gfx/convergence_td2003.pdf}
+\caption{Convergence of ListNet on query-level and globally normalised versions of TD2003}
+\label{fig:convergence_td2003}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[trim=5cm 5cm 5cm 5cm, scale=0.85]{gfx/convergence_mslr_web10k.pdf}
+\caption{Convergence of ListNet on normalised and unnormalised versions of MSLR-WEB10k}
+\label{fig:convergence_mslrweb10k}
+\end{figure}
 
@@
 \centering
-\includegraphics[trim=5cm 5cm 0cm 5cm, scale=0.7]{gfx/speedup_faceted.pdf}
+\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.7]{gfx/speedup_faceted.pdf}
 \caption{Processing time of a single ListNet training iteration as a function of the number of data nodes in a cluster}
@@
 \centering
-\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.8]{gfx/processing_speed_single_logx.pdf}
+\includegraphics[trim=5cm 5cm 0cm 5cm, scale=0.8]{gfx/processing_speed_single_logx.pdf}
 \caption{Processing speed of a single ListNet training iteration on various data sets}

### 2014-11-21 measurements/convergence_letor.R

@@ -1,0 +1,22 @@
+setwd("C:/Git-data/Afstuderen/implementation/RankLib/")
+library(ggplot2)
+library(mgcv)
+library(stringr)
+library(scales)
+measure = read.csv("normalization_measurements_LETOR3.csv") # read csv file
+measure = measure[measure$dataset=='HP2003',]
+
+d <- ggplot(data=measure, aes(x=iters, y=ndcg, colour=norm, shape=norm))
+d <- d + geom_point(size=2) + 
+	theme_bw() + 
+	theme(
+		panel.grid.major = element_blank(),
+		panel.grid.minor = element_blank(),
+		panel.border     = element_blank(),
+		panel.background = element_blank(),
+		axis.line = element_line(color='black')
+	) +
+	xlab("# of iterations") +
+	ylab("NDCG@10 on test set") +
+	labs(colour = "Execution mode", shape = "Execution mode") 
+ggplot_build(d)

### 2014-11-21 measurements/convergence_mslr10k.R

@@ -1,0 +1,21 @@
+setwd("C:/Git-data/Afstuderen/implementation/RankLib/")
+library(ggplot2)
+library(mgcv)
+library(stringr)
+library(scales)
+measure = read.csv("normalization_measurements_mslr_web_10k.csv") # read csv file
+
+d <- ggplot(data=measure, aes(x=iters, y=ndcg, colour=norm, shape=norm))
+d <- d + geom_point(size=2) + 
+	theme_bw() + 
+	theme(
+		panel.grid.major = element_blank(),
+		panel.grid.minor = element_blank(),
+		panel.border     = element_blank(),
+		panel.background = element_blank(),
+		axis.line = element_line(color='black')
+	) +
+	xlab("# of iterations") +
+	ylab("NDCG@10 on test set") +
+	labs(colour = "Execution mode", shape = "Execution mode") 
+ggplot_build(d)

### 2015-05-05 paper/IPM/paper.tex
nieuw werk

@@
 \title{A Cross-Benchmark Comparison of 87 Learning to Rank Methods}
-\author[avanade,ut]{Niek Tax\corref{cor}\fnref{fn}}
-\ead{niek.tax@avanade.com}
+\author[avanade,tue]{Niek Tax\corref{cor}\fnref{fn}}
+\ead{n.tax@tue.nl}
 \author[avanade]{Sander Bockting}
@@
 \ead{d.hiemstra@utwente.nl}
-\cortext[cor]{Corresponding Author: Niek Tax, Avanade Netherlands B.V., Versterkerstraat 6,
-1322AP Almere, the Netherlands; Email, niek.tax@avanade.com; Phone, +31634085760}
-\fntext[fn]{Author is affiliated with the University of Twente, but this paper was written during his stay at Avanade Netherlands B.V.}
+\cortext[cor]{Corresponding Author: Niek Tax, Eindhoven University of Technology, Department of Mathematics and Computer Science, P.O.
+Box 513, 5600MB Eindhoven, the Netherlands; Email, n.tax@tue.nl; Phone, +31634085760}
+\fntext[fn]{Author is affiliated with Eindhoven University of Technology, but this paper was written during his stay at Avanade Netherlands B.V.}
 \address[avanade]{Avanade Netherlands B.V., Versterkerstraat 6,
@@
 the Netherlands}
+\address[tue]{Eindhoven University of Technology, P.O. Box 513, 5600MB Eindhoven, the Netherlands}
 \begin{abstract}

### 2015-05-05 method_selection/bagplot_dataset_sensitivity.R

@@ -1,4 +1,5 @@
 library(aplpack)
+setwd("E:/Git/Afstuderen/method_selection/")
 listnet     = read.csv("listnet_dataset_sensitivity.csv")
 smoothrank  = read.csv("smoothrank_dataset_sensitivity.csv")
 fenchelrank = read.csv("fenchelrank_dataset_sensitivity.csv")

### 2015-05-05 method_selection/winning_number.R

@@ -101,13 +101,13 @@ combined_methods <- NULL
 combined_methods <- c(combined_methods, as.character(map_data[,1]))
 combined_methods <- c(combined_methods, as.character(ndcg3_data[,1]))
 combined_methods <- c(combined_methods, as.character(ndcg5_data[,1]))
-combined_methods <- c(combined_methods, as.character(ndcg10_data[,1]))
+#combined_methods <- c(combined_methods, as.character(ndcg10_data[,1]))
 
 combined_data <- NULL
 combined_data <- cbind(map_summed_winnum,map_summed_ideal_winnum)
 combined_data <- rbind(combined_data, cbind(ndcg3_summed_winnum, ndcg3_summed_ideal_winnum))
 combined_data <- rbind(combined_data, cbind(ndcg5_summed_winnum, ndcg5_summed_ideal_winnum))
-combined_data <- rbind(combined_data, cbind(ndcg10_summed_winnum, ndcg10_summed_ideal_winnum))
+#combined_data <- rbind(combined_data, cbind(ndcg10_summed_winnum, ndcg10_summed_ideal_winnum))
 
 combined <- cbind.data.frame(combined_methods,combined_data)
 colnames(combined) <- c("Method", "Summed_winnum", "Summed_ideal_winnum")

### 2015-05-05 method_selection/winning_number_sensitivity_dataset.R

@@ -2,7 +2,7 @@ setwd("E:/Git/Afstuderen/method_selection")
 require(ggplot2)
 require(plyr)
 
-lo <- 10
+lo <- 25
 lo <- lo+1
 
 ### MAP ###

### 2015-05-05 paper/IPM/paper.tex
Pareto optimality geintroduceerd

@@
 \usepackage{float}
+\usepackage{adjustbox}
 \renewcommand{\dblfloatpagefraction}{2.0}
@@
 Evaluation results of 87 learning to rank methods on 20 well-known benchmark datasets are collected through a structured literature search.
-ListNet, SmoothRank, FenchelRank, FSMRank, LRUF and LARF are the best performing learning to rank methods in increasing order of Normalized Winning Number and decreasing order of Ideal Winning Number.
+ListNet, SmoothRank, FenchelRank, FSMRank, LRUF and LARF are Pareto optimal learning to rank methods in the Normalized Winning Number and Ideal Winning Number dimensions, listed in increasing order of Normalized Winning Number and decreasing order of Ideal Winning Number.
 \end{abstract}
@@
 
-Research in the field of ranking models has been based on manually designed ranking functions, such as the well-known BM25 model \citep{Robertson1994}.
+Research in the field of ranking models has long been based on manually designed ranking functions, such as the well-known BM25 model \citep{Robertson1994}.
 Increased amounts of potential training data have recently made it possible to leverage machine learning methods to obtain more effective ranking models.
@@
 Table \ref{tbl:ltr_benchmark_collections} presents an overview of the benchmark collections included in the meta-analysis.
+Note that all these datasets offer feature set representations of the to-be-ranked documents instead of the documents themselves.
+Therefore, any difference in ranking performance is due to the ranking algorithm and not the features used.
 
@@
 The following enumeration enumerates those properties that made one or more studies unusable for the meta-analysis.
-Between brackets are the studies that these properties apply to.
+The references between brackets are the studies to which these properties apply.
 
@@
 \cite{Qin2010} state that it may differ between datasets what the most accurate ranking methods are.
-They propose a measure they call \emph{Winning Number} to evaluate the overall performance of learning to rank methods over the multiple datasets in the LETOR 3.0 collection.
+They propose a measure they call \emph{Winning Number} to evaluate the overall performance of learning to rank methods over the datasets included in the LETOR 3.0 collection.
 Winning Number is defined as the number of other algorithms that an algorithm can beat over the set of datasets, or more formally\\
@@
 The Winning Number evaluation metric relies on the denseness of the evaluation results set.
-In contrast to the LETOR 3.0 comparison, our set of evaluation results will be sparse.
+In contrast to the LETOR 3.0 comparison, our evaluation results will be a \emph{sparse} set.
 We propose a normalized version of the Winning Number metric to enable comparison of a sparse set of evaluation results.
-This Normalized Winning Number takes only those datasets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it it would have been the most accurate algorithm on all datasets on which it has been evaluated.
+This Normalized Winning Number takes only those datasets into account that an algorithm is evaluated on and divides this by the theoretically highest Winning Number that an algorithm would have had in case it would have been the most accurate algorithm on all datasets on which it has been evaluated.
 We will redefine the indicator function $I$ in order to only take into account those datasets that an algorithm is evaluated on, as\\
@@
 
-From now on this adjusted version of Winning Number will be references to as \emph{Normalized Winning Number (NWN)}. The mathematical definition of Normalized Winning Number is\\
+From now on this adjusted version of Winning Number will be references to as \emph{Normalized Winning Number (NWN)}. The formal definition of Normalized Winning Number is\\
 
@@
 Moving to the right, certainty on the performance of the method increases.
-The methods for which it holds that there is no other method that has 1) a higher NWN and 2) a higher number datasets evaluated, are identified as the best performing methods and are labeled.
+The Pareto optimal learning to rank methods, that is, the learning to rank methods for which it holds that there is no other method that has 1) a higher NWN and 2) a higher number datasets evaluated, are identified as the best performing methods and are labeled.
 Table \ref{tab:raw_data} in the appendix provides raw NWN data for the learning to rank methods at NDCG@$\{3,5,10\}$ and MAP and their cross-metric weighted average.
@@
 Figure \ref{fig:normalized_winning_number_ndcg3} shows the NWN of learning to rank methods based on NDCG@3 results.
-LambdaNeuralRank and CoList both acquired a NWN score of 1.0 by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
-LARF and LRUF scored very high scores of near 1.0 on three of the LETOR 3.0 datasets, which results in more certainty on these methods' performance because they are validated on three datasets that additionally are more relevant than AOL and Yahoo Set 2 (number of evaluation results for LETOR 3.0 are higher than those for AOL and Yahoo set 2).
+LambdaNeuralRank and CoList both acquired a NWN score of $1.0$ by beating all other algorithms on one dataset, with LambdaNeuralRank winning on the AOL dataset and CoList winning on Yahoo Set 2.
+LARF and LRUF scored very high scores of near $1.0$ on three of the LETOR 3.0 datasets, which results in more certainty on these methods' performance because they are validated on three datasets that additionally are more relevant than AOL and Yahoo Set 2 (number of evaluation results for LETOR 3.0 are higher than those for AOL and Yahoo set 2).
 FenchelRank, OWPC, SmoothRank, DCMP and ListNet are ordered decreasingly by NWN and at the same time increasingly in number of datasets that they are evaluated on, resulting in a higher degree of certainty on the accuracy of the algorithms.\\
@@
 LambdaNeuralRank, CoList, LARF, LRUF, OWPC and DCMP evaluation results are all based on one study, therefore are subjected to the risk of one overly optimistic study producing those results.
-FenchelRank evaluation result are based combined result from two studies, although those studies have overlap in authors.
+FenchelRank evaluation result are the combined result from two studies, although those studies have overlap in authors.
 SmoothRank and ListNet have the most reliable evaluation result source, as they were official LETOR baseline runs.
@@
 Figure \ref{fig:normalized_winning_number_ndcg5} shows the NWN of learning to rank methods based on NDCG@5 results.
-LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a NWN of 1.0.
+LambdaNeuralRank again beat all other methods solely with results on the AOL dataset scoring a NWN of $1.0$.
 LARF, LRUF, FenchelRank, SmoothRank, DCMP and ListNet are from left to right evaluated on an increasing number of datasets, but score decreasingly well in terms of NWN.
@@
 Figure \ref{fig:normalized_winning_number_ndcg10} shows the NWN of learning to rank methods based on NDCG@10 results.
-LambdaMART and LambdaNeuralRank score a NWN of 1.0 on the NDCG@10 comparison.
+LambdaMART and LambdaNeuralRank score a NWN of $1.0$ on the NDCG@10 comparison.
 For LambdaNeuralRank these results are again based on AOL dataset measurements.
 LambdaMART showed the highest NDCG@10 performance for the MSLR-WEB10k dataset.
-The set of algorithms for which there is no other algorithm with both a higher NWN and number of datasets evaluated on is partly in agreement with those for the NDCG@3 and @5 comparisons: {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
-In contrast to the NDCG@3 and @5 comparison, DCMP is not among the best algorithms.
+The set of Pareto optimal learning to rank algorithms is partly in agreement with the set of Pareto optimal methods for the NDCG@3 and @5 comparisons, both include {LARF, LRUF, FSMRank, SmoothRank, ListNet, RankSVM}.
+In contrast to the NDCG@3 and @5 comparisons, DCMP is not a Pareto optimal ranking method in the NDCG@10 comparison.
 
@@
 Figure \ref{fig:normalized_winning_number_map} shows the NWN of learning to rank methods based on MAP results.
-Comparisons on the NDCG metrics where highly in agreement on the best performing algorithms, MAP-based NWN results show different results.
-RankDE scores a NWN of 1.0 on one dataset, which is achieved by obtaining highest MAP-score on the LETOR 2.0 TD2003 which has many evaluation results are evaluated.\\
+Comparisons on the NDCG metrics where highly in agreement on the Pareto optimal algorithms, MAP-based NWN results show different results.
+RankDE scores a NWN of $1.0$ on one dataset, which is achieved by obtaining highest MAP-score on the LETOR 2.0 TD2003 which has many evaluation results are evaluated.\\
 
 LARF and LRUF score very high NWN scores, but based on only few datasets, just as in the NDCG-based comparisons.
-Notable is that low performance of SmoothRank and ListNet, given that those methods were top performing methods in the NDCG-based comparisons.
+Notable is the low performance of SmoothRank and ListNet, given that those methods were top performing methods in the NDCG-based comparisons.
 Table \ref{tab:raw_data} in the appendix shows that LAC-MR-OR is evaluated on more datasets on MAP than on NDCG, thereby LAC-MR-OR obtained equal certainty to ListNet with a higher NWN.
-SmoothRank performed a NWN of around 0.53 over 7 datasets, which is still good in both certainty and accuracy, but not among the top methods.
+SmoothRank performed a NWN of around $0.53$ on 7 datasets, which is good in both certainty and accuracy, but not a Pareto optimum.
 RE-QR is one of the best performers in the MAP comparison with a reasonable amount of benchmark evaluations.
@@
 \end{figure}
-Figure \ref{fig:normalized_winning_number_all} shows the NWN as function of IWN for the methods described in Table \ref{tab:ltr_methods_used}.
-The cross-metric comparison is based on the NDCG@$\{3,5,10\}$ and MAP comparisons combined, which justifies analysing the comparison more thoroughly.
-Figure \ref{fig:normalized_winning_number_all} labels the algorithms with no other algorithm having a higher value on both the horizontal axis and vertical axis, but also labels the algorithms with exactly one algorithm having a higher value on both axes with smaller font size.
-In addition, Linear Regression and the ranking method of simply sorting on the best single feature are labelled as baselines.\\
+Figure \ref{fig:normalized_winning_number_all} shows NWN as function of IWN for the methods listed in Table \ref{tab:ltr_methods_used}.
+The cross-metric comparison is based on the NDCG@$\{3,5,10\}$ and MAP comparisons combined.
+Figure \ref{fig:normalized_winning_number_all} labels the Pareto optimal algorithms, but also the Rank-2 Pareto optima, which are the labels the algorithms with exactly one algorithm having a higher value on both axes.
+Pareto optimal are labeled in large font while Rank-2 Pareto optima are labeled using a smaller font size.
+In addition, Linear Regression and the ranking method of simply sorting on the best single feature are labeled as baselines.\\
 
 LRUF, FSMRank, FenchelRank, SmoothRank and ListNet showed to be the methods that have no other method superior to them in both IWN and NWN.
-LRUF is the only method that achieved this in all NDCG comparisons, the MAP comparison as well as the cross-metric comparison.
-With FenchelRank, FSMRank, SmoothRank and ListNet being among the top performing methods in all NDCG comparisons as well as in the cross-metric comparison, it can be concluded that the cross-metric results are highly defined by the NDCG performance as opposed to the MAP performance.
+LRUF is the only method that achieved Pareto optimality in all NDCG comparisons, the MAP comparison as well as the cross-metric comparison.
+With FenchelRank, FSMRank, SmoothRank and ListNet being Pareto optimal in all NDCG comparisons as well as in the cross-metric comparison, it can be concluded that the cross-metric results are highly defined by the NDCG performance as opposed to the MAP performance.
 This was to be expected, because the cross-metric comparison input data of three NDCG entries (@3, @5, and @10) enables it to have up to three times as many as many weight as the MAP comparison.\\
 
-LARF, IPRank and DCMP and several variants of RankSVM performed very well on the cross-metric comparison, with all having only one method in its top right quadrant.
-LARF also performed among the top methods on the NDCG and MAP comparisons and DCMP was a top performer in a few of the NDCG comparisons.
+LARF, IPRank and DCMP and several variants of RankSVM are the Rank-2 Pareto optima of the cross-metric comparison.
+LARF was also a Pareto optima on the NDCG and MAP comparisons and DCMP was a Pareto optimal ranker in a few of the NDCG comparisons.
 C-CRF, DirectRank, FP-Rank, RankCSA, LambdaNeuralRank and VFLR all have a near-perfect NWN value, but have a low IWN value.
 Further evaluation runs of these methods on benchmark datasets that they are not yet evaluated on are desirable.
-The DirectRank paper \cite{Tan2013} shows that the method  is evaluated on more datasets than the number of datasets that we included evaluation results for in this meta-analysis.
+The DirectRank paper (\cite{Tan2013}) shows that the method  is evaluated on more datasets than the number of datasets that we included evaluation results for in this meta-analysis.
 Some of the DirectRank measurements could not be used because measurements on some datasets were only available in graphical form and not in raw data.\\
@@
 LAC-MR-OR and RE-QR showed very good ranking accuracy in the MAP comparison on multiple datasets.
-Because LAC-MR-OR is only evaluated on two datasets for NDCG@10 and RE-QR is not evaluated for NDCG at all, LAC-MR-OR and RE-QR are not within the top performing methods in the cross-metric comparison.
+Because LAC-MR-OR is only evaluated on two datasets for NDCG@10 and RE-QR is not evaluated for NDCG at all, LAC-MR-OR and RE-QR are not within the Pareto front of rankers in the cross-metric comparison.
+
+\section{Sensitivity Analysis}
+In this section we evaluate the stability of the obtained results when one of the evaluation measures (\ref{subsec:sensitivity_evaluation_measure}) or one of the datasets (\ref{subsec:sensitivity_data_set}) are left out of the comparison.
+We scope this sensitivity analysis to those ranking methods that showed to be Pareto optimal in the trade-off between IWN and NWN: ListNet, SmoothRank, FenchelRank, FSMRank and LRUF.
+\subsection{Sensitivity in the evaluation measure dimension}
+\label{subsec:sensitivity_evaluation_measure}
+To analyze the sensitivity of the comparison method in the evaluation measure dimension we repeated the NWN and IWN calculation while leaving one evaluation measure.
+Table \ref{tab:sensitivity_evaluation_measure} shows the NWN and IWN results when all evaluation measures are included in the computation and when MAP, NDCG@3, NDCG@5 or NDCG@10 are left out respectively.
+From this table we can infer that FSMRank is not a Pareto optimal ranking method when MAP is left out of the comparison (LRUF scores higher on both NWN and IWN) and FenchelRank is not a Pareto optimal ranking method when either NDCG@3 or NDCG@5 are left out (FSMRank scores higher on both NWN and IWN).
+All other orderings of ranking methods on NWN and IWN stay intact when one of the evaluation measures is left out of the comparison.\\
+
+Notable is that all Pareto optimal ranking methods have the largest increase in IWN as well as the largest decrease in NWN when the MAP measure is left out of the comparison.
+The NWN score of FSMRank increased almost $0.1$ when the MAP evaluation measure was left out, which is the highest deviation in NWN score seen in this sensitivity analysis.
+Note that MAP uses a binary notion of relevance, where NDCG uses graded relevance.
+The fact that all Pareto optimal rankers obtain an even higher NWN score when the MAP measure is left out shows that apparently the Pareto optimal rankers perform even better on ranking on graded relevance, compared to non-Pareto-optimal rankers.
+\begin{table}[!h]
+	\begin{adjustbox}{max width=\textwidth}
+\begin{tabular}{lll|ll|ll|ll|ll}
+	& \multicolumn{2}{c}{All} & \multicolumn{2}{c}{MAP} & \multicolumn{2}{c}{NDCG@3} & \multicolumn{2}{c}{NDCG@5} & \multicolumn{2}{c}{NDCG@10} \\ 
+	& NWN & IWN & NWN & IWN & NWN & IWN & NWN & IWN & NWN & IWN \\
+	\hline
+	ListNet & 0.4952 & 931 & 0.5127 & 669 & 0.5099 & 710 & 0.4965 & 707 & 0.4625 & 707 \\ 
+	SmoothRank & 0.6003 & 653 & 0.6266 & 474 & 0.5988 & 491 & 0.5900 & 500 & 0.5870 & 494 \\ 
+	FenchelRank & 0.7307 & 505 & 0.7628 & 371 & 0.7158 & 380 & 0.7244 & 381 & 0.7206 & 383 \\ 
+	FSMRank & 0.7593 & 482 & 0.8585 & 311 & 0.7403 & 385 & 0.7292 & 384 & 0.7268 & 366 \\ 
+	LRUF & 0.9783 & 460 & 0.9821 & 335 & 0.9767 & 344 & 0.9772 & 351 & 0.9771 & 350 \\ 
+	LARF & 0.9868 & 379 & 0.9891 & 275 & 0.9859 & 283 & 0.9861 & 288 & 0.9863 & 291 \\ 
+\end{tabular}
+\end{adjustbox}
+\label{tab:sensitivity_evaluation_measure}
+\caption{NWN and IWN scores of the Pareto optimal rankers on all evaluation metrics, and with MAP, NDCG@3, NDCG@5 or NDCG@10 left out of the comparison respectively}
+\end{table}
+
+\subsection{Sensitivity in the dataset dimension}
+\label{subsec:sensitivity_data_set}
+In Table \ref{tbl:ltr_benchmark_collections} we showed to include 20 datasets in our comparison, originating from eight data collections.
+We analyzed the variance in NWN and IWN scores of the Pareto optimal rankers for the situations where one of the 20 datasets is not included in the NWN and IWN computation.
+The results are visualized in Figure \ref{fig:sensitivity_data_set} in a series of bagplots, which is a bivariate generalization of the boxplot proposed by \cite{Rousseeuw1999}.
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.5]{gfx/bagplots}
+	\caption{Bagplots showing the variance in NWN and IWN of the Pareto optimal rankers when a dataset is left out of the comparison}
+	\label{fig:sensitivity_data_set}
+\end{figure}
 
@@
 In the NWN calculation, the weight of each benchmark on the total score is determined by the number of evaluation measurements on this benchmark.
-By calculating it in this way, we implicitly make the assumption that the learning to rank methods are (approximately) distributed uniformly over the benchmarks, such that the average learning to rank method tested are approximately equally hard for each data set.
+By calculating it in this way, we implicitly make the assumption that the learning to rank methods are (approximately) distributed uniformly over the benchmarks, such that the average learning to rank method tested are approximately equally hard for each dataset.
 It could be the case however that this assumption is false and that the accurateness of the learning to rank methods on a dataset is not dataset independent.\\

### 2015-05-05 paper/IPM/paper.tex
opgestuurd naar Djoerd

@@
 The fact that all Pareto optimal rankers obtain an even higher NWN score when the MAP measure is left out shows that apparently the Pareto optimal rankers perform even better on ranking on graded relevance, compared to non-Pareto-optimal rankers.
-\begin{table}[!h]
+\begin{table}
 	\begin{adjustbox}{max width=\textwidth}
@@
 The results are visualized in Figure \ref{fig:sensitivity_data_set} in a series of bagplots, which is a bivariate generalization of the boxplot proposed by \cite{Rousseeuw1999}.
-\begin{figure}[!h]
+Bagplot extends the univariate concept of rank as used in a boxplot to a halfspace location depth.
+The \emph{depth median}, shown in orange, is the deepest location.
+Surrounding it is a \emph{bag}, the dark blue area in Figure \ref{fig:sensitivity_data_set}, containing $\frac{n}{2}$ observations with the largest depth.
+The light blue area represents the \emph{fence}, which magnifies the bag by a factor $3$.\\
+
+Note that the number of unique observations on which the bagplots are created is equal to the number of dataset on which a ranking method is evaluated (in any of the evaluation measures), as removing a dataset on which a ranking algorithm is not evaluated does not have any effect on the NWN and IWN scores.
+The difference between the leftmost and the rightmost points of the bags seems to be more or less equal for all ranking methods while the NWN means are decreasing from top-to-bottom and and left-to-right.
+Therefore, the variance-to-mean ratio increases from top-top-bottom and from left-to-right.
+On the IWN dimension it is notable that LRUF and LARF has very low variance.
+It is important to stress that this does not imply high certainty about the level of ranking performance of these ranking methods, it solely shows the low variance in the evaluation results available for the ranking methods.
+
+\begin{figure}
 	\centering
-	\includegraphics[scale=0.5]{gfx/bagplots}
+	\includegraphics[scale=0.44]{gfx/bagplots}
 	\caption{Bagplots showing the variance in NWN and IWN of the Pareto optimal rankers when a dataset is left out of the comparison}
@@
 However, no closing arguments can be formulated on which learning to rank methods are most accurate.
-LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were the learning to rank algorithms for which it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy.
+LRUF, FSMRank, FenchelRank, SmoothRank and ListNet were found to be the Pareto optimal learning to rank algorithms in the NWN and IWN dimensions: for these ranking algorithm it holds that no other algorithm produced both more accurate rankings (NWN) and a higher degree of certainty of ranking accuracy (IWN).
 From left to right, the ranking accuracy of these methods decreases while the certainty of the ranking accuracy increases.\\

### 2015-05-05 method_selection/bagplot_dataset_sensitivity.R

@@ -8,7 +8,7 @@ lruf        = read.csv("lruf_dataset_sensitivity.csv")
 larf		= read.csv("larf_dataset_sensitivity.csv")
 
 layout(matrix(c(1,2,3,4,5,6),3,2,byrow=TRUE)) 
-bagplot(listnet, show.whiskers=FALSE)
+bagplot(listnet, show.whiskers=FALSE, show.loophull=FALSE)
 title("ListNet")
 bagplot(smoothrank, show.whiskers=FALSE)
 title("SmoothRank")
